### 2020-01-04 combinatorics.tex
initial

\begin{refsection}
\newcommand{\pifun}[2]{\prod_{\left( \mathit{ #1 } : #2 \right)}}
\chapter{Combinatorics}
\begin{quote}
  \textbf{Abstract}
  We explore five notions of finiteness in Homotopy Type Theory \cite{hottbook}.
  We prove closure properties about all of these notions, culminating in a proof
  that decidable Kuratowski-finite sets form a topos.

  We extend the definitions to include infinite types, developing a similar
  classification of countable types.

  We use the definition of finiteness to formalise \emph{species}, in much
  the same way as in \cite{yorgeyCombinatorialSpeciesLabelled2014}.
  A clear duality with containers
  \cite{abbottContainersConstructingStrictly2005} falls out naturally from our
  definition.

  We formalise our work in Cubical Agda
  \cite{vezzosiCubicalAgdaDependently2019}, and we implement a library for proof
  search (including combinators for level-polymorphic fully generic currying),
  and demonstrate how it can be used to both prove properties and synthesise
  full functions given desired properties.
\end{quote}
\section{Introduction}
Proofs in a constructive setting are more substantial things than their
classical counterparts.
Rather than just evidence for some fact, they are objects in their own right,
which we can manipulate and extract information from.

Sometimes, the extra detail can reveal interesting subtleties: fissures which
divide homogeneous-seeming definitions into a variety of related concepts.
Finiteness is a prime example of this phenomenon: in classical set theory, a
finite set is a finite set, and there's not much else that we can say.
In constructive type theory, we start the bidding at no fewer than \emph{four}
different predicates representing finiteness
\cite{spiwackConstructivelyFinite2010}, upped to eight in
\cite{firsovVariationsNoetherianness2016a}: all of which differ in significant
and interesting ways.

Other times, the extra baggage is unwanted.
Maybe we want to enforce that a particular piece of information should be
hidden, or treated as irrelevant: ``yes, I have proven that there is a
particular number with this property, but I don't want to reveal \emph{which}
number.''

This is where Homotopy Type Theory \cite{hottbook} can step in.
Through primarily a generalisation of equality, we gain access to a number of
exotic types which can, among other things, allow us to perform this ``hiding''.
In the context of finiteness, the act of hiding is worth studying in its own
right, and further delineates several new finiteness predicates which collapse
even in standard Martin-Löf type theory
\cite{martin-lofIntuitionisticTypeTheory1980}.

The other headline feature of HoTT's generalised equality is univalence: put
simply, it allows us to treat isomorphic types as equal, giving us in the
constructive world access to a technique that is commonplace classically.
It turns out that finite types have a lot to say about isomorphisms, and we will
rely heavily on this aspect of HoTT to work with them.
\subsection{Contributions}
In this work, we will explore finite types in Cubical Type Theory
\cite{cohenCubicalTypeTheory2016}, and expose their relationship to infinite
types, species, and demonstrate their practical uses for proofs in dependently
typed programming languages.

In section~\ref{strong-finiteness}, we will explore the ``strong'' notions of
finiteness (i.e. those at least as strong as Kuratowski finiteness
\cite{kuratowskiNotionEnsembleFini1920}), with a special focus on cardinal
finiteness.
We will delineate some subtly different finiteness predicates, and prove the
relationship between each.
We will prove, then, that every discrete Kuratowski finite type is cardinal
finite, and subsequently will show that discrete Kuratowski finite sets form a
topos.

In section~\ref{species-and-containers}, we will redefine our finiteness
predicates using \emph{containers}, and prove that the new definitions are
equivalent to the old.
Using this, we will prove much stronger relationships between our finiteness
predicates and relation-based definitions of finiteness.
Finally, We will show how the species-container duality falls out naturally from
these new definitions.

In section~\ref{infinite-cardinalities}, we will extend our study of finite
types to infinite but countable types.
We will see that the finiteness predicates are mirrored with countable
counterparts, and we will prove closure under the Kleene star and plus.

Proofs of finiteness have well-known practical applications in
constructive mathematics \cite{firsovDependentlyTypedProgramming2015}.
In section~\ref{practical}, we build a library which exploits these uses in
Cubical Agda \cite{vezzosiCubicalAgdaDependently2019}, allowing automation of
complex proofs over finite types.
We frame this in terms of the principle of omniscience for finite types.
Thanks to the flexibility afforded to us by Cubical Type Theory, we are able to
go further than the usual examples of this kind of proof automation: as well as
proving properties about functions, we can synthesise functions whole-cloth from
their desired properties.
Through the unified interface for finite and countable types, we can
reuse the automation machinery for \emph{partial} proof search over infinite
search spaces.
Along the way, we extend the work in \cite{allaisGenericLevelPolymorphic2019} to
prove isomorphisms between the curried and uncurried forms of \(n\)-ary
dependent functions.
\section{Strong Finiteness} \label{strong-finiteness}
\begin{figure}[h]
  \centering
  \begin{tikzcd}[cramped, row sep=small]
    &
    \ar[ddddd, dash, start anchor={[yshift=4ex]}, end anchor={[yshift=-4ex]}] &
    \text{Ordered} &
    \ar[ddddd, dash, dashed, start anchor={[yshift=4ex]}, end anchor={[yshift=-4ex]}] &
    \text{Unordered}
    \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \ar[rrrr, dash, start anchor={[xshift=-6ex]}, end anchor={[xshift=10ex]}] &
    &
    &
    &
    {}
    \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    &
    &
    \text{Split Enumerable}
      \; \ref{split-enumerable}
      \ar[d, xshift=-1ex, bend right=30]
      &
    &
    \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \text{Discrete} &
    &
    \text{Manifest Bishop}
      \ar[u, xshift=1ex, bend right=30]
      \ar[rr, end anchor=south west, start anchor=south east, bend right=10, crossing over]
      \; \ref{manifest-bishop}
      &
    &
    \text{Cardinal}
      \; \ref{cardinal}
      \ar[ll, start anchor=north west, end anchor=north east, bend right=10, "\text{linOrd}" description]
    \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \ar[rrrr, dash, dashed, start anchor={[xshift=-6ex]}, end anchor={[xshift=10ex]}] &
    &
    &
    &
    {}
    \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \text{Non Discrete} &
    {} &
    \text{Manifest Enumerable}
      \; \ref{manifest-enumerability}
      \ar[from=uu, xshift=-1ex, bend right=30, crossing over]
      \ar[uu, bend right=30,  "\text{Discrete}" description]
      \ar[rr, start anchor=south east, end anchor=south west, bend right=10, crossing over]
      &
    {}
    &
    \text{Kuratowski}
      \; \ref{kuratowski}
      \ar[ll, start anchor=north west, end anchor=north east, bend right=10,  "\text{linOrd}" description]
      \ar[uu, xshift=1ex, bend right=30,  "\text{Discrete}" description]
      \ar[from=uu, xshift=-1ex, bend right=30, crossing over]
  \end{tikzcd}
  \caption{Classification of the strong definitions of finiteness, according to
    whether they are discrete (imply decidable equality) and whether they induce
    a linear order.}
  \label{finite-classification}
\end{figure}

In this section, we will explore five predicates for finiteness, classified in
figure~\ref{finite-classification}.
Of these, manifest enumerability is new, to our knowledge.

Figure~\ref{finite-classification} organises the predicates according to
their ``strength''; i.e. how much information they provide about a conforming
type.
For instance, a proof that some type \(A\) is manifestly Bishop finite (the
strongest of the notions, explored in section~\ref{manifest-bishop}) also tells
us that \(A\) is discrete (has decidable equality), and gives us a linear order
on the type.
A type that is Kuratowski finite (section~\ref{kuratowski}) has no such extra
features: indeed, we will see examples of Kuratowski finite types which are not
even sets, never mind discrete ones.

In the following sections, we will go through each of the predicates, proving
how to weaken each (i.e. we will provide a proof that every cardinally finite
type is Kuratowski finite), and how to strengthen them, given the required
property.
In terms of figure~\ref{finite-classification}, this amounts to providing proofs
for each arrow.

For each predicate, we will also prove its closure over sums and products in
both dependent and non-dependent forms, if such a closure exists.
This will culminate in our main result for this section: the formal proof that
decidable Kuratowski finite sets form a topos. \todo[noline]{reference for this}
More specifically, we show that cardinal finite sets form a topos, that
decidable Kuratowski finite sets are equivalent in strength to cardinal finite
sets, and carry the proof over.
This result relies on proofs on each of the other finiteness predicates.
\subsection{Split Enumerability} \label{split-enumerable}
We will start with the simplest definition of finiteness: we say a set is
enumerable if there is a list of its elements which contains every element in
the set.
More formally:
\begin{definition}[Split Enumerable Set]
  \begin{equation}
    \mathcal{E}!(A) = \sum_{(\mathit{xs} : \textbf{List} A)} , \prod_{(x : A)} , x \in xs
  \end{equation}
  We call the first component of this pair the ``support'' list, and the second
  component the ``cover'' proof.
\end{definition}

The term \(x \in \mathit{xs}\) hides a subtle detail of this definition.
Depending on how we choose to represent \(x \in \mathit{xs}\), the resulting
predicate will be equivalent in strength to either manifest Bishop finiteness
(section~\ref{manifest-bishop}) or manifest enumerability
(section~\ref{manifest-enumerability}).
For now, we will define it like so:

\begin{definition}[List Membership] \label{list-membership}
  \begin{equation}
    x \in \mathit{xs} = \sum_{(i : \mathbf{Fin}(\text{length}(\mathit{xs})))} , \mathit{xs} \ind i \equiv x
  \end{equation}
  Here, \(\mathbf{Fin}(n)\) is the type of natural numbers smaller than \(n\),
  and \(\ind\) is an indexing function.
  In other words, to say that some item \(x\) is in some list \(\mathit{xs}\),
  we say that there is an index into \(\mathit{xs}\) which selects an element of
  \(\mathit{xs}\) which is equivalent to \(x\).
\end{definition}

Clearly, this definition of list membership tells us not just that an item is in
a list, but also its \emph{position} in that list.
That piece of information gives use decidable equality, restricting the
predicate to discrete sets.
In section~\ref{manifest-enumerability}, we will show how to hide the position,
giving us a variant of the predicate which allows for non-set finite types.
\subsubsection{Decidable Equality}
We will now prove that split enumerability implies decidable equality.
We first need to prove a lemma regarding surjections to split enumerable types.
\begin{lemma} \label{split-enum-surj}
  Any split enumerable type has a split surjection from a finite prefix of the
  natural numbers.
\end{lemma}
\begin{proof} 
  Let \(E\) be a proof of type \(\mathcal{E}!(A)\) for some type \(A\), and
  \(n\) be the length of the support list of \(E\).
  Then let \(f : \textbf{Fin}(n) \rightarrow A\) be defined as the indexing
  function into the support list of \(E\).

  We then prove this function to be a split surjection by providing a fiber over
  all of the elements of its domain: this fiber is the second component of \(E\).

  In Agda:
  \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{to-surj}
\end{proof}

Because of careful arrangement of previous definitions, this proof is only three
lines of Agda.
It is not difficult to derive decidable equality from this fact.
\begin{lemma} \label{split-enum-discrete}
  Any split enumerable type has decidable equality (is discrete).
\end{lemma}
\begin{proof}
  We use a corollary, proven in the appendix, that if there is a
  split-surjection from \(A\) to \(B\), and \(A\) is discrete, then \(B\) is
  also discrete.
  \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{discrete}
\end{proof}

This also tells us that any split enumerable type must be a set.
\begin{lemma}
  Any split enumerable type is a set.
\end{lemma}
\begin{proof}
  By Hedberg's theorem \cite{hedbergCoherenceTheoremMartinLof1998}, since split
  enumerable types have decidable equality
  (proposition~\ref{split-enum-discrete}), they are sets.
\end{proof}

Though it's not strictly needed for the rest of the construction, it is
instructive to also prove inverse regarding surjections; i.e. that we can
derive \(\mathcal{E}!\) from a surjection from another finite type.
\begin{lemma} \label{split-enum-from-surj}
  Any type with a split surjection from another split enumerable type is itself
  split enumerable.
\end{lemma}
\begin{proof}
  Let \(E\) be a proof of split enumerability for some set \(A\), and let \(f\)
  be a surjection from \(A\) to \(B\).
  The support list for the proof of split enumerability over \(B\) can be
  obtained by mapping the surjective function over the support list of \(E\).

  To obtain the cover proof, let \(f^{-1}\) be the inverse function of \(f\)
  (which we can obtain because it is a split-surjection).
  Using the cover proof of \(E\), we can show that, for all \(y : B\), \(f^{-1}(y)\)
  is in the support list for \(E\).
  Using the following lemma (proven in the appendix):
  \begin{equation}
    x \in xs \rightarrow f (x) \in \text{map} \; f \; \mathit{xs}
  \end{equation}
  We obtain a proof that:
  \begin{equation}
    \forall (y : B). f (f^{-1}(y)) \in \text{map} \; f \; \mathit{xs}
  \end{equation}
  And, since \(f (f^{-1}(y)) \equiv y\) (by the definition of a surjection), we
  have proven our goal.

  In Agda:
  \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{from-surj}
\end{proof}
\subsubsection{Closure}
In this section we will prove closure under various operations for split
enumerable sets.
We are working towards a topos proof, which requires us to prove closure under
a variety of operations: for now, we only have enough machinery to demonstrate
the semiring operations, and dependent sums.
in order to show closure under exponentials (function arrows), we will need an
equivalence with \(\mathbf{Fin}\), which will be provided in
section~\ref{manifest-bishop}.
\begin{lemma}
  \(\bot\) is split enumerable.
\end{lemma}
\begin{proof}
  The support list for \(\bot\) is the empty list.
  The cover proof is present via the principle of explosion.
\end{proof}
\begin{lemma}
  \(\top\) is split enumerable.
\end{lemma}
\begin{proof}
  The support list for \(\top\) is the singleton list containing \(\text{tt}\).
  The cover proof is a constant function which returns an index into the first
  item in the list.
\end{proof}
Both of those cases were relatively simple.
Next, we will look into how to combine proofs of split enumerability.
\begin{theorem}
  Split-enumerability is closed under \(\sum\).
\end{theorem}
\begin{proof}
  Let \(E_A\) be a proof of split enumerability for some type \(A\), and \(E_U\)
  be a function of the type:
  \begin{equation}
    E_U : \prod_{(x : A)} , \mathcal{E}!(U(x))
  \end{equation}
  In other words, a function which returns a proof of split enumerability for
  each member of the family \(U\).

  To obtain the support list, we concatenate the support lists of all the proofs
  of split-finiteness for \(U\) over the support list of \(E_A\).
  In Agda:
  \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{sigma-enum}
  ``do-notation'' is available to us as we're working in the list monad.

  The cover proof follows in a simple, if tedious, way.
\end{proof}
\begin{lemma}
  Split-enumerability is closed under Cartesian product (non-dependent product).
\end{lemma} 
\begin{proof}
  Since non-dependent products are simply a special case of the \(\sum\) type,
  we can reuse the proof of closure over \(\sum\) here.
\end{proof}
\begin{lemma}
  Split-enumerability is closed under disjoint union (non-dependent sum).
\end{lemma}
\begin{proof}
  Again, this can be derived from closure of the \(\sum\) type.
  Disjoint union between two types \(A\) and \(B\) can be represented by the
  following type:
  \begin{equation}
    A + B = \sum_{(x : \mathbf{Bool})} , \text{if} \; x \; \text{then} \; A \; \text{else} \; B
  \end{equation}
  Then, since all of \(\mathbf{Bool}\), \(A\), and \(B\) are split enumerable,
  the type \(A + B\) is split enumerable.
\end{proof}
The final case can be derived from the previous closures; we give it here anyway
for clarity.
\begin{lemma} \label{fin-split-enum}
  Any finite prefix of the natural numbers (what we have been calling
  \(\textbf{Fin}\)) is split enumerable.
\end{lemma}
\begin{proof}
  For the support list, we will make use of the function tabulate, which returns
  a list consisting of the application of \(f\) to every \(\textbf{Fin}(n)\),
  given \(n\).
  \begin{equation}
    \text{tabulate}(n , f) = \left[ f(0) , f(1) \dots f(n - 1) \right]
  \end{equation}
  In Agda:
  \ExecuteMetaData[agda/Data/List.tex]{tabulate}
  Tabulation with the identity function gives us our cover list.

  The cover proof is obviously true given the definition of tabulate above; in
  Agda it requires a little more work to show.
  \ExecuteMetaData[agda/Data/List/Membership.tex]{fin-in-tabulate}
\end{proof}
\subsection{Manifest Bishop Finiteness} \label{manifest-bishop}
\begin{definition}[Manifest Bishop Finiteness]
  \begin{equation}
    \mathcal{B}(A) = \sum_{(\mathit{xs} : \textbf{List}(A))} , \prod_{(x : A)} , x \inunique xs
  \end{equation}
\end{definition}
The only difference between this predicate and split enumerability is the list
membership term: we use \(\inunique\) here, where \(x \inunique \mathit{xs}\) is
to be read as ``\(x\) occurs exactly once in \(\mathit{xs}\)''.
It is defined in terms of our previous list membership term
(definition~\ref{list-membership}).
\begin{definition}[Unique List Membership]
  We say an item \(x\) is ``uniquely in'' some list \(\mathit{xs}\) if its
  membership in that list is a \emph{contraction}; i.e. its membership proof
  exists, and all such proofs are equal.
  \begin{equation}
    x \inunique \mathit{xs} = \text{isContr}(x \in \mathit{xs})
  \end{equation}
\end{definition}

A nice consequence of prohibiting duplicates is that now the length of the
support list is the same as the cardinality of the set.
\subsubsection{Equivalence}
While manifest Bishop finiteness may seem stricter than split enumerability, we
will see soon that it is actually equivalent in strength.
Why study it separately to split enumerability at all, then?
What purpose does it serve on our road to Kuratowski finiteness?
Those questions are answered by the following theorem:
\begin{lemma} \label{bishop-equiv}
  Any manifestly Bishop finite type has an equivalence with a finite prefix of
  the natural numbers.
\end{lemma}
\begin{proof}
  There are many equivalent definitions of equivalence in HoTT.
  Here we take the version preferred in the Cubical Agda library: contractible
  maps \cite[Definition~4.4.1]{hottbook}. \todo[noline]{Need this in appendix also}
  Because of the parallels between contractible maps and split surjections,
  the proof proceeds much the same as \ref{split-enum-surj}.

  First, we define the map as the indexing function into the support list.
  Proving this map is a contraction is precisely the cover proof.

  \ExecuteMetaData[agda/Cardinality/Finite/ManifestBishop.tex]{equiv}
\end{proof}

As we can see, our list-based definition of manifest Bishop finiteness is really
an equivalence with \(\mathbf{Fin}\) in disguise.
This will be the key component in establishing closure under exponentials.

We would like to draw attention to the ``power-to-weight ratio'' of this proof.
We have mentioned how important the equivalence will be for our later proofs: in
general, the ability to transport via univalence opens up many opportunities

In Cubical Type Theory, equivalences are much more useful than their counterpart
(isomorphisms) in Martin-Löf  Type Theory: arbitrary proofs and functions can be
\emph{transported} along equivalences, giving a constructive justification for
an intuitive and useful technique which is commonplace classically.
The proof itself, however, is three lines long: thanks to careful definitions,
the fact that manifest Bishop finiteness is the ``same'' as equivalences to
\(\mathbf{Fin}\) becomes obviously true.
Later we will see, by defining lists as containers, a proof that manifest Bishop
finiteness is equivalent (in the formal sense) to equivalences with
\(\mathbf{Fin}\). \todo[noline]{This needs to be added to the containers
  section.}
\todo[noline]{More on this? }
\subsubsection{Relationship to Split Enumerability}
We now show that manifest Bishop finiteness has equal strength to split
enumerability.
\begin{lemma}
  Any manifest Bishop finite set is split enumerable.
\end{lemma}
\begin{proof}
  The support set carries over simply, and the cover proof can be taken from the
  first component of the cover proof from the proof of manifest Bishop
  finiteness.
  \ExecuteMetaData[agda/Cardinality/Finite/ManifestBishop.tex]{bishop-to-split-enum}
\end{proof}
\begin{theorem} \label{split-enum-to-manifest-bishop}
  Any split enumerable set is manifest Bishop finite.
\end{theorem}
\begin{proof}
  Let \(E\) be a proof of split enumerability for some set \(A\).
  From proposition~\ref{split-enum-discrete} we can derive decidable equality on
  \(A\), and using this we can define a function \(\text{uniques}\) which filters
  out duplicates from lists of \(A\)s.
  \begin{equation}
    \text{uniques} : \textbf{List}(A) \rightarrow \textbf{List}(A)
  \end{equation}
  This gives us our support list.

  It suffices now to prove the following:
  \begin{equation}
    \prod_{(x : A)} , \prod_{(\mathit{xs} : \textbf{List}(A))} , x \in \mathit{xs} \rightarrow x \inunique \text{uniques}(\mathit{xs})
  \end{equation}
  And from that we can generate our cover proof.

  Full details of the formal proof are provided in the appendix.
\end{proof}
Note that while manifest Bishop finiteness as split enumerability are equivalent
in ``strength'', the two types are not equivalent.
In particular, there are infinitely many inhabitants of \(\mathcal{E}!\), while
for a type \(A\) with \(n\) inhabitants, there are only \(n!\) inhabitants of
\(\mathcal{B}(A)\).
\subsubsection{Closure}
Proving equal strength of split enumerability and manifest Bishop finiteness
allows us to carry all of the previous proofs of closure over to manifest Bishop
finite sets (and vice-versa).
Missing from our previous proofs was a proof of closure of functions.
We remedy that here.
\begin{theorem}
  Manifest bishop finiteness is closed over dependent functions (\(\prod\)-types).

  Formally, given a type \(A\) and a type family \(B\) on \(A\), when \(A\) is
  manifestly Bishop finite:
  \begin{equation}
    \mathcal{B}\left( A \right)
  \end{equation}
  And \(B\) is manifestly Bishop finite over all points of \(A\):
  \begin{equation}
    \prod_{(x : A)} , \mathcal{B}\left( B(x) \right)
  \end{equation}
  Then we have the following:
  \begin{equation}
    \mathcal{B}\left(\prod_{(x : A)} , B(x)\right)
  \end{equation}
\end{theorem}
\begin{proof}
  This proof is essentially the composition of two transport operations, made
  available to us via univalence.

  First, we will simplify things slightly by working only with split
  enumerability.
  As this is equal in strength to manifest Bishop finiteness, any closure proofs
  carry over.

  Secondly, we will replace \(A\) in all places with \(\mathbf{Fin}(n)\).
  Since we have already seen an equivalence between these two types, we are
  permitted to transport along these lines.
  This is the first transport operation.

  The bulk of the proof now is concerned with proving the following:
  \begin{equation}
    \left( \prod_{(x : \mathbf{Fin}(n))} , \mathcal{E}!(A(x)) \right) \rightarrow \mathcal{E}!\left( \prod_{(x : \mathbf{Fin}(n))} , A(x) \right)
  \end{equation}
  Our strategy to accomplish this will be to consider functions from
  \(\mathbf{Fin}(n)\) as \(n\)-tuples over some type family \(T\).
  \begin{equation}
    \begin{aligned}
      \mathbf{Tuple}(T, 0)   &= \top \\
      \mathbf{Tuple}(T, n+1) &= A(0) \times \mathbf{Tuple}(T \circ \text{suc}, n)
    \end{aligned}
  \end{equation}
  This type is manifestly Bishop finite, as it is constructed only from products
  and the unit type.

  We then prove an isomorphism between this representation and \(\Pi\)-types.
  \begin{equation}
    \mathbf{Tuple}(T, n) \iff \prod_{(x : \mathbf{Fin}(n))} , T(x)
  \end{equation}
  This allows us to transport our proof of finiteness on tuples to one on
  functions from \(\mathbf{Fin}\) (our second transport operation), proving our
  goal.

  \ExecuteMetaData[agda/Cardinality/Finite/ManifestBishop.tex]{pi-clos}
\end{proof}
\subsection{Manifest Enumerability} \label{manifest-enumerability}
As we said in section~\ref{split-enumerable}, we will now retrieve a distinction
between enumerability and manifest Bishop finiteness.
We do so by propositionally truncating the membership proof:
\begin{definition}[Manifest Enumerability]
  \begin{equation}
    \mathcal{E}(A) = \sum_{(\mathit{xs} : \mathbf{List}(A))} , \prod_{(x : A)} , \lVert x \in \mathit{xs} \rVert
  \end{equation}
  \ExecuteMetaData[agda/Cardinality/Finite/ManifestEnumerable.tex]{def}
\end{definition}

There is another definition, which lies between manifest enumerability and
bishop finiteness, where we would truncate just the equality, but reveal the
position in the membership proof.
\begin{equation}
  \mathcal{E}'(A) = \sum_{(\mathit{xs} : \mathbf{List}(A))} , \prod_{(x : A)} , \sum_{(i : \textbf{Fin}(\text{length}(\mathit{xs})))} , \lVert \mathit{xs} \ind i \equiv x \rVert
\end{equation}
However, this simply amounts to a proof of manifest Bishop finiteness over
set-truncated types, rather than a significantly different construction, so it
does not interest us here.
\subsubsection{Higher Homotopy Levels}
By hiding the position, we have essentially removed the ``decidable'' component
from split enumerability.
Our predicate now becomes general enough to work with non-sets: we will show
here that the circle is manifestly enumerable.
\begin{theorem}
  The circle \(S^1\) is manifestly enumerable.
\end{theorem}
\begin{proof}
  As the cover proof is a truncated proposition, we need only consider the point
  constructors, making this poof the same as the proof of split enumerability on
  \(\top\).

  The support list will be a singleton list containing the single point in the
  type, and the cover proof will be an index pointing at the first element in
  the support list.

  \ExecuteMetaData[agda/Cardinality/Finite/ManifestEnumerable.tex]{circ-inst}
\end{proof}
\subsubsection{Surjections}
This predicates relation to surjectivity is much the same as split
enumerability's relation to \emph{split} surjectivity.
Deriving a surjection from \(\mathbf{Fin}\) is straightforward:
\begin{lemma}
  If a type \(A\) is manifestly enumerable, then there is a surjection from
  \(\mathbf{Fin}\) into \(A\).
\end{lemma}
\begin{proof}
  \todo[inline]{Proof}
\end{proof}
And building from a surjection is similarly not too difficult:
\begin{lemma}
  If there is a surjection from a manifestly enumerable type \(A\) to another
  type \(B\), then \(B\) is manifestly enumerable.
\end{lemma}
\begin{proof}
  The proof proceeds in much the same way as in
  proposition~\ref{split-enum-from-surj}.
  The only difference is that the second component contains a truncation, but
  since we can always eliminate from a truncation into a truncation this is not
  a problem.

  In Agda:
  \ExecuteMetaData[agda/Cardinality/Finite/ManifestEnumerable.tex]{from-surj}
\end{proof}
\subsubsection{Relation to Split Enumerability}
As split enumerability is stronger than manifest enumerability, we can convert
from one to the other easily.
\begin{lemma}
  Any split enumerable type is also manifestly enumerable.
\end{lemma}
\begin{proof}
  The proof carries over via truncation of the cover proof.
\end{proof}
The main missing piece from manifest enumerability is decidable equality: by
truncating the membership proof, we have removed the ability to distinguish
(decidably) between members of the type.
Indeed, resupplying decidable equality is precisely sufficient to recover split
enumerability.

\begin{theorem} \label{manifest-enum-to-split}
  A manifestly enumerable type with decidable equality is split enumerable.
\end{theorem}
\begin{proof}
  The support list stays the same between both enumerability proofs.

  For the cover proof, we need a way to get the contents out of a truncated
  value.
  We can do exactly that with the following lemma, called recompute:
  Given a decision procedure for some type \(A\), and a truncation for the
  proposition \(A\), we can discount the possibility of \(A\) being false, and
  therefore extract the true decision.

  For the cover proof, our obligation now becomes constructing a decision for
  membership of the support list.
  This is straightforward given decidable equality.

  \ExecuteMetaData[agda/Cardinality/Finite/ManifestEnumerable.tex]{to-split-enum}
\end{proof}
By theorem~\ref{split-enum-to-manifest-bishop}, we also can derive that any
manifestly enumerable type with decidable equality is Bishop finite.
\subsubsection{Closure}
\begin{lemma}
  Manifest enumerability is closed under dependent sum, disjoint union
  (non-dependent sum), and Cartesian product (non-dependent product).
\end{lemma}
\begin{proof}
  For these three closures, the proofs on split enumerability consisted of a
  list manipulation followed by a proof that membership was preserved by the
  list manipulation.
  Because we separate these two concerns, the proofs carry over onto manifest
  enumerability: the support list manipulation stays the same, and the
  cover proofs are performed ``under'' the truncation.
\end{proof}
Notice that we do not have closure under functions: without decidability,
manifest enumerability is not closed under function arrows.
\subsection{Cardinal Finiteness} \label{cardinal}
We now turn our attention to notions of finiteness which do not induce a linear
order.
\begin{definition}[Cardinal Finiteness]
  A type \(A\) is cardinally finite, \(\mathcal{C}\), if it has a
  propositionally-truncated proof of bishop finiteness.
  \begin{equation}
    \mathcal{C}(A) = \lVert \mathcal{B}(A) \rVert
  \end{equation}
\end{definition}
\subsubsection{Characterisation}
Though we have chosen Bishop finiteness as our underlying finiteness proof to
truncate, we could have in fact used any equal-strength proof
(lemma~\ref{equal-strength}).
Because this predicate is a mere proposition, what we called before ``equal
strength'' now implies full equivalence.
With this proven, we can use any one of our equal-strength finite definitions
(\(\mathcal{E}!\), \(\mathcal{B}\)).
In particular, we have:
\begin{lemma}
  Cardinal finiteness is equivalent to a propositionally truncated equivalence
  with some \(\textbf{Fin}\).
  \begin{equation}
    \mathcal{C}(A) \simeq \lVert \sum_{(n : \mathbb{N})} , \textbf{Fin}(n) \simeq A \rVert
  \end{equation}
\end{lemma}
\begin{proof}
  By proposition~\ref{equal-strength}, we only need to provide conversion
  functions between the two sides of the equivalence.
  For left to right, we have proposition~\ref{bishop-equiv}.
  For right to left, we can transport along the equivalence, meaning we need
  only prove that \(\textbf{Fin}(n)\) is Bishop finite.
  We know this to be true, since it is split enumerable
  (proposition-\ref{fin-split-enum}).
\end{proof}
\subsubsection{Closure}
The closure proofs for cardinal finiteness are especially easy.
In contrast to manifest enumerability, under the propositional truncation we
have a full proof of bishop finiteness, meaning that all of the closure proofs
carry over.
\begin{lemma}
  Cardinal finiteness is closed under dependent and non-dependent sums,
  products, and functions.
\end{lemma}
\begin{proof}
  All closure functions can be lifted under propositional truncation.
  Therefore, cardinal finiteness has the same closure properties as manifest
  bishop finiteness.
\end{proof}
\subsubsection{Strength}
We can eliminate from cardinal finiteness in a coherently constant way:
effectively, anything which doesn't ``notice'' the order of the internal support
list is usable.

First, even though the cardinality of the type is hidden under the truncation,
we can pull it out, as it is constant through permutations of the list.
\begin{theorem}
  Given a cardinally finite type, we can derive the type's cardinality, as well
  as a propositionally truncated proof of equivalence with \(\textbf{Fin}\)s of
  the same cardinality.
  \begin{equation}
    \mathcal{C}(A) \rightarrow \sum_{(n : \mathbb{N})} , \lVert \textbf{Fin}(n) \simeq A \rVert
  \end{equation}
\end{theorem}
\begin{proof}
  We eliminate from the proof of cardinal finiteness via the eliminator
  described in proposition~\ref{elim-non-prop}.
  We pass a function \(\text{alg}\) which preserves the first part of the pair,
  and truncates the second.
  In Agda:
  \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{trunc-alg}

  It remains only to show that this function is coherently constant.
  \begin{equation}
    \prod_{(x : \sum_{(n : \mathbb{N})} , \textbf{Fin}(n) \simeq A)} ,
    \prod_{(y : \sum_{(m : \mathbb{N})} , \textbf{Fin}(m) \simeq A)} ,
    \text{alg}(x) \equiv \text{alg}(y)
  \end{equation}

  We will first address the first component of the output.
  To show that \(n\) and \(m\) are equal, we first observe that
  \begin{equation}
    \textbf{Fin}(n) \equiv \textbf{Fin}(m)
  \end{equation}
  since both are equivalent to \(A\).
  Relying on the fact that \(\textbf{Fin}\) is injective
  (proposition~\ref{fin-inj}), we can derive that \(n\) and \(m\) are equal.

  Since the second components of the output are propositions, they are
  definitionally equal.
  \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{const-alg}
\end{proof}

\begin{lemma} \label{fin-inj}
  \(\textbf{Fin}\) is injective.
\end{lemma}
\begin{proof}
  Let \(n\) and \(m\) be natural numbers.
  We want to derive a proof of \(n \equiv m\) from \(\textbf{Fin}(n) \equiv
  \textbf{Fin}(m)\).

  We prove by contradiction (since the property we are interested in is
  decidable, this is valid constructively).
  When \(n\) does not equal \(m\), there are two possible cases.
  \begin{equation}
    \neg (n \equiv m) \implies
    \begin{cases}
      m \equiv 1 + n + k, n < m  \\
      n \equiv 1 + m + k, m < n
    \end{cases}
  \end{equation}
  Take the first case, without loss of generality.
  \todo[inline]{Finish prose for this proof}

  The Agda version of this proof is in the appendix (\ref{agda-fin}).
\end{proof}

Another property which does not rely on an internal order is decidable equality.
\begin{theorem} \label{cardinal-finite-discrete}
  Any cardinal-finite set has decidable equality.
\end{theorem}
\begin{proof}
  We first show that decidable equality is a proposition itself.
  We know that if a type \(A\) is a proposition, then the decision over that
  type is also a proposition.
  Then, via Hedberg's theorem, we know that any type with decidable equality is
  a set, meaning that paths in that type are themselves propositions.
  Therefore we can derive that a decision of equality on elements with decidable
  equality is a proposition, and by function extensionality we see that
  decidable equality is itself a proposition.

  That is enough to pull it out of the truncation: since any Bishop-finite type
  has decidable equality, and decidable equality is a proposition, we conclude
  that cardinal-finite types have decidable equality.
\end{proof}
\subsubsection{Relation to Manifest Bishop Finiteness}
Cardinal finiteness tells us that there is an isomorphism between a type and
\(\mathbf{Fin}\); it just doesn't tell us \emph{which} isomorphism.
To take a simple example, \(\mathbf{Bool}\) has 2 possible isomorphisms with the
set \(\mathbf{Fin}(2)\): one where false maps to 0, and true to 1; and another
where false maps to 1 and true to 0.

To convert from Cardinal finiteness to Bishop finiteness, then, requires that we
supply enough information to identity a particular isomorphism.
A total order is sufficient here: it will give us enough to uniquely order the
support list invariant under permutations.
This tells us what we already knew in the introduction: manifest Bishop
finiteness is cardinal finiteness plus an order.

\begin{theorem}
  Any cardinal finite type with a (decidable) total order is manifestly Bishop
  finite.
\end{theorem}
\begin{proof}
  This proof is quite involved, and will rely on several subsequent lemmas, so
  we will give only its outline here.
  \begin{itemize}
    \item First, we will convert to manifest enumerability: knowing that the
      underlying type is discrete (theorem~\ref{cardinal-finite-discrete}) we can
      go from manifest enumerability to split enumerability
      (lemma~\ref{manifest-enum-to-split}), and subsequently to manifest Bishop
      finiteness (lemma~\ref{split-enum-to-manifest-bishop}).
    \item To convert to manifest enumerability, we need to provide a support
      list: this cannot simply be the support list hidden under the truncation,
      since that would violate the hiding promised by the truncation.
      Instead, we sort the list (using insertion sort).
      We must, therefore, prove that insertion sort is invariant under all
      support lists in cardinal finiteness proofs.
    \item We show that all support lists in cardinal finiteness proofs are
      permutations of each other.
    \item And then we show that insertion sort is invariant under permutations.
  \end{itemize} 
\end{proof}
Now we will build up the toolkit we need to perform the above steps.
First, permutations.
\begin{definition}[List Permutations]
  We say that two lists are permutations of each other if there is an
  isomorphism between membership proofs.
  \begin{equation}
    \mathit{xs} \leftrightsquigarrow \mathit{ys} = \prod_{(x : A)} , x \in \mathit{xs} \iff x \in \mathit{ys}
  \end{equation}
\end{definition}
We also prove some of the identities you might expect with regards to
permutations, all in the appendix.
In particular, we prove that they form an equivalence relation.

Next, we will use the following definition of insertion sort:
\ExecuteMetaData[agda/Data/List/Sort.tex]{sort}
\begin{lemma}
  Insertion sort is invariant under permutations.
  \begin{equation}
    \mathit{xs} \leftrightsquigarrow \mathit{ys} \implies \text{sort}(\mathit{xs}) \equiv \text{sort}(\mathit{ys})
  \end{equation}
\end{lemma}
\begin{proof}
  We first prove that insertion sort does indeed sort its input: i.e. it returns
  a sorted list that is a permutation of its input.
  Then we show that any two sorted lists which are permutations of each other
  are equal.
\end{proof}
\todo[inline]{Agda code for proofs}
\subsection{Kuratowski Finiteness} \label{kuratowski}
Finally we arrive at Kuratowski finiteness
\cite{kuratowskiNotionEnsembleFini1920}.
\subsubsection{The Kuratowski Set as a Higher Inductive Type}
We start with the Kuratowski finite sets, which are free join semilattices (or,
equivalently, free commutative idempotent monoids).
HITs are required to define this type
\cite{altenkirchDefinableQuotientsType2011}; however we have two possible
candidates for the definition.
The first is a direct translation of ``free commutative idempotent monoid'' into
a type:
\begin{equation}
\begin{aligned}
  \mathcal{K}(A) &=  \\
    &\left\vert \; \text{singleton} : A \rightarrow \mathcal{K}(A) \right; \\
    &\left\vert \; \cdot \cup \cdot : \mathcal{K}(A) \times \mathcal{K}(A) \rightarrow \mathcal{K}(A) \right; \\
    &\left\vert \; \emptyset : \mathcal{K}(A) \right; \\
    &\left\vert \; \cup \text{-assoc} : \Pi \left( \mathit{xs}, \mathit{ys}, \mathit{zs}:\mathcal{K}(A)\right) , (\mathit{xs} \cup \mathit{ys}) \cup \mathit{zs} \equiv \mathit{xs} \cup (\mathit{ys} \cup \mathit{zs}) \right; \\
    &\left\vert \; \cup \text{-commutative} : \Pi (\mathit{xs}, \mathit{ys}: \mathcal{K}(A)) , \mathit{xs} \cup \mathit{ys} \equiv \mathit{ys} \cup \mathit{xs} \right; \\
    &\left\vert \; \cup \text{-idempotent} : \Pi (\mathit{xs}: \mathcal{K}(A)) , \mathit{xs} \cup \mathit{xs} \equiv \mathit{xs} \right; \\
    &\left\vert \; \cup \text{-identity} : \Pi (\mathit{xs}: \mathcal{K}(A)) , \emptyset \cup \mathit{xs} \equiv \mathit{xs} \right; \\
    &\left\vert \; \text{trunc} : \Pi (\mathit{xs}, \mathit{ys}: \mathcal{K}(A)) , \Pi (p, q : \mathit{xs} \equiv \mathit{ys}) , p \equiv q \right;
\end{aligned}
\end{equation}
\todo[inline]{Have we already described the trunc constructor?}

However this proves cumbersome to work with.
Instead, we build on the free monoid (the list), and add the required equations
we need.
\begin{equation}
\begin{aligned}
  \mathcal{K}(A) &=  \\
    &\left\vert \; \cdot \dblcolon \cdot : A \times \mathcal{K}(A) \rightarrow \mathcal{K}(A) \right; \\
    &\left\vert \; [] : \mathcal{K}(A) \right; \\
    &\left\vert \; \text{com} : \Pi (x, y: A) , \Pi (\mathit{xs}: \mathcal{K}(A)) , x \dblcolon y \dblcolon \mathit{xs} \equiv y \dblcolon x \dblcolon \mathit{xs} \right; \\
    &\left\vert \; \text{dup} : \Pi (\mathit{x}: A) , \Pi (\mathit{xs}: \mathcal{K}(A)) , x \dblcolon x \dblcolon \mathit{xs} \equiv x \dblcolon \mathit{xs} \right; \\
    &\left\vert \; \text{trunc} : \Pi (\mathit{xs}, \mathit{ys}: \mathcal{K}(A)) , \Pi (p, q : \mathit{xs} \equiv \mathit{ys}) , p \equiv q \right;
\end{aligned}
\end{equation}
Since associativity and identity (the monoid laws) are automatically satisfied
by the list structure, we do not need to specify them as paths.
Furthermore, it is simpler to specify the rest of the paths on the heads of the
lists, rather than on the whole structure.
That said, the two definitions are equivalent.
\todo[inline]{Maybe add sketch of proof here?}
\subsubsection{Elimination}
With Agda's dependent pattern matching, we get an eliminator automatically from
the definition of an inductive type.
We will define two further eliminators here.

\begin{equation}
  \begin{aligned}
    \phi &: \mathcal{K}(A) \rightarrow \text{Type} \\
    \text{trunc}^\phi &: \pifun{x}{\mathcal{K}(A)} , \text{isSet}(\phi(x)) \\
    []^\phi &: \phi([]) \\
    \dblcolon^\phi &: \pifun{x}{A} , \pifun{xs}{\mathcal{K}(A)} , \pifun{p}{\phi(\mathit{xs})} , \phi(x \dblcolon \mathit{xs}) \\
    \text{com}^\phi &: \prod_{(x , y : A)} , \pifun{xs}{\mathcal{K}(A)} , \pifun{p}{\phi(\matit{xs})} , x \dblcolon^\phi (y \dblcolon \mathit{xs}) (y \dblcolon^\phi \mathit{xs} \; p) \equiv y \dblcolon^\phi (x \dblcolon \mathit{xs}) (x \dblcolon^\phi \mathit{xs} \; p) \\
    \text{dup}^\phi &: \pifun{x}{A} , \pifun{xs}{\mathcal{K}(A)} , \pifun{p}{\phi(\mathit{xs})} , x \dblcolon^\phi (x \dblcolon \mathit{xs}) (x \dblcolon^\phi \mathit{xs} \; p) \equiv x \dblcolon^\phi \mathi{xs} \; p
  \end{aligned}
\end{equation}

\begin{equation} \label{kuratowski-rec}
  \text{rec} \left\{ \begin{aligned}
    \phi &: \text{Type} \\
    \text{trunc}^\phi &: \text{isSet}(\phi) \\
    []^\phi &: \phi \\
    \dblcolon^\phi &: \pifun{x}{A} , \pifun{xs}{\mathcal{K}(A)} , \phi \\
    \text{com}^\phi &: \prod_{(x , y : A)} , \pifun{xs}{\mathcal{K}(A)} , x \dblcolon^\phi (y \dblcolon^\phi \mathit{xs}) \equiv y \dblcolon^\phi (x \dblcolon^\phi \mathit{xs}) \\
    \text{dup}^\phi &: \pifun{x}{A} , \pifun{xs}{\mathcal{K}(A)} , x \dblcolon^\phi (x \dblcolon^\phi \mathit{xs}) \equiv x \dblcolon^\phi \mathit{xs}
  \end{aligned} \right.
\end{equation}

\todo[inline]{Expand on these eliminators}

\subsubsection{Membership}
We can define membership of a Kuratowski set using these eliminators.
First we define the \(\Diamond\) type, described in the appendix.
\begin{equation}
  \Diamond P = \left\{ \begin{aligned}
    \phi &= \sum_{(x : \text{Type})} , \text{isProp}(x) \\
    \text{trunc}^\phi &= \text{isSetHProp} \\
    []^\phi &= \bot , \text{isProp}\bot \\
    x \dblcolon^\phi \mathit{xs} &= \lVert P(x) \uplus \text{fst}(\mathit{xs}) \rVert , \text{squash} \\
    \text{com}^\phi &= \dots \\
    \text{dup}^\phi &= \dots
  \end{aligned} \right.
\end{equation}
We can define membership then in terms of this.
\begin{definition}[Kuratowski Membership]
  \begin{equation}
    x \in \mathit{xs} = \Diamond (\equiv x) \mathit{xs}
  \end{equation}
\end{definition}

From this we can define Kuratowski finiteness.
\begin{definition}[Kuratowski Finiteness]
  A type is Kuratowski finite iff there exists a Kuratowski Set which contains
  all of its elements.
  \begin{equation}
    \mathcal{K}^{f}(A) = \sum_{(\text{support} : \mathcal{K}(A))} , \pifun{x}{A} , x \in \text{support}
  \end{equation}
\end{definition}

\subsubsection{Strength}
Since the Kuratowski set is a departure in structure from our previous
list-based notions of finiteness, it makes sense to first look for the closest
list-based analogue.
As it turns out, that analogue is manifestly enumerable finiteness, with the
order removed.
\begin{theorem}
  A proof of Kuratowski finiteness is equivalent to a propositionally truncated
  proof of enumerability.
  \begin{equation}
    \mathcal{K}^f(A) \simeq \lVert \mathcal{E}(A) \rVert
  \end{equation}
\end{theorem}
\begin{proof}
  We prove by way of an isomorphism.
  In the first direction (from \(\mathcal{K}\) to \(\mathcal{E}\)), we use the
  eliminator in \ref{kuratowski-rec}.
  Because we are eliminating into a proposition, we need not prove that the
  function obeys the path constructors.
  The remaining cases to deal with involve converting the Kuratowski set to a
  list, and the cover proof to its equivalent cover proof on lists.
  \ExecuteMetaData[agda/Cardinality/Finite/Kuratowski.tex]{to-enum}

  To go the other direction, we first need to prove that a proof of Kuratowski
  finiteness is a proposition.
  \todo[inline]{Explain rest of this proof}
\end{proof}
Based on previous proofs, we can derive that if we add decidability to a
Kuratowski finite type we retrieve cardinal finiteness.
\begin{lemma}
  Any Kuratowski finite set with decidable equality is cardinally finite.
\end{lemma}
\begin{proof}
  \todo[inline]{Proof}
\end{proof}
\begin{lemma}
  Kuratowski finiteness is equivalent to a truncated surjection from some prefix
  of the natural numbers.
  \begin{equation}
    \mathcal{K}^f(A) \simeq \left\Vert \sum_{(n : \mathbb{N})} , \left(  \mathbf{Fin}(n) \twoheadrightarrow A \right) \right\Vert
  \end{equation}
\end{lemma}
\begin{proof}
  \todo[inline]{Prose for proof}
\end{proof}
\subsubsection{Topos}
At this point, we see that a ``decidable Kuratowski finite set'' is precisely
equivalent to a cardinal finite set.
From this, we can lift over all of the properties of cardinal finite sets.
In particular, we see that decidable Kuratowski finite sets form a \emph{topos}.
\todo[inline]{Fill in rest}
\subsubsection{Closure}
\begin{theorem}
  Kuratowski finite sets are closed under \(\sum\).
\end{theorem}
\begin{proof}
  \todo[inline]{Proof}
\end{proof}
\section{Species and Containers} \label{species-and-containers}
Containers \cite{abbottContainersConstructingStrictly2005} give us a general way
to construct functors from generic building-blocks.
\begin{definition}[Container]
  A container is a dependent pair of shapes, and positions indexed by those
  shapes.
  \begin{equation}
    \mathbf{Container} = \sum_{(\text{shape} : \text{Type})} , \text{shape} \rightarrow \text{Type}
  \end{equation}
  \ExecuteMetaData[agda/Data/Container.tex]{def}
  We ``interpret'' a container into a pair of a specific shape, and a function
  which goes from positions to contents.
  \begin{equation}
    \llbracket S , P \rrbracket = \prod_{(X : \text{Type})} , \sum_{(s : S)} , \left( P(s) \rightarrow X \right)
  \end{equation}
  \ExecuteMetaData[agda/Data/Container.tex]{interp}
\end{definition}
For instance, the shape of lists is given by \(\Nat\), and the positions are then
\(\AgdaDatatype{Fin}\)s indexed by that \(\Nat\).
\ExecuteMetaData[agda/Data/Container/List.tex]{def}
And here is the definition of foldr for container-defined lists:
\ExecuteMetaData[agda/Data/Container/List.tex]{foldr}
\subsection{Containers For Finiteness}
The definition for the container-based finiteness predicates are almost
identical to their list-based counterparts.
\ExecuteMetaData[agda/Cardinality/Container/Finite.tex]{split-enum}
One advantage that they have, however, is that conversion back and forth between
the relation-based finiteness predicates is very straightforward.
This allows us to prove the following.
\begin{theorem}
  Manifest bishop finiteness is equivalent to an equivalence with some \(\mathbf{Fin}\).
  \begin{equation}
    \mathcal{B}(A) \simeq \sum_{(n : \mathbb{N})} , \mathbf{Fin}(n) \simeq A
  \end{equation}
\end{theorem}
\begin{proof}
  This proof is the composition of two equivalences.
  First, we must show that the list-based manifest Bishop finiteness is
  equivalent to the container-based definition, and then we must show that the
  container-based definition of manifest Bishop finiteness is equivalent to the
  equivalence.
  \todo[inline]{proof}
\end{proof}

\begin{lemma}
  Manifest enumerability is equivalent to a surjection from some \(\mathbf{Fin}\).
  \begin{equation}
    \mathcal{E}(A) \simeq \sum_{(n : \mathbb{N})} , \left(  \mathbf{Fin}(n) \twoheadrightarrow A \right)
  \end{equation}
\end{lemma}
\begin{proof}
\end{proof}

\begin{lemma}
  Split enumerability is equivalent to a split surjection from some \(\mathbf{Fin}\).
  \begin{equation}
    \mathcal{E}!(A) \simeq \sum_{(n : \mathbb{N})} , \left(  \mathbf{Fin}(n) \twoheadrightarrow ! A \right)
  \end{equation}
\end{lemma}
\begin{proof}
\end{proof}

\section{Infinite Cardinalities} \label{infinite-cardinalities}
While the previous section purported to be about finite sets, we can now see
that it was really only studying surjections and isomorphisms of different
flavours between types and \(\mathbf{Fin}\).
The natural next question which arises, then, is if we can extend that work to
surjections and isomorphisms with \(\mathbb{N}\).
In more standard language, what we're referring to here is of course
countable infinity and related concepts.
However, as with finite types, the ``countably infinite'' types have more
varieties in constructive mathematics than their classical counterparts.
\subsection{Split Countable Types}
Our first foray into the world of countable types will be a straightforward
analogue to the split enumerable types.
We need change only one element: instead of a support \emph{list}, we instead
have a support \emph{stream}, which is its infinite, coinductive counterpart.
\begin{definition}[Stream]
  We will work with two isomorphic definitions of streams.
  The first is the following:
  \begin{equation}
    \mathbf{Stream}(A) = \mathbb{N} \rightarrow A
  \end{equation}
  Conceptually, a stream is like a list without an end.
  Of course, such a type can not be defined in the same way as a list: it would
  be impossible to construct a value, as inductive types do not admit
  infinitely-sized inhabitants.
\end{definition}
We can now define the split countable types.
\begin{definition}[Split Countability]
  \begin{equation}
    \mathcal{E}!(A) = \sum_{(\mathit{xs} : \mathbf{Stream}(A))} , \prod_{(x : A)} , x \in \mathit{xs}
  \end{equation}
\end{definition}
\subsubsection{\(\Sigma\) Closure}
We know that countable infinity is not closed under the exponential (function
arrow), so the only closure we need to prove is \(\Sigma\) to cover all of
what's left.
To do this we have to take a slightly different approach to the functions we
defined before.
Figure~\ref{pairings} illustrates the reason why: previously, we used the
``Cartesian'' product pairing for each support list.
This diverges if the first list is infinite, never exploring anything other than
the first element in the second list.
Instead, we use here the cantor pairing function, which performs a breadth-first
search of the pairings of both lists.
\begin{figure}
  \centering
  \begin{subfigure}[b]{.5\linewidth}
    \begin{tikzcd}[sep=tiny,font=\footnotesize]
      (1,e) \ar[rdddd, out=-45, in=135] & (2,e) \ar[rdddd, out=-45, in=135] & (3,e) \ar[rdddd, out=-45, in=135] & (4,e) \ar[rdddd, out=-45, in=135] & (5,e)        \\
      (1,d) \ar[u]     & (2,d) \ar[u]    & (3,d) \ar[u]    & (4,d) \ar[u]    & (5,d) \ar[u] \\
      (1,c) \ar[u]     & (2,c) \ar[u]    & (3,c) \ar[u]    & (4,c) \ar[u]    & (5,c) \ar[u] \\
      (1,b) \ar[u]     & (2,b) \ar[u]    & (3,b) \ar[u]    & (4,b) \ar[u]    & (5,b) \ar[u] \\
      (1,a) \ar[u]     & (2,a) \ar[u]    & (3,a) \ar[u]    & (4,a) \ar[u]    & (5,a) \ar[u]
    \end{tikzcd}
    \caption{Cartesian}
    \label{cartesian}
  \end{subfigure}%
  \begin{subfigure}[b]{.5\linewidth}
    \begin{tikzcd}[sep=tiny,font=\footnotesize]
      (1,e) \ar[dr] & (2,e) \ar[dr]  & (3,e) \ar[dr]    & (4,e) \ar[dr] & (5,e) \\
      (1,d) \ar[dr] & (2,d) \ar[dr]  & (3,d) \ar[dr]    & (4,d) \ar[dr] & (5,d) \ar[u] \\
      (1,c) \ar[dr] & (2,c) \ar[dr]  & (3,c) \ar[dr]    & (4,c) \ar[dr] & (5,c) \ar[uul] \\
      (1,b) \ar[dr] & (2,b) \ar[dr]  & (3,b) \ar[dr]    & (4,b) \ar[dr] & (5,b) \ar[uuull, out=130, in=-50] \\
      (1,a) \ar[u]  & (2,a) \ar[uul, out=130, in=-50] & (3,a) \ar[uuull, out=130, in=-50] & (4,a) \ar[uuuulll, out=130, in=-50] & (5,a) \ar[uuuulll, out=130, in=-50]
    \end{tikzcd}
    \caption{Cantor}
    \label{cantor}
  \end{subfigure}
  \caption{Two possible products for the sets \(\left[ 1 \dots 5 \right]\) and
    \(\left[  a \dots e \right]\)}
  \label{pairings}
\end{figure}
\begin{theorem} \label{split-countability-sigma}
  Split countability is closed under \(\Sigma\).
\end{theorem}
\begin{proof}
  Let \(\mathcal{X}\) be the proof of split countability on \(A\), and
  \(\mathcal{Y}\) be a function of the following type:
  \begin{equation}
    \mathcal{Y} : \prod_{(x : A)} , \mathcal{E}!(U(x))
  \end{equation}
  Where \(U\) is a type family on \(A\).
  Our task is to provide the following:
  \begin{equation}
    \mathcal{E}!\left(\sum_{(x : A)} , U(x)\right)
  \end{equation}

  This final proof consists of the support stream, and the proof that the
  support stream covers the input.

  As mentioned, we will have to use a more sophisticated pairing function than
  the Cartesian product we used before.
  We instead will mirror the pattern in figure~\ref{cantor}.
  To greatly simplify the algorithm, we will produce an intermediate stream of
  lists which consists of the diagonals in the diagram.
  We then concatenate these streams into the final support stream.

  Here is the algorithm in Agda:
  \ExecuteMetaData[agda/Cardinality/Container/Infinite.tex]{conv}
  \todo[inline]{Finish prose for this proof}
\end{proof}
\begin{lemma}
  Split countability is closed under non-dependent product and sum.
\end{lemma}
\begin{proof}
  Follows from theorem~\ref{split-countability-sigma}.
\end{proof}
\subsubsection{Kleene Star and Plus}
While we lose some closures with the inclusion of infinite types, we gain some
others.
In particular, we have the Kleene star and plus.
These allow us to introduce recursion into countable types, by way of lists.
\begin{definition}[Kleene Lists]
  It is often useful to have two mutually-defined list types (one for
  possibly-empty lists and one for nonempty lists).
  Such a definition closely mirrors the Kleene star and plus.
  \begin{equation}
    \begin{aligned}
      A^+ &= \\
          &\left\vert \; \text{head} : A^+ \rightarrow A \right; \\
          &\left\vert \; \text{tail} : A^+ \rightarrow A^\star \right; \\
      A^\star &= \\
          &\left\vert \; [] : A^\star \right; \\
          &\left\vert \; [\cdot] : A^+ \rightarrow A^\star \right;
    \end{aligned}
  \end{equation}
\end{definition}
Our definition of split countability is closed under these types in particular.
\begin{theorem}
  Split countability is closed under Kleene star and plus.
\end{theorem}
\begin{proof}
  \tot[inline]{Proof}
\end{proof}
\subsection{Manifest Countability}
As we can quotient out the position information with finite types, so can we with
countable types.
\todo[inline]{fill in rest here}
\section{Practical Uses} \label{practical}
\todo[inline]{%
\begin{itemize}
  \item Synthesize functions according to specifications.
  \item Nonexhaustive search for the infinite types.
  \item Lift quantifiers over types other than bishop (i.e. K). Show that
    quantifiers have to be different (i.e. existence has to be unique).
  \item Big operators\cite{gustafssonCountingTypeIsomorphisms}
    \cite{gustafssonFoldableContainersDependent}
\end{itemize}
}
The theory of finite types in constructive mathematics, and in HoTT in
particular, is rich and interesting, as we hope we have demonstrated thus far.
As well as being theoretically interesting, however, the proofs and combinators
we have defined here are \emph{practically} useful.
\subsection{Omniscience}
\todo[inline]{Put definition for decidable somewhere}
In this section we are interested in restricted forms of the limited principle
of omniscience \cite{myhillErrettBishopFoundations1972}.
\begin{definition}[Limited Principle of Omniscience]
  For any type \(A\) and predicate \(P\) on \(A\), the limited principle of
  omniscience is as follows:
  \begin{equation}
    \left( \prod_{(x : A)} , \mathbf{Dec}(P(x)) \right) \rightarrow \mathbf{Dec} \left( \sum_{(x : A)} , P(x) \right)
  \end{equation}
  In other words, for any decidable predicate the existential quantification of
  that predicate is also decidable.
\end{definition}
The limited principle of omniscience is non-constructive, but individual types
can themselves satisfy omniscience.
In particular, \emph{finite} types are omniscient.
Omniscience is defined in Agda like so:
\ExecuteMetaData[agda/Relation/Nullary/Omniscience.tex]{omniscient}

There is also a universal form of omniscience, which we call exhaustibility.
\begin{definition}[Exhaustibility]
  We say a type \(A\) is exhaustible if, for any decidable predicate \(P\) on
  \(A\), the universal quantification of the predicate is decidable.
  \begin{equation}
    \left( \prod_{(x : A)} , \mathbf{Dec}(P(x)) \right) \rightarrow \mathbf{Dec} \left( \prod_{(x : A)} , P(x) \right)
  \end{equation}
  \ExecuteMetaData[agda/Relation/Nullary/Omniscience.tex]{exhaustible}
\end{definition}
Omniscience is stronger than exhaustibility, as we can derive the latter from
the former:
\begin{lemma} \label{omniscient-is-exhaustible}
  Any omniscient type is exhaustible.
\end{lemma}
\begin{proof}
  \todo[inline]{Proof}
  \ExecuteMetaData[agda/Relation/Nullary/Omniscience.tex]{weaken}
  \todo[noline]{stability in appendix}
\end{proof}
We cannot derive, however, that any exhaustible type is omniscient: to do so
would require us to choose a representative from an arbitrary type, which is not
possible constructively.

\subsubsection{Split Enumerability}
We now show that split enumerable types are omniscient.
\begin{theorem}
  Any split enumerable type is omniscient.
\end{theorem}
\begin{proof}
  Let \(E_A\) be a proof of split enumerability on a type \(A\), and \(P\) be a
  decidable predicate on the type \(A\).
  First, we lift the decidable predicate onto one on lists, using the
  \(\Diamond\) type described earlier.
  Then, we run the predicate on the support list of \(E_A\).

  In the case that it is true for any member of the list, we return that element
  along with its proof.

  If the predicate fails for every member of the list, we say that the whole
  test has failed.
  To support this conclusion, we reason that if a predicate is false for every
  item in a list \(\mathit{xs}\), and some element \(x\) is in that list, then
  the predicate must be false for that item.

  In Agda:
  \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{exists-quant}
\end{proof}
\begin{theorem}
  Any split enumerable type is exhaustible.
\end{theorem}
\begin{proof}
  Via lemma~\ref{omniscient-is-exhaustible}.
\end{proof}

\subsubsection{Manifest Enumerability}
We now reach for the same quantification combinators we used for
split enumerable types.
Because neither of these quantifications ``care'' about the specific proof of
membership used, they can both be used.
\begin{theorem}
  For a decidable predicate \(P\) on a type \(A\), if the type is manifestly
  enumerable, then we can decide the existential of \(P\).
\end{theorem}
\begin{proof}
  \todo[inline]{Proof}
\end{proof}

\begin{theorem}
  For a decidable predicate \(P\) on a type \(A\), if the type is manifestly
  enumerable, then we can decide the universal of \(P\).
\end{theorem}
\begin{proof}
  \todo[inline]{Proof}
\end{proof}
\subsubsection{Cardinal Finiteness}
The quantification operations partially carry over.
\begin{theorem}
  For a decidable predicate \(P\) on a type \(A\), if the type is cardinal
  finite, then we can decide the universal of \(P\).
\end{theorem}
\begin{proof}
  \todo[inline]{Proof}
\end{proof}

An existential predicate is harder to express.
In previous sections, we simply tested each member of the set in turn, and
returned the first member which passed the test.
Clearly this operation can depend on the underlying order.
\begin{theorem}
  For a decidable predicate \(P\) on a type \(A\), if the type is cardinal
  finite, then we can decide the propositionally truncated existential of \(P\).
\end{theorem}
\begin{proof}
  \todo[inline]{proof}
\end{proof}
\begin{theorem}
  For a decidable predicate \(P\) on a type \(A\), which uniquely determines an
  input, we can decide the existential of \(P\).
\end{theorem}
\begin{proof}
  \todo[inline]{proof}
\end{proof}
\subsubsection{Kuratowski}
\begin{itemize}
\item Forall
\item Exists
\end{itemize}

\subsection{Synthesising Pattern-Matching Proofs}
In particular, they can automate large proofs by analysing every possible case.
In \cite{firsovDependentlyTypedProgramming2015}, the \(\AgdaDatatype{Pauli}\)
group is used as an example.
\ExecuteMetaData[agda/Data/Pauli.tex]{def}
For this type, there are several practical tasks we hope to achieve with the
help of our finiteness combinators:
\begin{itemize}
  \item Define decidable equality on the type
  \item Define the group operation on the type
  \item Prove properties about the group operation
\end{itemize}
Unfortunately, the simple pattern-matching way to do many of these tasks is
prohibitively verbose.
As \(\AgdaDatatype{Pauli}\) has 4 constructors, \(n\)-ary functions on
\(\AgdaDatatype{Pauli}\) may require up to \(4^n\) cases.
A proof of decidable equality is one such function: it can be seen fully
worked-through in the appendix.

The alternative is to derive the things we need from \(\Enum\) somehow.
First, then, we need an instance for \(\AgdaDatatype{Pauli}\):
\ExecuteMetaData[agda/Data/Pauli.tex]{inst}
From here we can already derive decidable equality:
\ExecuteMetaData[agda/Data/Pauli.tex]{good-eq}

Next, we want to prove some things about the group operation itself:
\ExecuteMetaData[agda/Data/Pauli.tex]{group-op}
The first property is the following:
\ExecuteMetaData[agda/Data/Pauli.tex]{cancel}
Tough the proof is simple enough to write by hand, it's a good place to start
for demonstrating the automation technique.

First, we will witness the fact that if a property is true for every element in
a finite type's support list, it is true for every element in the type (and vice
versa).
\ExecuteMetaData[agda/Cardinality/Finite/Search.tex]{all-comb}
We can also lift a decidable property on elements of the type to a decidable
property on \emph{lists} of elements of the type.
\ExecuteMetaData[agda/Data/List/Relation/Unary.tex]{dec-all}
Combining these, we can turn a decidable predicate on a set into a decidable
\emph{universal property} on the set.
\ExecuteMetaData[agda/Cardinality/Finite/Search.tex]{lift-forall}

To use this as a proof, we can run the decision procedure at type checking,
using the following:
\ExecuteMetaData[agda/Cardinality/Finite/Search.tex]{auto}
This function uses the \(\AgdaDatatype{True}\) type, defined like so:
\ExecuteMetaData[agda/Relation/Nullary.tex]{true-def}
For a decision \(d\), \(\AgdaDatatype{True}\; d\) will resolve to \(\top\) when
the decision is \(\AgdaInductiveConstructor{yes}\) (i.e. when the proposition is
true), or \(\bot\) if the decision is \(\AgdaInductiveConstructor{no}\).
Then, we require an instance of the resulting type to be in scope: this
requirement can only be satisfied for \(\top\), thereby providing a proof that
the decision was indeed \(\AgdaInductiveConstructor{yes}\).

And finally we can prove the property we wanted to on \(\AgdaDatatype{Pauli}\)
like so:
\ExecuteMetaData[agda/Data/Pauli.tex]{noninst-prf}

As a quick aside, when the property is \emph{not} true, for instance:
\ExecuteMetaData[agda/Data/Pauli.tex]{contra}
Agda will fail by not finding an instance of \(\bot\).
The error message specifically is:
\begin{displayquote}
  \verb+No instance of type ⊥ was found in scope.+
\end{displayquote}
We can actually display a counterexample, by defining a custom empty type
parameterised by the counterexample itself.
\ExecuteMetaData[agda/Cardinality/Finite/Search.tex]{counterexample-def}
Combined with some other changes to the combinators, this will give the
following more helpful error message:
\begin{displayquote}
  \verb+No instance of type Counterexample X was found in scope.+
\end{displayquote}
\subsection{Instances}
Running decision procedures during typechecking isn't the only use of instances:
we can also use them similarly to how typeclasses are used in Haskell, to
automate simple derivation procedures.
For instance, in Haskell we can define a class for generating textual
representations of data:
\ExecuteMetaData[haskell/Introduction.tex]{show-cls}
With instances that look like this:
\ExecuteMetaData[haskell/Introduction.tex]{show-inst}
We can have instances \emph{depend} on other instances, like the following:
\ExecuteMetaData[haskell/Introduction.tex]{show-deriv}
In this way, it gets rid of a tremendous about of boilerplate for more complex
types, automatically picking out the correct instance when it can.

In Agda, as in Haskell, the problem of overlapping instances often presents
itself when more complex typeclass-based machinery is used.
While there are ways to deal with overlapping instances, it is usually better to
avoid them altogether, which is precisely what we do in our approach.
We provide instances for the semiring-based combinators we defined in
\todo[noline]{reference here}
, and require the user to provide an instance for their own type.
We don't, for example, provide an instance that itself searches for surjections
from other instances: this instance is not uniquely determined.

After all of that, the change to our \(\AgdaFunction{∀↯}\) is simple: we just
change the first parameter from explicit to an instance.
\ExecuteMetaData[agda/Cardinality/Finite/Search.tex]{auto-inst}

The change to the \(\AgdaDatatype{Pauli}\) proof of cancellation isn't
groundbreaking at first:
\ExecuteMetaData[agda/Data/Pauli.tex]{cancel-auto}
But the real benefit is now we can automate proofs over \emph{tuples}, allowing
us to prove, for instance, commutativity.
\ExecuteMetaData[agda/Data/Pauli.tex]{comm-uncurried}
\subsection{Generic Currying and Uncurrying}
While we have arguably removed the bulk of the boilerplate from the automated
proofs, there is still the case of the ugly noise of currying and uncurrying.
In this section, we take inspiration from
\cite{allaisGenericLevelPolymorphic2019} to develop a small interface to generic
\(n\)-ary functions and properties.
What we provide here differs from that work in the following ways:
\begin{itemize}
  \item Our generic representation can handle dependent \(\sum\) and \(\prod\)
    types (rather than their non-dependent counterparts, \(\times\) and
    \(\rightarrow\)).
    This extension was necessary for our use case: it is mentioned in the paper
    as the obvious next step. 
  \item We implement the curry-uncurry combinators as (verified) isomorphisms.
    Since we are in a cubical setting, this gives us equivalences between the
    types, a feature not available in standard Agda.
  \item We deal with implicit and instance arguments generically.
\end{itemize}
A full explanation of our implementation is beyond the scope of this work, but
we will mention the key parts here.
First, we define a function arrow generic over the application method:
\ExecuteMetaData[agda/Data/Product/N-ary.tex]{fun-arr}
Then, we prove that it is isomorphic to the normal function arrow:
\ExecuteMetaData[agda/Data/Product/N-ary.tex]{fun-iso}
This step will allow us to write the curry-uncurry proofs once, and then extend
them to the three different argument forms without difficulty.

We do the same with dependent function types:
\ExecuteMetaData[agda/Data/Product/N-ary.tex]{pi-iso}

Next, we need to make our machinery for multiple arguments.
Here we will define a generic tuple, indexed by some \(\Nat\).
As observed in \cite{allaisGenericLevelPolymorphic2019}, it's important to
\emph{not} implement this as an inductively defined vector, e.g.
\ExecuteMetaData[agda/Data/Vec.tex]{def}
As this will not give us the correct \(\eta\)-equality we need for unification.
\todo[noline]{Stick reference to iterated Vec definition}
Once we have a unification-friendly vector type, we can use it to implement our
generic (and level-polymorphic) tuples.
\ExecuteMetaData[agda/Data/Product/N-ary.tex]{tuple}
We decided against \(\top\)-terminated tuples, as the actual contents of the
tuple type are exposed in the interface.

Finally, we can prove isomorphisms between the curried and uncurried versions of
functions:
\ExecuteMetaData[agda/Data/Product/N-ary.tex]{curry-n-iso}
And dependent functions:
\ExecuteMetaData[agda/Data/Product/N-ary.tex]{curry-n-pi-iso}
With these combinators, we can implement the search functions in an
arity-generic way:
\ExecuteMetaData[agda/Cardinality/Finite/Search.tex]{all-nary}
And automate away our proofs:
\ExecuteMetaData[agda/Data/Pauli.tex]{prf-curried}

\begin{subappendices}

\section{Notation, Cubical Type Theory, and Agda}
Our chosen setting is Cubical Type Theory (CuTT)
\cite{cohenCubicalTypeTheory2016}, which is closely related to Homotopy Type
Theory (HoTT) \cite{hottbook}.
We will go through the basics and some notation here.
\subsection{Martin-Löf Type Theory}
CuTT has its roots in Martin-Löf intuitionistic type theory
\cite{martin-lofIntuitionisticTypeTheory1980}.
We will describe our notation for the MLTT parts of CuTT here (although our
notation is quite standard).

The empty or \(\mathbf{0}\) type will be denoted by \(\bot\), and the
\(\mathb{1}\) type by \(\top\).
The single inhabitant of \(\top\) is called \(\text{tt}\).

``not'' is expressed as follows:
\begin{equation}
  \neg A = A \rightarrow \bot
\end{equation}
In other words, to say a type is impossible we say we can derive \(\bot\) from
it.

We also use the \(\mathbf{2}\) type, which we call \(\textbf{Bool}\).
Its two canonical inhabitants are \(\text{true}\) and \(\text{false}\).

We will use the term type family (or simply ``family'') to mean a non-dependent
function with a codomain in Type.

Dependent function types, or \(\prod\)-types, use the following notation:
\begin{equation}
  \prod_{(x : A)} , B (x)
\end{equation}
In Agda we use a single arrow to denote the same:
\ExecuteMetaData[agda/Snippets/Introduction.tex]{pi}
When the type of the left-hand-side can be inferred, we are permitted to use the
following notation:
\ExecuteMetaData[agda/Snippets/Introduction.tex]{forall}
For non-dependent functions we use the single arrow (\(\rightarrow\)).

Dependent pairs, or \(\sum\)-types, use a similar notation to \(\prod\)-types.
\begin{equation}
  \sum_{(x : A)} , B (x)
\end{equation}
In Agda:
\ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma}
Again, we get a shorthand when the type of the first element can be inferred:
\ExecuteMetaData[agda/Snippets/Introduction.tex]{exists}
For non-dependent pairs we use the \(\times\) symbol.
We access the two fields of \(\sum\) types with \(\text{fst}\) and
\(\text{snd}\):
\begin{align}
  \text{fst} &: \sum_{(x : A)} , B (x) \rightarrow A \\
  \text{snd} &: \prod_{(x : \sum_{(y : A)} , B (y))} , B (\text{fst}(x))
\end{align}

Finally, the identity type is denoted by the triple equals (\(\equiv\)).
The normal equals (\(=\)) is reserved for definitions.

\subsection{Agda Basics}
Agda \cite{norellDependentlyTypedProgramming2008} is a dependently-typed
functional programming language, with roots in Haskell and ML.
The core type theory is based on MLTT, though support for CuTT was added in the
last year \cite{vezzosiCubicalAgdaDependently2019}.
\subsubsection{Declarations}
The anatomy of a simple Agda declaration is thus:
\ExecuteMetaData[agda/Snippets/Introduction.tex]{abool}
We first have a type signature: the name of the variable is to the left of the
colon, and its type to the right.
Below, we give an equation: as Agda is pure, it suffices to imagine that
everywhere the name on the left of the equals sign is seen, we can replace it
with the value on the right.
\subsubsection{Functions}
Functions are usually declared with pattern matching.
\ExecuteMetaData[agda/Snippets/Introduction.tex]{not}
Again, each clause is an equation.
Agda's purity will ensure that these clauses hold true definitionally\footnotemark.
\footnotetext{
  Actually, it is possible to construct functions whose equations
  hold propositionally, but not definitionally.
  An example might be the following:
  \ExecuteMetaData[agda/Snippets/Introduction.tex]{non-exact-split}
  This is actually attempted by a lot of Agda beginners: often the compiler will
  complain ``couldn't match type \(n\) with \(n + 0\)'', so the programmer will
  simply add it as a clause.
  Unfortunately, this actually over complicates things in the long run.
  Neither the second nor third clause here will be definitionally equal: if you
  have an obligation \(n \oplus 0 \qeq n\), the first pattern (\(0 \oplus m = m\))
  has to be ruled out first, which means you'll have to do a manual pattern
  match in your proof.
  Similarly for the third pattern: to prove the obligation (\((1 + n) \oplus m
  \qeq 1 + (n \oplus m)\)), \emph{both} preceding patterns have to be checked
  against first.

  This ordering requirement is introduced by overlapping patterns, and Agda
  warns against it.
  Knowing when to avoid it and when it doesn't matter is one of the aspects of
  programming and proving in Agda which is more of an art form.
}

As Agda is a functional language, functions can also be defined with lambda
expressions.
\ExecuteMetaData[agda/Snippets/Introduction.tex]{id}

Finally, ``mixfix'' operators play a large role in Agda's syntax.
We can have binary operators:
\ExecuteMetaData[agda/Snippets/Introduction.tex]{add}
Unary operators (here, a prefix increment operator):
\ExecuteMetaData[agda/Snippets/Introduction.tex]{unary-inc}
Ternary operators:
\ExecuteMetaData[agda/Snippets/Introduction.tex]{ifte}
And so on.
\subsubsection{Data Type Definitions}
We can define new data types in Agda by listing their constructors:
\ExecuteMetaData[agda/Snippets/Introduction.tex]{maybe}
\subsubsection{Termination}
\todo[inline]{Coinductive types, termination checking, etc.}
\subsubsection{Universe Levels}
All dependently-typed programming languages have to contend, at some point or
another, with Girard's paradox
\cite{girardInterpretationFonctionelleElimination1972}.
The paradox concerns the type of types: it's the type-theory analogue of
Russell's paradox.

Unfortunately, most of the ways to avoid the paradox add significant clutter to
the type theory, and increase the complexity of even simple declarations.
Agda's answer is no exception.

In Agda, every type is tagged with a ``universe level''.
The basic types like \(\AgdaDatatype{Bool}\) live in universe level 0, the type
of types lives in universe level 1, the type of types in universe level 1 lives
in universe level 2, and so on.
We won't go into it much more than that: usually there is an obvious way to
handle the universe level of a given expression ``properly'', and cases where
universe levels do something interesting come up rarely (although
section~\ref{practical} does touch on one).
We only mention it here to explain the expressions like \(A :
\AgdaDatatype{Type} \; a\) which show up in many examples: \(a\) there is the
universe level of the type \(A\).
\subsection{Cubical Basics}
This will not serve as an introduction CuTT,
but rather a brief overview of the parts of the theory we will be using.
For a fuller reference we direct the reader first to the ``HoTT Book''
\cite{hottbook}, then the paper on Cubical Type Theory
\cite{cohenCubicalTypeTheory2016}, and finally the paper on Cubical Agda
\cite{vezzosiCubicalAgdaDependently2019}.
\subsubsection{Paths}
Central to CuTT is the \emph{path} type.
This is a path in the topological sense, although we use it as a representation
for equality (which makes sense: central to HoTT is the
equivalence of these two concepts).
In most of the code presented here this will not behave in any different way to
the traditional identity type in Martin-Löf type theory: one exception to this
is pattern-matching, which is not available on paths in Cubical Agda currently.
As a result, we have to use its eliminator directly (if needed), although
usually higher-level combinators like \(\AgdaFunction{subst}\) are sufficient.

We will use \(\equiv\) as a notation for path types as well as the identity
types.
Although an identity type does exist in Cubical Type Theory, we will not use it,
so we do not fear much confusion on this point.
\subsubsection{Homotopy Levels}
Next, we have the concept of ``homotopy levels''.
These classify types in homotopy type theory according to their degree of
truncation.
Homotopy level 1, for instance, is the ``mere propositions'': the types for
which every element is equal to every other.
As a predicate, we can define \(\AgdaFunction{isProp}\) like so:
\begin{equation}
  \text{isProp}(A) = \prod_{(x : A)} , \prod_{(y : A)} , x \equiv y
\end{equation}
In Agda:
\ExecuteMetaData[agda/Snippets/Introduction.tex]{isProp}

One level up from the propositions are the (homotopy) sets.
These are types for which all paths are themselves propositions.
\begin{equation}
  \text{isSet}(A) = \prod_{(x : A)} , \prod_{(y : A)} , \text{isProp}(x \equiv y)
\end{equation}
In Agda:
\ExecuteMetaData[agda/Snippets/Introduction.tex]{isSet}

Finally, one level down from the propositions are the contractions.
These are types for which every element is equal to some specific element.
\begin{equation}
  \text{isContr}(A) = \sum_{(x : A)} , \prod_{(y : A)} , x \equiv y
\end{equation}
\ExecuteMetaData[agda/Snippets/Introduction.tex]{isContr}
\subsubsection{Univalence}
\todo[inline]{Flesh out}
To describe this we first need fibers:
\begin{equation}
  \text{fiber}_f(y) = \sum_{(x : A)} , f (x) \equiv y
\end{equation}
\ExecuteMetaData[agda/Snippets/Introduction.tex]{fiber}
From this we build the notion of an ``equivalence'': a function whose fibers are
all contractions.
\begin{equation}
  A \simeq B = \sum_{(f : A \rightarrow B)}, \prod_{(y : B)} , \text{isContr}(\text{fiber}_f(y))
\end{equation}
\ExecuteMetaData[agda/Snippets/Introduction.tex]{equiv}
The univalence axiom is that all equivalences are equivalent to the identity
type.
\begin{definition}[The Univalence Axiom]
  \begin{equation}
    \simeq \; \simeq \; \equiv
  \end{equation}
\end{definition}
In practice this means that we can prove equivalence between two types and then
transport code on one to code on the other.
It also means that isomorphisms can be upgraded to fully-fledged equalities.
\begin{definition}[Isomorphism]
  An isomorphism between two types \(A\) and \(B\) is a pair of functions:
  \begin{align}
    f : A \rightarrow B \\
    g : B \rightarrow A
  \end{align}
  Such that they are inverses of each other.
  \begin{align}
    \prod_{(x : A)} , g(f(x)) \equiv x \\
    \prod_{(y : B)} , f(g(y)) \equiv y
  \end{align}
\end{definition}
\ExecuteMetaData[agda/Snippets/Introduction.tex]{iso}
\subsubsection{Higher Inductive Types}
Along with univalence, the other significant benefit we get from path types is
the ability to define new paths on types.
A direct use of this feature is synthetic topology.
We can, for instance, define the circle \(S^1\) as a higher-inductive type.
\ExecuteMetaData[agda/Snippets/Introduction.tex]{circle}

Another important use of higher inductive types is quotient types.
These types are unavailable in ML type theory, but in Cubical Type Theory we can
declare a quotient (here, the type of Kuratowski finite sets) like so:
\ExecuteMetaData[agda/Algebra/Construct/Free/Semilattice.tex]{def}
The first two constructors are standard data constructors, which mimic those for
the list in Haskell.
The next two add the equations we want our type to obey: we don't want it to
care about order or duplicates.
The last constructor \emph{truncates} the type, ensuring that it is a set.
If we hadn't included it paths would not be mere propositions: there would be
more than one path between any given elements of a type.
The ability to have multiple paths is what allows Cubical Type Theory to do the
more advanced things in topology etc., but we don't need its power here.

One HIT in particular will be of great use to us: the HIT for propositional
truncation, denoted with \(\lVert \cdot \rVert\).
We will go through it in detail in section~\ref{prop-trunc}.
\section{Supplementary Definitions}
This section will serve as a reference for any definitions used in the text
which are either not important enough to be examined inline, or common enough to
justify elision.
Many of the definitions come straight from the ``HoTT Book'' \cite{hottbook}.
\subsection{Decidable Types}
It is important to be able to talk about ``decisions'' for things like decidable
equality, and so on.
\begin{definition}[Decision]
  A decision for some type \(P\) is represented by the type \(\mathbf{Dec}(A)\),
  which has two constructors:
  \begin{equation}
    \begin{alignedat}{2}
      \text{yes} &:      &A \rightarrow \mathbf{Dec}(A) \\
      \text{no}  &: \neg &A \rightarrow \mathbf{Dec}(A)
    \end{alignedat}
  \end{equation}
\end{definition}

Internally, we actually use a rather more complicated type, which has preferable
computational behaviour.
The interface is the same, however.
\subsection{Propositional Truncation} \label{prop-trunc}
Essentially, a value of type \(\lVert A \rVert\) proof that an \(A\) exists,
without revealing which \(A\) in particular.
In HoTT, it allows us to work in a setting similar to classical logic.
In Agda its definition is:
\ExecuteMetaData[agda/Snippets/Introduction.tex]{prop-trunc}
Any eliminators from this type must not distinguish between different values of
\(A\).

\begin{lemma}[Propositional Elimination]
  We can eliminate from the type \(\lVert A \rVert\) to \(B\) with a function
  \(A \rightarrow B\) if B is a proposition.
\end{lemma}
\begin{proof}
  By case-analysis on \(\lVert A \rVert\).
  In the first case, we apply our function.
  In the second, we are required to show that that application agrees on any two
  values of \(\lVert A \rVert\) \(\lVert A \rVert\).
  This is true as \(B\) is a proposition.
  In Agda:
  \ExecuteMetaData[agda/HITs/PropositionalTruncation/Proofs.tex]{rec}
\end{proof}

Since \(\lVert A \rVert\) is itself a proposition, we can work ``under'' the
truncation as much as we like.
In fact the type \(\lVert \cdot \rVert\) itself forms a monad, allowing us to
use monadic notation.
\begin{lemma}
  \(\lVert \cdot \rVert\) is a monad, with bind being the eliminator defined
  above, and return being the \(\lvert \cdot \rvert\) constructor.
\end{lemma}
\begin{proof}
  The monad laws are satisfied trivially, as \(\lVert \cdot \rVert\) is a
  proposition.
\end{proof}

We can further eliminate into a type that isn't a proposition, as long as we
prove that the function used in the eliminator doesn't discriminate between
different \(A\)s in \(\lVert A \rVert\).
\begin{lemma}[Propositional Elimination by Coherence] \label{elim-non-prop}
  We can eliminate from a propositionally-truncated type into a non-proposition,
  as long as the function is coherently constant
  \cite{krausGeneralUniversalProperty2015}.
\end{lemma}
\begin{proof}
  \todo[inline]{sketch proof}
\end{proof}

From these eliminators, we can derive some useful lemmas.
\begin{lemma}[Propositional Refutation]
  While a propositionally truncated proof is not enough to prove the
  non-truncated type, it is enough to \emph{refute} the non-truncated negation
  of the type.
  \begin{equation}
    \lVert A \rVert \rightarrow \neg (\neg A)
  \end{equation}
\end{lemma}
\begin{proof}
  We will use propositional elimination.
  First, to go from \(A\) to \((A \rightarrow \bot) \rightarrow \bot\) is
  trivial.
  Secondly, to show that \((A \rightarrow \bot) \rightarrow \bot\) is a
  proposition, we can use the fact that \(\bot\) is a proposition, and then
  derive (via function extensionality) that any \(A \rightarrow \bot\) is also a
  proposition.
\end{proof}

\begin{lemma}[Recomputation]
  If a type \(A\) is decidable, and there exists a proof of the type \(\lVert A
  \rVert\), we can derive a proof of \(A\).
  In other words, the following function exists:
  \begin{equation}
    \text{recompute} : \textbf{Dec}(A) \times \lVert A \rVert \rightarrow A
  \end{equation}
\end{lemma}
\begin{proof}
  By case analysis of the decision over \(A\), we know that the false case must
  be untrue (since we can refute it using propositional refutation), and we can
  therefore extract the true case.
\end{proof}

Occasionally we will call families ``predicates'' when we mean to treat the
family as a property on the domain.
Formally this might imply that the codomain is a proposition (i.e. isomorphic to
either \(\top\) or \(\bot\)), but it is not always necessary to prove this fact,
and indeed occasionally we will want to use some of the information stored in
the predicate, making it not propositional.
For instance, a natural number being ``not prime'' we would refer to as a
predicate, but the proof of this fact may be a list of the number's prime
factors, so we will not truncate it away.

When we say that two predicates \(P\) and \(Q\) are ``equal strength'', we mean
there are a pair of functions \(P \rightarrow Q\) and \(Q \rightarrow P\).
This is distinct from \(P\) and \(Q\) being isomorphic, as we haven't shown
anything regarding the relationship between those two functions.
However, if \(P\) and \(Q\) are themselves propositions, this does indeed imply
an isomorphism.
\begin{lemma} \label{equal-strength}
  For any two types \(A\) and \(B\), if there are functions \(f : A \rightarrow
  B\) and \(g : B \rightarrow A\), \(\lVert A \rVert\) and \(\lVert B \rVert\)
  are equivalent.
\end{lemma}
\begin{proof}
  We prove equivalence by way of isomorphism.

  Let \(A\), \(B\), \(f\), and \(g\) be defined as above.
  For isomorphism, we first provide a pair of functions \(\lVert f \rVert\) and
  \(\lVert g \rVert\), which are defined by lifting \(f\) and \(g\) under
  propositional truncation, respectively.

  Next, we need to prove that \(g\) is a section of \(f\).
  \begin{equation}
    \text{section}(f, g) = f \circ g \equiv id
  \end{equation}
  Since \(\lVert B \rVert\) is a proposition, this is true by function
  extensionality.

  Finally, \(f\) is a retraction of \(g\).
  \begin{equation}
    \text{retraction}(f, g) = g \circ f \equiv id
  \end{equation}
  Again, since \(\lVert A \rVert\) is a proposition, this is true by function
  extensionality.

  In Agda:
  \ExecuteMetaData[agda/HLevels/Proposition.tex]{prop-iff}
\end{proof}
\subsection{Functions}
\subsubsection{Surjections}
The following is the definition of surjectivity in HoTT:
\begin{equation}
  \text{Surjective}(f) = \prod_{(y : B)} , \lVert \text{fiber}_f(y) \rVert
\end{equation}
\ExecuteMetaData[agda/Function/Surjective.tex]{surj-def}
We are also interested in \emph{split} surjectivity, where we don't truncate the
fiber:

\begin{equation}
  \text{SplitSurjective}(f) = \prod_{(y : B)} , \text{fiber}_f(y)
\end{equation}
\ExecuteMetaData[agda/Function/Surjective.tex]{split-surj-def}

A surject\emph{ion}, then, is a function which is surject\emph{ive}:
\begin{equation}
  A \twoheadrightarrow B = \sum_{(f : A \rightarrow B)} \text{SplitSurjective}(f)
\end{equation}
\ExecuteMetaData[agda/Function/Surjective.tex]{surjection}
\subsubsection{Injections}
The standard definition of an injective function:
\ExecuteMetaData[agda/Function/Injective.tex]{injective-def}
An injection:
\ExecuteMetaData[agda/Function/Injective.tex]{injection-def}
\subsection{Fin}
\(\textbf{Fin}\) is a finite prefix of the natural numbers.
Semi-formally:
\begin{equation}
  \textbf{Fin}(n) = \left[ 0 \dots n \right)
\end{equation}
There are actually a number of ways to define \(\textbf{Fin}\) in type theory.
The method we use here is as follows:
\ExecuteMetaData[agda/Data/Fin.tex]{def}
\subsection{Split Enumerability}
Here is how we define it in Agda:
\ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{definition}
\subsection{Predicates Over Lists}
These two well-known dual types for quantifying a predicate over a list are used
extensively in this work.
We will define them using the indexing operator \(!\), which has the following
type:
\begin{equation}
  ! : \prod_{(\mathit{xs} : \textbf{List}(A))} , \textbf{Fin}(\text{length}(\mathit{xs})) \rightarrow A
\end{equation}
For some predicate \(P\), the two definitions are as follows:
\begin{definition}[Any]
  \begin{equation}
    \Diamond P \mathit{xs} = \sum_{(i : \textbf{Fin}(\text{length}(\mathit{xs})))} , P(xs \ind i)
  \end{equation}
  \ExecuteMetaData[agda/Data/List/Relation/Unary.tex]{any-def}
\end{definition}
\begin{definition}[All]
  \begin{equation}
    \Box P \mathit{xs} = \prod_{(i : \textbf{Fin}(\text{length}(\mathit{xs})))} , P(xs \ind i)
  \end{equation}
  \ExecuteMetaData[agda/Data/List/Relation/Unary.tex]{all-def}
\end{definition}
We are using the tradition modal logic operators here: \(\Diamond\)
means ``any'', and \(\Box\) means ``all''.
\subsection{Membership}
Several other constructions are defined in terms of \(\AgdaDatatype{◇}\),
starting with \(\AgdaDatatype{∈}\):
\begin{equation}
  x \in \mathit{xs} = \Diamond (\equiv x) \mathit{xs}
\end{equation}
\ExecuteMetaData[agda/Data/List/Membership.tex]{in-def}

To show \emph{unique} membership, the membership proof itself must be a
contraction.
\begin{equation}
  x \inunique \mathit{xs} = \text{isContr}(x \in \mathit{xs})
\end{equation}
\ExecuteMetaData[agda/Data/List/Membership.tex]{in-uniq}
\section{Supplementary Code}
\subsection{Permutations}
Swap:
\ExecuteMetaData[agda/Data/List/Relation/Binary/Permutation.tex]{swap}
Cons:
\ExecuteMetaData[agda/Data/List/Relation/Binary/Permutation.tex]{cons}
Tail:
\ExecuteMetaData[agda/Data/List/Relation/Binary/Permutation.tex]{tail}
\subsection{Non-Dependent Sum (Disjoint Union)}
\ExecuteMetaData[agda/Snippets/Introduction.tex]{sum-def}
\subsection{Split Enumerability}
Closure over \(\bot\):
\ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{empty}
Closure over \(\top\):
\ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{unit}
Closure over non-dependent sum:
\ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{sum-prf}
\subsection{Lists}
\subsubsection{Indexing}
\ExecuteMetaData[agda/Data/List.tex]{indexing}
\subsubsection{Removing Duplicates}
\ExecuteMetaData[agda/Data/List/Membership.tex]{uniques}
\subsection{Functions}
If there is an injection from \(A\) into a set with decidable equality, \(A\)
has decidable equality.
\ExecuteMetaData[agda/Function/Injective.tex]{discrete}
If there is a surjection from \(A\) to \(B\), there is an injection from \(B\)
to \(A\).
\ExecuteMetaData[agda/Function/Surjective.tex]{surj-to-inj}
\subsection{Membership Proofs}
All iso:
\ExecuteMetaData[agda/Data/List/Relation/Unary.tex]{all-iso}
Any iso:
\ExecuteMetaData[agda/Data/List/Relation/Unary.tex]{any-iso}
A small function to make membership proofs shorter:
\ExecuteMetaData[agda/Data/List/Membership.tex]{at}
Pushing and pulling: going between \(x \; \AgdaDatatype{∈} \; xs\) and \(x
\AgdaDatatype{∈} \; y \; \AgdaInductiveConstructor{∷} \; xs\).
\ExecuteMetaData[agda/Data/List/Relation/Unary.tex]{push-pull}
Push-pull for unique proofs of membership.
\ExecuteMetaData[agda/Data/List/Relation/Unary.tex]{push-pull-unique}
Proving that an element at the head of a list is in that list.
\ExecuteMetaData[agda/Data/List/Relation/Unary.tex]{here-unique}
\(\AgdaFunction{map}\) distributes over \(\AgdaDatatype{∈}\)
\ExecuteMetaData[agda/Data/List/Relation/Unary.tex]{cong-any}
\ExecuteMetaData[agda/Data/List/Membership.tex]{cong-in}
An element is not in a list it's been removed from.
\ExecuteMetaData[agda/Data/List/Membership.tex]{not-in-removed}
\ExecuteMetaData[agda/Data/List/Membership.tex]{cons-remove}
If \(x\) isn't in a list, it is not in that list with another element removed.
\ExecuteMetaData[agda/Data/List/Membership.tex]{remove-other}
Removing another element from a list doesn't affect \(x\)'s proof of unique
membership.
\ExecuteMetaData[agda/Data/List/Membership.tex]{unique-remove-other}
If \(x\) is in \(xs\), \(x\) is uniquely in the unique elements of \(xs\).
\ExecuteMetaData[agda/Data/List/Membership.tex]{unique-preserves}
% \subsection{Convolution}
% Convolution of lists.
% \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{convolve}
% Convolution of streams.
% \ExecuteMetaData[agda/Cardinality/Infinite.tex]{convolve}
% Convolution of colists.
% \ExecuteMetaData[agda/Cardinality/PossiblyInfinite.tex]{convolve}
% \subsection{Decidable Predicates}
% Rather than the ``usual'' formulation of \(\AgdaDatatype{Dec}\), we use the
% following definition:
% \ExecuteMetaData[agda/Relation/Nullary.tex]{dec-def}
% \subsection{Pauli Group Functions}
% \ExecuteMetaData[agda/Data/Pauli.tex]{bad-eq}
\subsection{Fin} \label{agda-fin}
Injectivity of the \(\textbf{Fin}\) type constructor.
\ExecuteMetaData[agda/Data/Fin/Properties.tex]{fin-inj}
\section{References}
\printbibliography[heading=none]
\end{subappendices}
\end{refsection}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../paper"
%%% End:

### 2020-01-04 paper.tex
initial

\input{preamble}
\title{Semirings and Semimodules: A Substrate for Computation}
\author{Donnacha Oisín Kidney}
\begin{document}
\maketitle
\listoftodos
\tableofcontents
\include{sections/maths}
\include{sections/type-theory}
\include{sections/implementation}
\include{sections/free-semimodule}
\include{sections/univalence-for-privacy}
\include{sections/combinatorics}
\include{sections/ordered-monoids-and-search}
\include{sections/semirings-on-the-gpu}
\end{document}

### 2020-01-04 paper.tex
bib  etc

@@
 \input{preamble}
-\title{Semirings and Semimodules: A Substrate for Computation}
+\title{Finiteness, Cardinality, and Combinatorics in Homotopy Type Theory}
 \author{Donnacha Oisín Kidney}
@@
 \tableofcontents
-\include{sections/maths}
-\include{sections/type-theory}
-\include{sections/implementation}
-\include{sections/free-semimodule}
-\include{sections/univalence-for-privacy}
-\include{sections/combinatorics}
-\include{sections/ordered-monoids-and-search}
-\include{sections/semirings-on-the-gpu}
+\newcommand{\pifun}[2]{\prod_{\left( \mathit{ #1 } : #2 \right)}}
+\begin{quote}
+  \textbf{Abstract}
+  We explore five notions of finiteness in Homotopy Type Theory \cite{hottbook}.
+  We prove closure properties about all of these notions, culminating in a proof
+  that decidable Kuratowski-finite sets form a topos.
+
+  We extend the definitions to include infinite types, developing a similar
+  classification of countable types.
+
+  We use the definition of finiteness to formalise \emph{species}, in much
+  the same way as in \cite{yorgeyCombinatorialSpeciesLabelled2014}.
+  A clear duality with containers
+  \cite{abbottContainersConstructingStrictly2005} falls out naturally from our
+  definition.
+
+  We formalise our work in Cubical Agda
+  \cite{vezzosiCubicalAgdaDependently2019}, and we implement a library for proof
+  search (including combinators for level-polymorphic fully generic currying),
+  and demonstrate how it can be used to both prove properties and synthesise
+  full functions given desired properties.
+\end{quote}
+\section{Introduction}
+Proofs in a constructive setting are more substantial things than their
+classical counterparts.
+Rather than just evidence for some fact, they are objects in their own right,
+which we can manipulate and extract information from.
+
+Sometimes, the extra detail can reveal interesting subtleties: fissures which
+divide homogeneous-seeming definitions into a variety of related concepts.
+Finiteness is a prime example of this phenomenon: in classical set theory, a
+finite set is a finite set, and there's not much else that we can say.
+In constructive type theory, we start the bidding at no fewer than \emph{four}
+different predicates representing finiteness
+\cite{spiwackConstructivelyFinite2010}, upped to eight in
+\cite{firsovVariationsNoetherianness2016a}: all of which differ in significant
+and interesting ways.
+
+Other times, the extra baggage is unwanted.
+Maybe we want to enforce that a particular piece of information should be
+hidden, or treated as irrelevant: ``yes, I have proven that there is a
+particular number with this property, but I don't want to reveal \emph{which}
+number.''
+
+This is where Homotopy Type Theory \cite{hottbook} can step in.
+Through primarily a generalisation of equality, we gain access to a number of
+exotic types which can, among other things, allow us to perform this ``hiding''.
+In the context of finiteness, the act of hiding is worth studying in its own
+right, and further delineates several new finiteness predicates which collapse
+even in standard Martin-Löf type theory
+\cite{martin-lofIntuitionisticTypeTheory1980}.
+
+The other headline feature of HoTT's generalised equality is univalence: put
+simply, it allows us to treat isomorphic types as equal, giving us in the
+constructive world access to a technique that is commonplace classically.
+It turns out that finite types have a lot to say about isomorphisms, and we will
+rely heavily on this aspect of HoTT to work with them.
+\subsection{Contributions}
+In this work, we will explore finite types in Cubical Type Theory
+\cite{cohenCubicalTypeTheory2016}, and expose their relationship to infinite
+types, species, and demonstrate their practical uses for proofs in dependently
+typed programming languages.
+
+In section~\ref{strong-finiteness}, we will explore the ``strong'' notions of
+finiteness (i.e. those at least as strong as Kuratowski finiteness
+\cite{kuratowskiNotionEnsembleFini1920}), with a special focus on cardinal
+finiteness.
+We will delineate some subtly different finiteness predicates, and prove the
+relationship between each.
+We will prove, then, that every discrete Kuratowski finite type is cardinal
+finite, and subsequently will show that discrete Kuratowski finite sets form a
+topos.
+
+In section~\ref{species-and-containers}, we will redefine our finiteness
+predicates using \emph{containers}, and prove that the new definitions are
+equivalent to the old.
+Using this, we will prove much stronger relationships between our finiteness
+predicates and relation-based definitions of finiteness.
+Finally, We will show how the species-container duality falls out naturally from
+these new definitions.
+
+In section~\ref{infinite-cardinalities}, we will extend our study of finite
+types to infinite but countable types.
+We will see that the finiteness predicates are mirrored with countable
+counterparts, and we will prove closure under the Kleene star and plus.
+
+Proofs of finiteness have well-known practical applications in
+constructive mathematics \cite{firsovDependentlyTypedProgramming2015}.
+In section~\ref{practical}, we build a library which exploits these uses in
+Cubical Agda \cite{vezzosiCubicalAgdaDependently2019}, allowing automation of
+complex proofs over finite types.
+We frame this in terms of the principle of omniscience for finite types.
+Thanks to the flexibility afforded to us by Cubical Type Theory, we are able to
+go further than the usual examples of this kind of proof automation: as well as
+proving properties about functions, we can synthesise functions whole-cloth from
+their desired properties.
+Through the unified interface for finite and countable types, we can
+reuse the automation machinery for \emph{partial} proof search over infinite
+search spaces.
+Along the way, we extend the work in \cite{allaisGenericLevelPolymorphic2019} to
+prove isomorphisms between the curried and uncurried forms of \(n\)-ary
+dependent functions.
+\section{Strong Finiteness} \label{strong-finiteness}
+\begin{figure}[h]
+  \centering
+  \begin{tikzcd}[cramped, row sep=small]
+    &
+    \ar[ddddd, dash, start anchor={[yshift=4ex]}, end anchor={[yshift=-4ex]}] &
+    \text{Ordered} &
+    \ar[ddddd, dash, dashed, start anchor={[yshift=4ex]}, end anchor={[yshift=-4ex]}] &
+    \text{Unordered}
+    \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \ar[rrrr, dash, start anchor={[xshift=-6ex]}, end anchor={[xshift=10ex]}] &
+    &
+    &
+    &
+    {}
+    \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    &
+    &
+    \text{Split Enumerable}
+      \; \ref{split-enumerable}
+      \ar[d, xshift=-1ex, bend right=30]
+      &
+    &
+    \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \text{Discrete} &
+    &
+    \text{Manifest Bishop}
+      \ar[u, xshift=1ex, bend right=30]
+      \ar[rr, end anchor=south west, start anchor=south east, bend right=10, crossing over]
+      \; \ref{manifest-bishop}
+      &
+    &
+    \text{Cardinal}
+      \; \ref{cardinal}
+      \ar[ll, start anchor=north west, end anchor=north east, bend right=10, "\text{linOrd}" description]
+    \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \ar[rrrr, dash, dashed, start anchor={[xshift=-6ex]}, end anchor={[xshift=10ex]}] &
+    &
+    &
+    &
+    {}
+    \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \text{Non Discrete} &
+    {} &
+    \text{Manifest Enumerable}
+      \; \ref{manifest-enumerability}
+      \ar[from=uu, xshift=-1ex, bend right=30, crossing over]
+      \ar[uu, bend right=30,  "\text{Discrete}" description]
+      \ar[rr, start anchor=south east, end anchor=south west, bend right=10, crossing over]
+      &
+    {}
+    &
+    \text{Kuratowski}
+      \; \ref{kuratowski}
+      \ar[ll, start anchor=north west, end anchor=north east, bend right=10,  "\text{linOrd}" description]
+      \ar[uu, xshift=1ex, bend right=30,  "\text{Discrete}" description]
+      \ar[from=uu, xshift=-1ex, bend right=30, crossing over]
+  \end{tikzcd}
+  \caption{Classification of the strong definitions of finiteness, according to
+    whether they are discrete (imply decidable equality) and whether they induce
+    a linear order.}
+  \label{finite-classification}
+\end{figure}
+
+In this section, we will explore five predicates for finiteness, classified in
+figure~\ref{finite-classification}.
+Of these, manifest enumerability is new, to our knowledge.
+
+Figure~\ref{finite-classification} organises the predicates according to
+their ``strength''; i.e. how much information they provide about a conforming
+type.
+For instance, a proof that some type \(A\) is manifestly Bishop finite (the
+strongest of the notions, explored in section~\ref{manifest-bishop}) also tells
+us that \(A\) is discrete (has decidable equality), and gives us a linear order
+on the type.
+A type that is Kuratowski finite (section~\ref{kuratowski}) has no such extra
+features: indeed, we will see examples of Kuratowski finite types which are not
+even sets, never mind discrete ones.
+
+In the following sections, we will go through each of the predicates, proving
+how to weaken each (i.e. we will provide a proof that every cardinally finite
+type is Kuratowski finite), and how to strengthen them, given the required
+property.
+In terms of figure~\ref{finite-classification}, this amounts to providing proofs
+for each arrow.
+
+For each predicate, we will also prove its closure over sums and products in
+both dependent and non-dependent forms, if such a closure exists.
+This will culminate in our main result for this section: the formal proof that
+decidable Kuratowski finite sets form a topos. \todo[noline]{reference for this}
+More specifically, we show that cardinal finite sets form a topos, that
+decidable Kuratowski finite sets are equivalent in strength to cardinal finite
+sets, and carry the proof over.
+This result relies on proofs on each of the other finiteness predicates.
+\subsection{Split Enumerability} \label{split-enumerable}
+We will start with the simplest definition of finiteness: we say a set is
+enumerable if there is a list of its elements which contains every element in
+the set.
+More formally:
+\begin{definition}[Split Enumerable Set]
+  \begin{equation}
+    \mathcal{E}!(A) = \sum_{(\mathit{xs} : \textbf{List} A)} , \prod_{(x : A)} , x \in xs
+  \end{equation}
+  We call the first component of this pair the ``support'' list, and the second
+  component the ``cover'' proof.
+\end{definition}
+
+The term \(x \in \mathit{xs}\) hides a subtle detail of this definition.
+Depending on how we choose to represent \(x \in \mathit{xs}\), the resulting
+predicate will be equivalent in strength to either manifest Bishop finiteness
+(section~\ref{manifest-bishop}) or manifest enumerability
+(section~\ref{manifest-enumerability}).
+For now, we will define it like so:
+
+\begin{definition}[List Membership] \label{list-membership}
+  \begin{equation}
+    x \in \mathit{xs} = \sum_{(i : \mathbf{Fin}(\text{length}(\mathit{xs})))} , \mathit{xs} \ind i \equiv x
+  \end{equation}
+  Here, \(\mathbf{Fin}(n)\) is the type of natural numbers smaller than \(n\),
+  and \(\ind\) is an indexing function.
+  In other words, to say that some item \(x\) is in some list \(\mathit{xs}\),
+  we say that there is an index into \(\mathit{xs}\) which selects an element of
+  \(\mathit{xs}\) which is equivalent to \(x\).
+\end{definition}
+
+Clearly, this definition of list membership tells us not just that an item is in
+a list, but also its \emph{position} in that list.
+That piece of information gives use decidable equality, restricting the
+predicate to discrete sets.
+In section~\ref{manifest-enumerability}, we will show how to hide the position,
+giving us a variant of the predicate which allows for non-set finite types.
+\subsubsection{Decidable Equality}
+We will now prove that split enumerability implies decidable equality.
+We first need to prove a lemma regarding surjections to split enumerable types.
+\begin{lemma} \label{split-enum-surj}
+  Any split enumerable type has a split surjection from a finite prefix of the
+  natural numbers.
+\end{lemma}
+\begin{proof} 
+  Let \(E\) be a proof of type \(\mathcal{E}!(A)\) for some type \(A\), and
+  \(n\) be the length of the support list of \(E\).
+  Then let \(f : \textbf{Fin}(n) \rightarrow A\) be defined as the indexing
+  function into the support list of \(E\).
+
+  We then prove this function to be a split surjection by providing a fiber over
+  all of the elements of its domain: this fiber is the second component of \(E\).
+
+  In Agda:
+  \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{to-surj}
+\end{proof}
+
+Because of careful arrangement of previous definitions, this proof is only three
+lines of Agda.
+It is not difficult to derive decidable equality from this fact.
+\begin{lemma} \label{split-enum-discrete}
+  Any split enumerable type has decidable equality (is discrete).
+\end{lemma}
+\begin{proof}
+  We use a corollary, proven in the appendix, that if there is a
+  split-surjection from \(A\) to \(B\), and \(A\) is discrete, then \(B\) is
+  also discrete.
+  \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{discrete}
+\end{proof}
+
+This also tells us that any split enumerable type must be a set.
+\begin{lemma}
+  Any split enumerable type is a set.
+\end{lemma}
+\begin{proof}
+  By Hedberg's theorem \cite{hedbergCoherenceTheoremMartinLof1998}, since split
+  enumerable types have decidable equality
+  (proposition~\ref{split-enum-discrete}), they are sets.
+\end{proof}
+
+Though it's not strictly needed for the rest of the construction, it is
+instructive to also prove inverse regarding surjections; i.e. that we can
+derive \(\mathcal{E}!\) from a surjection from another finite type.
+\begin{lemma} \label{split-enum-from-surj}
+  Any type with a split surjection from another split enumerable type is itself
+  split enumerable.
+\end{lemma}
+\begin{proof}
+  Let \(E\) be a proof of split enumerability for some set \(A\), and let \(f\)
+  be a surjection from \(A\) to \(B\).
+  The support list for the proof of split enumerability over \(B\) can be
+  obtained by mapping the surjective function over the support list of \(E\).
+
+  To obtain the cover proof, let \(f^{-1}\) be the inverse function of \(f\)
+  (which we can obtain because it is a split-surjection).
+  Using the cover proof of \(E\), we can show that, for all \(y : B\), \(f^{-1}(y)\)
+  is in the support list for \(E\).
+  Using the following lemma (proven in the appendix):
+  \begin{equation}
+    x \in xs \rightarrow f (x) \in \text{map} \; f \; \mathit{xs}
+  \end{equation}
+  We obtain a proof that:
+  \begin{equation}
+    \forall (y : B). f (f^{-1}(y)) \in \text{map} \; f \; \mathit{xs}
+  \end{equation}
+  And, since \(f (f^{-1}(y)) \equiv y\) (by the definition of a surjection), we
+  have proven our goal.
+
+  In Agda:
+  \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{from-surj}
+\end{proof}
+\subsubsection{Closure}
+In this section we will prove closure under various operations for split
+enumerable sets.
+We are working towards a topos proof, which requires us to prove closure under
+a variety of operations: for now, we only have enough machinery to demonstrate
+the semiring operations, and dependent sums.
+in order to show closure under exponentials (function arrows), we will need an
+equivalence with \(\mathbf{Fin}\), which will be provided in
+section~\ref{manifest-bishop}.
+\begin{lemma}
+  \(\bot\) is split enumerable.
+\end{lemma}
+\begin{proof}
+  The support list for \(\bot\) is the empty list.
+  The cover proof is present via the principle of explosion.
+\end{proof}
+\begin{lemma}
+  \(\top\) is split enumerable.
+\end{lemma}
+\begin{proof}
+  The support list for \(\top\) is the singleton list containing \(\text{tt}\).
+  The cover proof is a constant function which returns an index into the first
+  item in the list.
+\end{proof}
+Both of those cases were relatively simple.
+Next, we will look into how to combine proofs of split enumerability.
+\begin{theorem}
+  Split-enumerability is closed under \(\sum\).
+\end{theorem}
+\begin{proof}
+  Let \(E_A\) be a proof of split enumerability for some type \(A\), and \(E_U\)
+  be a function of the type:
+  \begin{equation}
+    E_U : \prod_{(x : A)} , \mathcal{E}!(U(x))
+  \end{equation}
+  In other words, a function which returns a proof of split enumerability for
+  each member of the family \(U\).
+
+  To obtain the support list, we concatenate the support lists of all the proofs
+  of split-finiteness for \(U\) over the support list of \(E_A\).
+  In Agda:
+  \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{sigma-enum}
+  ``do-notation'' is available to us as we're working in the list monad.
+
+  The cover proof follows in a simple, if tedious, way.
+\end{proof}
+\begin{lemma}
+  Split-enumerability is closed under Cartesian product (non-dependent product).
+\end{lemma} 
+\begin{proof}
+  Since non-dependent products are simply a special case of the \(\sum\) type,
+  we can reuse the proof of closure over \(\sum\) here.
+\end{proof}
+\begin{lemma}
+  Split-enumerability is closed under disjoint union (non-dependent sum).
+\end{lemma}
+\begin{proof}
+  Again, this can be derived from closure of the \(\sum\) type.
+  Disjoint union between two types \(A\) and \(B\) can be represented by the
+  following type:
+  \begin{equation}
+    A + B = \sum_{(x : \mathbf{Bool})} , \text{if} \; x \; \text{then} \; A \; \text{else} \; B
+  \end{equation}
+  Then, since all of \(\mathbf{Bool}\), \(A\), and \(B\) are split enumerable,
+  the type \(A + B\) is split enumerable.
+\end{proof}
+The final case can be derived from the previous closures; we give it here anyway
+for clarity.
+\begin{lemma} \label{fin-split-enum}
+  Any finite prefix of the natural numbers (what we have been calling
+  \(\textbf{Fin}\)) is split enumerable.
+\end{lemma}
+\begin{proof}
+  For the support list, we will make use of the function tabulate, which returns
+  a list consisting of the application of \(f\) to every \(\textbf{Fin}(n)\),
+  given \(n\).
+  \begin{equation}
+    \text{tabulate}(n , f) = \left[ f(0) , f(1) \dots f(n - 1) \right]
+  \end{equation}
+  In Agda:
+  \ExecuteMetaData[agda/Data/List.tex]{tabulate}
+  Tabulation with the identity function gives us our cover list.
+
+  The cover proof is obviously true given the definition of tabulate above; in
+  Agda it requires a little more work to show.
+  \ExecuteMetaData[agda/Data/List/Membership.tex]{fin-in-tabulate}
+\end{proof}
+\subsection{Manifest Bishop Finiteness} \label{manifest-bishop}
+\begin{definition}[Manifest Bishop Finiteness]
+  \begin{equation}
+    \mathcal{B}(A) = \sum_{(\mathit{xs} : \textbf{List}(A))} , \prod_{(x : A)} , x \inunique xs
+  \end{equation}
+\end{definition}
+The only difference between this predicate and split enumerability is the list
+membership term: we use \(\inunique\) here, where \(x \inunique \mathit{xs}\) is
+to be read as ``\(x\) occurs exactly once in \(\mathit{xs}\)''.
+It is defined in terms of our previous list membership term
+(definition~\ref{list-membership}).
+\begin{definition}[Unique List Membership]
+  We say an item \(x\) is ``uniquely in'' some list \(\mathit{xs}\) if its
+  membership in that list is a \emph{contraction}; i.e. its membership proof
+  exists, and all such proofs are equal.
+  \begin{equation}
+    x \inunique \mathit{xs} = \text{isContr}(x \in \mathit{xs})
+  \end{equation}
+\end{definition}
+
+A nice consequence of prohibiting duplicates is that now the length of the
+support list is the same as the cardinality of the set.
+\subsubsection{Equivalence}
+While manifest Bishop finiteness may seem stricter than split enumerability, we
+will see soon that it is actually equivalent in strength.
+Why study it separately to split enumerability at all, then?
+What purpose does it serve on our road to Kuratowski finiteness?
+Those questions are answered by the following theorem:
+\begin{lemma} \label{bishop-equiv}
+  Any manifestly Bishop finite type has an equivalence with a finite prefix of
+  the natural numbers.
+\end{lemma}
+\begin{proof}
+  There are many equivalent definitions of equivalence in HoTT.
+  Here we take the version preferred in the Cubical Agda library: contractible
+  maps \cite[Definition~4.4.1]{hottbook}. \todo[noline]{Need this in appendix also}
+  Because of the parallels between contractible maps and split surjections,
+  the proof proceeds much the same as \ref{split-enum-surj}.
+
+  First, we define the map as the indexing function into the support list.
+  Proving this map is a contraction is precisely the cover proof.
+
+  \ExecuteMetaData[agda/Cardinality/Finite/ManifestBishop.tex]{equiv}
+\end{proof}
+
+As we can see, our list-based definition of manifest Bishop finiteness is really
+an equivalence with \(\mathbf{Fin}\) in disguise.
+This will be the key component in establishing closure under exponentials.
+
+We would like to draw attention to the ``power-to-weight ratio'' of this proof.
+We have mentioned how important the equivalence will be for our later proofs: in
+general, the ability to transport via univalence opens up many opportunities
+
+In Cubical Type Theory, equivalences are much more useful than their counterpart
+(isomorphisms) in Martin-Löf  Type Theory: arbitrary proofs and functions can be
+\emph{transported} along equivalences, giving a constructive justification for
+an intuitive and useful technique which is commonplace classically.
+The proof itself, however, is three lines long: thanks to careful definitions,
+the fact that manifest Bishop finiteness is the ``same'' as equivalences to
+\(\mathbf{Fin}\) becomes obviously true.
+Later we will see, by defining lists as containers, a proof that manifest Bishop
+finiteness is equivalent (in the formal sense) to equivalences with
+\(\mathbf{Fin}\). \todo[noline]{This needs to be added to the containers
+  section.}
+\todo[noline]{More on this? }
+\subsubsection{Relationship to Split Enumerability}
+We now show that manifest Bishop finiteness has equal strength to split
+enumerability.
+\begin{lemma}
+  Any manifest Bishop finite set is split enumerable.
+\end{lemma}
+\begin{proof}
+  The support set carries over simply, and the cover proof can be taken from the
+  first component of the cover proof from the proof of manifest Bishop
+  finiteness.
+  \ExecuteMetaData[agda/Cardinality/Finite/ManifestBishop.tex]{bishop-to-split-enum}
+\end{proof}
+\begin{theorem} \label{split-enum-to-manifest-bishop}
+  Any split enumerable set is manifest Bishop finite.
+\end{theorem}
+\begin{proof}
+  Let \(E\) be a proof of split enumerability for some set \(A\).
+  From proposition~\ref{split-enum-discrete} we can derive decidable equality on
+  \(A\), and using this we can define a function \(\text{uniques}\) which filters
+  out duplicates from lists of \(A\)s.
+  \begin{equation}
+    \text{uniques} : \textbf{List}(A) \rightarrow \textbf{List}(A)
+  \end{equation}
+  This gives us our support list.
+
+  It suffices now to prove the following:
+  \begin{equation}
+    \prod_{(x : A)} , \prod_{(\mathit{xs} : \textbf{List}(A))} , x \in \mathit{xs} \rightarrow x \inunique \text{uniques}(\mathit{xs})
+  \end{equation}
+  And from that we can generate our cover proof.
+
+  Full details of the formal proof are provided in the appendix.
+\end{proof}
+Note that while manifest Bishop finiteness as split enumerability are equivalent
+in ``strength'', the two types are not equivalent.
+In particular, there are infinitely many inhabitants of \(\mathcal{E}!\), while
+for a type \(A\) with \(n\) inhabitants, there are only \(n!\) inhabitants of
+\(\mathcal{B}(A)\).
+\subsubsection{Closure}
+Proving equal strength of split enumerability and manifest Bishop finiteness
+allows us to carry all of the previous proofs of closure over to manifest Bishop
+finite sets (and vice-versa).
+Missing from our previous proofs was a proof of closure of functions.
+We remedy that here.
+\begin{theorem}
+  Manifest bishop finiteness is closed over dependent functions (\(\prod\)-types).
+
+  Formally, given a type \(A\) and a type family \(B\) on \(A\), when \(A\) is
+  manifestly Bishop finite:
+  \begin{equation}
+    \mathcal{B}\left( A \right)
+  \end{equation}
+  And \(B\) is manifestly Bishop finite over all points of \(A\):
+  \begin{equation}
+    \prod_{(x : A)} , \mathcal{B}\left( B(x) \right)
+  \end{equation}
+  Then we have the following:
+  \begin{equation}
+    \mathcal{B}\left(\prod_{(x : A)} , B(x)\right)
+  \end{equation}
+\end{theorem}
+\begin{proof}
+  This proof is essentially the composition of two transport operations, made
+  available to us via univalence.
+
+  First, we will simplify things slightly by working only with split
+  enumerability.
+  As this is equal in strength to manifest Bishop finiteness, any closure proofs
+  carry over.
+
+  Secondly, we will replace \(A\) in all places with \(\mathbf{Fin}(n)\).
+  Since we have already seen an equivalence between these two types, we are
+  permitted to transport along these lines.
+  This is the first transport operation.
+
+  The bulk of the proof now is concerned with proving the following:
+  \begin{equation}
+    \left( \prod_{(x : \mathbf{Fin}(n))} , \mathcal{E}!(A(x)) \right) \rightarrow \mathcal{E}!\left( \prod_{(x : \mathbf{Fin}(n))} , A(x) \right)
+  \end{equation}
+  Our strategy to accomplish this will be to consider functions from
+  \(\mathbf{Fin}(n)\) as \(n\)-tuples over some type family \(T\).
+  \begin{equation}
+    \begin{aligned}
+      \mathbf{Tuple}(T, 0)   &= \top \\
+      \mathbf{Tuple}(T, n+1) &= A(0) \times \mathbf{Tuple}(T \circ \text{suc}, n)
+    \end{aligned}
+  \end{equation}
+  This type is manifestly Bishop finite, as it is constructed only from products
+  and the unit type.
+
+  We then prove an isomorphism between this representation and \(\Pi\)-types.
+  \begin{equation}
+    \mathbf{Tuple}(T, n) \iff \prod_{(x : \mathbf{Fin}(n))} , T(x)
+  \end{equation}
+  This allows us to transport our proof of finiteness on tuples to one on
+  functions from \(\mathbf{Fin}\) (our second transport operation), proving our
+  goal.
+
+  \ExecuteMetaData[agda/Cardinality/Finite/ManifestBishop.tex]{pi-clos}
+\end{proof}
+\subsection{Manifest Enumerability} \label{manifest-enumerability}
+As we said in section~\ref{split-enumerable}, we will now retrieve a distinction
+between enumerability and manifest Bishop finiteness.
+We do so by propositionally truncating the membership proof:
+\begin{definition}[Manifest Enumerability]
+  \begin{equation}
+    \mathcal{E}(A) = \sum_{(\mathit{xs} : \mathbf{List}(A))} , \prod_{(x : A)} , \lVert x \in \mathit{xs} \rVert
+  \end{equation}
+  \ExecuteMetaData[agda/Cardinality/Finite/ManifestEnumerable.tex]{def}
+\end{definition}
+
+There is another definition, which lies between manifest enumerability and
+bishop finiteness, where we would truncate just the equality, but reveal the
+position in the membership proof.
+\begin{equation}
+  \mathcal{E}'(A) = \sum_{(\mathit{xs} : \mathbf{List}(A))} , \prod_{(x : A)} , \sum_{(i : \textbf{Fin}(\text{length}(\mathit{xs})))} , \lVert \mathit{xs} \ind i \equiv x \rVert
+\end{equation}
+However, this simply amounts to a proof of manifest Bishop finiteness over
+set-truncated types, rather than a significantly different construction, so it
+does not interest us here.
+\subsubsection{Higher Homotopy Levels}
+By hiding the position, we have essentially removed the ``decidable'' component
+from split enumerability.
+Our predicate now becomes general enough to work with non-sets: we will show
+here that the circle is manifestly enumerable.
+\begin{theorem}
+  The circle \(S^1\) is manifestly enumerable.
+\end{theorem}
+\begin{proof}
+  As the cover proof is a truncated proposition, we need only consider the point
+  constructors, making this poof the same as the proof of split enumerability on
+  \(\top\).
+
+  The support list will be a singleton list containing the single point in the
+  type, and the cover proof will be an index pointing at the first element in
+  the support list.
+
+  \ExecuteMetaData[agda/Cardinality/Finite/ManifestEnumerable.tex]{circ-inst}
+\end{proof}
+\subsubsection{Surjections}
+This predicates relation to surjectivity is much the same as split
+enumerability's relation to \emph{split} surjectivity.
+Deriving a surjection from \(\mathbf{Fin}\) is straightforward:
+\begin{lemma}
+  If a type \(A\) is manifestly enumerable, then there is a surjection from
+  \(\mathbf{Fin}\) into \(A\).
+\end{lemma}
+\begin{proof}
+  \todo[inline]{Proof}
+\end{proof}
+And building from a surjection is similarly not too difficult:
+\begin{lemma}
+  If there is a surjection from a manifestly enumerable type \(A\) to another
+  type \(B\), then \(B\) is manifestly enumerable.
+\end{lemma}
+\begin{proof}
+  The proof proceeds in much the same way as in
+  proposition~\ref{split-enum-from-surj}.
+  The only difference is that the second component contains a truncation, but
+  since we can always eliminate from a truncation into a truncation this is not
+  a problem.
+
+  In Agda:
+  \ExecuteMetaData[agda/Cardinality/Finite/ManifestEnumerable.tex]{from-surj}
+\end{proof}
+\subsubsection{Relation to Split Enumerability}
+As split enumerability is stronger than manifest enumerability, we can convert
+from one to the other easily.
+\begin{lemma}
+  Any split enumerable type is also manifestly enumerable.
+\end{lemma}
+\begin{proof}
+  The proof carries over via truncation of the cover proof.
+\end{proof}
+The main missing piece from manifest enumerability is decidable equality: by
+truncating the membership proof, we have removed the ability to distinguish
+(decidably) between members of the type.
+Indeed, resupplying decidable equality is precisely sufficient to recover split
+enumerability.
+
+\begin{theorem} \label{manifest-enum-to-split}
+  A manifestly enumerable type with decidable equality is split enumerable.
+\end{theorem}
+\begin{proof}
+  The support list stays the same between both enumerability proofs.
+
+  For the cover proof, we need a way to get the contents out of a truncated
+  value.
+  We can do exactly that with the following lemma, called recompute:
+  Given a decision procedure for some type \(A\), and a truncation for the
+  proposition \(A\), we can discount the possibility of \(A\) being false, and
+  therefore extract the true decision.
+
+  For the cover proof, our obligation now becomes constructing a decision for
+  membership of the support list.
+  This is straightforward given decidable equality.
+
+  \ExecuteMetaData[agda/Cardinality/Finite/ManifestEnumerable.tex]{to-split-enum}
+\end{proof}
+By theorem~\ref{split-enum-to-manifest-bishop}, we also can derive that any
+manifestly enumerable type with decidable equality is Bishop finite.
+\subsubsection{Closure}
+\begin{lemma}
+  Manifest enumerability is closed under dependent sum, disjoint union
+  (non-dependent sum), and Cartesian product (non-dependent product).
+\end{lemma}
+\begin{proof}
+  For these three closures, the proofs on split enumerability consisted of a
+  list manipulation followed by a proof that membership was preserved by the
+  list manipulation.
+  Because we separate these two concerns, the proofs carry over onto manifest
+  enumerability: the support list manipulation stays the same, and the
+  cover proofs are performed ``under'' the truncation.
+\end{proof}
+Notice that we do not have closure under functions: without decidability,
+manifest enumerability is not closed under function arrows.
+\subsection{Cardinal Finiteness} \label{cardinal}
+We now turn our attention to notions of finiteness which do not induce a linear
+order.
+\begin{definition}[Cardinal Finiteness]
+  A type \(A\) is cardinally finite, \(\mathcal{C}\), if it has a
+  propositionally-truncated proof of bishop finiteness.
+  \begin{equation}
+    \mathcal{C}(A) = \lVert \mathcal{B}(A) \rVert
+  \end{equation}
+\end{definition}
+\subsubsection{Characterisation}
+Though we have chosen Bishop finiteness as our underlying finiteness proof to
+truncate, we could have in fact used any equal-strength proof
+(lemma~\ref{equal-strength}).
+Because this predicate is a mere proposition, what we called before ``equal
+strength'' now implies full equivalence.
+With this proven, we can use any one of our equal-strength finite definitions
+(\(\mathcal{E}!\), \(\mathcal{B}\)).
+In particular, we have:
+\begin{lemma}
+  Cardinal finiteness is equivalent to a propositionally truncated equivalence
+  with some \(\textbf{Fin}\).
+  \begin{equation}
+    \mathcal{C}(A) \simeq \lVert \sum_{(n : \mathbb{N})} , \textbf{Fin}(n) \simeq A \rVert
+  \end{equation}
+\end{lemma}
+\begin{proof}
+  By proposition~\ref{equal-strength}, we only need to provide conversion
+  functions between the two sides of the equivalence.
+  For left to right, we have proposition~\ref{bishop-equiv}.
+  For right to left, we can transport along the equivalence, meaning we need
+  only prove that \(\textbf{Fin}(n)\) is Bishop finite.
+  We know this to be true, since it is split enumerable
+  (proposition-\ref{fin-split-enum}).
+\end{proof}
+\subsubsection{Closure}
+The closure proofs for cardinal finiteness are especially easy.
+In contrast to manifest enumerability, under the propositional truncation we
+have a full proof of bishop finiteness, meaning that all of the closure proofs
+carry over.
+\begin{lemma}
+  Cardinal finiteness is closed under dependent and non-dependent sums,
+  products, and functions.
+\end{lemma}
+\begin{proof}
+  All closure functions can be lifted under propositional truncation.
+  Therefore, cardinal finiteness has the same closure properties as manifest
+  bishop finiteness.
+\end{proof}
+\subsubsection{Strength}
+We can eliminate from cardinal finiteness in a coherently constant way:
+effectively, anything which doesn't ``notice'' the order of the internal support
+list is usable.
+
+First, even though the cardinality of the type is hidden under the truncation,
+we can pull it out, as it is constant through permutations of the list.
+\begin{theorem}
+  Given a cardinally finite type, we can derive the type's cardinality, as well
+  as a propositionally truncated proof of equivalence with \(\textbf{Fin}\)s of
+  the same cardinality.
+  \begin{equation}
+    \mathcal{C}(A) \rightarrow \sum_{(n : \mathbb{N})} , \lVert \textbf{Fin}(n) \simeq A \rVert
+  \end{equation}
+\end{theorem}
+\begin{proof}
+  We eliminate from the proof of cardinal finiteness via the eliminator
+  described in proposition~\ref{elim-non-prop}.
+  We pass a function \(\text{alg}\) which preserves the first part of the pair,
+  and truncates the second.
+  In Agda:
+  \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{trunc-alg}
+
+  It remains only to show that this function is coherently constant.
+  \begin{equation}
+    \prod_{(x : \sum_{(n : \mathbb{N})} , \textbf{Fin}(n) \simeq A)} ,
+    \prod_{(y : \sum_{(m : \mathbb{N})} , \textbf{Fin}(m) \simeq A)} ,
+    \text{alg}(x) \equiv \text{alg}(y)
+  \end{equation}
+
+  We will first address the first component of the output.
+  To show that \(n\) and \(m\) are equal, we first observe that
+  \begin{equation}
+    \textbf{Fin}(n) \equiv \textbf{Fin}(m)
+  \end{equation}
+  since both are equivalent to \(A\).
+  Relying on the fact that \(\textbf{Fin}\) is injective
+  (proposition~\ref{fin-inj}), we can derive that \(n\) and \(m\) are equal.
+
+  Since the second components of the output are propositions, they are
+  definitionally equal.
+  \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{const-alg}
+\end{proof}
+
+\begin{lemma} \label{fin-inj}
+  \(\textbf{Fin}\) is injective.
+\end{lemma}
+\begin{proof}
+  Let \(n\) and \(m\) be natural numbers.
+  We want to derive a proof of \(n \equiv m\) from \(\textbf{Fin}(n) \equiv
+  \textbf{Fin}(m)\).
+
+  We prove by contradiction (since the property we are interested in is
+  decidable, this is valid constructively).
+  When \(n\) does not equal \(m\), there are two possible cases.
+  \begin{equation}
+    \neg (n \equiv m) \implies
+    \begin{cases}
+      m \equiv 1 + n + k, n < m  \\
+      n \equiv 1 + m + k, m < n
+    \end{cases}
+  \end{equation}
+  Take the first case, without loss of generality.
+  \todo[inline]{Finish prose for this proof}
+
+  The Agda version of this proof is in the appendix (\ref{agda-fin}).
+\end{proof}
+
+Another property which does not rely on an internal order is decidable equality.
+\begin{theorem} \label{cardinal-finite-discrete}
+  Any cardinal-finite set has decidable equality.
+\end{theorem}
+\begin{proof}
+  We first show that decidable equality is a proposition itself.
+  We know that if a type \(A\) is a proposition, then the decision over that
+  type is also a proposition.
+  Then, via Hedberg's theorem, we know that any type with decidable equality is
+  a set, meaning that paths in that type are themselves propositions.
+  Therefore we can derive that a decision of equality on elements with decidable
+  equality is a proposition, and by function extensionality we see that
+  decidable equality is itself a proposition.
+
+  That is enough to pull it out of the truncation: since any Bishop-finite type
+  has decidable equality, and decidable equality is a proposition, we conclude
+  that cardinal-finite types have decidable equality.
+\end{proof}
+\subsubsection{Relation to Manifest Bishop Finiteness}
+Cardinal finiteness tells us that there is an isomorphism between a type and
+\(\mathbf{Fin}\); it just doesn't tell us \emph{which} isomorphism.
+To take a simple example, \(\mathbf{Bool}\) has 2 possible isomorphisms with the
+set \(\mathbf{Fin}(2)\): one where false maps to 0, and true to 1; and another
+where false maps to 1 and true to 0.
+
+To convert from Cardinal finiteness to Bishop finiteness, then, requires that we
+supply enough information to identity a particular isomorphism.
+A total order is sufficient here: it will give us enough to uniquely order the
+support list invariant under permutations.
+This tells us what we already knew in the introduction: manifest Bishop
+finiteness is cardinal finiteness plus an order.
+
+\begin{theorem}
+  Any cardinal finite type with a (decidable) total order is manifestly Bishop
+  finite.
+\end{theorem}
+\begin{proof}
+  This proof is quite involved, and will rely on several subsequent lemmas, so
+  we will give only its outline here.
+  \begin{itemize}
+    \item First, we will convert to manifest enumerability: knowing that the
+      underlying type is discrete (theorem~\ref{cardinal-finite-discrete}) we can
+      go from manifest enumerability to split enumerability
+      (lemma~\ref{manifest-enum-to-split}), and subsequently to manifest Bishop
+      finiteness (lemma~\ref{split-enum-to-manifest-bishop}).
+    \item To convert to manifest enumerability, we need to provide a support
+      list: this cannot simply be the support list hidden under the truncation,
+      since that would violate the hiding promised by the truncation.
+      Instead, we sort the list (using insertion sort).
+      We must, therefore, prove that insertion sort is invariant under all
+      support lists in cardinal finiteness proofs.
+    \item We show that all support lists in cardinal finiteness proofs are
+      permutations of each other.
+    \item And then we show that insertion sort is invariant under permutations.
+  \end{itemize} 
+\end{proof}
+Now we will build up the toolkit we need to perform the above steps.
+First, permutations.
+\begin{definition}[List Permutations]
+  We say that two lists are permutations of each other if there is an
+  isomorphism between membership proofs.
+  \begin{equation}
+    \mathit{xs} \leftrightsquigarrow \mathit{ys} = \prod_{(x : A)} , x \in \mathit{xs} \iff x \in \mathit{ys}
+  \end{equation}
+\end{definition}
+We also prove some of the identities you might expect with regards to
+permutations, all in the appendix.
+In particular, we prove that they form an equivalence relation.
+
+Next, we will use the following definition of insertion sort:
+\ExecuteMetaData[agda/Data/List/Sort.tex]{sort}
+\begin{lemma}
+  Insertion sort is invariant under permutations.
+  \begin{equation}
+    \mathit{xs} \leftrightsquigarrow \mathit{ys} \implies \text{sort}(\mathit{xs}) \equiv \text{sort}(\mathit{ys})
+  \end{equation}
+\end{lemma}
+\begin{proof}
+  We first prove that insertion sort does indeed sort its input: i.e. it returns
+  a sorted list that is a permutation of its input.
+  Then we show that any two sorted lists which are permutations of each other
+  are equal.
+\end{proof}
+\todo[inline]{Agda code for proofs}
+\subsection{Kuratowski Finiteness} \label{kuratowski}
+Finally we arrive at Kuratowski finiteness
+\cite{kuratowskiNotionEnsembleFini1920}.
+\subsubsection{The Kuratowski Set as a Higher Inductive Type}
+We start with the Kuratowski finite sets, which are free join semilattices (or,
+equivalently, free commutative idempotent monoids).
+HITs are required to define this type
+\cite{altenkirchDefinableQuotientsType2011}; however we have two possible
+candidates for the definition.
+The first is a direct translation of ``free commutative idempotent monoid'' into
+a type:
+\begin{equation}
+\begin{aligned}
+  \mathcal{K}(A) &=  \\
+    &\left\vert \; \text{singleton} : A \rightarrow \mathcal{K}(A) \right; \\
+    &\left\vert \; \cdot \cup \cdot : \mathcal{K}(A) \times \mathcal{K}(A) \rightarrow \mathcal{K}(A) \right; \\
+    &\left\vert \; \emptyset : \mathcal{K}(A) \right; \\
+    &\left\vert \; \cup \text{-assoc} : \Pi \left( \mathit{xs}, \mathit{ys}, \mathit{zs}:\mathcal{K}(A)\right) , (\mathit{xs} \cup \mathit{ys}) \cup \mathit{zs} \equiv \mathit{xs} \cup (\mathit{ys} \cup \mathit{zs}) \right; \\
+    &\left\vert \; \cup \text{-commutative} : \Pi (\mathit{xs}, \mathit{ys}: \mathcal{K}(A)) , \mathit{xs} \cup \mathit{ys} \equiv \mathit{ys} \cup \mathit{xs} \right; \\
+    &\left\vert \; \cup \text{-idempotent} : \Pi (\mathit{xs}: \mathcal{K}(A)) , \mathit{xs} \cup \mathit{xs} \equiv \mathit{xs} \right; \\
+    &\left\vert \; \cup \text{-identity} : \Pi (\mathit{xs}: \mathcal{K}(A)) , \emptyset \cup \mathit{xs} \equiv \mathit{xs} \right; \\
+    &\left\vert \; \text{trunc} : \Pi (\mathit{xs}, \mathit{ys}: \mathcal{K}(A)) , \Pi (p, q : \mathit{xs} \equiv \mathit{ys}) , p \equiv q \right;
+\end{aligned}
+\end{equation}
+\todo[inline]{Have we already described the trunc constructor?}
+
+However this proves cumbersome to work with.
+Instead, we build on the free monoid (the list), and add the required equations
+we need.
+\begin{equation}
+\begin{aligned}
+  \mathcal{K}(A) &=  \\
+    &\left\vert \; \cdot \dblcolon \cdot : A \times \mathcal{K}(A) \rightarrow \mathcal{K}(A) \right; \\
+    &\left\vert \; [] : \mathcal{K}(A) \right; \\
+    &\left\vert \; \text{com} : \Pi (x, y: A) , \Pi (\mathit{xs}: \mathcal{K}(A)) , x \dblcolon y \dblcolon \mathit{xs} \equiv y \dblcolon x \dblcolon \mathit{xs} \right; \\
+    &\left\vert \; \text{dup} : \Pi (\mathit{x}: A) , \Pi (\mathit{xs}: \mathcal{K}(A)) , x \dblcolon x \dblcolon \mathit{xs} \equiv x \dblcolon \mathit{xs} \right; \\
+    &\left\vert \; \text{trunc} : \Pi (\mathit{xs}, \mathit{ys}: \mathcal{K}(A)) , \Pi (p, q : \mathit{xs} \equiv \mathit{ys}) , p \equiv q \right;
+\end{aligned}
+\end{equation}
+Since associativity and identity (the monoid laws) are automatically satisfied
+by the list structure, we do not need to specify them as paths.
+Furthermore, it is simpler to specify the rest of the paths on the heads of the
+lists, rather than on the whole structure.
+That said, the two definitions are equivalent.
+\todo[inline]{Maybe add sketch of proof here?}
+\subsubsection{Elimination}
+With Agda's dependent pattern matching, we get an eliminator automatically from
+the definition of an inductive type.
+We will define two further eliminators here.
+
+\begin{equation}
+  \begin{aligned}
+    \phi &: \mathcal{K}(A) \rightarrow \text{Type} \\
+    \text{trunc}^\phi &: \pifun{x}{\mathcal{K}(A)} , \text{isSet}(\phi(x)) \\
+    []^\phi &: \phi([]) \\
+    \dblcolon^\phi &: \pifun{x}{A} , \pifun{xs}{\mathcal{K}(A)} , \pifun{p}{\phi(\mathit{xs})} , \phi(x \dblcolon \mathit{xs}) \\
+    \text{com}^\phi &: \prod_{(x , y : A)} , \pifun{xs}{\mathcal{K}(A)} , \pifun{p}{\phi(\matit{xs})} , x \dblcolon^\phi (y \dblcolon \mathit{xs}) (y \dblcolon^\phi \mathit{xs} \; p) \equiv y \dblcolon^\phi (x \dblcolon \mathit{xs}) (x \dblcolon^\phi \mathit{xs} \; p) \\
+    \text{dup}^\phi &: \pifun{x}{A} , \pifun{xs}{\mathcal{K}(A)} , \pifun{p}{\phi(\mathit{xs})} , x \dblcolon^\phi (x \dblcolon \mathit{xs}) (x \dblcolon^\phi \mathit{xs} \; p) \equiv x \dblcolon^\phi \mathi{xs} \; p
+  \end{aligned}
+\end{equation}
+
+\begin{equation} \label{kuratowski-rec}
+  \text{rec} \left\{ \begin{aligned}
+    \phi &: \text{Type} \\
+    \text{trunc}^\phi &: \text{isSet}(\phi) \\
+    []^\phi &: \phi \\
+    \dblcolon^\phi &: \pifun{x}{A} , \pifun{xs}{\mathcal{K}(A)} , \phi \\
+    \text{com}^\phi &: \prod_{(x , y : A)} , \pifun{xs}{\mathcal{K}(A)} , x \dblcolon^\phi (y \dblcolon^\phi \mathit{xs}) \equiv y \dblcolon^\phi (x \dblcolon^\phi \mathit{xs}) \\
+    \text{dup}^\phi &: \pifun{x}{A} , \pifun{xs}{\mathcal{K}(A)} , x \dblcolon^\phi (x \dblcolon^\phi \mathit{xs}) \equiv x \dblcolon^\phi \mathit{xs}
+  \end{aligned} \right.
+\end{equation}
+
+\todo[inline]{Expand on these eliminators}
+
+\subsubsection{Membership}
+We can define membership of a Kuratowski set using these eliminators.
+First we define the \(\Diamond\) type, described in the appendix.
+\begin{equation}
+  \Diamond P = \left\{ \begin{aligned}
+    \phi &= \sum_{(x : \text{Type})} , \text{isProp}(x) \\
+    \text{trunc}^\phi &= \text{isSetHProp} \\
+    []^\phi &= \bot , \text{isProp}\bot \\
+    x \dblcolon^\phi \mathit{xs} &= \lVert P(x) \uplus \text{fst}(\mathit{xs}) \rVert , \text{squash} \\
+    \text{com}^\phi &= \dots \\
+    \text{dup}^\phi &= \dots
+  \end{aligned} \right.
+\end{equation}
+We can define membership then in terms of this.
+\begin{definition}[Kuratowski Membership]
+  \begin{equation}
+    x \in \mathit{xs} = \Diamond (\equiv x) \mathit{xs}
+  \end{equation}
+\end{definition}
+
+From this we can define Kuratowski finiteness.
+\begin{definition}[Kuratowski Finiteness]
+  A type is Kuratowski finite iff there exists a Kuratowski Set which contains
+  all of its elements.
+  \begin{equation}
+    \mathcal{K}^{f}(A) = \sum_{(\text{support} : \mathcal{K}(A))} , \pifun{x}{A} , x \in \text{support}
+  \end{equation}
+\end{definition}
+
+\subsubsection{Strength}
+Since the Kuratowski set is a departure in structure from our previous
+list-based notions of finiteness, it makes sense to first look for the closest
+list-based analogue.
+As it turns out, that analogue is manifestly enumerable finiteness, with the
+order removed.
+\begin{theorem}
+  A proof of Kuratowski finiteness is equivalent to a propositionally truncated
+  proof of enumerability.
+  \begin{equation}
+    \mathcal{K}^f(A) \simeq \lVert \mathcal{E}(A) \rVert
+  \end{equation}
+\end{theorem}
+\begin{proof}
+  We prove by way of an isomorphism.
+  In the first direction (from \(\mathcal{K}\) to \(\mathcal{E}\)), we use the
+  eliminator in \ref{kuratowski-rec}.
+  Because we are eliminating into a proposition, we need not prove that the
+  function obeys the path constructors.
+  The remaining cases to deal with involve converting the Kuratowski set to a
+  list, and the cover proof to its equivalent cover proof on lists.
+  \ExecuteMetaData[agda/Cardinality/Finite/Kuratowski.tex]{to-enum}
+
+  To go the other direction, we first need to prove that a proof of Kuratowski
+  finiteness is a proposition.
+  \todo[inline]{Explain rest of this proof}
+\end{proof}
+Based on previous proofs, we can derive that if we add decidability to a
+Kuratowski finite type we retrieve cardinal finiteness.
+\begin{lemma}
+  Any Kuratowski finite set with decidable equality is cardinally finite.
+\end{lemma}
+\begin{proof}
+  \todo[inline]{Proof}
+\end{proof}
+\begin{lemma}
+  Kuratowski finiteness is equivalent to a truncated surjection from some prefix
+  of the natural numbers.
+  \begin{equation}
+    \mathcal{K}^f(A) \simeq \left\Vert \sum_{(n : \mathbb{N})} , \left(  \mathbf{Fin}(n) \twoheadrightarrow A \right) \right\Vert
+  \end{equation}
+\end{lemma}
+\begin{proof}
+  \todo[inline]{Prose for proof}
+\end{proof}
+\subsubsection{Topos}
+At this point, we see that a ``decidable Kuratowski finite set'' is precisely
+equivalent to a cardinal finite set.
+From this, we can lift over all of the properties of cardinal finite sets.
+In particular, we see that decidable Kuratowski finite sets form a \emph{topos}.
+\todo[inline]{Fill in rest}
+\subsubsection{Closure}
+\begin{theorem}
+  Kuratowski finite sets are closed under \(\sum\).
+\end{theorem}
+\begin{proof}
+  \todo[inline]{Proof}
+\end{proof}
+\section{Species and Containers} \label{species-and-containers}
+Containers \cite{abbottContainersConstructingStrictly2005} give us a general way
+to construct functors from generic building-blocks.
+\begin{definition}[Container]
+  A container is a dependent pair of shapes, and positions indexed by those
+  shapes.
+  \begin{equation}
+    \mathbf{Container} = \sum_{(\text{shape} : \text{Type})} , \text{shape} \rightarrow \text{Type}
+  \end{equation}
+  \ExecuteMetaData[agda/Data/Container.tex]{def}
+  We ``interpret'' a container into a pair of a specific shape, and a function
+  which goes from positions to contents.
+  \begin{equation}
+    \llbracket S , P \rrbracket = \prod_{(X : \text{Type})} , \sum_{(s : S)} , \left( P(s) \rightarrow X \right)
+  \end{equation}
+  \ExecuteMetaData[agda/Data/Container.tex]{interp}
+\end{definition}
+For instance, the shape of lists is given by \(\Nat\), and the positions are then
+\(\AgdaDatatype{Fin}\)s indexed by that \(\Nat\).
+\ExecuteMetaData[agda/Data/Container/List.tex]{def}
+And here is the definition of foldr for container-defined lists:
+\ExecuteMetaData[agda/Data/Container/List.tex]{foldr}
+\subsection{Containers For Finiteness}
+The definition for the container-based finiteness predicates are almost
+identical to their list-based counterparts.
+\ExecuteMetaData[agda/Cardinality/Container/Finite.tex]{split-enum}
+One advantage that they have, however, is that conversion back and forth between
+the relation-based finiteness predicates is very straightforward.
+This allows us to prove the following.
+\begin{theorem}
+  Manifest bishop finiteness is equivalent to an equivalence with some \(\mathbf{Fin}\).
+  \begin{equation}
+    \mathcal{B}(A) \simeq \sum_{(n : \mathbb{N})} , \mathbf{Fin}(n) \simeq A
+  \end{equation}
+\end{theorem}
+\begin{proof}
+  This proof is the composition of two equivalences.
+  First, we must show that the list-based manifest Bishop finiteness is
+  equivalent to the container-based definition, and then we must show that the
+  container-based definition of manifest Bishop finiteness is equivalent to the
+  equivalence.
+  \todo[inline]{proof}
+\end{proof}
+
+\begin{lemma}
+  Manifest enumerability is equivalent to a surjection from some \(\mathbf{Fin}\).
+  \begin{equation}
+    \mathcal{E}(A) \simeq \sum_{(n : \mathbb{N})} , \left(  \mathbf{Fin}(n) \twoheadrightarrow A \right)
+  \end{equation}
+\end{lemma}
+\begin{proof}
+\end{proof}
+
+\begin{lemma}
+  Split enumerability is equivalent to a split surjection from some \(\mathbf{Fin}\).
+  \begin{equation}
+    \mathcal{E}!(A) \simeq \sum_{(n : \mathbb{N})} , \left(  \mathbf{Fin}(n) \twoheadrightarrow ! A \right)
+  \end{equation}
+\end{lemma}
+\begin{proof}
+\end{proof}
+
+\section{Infinite Cardinalities} \label{infinite-cardinalities}
+While the previous section purported to be about finite sets, we can now see
+that it was really only studying surjections and isomorphisms of different
+flavours between types and \(\mathbf{Fin}\).
+The natural next question which arises, then, is if we can extend that work to
+surjections and isomorphisms with \(\mathbb{N}\).
+In more standard language, what we're referring to here is of course
+countable infinity and related concepts.
+However, as with finite types, the ``countably infinite'' types have more
+varieties in constructive mathematics than their classical counterparts.
+\subsection{Split Countable Types}
+Our first foray into the world of countable types will be a straightforward
+analogue to the split enumerable types.
+We need change only one element: instead of a support \emph{list}, we instead
+have a support \emph{stream}, which is its infinite, coinductive counterpart.
+\begin{definition}[Stream]
+  We will work with two isomorphic definitions of streams.
+  The first is the following:
+  \begin{equation}
+    \mathbf{Stream}(A) = \mathbb{N} \rightarrow A
+  \end{equation}
+  Conceptually, a stream is like a list without an end.
+  Of course, such a type can not be defined in the same way as a list: it would
+  be impossible to construct a value, as inductive types do not admit
+  infinitely-sized inhabitants.
+\end{definition}
+We can now define the split countable types.
+\begin{definition}[Split Countability]
+  \begin{equation}
+    \mathcal{E}!(A) = \sum_{(\mathit{xs} : \mathbf{Stream}(A))} , \prod_{(x : A)} , x \in \mathit{xs}
+  \end{equation}
+\end{definition}
+\subsubsection{\(\Sigma\) Closure}
+We know that countable infinity is not closed under the exponential (function
+arrow), so the only closure we need to prove is \(\Sigma\) to cover all of
+what's left.
+To do this we have to take a slightly different approach to the functions we
+defined before.
+Figure~\ref{pairings} illustrates the reason why: previously, we used the
+``Cartesian'' product pairing for each support list.
+This diverges if the first list is infinite, never exploring anything other than
+the first element in the second list.
+Instead, we use here the cantor pairing function, which performs a breadth-first
+search of the pairings of both lists.
+\begin{figure}
+  \centering
+  \begin{subfigure}[b]{.5\linewidth}
+    \begin{tikzcd}[sep=tiny,font=\footnotesize]
+      (1,e) \ar[rdddd, out=-45, in=135] & (2,e) \ar[rdddd, out=-45, in=135] & (3,e) \ar[rdddd, out=-45, in=135] & (4,e) \ar[rdddd, out=-45, in=135] & (5,e)        \\
+      (1,d) \ar[u]     & (2,d) \ar[u]    & (3,d) \ar[u]    & (4,d) \ar[u]    & (5,d) \ar[u] \\
+      (1,c) \ar[u]     & (2,c) \ar[u]    & (3,c) \ar[u]    & (4,c) \ar[u]    & (5,c) \ar[u] \\
+      (1,b) \ar[u]     & (2,b) \ar[u]    & (3,b) \ar[u]    & (4,b) \ar[u]    & (5,b) \ar[u] \\
+      (1,a) \ar[u]     & (2,a) \ar[u]    & (3,a) \ar[u]    & (4,a) \ar[u]    & (5,a) \ar[u]
+    \end{tikzcd}
+    \caption{Cartesian}
+    \label{cartesian}
+  \end{subfigure}%
+  \begin{subfigure}[b]{.5\linewidth}
+    \begin{tikzcd}[sep=tiny,font=\footnotesize]
+      (1,e) \ar[dr] & (2,e) \ar[dr]  & (3,e) \ar[dr]    & (4,e) \ar[dr] & (5,e) \\
+      (1,d) \ar[dr] & (2,d) \ar[dr]  & (3,d) \ar[dr]    & (4,d) \ar[dr] & (5,d) \ar[u] \\
+      (1,c) \ar[dr] & (2,c) \ar[dr]  & (3,c) \ar[dr]    & (4,c) \ar[dr] & (5,c) \ar[uul] \\
+      (1,b) \ar[dr] & (2,b) \ar[dr]  & (3,b) \ar[dr]    & (4,b) \ar[dr] & (5,b) \ar[uuull, out=130, in=-50] \\
+      (1,a) \ar[u]  & (2,a) \ar[uul, out=130, in=-50] & (3,a) \ar[uuull, out=130, in=-50] & (4,a) \ar[uuuulll, out=130, in=-50] & (5,a) \ar[uuuulll, out=130, in=-50]
+    \end{tikzcd}
+    \caption{Cantor}
+    \label{cantor}
+  \end{subfigure}
+  \caption{Two possible products for the sets \(\left[ 1 \dots 5 \right]\) and
+    \(\left[  a \dots e \right]\)}
+  \label{pairings}
+\end{figure}
+\begin{theorem} \label{split-countability-sigma}
+  Split countability is closed under \(\Sigma\).
+\end{theorem}
+\begin{proof}
+  Let \(\mathcal{X}\) be the proof of split countability on \(A\), and
+  \(\mathcal{Y}\) be a function of the following type:
+  \begin{equation}
+    \mathcal{Y} : \prod_{(x : A)} , \mathcal{E}!(U(x))
+  \end{equation}
+  Where \(U\) is a type family on \(A\).
+  Our task is to provide the following:
+  \begin{equation}
+    \mathcal{E}!\left(\sum_{(x : A)} , U(x)\right)
+  \end{equation}
+
+  This final proof consists of the support stream, and the proof that the
+  support stream covers the input.
+
+  As mentioned, we will have to use a more sophisticated pairing function than
+  the Cartesian product we used before.
+  We instead will mirror the pattern in figure~\ref{cantor}.
+  To greatly simplify the algorithm, we will produce an intermediate stream of
+  lists which consists of the diagonals in the diagram.
+  We then concatenate these streams into the final support stream.
+
+  Here is the algorithm in Agda:
+  \ExecuteMetaData[agda/Cardinality/Container/Infinite.tex]{conv}
+  \todo[inline]{Finish prose for this proof}
+\end{proof}
+\begin{lemma}
+  Split countability is closed under non-dependent product and sum.
+\end{lemma}
+\begin{proof}
+  Follows from theorem~\ref{split-countability-sigma}.
+\end{proof}
+\subsubsection{Kleene Star and Plus}
+While we lose some closures with the inclusion of infinite types, we gain some
+others.
+In particular, we have the Kleene star and plus.
+These allow us to introduce recursion into countable types, by way of lists.
+\begin{definition}[Kleene Lists]
+  It is often useful to have two mutually-defined list types (one for
+  possibly-empty lists and one for nonempty lists).
+  Such a definition closely mirrors the Kleene star and plus.
+  \begin{equation}
+    \begin{aligned}
+      A^+ &= \\
+          &\left\vert \; \text{head} : A^+ \rightarrow A \right; \\
+          &\left\vert \; \text{tail} : A^+ \rightarrow A^\star \right; \\
+      A^\star &= \\
+          &\left\vert \; [] : A^\star \right; \\
+          &\left\vert \; [\cdot] : A^+ \rightarrow A^\star \right;
+    \end{aligned}
+  \end{equation}
+\end{definition}
+Our definition of split countability is closed under these types in particular.
+\begin{theorem}
+  Split countability is closed under Kleene star and plus.
+\end{theorem}
+\begin{proof}
+  \tot[inline]{Proof}
+\end{proof}
+\subsection{Manifest Countability}
+As we can quotient out the position information with finite types, so can we with
+countable types.
+\todo[inline]{fill in rest here}
+\section{Practical Uses} \label{practical}
+\todo[inline]{%
+\begin{itemize}
+  \item Synthesize functions according to specifications.
+  \item Nonexhaustive search for the infinite types.
+  \item Lift quantifiers over types other than bishop (i.e. K). Show that
+    quantifiers have to be different (i.e. existence has to be unique).
+  \item Big operators\cite{gustafssonCountingTypeIsomorphisms}
+    \cite{gustafssonFoldableContainersDependent}
+\end{itemize}
+}
+The theory of finite types in constructive mathematics, and in HoTT in
+particular, is rich and interesting, as we hope we have demonstrated thus far.
+As well as being theoretically interesting, however, the proofs and combinators
+we have defined here are \emph{practically} useful.
+\subsection{Omniscience}
+\todo[inline]{Put definition for decidable somewhere}
+In this section we are interested in restricted forms of the limited principle
+of omniscience \cite{myhillErrettBishopFoundations1972}.
+\begin{definition}[Limited Principle of Omniscience]
+  For any type \(A\) and predicate \(P\) on \(A\), the limited principle of
+  omniscience is as follows:
+  \begin{equation}
+    \left( \prod_{(x : A)} , \mathbf{Dec}(P(x)) \right) \rightarrow \mathbf{Dec} \left( \sum_{(x : A)} , P(x) \right)
+  \end{equation}
+  In other words, for any decidable predicate the existential quantification of
+  that predicate is also decidable.
+\end{definition}
+The limited principle of omniscience is non-constructive, but individual types
+can themselves satisfy omniscience.
+In particular, \emph{finite} types are omniscient.
+Omniscience is defined in Agda like so:
+\ExecuteMetaData[agda/Relation/Nullary/Omniscience.tex]{omniscient}
+
+There is also a universal form of omniscience, which we call exhaustibility.
+\begin{definition}[Exhaustibility]
+  We say a type \(A\) is exhaustible if, for any decidable predicate \(P\) on
+  \(A\), the universal quantification of the predicate is decidable.
+  \begin{equation}
+    \left( \prod_{(x : A)} , \mathbf{Dec}(P(x)) \right) \rightarrow \mathbf{Dec} \left( \prod_{(x : A)} , P(x) \right)
+  \end{equation}
+  \ExecuteMetaData[agda/Relation/Nullary/Omniscience.tex]{exhaustible}
+\end{definition}
+Omniscience is stronger than exhaustibility, as we can derive the latter from
+the former:
+\begin{lemma} \label{omniscient-is-exhaustible}
+  Any omniscient type is exhaustible.
+\end{lemma}
+\begin{proof}
+  \todo[inline]{Proof}
+  \ExecuteMetaData[agda/Relation/Nullary/Omniscience.tex]{weaken}
+  \todo[noline]{stability in appendix}
+\end{proof}
+We cannot derive, however, that any exhaustible type is omniscient: to do so
+would require us to choose a representative from an arbitrary type, which is not
+possible constructively.
+
+\subsubsection{Split Enumerability}
+We now show that split enumerable types are omniscient.
+\begin{theorem}
+  Any split enumerable type is omniscient.
+\end{theorem}
+\begin{proof}
+  Let \(E_A\) be a proof of split enumerability on a type \(A\), and \(P\) be a
+  decidable predicate on the type \(A\).
+  First, we lift the decidable predicate onto one on lists, using the
+  \(\Diamond\) type described earlier.
+  Then, we run the predicate on the support list of \(E_A\).
+
+  In the case that it is true for any member of the list, we return that element
+  along with its proof.
+
+  If the predicate fails for every member of the list, we say that the whole
+  test has failed.
+  To support this conclusion, we reason that if a predicate is false for every
+  item in a list \(\mathit{xs}\), and some element \(x\) is in that list, then
+  the predicate must be false for that item.
+
+  In Agda:
+  \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{exists-quant}
+\end{proof}
+\begin{theorem}
+  Any split enumerable type is exhaustible.
+\end{theorem}
+\begin{proof}
+  Via lemma~\ref{omniscient-is-exhaustible}.
+\end{proof}
+
+\subsubsection{Manifest Enumerability}
+We now reach for the same quantification combinators we used for
+split enumerable types.
+Because neither of these quantifications ``care'' about the specific proof of
+membership used, they can both be used.
+\begin{theorem}
+  For a decidable predicate \(P\) on a type \(A\), if the type is manifestly
+  enumerable, then we can decide the existential of \(P\).
+\end{theorem}
+\begin{proof}
+  \todo[inline]{Proof}
+\end{proof}
+
+\begin{theorem}
+  For a decidable predicate \(P\) on a type \(A\), if the type is manifestly
+  enumerable, then we can decide the universal of \(P\).
+\end{theorem}
+\begin{proof}
+  \todo[inline]{Proof}
+\end{proof}
+\subsubsection{Cardinal Finiteness}
+The quantification operations partially carry over.
+\begin{theorem}
+  For a decidable predicate \(P\) on a type \(A\), if the type is cardinal
+  finite, then we can decide the universal of \(P\).
+\end{theorem}
+\begin{proof}
+  \todo[inline]{Proof}
+\end{proof}
+
+An existential predicate is harder to express.
+In previous sections, we simply tested each member of the set in turn, and
+returned the first member which passed the test.
+Clearly this operation can depend on the underlying order.
+\begin{theorem}
+  For a decidable predicate \(P\) on a type \(A\), if the type is cardinal
+  finite, then we can decide the propositionally truncated existential of \(P\).
+\end{theorem}
+\begin{proof}
+  \todo[inline]{proof}
+\end{proof}
+\begin{theorem}
+  For a decidable predicate \(P\) on a type \(A\), which uniquely determines an
+  input, we can decide the existential of \(P\).
+\end{theorem}
+\begin{proof}
+  \todo[inline]{proof}
+\end{proof}
+\subsubsection{Kuratowski}
+\begin{itemize}
+\item Forall
+\item Exists
+\end{itemize}
+
+\subsection{Synthesising Pattern-Matching Proofs}
+In particular, they can automate large proofs by analysing every possible case.
+In \cite{firsovDependentlyTypedProgramming2015}, the \(\AgdaDatatype{Pauli}\)
+group is used as an example.
+\ExecuteMetaData[agda/Data/Pauli.tex]{def}
+For this type, there are several practical tasks we hope to achieve with the
+help of our finiteness combinators:
+\begin{itemize}
+  \item Define decidable equality on the type
+  \item Define the group operation on the type
+  \item Prove properties about the group operation
+\end{itemize}
+Unfortunately, the simple pattern-matching way to do many of these tasks is
+prohibitively verbose.
+As \(\AgdaDatatype{Pauli}\) has 4 constructors, \(n\)-ary functions on
+\(\AgdaDatatype{Pauli}\) may require up to \(4^n\) cases.
+A proof of decidable equality is one such function: it can be seen fully
+worked-through in the appendix.
+
+The alternative is to derive the things we need from \(\Enum\) somehow.
+First, then, we need an instance for \(\AgdaDatatype{Pauli}\):
+\ExecuteMetaData[agda/Data/Pauli.tex]{inst}
+From here we can already derive decidable equality:
+\ExecuteMetaData[agda/Data/Pauli.tex]{good-eq}
+
+Next, we want to prove some things about the group operation itself:
+\ExecuteMetaData[agda/Data/Pauli.tex]{group-op}
+The first property is the following:
+\ExecuteMetaData[agda/Data/Pauli.tex]{cancel}
+Tough the proof is simple enough to write by hand, it's a good place to start
+for demonstrating the automation technique.
+
+First, we will witness the fact that if a property is true for every element in
+a finite type's support list, it is true for every element in the type (and vice
+versa).
+\ExecuteMetaData[agda/Cardinality/Finite/Search.tex]{all-comb}
+We can also lift a decidable property on elements of the type to a decidable
+property on \emph{lists} of elements of the type.
+\ExecuteMetaData[agda/Data/List/Relation/Unary.tex]{dec-all}
+Combining these, we can turn a decidable predicate on a set into a decidable
+\emph{universal property} on the set.
+\ExecuteMetaData[agda/Cardinality/Finite/Search.tex]{lift-forall}
+
+To use this as a proof, we can run the decision procedure at type checking,
+using the following:
+\ExecuteMetaData[agda/Cardinality/Finite/Search.tex]{auto}
+This function uses the \(\AgdaDatatype{True}\) type, defined like so:
+\ExecuteMetaData[agda/Relation/Nullary.tex]{true-def}
+For a decision \(d\), \(\AgdaDatatype{True}\; d\) will resolve to \(\top\) when
+the decision is \(\AgdaInductiveConstructor{yes}\) (i.e. when the proposition is
+true), or \(\bot\) if the decision is \(\AgdaInductiveConstructor{no}\).
+Then, we require an instance of the resulting type to be in scope: this
+requirement can only be satisfied for \(\top\), thereby providing a proof that
+the decision was indeed \(\AgdaInductiveConstructor{yes}\).
+
+And finally we can prove the property we wanted to on \(\AgdaDatatype{Pauli}\)
+like so:
+\ExecuteMetaData[agda/Data/Pauli.tex]{noninst-prf}
+
+As a quick aside, when the property is \emph{not} true, for instance:
+\ExecuteMetaData[agda/Data/Pauli.tex]{contra}
+Agda will fail by not finding an instance of \(\bot\).
+The error message specifically is:
+\begin{displayquote}
+  \verb+No instance of type ⊥ was found in scope.+
+\end{displayquote}
+We can actually display a counterexample, by defining a custom empty type
+parameterised by the counterexample itself.
+\ExecuteMetaData[agda/Cardinality/Finite/Search.tex]{counterexample-def}
+Combined with some other changes to the combinators, this will give the
+following more helpful error message:
+\begin{displayquote}
+  \verb+No instance of type Counterexample X was found in scope.+
+\end{displayquote}
+\subsection{Instances}
+Running decision procedures during typechecking isn't the only use of instances:
+we can also use them similarly to how typeclasses are used in Haskell, to
+automate simple derivation procedures.
+For instance, in Haskell we can define a class for generating textual
+representations of data:
+\ExecuteMetaData[haskell/Introduction.tex]{show-cls}
+With instances that look like this:
+\ExecuteMetaData[haskell/Introduction.tex]{show-inst}
+We can have instances \emph{depend} on other instances, like the following:
+\ExecuteMetaData[haskell/Introduction.tex]{show-deriv}
+In this way, it gets rid of a tremendous about of boilerplate for more complex
+types, automatically picking out the correct instance when it can.
+
+In Agda, as in Haskell, the problem of overlapping instances often presents
+itself when more complex typeclass-based machinery is used.
+While there are ways to deal with overlapping instances, it is usually better to
+avoid them altogether, which is precisely what we do in our approach.
+We provide instances for the semiring-based combinators we defined in
+\todo[noline]{reference here}
+, and require the user to provide an instance for their own type.
+We don't, for example, provide an instance that itself searches for surjections
+from other instances: this instance is not uniquely determined.
+
+After all of that, the change to our \(\AgdaFunction{∀↯}\) is simple: we just
+change the first parameter from explicit to an instance.
+\ExecuteMetaData[agda/Cardinality/Finite/Search.tex]{auto-inst}
+
+The change to the \(\AgdaDatatype{Pauli}\) proof of cancellation isn't
+groundbreaking at first:
+\ExecuteMetaData[agda/Data/Pauli.tex]{cancel-auto}
+But the real benefit is now we can automate proofs over \emph{tuples}, allowing
+us to prove, for instance, commutativity.
+\ExecuteMetaData[agda/Data/Pauli.tex]{comm-uncurried}
+\subsection{Generic Currying and Uncurrying}
+While we have arguably removed the bulk of the boilerplate from the automated
+proofs, there is still the case of the ugly noise of currying and uncurrying.
+In this section, we take inspiration from
+\cite{allaisGenericLevelPolymorphic2019} to develop a small interface to generic
+\(n\)-ary functions and properties.
+What we provide here differs from that work in the following ways:
+\begin{itemize}
+  \item Our generic representation can handle dependent \(\sum\) and \(\prod\)
+    types (rather than their non-dependent counterparts, \(\times\) and
+    \(\rightarrow\)).
+    This extension was necessary for our use case: it is mentioned in the paper
+    as the obvious next step. 
+  \item We implement the curry-uncurry combinators as (verified) isomorphisms.
+    Since we are in a cubical setting, this gives us equivalences between the
+    types, a feature not available in standard Agda.
+  \item We deal with implicit and instance arguments generically.
+\end{itemize}
+A full explanation of our implementation is beyond the scope of this work, but
+we will mention the key parts here.
+First, we define a function arrow generic over the application method:
+\ExecuteMetaData[agda/Data/Product/N-ary.tex]{fun-arr}
+Then, we prove that it is isomorphic to the normal function arrow:
+\ExecuteMetaData[agda/Data/Product/N-ary.tex]{fun-iso}
+This step will allow us to write the curry-uncurry proofs once, and then extend
+them to the three different argument forms without difficulty.
+
+We do the same with dependent function types:
+\ExecuteMetaData[agda/Data/Product/N-ary.tex]{pi-iso}
+
+Next, we need to make our machinery for multiple arguments.
+Here we will define a generic tuple, indexed by some \(\Nat\).
+As observed in \cite{allaisGenericLevelPolymorphic2019}, it's important to
+\emph{not} implement this as an inductively defined vector, e.g.
+\ExecuteMetaData[agda/Data/Vec.tex]{def}
+As this will not give us the correct \(\eta\)-equality we need for unification.
+\todo[noline]{Stick reference to iterated Vec definition}
+Once we have a unification-friendly vector type, we can use it to implement our
+generic (and level-polymorphic) tuples.
+\ExecuteMetaData[agda/Data/Product/N-ary.tex]{tuple}
+We decided against \(\top\)-terminated tuples, as the actual contents of the
+tuple type are exposed in the interface.
+
+Finally, we can prove isomorphisms between the curried and uncurried versions of
+functions:
+\ExecuteMetaData[agda/Data/Product/N-ary.tex]{curry-n-iso}
+And dependent functions:
+\ExecuteMetaData[agda/Data/Product/N-ary.tex]{curry-n-pi-iso}
+With these combinators, we can implement the search functions in an
+arity-generic way:
+\ExecuteMetaData[agda/Cardinality/Finite/Search.tex]{all-nary}
+And automate away our proofs:
+\ExecuteMetaData[agda/Data/Pauli.tex]{prf-curried}
+
+\begin{subappendices}
+
+\section{Notation, Cubical Type Theory, and Agda}
+Our chosen setting is Cubical Type Theory (CuTT)
+\cite{cohenCubicalTypeTheory2016}, which is closely related to Homotopy Type
+Theory (HoTT) \cite{hottbook}.
+We will go through the basics and some notation here.
+\subsection{Martin-Löf Type Theory}
+CuTT has its roots in Martin-Löf intuitionistic type theory
+\cite{martin-lofIntuitionisticTypeTheory1980}.
+We will describe our notation for the MLTT parts of CuTT here (although our
+notation is quite standard).
+
+The empty or \(\mathbf{0}\) type will be denoted by \(\bot\), and the
+\(\mathb{1}\) type by \(\top\).
+The single inhabitant of \(\top\) is called \(\text{tt}\).
+
+``not'' is expressed as follows:
+\begin{equation}
+  \neg A = A \rightarrow \bot
+\end{equation}
+In other words, to say a type is impossible we say we can derive \(\bot\) from
+it.
+
+We also use the \(\mathbf{2}\) type, which we call \(\textbf{Bool}\).
+Its two canonical inhabitants are \(\text{true}\) and \(\text{false}\).
+
+We will use the term type family (or simply ``family'') to mean a non-dependent
+function with a codomain in Type.
+
+Dependent function types, or \(\prod\)-types, use the following notation:
+\begin{equation}
+  \prod_{(x : A)} , B (x)
+\end{equation}
+In Agda we use a single arrow to denote the same:
+\ExecuteMetaData[agda/Snippets/Introduction.tex]{pi}
+When the type of the left-hand-side can be inferred, we are permitted to use the
+following notation:
+\ExecuteMetaData[agda/Snippets/Introduction.tex]{forall}
+For non-dependent functions we use the single arrow (\(\rightarrow\)).
+
+Dependent pairs, or \(\sum\)-types, use a similar notation to \(\prod\)-types.
+\begin{equation}
+  \sum_{(x : A)} , B (x)
+\end{equation}
+In Agda:
+\ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma}
+Again, we get a shorthand when the type of the first element can be inferred:
+\ExecuteMetaData[agda/Snippets/Introduction.tex]{exists}
+For non-dependent pairs we use the \(\times\) symbol.
+We access the two fields of \(\sum\) types with \(\text{fst}\) and
+\(\text{snd}\):
+\begin{align}
+  \text{fst} &: \sum_{(x : A)} , B (x) \rightarrow A \\
+  \text{snd} &: \prod_{(x : \sum_{(y : A)} , B (y))} , B (\text{fst}(x))
+\end{align}
+
+Finally, the identity type is denoted by the triple equals (\(\equiv\)).
+The normal equals (\(=\)) is reserved for definitions.
+
+\subsection{Agda Basics}
+Agda \cite{norellDependentlyTypedProgramming2008} is a dependently-typed
+functional programming language, with roots in Haskell and ML.
+The core type theory is based on MLTT, though support for CuTT was added in the
+last year \cite{vezzosiCubicalAgdaDependently2019}.
+\subsubsection{Declarations}
+The anatomy of a simple Agda declaration is thus:
+\ExecuteMetaData[agda/Snippets/Introduction.tex]{abool}
+We first have a type signature: the name of the variable is to the left of the
+colon, and its type to the right.
+Below, we give an equation: as Agda is pure, it suffices to imagine that
+everywhere the name on the left of the equals sign is seen, we can replace it
+with the value on the right.
+\subsubsection{Functions}
+Functions are usually declared with pattern matching.
+\ExecuteMetaData[agda/Snippets/Introduction.tex]{not}
+Again, each clause is an equation.
+Agda's purity will ensure that these clauses hold true definitionally\footnotemark.
+\footnotetext{
+  Actually, it is possible to construct functions whose equations
+  hold propositionally, but not definitionally.
+  An example might be the following:
+  \ExecuteMetaData[agda/Snippets/Introduction.tex]{non-exact-split}
+  This is actually attempted by a lot of Agda beginners: often the compiler will
+  complain ``couldn't match type \(n\) with \(n + 0\)'', so the programmer will
+  simply add it as a clause.
+  Unfortunately, this actually over complicates things in the long run.
+  Neither the second nor third clause here will be definitionally equal: if you
+  have an obligation \(n \oplus 0 \qeq n\), the first pattern (\(0 \oplus m = m\))
+  has to be ruled out first, which means you'll have to do a manual pattern
+  match in your proof.
+  Similarly for the third pattern: to prove the obligation (\((1 + n) \oplus m
+  \qeq 1 + (n \oplus m)\)), \emph{both} preceding patterns have to be checked
+  against first.
+
+  This ordering requirement is introduced by overlapping patterns, and Agda
+  warns against it.
+  Knowing when to avoid it and when it doesn't matter is one of the aspects of
+  programming and proving in Agda which is more of an art form.
+}
+
+As Agda is a functional language, functions can also be defined with lambda
+expressions.
+\ExecuteMetaData[agda/Snippets/Introduction.tex]{id}
+
+Finally, ``mixfix'' operators play a large role in Agda's syntax.
+We can have binary operators:
+\ExecuteMetaData[agda/Snippets/Introduction.tex]{add}
+Unary operators (here, a prefix increment operator):
+\ExecuteMetaData[agda/Snippets/Introduction.tex]{unary-inc}
+Ternary operators:
+\ExecuteMetaData[agda/Snippets/Introduction.tex]{ifte}
+And so on.
+\subsubsection{Data Type Definitions}
+We can define new data types in Agda by listing their constructors:
+\ExecuteMetaData[agda/Snippets/Introduction.tex]{maybe}
+\subsubsection{Termination}
+\todo[inline]{Coinductive types, termination checking, etc.}
+\subsubsection{Universe Levels}
+All dependently-typed programming languages have to contend, at some point or
+another, with Girard's paradox
+\cite{girardInterpretationFonctionelleElimination1972}.
+The paradox concerns the type of types: it's the type-theory analogue of
+Russell's paradox.
+
+Unfortunately, most of the ways to avoid the paradox add significant clutter to
+the type theory, and increase the complexity of even simple declarations.
+Agda's answer is no exception.
+
+In Agda, every type is tagged with a ``universe level''.
+The basic types like \(\AgdaDatatype{Bool}\) live in universe level 0, the type
+of types lives in universe level 1, the type of types in universe level 1 lives
+in universe level 2, and so on.
+We won't go into it much more than that: usually there is an obvious way to
+handle the universe level of a given expression ``properly'', and cases where
+universe levels do something interesting come up rarely (although
+section~\ref{practical} does touch on one).
+We only mention it here to explain the expressions like \(A :
+\AgdaDatatype{Type} \; a\) which show up in many examples: \(a\) there is the
+universe level of the type \(A\).
+\subsection{Cubical Basics}
+This will not serve as an introduction CuTT,
+but rather a brief overview of the parts of the theory we will be using.
+For a fuller reference we direct the reader first to the ``HoTT Book''
+\cite{hottbook}, then the paper on Cubical Type Theory
+\cite{cohenCubicalTypeTheory2016}, and finally the paper on Cubical Agda
+\cite{vezzosiCubicalAgdaDependently2019}.
+\subsubsection{Paths}
+Central to CuTT is the \emph{path} type.
+This is a path in the topological sense, although we use it as a representation
+for equality (which makes sense: central to HoTT is the
+equivalence of these two concepts).
+In most of the code presented here this will not behave in any different way to
+the traditional identity type in Martin-Löf type theory: one exception to this
+is pattern-matching, which is not available on paths in Cubical Agda currently.
+As a result, we have to use its eliminator directly (if needed), although
+usually higher-level combinators like \(\AgdaFunction{subst}\) are sufficient.
+
+We will use \(\equiv\) as a notation for path types as well as the identity
+types.
+Although an identity type does exist in Cubical Type Theory, we will not use it,
+so we do not fear much confusion on this point.
+\subsubsection{Homotopy Levels}
+Next, we have the concept of ``homotopy levels''.
+These classify types in homotopy type theory according to their degree of
+truncation.
+Homotopy level 1, for instance, is the ``mere propositions'': the types for
+which every element is equal to every other.
+As a predicate, we can define \(\AgdaFunction{isProp}\) like so:
+\begin{equation}
+  \text{isProp}(A) = \prod_{(x : A)} , \prod_{(y : A)} , x \equiv y
+\end{equation}
+In Agda:
+\ExecuteMetaData[agda/Snippets/Introduction.tex]{isProp}
+
+One level up from the propositions are the (homotopy) sets.
+These are types for which all paths are themselves propositions.
+\begin{equation}
+  \text{isSet}(A) = \prod_{(x : A)} , \prod_{(y : A)} , \text{isProp}(x \equiv y)
+\end{equation}
+In Agda:
+\ExecuteMetaData[agda/Snippets/Introduction.tex]{isSet}
+
+Finally, one level down from the propositions are the contractions.
+These are types for which every element is equal to some specific element.
+\begin{equation}
+  \text{isContr}(A) = \sum_{(x : A)} , \prod_{(y : A)} , x \equiv y
+\end{equation}
+\ExecuteMetaData[agda/Snippets/Introduction.tex]{isContr}
+\subsubsection{Univalence}
+\todo[inline]{Flesh out}
+To describe this we first need fibers:
+\begin{equation}
+  \text{fiber}_f(y) = \sum_{(x : A)} , f (x) \equiv y
+\end{equation}
+\ExecuteMetaData[agda/Snippets/Introduction.tex]{fiber}
+From this we build the notion of an ``equivalence'': a function whose fibers are
+all contractions.
+\begin{equation}
+  A \simeq B = \sum_{(f : A \rightarrow B)}, \prod_{(y : B)} , \text{isContr}(\text{fiber}_f(y))
+\end{equation}
+\ExecuteMetaData[agda/Snippets/Introduction.tex]{equiv}
+The univalence axiom is that all equivalences are equivalent to the identity
+type.
+\begin{definition}[The Univalence Axiom]
+  \begin{equation}
+    \simeq \; \simeq \; \equiv
+  \end{equation}
+\end{definition}
+In practice this means that we can prove equivalence between two types and then
+transport code on one to code on the other.
+It also means that isomorphisms can be upgraded to fully-fledged equalities.
+\begin{definition}[Isomorphism]
+  An isomorphism between two types \(A\) and \(B\) is a pair of functions:
+  \begin{align}
+    f : A \rightarrow B \\
+    g : B \rightarrow A
+  \end{align}
+  Such that they are inverses of each other.
+  \begin{align}
+    \prod_{(x : A)} , g(f(x)) \equiv x \\
+    \prod_{(y : B)} , f(g(y)) \equiv y
+  \end{align}
+\end{definition}
+\ExecuteMetaData[agda/Snippets/Introduction.tex]{iso}
+\subsubsection{Higher Inductive Types}
+Along with univalence, the other significant benefit we get from path types is
+the ability to define new paths on types.
+A direct use of this feature is synthetic topology.
+We can, for instance, define the circle \(S^1\) as a higher-inductive type.
+\ExecuteMetaData[agda/Snippets/Introduction.tex]{circle}
+
+Another important use of higher inductive types is quotient types.
+These types are unavailable in ML type theory, but in Cubical Type Theory we can
+declare a quotient (here, the type of Kuratowski finite sets) like so:
+\ExecuteMetaData[agda/Algebra/Construct/Free/Semilattice.tex]{def}
+The first two constructors are standard data constructors, which mimic those for
+the list in Haskell.
+The next two add the equations we want our type to obey: we don't want it to
+care about order or duplicates.
+The last constructor \emph{truncates} the type, ensuring that it is a set.
+If we hadn't included it paths would not be mere propositions: there would be
+more than one path between any given elements of a type.
+The ability to have multiple paths is what allows Cubical Type Theory to do the
+more advanced things in topology etc., but we don't need its power here.
+
+One HIT in particular will be of great use to us: the HIT for propositional
+truncation, denoted with \(\lVert \cdot \rVert\).
+We will go through it in detail in section~\ref{prop-trunc}.
+\section{Supplementary Definitions}
+This section will serve as a reference for any definitions used in the text
+which are either not important enough to be examined inline, or common enough to
+justify elision.
+Many of the definitions come straight from the ``HoTT Book'' \cite{hottbook}.
+\subsection{Decidable Types}
+It is important to be able to talk about ``decisions'' for things like decidable
+equality, and so on.
+\begin{definition}[Decision]
+  A decision for some type \(P\) is represented by the type \(\mathbf{Dec}(A)\),
+  which has two constructors:
+  \begin{equation}
+    \begin{alignedat}{2}
+      \text{yes} &:      &A \rightarrow \mathbf{Dec}(A) \\
+      \text{no}  &: \neg &A \rightarrow \mathbf{Dec}(A)
+    \end{alignedat}
+  \end{equation}
+\end{definition}
+
+Internally, we actually use a rather more complicated type, which has preferable
+computational behaviour.
+The interface is the same, however.
+\subsection{Propositional Truncation} \label{prop-trunc}
+Essentially, a value of type \(\lVert A \rVert\) proof that an \(A\) exists,
+without revealing which \(A\) in particular.
+In HoTT, it allows us to work in a setting similar to classical logic.
+In Agda its definition is:
+\ExecuteMetaData[agda/Snippets/Introduction.tex]{prop-trunc}
+Any eliminators from this type must not distinguish between different values of
+\(A\).
+
+\begin{lemma}[Propositional Elimination]
+  We can eliminate from the type \(\lVert A \rVert\) to \(B\) with a function
+  \(A \rightarrow B\) if B is a proposition.
+\end{lemma}
+\begin{proof}
+  By case-analysis on \(\lVert A \rVert\).
+  In the first case, we apply our function.
+  In the second, we are required to show that that application agrees on any two
+  values of \(\lVert A \rVert\) \(\lVert A \rVert\).
+  This is true as \(B\) is a proposition.
+  In Agda:
+  \ExecuteMetaData[agda/HITs/PropositionalTruncation/Proofs.tex]{rec}
+\end{proof}
+
+Since \(\lVert A \rVert\) is itself a proposition, we can work ``under'' the
+truncation as much as we like.
+In fact the type \(\lVert \cdot \rVert\) itself forms a monad, allowing us to
+use monadic notation.
+\begin{lemma}
+  \(\lVert \cdot \rVert\) is a monad, with bind being the eliminator defined
+  above, and return being the \(\lvert \cdot \rvert\) constructor.
+\end{lemma}
+\begin{proof}
+  The monad laws are satisfied trivially, as \(\lVert \cdot \rVert\) is a
+  proposition.
+\end{proof}
+
+We can further eliminate into a type that isn't a proposition, as long as we
+prove that the function used in the eliminator doesn't discriminate between
+different \(A\)s in \(\lVert A \rVert\).
+\begin{lemma}[Propositional Elimination by Coherence] \label{elim-non-prop}
+  We can eliminate from a propositionally-truncated type into a non-proposition,
+  as long as the function is coherently constant
+  \cite{krausGeneralUniversalProperty2015}.
+\end{lemma}
+\begin{proof}
+  \todo[inline]{sketch proof}
+\end{proof}
+
+From these eliminators, we can derive some useful lemmas.
+\begin{lemma}[Propositional Refutation]
+  While a propositionally truncated proof is not enough to prove the
+  non-truncated type, it is enough to \emph{refute} the non-truncated negation
+  of the type.
+  \begin{equation}
+    \lVert A \rVert \rightarrow \neg (\neg A)
+  \end{equation}
+\end{lemma}
+\begin{proof}
+  We will use propositional elimination.
+  First, to go from \(A\) to \((A \rightarrow \bot) \rightarrow \bot\) is
+  trivial.
+  Secondly, to show that \((A \rightarrow \bot) \rightarrow \bot\) is a
+  proposition, we can use the fact that \(\bot\) is a proposition, and then
+  derive (via function extensionality) that any \(A \rightarrow \bot\) is also a
+  proposition.
+\end{proof}
+
+\begin{lemma}[Recomputation]
+  If a type \(A\) is decidable, and there exists a proof of the type \(\lVert A
+  \rVert\), we can derive a proof of \(A\).
+  In other words, the following function exists:
+  \begin{equation}
+    \text{recompute} : \textbf{Dec}(A) \times \lVert A \rVert \rightarrow A
+  \end{equation}
+\end{lemma}
+\begin{proof}
+  By case analysis of the decision over \(A\), we know that the false case must
+  be untrue (since we can refute it using propositional refutation), and we can
+  therefore extract the true case.
+\end{proof}
+
+Occasionally we will call families ``predicates'' when we mean to treat the
+family as a property on the domain.
+Formally this might imply that the codomain is a proposition (i.e. isomorphic to
+either \(\top\) or \(\bot\)), but it is not always necessary to prove this fact,
+and indeed occasionally we will want to use some of the information stored in
+the predicate, making it not propositional.
+For instance, a natural number being ``not prime'' we would refer to as a
+predicate, but the proof of this fact may be a list of the number's prime
+factors, so we will not truncate it away.
+
+When we say that two predicates \(P\) and \(Q\) are ``equal strength'', we mean
+there are a pair of functions \(P \rightarrow Q\) and \(Q \rightarrow P\).
+This is distinct from \(P\) and \(Q\) being isomorphic, as we haven't shown
+anything regarding the relationship between those two functions.
+However, if \(P\) and \(Q\) are themselves propositions, this does indeed imply
+an isomorphism.
+\begin{lemma} \label{equal-strength}
+  For any two types \(A\) and \(B\), if there are functions \(f : A \rightarrow
+  B\) and \(g : B \rightarrow A\), \(\lVert A \rVert\) and \(\lVert B \rVert\)
+  are equivalent.
+\end{lemma}
+\begin{proof}
+  We prove equivalence by way of isomorphism.
+
+  Let \(A\), \(B\), \(f\), and \(g\) be defined as above.
+  For isomorphism, we first provide a pair of functions \(\lVert f \rVert\) and
+  \(\lVert g \rVert\), which are defined by lifting \(f\) and \(g\) under
+  propositional truncation, respectively.
+
+  Next, we need to prove that \(g\) is a section of \(f\).
+  \begin{equation}
+    \text{section}(f, g) = f \circ g \equiv id
+  \end{equation}
+  Since \(\lVert B \rVert\) is a proposition, this is true by function
+  extensionality.
+
+  Finally, \(f\) is a retraction of \(g\).
+  \begin{equation}
+    \text{retraction}(f, g) = g \circ f \equiv id
+  \end{equation}
+  Again, since \(\lVert A \rVert\) is a proposition, this is true by function
+  extensionality.
+
+  In Agda:
+  \ExecuteMetaData[agda/HLevels/Proposition.tex]{prop-iff}
+\end{proof}
+\subsection{Functions}
+\subsubsection{Surjections}
+The following is the definition of surjectivity in HoTT:
+\begin{equation}
+  \text{Surjective}(f) = \prod_{(y : B)} , \lVert \text{fiber}_f(y) \rVert
+\end{equation}
+\ExecuteMetaData[agda/Function/Surjective.tex]{surj-def}
+We are also interested in \emph{split} surjectivity, where we don't truncate the
+fiber:
+
+\begin{equation}
+  \text{SplitSurjective}(f) = \prod_{(y : B)} , \text{fiber}_f(y)
+\end{equation}
+\ExecuteMetaData[agda/Function/Surjective.tex]{split-surj-def}
+
+A surject\emph{ion}, then, is a function which is surject\emph{ive}:
+\begin{equation}
+  A \twoheadrightarrow B = \sum_{(f : A \rightarrow B)} \text{SplitSurjective}(f)
+\end{equation}
+\ExecuteMetaData[agda/Function/Surjective.tex]{surjection}
+\subsubsection{Injections}
+The standard definition of an injective function:
+\ExecuteMetaData[agda/Function/Injective.tex]{injective-def}
+An injection:
+\ExecuteMetaData[agda/Function/Injective.tex]{injection-def}
+\subsection{Fin}
+\(\textbf{Fin}\) is a finite prefix of the natural numbers.
+Semi-formally:
+\begin{equation}
+  \textbf{Fin}(n) = \left[ 0 \dots n \right)
+\end{equation}
+There are actually a number of ways to define \(\textbf{Fin}\) in type theory.
+The method we use here is as follows:
+\ExecuteMetaData[agda/Data/Fin.tex]{def}
+\subsection{Split Enumerability}
+Here is how we define it in Agda:
+\ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{definition}
+\subsection{Predicates Over Lists}
+These two well-known dual types for quantifying a predicate over a list are used
+extensively in this work.
+We will define them using the indexing operator \(!\), which has the following
+type:
+\begin{equation}
+  ! : \prod_{(\mathit{xs} : \textbf{List}(A))} , \textbf{Fin}(\text{length}(\mathit{xs})) \rightarrow A
+\end{equation}
+For some predicate \(P\), the two definitions are as follows:
+\begin{definition}[Any]
+  \begin{equation}
+    \Diamond P \mathit{xs} = \sum_{(i : \textbf{Fin}(\text{length}(\mathit{xs})))} , P(xs \ind i)
+  \end{equation}
+  \ExecuteMetaData[agda/Data/List/Relation/Unary.tex]{any-def}
+\end{definition}
+\begin{definition}[All]
+  \begin{equation}
+    \Box P \mathit{xs} = \prod_{(i : \textbf{Fin}(\text{length}(\mathit{xs})))} , P(xs \ind i)
+  \end{equation}
+  \ExecuteMetaData[agda/Data/List/Relation/Unary.tex]{all-def}
+\end{definition}
+We are using the tradition modal logic operators here: \(\Diamond\)
+means ``any'', and \(\Box\) means ``all''.
+\subsection{Membership}
+Several other constructions are defined in terms of \(\AgdaDatatype{◇}\),
+starting with \(\AgdaDatatype{∈}\):
+\begin{equation}
+  x \in \mathit{xs} = \Diamond (\equiv x) \mathit{xs}
+\end{equation}
+\ExecuteMetaData[agda/Data/List/Membership.tex]{in-def}
+
+To show \emph{unique} membership, the membership proof itself must be a
+contraction.
+\begin{equation}
+  x \inunique \mathit{xs} = \text{isContr}(x \in \mathit{xs})
+\end{equation}
+\ExecuteMetaData[agda/Data/List/Membership.tex]{in-uniq}
+\section{Supplementary Code}
+\subsection{Permutations}
+Swap:
+\ExecuteMetaData[agda/Data/List/Relation/Binary/Permutation.tex]{swap}
+Cons:
+\ExecuteMetaData[agda/Data/List/Relation/Binary/Permutation.tex]{cons}
+Tail:
+\ExecuteMetaData[agda/Data/List/Relation/Binary/Permutation.tex]{tail}
+\subsection{Non-Dependent Sum (Disjoint Union)}
+\ExecuteMetaData[agda/Snippets/Introduction.tex]{sum-def}
+\subsection{Split Enumerability}
+Closure over \(\bot\):
+\ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{empty}
+Closure over \(\top\):
+\ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{unit}
+Closure over non-dependent sum:
+\ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{sum-prf}
+\subsection{Lists}
+\subsubsection{Indexing}
+\ExecuteMetaData[agda/Data/List.tex]{indexing}
+\subsubsection{Removing Duplicates}
+\ExecuteMetaData[agda/Data/List/Membership.tex]{uniques}
+\subsection{Functions}
+If there is an injection from \(A\) into a set with decidable equality, \(A\)
+has decidable equality.
+\ExecuteMetaData[agda/Function/Injective.tex]{discrete}
+If there is a surjection from \(A\) to \(B\), there is an injection from \(B\)
+to \(A\).
+\ExecuteMetaData[agda/Function/Surjective.tex]{surj-to-inj}
+\subsection{Membership Proofs}
+All iso:
+\ExecuteMetaData[agda/Data/List/Relation/Unary.tex]{all-iso}
+Any iso:
+\ExecuteMetaData[agda/Data/List/Relation/Unary.tex]{any-iso}
+A small function to make membership proofs shorter:
+\ExecuteMetaData[agda/Data/List/Membership.tex]{at}
+Pushing and pulling: going between \(x \; \AgdaDatatype{∈} \; xs\) and \(x
+\AgdaDatatype{∈} \; y \; \AgdaInductiveConstructor{∷} \; xs\).
+\ExecuteMetaData[agda/Data/List/Relation/Unary.tex]{push-pull}
+Push-pull for unique proofs of membership.
+\ExecuteMetaData[agda/Data/List/Relation/Unary.tex]{push-pull-unique}
+Proving that an element at the head of a list is in that list.
+\ExecuteMetaData[agda/Data/List/Relation/Unary.tex]{here-unique}
+\(\AgdaFunction{map}\) distributes over \(\AgdaDatatype{∈}\)
+\ExecuteMetaData[agda/Data/List/Relation/Unary.tex]{cong-any}
+\ExecuteMetaData[agda/Data/List/Membership.tex]{cong-in}
+An element is not in a list it's been removed from.
+\ExecuteMetaData[agda/Data/List/Membership.tex]{not-in-removed}
+\ExecuteMetaData[agda/Data/List/Membership.tex]{cons-remove}
+If \(x\) isn't in a list, it is not in that list with another element removed.
+\ExecuteMetaData[agda/Data/List/Membership.tex]{remove-other}
+Removing another element from a list doesn't affect \(x\)'s proof of unique
+membership.
+\ExecuteMetaData[agda/Data/List/Membership.tex]{unique-remove-other}
+If \(x\) is in \(xs\), \(x\) is uniquely in the unique elements of \(xs\).
+\ExecuteMetaData[agda/Data/List/Membership.tex]{unique-preserves}
+% \subsection{Convolution}
+% Convolution of lists.
+% \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{convolve}
+% Convolution of streams.
+% \ExecuteMetaData[agda/Cardinality/Infinite.tex]{convolve}
+% Convolution of colists.
+% \ExecuteMetaData[agda/Cardinality/PossiblyInfinite.tex]{convolve}
+% \subsection{Decidable Predicates}
+% Rather than the ``usual'' formulation of \(\AgdaDatatype{Dec}\), we use the
+% following definition:
+% \ExecuteMetaData[agda/Relation/Nullary.tex]{dec-def}
+% \subsection{Pauli Group Functions}
+% \ExecuteMetaData[agda/Data/Pauli.tex]{bad-eq}
+\subsection{Fin} \label{agda-fin}
+Injectivity of the \(\textbf{Fin}\) type constructor.
+\ExecuteMetaData[agda/Data/Fin/Properties.tex]{fin-inj}
+\section{References}
+\printbibliography[heading=none]
+\end{subappendices}
 \end{document}

### 2020-01-04 paper.tex
finally compiles

@@
 \author{Donnacha Oisín Kidney}
+\authorrunning{D. O. Kidney}
+\institute{University College Cork \and
+  \email{o.kidney@cs.ucc.ie}}
 \begin{document}
 \maketitle
-\listoftodos
-\tableofcontents
-\newcommand{\pifun}[2]{\prod_{\left( \mathit{ #1 } : #2 \right)}}
-\begin{quote}
-  \textbf{Abstract}
+\begin{abstract}
   We explore five notions of finiteness in Homotopy Type Theory \cite{hottbook}.
@@
   full functions given desired properties.
-\end{quote}
+\end{abstract}
 \section{Introduction}
@@
   Here we take the version preferred in the Cubical Agda library: contractible
-  maps \cite[Definition~4.4.1]{hottbook}. \todo[noline]{Need this in appendix also}
+  maps \cite{hottbook}. \todo[noline]{Need this in appendix also}
   Because of the parallels between contractible maps and split surjections,
@@
 \ExecuteMetaData[agda/Data/Pauli.tex]{prf-curried}
-
-\begin{subappendices}
-
-\section{Notation, Cubical Type Theory, and Agda}
-Our chosen setting is Cubical Type Theory (CuTT)
-\cite{cohenCubicalTypeTheory2016}, which is closely related to Homotopy Type
-Theory (HoTT) \cite{hottbook}.
-We will go through the basics and some notation here.
-\subsection{Martin-Löf Type Theory}
-CuTT has its roots in Martin-Löf intuitionistic type theory
-\cite{martin-lofIntuitionisticTypeTheory1980}.
-We will describe our notation for the MLTT parts of CuTT here (although our
-notation is quite standard).
-
-The empty or \(\mathbf{0}\) type will be denoted by \(\bot\), and the
-\(\mathb{1}\) type by \(\top\).
-The single inhabitant of \(\top\) is called \(\text{tt}\).
-
-``not'' is expressed as follows:
-\begin{equation}
-  \neg A = A \rightarrow \bot
-\end{equation}
-In other words, to say a type is impossible we say we can derive \(\bot\) from
-it.
-
-We also use the \(\mathbf{2}\) type, which we call \(\textbf{Bool}\).
-Its two canonical inhabitants are \(\text{true}\) and \(\text{false}\).
-
-We will use the term type family (or simply ``family'') to mean a non-dependent
-function with a codomain in Type.
-
-Dependent function types, or \(\prod\)-types, use the following notation:
-\begin{equation}
-  \prod_{(x : A)} , B (x)
-\end{equation}
-In Agda we use a single arrow to denote the same:
-\ExecuteMetaData[agda/Snippets/Introduction.tex]{pi}
-When the type of the left-hand-side can be inferred, we are permitted to use the
-following notation:
-\ExecuteMetaData[agda/Snippets/Introduction.tex]{forall}
-For non-dependent functions we use the single arrow (\(\rightarrow\)).
-
-Dependent pairs, or \(\sum\)-types, use a similar notation to \(\prod\)-types.
-\begin{equation}
-  \sum_{(x : A)} , B (x)
-\end{equation}
-In Agda:
-\ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma}
-Again, we get a shorthand when the type of the first element can be inferred:
-\ExecuteMetaData[agda/Snippets/Introduction.tex]{exists}
-For non-dependent pairs we use the \(\times\) symbol.
-We access the two fields of \(\sum\) types with \(\text{fst}\) and
-\(\text{snd}\):
-\begin{align}
-  \text{fst} &: \sum_{(x : A)} , B (x) \rightarrow A \\
-  \text{snd} &: \prod_{(x : \sum_{(y : A)} , B (y))} , B (\text{fst}(x))
-\end{align}
-
-Finally, the identity type is denoted by the triple equals (\(\equiv\)).
-The normal equals (\(=\)) is reserved for definitions.
-
-\subsection{Agda Basics}
-Agda \cite{norellDependentlyTypedProgramming2008} is a dependently-typed
-functional programming language, with roots in Haskell and ML.
-The core type theory is based on MLTT, though support for CuTT was added in the
-last year \cite{vezzosiCubicalAgdaDependently2019}.
-\subsubsection{Declarations}
-The anatomy of a simple Agda declaration is thus:
-\ExecuteMetaData[agda/Snippets/Introduction.tex]{abool}
-We first have a type signature: the name of the variable is to the left of the
-colon, and its type to the right.
-Below, we give an equation: as Agda is pure, it suffices to imagine that
-everywhere the name on the left of the equals sign is seen, we can replace it
-with the value on the right.
-\subsubsection{Functions}
-Functions are usually declared with pattern matching.
-\ExecuteMetaData[agda/Snippets/Introduction.tex]{not}
-Again, each clause is an equation.
-Agda's purity will ensure that these clauses hold true definitionally\footnotemark.
-\footnotetext{
-  Actually, it is possible to construct functions whose equations
-  hold propositionally, but not definitionally.
-  An example might be the following:
-  \ExecuteMetaData[agda/Snippets/Introduction.tex]{non-exact-split}
-  This is actually attempted by a lot of Agda beginners: often the compiler will
-  complain ``couldn't match type \(n\) with \(n + 0\)'', so the programmer will
-  simply add it as a clause.
-  Unfortunately, this actually over complicates things in the long run.
-  Neither the second nor third clause here will be definitionally equal: if you
-  have an obligation \(n \oplus 0 \qeq n\), the first pattern (\(0 \oplus m = m\))
-  has to be ruled out first, which means you'll have to do a manual pattern
-  match in your proof.
-  Similarly for the third pattern: to prove the obligation (\((1 + n) \oplus m
-  \qeq 1 + (n \oplus m)\)), \emph{both} preceding patterns have to be checked
-  against first.
-
-  This ordering requirement is introduced by overlapping patterns, and Agda
-  warns against it.
-  Knowing when to avoid it and when it doesn't matter is one of the aspects of
-  programming and proving in Agda which is more of an art form.
-}
-
-As Agda is a functional language, functions can also be defined with lambda
-expressions.
-\ExecuteMetaData[agda/Snippets/Introduction.tex]{id}
-
-Finally, ``mixfix'' operators play a large role in Agda's syntax.
-We can have binary operators:
-\ExecuteMetaData[agda/Snippets/Introduction.tex]{add}
-Unary operators (here, a prefix increment operator):
-\ExecuteMetaData[agda/Snippets/Introduction.tex]{unary-inc}
-Ternary operators:
-\ExecuteMetaData[agda/Snippets/Introduction.tex]{ifte}
-And so on.
-\subsubsection{Data Type Definitions}
-We can define new data types in Agda by listing their constructors:
-\ExecuteMetaData[agda/Snippets/Introduction.tex]{maybe}
-\subsubsection{Termination}
-\todo[inline]{Coinductive types, termination checking, etc.}
-\subsubsection{Universe Levels}
-All dependently-typed programming languages have to contend, at some point or
-another, with Girard's paradox
-\cite{girardInterpretationFonctionelleElimination1972}.
-The paradox concerns the type of types: it's the type-theory analogue of
-Russell's paradox.
-
-Unfortunately, most of the ways to avoid the paradox add significant clutter to
-the type theory, and increase the complexity of even simple declarations.
-Agda's answer is no exception.
-
-In Agda, every type is tagged with a ``universe level''.
-The basic types like \(\AgdaDatatype{Bool}\) live in universe level 0, the type
-of types lives in universe level 1, the type of types in universe level 1 lives
-in universe level 2, and so on.
-We won't go into it much more than that: usually there is an obvious way to
-handle the universe level of a given expression ``properly'', and cases where
-universe levels do something interesting come up rarely (although
-section~\ref{practical} does touch on one).
-We only mention it here to explain the expressions like \(A :
-\AgdaDatatype{Type} \; a\) which show up in many examples: \(a\) there is the
-universe level of the type \(A\).
-\subsection{Cubical Basics}
-This will not serve as an introduction CuTT,
-but rather a brief overview of the parts of the theory we will be using.
-For a fuller reference we direct the reader first to the ``HoTT Book''
-\cite{hottbook}, then the paper on Cubical Type Theory
-\cite{cohenCubicalTypeTheory2016}, and finally the paper on Cubical Agda
-\cite{vezzosiCubicalAgdaDependently2019}.
-\subsubsection{Paths}
-Central to CuTT is the \emph{path} type.
-This is a path in the topological sense, although we use it as a representation
-for equality (which makes sense: central to HoTT is the
-equivalence of these two concepts).
-In most of the code presented here this will not behave in any different way to
-the traditional identity type in Martin-Löf type theory: one exception to this
-is pattern-matching, which is not available on paths in Cubical Agda currently.
-As a result, we have to use its eliminator directly (if needed), although
-usually higher-level combinators like \(\AgdaFunction{subst}\) are sufficient.
-
-We will use \(\equiv\) as a notation for path types as well as the identity
-types.
-Although an identity type does exist in Cubical Type Theory, we will not use it,
-so we do not fear much confusion on this point.
-\subsubsection{Homotopy Levels}
-Next, we have the concept of ``homotopy levels''.
-These classify types in homotopy type theory according to their degree of
-truncation.
-Homotopy level 1, for instance, is the ``mere propositions'': the types for
-which every element is equal to every other.
-As a predicate, we can define \(\AgdaFunction{isProp}\) like so:
-\begin{equation}
-  \text{isProp}(A) = \prod_{(x : A)} , \prod_{(y : A)} , x \equiv y
-\end{equation}
-In Agda:
-\ExecuteMetaData[agda/Snippets/Introduction.tex]{isProp}
-
-One level up from the propositions are the (homotopy) sets.
-These are types for which all paths are themselves propositions.
-\begin{equation}
-  \text{isSet}(A) = \prod_{(x : A)} , \prod_{(y : A)} , \text{isProp}(x \equiv y)
-\end{equation}
-In Agda:
-\ExecuteMetaData[agda/Snippets/Introduction.tex]{isSet}
-
-Finally, one level down from the propositions are the contractions.
-These are types for which every element is equal to some specific element.
-\begin{equation}
-  \text{isContr}(A) = \sum_{(x : A)} , \prod_{(y : A)} , x \equiv y
-\end{equation}
-\ExecuteMetaData[agda/Snippets/Introduction.tex]{isContr}
-\subsubsection{Univalence}
-\todo[inline]{Flesh out}
-To describe this we first need fibers:
-\begin{equation}
-  \text{fiber}_f(y) = \sum_{(x : A)} , f (x) \equiv y
-\end{equation}
-\ExecuteMetaData[agda/Snippets/Introduction.tex]{fiber}
-From this we build the notion of an ``equivalence'': a function whose fibers are
-all contractions.
-\begin{equation}
-  A \simeq B = \sum_{(f : A \rightarrow B)}, \prod_{(y : B)} , \text{isContr}(\text{fiber}_f(y))
-\end{equation}
-\ExecuteMetaData[agda/Snippets/Introduction.tex]{equiv}
-The univalence axiom is that all equivalences are equivalent to the identity
-type.
-\begin{definition}[The Univalence Axiom]
-  \begin{equation}
-    \simeq \; \simeq \; \equiv
-  \end{equation}
-\end{definition}
-In practice this means that we can prove equivalence between two types and then
-transport code on one to code on the other.
-It also means that isomorphisms can be upgraded to fully-fledged equalities.
-\begin{definition}[Isomorphism]
-  An isomorphism between two types \(A\) and \(B\) is a pair of functions:
-  \begin{align}
-    f : A \rightarrow B \\
-    g : B \rightarrow A
-  \end{align}
-  Such that they are inverses of each other.
-  \begin{align}
-    \prod_{(x : A)} , g(f(x)) \equiv x \\
-    \prod_{(y : B)} , f(g(y)) \equiv y
-  \end{align}
-\end{definition}
-\ExecuteMetaData[agda/Snippets/Introduction.tex]{iso}
-\subsubsection{Higher Inductive Types}
-Along with univalence, the other significant benefit we get from path types is
-the ability to define new paths on types.
-A direct use of this feature is synthetic topology.
-We can, for instance, define the circle \(S^1\) as a higher-inductive type.
-\ExecuteMetaData[agda/Snippets/Introduction.tex]{circle}
-
-Another important use of higher inductive types is quotient types.
-These types are unavailable in ML type theory, but in Cubical Type Theory we can
-declare a quotient (here, the type of Kuratowski finite sets) like so:
-\ExecuteMetaData[agda/Algebra/Construct/Free/Semilattice.tex]{def}
-The first two constructors are standard data constructors, which mimic those for
-the list in Haskell.
-The next two add the equations we want our type to obey: we don't want it to
-care about order or duplicates.
-The last constructor \emph{truncates} the type, ensuring that it is a set.
-If we hadn't included it paths would not be mere propositions: there would be
-more than one path between any given elements of a type.
-The ability to have multiple paths is what allows Cubical Type Theory to do the
-more advanced things in topology etc., but we don't need its power here.
-
-One HIT in particular will be of great use to us: the HIT for propositional
-truncation, denoted with \(\lVert \cdot \rVert\).
-We will go through it in detail in section~\ref{prop-trunc}.
-\section{Supplementary Definitions}
-This section will serve as a reference for any definitions used in the text
-which are either not important enough to be examined inline, or common enough to
-justify elision.
-Many of the definitions come straight from the ``HoTT Book'' \cite{hottbook}.
-\subsection{Decidable Types}
-It is important to be able to talk about ``decisions'' for things like decidable
-equality, and so on.
-\begin{definition}[Decision]
-  A decision for some type \(P\) is represented by the type \(\mathbf{Dec}(A)\),
-  which has two constructors:
-  \begin{equation}
-    \begin{alignedat}{2}
-      \text{yes} &:      &A \rightarrow \mathbf{Dec}(A) \\
-      \text{no}  &: \neg &A \rightarrow \mathbf{Dec}(A)
-    \end{alignedat}
-  \end{equation}
-\end{definition}
-
-Internally, we actually use a rather more complicated type, which has preferable
-computational behaviour.
-The interface is the same, however.
-\subsection{Propositional Truncation} \label{prop-trunc}
-Essentially, a value of type \(\lVert A \rVert\) proof that an \(A\) exists,
-without revealing which \(A\) in particular.
-In HoTT, it allows us to work in a setting similar to classical logic.
-In Agda its definition is:
-\ExecuteMetaData[agda/Snippets/Introduction.tex]{prop-trunc}
-Any eliminators from this type must not distinguish between different values of
-\(A\).
-
-\begin{lemma}[Propositional Elimination]
-  We can eliminate from the type \(\lVert A \rVert\) to \(B\) with a function
-  \(A \rightarrow B\) if B is a proposition.
-\end{lemma}
-\begin{proof}
-  By case-analysis on \(\lVert A \rVert\).
-  In the first case, we apply our function.
-  In the second, we are required to show that that application agrees on any two
-  values of \(\lVert A \rVert\) \(\lVert A \rVert\).
-  This is true as \(B\) is a proposition.
-  In Agda:
-  \ExecuteMetaData[agda/HITs/PropositionalTruncation/Proofs.tex]{rec}
-\end{proof}
-
-Since \(\lVert A \rVert\) is itself a proposition, we can work ``under'' the
-truncation as much as we like.
-In fact the type \(\lVert \cdot \rVert\) itself forms a monad, allowing us to
-use monadic notation.
-\begin{lemma}
-  \(\lVert \cdot \rVert\) is a monad, with bind being the eliminator defined
-  above, and return being the \(\lvert \cdot \rvert\) constructor.
-\end{lemma}
-\begin{proof}
-  The monad laws are satisfied trivially, as \(\lVert \cdot \rVert\) is a
-  proposition.
-\end{proof}
-
-We can further eliminate into a type that isn't a proposition, as long as we
-prove that the function used in the eliminator doesn't discriminate between
-different \(A\)s in \(\lVert A \rVert\).
-\begin{lemma}[Propositional Elimination by Coherence] \label{elim-non-prop}
-  We can eliminate from a propositionally-truncated type into a non-proposition,
-  as long as the function is coherently constant
-  \cite{krausGeneralUniversalProperty2015}.
-\end{lemma}
-\begin{proof}
-  \todo[inline]{sketch proof}
-\end{proof}
-
-From these eliminators, we can derive some useful lemmas.
-\begin{lemma}[Propositional Refutation]
-  While a propositionally truncated proof is not enough to prove the
-  non-truncated type, it is enough to \emph{refute} the non-truncated negation
-  of the type.
-  \begin{equation}
-    \lVert A \rVert \rightarrow \neg (\neg A)
-  \end{equation}
-\end{lemma}
-\begin{proof}
-  We will use propositional elimination.
-  First, to go from \(A\) to \((A \rightarrow \bot) \rightarrow \bot\) is
-  trivial.
-  Secondly, to show that \((A \rightarrow \bot) \rightarrow \bot\) is a
-  proposition, we can use the fact that \(\bot\) is a proposition, and then
-  derive (via function extensionality) that any \(A \rightarrow \bot\) is also a
-  proposition.
-\end{proof}
-
-\begin{lemma}[Recomputation]
-  If a type \(A\) is decidable, and there exists a proof of the type \(\lVert A
-  \rVert\), we can derive a proof of \(A\).
-  In other words, the following function exists:
-  \begin{equation}
-    \text{recompute} : \textbf{Dec}(A) \times \lVert A \rVert \rightarrow A
-  \end{equation}
-\end{lemma}
-\begin{proof}
-  By case analysis of the decision over \(A\), we know that the false case must
-  be untrue (since we can refute it using propositional refutation), and we can
-  therefore extract the true case.
-\end{proof}
-
-Occasionally we will call families ``predicates'' when we mean to treat the
-family as a property on the domain.
-Formally this might imply that the codomain is a proposition (i.e. isomorphic to
-either \(\top\) or \(\bot\)), but it is not always necessary to prove this fact,
-and indeed occasionally we will want to use some of the information stored in
-the predicate, making it not propositional.
-For instance, a natural number being ``not prime'' we would refer to as a
-predicate, but the proof of this fact may be a list of the number's prime
-factors, so we will not truncate it away.
-
-When we say that two predicates \(P\) and \(Q\) are ``equal strength'', we mean
-there are a pair of functions \(P \rightarrow Q\) and \(Q \rightarrow P\).
-This is distinct from \(P\) and \(Q\) being isomorphic, as we haven't shown
-anything regarding the relationship between those two functions.
-However, if \(P\) and \(Q\) are themselves propositions, this does indeed imply
-an isomorphism.
-\begin{lemma} \label{equal-strength}
-  For any two types \(A\) and \(B\), if there are functions \(f : A \rightarrow
-  B\) and \(g : B \rightarrow A\), \(\lVert A \rVert\) and \(\lVert B \rVert\)
-  are equivalent.
-\end{lemma}
-\begin{proof}
-  We prove equivalence by way of isomorphism.
-
-  Let \(A\), \(B\), \(f\), and \(g\) be defined as above.
-  For isomorphism, we first provide a pair of functions \(\lVert f \rVert\) and
-  \(\lVert g \rVert\), which are defined by lifting \(f\) and \(g\) under
-  propositional truncation, respectively.
-
-  Next, we need to prove that \(g\) is a section of \(f\).
-  \begin{equation}
-    \text{section}(f, g) = f \circ g \equiv id
-  \end{equation}
-  Since \(\lVert B \rVert\) is a proposition, this is true by function
-  extensionality.
-
-  Finally, \(f\) is a retraction of \(g\).
-  \begin{equation}
-    \text{retraction}(f, g) = g \circ f \equiv id
-  \end{equation}
-  Again, since \(\lVert A \rVert\) is a proposition, this is true by function
-  extensionality.
-
-  In Agda:
-  \ExecuteMetaData[agda/HLevels/Proposition.tex]{prop-iff}
-\end{proof}
-\subsection{Functions}
-\subsubsection{Surjections}
-The following is the definition of surjectivity in HoTT:
-\begin{equation}
-  \text{Surjective}(f) = \prod_{(y : B)} , \lVert \text{fiber}_f(y) \rVert
-\end{equation}
-\ExecuteMetaData[agda/Function/Surjective.tex]{surj-def}
-We are also interested in \emph{split} surjectivity, where we don't truncate the
-fiber:
-
-\begin{equation}
-  \text{SplitSurjective}(f) = \prod_{(y : B)} , \text{fiber}_f(y)
-\end{equation}
-\ExecuteMetaData[agda/Function/Surjective.tex]{split-surj-def}
-
-A surject\emph{ion}, then, is a function which is surject\emph{ive}:
-\begin{equation}
-  A \twoheadrightarrow B = \sum_{(f : A \rightarrow B)} \text{SplitSurjective}(f)
-\end{equation}
-\ExecuteMetaData[agda/Function/Surjective.tex]{surjection}
-\subsubsection{Injections}
-The standard definition of an injective function:
-\ExecuteMetaData[agda/Function/Injective.tex]{injective-def}
-An injection:
-\ExecuteMetaData[agda/Function/Injective.tex]{injection-def}
-\subsection{Fin}
-\(\textbf{Fin}\) is a finite prefix of the natural numbers.
-Semi-formally:
-\begin{equation}
-  \textbf{Fin}(n) = \left[ 0 \dots n \right)
-\end{equation}
-There are actually a number of ways to define \(\textbf{Fin}\) in type theory.
-The method we use here is as follows:
-\ExecuteMetaData[agda/Data/Fin.tex]{def}
-\subsection{Split Enumerability}
-Here is how we define it in Agda:
-\ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{definition}
-\subsection{Predicates Over Lists}
-These two well-known dual types for quantifying a predicate over a list are used
-extensively in this work.
-We will define them using the indexing operator \(!\), which has the following
-type:
-\begin{equation}
-  ! : \prod_{(\mathit{xs} : \textbf{List}(A))} , \textbf{Fin}(\text{length}(\mathit{xs})) \rightarrow A
-\end{equation}
-For some predicate \(P\), the two definitions are as follows:
-\begin{definition}[Any]
-  \begin{equation}
-    \Diamond P \mathit{xs} = \sum_{(i : \textbf{Fin}(\text{length}(\mathit{xs})))} , P(xs \ind i)
-  \end{equation}
-  \ExecuteMetaData[agda/Data/List/Relation/Unary.tex]{any-def}
-\end{definition}
-\begin{definition}[All]
-  \begin{equation}
-    \Box P \mathit{xs} = \prod_{(i : \textbf{Fin}(\text{length}(\mathit{xs})))} , P(xs \ind i)
-  \end{equation}
-  \ExecuteMetaData[agda/Data/List/Relation/Unary.tex]{all-def}
-\end{definition}
-We are using the tradition modal logic operators here: \(\Diamond\)
-means ``any'', and \(\Box\) means ``all''.
-\subsection{Membership}
-Several other constructions are defined in terms of \(\AgdaDatatype{◇}\),
-starting with \(\AgdaDatatype{∈}\):
-\begin{equation}
-  x \in \mathit{xs} = \Diamond (\equiv x) \mathit{xs}
-\end{equation}
-\ExecuteMetaData[agda/Data/List/Membership.tex]{in-def}
-
-To show \emph{unique} membership, the membership proof itself must be a
-contraction.
-\begin{equation}
-  x \inunique \mathit{xs} = \text{isContr}(x \in \mathit{xs})
-\end{equation}
-\ExecuteMetaData[agda/Data/List/Membership.tex]{in-uniq}
-\section{Supplementary Code}
-\subsection{Permutations}
-Swap:
-\ExecuteMetaData[agda/Data/List/Relation/Binary/Permutation.tex]{swap}
-Cons:
-\ExecuteMetaData[agda/Data/List/Relation/Binary/Permutation.tex]{cons}
-Tail:
-\ExecuteMetaData[agda/Data/List/Relation/Binary/Permutation.tex]{tail}
-\subsection{Non-Dependent Sum (Disjoint Union)}
-\ExecuteMetaData[agda/Snippets/Introduction.tex]{sum-def}
-\subsection{Split Enumerability}
-Closure over \(\bot\):
-\ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{empty}
-Closure over \(\top\):
-\ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{unit}
-Closure over non-dependent sum:
-\ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{sum-prf}
-\subsection{Lists}
-\subsubsection{Indexing}
-\ExecuteMetaData[agda/Data/List.tex]{indexing}
-\subsubsection{Removing Duplicates}
-\ExecuteMetaData[agda/Data/List/Membership.tex]{uniques}
-\subsection{Functions}
-If there is an injection from \(A\) into a set with decidable equality, \(A\)
-has decidable equality.
-\ExecuteMetaData[agda/Function/Injective.tex]{discrete}
-If there is a surjection from \(A\) to \(B\), there is an injection from \(B\)
-to \(A\).
-\ExecuteMetaData[agda/Function/Surjective.tex]{surj-to-inj}
-\subsection{Membership Proofs}
-All iso:
-\ExecuteMetaData[agda/Data/List/Relation/Unary.tex]{all-iso}
-Any iso:
-\ExecuteMetaData[agda/Data/List/Relation/Unary.tex]{any-iso}
-A small function to make membership proofs shorter:
-\ExecuteMetaData[agda/Data/List/Membership.tex]{at}
-Pushing and pulling: going between \(x \; \AgdaDatatype{∈} \; xs\) and \(x
-\AgdaDatatype{∈} \; y \; \AgdaInductiveConstructor{∷} \; xs\).
-\ExecuteMetaData[agda/Data/List/Relation/Unary.tex]{push-pull}
-Push-pull for unique proofs of membership.
-\ExecuteMetaData[agda/Data/List/Relation/Unary.tex]{push-pull-unique}
-Proving that an element at the head of a list is in that list.
-\ExecuteMetaData[agda/Data/List/Relation/Unary.tex]{here-unique}
-\(\AgdaFunction{map}\) distributes over \(\AgdaDatatype{∈}\)
-\ExecuteMetaData[agda/Data/List/Relation/Unary.tex]{cong-any}
-\ExecuteMetaData[agda/Data/List/Membership.tex]{cong-in}
-An element is not in a list it's been removed from.
-\ExecuteMetaData[agda/Data/List/Membership.tex]{not-in-removed}
-\ExecuteMetaData[agda/Data/List/Membership.tex]{cons-remove}
-If \(x\) isn't in a list, it is not in that list with another element removed.
-\ExecuteMetaData[agda/Data/List/Membership.tex]{remove-other}
-Removing another element from a list doesn't affect \(x\)'s proof of unique
-membership.
-\ExecuteMetaData[agda/Data/List/Membership.tex]{unique-remove-other}
-If \(x\) is in \(xs\), \(x\) is uniquely in the unique elements of \(xs\).
-\ExecuteMetaData[agda/Data/List/Membership.tex]{unique-preserves}
-% \subsection{Convolution}
-% Convolution of lists.
-% \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{convolve}
-% Convolution of streams.
-% \ExecuteMetaData[agda/Cardinality/Infinite.tex]{convolve}
-% Convolution of colists.
-% \ExecuteMetaData[agda/Cardinality/PossiblyInfinite.tex]{convolve}
-% \subsection{Decidable Predicates}
-% Rather than the ``usual'' formulation of \(\AgdaDatatype{Dec}\), we use the
-% following definition:
-% \ExecuteMetaData[agda/Relation/Nullary.tex]{dec-def}
-% \subsection{Pauli Group Functions}
-% \ExecuteMetaData[agda/Data/Pauli.tex]{bad-eq}
-\subsection{Fin} \label{agda-fin}
-Injectivity of the \(\textbf{Fin}\) type constructor.
-\ExecuteMetaData[agda/Data/Fin/Properties.tex]{fin-inj}
-\section{References}
-\printbibliography[heading=none]
-\end{subappendices}
+\bibliographystyle{splncs04}
+\bibliography{bibliography.bib}
 \end{document}

### 2020-01-04 agda-from-toplevel.sh

@@ -1,0 +1,5 @@
+#!/bin/bash
+withcode=$1
+nocode=${withcode#"agda/"}
+cd agda || exit
+agda --latex --latex-dir=. "$nocode"

### 2020-01-04 check-all.sh

@@ -1,0 +1,11 @@
+#!/bin/bash
+( cd agda
+echo '{-# OPTIONS --cubical #-}
+' > Everything
+echo 'module Everything where
+' >> Everything
+find -E . -type f -regex '.*\.l?agda' | cut -c 3- | cut -f1 -d'.' | sed 's/\//\./g' | sed 's/^/import /' >> Everything
+mv Everything Everything.agda
+agda Everything.agda
+rm Everything.agda
+)

### 2020-01-04 haskell-from-toplevel.sh

@@ -1,0 +1,7 @@
+#!/bin/bash
+withcode=$1
+nocode=${withcode#"haskell/"}
+cd haskell || exit
+dir=$(dirname "$nocode")
+file=$(basename "$nocode" .lhs).tex
+lhs2TeX -o "$dir/$file" "$nocode"

### 2020-01-04 init-lhs.sh

@@ -1,0 +1,18 @@
+#!/bin/bash
+
+if [ ! -f "locallhs2TeX.lhs" ]; then
+    echo "%include polycode.fmt" > "locallhs2TeX.lhs"
+fi
+
+if [ ! -f "locallhs2TeX.sty" ]; then
+    lhs2TeX -o "locallhs2TeX.sty" "locallhs2TeX.lhs"
+fi
+
+find haskell -type f -name '*.lhs' | while read -r code ; do
+    dir=$(dirname "$code")
+    file="$dir"/$(basename "$code" .lhs).tex
+    if [ ! -e "$file" ]
+    then
+        ./haskell-from-toplevel.sh "$code"
+    fi
+done

### 2020-01-04 init-missing-lagda.sh

@@ -1,0 +1,10 @@
+#!/bin/bash
+cd agda || exit
+find . -type f -name '*.lagda' | while read -r code ; do
+    dir=$(dirname "$code")
+    file=$(basename "$code" .lagda).tex
+    if [ ! -e "$dir/$file" ]
+    then
+        agda --latex --latex-dir=. "$code"
+    fi
+done

### 2020-01-04 figures/finite-classification.tex
lots

--- 2020-01-04 paper.tex
@@
-\input{preamble}
-\title{Semirings and Semimodules: A Substrate for Computation}
-\author{Donnacha Oisín Kidney}
-\begin{document}
-\maketitle
-\listoftodos
-\tableofcontents
-\include{sections/maths}
-\include{sections/type-theory}
-\include{sections/implementation}
-\include{sections/free-semimodule}
-\include{sections/univalence-for-privacy}
-\include{sections/combinatorics}
-\include{sections/ordered-monoids-and-search}
-\include{sections/semirings-on-the-gpu}
-\end{document}
+\begin{figure}[h]
+  \centering
+  \begin{tikzcd}[cramped, row sep=small]
+    &
+    \ar[ddddd, dash, start anchor={[yshift=4ex]}, end anchor={[yshift=-4ex]}] &
+    \text{Ordered} &
+    \ar[ddddd, dash, dashed, start anchor={[yshift=4ex]}, end anchor={[yshift=-4ex]}] &
+    \text{Unordered}
+    \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \ar[rrrr, dash, start anchor={[xshift=-6ex]}, end anchor={[xshift=10ex]}] &
+    &
+    &
+    &
+    {}
+    \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    &
+    &
+    \text{Split Enumerable}
+      \; \ref{split-enumerable}
+      \ar[d, xshift=-1ex, bend right=30]
+      &
+    &
+    \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \text{Discrete} &
+    &
+    \text{Manifest Bishop}
+      \ar[u, xshift=1ex, bend right=30]
+      \ar[rr, end anchor=south west, start anchor=south east, bend right=10, crossing over]
+      \; \ref{manifest-bishop}
+      &
+    &
+    \text{Cardinal}
+      \; \ref{cardinal}
+      \ar[ll, start anchor=north west, end anchor=north east, bend right=10, "\text{linOrd}" description]
+    \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \ar[rrrr, dash, dashed, start anchor={[xshift=-6ex]}, end anchor={[xshift=10ex]}] &
+    &
+    &
+    &
+    {}
+    \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \text{Non Discrete} &
+    {} &
+    \text{Manifest Enumerable}
+      \; \ref{manifest-enumerability}
+      \ar[from=uu, xshift=-1ex, bend right=30, crossing over]
+      \ar[uu, bend right=30,  "\text{Discrete}" description]
+      \ar[rr, start anchor=south east, end anchor=south west, bend right=10, crossing over]
+      &
+    {}
+    &
+    \text{Kuratowski}
+      \; \ref{kuratowski}
+      \ar[ll, start anchor=north west, end anchor=north east, bend right=10,  "\text{linOrd}" description]
+      \ar[uu, xshift=1ex, bend right=30,  "\text{Discrete}" description]
+      \ar[from=uu, xshift=-1ex, bend right=30, crossing over]
+  \end{tikzcd}
+  \caption{Classification of the strong definitions of finiteness, according to
+    whether they are discrete (imply decidable equality) and whether they induce
+    a linear order.}
+  \label{finite-classification}
+\end{figure}

### 2020-01-04 paper.tex
lots

@@
 typed programming languages.
-
-In section~\ref{strong-finiteness}, we will explore the ``strong'' notions of
-finiteness (i.e. those at least as strong as Kuratowski finiteness
-\cite{kuratowskiNotionEnsembleFini1920}), with a special focus on cardinal
-finiteness.
-We will delineate some subtly different finiteness predicates, and prove the
-relationship between each.
-We will prove, then, that every discrete Kuratowski finite type is cardinal
-finite, and subsequently will show that discrete Kuratowski finite sets form a
-topos.
-
+\subsubsection{Strong Finiteness Predicates}
+We will first explore the ``strong'' notions of finiteness (i.e. those at least
+as strong as Kuratowski finiteness \cite{kuratowskiNotionEnsembleFini1920}),
+with a special focus on cardinal finiteness (section~\ref{cardinal}), and
+manifest enumerability (section~\ref{manifest-enumerability}), which is new, to
+our knowledge.
+
+\input{figures/finite-classification}
+Figure~\ref{finite-classification} organises the predicates according to
+their ``strength''; i.e. how much information they provide about a conforming
+type.
+For instance, a proof that some type \(A\) is manifestly Bishop finite (the
+strongest of the notions, explored in section~\ref{manifest-bishop}) also tells
+us that \(A\) is discrete (has decidable equality), and gives us a linear order
+on the type.
+A type that is Kuratowski finite (section~\ref{kuratowski}) has no such extra
+features: indeed, we will see examples of Kuratowski finite types which are not
+even sets, never mind discrete ones.
+
+We will go through each of the predicates, proving how to weaken each (i.e. we
+will provide a proof that every cardinally finite type is Kuratowski finite),
+and how to strengthen them, given the required property.
+In terms of figure~\ref{finite-classification}, this amounts to providing proofs
+for each arrow.
+
+We will---through the use of containers
+\cite{abbottContainersConstructingStrictly2005}---formally prove the equivalence
+these predicates have with the usual function relations i.e. we will show that a
+proof of manifest enumerability is precisely equivalent to a surjection from a
+finite prefix of the natural numbers.
+
+For each predicate, we will also prove its closure over sums and products in
+both dependent and non-dependent forms, if such a closure exists.
+This will culminate in our main result for this section: the formal proof that
+decidable Kuratowski finite sets form a topos. \todo[noline]{reference for this}
+More specifically, we show that cardinal finite sets form a topos, that
+decidable Kuratowski finite sets are equivalent in strength to cardinal finite
+sets, and carry the proof over.
+This result relies on proofs on each of the other finiteness predicates.
+\subsubsection{Species}
+\todo[inline]{Redo this next paragraph}
 In section~\ref{species-and-containers}, we will redefine our finiteness
@@
 
+\subsubsection{Infinite Types}
 In section~\ref{infinite-cardinalities}, we will extend our study of finite
@@
 
+\subsubsection{Practical Uses of Finiteness}
 Proofs of finiteness have well-known practical applications in
@@
 dependent functions.
-\section{Strong Finiteness} \label{strong-finiteness}
-\begin{figure}[h]
-  \centering
-  \begin{tikzcd}[cramped, row sep=small]
-    &
-    \ar[ddddd, dash, start anchor={[yshift=4ex]}, end anchor={[yshift=-4ex]}] &
-    \text{Ordered} &
-    \ar[ddddd, dash, dashed, start anchor={[yshift=4ex]}, end anchor={[yshift=-4ex]}] &
-    \text{Unordered}
-    \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \ar[rrrr, dash, start anchor={[xshift=-6ex]}, end anchor={[xshift=10ex]}] &
-    &
-    &
-    &
-    {}
-    \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    &
-    &
-    \text{Split Enumerable}
-      \; \ref{split-enumerable}
-      \ar[d, xshift=-1ex, bend right=30]
-      &
-    &
-    \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \text{Discrete} &
-    &
-    \text{Manifest Bishop}
-      \ar[u, xshift=1ex, bend right=30]
-      \ar[rr, end anchor=south west, start anchor=south east, bend right=10, crossing over]
-      \; \ref{manifest-bishop}
-      &
-    &
-    \text{Cardinal}
-      \; \ref{cardinal}
-      \ar[ll, start anchor=north west, end anchor=north east, bend right=10, "\text{linOrd}" description]
-    \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \ar[rrrr, dash, dashed, start anchor={[xshift=-6ex]}, end anchor={[xshift=10ex]}] &
-    &
-    &
-    &
-    {}
-    \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \text{Non Discrete} &
-    {} &
-    \text{Manifest Enumerable}
-      \; \ref{manifest-enumerability}
-      \ar[from=uu, xshift=-1ex, bend right=30, crossing over]
-      \ar[uu, bend right=30,  "\text{Discrete}" description]
-      \ar[rr, start anchor=south east, end anchor=south west, bend right=10, crossing over]
-      &
-    {}
-    &
-    \text{Kuratowski}
-      \; \ref{kuratowski}
-      \ar[ll, start anchor=north west, end anchor=north east, bend right=10,  "\text{linOrd}" description]
-      \ar[uu, xshift=1ex, bend right=30,  "\text{Discrete}" description]
-      \ar[from=uu, xshift=-1ex, bend right=30, crossing over]
-  \end{tikzcd}
-  \caption{Classification of the strong definitions of finiteness, according to
-    whether they are discrete (imply decidable equality) and whether they induce
-    a linear order.}
-  \label{finite-classification}
-\end{figure}
-
-In this section, we will explore five predicates for finiteness, classified in
-figure~\ref{finite-classification}.
-Of these, manifest enumerability is new, to our knowledge.
-
-Figure~\ref{finite-classification} organises the predicates according to
-their ``strength''; i.e. how much information they provide about a conforming
-type.
-For instance, a proof that some type \(A\) is manifestly Bishop finite (the
-strongest of the notions, explored in section~\ref{manifest-bishop}) also tells
-us that \(A\) is discrete (has decidable equality), and gives us a linear order
-on the type.
-A type that is Kuratowski finite (section~\ref{kuratowski}) has no such extra
-features: indeed, we will see examples of Kuratowski finite types which are not
-even sets, never mind discrete ones.
-
-In the following sections, we will go through each of the predicates, proving
-how to weaken each (i.e. we will provide a proof that every cardinally finite
-type is Kuratowski finite), and how to strengthen them, given the required
-property.
-In terms of figure~\ref{finite-classification}, this amounts to providing proofs
-for each arrow.
-
-For each predicate, we will also prove its closure over sums and products in
-both dependent and non-dependent forms, if such a closure exists.
-This will culminate in our main result for this section: the formal proof that
-decidable Kuratowski finite sets form a topos. \todo[noline]{reference for this}
-More specifically, we show that cardinal finite sets form a topos, that
-decidable Kuratowski finite sets are equivalent in strength to cardinal finite
-sets, and carry the proof over.
-This result relies on proofs on each of the other finiteness predicates.
-\subsection{Split Enumerability} \label{split-enumerable}
+\section{Split Enumerability} \label{split-enumerable}
 We will start with the simplest definition of finiteness: we say a set is
@@
   \begin{equation}
-    \mathcal{E}!(A) = \sum_{(\mathit{xs} : \textbf{List} A)} , \prod_{(x : A)} , x \in xs
+    \mathcal{E}!(A) = \sum_{(\mathit{xs} : \textbf{List}(A))} , \prod_{(x : A)} , x \in xs
   \end{equation}
@@
   \end{equation}
-  Here, \(\mathbf{Fin}(n)\) is the type of natural numbers smaller than \(n\),
-  and \(\ind\) is an indexing function.
+  Here, \(\ind\) is a safe indexing function.
   In other words, to say that some item \(x\) is in some list \(\mathit{xs}\),
@@
   \(\mathit{xs}\) which is equivalent to \(x\).
+\end{definition}
+\begin{definition}[\mathbf{Fin}]
+  \(\mathbf{Fin}\)(n) is the type of natural numbers smaller than \(n\).
+  It is defined inductively, as follows:
+  \begin{equation}
+    \begin{aligned}
+      \mathbf{Fin}(0) &= \bot \\
+      \mathbf{Fin}(n + 1) &= 1 + \mathbf{Fin}(n)
+    \end{aligned}
+  \end{equation}
 \end{definition}
@@
 giving us a variant of the predicate which allows for non-set finite types.
-\subsubsection{Decidable Equality}
-We will now prove that split enumerability implies decidable equality.
-We first need to prove a lemma regarding surjections to split enumerable types.
-\begin{lemma} \label{split-enum-surj}
-  Any split enumerable type has a split surjection from a finite prefix of the
-  natural numbers.
-\end{lemma}
-\begin{proof} 
-  Let \(E\) be a proof of type \(\mathcal{E}!(A)\) for some type \(A\), and
-  \(n\) be the length of the support list of \(E\).
-  Then let \(f : \textbf{Fin}(n) \rightarrow A\) be defined as the indexing
-  function into the support list of \(E\).
-
-  We then prove this function to be a split surjection by providing a fiber over
-  all of the elements of its domain: this fiber is the second component of \(E\).
-
-  In Agda:
-  \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{to-surj}
-\end{proof}
-
-Because of careful arrangement of previous definitions, this proof is only three
-lines of Agda.
-It is not difficult to derive decidable equality from this fact.
+\subsection{Proving Equivalence with Split Surjections}
+Another, equivalent way to define ``finiteness'' is via a (split) surjection
+from a finite prefix of the natural numbers.
+Here, we will formalise this equivalence, by first proving an equivalence with a
+container-based definition of split enumerability.
+
+Containers \cite{abbottContainersConstructingStrictly2005} give us a general way
+to construct functors from generic building-blocks.
+\begin{definition}[Container]
+  A container is a dependent pair of shapes, and positions indexed by those
+  shapes.
+  \begin{equation}
+    \mathbf{Container} = \sum_{(\text{shape} : \text{Type})} , \text{shape} \rightarrow \text{Type}
+  \end{equation}
+  We ``interpret'' a container into a pair of a specific shape, and a function
+  which goes from positions to contents.
+  \begin{equation}
+    \llbracket S , P \rrbracket = \prod_{(X : \text{Type})} , \sum_{(s : S)} , \left( P(s) \rightarrow X \right)
+  \end{equation}
+\end{definition}
+
+For instance, the shape of lists is given by \(\mathbb{N}\), and the positions are then
+\(\mathbf{Fin}\)s indexed by that \(\mathbb{N}\).
+If we swap out wherever we used ``list'' in the definition of split
+enumerability with this container-based list, we get an equivalent finiteness
+predicate.
+\begin{lemma} \label{container-split-enum-equiv}
+  The container-based definition of split enumerability is equivalent to the
+  inductive type version.
+  \begin{equation}
+    \sum_{(\mathit{xs} : \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket A)} , \prod_{(x : A)} , x \in xs \simeq
+    \sum_{(\mathit{xs} : \textbf{List}(A))} , \prod_{(x : A)} , x \in xs
+  \end{equation}
+  Membership for container-defined lists is a fiber.
+  \begin{equation}
+    x \in \mathit{xs} = \text{fiber}(\text{snd}(\mathit{xs}), x)
+  \end{equation}
+\end{lemma}
+\begin{proof}
+  \todo[inline]{Proof}
+\end{proof}
+
+This version of split enumerability is equivalent to a split surjection from a
+finite prefix of the natural numbers. 
+\begin{lemma} \label{container-split-enum-surj}
+  Container-defined split enumerability is equivalent to a split surjection from
+  some \(\mathbf{Fin}\).
+  \begin{equation}
+    \mathcal{E}!(A) \simeq \sum_{(n : \mathbb{N})} , \mathbf{Fin}(n) \twoheadrightarrow ! A
+  \end{equation}
+\end{lemma}
+\begin{proof}
+  The proof is surprisingly short: after sufficient inlining, it emerges that
+  our goal is simply a reassociation.
+  \ExecuteMetaData[agda/Cardinality/Container/Finite.tex]{split-surj}
+\end{proof}
+This lets us prove our overall goal for this section.
+\begin{theorem} \label{split-enum-surj}
+  Split enumerability is equivalent to a split surjection from a finite prefix
+  of the natural numbers.
+\end{theorem}
+\begin{proof}
+  This follows from lemmas \ref{container-split-enum-equiv} and
+  \ref{container-split-enum-surj}.
+\end{proof}
+\subsection{Decidable Equality}
 \begin{lemma} \label{split-enum-discrete}
@@
   \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{discrete}
+  \todo[inline]{Rework this proof to rely on the split surjection equivalence
+    from before}
 \end{proof}
@@
 \end{proof}
-
-Though it's not strictly needed for the rest of the construction, it is
-instructive to also prove inverse regarding surjections; i.e. that we can
-derive \(\mathcal{E}!\) from a surjection from another finite type.
-\begin{lemma} \label{split-enum-from-surj}
-  Any type with a split surjection from another split enumerable type is itself
-  split enumerable.
-\end{lemma}
-\begin{proof}
-  Let \(E\) be a proof of split enumerability for some set \(A\), and let \(f\)
-  be a surjection from \(A\) to \(B\).
-  The support list for the proof of split enumerability over \(B\) can be
-  obtained by mapping the surjective function over the support list of \(E\).
-
-  To obtain the cover proof, let \(f^{-1}\) be the inverse function of \(f\)
-  (which we can obtain because it is a split-surjection).
-  Using the cover proof of \(E\), we can show that, for all \(y : B\), \(f^{-1}(y)\)
-  is in the support list for \(E\).
-  Using the following lemma (proven in the appendix):
-  \begin{equation}
-    x \in xs \rightarrow f (x) \in \text{map} \; f \; \mathit{xs}
-  \end{equation}
-  We obtain a proof that:
-  \begin{equation}
-    \forall (y : B). f (f^{-1}(y)) \in \text{map} \; f \; \mathit{xs}
-  \end{equation}
-  And, since \(f (f^{-1}(y)) \equiv y\) (by the definition of a surjection), we
-  have proven our goal.
-
-  In Agda:
-  \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{from-surj}
-\end{proof}
-\subsubsection{Closure}
+\subsection{Closure}
 In this section we will prove closure under various operations for split
@@
 \end{proof}
-The final case can be derived from the previous closures; we give it here anyway
-for clarity.
-\begin{lemma} \label{fin-split-enum}
-  Any finite prefix of the natural numbers (what we have been calling
-  \(\textbf{Fin}\)) is split enumerable.
-\end{lemma}
-\begin{proof}
-  For the support list, we will make use of the function tabulate, which returns
-  a list consisting of the application of \(f\) to every \(\textbf{Fin}(n)\),
-  given \(n\).
-  \begin{equation}
-    \text{tabulate}(n , f) = \left[ f(0) , f(1) \dots f(n - 1) \right]
-  \end{equation}
-  In Agda:
-  \ExecuteMetaData[agda/Data/List.tex]{tabulate}
-  Tabulation with the identity function gives us our cover list.
-
-  The cover proof is obviously true given the definition of tabulate above; in
-  Agda it requires a little more work to show.
-  \ExecuteMetaData[agda/Data/List/Membership.tex]{fin-in-tabulate}
-\end{proof}
-\subsection{Manifest Bishop Finiteness} \label{manifest-bishop}
+\section{Manifest Bishop Finiteness} \label{manifest-bishop}
 \begin{definition}[Manifest Bishop Finiteness]
@@
 support list is the same as the cardinality of the set.
-\subsubsection{Equivalence}
+\subsection{Equivalence}
 While manifest Bishop finiteness may seem stricter than split enumerability, we
@@
 \begin{lemma} \label{bishop-equiv}
-  Any manifestly Bishop finite type has an equivalence with a finite prefix of
-  the natural numbers.
+  A proof of manifest Bishop finiteness is equivalent to an equivalence with a
+  finite prefix of the natural numbers.
+  \begin{equation}
+    \mathcal{B}(A) \simeq \sum_{(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \simeq A \right)
+  \end{equation}
 \end{lemma}
@@
   Here we take the version preferred in the Cubical Agda library: contractible
-  maps \cite{hottbook}. \todo[noline]{Need this in appendix also}
+  maps \cite{hottbook}.
   Because of the parallels between contractible maps and split surjections,
   the proof proceeds much the same as \ref{split-enum-surj}.
-
-  First, we define the map as the indexing function into the support list.
-  Proving this map is a contraction is precisely the cover proof.
-
-  \ExecuteMetaData[agda/Cardinality/Finite/ManifestBishop.tex]{equiv}
-\end{proof}
-
-As we can see, our list-based definition of manifest Bishop finiteness is really
-an equivalence with \(\mathbf{Fin}\) in disguise.
-This will be the key component in establishing closure under exponentials.
-
-We would like to draw attention to the ``power-to-weight ratio'' of this proof.
-We have mentioned how important the equivalence will be for our later proofs: in
-general, the ability to transport via univalence opens up many opportunities
-
-In Cubical Type Theory, equivalences are much more useful than their counterpart
-(isomorphisms) in Martin-Löf  Type Theory: arbitrary proofs and functions can be
-\emph{transported} along equivalences, giving a constructive justification for
-an intuitive and useful technique which is commonplace classically.
-The proof itself, however, is three lines long: thanks to careful definitions,
-the fact that manifest Bishop finiteness is the ``same'' as equivalences to
-\(\mathbf{Fin}\) becomes obviously true.
-Later we will see, by defining lists as containers, a proof that manifest Bishop
-finiteness is equivalent (in the formal sense) to equivalences with
-\(\mathbf{Fin}\). \todo[noline]{This needs to be added to the containers
-  section.}
-\todo[noline]{More on this? }
-\subsubsection{Relationship to Split Enumerability}
+  \todo[inline]{Finish prose for this proof}
+\end{proof}
+\subsection{Relationship to Split Enumerability}
 We now show that manifest Bishop finiteness has equal strength to split
@@
 \(\mathcal{B}(A)\).
-\subsubsection{Closure}
+\subsection{Closure}
 Proving equal strength of split enumerability and manifest Bishop finiteness
@@
 
-  Formally, given a type \(A\) and a type family \(B\) on \(A\), when \(A\) is
+  Formally, given a type \(A\) and a type family \(U\) on \(A\), when \(A\) is
   manifestly Bishop finite:
@@
   \end{equation}
-  And \(B\) is manifestly Bishop finite over all points of \(A\):
-  \begin{equation}
-    \prod_{(x : A)} , \mathcal{B}\left( B(x) \right)
+  And \(U\) is manifestly Bishop finite over all points of \(A\):
+  \begin{equation}
+    \prod_{(x : A)} , \mathcal{B}\left( U(x) \right)
   \end{equation}
@@
   \begin{equation}
-    \mathcal{B}\left(\prod_{(x : A)} , B(x)\right)
+    \mathcal{B}\left(\prod_{(x : A)} , U(x)\right)
   \end{equation}
@@
 \end{proof}
-\subsection{Manifest Enumerability} \label{manifest-enumerability}
+\section{Manifest Enumerability} \label{manifest-enumerability}
 As we said in section~\ref{split-enumerable}, we will now retrieve a distinction
@@
 does not interest us here.
-\subsubsection{Higher Homotopy Levels}
+\subsection{Higher Homotopy Levels}
 By hiding the position, we have essentially removed the ``decidable'' component
@@
 \end{proof}
-\subsubsection{Surjections}
+\subsection{Surjections}
 This predicates relation to surjectivity is much the same as split
@@
 \begin{lemma}
-  If a type \(A\) is manifestly enumerable, then there is a surjection from
-  \(\mathbf{Fin}\) into \(A\).
+  A proof of manifest enumerability is equivalent to a surjection from a finite
+  prefix of the natural numbers.
 \end{lemma}
@@
 \end{proof}
-And building from a surjection is similarly not too difficult:
-\begin{lemma}
-  If there is a surjection from a manifestly enumerable type \(A\) to another
-  type \(B\), then \(B\) is manifestly enumerable.
-\end{lemma}
-\begin{proof}
-  The proof proceeds in much the same way as in
-  proposition~\ref{split-enum-from-surj}.
-  The only difference is that the second component contains a truncation, but
-  since we can always eliminate from a truncation into a truncation this is not
-  a problem.
-
-  In Agda:
-  \ExecuteMetaData[agda/Cardinality/Finite/ManifestEnumerable.tex]{from-surj}
-\end{proof}
-\subsubsection{Relation to Split Enumerability}
+\subsection{Relation to Split Enumerability}
 As split enumerability is stronger than manifest enumerability, we can convert
@@
 manifestly enumerable type with decidable equality is Bishop finite.
-\subsubsection{Closure}
+\subsection{Closure}
 \begin{lemma}
@@
 manifest enumerability is not closed under function arrows.
-\subsection{Cardinal Finiteness} \label{cardinal}
+\section{Cardinal Finiteness} \label{cardinal}
 We now turn our attention to notions of finiteness which do not induce a linear
@@
 \end{definition}
-\subsubsection{Characterisation}
-Though we have chosen Bishop finiteness as our underlying finiteness proof to
-truncate, we could have in fact used any equal-strength proof
-(lemma~\ref{equal-strength}).
-Because this predicate is a mere proposition, what we called before ``equal
-strength'' now implies full equivalence.
-With this proven, we can use any one of our equal-strength finite definitions
-(\(\mathcal{E}!\), \(\mathcal{B}\)).
-In particular, we have:
-\begin{lemma}
-  Cardinal finiteness is equivalent to a propositionally truncated equivalence
-  with some \(\textbf{Fin}\).
-  \begin{equation}
-    \mathcal{C}(A) \simeq \lVert \sum_{(n : \mathbb{N})} , \textbf{Fin}(n) \simeq A \rVert
-  \end{equation}
-\end{lemma}
-\begin{proof}
-  By proposition~\ref{equal-strength}, we only need to provide conversion
-  functions between the two sides of the equivalence.
-  For left to right, we have proposition~\ref{bishop-equiv}.
-  For right to left, we can transport along the equivalence, meaning we need
-  only prove that \(\textbf{Fin}(n)\) is Bishop finite.
-  We know this to be true, since it is split enumerable
-  (proposition-\ref{fin-split-enum}).
-\end{proof}
-\subsubsection{Closure}
+\subsection{Closure}
 The closure proofs for cardinal finiteness are especially easy.
@@
 \end{proof}
-\subsubsection{Strength}
+\subsection{Strength}
 We can eliminate from cardinal finiteness in a coherently constant way:
@@
 \end{proof}
-\subsubsection{Relation to Manifest Bishop Finiteness}
+\subsection{Relation to Manifest Bishop Finiteness}
 Cardinal finiteness tells us that there is an isomorphism between a type and
@@
 \todo[inline]{Agda code for proofs}
-\subsection{Kuratowski Finiteness} \label{kuratowski}
+\section{Kuratowski Finiteness} \label{kuratowski}
 Finally we arrive at Kuratowski finiteness
 \cite{kuratowskiNotionEnsembleFini1920}.
-\subsubsection{The Kuratowski Set as a Higher Inductive Type}
+\subsection{The Kuratowski Set as a Higher Inductive Type}
 We start with the Kuratowski finite sets, which are free join semilattices (or,
@@
 \todo[inline]{Maybe add sketch of proof here?}
-\subsubsection{Elimination}
+\subsection{Elimination}
 With Agda's dependent pattern matching, we get an eliminator automatically from
@@
 
-\subsubsection{Membership}
+\subsection{Membership}
 We can define membership of a Kuratowski set using these eliminators.
@@
 
-\subsubsection{Strength}
+\subsection{Strength}
 Since the Kuratowski set is a departure in structure from our previous
@@
 \end{proof}
-\subsubsection{Topos}
+\subsection{Topos}
 At this point, we see that a ``decidable Kuratowski finite set'' is precisely
@@
 \todo[inline]{Fill in rest}
-\subsubsection{Closure}
+\subsection{Closure}
 \begin{theorem}
@@
 \section{Species and Containers} \label{species-and-containers}
-Containers \cite{abbottContainersConstructingStrictly2005} give us a general way
-to construct functors from generic building-blocks.
-\begin{definition}[Container]
-  A container is a dependent pair of shapes, and positions indexed by those
-  shapes.
-  \begin{equation}
-    \mathbf{Container} = \sum_{(\text{shape} : \text{Type})} , \text{shape} \rightarrow \text{Type}
-  \end{equation}
-  \ExecuteMetaData[agda/Data/Container.tex]{def}
-  We ``interpret'' a container into a pair of a specific shape, and a function
-  which goes from positions to contents.
-  \begin{equation}
-    \llbracket S , P \rrbracket = \prod_{(X : \text{Type})} , \sum_{(s : S)} , \left( P(s) \rightarrow X \right)
-  \end{equation}
-  \ExecuteMetaData[agda/Data/Container.tex]{interp}
-\end{definition}
-For instance, the shape of lists is given by \(\Nat\), and the positions are then
-\(\AgdaDatatype{Fin}\)s indexed by that \(\Nat\).
-\ExecuteMetaData[agda/Data/Container/List.tex]{def}
-And here is the definition of foldr for container-defined lists:
-\ExecuteMetaData[agda/Data/Container/List.tex]{foldr}
 \subsection{Containers For Finiteness}

### 2020-01-05 paper.tex
more

@@
 giving us a variant of the predicate which allows for non-set finite types.
-\subsection{Proving Equivalence with Split Surjections}
+\subsection{Split Surjections}
 Another, equivalent way to define ``finiteness'' is via a (split) surjection
 from a finite prefix of the natural numbers.
-Here, we will formalise this equivalence, by first proving an equivalence with a
-container-based definition of split enumerability.
-
+In this section, we will prove that equivalence, formally.
+
+\begin{theorem} \label{split-enum-surj}
+  Split enumerability is equivalent to a split surjection from a finite prefix
+  of the natural numbers.
+  \begin{equation}
+    \mathcal{E}!(A) \simeq \sum_{(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \twoheadrightarrow ! \; A \right)
+  \end{equation}
+\end{theorem}
+\begin{proof}
+  This proof is the composition of two intermediate equivalences: lemmas
+  \ref{container-split-enum-equiv} and \ref{container-split-enum-surj}.
+  This first of these proves that our definition of finiteness is equivalent to
+  another definition which uses lists defined as a \emph{container}, rather than
+  the usual inductive type definition.
+  The second of these shows that the container-based finiteness predicate is
+  equivalent to a split surjection from some \(\mathbf{Fin}\).
+\end{proof}
+
+To do so, we will first need to take a small detour to look at Containers.
 Containers \cite{abbottContainersConstructingStrictly2005} give us a general way
@@
     \llbracket S , P \rrbracket = \prod_{(X : \text{Type})} , \sum_{(s : S)} , \left( P(s) \rightarrow X \right)
+  \end{equation}
+  Membership of a container can be defined like so:
+  \begin{equation}
+    x \in \mathit{xs} = \text{fiber}(\text{snd}(\mathit{xs}), x)
   \end{equation}
@@
   \end{equation}
-  Membership for container-defined lists is a fiber.
-  \begin{equation}
-    x \in \mathit{xs} = \text{fiber}(\text{snd}(\mathit{xs}), x)
-  \end{equation}
 \end{lemma}
@@
   \begin{equation}
-    \mathcal{E}!(A) \simeq \sum_{(n : \mathbb{N})} , \mathbf{Fin}(n) \twoheadrightarrow ! A
+    \mathcal{E}!(A) \simeq \sum_{(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \twoheadrightarrow ! \; A \right)
   \end{equation}
@@
 \end{proof}
-This lets us prove our overall goal for this section.
-\begin{theorem} \label{split-enum-surj}
-  Split enumerability is equivalent to a split surjection from a finite prefix
-  of the natural numbers.
-\end{theorem}
-\begin{proof}
-  This follows from lemmas \ref{container-split-enum-equiv} and
-  \ref{container-split-enum-surj}.
-\end{proof}
 \subsection{Decidable Equality}
@@
 \begin{proof}
-  We use a corollary, proven in the appendix, that if there is a
-  split-surjection from \(A\) to \(B\), and \(A\) is discrete, then \(B\) is
-  also discrete.
+  We use a corollary that if there is a split-surjection from \(A\) to \(B\),
+  and \(A\) is discrete, then \(B\) is also discrete.
   \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{discrete}
-  \todo[inline]{Rework this proof to rely on the split surjection equivalence
-    from before}
-\end{proof}
-
+\end{proof}
 This also tells us that any split enumerable type must be a set.
@@
   the proof proceeds much the same as \ref{split-enum-surj}.
-  \todo[inline]{Finish prose for this proof}
+  In other words, we can define a container-based version of bishop finiteness
+  which is a reassociation of a contractible map, and then show that the
+  container-based predicate is equivalent to the list-based.
 \end{proof}
@@
   And from that we can generate our cover proof.
-
-  Full details of the formal proof are provided in the appendix.
 \end{proof}

### 2020-01-05 paper.tex
Update paper.tex
more

@@
   This first of these proves that our definition of finiteness is equivalent to
-  another definition which uses lists defined as a \emph{container}, rather than
-  the usual inductive type definition.
+  another predicate which uses lists defined as a \emph{container}
+  (definition~\ref{container-def}).
   The second of these shows that the container-based finiteness predicate is
@@
 
-To do so, we will first need to take a small detour to look at Containers.
-Containers \cite{abbottContainersConstructingStrictly2005} give us a general way
-to construct functors from generic building-blocks.
-\begin{definition}[Container]
-  A container is a dependent pair of shapes, and positions indexed by those
-  shapes.
+\begin{definition}[Container] \label{container-def}
+  A container \cite{abbottContainersConstructingStrictly2005} is a dependent
+  pair of shapes, and positions indexed by those shapes. 
   \begin{equation}
@@
 \end{definition}
-
-For instance, the shape of lists is given by \(\mathbb{N}\), and the positions are then
-\(\mathbf{Fin}\)s indexed by that \(\mathbb{N}\).
-If we swap out wherever we used ``list'' in the definition of split
-enumerability with this container-based list, we get an equivalent finiteness
-predicate.
+We often use containers to define types like lists or trees in a more generic
+way.
+For instance, while lists would usually be defined like so:
+\begin{equation}
+  \begin{aligned}
+    \mathbf{List}(A) &= \\
+      & \left| \; \text{nil} : \mathbf{List}(A)   \right; \\
+      & \left| \; \text{cons} : A \rightarrow \mathbf{List}(A) \rightarrow \mathbf{List}(A) \right;
+  \end{aligned}
+\end{equation}
+We could equivalently define them as a container:
+\begin{equation}
+  \mathbf{List}(A) = \left\llbracket \mathbb{N} , \mathbf{Fin} \right\rrbracket
+\end{equation}
+
+The two representations are equivalent, and they lead to two versions of split
+enumerability which are equivalent.
 \begin{lemma} \label{container-split-enum-equiv}
@@
 
-This version of split enumerability is equivalent to a split surjection from a
-finite prefix of the natural numbers. 
+Our interest in the container-based finiteness predicate comes from the fact
+that it is very straightforward to prove it equivalent to a split surjection
+from some \(\mathbf{Fin}\).
 \begin{lemma} \label{container-split-enum-surj}
@@
   and \(A\) is discrete, then \(B\) is also discrete.
-  \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{discrete}
-\end{proof}
-This also tells us that any split enumerable type must be a set.
+\end{proof}
 \begin{lemma}
@@
   \end{equation}
-  \ExecuteMetaData[agda/Cardinality/Finite/ManifestEnumerable.tex]{def}
-\end{definition}
-
-There is another definition, which lies between manifest enumerability and
-bishop finiteness, where we would truncate just the equality, but reveal the
-position in the membership proof.
-\begin{equation}
-  \mathcal{E}'(A) = \sum_{(\mathit{xs} : \mathbf{List}(A))} , \prod_{(x : A)} , \sum_{(i : \textbf{Fin}(\text{length}(\mathit{xs})))} , \lVert \mathit{xs} \ind i \equiv x \rVert
-\end{equation}
-However, this simply amounts to a proof of manifest Bishop finiteness over
-set-truncated types, rather than a significantly different construction, so it
-does not interest us here.
+\end{definition}
 \subsection{Higher Homotopy Levels}

### 2020-01-05 paper.tex
non it theorems etc

@@
 giving us a variant of the predicate which allows for non-set finite types.
+\subsection{Containers}
+One last part is missing from the finiteness definition above: the definition of
+\(\mathbf{List}\).
+The standard definition is as follows:
+\begin{definition}[Lists as an Inductive Type]
+  \begin{equation}
+    \begin{aligned}
+      \mathbf{List}(A) &= \\
+      & \left| \; \text{nil} : \mathbf{List}(A)   \right; \\
+      & \left| \; \text{cons} : A \rightarrow \mathbf{List}(A) \rightarrow \mathbf{List}(A) \right;
+    \end{aligned}
+  \end{equation}
+\end{definition}
+
+However, as we will see, it will be occasionally more convenient to work with an
+equivalent type defined as a \emph{container}.
+\begin{definition}[Container] \label{container-def}
+  A container \cite{abbottContainersConstructingStrictly2005} is a dependent
+  pair of shapes, and positions indexed by those shapes. 
+  \begin{equation}
+    \mathbf{Container} = \sum_{(\text{shape} : \text{Type})} , \text{shape} \rightarrow \text{Type}
+  \end{equation}
+  We ``interpret'' a container into a pair of a specific shape, and a function
+  which goes from positions to contents.
+  \begin{equation}
+    \llbracket S , P \rrbracket = \prod_{(X : \text{Type})} , \sum_{(s : S)} , \left( P(s) \rightarrow X \right)
+  \end{equation}
+  Membership of a container can be defined like so:
+  \begin{equation}
+    x \in \mathit{xs} = \text{fiber}(\text{snd}(\mathit{xs}), x)
+  \end{equation}
+\end{definition}
+
+We often use containers to define types like lists or trees in a more generic
+way.
+While they tend to be more cumbersome to work with (especially for termination
+checking) they have a number of distinct advantages.
+In particular, when finiteness predicates are defined in terms of containers,
+certain proofs (we will point them out as we come upon them) become trivial.
+
+Happily, we can have our cake and eat it too.
+Since we have access to the univalence axiom, we can transport along any
+equivalence to a more convenient representation with good justification.
+
+\begin{lemma}
+  The two definitions of lists---the traditional inductive type, and the
+  container---are equivalent.
+  \begin{equation}
+    \mathbf{List} \simeq \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket
+  \end{equation}
+\end{lemma}
+\begin{proof}
+\todo[inline]{Agda code for proof?}
+\end{proof}
+
 \subsection{Split Surjections}
@@
   equivalent to a split surjection from some \(\mathbf{Fin}\).
-\end{proof}
-
-\begin{definition}[Container] \label{container-def}
-  A container \cite{abbottContainersConstructingStrictly2005} is a dependent
-  pair of shapes, and positions indexed by those shapes. 
-  \begin{equation}
-    \mathbf{Container} = \sum_{(\text{shape} : \text{Type})} , \text{shape} \rightarrow \text{Type}
-  \end{equation}
-  We ``interpret'' a container into a pair of a specific shape, and a function
-  which goes from positions to contents.
-  \begin{equation}
-    \llbracket S , P \rrbracket = \prod_{(X : \text{Type})} , \sum_{(s : S)} , \left( P(s) \rightarrow X \right)
-  \end{equation}
-  Membership of a container can be defined like so:
-  \begin{equation}
-    x \in \mathit{xs} = \text{fiber}(\text{snd}(\mathit{xs}), x)
-  \end{equation}
-\end{definition}
-We often use containers to define types like lists or trees in a more generic
-way.
-For instance, while lists would usually be defined like so:
-\begin{equation}
-  \begin{aligned}
-    \mathbf{List}(A) &= \\
-      & \left| \; \text{nil} : \mathbf{List}(A)   \right; \\
-      & \left| \; \text{cons} : A \rightarrow \mathbf{List}(A) \rightarrow \mathbf{List}(A) \right;
-  \end{aligned}
-\end{equation}
-We could equivalently define them as a container:
-\begin{equation}
-  \mathbf{List}(A) = \left\llbracket \mathbb{N} , \mathbf{Fin} \right\rrbracket
-\end{equation}
-
-The two representations are equivalent, and they lead to two versions of split
-enumerability which are equivalent.
-\begin{lemma} \label{container-split-enum-equiv}
-  The container-based definition of split enumerability is equivalent to the
-  inductive type version.
-  \begin{equation}
-    \sum_{(\mathit{xs} : \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket A)} , \prod_{(x : A)} , x \in xs \simeq
-    \sum_{(\mathit{xs} : \textbf{List}(A))} , \prod_{(x : A)} , x \in xs
-  \end{equation}
-\end{lemma}
-\begin{proof}
-  \todo[inline]{Proof}
 \end{proof}

### 2020-01-05 paper.tex
rm theroems, definitions, lemmas

@@
 More formally:
-\begin{definition}[Split Enumerable Set]
+\begin{rm-definition}[Split Enumerable Set]
   \begin{equation}
@@
   component the ``cover'' proof.
-\end{definition}
+\end{rm-definition}
 
@@
 
-\begin{definition}[List Membership] \label{list-membership}
+\begin{rm-definition}[List Membership] \label{list-membership}
   \begin{equation}
@@
   \(\mathit{xs}\) which is equivalent to \(x\).
-\end{definition}
-\begin{definition}[\mathbf{Fin}]
+\end{rm-definition}
+\begin{rm-definition}[\mathbf{Fin}]
   \(\mathbf{Fin}\)(n) is the type of natural numbers smaller than \(n\).
@@
   \end{equation}
-\end{definition}
+\end{rm-definition}
 
@@
 The standard definition is as follows:
-\begin{definition}[Lists as an Inductive Type]
+\begin{rm-definition}[Lists as an Inductive Type]
   \begin{equation}
@@
 equivalent type defined as a \emph{container}.
-\begin{definition}[Container] \label{container-def}
+\begin{rm-definition}[Container] \label{container-def}
   A container \cite{abbottContainersConstructingStrictly2005} is a dependent
@@
 
-\begin{lemma}
+\begin{rm-lemma}
   The two definitions of lists---the traditional inductive type, and the
@@
   \end{equation}
-\end{lemma}
+\end{rm-lemma}
 \begin{proof}
@@
 
-\begin{theorem} \label{split-enum-surj}
+\begin{rm-theorem} \label{split-enum-surj}
   Split enumerability is equivalent to a split surjection from a finite prefix
@@
   \end{equation}
-\end{theorem}
+\end{rm-theorem}
 \begin{proof}
@@
 from some \(\mathbf{Fin}\).
-\begin{lemma} \label{container-split-enum-surj}
+\begin{rm-lemma} \label{container-split-enum-surj}
   Container-defined split enumerability is equivalent to a split surjection from
@@
 \subsection{Decidable Equality}
-\begin{lemma} \label{split-enum-discrete}
+\begin{rm-lemma} \label{split-enum-discrete}
   Any split enumerable type has decidable equality (is discrete).
-\end{lemma}
+\end{rm-lemma}
 \begin{proof}
@@
 \end{proof}
-\begin{lemma}
+\begin{rm-lemma}
   Any split enumerable type is a set.
-\end{lemma}
+\end{rm-lemma}
 \begin{proof}
@@
 section~\ref{manifest-bishop}.
-\begin{lemma}
+\begin{rm-lemma}
   \(\bot\) is split enumerable.
-\end{lemma}
+\end{rm-lemma}
 \begin{proof}
@@
 \end{proof}
-\begin{lemma}
+\begin{rm-lemma}
   \(\top\) is split enumerable.
-\end{lemma}
+\end{rm-lemma}
 \begin{proof}
@@
 Next, we will look into how to combine proofs of split enumerability.
-\begin{theorem}
+\begin{rm-theorem}
   Split-enumerability is closed under \(\sum\).
-\end{theorem}
+\end{rm-theorem}
 \begin{proof}
@@
 \end{proof}
-\begin{lemma}
+\begin{rm-lemma}
   Split-enumerability is closed under Cartesian product (non-dependent product).
-\end{lemma} 
+\end{rm-lemma} 
 \begin{proof}
@@
 \end{proof}
-\begin{lemma}
+\begin{rm-lemma}
   Split-enumerability is closed under disjoint union (non-dependent sum).
-\end{lemma}
+\end{rm-lemma}
 \begin{proof}
@@
 \section{Manifest Bishop Finiteness} \label{manifest-bishop}
-\begin{definition}[Manifest Bishop Finiteness]
+\begin{rm-definition}[Manifest Bishop Finiteness]
   \begin{equation}
@@
   \end{equation}
-\end{definition}
+\end{rm-definition}
 The only difference between this predicate and split enumerability is the list
@@
 (definition~\ref{list-membership}).
-\begin{definition}[Unique List Membership]
+\begin{rm-definition}[Unique List Membership]
   We say an item \(x\) is ``uniquely in'' some list \(\mathit{xs}\) if its
@@
 Those questions are answered by the following theorem:
-\begin{lemma} \label{bishop-equiv}
+\begin{rm-lemma} \label{bishop-equiv}
   A proof of manifest Bishop finiteness is equivalent to an equivalence with a
@@
 enumerability.
-\begin{lemma}
+\begin{rm-lemma}
   Any manifest Bishop finite set is split enumerable.
-\end{lemma}
+\end{rm-lemma}
 \begin{proof}
@@
 \end{proof}
-\begin{theorem} \label{split-enum-to-manifest-bishop}
+\begin{rm-theorem} \label{split-enum-to-manifest-bishop}
   Any split enumerable set is manifest Bishop finite.
-\end{theorem}
+\end{rm-theorem}
 \begin{proof}
@@
 We remedy that here.
-\begin{theorem}
+\begin{rm-theorem}
   Manifest bishop finiteness is closed over dependent functions (\(\prod\)-types).
@@
 We do so by propositionally truncating the membership proof:
-\begin{definition}[Manifest Enumerability]
+\begin{rm-definition}[Manifest Enumerability]
   \begin{equation}
@@
   \end{equation}
-\end{definition}
+\end{rm-definition}
 \subsection{Higher Homotopy Levels}
@@
 here that the circle is manifestly enumerable.
-\begin{theorem}
+\begin{rm-theorem}
   The circle \(S^1\) is manifestly enumerable.
-\end{theorem}
+\end{rm-theorem}
 \begin{proof}
@@
 Deriving a surjection from \(\mathbf{Fin}\) is straightforward:
-\begin{lemma}
+\begin{rm-lemma}
   A proof of manifest enumerability is equivalent to a surjection from a finite
   prefix of the natural numbers.
-\end{lemma}
+\end{rm-lemma}
 \begin{proof}
@@
 from one to the other easily.
-\begin{lemma}
+\begin{rm-lemma}
   Any split enumerable type is also manifestly enumerable.
-\end{lemma}
+\end{rm-lemma}
 \begin{proof}
@@
 
-\begin{theorem} \label{manifest-enum-to-split}
+\begin{rm-theorem} \label{manifest-enum-to-split}
   A manifestly enumerable type with decidable equality is split enumerable.
-\end{theorem}
+\end{rm-theorem}
 \begin{proof}
@@
 \subsection{Closure}
-\begin{lemma}
+\begin{rm-lemma}
   Manifest enumerability is closed under dependent sum, disjoint union
   (non-dependent sum), and Cartesian product (non-dependent product).
-\end{lemma}
+\end{rm-lemma}
 \begin{proof}
@@
 order.
-\begin{definition}[Cardinal Finiteness]
+\begin{rm-definition}[Cardinal Finiteness]
   A type \(A\) is cardinally finite, \(\mathcal{C}\), if it has a
@@
   \end{equation}
-\end{definition}
+\end{rm-definition}
 \subsection{Closure}
@@
 carry over.
-\begin{lemma}
+\begin{rm-lemma}
   Cardinal finiteness is closed under dependent and non-dependent sums,
   products, and functions.
-\end{lemma}
+\end{rm-lemma}
 \begin{proof}
@@
 we can pull it out, as it is constant through permutations of the list.
-\begin{theorem}
+\begin{rm-theorem}
   Given a cardinally finite type, we can derive the type's cardinality, as well
@@
 
-\begin{lemma} \label{fin-inj}
+\begin{rm-lemma} \label{fin-inj}
   \(\textbf{Fin}\) is injective.
-\end{lemma}
+\end{rm-lemma}
 \begin{proof}
@@
 Another property which does not rely on an internal order is decidable equality.
-\begin{theorem} \label{cardinal-finite-discrete}
+\begin{rm-theorem} \label{cardinal-finite-discrete}
   Any cardinal-finite set has decidable equality.
-\end{theorem}
+\end{rm-theorem}
 \begin{proof}
@@
 
-\begin{theorem}
+\begin{rm-theorem}
   Any cardinal finite type with a (decidable) total order is manifestly Bishop
   finite.
-\end{theorem}
+\end{rm-theorem}
 \begin{proof}
@@
 First, permutations.
-\begin{definition}[List Permutations]
+\begin{rm-definition}[List Permutations]
   We say that two lists are permutations of each other if there is an
@@
   \end{equation}
-\end{definition}
+\end{rm-definition}
 We also prove some of the identities you might expect with regards to
@@
 \ExecuteMetaData[agda/Data/List/Sort.tex]{sort}
-\begin{lemma}
+\begin{rm-lemma}
   Insertion sort is invariant under permutations.
@@
 We can define membership then in terms of this.
-\begin{definition}[Kuratowski Membership]
+\begin{rm-definition}[Kuratowski Membership]
   \begin{equation}
@@
   \end{equation}
-\end{definition}
+\end{rm-definition}
 
 From this we can define Kuratowski finiteness.
-\begin{definition}[Kuratowski Finiteness]
+\begin{rm-definition}[Kuratowski Finiteness]
   A type is Kuratowski finite iff there exists a Kuratowski Set which contains
@@
 order removed.
-\begin{theorem}
+\begin{rm-theorem}
   A proof of Kuratowski finiteness is equivalent to a propositionally truncated
@@
 Kuratowski finite type we retrieve cardinal finiteness.
-\begin{lemma}
+\begin{rm-lemma}
   Any Kuratowski finite set with decidable equality is cardinally finite.
-\end{lemma}
+\end{rm-lemma}
 \begin{proof}
@@
 \end{proof}
-\begin{lemma}
+\begin{rm-lemma}
   Kuratowski finiteness is equivalent to a truncated surjection from some prefix
@@
 \subsection{Closure}
-\begin{theorem}
+\begin{rm-theorem}
   Kuratowski finite sets are closed under \(\sum\).
-\end{theorem}
+\end{rm-theorem}
 \begin{proof}
@@
 This allows us to prove the following.
-\begin{theorem}
+\begin{rm-theorem}
   Manifest bishop finiteness is equivalent to an equivalence with some \(\mathbf{Fin}\).
@@
 
-\begin{lemma}
+\begin{rm-lemma}
   Manifest enumerability is equivalent to a surjection from some \(\mathbf{Fin}\).
@@
   \end{equation}
-\end{lemma}
-\begin{proof}
-\end{proof}
-
-\begin{lemma}
+\end{rm-lemma}
+\begin{proof}
+\end{proof}
+
+\begin{rm-lemma}
   Split enumerability is equivalent to a split surjection from some \(\mathbf{Fin}\).
@@
 have a support \emph{stream}, which is its infinite, coinductive counterpart.
-\begin{definition}[Stream]
+\begin{rm-definition}[Stream]
   We will work with two isomorphic definitions of streams.
@@
   infinitely-sized inhabitants.
-\end{definition}
+\end{rm-definition}
 We can now define the split countable types.
-\begin{definition}[Split Countability]
+\begin{rm-definition}[Split Countability]
   \begin{equation}
@@
   \end{equation}
-\end{definition}
+\end{rm-definition}
 \subsubsection{\(\Sigma\) Closure}
@@
 \end{figure}
-\begin{theorem} \label{split-countability-sigma}
+\begin{rm-theorem} \label{split-countability-sigma}
   Split countability is closed under \(\Sigma\).
-\end{theorem}
+\end{rm-theorem}
 \begin{proof}
@@
 \end{proof}
-\begin{lemma}
+\begin{rm-lemma}
   Split countability is closed under non-dependent product and sum.
-\end{lemma}
+\end{rm-lemma}
 \begin{proof}
@@
 These allow us to introduce recursion into countable types, by way of lists.
-\begin{definition}[Kleene Lists]
+\begin{rm-definition}[Kleene Lists]
   It is often useful to have two mutually-defined list types (one for
@@
   \end{equation}
-\end{definition}
+\end{rm-definition}
 Our definition of split countability is closed under these types in particular.
-\begin{theorem}
+\begin{rm-theorem}
   Split countability is closed under Kleene star and plus.
-\end{theorem}
+\end{rm-theorem}
 \begin{proof}
@@
 of omniscience \cite{myhillErrettBishopFoundations1972}.
-\begin{definition}[Limited Principle of Omniscience]
+\begin{rm-definition}[Limited Principle of Omniscience]
   For any type \(A\) and predicate \(P\) on \(A\), the limited principle of
@@
   that predicate is also decidable.
-\end{definition}
+\end{rm-definition}
 The limited principle of omniscience is non-constructive, but individual types
@@
 There is also a universal form of omniscience, which we call exhaustibility.
-\begin{definition}[Exhaustibility]
+\begin{rm-definition}[Exhaustibility]
   We say a type \(A\) is exhaustible if, for any decidable predicate \(P\) on
@@
   \ExecuteMetaData[agda/Relation/Nullary/Omniscience.tex]{exhaustible}
-\end{definition}
+\end{rm-definition}
 Omniscience is stronger than exhaustibility, as we can derive the latter from
 the former:
-\begin{lemma} \label{omniscient-is-exhaustible}
+\begin{rm-lemma} \label{omniscient-is-exhaustible}
   Any omniscient type is exhaustible.
-\end{lemma}
+\end{rm-lemma}
 \begin{proof}
@@
 We now show that split enumerable types are omniscient.
-\begin{theorem}
+\begin{rm-theorem}
   Any split enumerable type is omniscient.
-\end{theorem}
+\end{rm-theorem}
 \begin{proof}
@@
 \end{proof}
-\begin{theorem}
+\begin{rm-theorem}
   Any split enumerable type is exhaustible.
-\end{theorem}
+\end{rm-theorem}
 \begin{proof}
@@
 membership used, they can both be used.
-\begin{theorem}
+\begin{rm-theorem}
   For a decidable predicate \(P\) on a type \(A\), if the type is manifestly
   enumerable, then we can decide the existential of \(P\).
-\end{theorem}
+\end{rm-theorem}
 \begin{proof}
@@
 
-\begin{theorem}
+\begin{rm-theorem}
   For a decidable predicate \(P\) on a type \(A\), if the type is manifestly
   enumerable, then we can decide the universal of \(P\).
-\end{theorem}
+\end{rm-theorem}
 \begin{proof}
@@
 The quantification operations partially carry over.
-\begin{theorem}
+\begin{rm-theorem}
   For a decidable predicate \(P\) on a type \(A\), if the type is cardinal
   finite, then we can decide the universal of \(P\).
-\end{theorem}
+\end{rm-theorem}
 \begin{proof}
@@
 Clearly this operation can depend on the underlying order.
-\begin{theorem}
+\begin{rm-theorem}
   For a decidable predicate \(P\) on a type \(A\), if the type is cardinal
   finite, then we can decide the propositionally truncated existential of \(P\).
-\end{theorem}
+\end{rm-theorem}
 \begin{proof}
@@
 \end{proof}
-\begin{theorem}
+\begin{rm-theorem}
   For a decidable predicate \(P\) on a type \(A\), which uniquely determines an
   input, we can decide the existential of \(P\).
-\end{theorem}
+\end{rm-theorem}
 \begin{proof}

### 2020-01-06 paper.tex
shortened up

@@
 \end{abstract}
-\section{Introduction}
-Proofs in a constructive setting are more substantial things than their
-classical counterparts.
-Rather than just evidence for some fact, they are objects in their own right,
-which we can manipulate and extract information from.
-
-Sometimes, the extra detail can reveal interesting subtleties: fissures which
-divide homogeneous-seeming definitions into a variety of related concepts.
-Finiteness is a prime example of this phenomenon: in classical set theory, a
-finite set is a finite set, and there's not much else that we can say.
-In constructive type theory, we start the bidding at no fewer than \emph{four}
-different predicates representing finiteness
-\cite{spiwackConstructivelyFinite2010}, upped to eight in
-\cite{firsovVariationsNoetherianness2016a}: all of which differ in significant
-and interesting ways.
-
-Other times, the extra baggage is unwanted.
-Maybe we want to enforce that a particular piece of information should be
-hidden, or treated as irrelevant: ``yes, I have proven that there is a
-particular number with this property, but I don't want to reveal \emph{which}
-number.''
-
-This is where Homotopy Type Theory \cite{hottbook} can step in.
-Through primarily a generalisation of equality, we gain access to a number of
-exotic types which can, among other things, allow us to perform this ``hiding''.
-In the context of finiteness, the act of hiding is worth studying in its own
-right, and further delineates several new finiteness predicates which collapse
-even in standard Martin-Löf type theory
-\cite{martin-lofIntuitionisticTypeTheory1980}.
-
-The other headline feature of HoTT's generalised equality is univalence: put
-simply, it allows us to treat isomorphic types as equal, giving us in the
-constructive world access to a technique that is commonplace classically.
-It turns out that finite types have a lot to say about isomorphisms, and we will
-rely heavily on this aspect of HoTT to work with them.
-\subsection{Contributions}
-In this work, we will explore finite types in Cubical Type Theory
-\cite{cohenCubicalTypeTheory2016}, and expose their relationship to infinite
-types, species, and demonstrate their practical uses for proofs in dependently
-typed programming languages.
-\subsubsection{Strong Finiteness Predicates}
-We will first explore the ``strong'' notions of finiteness (i.e. those at least
-as strong as Kuratowski finiteness \cite{kuratowskiNotionEnsembleFini1920}),
-with a special focus on cardinal finiteness (section~\ref{cardinal}), and
-manifest enumerability (section~\ref{manifest-enumerability}), which is new, to
-our knowledge.
-
-\input{figures/finite-classification}
-Figure~\ref{finite-classification} organises the predicates according to
-their ``strength''; i.e. how much information they provide about a conforming
-type.
-For instance, a proof that some type \(A\) is manifestly Bishop finite (the
-strongest of the notions, explored in section~\ref{manifest-bishop}) also tells
-us that \(A\) is discrete (has decidable equality), and gives us a linear order
-on the type.
-A type that is Kuratowski finite (section~\ref{kuratowski}) has no such extra
-features: indeed, we will see examples of Kuratowski finite types which are not
-even sets, never mind discrete ones.
-
-We will go through each of the predicates, proving how to weaken each (i.e. we
-will provide a proof that every cardinally finite type is Kuratowski finite),
-and how to strengthen them, given the required property.
-In terms of figure~\ref{finite-classification}, this amounts to providing proofs
-for each arrow.
-
-We will---through the use of containers
-\cite{abbottContainersConstructingStrictly2005}---formally prove the equivalence
-these predicates have with the usual function relations i.e. we will show that a
-proof of manifest enumerability is precisely equivalent to a surjection from a
-finite prefix of the natural numbers.
-
-For each predicate, we will also prove its closure over sums and products in
-both dependent and non-dependent forms, if such a closure exists.
-This will culminate in our main result for this section: the formal proof that
-decidable Kuratowski finite sets form a topos. \todo[noline]{reference for this}
-More specifically, we show that cardinal finite sets form a topos, that
-decidable Kuratowski finite sets are equivalent in strength to cardinal finite
-sets, and carry the proof over.
-This result relies on proofs on each of the other finiteness predicates.
-\subsubsection{Species}
-\todo[inline]{Redo this next paragraph}
-In section~\ref{species-and-containers}, we will redefine our finiteness
-predicates using \emph{containers}, and prove that the new definitions are
-equivalent to the old.
-Using this, we will prove much stronger relationships between our finiteness
-predicates and relation-based definitions of finiteness.
-Finally, We will show how the species-container duality falls out naturally from
-these new definitions.
-
-\subsubsection{Infinite Types}
-In section~\ref{infinite-cardinalities}, we will extend our study of finite
-types to infinite but countable types.
-We will see that the finiteness predicates are mirrored with countable
-counterparts, and we will prove closure under the Kleene star and plus.
-
-\subsubsection{Practical Uses of Finiteness}
-Proofs of finiteness have well-known practical applications in
-constructive mathematics \cite{firsovDependentlyTypedProgramming2015}.
-In section~\ref{practical}, we build a library which exploits these uses in
-Cubical Agda \cite{vezzosiCubicalAgdaDependently2019}, allowing automation of
-complex proofs over finite types.
-We frame this in terms of the principle of omniscience for finite types.
-Thanks to the flexibility afforded to us by Cubical Type Theory, we are able to
-go further than the usual examples of this kind of proof automation: as well as
-proving properties about functions, we can synthesise functions whole-cloth from
-their desired properties.
-Through the unified interface for finite and countable types, we can
-reuse the automation machinery for \emph{partial} proof search over infinite
-search spaces.
-Along the way, we extend the work in \cite{allaisGenericLevelPolymorphic2019} to
-prove isomorphisms between the curried and uncurried forms of \(n\)-ary
-dependent functions.
+% \section{Introduction}
+% Proofs in a constructive setting are more substantial things than their
+% classical counterparts.
+% Rather than just evidence for some fact, they are objects in their own right,
+% which we can manipulate and extract information from.
+
+% Sometimes, the extra detail can reveal interesting subtleties: fissures which
+% divide homogeneous-seeming definitions into a variety of related concepts.
+% Finiteness is a prime example of this phenomenon: in classical set theory, a
+% finite set is a finite set, and there's not much else that we can say.
+% In constructive type theory, we start the bidding at no fewer than \emph{four}
+% different predicates representing finiteness
+% \cite{spiwackConstructivelyFinite2010}, upped to eight in
+% \cite{firsovVariationsNoetherianness2016a}: all of which differ in significant
+% and interesting ways.
+
+% Other times, the extra baggage is unwanted.
+% Maybe we want to enforce that a particular piece of information should be
+% hidden, or treated as irrelevant: ``yes, I have proven that there is a
+% particular number with this property, but I don't want to reveal \emph{which}
+% number.''
+
+% This is where Homotopy Type Theory \cite{hottbook} can step in.
+% Through primarily a generalisation of equality, we gain access to a number of
+% exotic types which can, among other things, allow us to perform this ``hiding''.
+% In the context of finiteness, the act of hiding is worth studying in its own
+% right, and further delineates several new finiteness predicates which collapse
+% even in standard Martin-Löf type theory
+% \cite{martin-lofIntuitionisticTypeTheory1980}.
+
+% The other headline feature of HoTT's generalised equality is univalence: put
+% simply, it allows us to treat isomorphic types as equal, giving us in the
+% constructive world access to a technique that is commonplace classically.
+% It turns out that finite types have a lot to say about isomorphisms, and we will
+% rely heavily on this aspect of HoTT to work with them.
+% \subsection{Contributions}
+% In this work, we will explore finite types in Cubical Type Theory
+% \cite{cohenCubicalTypeTheory2016}, and expose their relationship to infinite
+% types, species, and demonstrate their practical uses for proofs in dependently
+% typed programming languages.
+% \subsubsection{Strong Finiteness Predicates}
+% We will first explore the ``strong'' notions of finiteness (i.e. those at least
+% as strong as Kuratowski finiteness \cite{kuratowskiNotionEnsembleFini1920}),
+% with a special focus on cardinal finiteness (section~\ref{cardinal}), and
+% manifest enumerability (section~\ref{manifest-enumerability}), which is new, to
+% our knowledge.
+
+% \input{figures/finite-classification}
+% Figure~\ref{finite-classification} organises the predicates according to
+% their ``strength''; i.e. how much information they provide about a conforming
+% type.
+% For instance, a proof that some type \(A\) is manifestly Bishop finite (the
+% strongest of the notions, explored in section~\ref{manifest-bishop}) also tells
+% us that \(A\) is discrete (has decidable equality), and gives us a linear order
+% on the type.
+% A type that is Kuratowski finite (section~\ref{kuratowski}) has no such extra
+% features: indeed, we will see examples of Kuratowski finite types which are not
+% even sets, never mind discrete ones.
+
+% We will go through each of the predicates, proving how to weaken each (i.e. we
+% will provide a proof that every cardinally finite type is Kuratowski finite),
+% and how to strengthen them, given the required property.
+% In terms of figure~\ref{finite-classification}, this amounts to providing proofs
+% for each arrow.
+
+% We will---through the use of containers
+% \cite{abbottContainersConstructingStrictly2005}---formally prove the equivalence
+% these predicates have with the usual function relations i.e. we will show that a
+% proof of manifest enumerability is precisely equivalent to a surjection from a
+% finite prefix of the natural numbers.
+
+% For each predicate, we will also prove its closure over sums and products in
+% both dependent and non-dependent forms, if such a closure exists.
+% This will culminate in our main result for this section: the formal proof that
+% decidable Kuratowski finite sets form a topos. \todo[noline]{reference for this}
+% More specifically, we show that cardinal finite sets form a topos, that
+% decidable Kuratowski finite sets are equivalent in strength to cardinal finite
+% sets, and carry the proof over.
+% This result relies on proofs on each of the other finiteness predicates.
+% \subsubsection{Species}
+% \todo[inline]{Redo this next paragraph}
+% In section~\ref{species-and-containers}, we will redefine our finiteness
+% predicates using \emph{containers}, and prove that the new definitions are
+% equivalent to the old.
+% Using this, we will prove much stronger relationships between our finiteness
+% predicates and relation-based definitions of finiteness.
+% Finally, We will show how the species-container duality falls out naturally from
+% these new definitions.
+
+% \subsubsection{Infinite Types}
+% In section~\ref{infinite-cardinalities}, we will extend our study of finite
+% types to infinite but countable types.
+% We will see that the finiteness predicates are mirrored with countable
+% counterparts, and we will prove closure under the Kleene star and plus.
+
+% \subsubsection{Practical Uses of Finiteness}
+% Proofs of finiteness have well-known practical applications in
+% constructive mathematics \cite{firsovDependentlyTypedProgramming2015}.
+% In section~\ref{practical}, we build a library which exploits these uses in
+% Cubical Agda \cite{vezzosiCubicalAgdaDependently2019}, allowing automation of
+% complex proofs over finite types.
+% We frame this in terms of the principle of omniscience for finite types.
+% Thanks to the flexibility afforded to us by Cubical Type Theory, we are able to
+% go further than the usual examples of this kind of proof automation: as well as
+% proving properties about functions, we can synthesise functions whole-cloth from
+% their desired properties.
+% Through the unified interface for finite and countable types, we can
+% reuse the automation machinery for \emph{partial} proof search over infinite
+% search spaces.
+% Along the way, we extend the work in \cite{allaisGenericLevelPolymorphic2019} to
+% prove isomorphisms between the curried and uncurried forms of \(n\)-ary
+% dependent functions.
 \section{Split Enumerability} \label{split-enumerable}
@@
       \mathbf{Fin}(0) &= \bot \\
-      \mathbf{Fin}(n + 1) &= 1 + \mathbf{Fin}(n)
+      \mathbf{Fin}(n + 1) &= \top + \mathbf{Fin}(n)
     \end{aligned}
@@
 
+For each finiteness predicate, we have proven it equivalent to its
+container-based version.
+These proofs are included in the Agda formalisation.
 \subsection{Split Surjections}
@@
 \begin{proof}
-  This proof is the composition of two intermediate equivalences: lemmas
-  \ref{container-split-enum-equiv} and \ref{container-split-enum-surj}.
-  This first of these proves that our definition of finiteness is equivalent to
-  another predicate which uses lists defined as a \emph{container}
-  (definition~\ref{container-def}).
-  The second of these shows that the container-based finiteness predicate is
-  equivalent to a split surjection from some \(\mathbf{Fin}\).
-\end{proof}
-
-Our interest in the container-based finiteness predicate comes from the fact
-that it is very straightforward to prove it equivalent to a split surjection
-from some \(\mathbf{Fin}\).
-\begin{rm-lemma} \label{container-split-enum-surj}
-  Container-defined split enumerability is equivalent to a split surjection from
-  some \(\mathbf{Fin}\).
-  \begin{equation}
-    \mathcal{E}!(A) \simeq \sum_{(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \twoheadrightarrow ! \; A \right)
-  \end{equation}
-\end{rm-lemma}
-\begin{proof}
+  For this proof, we will work with the container-based finiteness predicate.
   The proof is surprisingly short: after sufficient inlining, it emerges that
@@
 \begin{proof}
-  Since non-dependent products are simply a special case of the \(\sum\) type,
-  we can reuse the proof of closure over \(\sum\) here.
-\end{proof}
-\begin{rm-lemma}
-  Split-enumerability is closed under disjoint union (non-dependent sum).
-\end{rm-lemma}
-\begin{proof}
-  Again, this can be derived from closure of the \(\sum\) type.
-  Disjoint union between two types \(A\) and \(B\) can be represented by the
+\end{proof}
+\begin{rm-lemma}
+  Split-enumerability is closed under disjoint union (non-dependent sum) and
+  Cartesian product (non-dependent product).
+\end{rm-lemma}
+\begin{proof}
+  Both of these closures can be derived from closure of the \(\sum\) type.
+  The non-dependent product is equivalent to a special case of \(\sum\):
+  \begin{equation}
+    A \times B \simeq \sum_{(x : A)} , B
+  \end{equation}
+  And disjoint union between two types \(A\) and \(B\) can be represented by the
   following type:
@@
 \subsection{Equivalence}
-While manifest Bishop finiteness may seem stricter than split enumerability, we
-will see soon that it is actually equivalent in strength.
-Why study it separately to split enumerability at all, then?
-What purpose does it serve on our road to Kuratowski finiteness?
-Those questions are answered by the following theorem:
+Where split enumerability was the enumeration form of a surjection from
+\(\mathbf{Fin}\), we see here that manifest Bishop finiteness is the enumeration
+form of an \emph{equivalence} with \(\mathbf{Fin}\).
 \begin{rm-lemma} \label{bishop-equiv}
@@
   finiteness.
-  \ExecuteMetaData[agda/Cardinality/Finite/ManifestBishop.tex]{bishop-to-split-enum}
 \end{proof}
@@
   We say that two lists are permutations of each other if there is an
-  isomorphism between membership proofs.
+  isomorphism between membership proofs
+  \cite{danielssonBagEquivalenceProofRelevant2012}.
   \begin{equation}
@@
 In particular, we prove that they form an equivalence relation.
-
-Next, we will use the following definition of insertion sort:
-\ExecuteMetaData[agda/Data/List/Sort.tex]{sort}
 \begin{rm-lemma}
@@
 \subsection{Omniscience}
-\todo[inline]{Put definition for decidable somewhere}
 In this section we are interested in restricted forms of the limited principle
@@
 \begin{rm-theorem}
-  Any split enumerable type is exhaustible.
-\end{rm-theorem}
-\begin{proof}
-  Via lemma~\ref{omniscient-is-exhaustible}.
-\end{proof}
-
-\subsubsection{Manifest Enumerability}
-We now reach for the same quantification combinators we used for
-split enumerable types.
-Because neither of these quantifications ``care'' about the specific proof of
-membership used, they can both be used.
+  Any manifestly enumerable type is omniscient.
+\end{rm-theorem}
+\begin{proof}
+  \todo[inline]{Proof}
+\end{proof}
 \begin{rm-theorem}
-  For a decidable predicate \(P\) on a type \(A\), if the type is manifestly
-  enumerable, then we can decide the existential of \(P\).
+  Any cardinal finite type is exhaustible.
 \end{rm-theorem}
@@
 \end{proof}
-
 \begin{rm-theorem}
-  For a decidable predicate \(P\) on a type \(A\), if the type is manifestly
-  enumerable, then we can decide the universal of \(P\).
-\end{rm-theorem}
-\begin{proof}
-  \todo[inline]{Proof}
-\end{proof}
-\subsubsection{Cardinal Finiteness}
-The quantification operations partially carry over.
+  Cardinally finite types are omniscient about prop-valued predicates.
+\end{rm-theorem}
 \begin{rm-theorem}
-  For a decidable predicate \(P\) on a type \(A\), if the type is cardinal
-  finite, then we can decide the universal of \(P\).
-\end{rm-theorem}
-\begin{proof}
-  \todo[inline]{Proof}
-\end{proof}
-
-An existential predicate is harder to express.
-In previous sections, we simply tested each member of the set in turn, and
-returned the first member which passed the test.
-Clearly this operation can depend on the underlying order.
-\begin{rm-theorem}
-  For a decidable predicate \(P\) on a type \(A\), if the type is cardinal
-  finite, then we can decide the propositionally truncated existential of \(P\).
-\end{rm-theorem}
-\begin{proof}
-  \todo[inline]{proof}
-\end{proof}
-\begin{rm-theorem}
-  For a decidable predicate \(P\) on a type \(A\), which uniquely determines an
-  input, we can decide the existential of \(P\).
-\end{rm-theorem}
-\begin{proof}
-  \todo[inline]{proof}
-\end{proof}
-\subsubsection{Kuratowski}
-\begin{itemize}
-\item Forall
-\item Exists
-\end{itemize}
+  Kuratowski finite types have the same omniscience properties as cardinal
+  finite types.
+\end{rm-theorem}
 

### 2020-01-06 paper.tex
tightened up a little

@@
 
-The term \(x \in \mathit{xs}\) hides a subtle detail of this definition.
-Depending on how we choose to represent \(x \in \mathit{xs}\), the resulting
-predicate will be equivalent in strength to either manifest Bishop finiteness
-(section~\ref{manifest-bishop}) or manifest enumerability
-(section~\ref{manifest-enumerability}).
-For now, we will define it like so:
-
-\begin{rm-definition}[List Membership] \label{list-membership}
-  \begin{equation}
-    x \in \mathit{xs} = \sum_{(i : \mathbf{Fin}(\text{length}(\mathit{xs})))} , \mathit{xs} \ind i \equiv x
-  \end{equation}
-  Here, \(\ind\) is a safe indexing function.
-  In other words, to say that some item \(x\) is in some list \(\mathit{xs}\),
-  we say that there is an index into \(\mathit{xs}\) which selects an element of
-  \(\mathit{xs}\) which is equivalent to \(x\).
-\end{rm-definition}
+We will at this point take a moment to define some of the types we used to
+define \(\mathcal{E}!\).
+Lists, and membership thereof, are defined using \emph{containers}.
+\begin{rm-definition}[Container] \label{container-def}
+  A container \cite{abbottContainersConstructingStrictly2005} is a dependent
+  pair of shapes, and positions indexed by those shapes. 
+  \begin{equation}
+    \mathbf{Container} = \sum_{(\text{shape} : \text{Type})} , \text{shape} \rightarrow \text{Type}
+  \end{equation}
+  We ``interpret'' a container into a pair of a specific shape, and a function
+  which goes from positions to contents.
+  \begin{equation}
+    \llbracket S , P \rrbracket = \prod_{(X : \text{Type})} , \sum_{(s : S)} , \left( P(s) \rightarrow X \right)
+  \end{equation}
+  Membership of a container can be defined like so:
+  \begin{equation}
+    x \in \mathit{xs} = \text{fiber}(\text{snd}(\mathit{xs}), x)
+  \end{equation}
+\end{rm-definition}
+Containers can be used to define a wide variety of functors (streams, trees,
+etc.): lists are all that interest us now.
 \begin{rm-definition}[\mathbf{Fin}]
@@
 \end{rm-definition}
-
-Clearly, this definition of list membership tells us not just that an item is in
-a list, but also its \emph{position} in that list.
-That piece of information gives use decidable equality, restricting the
-predicate to discrete sets.
-In section~\ref{manifest-enumerability}, we will show how to hide the position,
-giving us a variant of the predicate which allows for non-set finite types.
-\subsection{Containers}
-One last part is missing from the finiteness definition above: the definition of
-\(\mathbf{List}\).
-The standard definition is as follows:
-\begin{rm-definition}[Lists as an Inductive Type]
-  \begin{equation}
-    \begin{aligned}
-      \mathbf{List}(A) &= \\
-      & \left| \; \text{nil} : \mathbf{List}(A)   \right; \\
-      & \left| \; \text{cons} : A \rightarrow \mathbf{List}(A) \rightarrow \mathbf{List}(A) \right;
-    \end{aligned}
-  \end{equation}
-\end{rm-definition}
-
-However, as we will see, it will be occasionally more convenient to work with an
-equivalent type defined as a \emph{container}.
-\begin{rm-definition}[Container] \label{container-def}
-  A container \cite{abbottContainersConstructingStrictly2005} is a dependent
-  pair of shapes, and positions indexed by those shapes. 
-  \begin{equation}
-    \mathbf{Container} = \sum_{(\text{shape} : \text{Type})} , \text{shape} \rightarrow \text{Type}
-  \end{equation}
-  We ``interpret'' a container into a pair of a specific shape, and a function
-  which goes from positions to contents.
-  \begin{equation}
-    \llbracket S , P \rrbracket = \prod_{(X : \text{Type})} , \sum_{(s : S)} , \left( P(s) \rightarrow X \right)
-  \end{equation}
-  Membership of a container can be defined like so:
-  \begin{equation}
-    x \in \mathit{xs} = \text{fiber}(\text{snd}(\mathit{xs}), x)
-  \end{equation}
-\end{rm-definition}
-
-We often use containers to define types like lists or trees in a more generic
-way.
-While they tend to be more cumbersome to work with (especially for termination
-checking) they have a number of distinct advantages.
-In particular, when finiteness predicates are defined in terms of containers,
-certain proofs (we will point them out as we come upon them) become trivial.
-
-Happily, we can have our cake and eat it too.
-Since we have access to the univalence axiom, we can transport along any
-equivalence to a more convenient representation with good justification.
-
-\begin{rm-lemma}
-  The two definitions of lists---the traditional inductive type, and the
-  container---are equivalent.
-  \begin{equation}
-    \mathbf{List} \simeq \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket
-  \end{equation}
-\end{rm-lemma}
-\begin{proof}
-  \todo[inline]{Agda code for proof?}
-\end{proof}
-
-For each finiteness predicate, we have proven it equivalent to its
-container-based version.
-These proofs are included in the Agda formalisation.
+\begin{rm-definition}[Lists]
+  The ``shape'' of lists is \(\mathbb{N}\), indicating the length of the list in
+  question.
+  \begin{equation}
+    \mathbf{List} = \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket
+  \end{equation}
+\end{rm-definition}
+
+Internally, in our formalisation, we actually use the standard
+inductive definition of lists more often (it tends to work better in more
+complex algorithms, and functions on it seems to satisfy the termination
+checker more readily).
+However, since both types are equivalent, univalence allows us to transport to
+whichever representation is more convenient in a given situation.
+For the higher-level proofs we present here, though, the container-based
+definition greatly simplifies certain steps, which is why we have chosen it as
+our representation.
 \subsection{Split Surjections}
@@
 \begin{proof}
-  For this proof, we will work with the container-based finiteness predicate.
   The proof is surprisingly short: after sufficient inlining, it emerges that
@@
   ``do-notation'' is available to us as we're working in the list monad.
-
-  The cover proof follows in a simple, if tedious, way.
-\end{proof}
-\begin{rm-lemma}
-  Split-enumerability is closed under Cartesian product (non-dependent product).
-\end{rm-lemma} 
-\begin{proof}
 \end{proof}
@@
 to be read as ``\(x\) occurs exactly once in \(\mathit{xs}\)''.
-It is defined in terms of our previous list membership term
-(definition~\ref{list-membership}).
-\begin{rm-definition}[Unique List Membership]
-  We say an item \(x\) is ``uniquely in'' some list \(\mathit{xs}\) if its
+\begin{rm-definition}[Unique Membership]
+  We say an item \(x\) is ``uniquely in'' some container \(\mathit{xs}\) if its
   membership in that list is a \emph{contraction}; i.e. its membership proof

### 2020-01-06 paper.tex
more

@@
 \begin{rm-definition}[Container] \label{container-def}
-  A container \cite{abbottContainersConstructingStrictly2005} is a dependent
-  pair of shapes, and positions indexed by those shapes. 
-  \begin{equation}
-    \mathbf{Container} = \sum_{(\text{shape} : \text{Type})} , \text{shape} \rightarrow \text{Type}
-  \end{equation}
-  We ``interpret'' a container into a pair of a specific shape, and a function
-  which goes from positions to contents.
-  \begin{equation}
-    \llbracket S , P \rrbracket = \prod_{(X : \text{Type})} , \sum_{(s : S)} , \left( P(s) \rightarrow X \right)
+  A container \cite{abbottContainersConstructingStrictly2005} is a pair
+  \(S \triangleright P\) where \(S\) is a type, the elements of which are called
+  the \emph{shapes} of the container, and \(P\) is a type family on \(S\), where
+  the elements of \(P(s)\) are called the \emph{positions} of a container.
+  \begin{equation}
+    \begin{aligned}
+      S : \text{Type}, P : S \rightarrow \text{Type} \\
+      \mathbf{Container} = S \triangleright P
+    \end{aligned}
+  \end{equation}
+  We ``interpret'' a container into a functor defined like so:
+  \begin{equation}
+    \llbracket S \triangleright P \rrbracket = \prod_{(X : \text{Type})} , \sum_{(s : S)} , \left( P(s) \rightarrow X \right)
   \end{equation}
@@
 \begin{rm-lemma}
-  \(\bot\) is split enumerable.
-\end{rm-lemma}
-\begin{proof}
-  The support list for \(\bot\) is the empty list.
-  The cover proof is present via the principle of explosion.
-\end{proof}
-\begin{rm-lemma}
-  \(\top\) is split enumerable.
-\end{rm-lemma}
-\begin{proof}
-  The support list for \(\top\) is the singleton list containing \(\text{tt}\).
-  The cover proof is a constant function which returns an index into the first
-  item in the list.
-\end{proof}
-Both of those cases were relatively simple.
+  \(\bot\), \(\top\), and \(\mathbf{Bool}\) are all split enumerable.
+\end{rm-lemma}
+\begin{proof}
+  These non-recursive types have similar, simple proofs.
+  For each we first provide the support list: they are \([]\), \([\text{tt}]\),
+  and \([\text{false}, \text{true}]\) respectively.
+  The cover proof should return an index which points at the given element: for
+  \(\bot\) this function is present via the principle of explosion, for \(\top\)
+  we always return a \(0\), and for \(\mathbf{Bool}\) we return a \(0\) for
+  \(\text{false}\), and a \(1\) for \(\text{true}\).
+  \begin{multicols}{3}
+    \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{bool}
+    \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{unit}
+    \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{empty}
+  \end{multicols}
+\end{proof}
 Next, we will look into how to combine proofs of split enumerability.
@@
   the proof proceeds much the same as \ref{split-enum-surj}.
-  In other words, we can define a container-based version of bishop finiteness
-  which is a reassociation of a contractible map, and then show that the
-  container-based predicate is equivalent to the list-based.
+  In other words, the definition of Bishop finiteness is itself a reassociation
+  of a contractible map.
 \end{proof}
@@
   Our strategy to accomplish this will be to consider functions from
-  \(\mathbf{Fin}(n)\) as \(n\)-tuples over some type family \(T\).
+  \(\mathbf{Fin}(n)\) as \(n\)-tuples over some type family \(T : \mathbb{N}
+  \rightarrow \text{Type}\).
   \begin{equation}
@@
       \mathbf{Tuple}(T, 0)   &= \top \\
-      \mathbf{Tuple}(T, n+1) &= A(0) \times \mathbf{Tuple}(T \circ \text{suc}, n)
+      \mathbf{Tuple}(T, n+1) &= T(0) \times \mathbf{Tuple}(T \circ \text{suc}, n)
     \end{aligned}
@@
 \section{Manifest Enumerability} \label{manifest-enumerability}
-As we said in section~\ref{split-enumerable}, we will now retrieve a distinction
-between enumerability and manifest Bishop finiteness.
-We do so by propositionally truncating the membership proof:
+A defining feature of split enumerability and Bishop finiteness is decidability:
+both predicates imply a decidable equality function on the underlying type.
+To find a predicate for finiteness which doesn't imply this, and therefore
+works with types other than just sets, we \emph{truncate} the membership proof.
 \begin{rm-definition}[Manifest Enumerability]
@@
   \end{equation}
+\end{rm-definition}
+\subsection{Propositional Truncation}
+\begin{rm-definition}[Propositional Truncation]
+  The type \(\lVert A \rVert\) on some type \(A\) is a propositionally truncated
+  proof of \(A\).
+  In other words, it is a proof that some \(A\) exists, but it does not tell you
+  \emph{which} \(A\).
+
+  It is defined as a Higher Inductive Type:
+  \begin{equation}
+    \begin{aligned}
+      \lVert A \rVert &= \\
+        &| \; \lvert \cdot \rvert : A \rightarrow \lVert A \rVert ; \\
+        &| \; \text{squash} : \prod_{(x, y : \lVert A \rVert)} , x \equiv y  ; \\
+    \end{aligned}
+  \end{equation}
+
+  To eliminate from \(\lVert A \rVert\) we need to prove that we don't ``care''
+  about the specific choice of \(A\).
+  In other words, given a function \(f : A \rightarrow B\), to construct a
+  function \(\lVert A \rVert \rightarrow B\), \(f\) needs to satisfy the following:
+  \begin{equation}
+    \prod_{(x , y : A)} , f(x) \equiv f(y)
+  \end{equation}
+  Formally speaking, \(f\) needs to be ``coherently constant''
+  \cite{krausGeneralUniversalProperty2015}.
 \end{rm-definition}

### 2020-01-06 paper.tex
most of beginning done
much, much cleaner

@@
   \begin{equation}
-    \mathcal{E}!(A) = \sum_{(\mathit{xs} : \textbf{List}(A))} , \prod_{(x : A)} , x \in xs
+    \mathcal{E}!(A) = \Sigma {(\mathit{xs} : \textbf{List}(A))} , \Pi {(x : A)} , x \in xs
   \end{equation}
@@
   \begin{equation}
-    \llbracket S \triangleright P \rrbracket = \prod_{(X : \text{Type})} , \sum_{(s : S)} , \left( P(s) \rightarrow X \right)
+    \llbracket S \triangleright P \rrbracket = \Pi {(X : \text{Type})} , \Sigma {(s : S)} , \left( P(s) \rightarrow X \right)
   \end{equation}
@@
   \begin{equation}
-    \mathcal{E}!(A) \simeq \sum_{(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \twoheadrightarrow ! \; A \right)
+    \mathcal{E}!(A) \simeq \Sigma {(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \twoheadrightarrow ! \; A \right)
   \end{equation}
@@
   \ExecuteMetaData[agda/Cardinality/Container/Finite.tex]{split-surj}
+  To be clear: in Agda, the proof could simply be \(\AgdaFunction{reassoc}\); we
+  have written out the extra lines for clarity alone.
 \end{proof}
@@
   \begin{equation}
-    E_U : \prod_{(x : A)} , \mathcal{E}!(U(x))
+    E_U : \Pi {(x : A)} , \mathcal{E}!(U(x))
   \end{equation}
@@
   \begin{equation}
-    A \times B \simeq \sum_{(x : A)} , B
+    A \times B \simeq \Sigma {(x : A)} , B
   \end{equation}
@@
   \begin{equation}
-    A + B = \sum_{(x : \mathbf{Bool})} , \text{if} \; x \; \text{then} \; A \; \text{else} \; B
+    A + B = \Sigma {(x : \mathbf{Bool})} , \text{if} \; x \; \text{then} \; A \; \text{else} \; B
   \end{equation}
@@
   \begin{equation}
-    \mathcal{B}(A) = \sum_{(\mathit{xs} : \textbf{List}(A))} , \prod_{(x : A)} , x \inunique xs
+    \mathcal{B}(A) = \Sigma {(\mathit{xs} : \textbf{List}(A))} , \Pi {(x : A)} , x \inunique xs
   \end{equation}
@@
   \begin{equation}
-    \mathcal{B}(A) \simeq \sum_{(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \simeq A \right)
+    \mathcal{B}(A) \simeq \Sigma {(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \simeq A \right)
   \end{equation}
@@
   \begin{equation}
-    \prod_{(x : A)} , \prod_{(\mathit{xs} : \textbf{List}(A))} , x \in \mathit{xs} \rightarrow x \inunique \text{uniques}(\mathit{xs})
+    \Pi {(x : A)} , \Pi {(\mathit{xs} : \textbf{List}(A))} , x \in \mathit{xs} \rightarrow x \inunique \text{uniques}(\mathit{xs})
   \end{equation}
@@
   \begin{equation}
-    \prod_{(x : A)} , \mathcal{B}\left( U(x) \right)
+    \Pi {(x : A)} , \mathcal{B}\left( U(x) \right)
   \end{equation}
@@
   \begin{equation}
-    \mathcal{B}\left(\prod_{(x : A)} , U(x)\right)
+    \mathcal{B}\left(\Pi {(x : A)} , U(x)\right)
   \end{equation}
@@
   \begin{equation}
-    \left( \prod_{(x : \mathbf{Fin}(n))} , \mathcal{E}!(A(x)) \right) \rightarrow \mathcal{E}!\left( \prod_{(x : \mathbf{Fin}(n))} , A(x) \right)
+    \left( \Pi {(x : \mathbf{Fin}(n))} , \mathcal{E}!(A(x)) \right) \rightarrow \mathcal{E}!\left( \Pi {(x : \mathbf{Fin}(n))} , A(x) \right)
   \end{equation}
@@
   \begin{equation}
-    \mathbf{Tuple}(T, n) \iff \prod_{(x : \mathbf{Fin}(n))} , T(x)
+    \mathbf{Tuple}(T, n) \iff \Pi {(x : \mathbf{Fin}(n))} , T(x)
   \end{equation}
@@
   \begin{equation}
-    \mathcal{E}(A) = \sum_{(\mathit{xs} : \mathbf{List}(A))} , \prod_{(x : A)} , \lVert x \in \mathit{xs} \rVert
-  \end{equation}
-\end{rm-definition}
-\subsection{Propositional Truncation}
-\begin{rm-definition}[Propositional Truncation]
+    \mathcal{E}(A) = \Sigma {(\mathit{xs} : \mathbf{List}(A))} , \Pi {(x : A)} , \lVert x \in \mathit{xs} \rVert
+  \end{equation}
+\end{rm-definition}
+\begin{rm-definition}[Propositional Truncation]a \label{prop-trunc}
   The type \(\lVert A \rVert\) on some type \(A\) is a propositionally truncated
@@
         &| \; \lvert \cdot \rvert : A \rightarrow \lVert A \rVert ; \\
-        &| \; \text{squash} : \prod_{(x, y : \lVert A \rVert)} , x \equiv y  ; \\
+        &| \; \text{squash} : \Pi {(x, y : \lVert A \rVert)} , x \equiv y  ; \\
     \end{aligned}
@@
 
-  To eliminate from \(\lVert A \rVert\) we need to prove that we don't ``care''
-  about the specific choice of \(A\).
-  In other words, given a function \(f : A \rightarrow B\), to construct a
-  function \(\lVert A \rVert \rightarrow B\), \(f\) needs to satisfy the following:
-  \begin{equation}
-    \prod_{(x , y : A)} , f(x) \equiv f(y)
+  We will use and consume values of the type \(\lVert A \rVert\) in three main
+  ways.
+  \begin{enumerate}
+    \item \label{elim-prop-prop} We can first eliminate from \(\lVert A \rVert\)
+      into any type which is a proposition.
+      In other words, given a function \(f : A \rightarrow B\), and a proof that
+      \(B\) is a proposition, we can construct a function \(\lVert A \rVert
+      \rightarrow B\).
+    \item \label{elim-prop-monad} As a consequence of this first point, we can
+      always eliminate into another propositionally truncated type.
+      As a result, \(\lVert \cdot \rVert\) forms a Monad: for our purposes, this
+      simply means tat we can work ``under'' a propositional truncation in an
+      ergonomic way.
+    \item \label{elim-prop-coh} We can eliminate from \(\lVert A \rVert\) with a
+      function \(f : A \rightarrow B\) iff \(f\) ``doesn't care'' about the
+      choice of \(A\).
+      \begin{equation}
+        \Pi {(x , y : A)} , f(x) \equiv f(y)
       \end{equation}
       Formally speaking, \(f\) needs to be ``coherently constant''
-  \cite{krausGeneralUniversalProperty2015}.
+      \cite{krausGeneralUniversalProperty2015}, and \(B\) needs to be an
+      \(n\)-type for some finite \(n\).
+  \end{enumerate}
 \end{rm-definition}
@@
   \(\top\).
-
-  The support list will be a singleton list containing the single point in the
-  type, and the cover proof will be an index pointing at the first element in
-  the support list.
-
   \ExecuteMetaData[agda/Cardinality/Finite/ManifestEnumerable.tex]{circ-inst}
@@
 enumerability's relation to \emph{split} surjectivity.
-Deriving a surjection from \(\mathbf{Fin}\) is straightforward:
 \begin{rm-lemma}
@@
 \begin{proof}
-  \todo[inline]{Proof}
+  As with the other surjection proof (lemma~\ref{split-enum-surj}), this is
+  simply a reassociation.
 \end{proof}
 \subsection{Relation to Split Enumerability}
-As split enumerability is stronger than manifest enumerability, we can convert
-from one to the other easily.
 \begin{rm-lemma}
@@
 \end{proof}
-The main missing piece from manifest enumerability is decidable equality: by
-truncating the membership proof, we have removed the ability to distinguish
-(decidably) between members of the type.
-Indeed, resupplying decidable equality is precisely sufficient to recover split
-enumerability.
-
 \begin{rm-theorem} \label{manifest-enum-to-split}
@@
   We can do exactly that with the following lemma, called recompute:
-  Given a decision procedure for some type \(A\), and a truncation for the
+  given a decision procedure for some type \(A\), and a truncation for the
   proposition \(A\), we can discount the possibility of \(A\) being false, and
@@
   This is straightforward given decidable equality.
-
-  \ExecuteMetaData[agda/Cardinality/Finite/ManifestEnumerable.tex]{to-split-enum}
-\end{proof}
-By theorem~\ref{split-enum-to-manifest-bishop}, we also can derive that any
-manifestly enumerable type with decidable equality is Bishop finite.
+\end{proof}
 \subsection{Closure}
@@
 \section{Cardinal Finiteness} \label{cardinal}
-We now turn our attention to notions of finiteness which do not induce a linear
-order.
+For manifest enumerability, we removed the need for decidable equality: in these
+next two finiteness predicates, we remove the need for a total order on the
+underlying type.
 \begin{rm-definition}[Cardinal Finiteness]
@@
   Therefore, cardinal finiteness has the same closure properties as manifest
-  bishop finiteness.
-\end{proof}
-\subsection{Strength}
-We can eliminate from cardinal finiteness in a coherently constant way:
-effectively, anything which doesn't ``notice'' the order of the internal support
-list is usable.
-
-First, even though the cardinality of the type is hidden under the truncation,
-we can pull it out, as it is constant through permutations of the list.
-\begin{rm-theorem}
-  Given a cardinally finite type, we can derive the type's cardinality, as well
-  as a propositionally truncated proof of equivalence with \(\textbf{Fin}\)s of
-  the same cardinality.
-  \begin{equation}
-    \mathcal{C}(A) \rightarrow \sum_{(n : \mathbb{N})} , \lVert \textbf{Fin}(n) \simeq A \rVert
-  \end{equation}
-\end{rm-theorem}
-\begin{proof}
-  We eliminate from the proof of cardinal finiteness via the eliminator
-  described in proposition~\ref{elim-non-prop}.
-  We pass a function \(\text{alg}\) which preserves the first part of the pair,
-  and truncates the second.
-  In Agda:
-  \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{trunc-alg}
-
-  It remains only to show that this function is coherently constant.
-  \begin{equation}
-    \prod_{(x : \sum_{(n : \mathbb{N})} , \textbf{Fin}(n) \simeq A)} ,
-    \prod_{(y : \sum_{(m : \mathbb{N})} , \textbf{Fin}(m) \simeq A)} ,
-    \text{alg}(x) \equiv \text{alg}(y)
-  \end{equation}
-
-  We will first address the first component of the output.
-  To show that \(n\) and \(m\) are equal, we first observe that
-  \begin{equation}
-    \textbf{Fin}(n) \equiv \textbf{Fin}(m)
-  \end{equation}
-  since both are equivalent to \(A\).
-  Relying on the fact that \(\textbf{Fin}\) is injective
-  (proposition~\ref{fin-inj}), we can derive that \(n\) and \(m\) are equal.
-
-  Since the second components of the output are propositions, they are
-  definitionally equal.
-  \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{const-alg}
-\end{proof}
-
-\begin{rm-lemma} \label{fin-inj}
-  \(\textbf{Fin}\) is injective.
-\end{rm-lemma}
-\begin{proof}
-  Let \(n\) and \(m\) be natural numbers.
-  We want to derive a proof of \(n \equiv m\) from \(\textbf{Fin}(n) \equiv
-  \textbf{Fin}(m)\).
-
-  We prove by contradiction (since the property we are interested in is
-  decidable, this is valid constructively).
-  When \(n\) does not equal \(m\), there are two possible cases.
-  \begin{equation}
-    \neg (n \equiv m) \implies
-    \begin{cases}
-      m \equiv 1 + n + k, n < m  \\
-      n \equiv 1 + m + k, m < n
-    \end{cases}
-  \end{equation}
-  Take the first case, without loss of generality.
-  \todo[inline]{Finish prose for this proof}
-
-  The Agda version of this proof is in the appendix (\ref{agda-fin}).
-\end{proof}
-
-Another property which does not rely on an internal order is decidable equality.
+  Bishop finiteness.
+\end{proof}
+\subsection{Decidable Equality}
 \begin{rm-theorem} \label{cardinal-finite-discrete}
@@
 \begin{proof}
-  We first show that decidable equality is a proposition itself.
+  We will use eliminator~\ref{elim-prop-prop} from definition~\ref{prop-trunc}.
+  Manifest Bishop finiteness implies decidable equality already, so our task
+  here is to prove that decidable equality itself is a proposition.
+
   We know that if a type \(A\) is a proposition, then the decision over that
@@
   decidable equality is itself a proposition.
-
-  That is enough to pull it out of the truncation: since any Bishop-finite type
-  has decidable equality, and decidable equality is a proposition, we conclude
-  that cardinal-finite types have decidable equality.
+\end{proof}
+\subsection{Cardinality}
+\begin{rm-theorem}
+  Given a cardinally finite type, we can derive the type's cardinality, as well
+  as a propositionally truncated proof of equivalence with \(\textbf{Fin}\)s of
+  the same cardinality.
+  \begin{equation}
+    \mathcal{C}(A) \rightarrow \Sigma {(n : \mathbb{N})} , \lVert \textbf{Fin}(n) \simeq A \rVert
+  \end{equation}
+\end{rm-theorem}
+\begin{proof}
+  Our task here is to ``pull out'' the cardinality of the set from under the
+  propositional truncation.
+  In effect, we need the following function:
+  \begin{equation}
+    \left\lVert \Sigma {(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \simeq A \right) \right\rVert \rightarrow
+    \Sigma {(n : \mathbb{N})} , \left\lVert \mathbf{Fin}(n) \simeq A \right\rVert
+  \end{equation}
+
+  We will use eliminator~\ref{elim-prop-coh} from definition~\ref{prop-trunc}.
+  We eliminate with the following function:
+  \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{trunc-alg}
+
+  To show that \(\text{alg}\) is coherently constant, we first notice that the
+  second element of the output pair is propositionally truncated, meaning that
+  it is trivially equal to any other element of the same type.
+  Our task, then, simplifies to demonstrating that the first element of the
+  output pair is coherently constant.
+  \begin{equation}
+    \Pi {(x : \Sigma {(n : \mathbb{N})} , \textbf{Fin}(n) \simeq A)} ,
+    \Pi {(y : \Sigma {(m : \mathbb{N})} , \textbf{Fin}(m) \simeq A)} ,
+    n \equiv m
+  \end{equation}
+
+  Notice that \(\mathbf{Fin}(n)\) and \(\mathbf{Fin}(m)\) are both equivalent to
+  \(A\): we can join these proofs together, giving us the following:
+  \begin{equation}
+    \textbf{Fin}(n) \equiv \textbf{Fin}(m)
+  \end{equation}
+  Relying on the fact that \(\textbf{Fin}\) is injective
+  (proposition~\ref{fin-inj}), we can derive that \(n\) and \(m\) are equal.
+\end{proof}
+
+\begin{rm-lemma} \label{fin-inj}
+  \(\textbf{Fin}\) is injective.
+\end{rm-lemma}
+\begin{proof}
+  Though this proof is surprisingly complex, it is a well-known puzzle in
+  Martin-Löf type theory.
+  Our proof does not differ significantly from standard approaches, so we will
+  not detail it here.
 \end{proof}
@@
       support lists in cardinal finiteness proofs.
-    \item We show that all support lists in cardinal finiteness proofs are
-      permutations of each other.
-    \item And then we show that insertion sort is invariant under permutations.
+    \item We do this by first showing that all support lists in cardinal
+      finiteness proofs are permutations of each other, and then that insertion
+      sort is invariant under permutations.
+    \item Given our particular definition of permutations, cover proofs
+      transfers naturally between lists which are permutations of each other.
   \end{itemize} 
@@
   \begin{equation}
-    \mathit{xs} \leftrightsquigarrow \mathit{ys} = \prod_{(x : A)} , x \in \mathit{xs} \iff x \in \mathit{ys}
-  \end{equation}
+    \mathit{xs} \leftrightsquigarrow \mathit{ys} = \Pi {(x : A)} , x \in \mathit{xs} \iff x \in \mathit{ys}
+  \end{equation}
+  Seen another way, a permutation is an isomorphism of cover proofs.
 \end{rm-definition}
 We also prove some of the identities you might expect with regards to
-permutations, all in the appendix.
-In particular, we prove that they form an equivalence relation.
+permutations.
 \begin{rm-lemma}
@@
 \begin{proof}
-  We first prove that insertion sort does indeed sort its input: i.e. it returns
-  a sorted list that is a permutation of its input.
-  Then we show that any two sorted lists which are permutations of each other
-  are equal.
-\end{proof}
-\todo[inline]{Agda code for proofs}
+  Again, this proof is quite complex, so we only give a sketch here.
+  First, we prove two properties about insertion sort:
+  \begin{enumerate}
+    \item It returns a sorted list.
+    \item It a returns a list that is a permutation of its input.
+  \end{enumerate}
+  The second of these points allows us to show that \(\text{sort}(\mathit{xs})\)
+  is a permutation of \(\text{sort}(\mathit{ys})\).
+  \begin{equation}
+    \text{sort}(\mathit{xs}) \leftrightsquigarrow \mathit{xs} \leftrightsquigarrow \mathit{ys} \leftrightsquigarrow \text{sort}(\mathit{ys})
+  \end{equation}
+  Then, we show that any lists which are both sorted and permutations of each
+  other are equal.
+  Both of these conditions are true for the output of \(\text{sort}\).
+\end{proof}
 \section{Kuratowski Finiteness} \label{kuratowski}
@@
 \cite{kuratowskiNotionEnsembleFini1920}.
-\subsection{The Kuratowski Set as a Higher Inductive Type}
-We start with the Kuratowski finite sets, which are free join semilattices (or,
-equivalently, free commutative idempotent monoids).
+\begin{rm-definition}[Kuratowski-Finite Set]
+  The Kuratowski finite set is a free join semilattice (or, equivalently, a free
+  commutative idempotent monoid).
   HITs are required to define this type
-\cite{altenkirchDefinableQuotientsType2011}; however we have two possible
-candidates for the definition.
-The first is a direct translation of ``free commutative idempotent monoid'' into
-a type:
-\begin{equation}
-\begin{aligned}
-  \mathcal{K}(A) &=  \\
-    &\left\vert \; \text{singleton} : A \rightarrow \mathcal{K}(A) \right; \\
-    &\left\vert \; \cdot \cup \cdot : \mathcal{K}(A) \times \mathcal{K}(A) \rightarrow \mathcal{K}(A) \right; \\
-    &\left\vert \; \emptyset : \mathcal{K}(A) \right; \\
-    &\left\vert \; \cup \text{-assoc} : \Pi \left( \mathit{xs}, \mathit{ys}, \mathit{zs}:\mathcal{K}(A)\right) , (\mathit{xs} \cup \mathit{ys}) \cup \mathit{zs} \equiv \mathit{xs} \cup (\mathit{ys} \cup \mathit{zs}) \right; \\
-    &\left\vert \; \cup \text{-commutative} : \Pi (\mathit{xs}, \mathit{ys}: \mathcal{K}(A)) , \mathit{xs} \cup \mathit{ys} \equiv \mathit{ys} \cup \mathit{xs} \right; \\
-    &\left\vert \; \cup \text{-idempotent} : \Pi (\mathit{xs}: \mathcal{K}(A)) , \mathit{xs} \cup \mathit{xs} \equiv \mathit{xs} \right; \\
-    &\left\vert \; \cup \text{-identity} : \Pi (\mathit{xs}: \mathcal{K}(A)) , \emptyset \cup \mathit{xs} \equiv \mathit{xs} \right; \\
-    &\left\vert \; \text{trunc} : \Pi (\mathit{xs}, \mathit{ys}: \mathcal{K}(A)) , \Pi (p, q : \mathit{xs} \equiv \mathit{ys}) , p \equiv q \right;
-\end{aligned}
-\end{equation}
-\todo[inline]{Have we already described the trunc constructor?}
-
-However this proves cumbersome to work with.
-Instead, we build on the free monoid (the list), and add the required equations
-we need.
+  \cite{altenkirchDefinableQuotientsType2011}:
   \begin{equation}
@@
   \end{equation}
-Since associativity and identity (the monoid laws) are automatically satisfied
-by the list structure, we do not need to specify them as paths.
-Furthermore, it is simpler to specify the rest of the paths on the heads of the
-lists, rather than on the whole structure.
-That said, the two definitions are equivalent.
-\todo[inline]{Maybe add sketch of proof here?}
-\subsection{Elimination}
-With Agda's dependent pattern matching, we get an eliminator automatically from
-the definition of an inductive type.
-We will define two further eliminators here.
-
-\begin{equation}
-  \begin{aligned}
-    \phi &: \mathcal{K}(A) \rightarrow \text{Type} \\
-    \text{trunc}^\phi &: \pifun{x}{\mathcal{K}(A)} , \text{isSet}(\phi(x)) \\
-    []^\phi &: \phi([]) \\
-    \dblcolon^\phi &: \pifun{x}{A} , \pifun{xs}{\mathcal{K}(A)} , \pifun{p}{\phi(\mathit{xs})} , \phi(x \dblcolon \mathit{xs}) \\
-    \text{com}^\phi &: \prod_{(x , y : A)} , \pifun{xs}{\mathcal{K}(A)} , \pifun{p}{\phi(\matit{xs})} , x \dblcolon^\phi (y \dblcolon \mathit{xs}) (y \dblcolon^\phi \mathit{xs} \; p) \equiv y \dblcolon^\phi (x \dblcolon \mathit{xs}) (x \dblcolon^\phi \mathit{xs} \; p) \\
-    \text{dup}^\phi &: \pifun{x}{A} , \pifun{xs}{\mathcal{K}(A)} , \pifun{p}{\phi(\mathit{xs})} , x \dblcolon^\phi (x \dblcolon \mathit{xs}) (x \dblcolon^\phi \mathit{xs} \; p) \equiv x \dblcolon^\phi \mathi{xs} \; p
-  \end{aligned}
-\end{equation}
-
-\begin{equation} \label{kuratowski-rec}
-  \text{rec} \left\{ \begin{aligned}
-    \phi &: \text{Type} \\
-    \text{trunc}^\phi &: \text{isSet}(\phi) \\
-    []^\phi &: \phi \\
-    \dblcolon^\phi &: \pifun{x}{A} , \pifun{xs}{\mathcal{K}(A)} , \phi \\
-    \text{com}^\phi &: \prod_{(x , y : A)} , \pifun{xs}{\mathcal{K}(A)} , x \dblcolon^\phi (y \dblcolon^\phi \mathit{xs}) \equiv y \dblcolon^\phi (x \dblcolon^\phi \mathit{xs}) \\
-    \text{dup}^\phi &: \pifun{x}{A} , \pifun{xs}{\mathcal{K}(A)} , x \dblcolon^\phi (x \dblcolon^\phi \mathit{xs}) \equiv x \dblcolon^\phi \mathit{xs}
-  \end{aligned} \right.
-\end{equation}
-
-\todo[inline]{Expand on these eliminators}
-
-\subsection{Membership}
-We can define membership of a Kuratowski set using these eliminators.
-First we define the \(\Diamond\) type, described in the appendix.
-\begin{equation}
-  \Diamond P = \left\{ \begin{aligned}
-    \phi &= \sum_{(x : \text{Type})} , \text{isProp}(x) \\
-    \text{trunc}^\phi &= \text{isSetHProp} \\
-    []^\phi &= \bot , \text{isProp}\bot \\
-    x \dblcolon^\phi \mathit{xs} &= \lVert P(x) \uplus \text{fst}(\mathit{xs}) \rVert , \text{squash} \\
-    \text{com}^\phi &= \dots \\
-    \text{dup}^\phi &= \dots
-  \end{aligned} \right.
-\end{equation}
-We can define membership then in terms of this.
-\begin{rm-definition}[Kuratowski Membership]
-  \begin{equation}
-    x \in \mathit{xs} = \Diamond (\equiv x) \mathit{xs}
-  \end{equation}
-\end{rm-definition}
-
-From this we can define Kuratowski finiteness.
+  The \(\text{com}\) and \(\text{dup}\) constructors effectively add
+  commutativity and idempotency to the free monoid (the list), which is made by
+  the first two constructors.
+  The last constructor makes \(\mathcal{K}(A)\) a set.
+
+  To eliminate from \(\mathcal{K}(A)\), we have to provide equations for each of
+  the point constructors which obey the equations of the path constructors.
+  For \(\text{com}\) and \(\text{dup}\), this means ensuring that the fold is
+  commutative and idempotent, whereas \(\text{trunc}\) means we can only
+  eliminate into sets.
+\end{rm-definition}
+
+Other representations of \(\mathcal{K}\) \cite{fruminFiniteSetsHomotopy2018}
+are more explicit constructions of the free join semilattice (i.e. there is a
+point constructors for union instead of cons, and then path constructors for
+the associativity and identity laws), but we have found this representation
+easier to work with.
+Nonetheless, the alternative representation is included in our formalisation,
+and proven equivalent to the representation here.
+\begin{rm-definition}[Membership of \mathcal{K}]
+  First, we need to provide equations for the two point constructors.
+  \begin{equation}
+    \begin{alignat}{2}
+      x &\in& [] &= \bot \\
+      x &\in& y \dblcolon \mathit{xs} &= \lVert (x \equiv y) + (x \in \mathit{xs}) \rVert
+    \end{alignat}
+  \end{equation}
+  The \(\text{com}\) and \(\text{dup}\) constructors are handled by proving that
+  the truncated form of \(+\) is itself commutative and idempotent.
+  The type of propositions is itself a set, satisfying the \(\text{trunc}\)
+  constructor.
+\end{rm-definition}
 \begin{rm-definition}[Kuratowski Finiteness]
@@
   \begin{equation}
-    \mathcal{K}^{f}(A) = \sum_{(\text{support} : \mathcal{K}(A))} , \pifun{x}{A} , x \in \text{support}
+    \mathcal{K}^{f}(A) = \Sigma {(\mathit{xs} : \mathcal{K}(A))} , \Pi (x : A) , x \in \mathit{xs}
   \end{equation}
@@
   We prove by way of an isomorphism.
-  In the first direction (from \(\mathcal{K}\) to \(\mathcal{E}\)), we use the
-  eliminator in \ref{kuratowski-rec}.
-  Because we are eliminating into a proposition, we need not prove that the
-  function obeys the path constructors.
-  The remaining cases to deal with involve converting the Kuratowski set to a
-  list, and the cover proof to its equivalent cover proof on lists.
-  \ExecuteMetaData[agda/Cardinality/Finite/Kuratowski.tex]{to-enum}
-
-  To go the other direction, we first need to prove that a proof of Kuratowski
-  finiteness is a proposition.
-  \todo[inline]{Explain rest of this proof}
+  In the first direction (from \(\mathcal{K}\) to \(\mathcal{E}\)), because we
+  are eliminating into a proposition, we need only deal with the point
+  constructors.
+  For these, we convert the \(\mathcal{K}\) cons to its list counterpart, and
+  similarly for the nil constructor.
+
+  The other direction is proven in \cite{fruminFiniteSetsHomotopy2018}, so we
+  will not describe it here.
 \end{proof}
@@
 \begin{proof}
-  \todo[inline]{Proof}
-\end{proof}
-\begin{rm-lemma}
-  Kuratowski finiteness is equivalent to a truncated surjection from some prefix
-  of the natural numbers.
-  \begin{equation}
-    \mathcal{K}^f(A) \simeq \left\Vert \sum_{(n : \mathbb{N})} , \left(  \mathbf{Fin}(n) \twoheadrightarrow A \right) \right\Vert
-  \end{equation}
-\end{rm-lemma}
-\begin{proof}
-  \todo[inline]{Prose for proof}
+  We first convert to a propositionally truncated proof of manifest
+  enumerability.
+  From here, we can convert to manifest Bishop finiteness under the truncation
+  with our decidable equality proof.
 \end{proof}
@@
 \begin{proof}
-  \todo[inline]{Proof}
+  This follows from the \(\sum\) closure proof on manifestly enumerable types.
 \end{proof}
 \section{Species and Containers} \label{species-and-containers}
-\subsection{Containers For Finiteness}
-The definition for the container-based finiteness predicates are almost
-identical to their list-based counterparts.
-\ExecuteMetaData[agda/Cardinality/Container/Finite.tex]{split-enum}
-One advantage that they have, however, is that conversion back and forth between
-the relation-based finiteness predicates is very straightforward.
-This allows us to prove the following.
-\begin{rm-theorem}
-  Manifest bishop finiteness is equivalent to an equivalence with some \(\mathbf{Fin}\).
-  \begin{equation}
-    \mathcal{B}(A) \simeq \sum_{(n : \mathbb{N})} , \mathbf{Fin}(n) \simeq A
-  \end{equation}
-\end{rm-theorem}
-\begin{proof}
-  This proof is the composition of two equivalences.
-  First, we must show that the list-based manifest Bishop finiteness is
-  equivalent to the container-based definition, and then we must show that the
-  container-based definition of manifest Bishop finiteness is equivalent to the
-  equivalence.
-  \todo[inline]{proof}
-\end{proof}
-
-\begin{rm-lemma}
-  Manifest enumerability is equivalent to a surjection from some \(\mathbf{Fin}\).
-  \begin{equation}
-    \mathcal{E}(A) \simeq \sum_{(n : \mathbb{N})} , \left(  \mathbf{Fin}(n) \twoheadrightarrow A \right)
-  \end{equation}
-\end{rm-lemma}
-\begin{proof}
-\end{proof}
-
-\begin{rm-lemma}
-  Split enumerability is equivalent to a split surjection from some \(\mathbf{Fin}\).
-  \begin{equation}
-    \mathcal{E}!(A) \simeq \sum_{(n : \mathbb{N})} , \left(  \mathbf{Fin}(n) \twoheadrightarrow ! A \right)
-  \end{equation}
-\end{rm-lemma}
-\begin{proof}
-\end{proof}
-
+\todo[inline]{Figure out this section}
 \section{Infinite Cardinalities} \label{infinite-cardinalities}
@@
   \begin{equation}
-    \mathcal{E}!(A) = \sum_{(\mathit{xs} : \mathbf{Stream}(A))} , \prod_{(x : A)} , x \in \mathit{xs}
+    \mathcal{E}!(A) = \Sigma {(\mathit{xs} : \mathbf{Stream}(A))} , \Pi {(x : A)} , x \in \mathit{xs}
   \end{equation}
@@
   \begin{equation}
-    \mathcal{Y} : \prod_{(x : A)} , \mathcal{E}!(U(x))
+    \mathcal{Y} : \Pi {(x : A)} , \mathcal{E}!(U(x))
   \end{equation}
@@
   \begin{equation}
-    \mathcal{E}!\left(\sum_{(x : A)} , U(x)\right)
+    \mathcal{E}!\left(\Sigma {(x : A)} , U(x)\right)
   \end{equation}
@@
   \begin{equation}
-    \left( \prod_{(x : A)} , \mathbf{Dec}(P(x)) \right) \rightarrow \mathbf{Dec} \left( \sum_{(x : A)} , P(x) \right)
+    \left( \Pi {(x : A)} , \mathbf{Dec}(P(x)) \right) \rightarrow \mathbf{Dec} \left( \Sigma {(x : A)} , P(x) \right)
   \end{equation}
@@
   \begin{equation}
-    \left( \prod_{(x : A)} , \mathbf{Dec}(P(x)) \right) \rightarrow \mathbf{Dec} \left( \prod_{(x : A)} , P(x) \right)
+    \left( \Pi {(x : A)} , \mathbf{Dec}(P(x)) \right) \rightarrow \mathbf{Dec} \left( \Pi {(x : A)} , P(x) \right)
   \end{equation}

### 2020-01-06 paper.tex
much shorter

@@
 \end{abstract}
-% \section{Introduction}
-% Proofs in a constructive setting are more substantial things than their
-% classical counterparts.
-% Rather than just evidence for some fact, they are objects in their own right,
-% which we can manipulate and extract information from.
-
-% Sometimes, the extra detail can reveal interesting subtleties: fissures which
-% divide homogeneous-seeming definitions into a variety of related concepts.
-% Finiteness is a prime example of this phenomenon: in classical set theory, a
-% finite set is a finite set, and there's not much else that we can say.
-% In constructive type theory, we start the bidding at no fewer than \emph{four}
-% different predicates representing finiteness
-% \cite{spiwackConstructivelyFinite2010}, upped to eight in
-% \cite{firsovVariationsNoetherianness2016a}: all of which differ in significant
-% and interesting ways.
-
-% Other times, the extra baggage is unwanted.
-% Maybe we want to enforce that a particular piece of information should be
-% hidden, or treated as irrelevant: ``yes, I have proven that there is a
-% particular number with this property, but I don't want to reveal \emph{which}
-% number.''
-
-% This is where Homotopy Type Theory \cite{hottbook} can step in.
-% Through primarily a generalisation of equality, we gain access to a number of
-% exotic types which can, among other things, allow us to perform this ``hiding''.
-% In the context of finiteness, the act of hiding is worth studying in its own
-% right, and further delineates several new finiteness predicates which collapse
-% even in standard Martin-Löf type theory
-% \cite{martin-lofIntuitionisticTypeTheory1980}.
-
-% The other headline feature of HoTT's generalised equality is univalence: put
-% simply, it allows us to treat isomorphic types as equal, giving us in the
-% constructive world access to a technique that is commonplace classically.
-% It turns out that finite types have a lot to say about isomorphisms, and we will
-% rely heavily on this aspect of HoTT to work with them.
-% \subsection{Contributions}
-% In this work, we will explore finite types in Cubical Type Theory
-% \cite{cohenCubicalTypeTheory2016}, and expose their relationship to infinite
-% types, species, and demonstrate their practical uses for proofs in dependently
-% typed programming languages.
-% \subsubsection{Strong Finiteness Predicates}
-% We will first explore the ``strong'' notions of finiteness (i.e. those at least
-% as strong as Kuratowski finiteness \cite{kuratowskiNotionEnsembleFini1920}),
-% with a special focus on cardinal finiteness (section~\ref{cardinal}), and
-% manifest enumerability (section~\ref{manifest-enumerability}), which is new, to
-% our knowledge.
-
-% \input{figures/finite-classification}
-% Figure~\ref{finite-classification} organises the predicates according to
-% their ``strength''; i.e. how much information they provide about a conforming
-% type.
-% For instance, a proof that some type \(A\) is manifestly Bishop finite (the
-% strongest of the notions, explored in section~\ref{manifest-bishop}) also tells
-% us that \(A\) is discrete (has decidable equality), and gives us a linear order
-% on the type.
-% A type that is Kuratowski finite (section~\ref{kuratowski}) has no such extra
-% features: indeed, we will see examples of Kuratowski finite types which are not
-% even sets, never mind discrete ones.
-
-% We will go through each of the predicates, proving how to weaken each (i.e. we
-% will provide a proof that every cardinally finite type is Kuratowski finite),
-% and how to strengthen them, given the required property.
-% In terms of figure~\ref{finite-classification}, this amounts to providing proofs
-% for each arrow.
-
-% We will---through the use of containers
-% \cite{abbottContainersConstructingStrictly2005}---formally prove the equivalence
-% these predicates have with the usual function relations i.e. we will show that a
-% proof of manifest enumerability is precisely equivalent to a surjection from a
-% finite prefix of the natural numbers.
-
-% For each predicate, we will also prove its closure over sums and products in
-% both dependent and non-dependent forms, if such a closure exists.
-% This will culminate in our main result for this section: the formal proof that
-% decidable Kuratowski finite sets form a topos. \todo[noline]{reference for this}
-% More specifically, we show that cardinal finite sets form a topos, that
-% decidable Kuratowski finite sets are equivalent in strength to cardinal finite
-% sets, and carry the proof over.
-% This result relies on proofs on each of the other finiteness predicates.
-% \subsubsection{Species}
-% \todo[inline]{Redo this next paragraph}
-% In section~\ref{species-and-containers}, we will redefine our finiteness
-% predicates using \emph{containers}, and prove that the new definitions are
-% equivalent to the old.
-% Using this, we will prove much stronger relationships between our finiteness
-% predicates and relation-based definitions of finiteness.
-% Finally, We will show how the species-container duality falls out naturally from
-% these new definitions.
-
-% \subsubsection{Infinite Types}
-% In section~\ref{infinite-cardinalities}, we will extend our study of finite
-% types to infinite but countable types.
-% We will see that the finiteness predicates are mirrored with countable
-% counterparts, and we will prove closure under the Kleene star and plus.
-
-% \subsubsection{Practical Uses of Finiteness}
-% Proofs of finiteness have well-known practical applications in
-% constructive mathematics \cite{firsovDependentlyTypedProgramming2015}.
-% In section~\ref{practical}, we build a library which exploits these uses in
-% Cubical Agda \cite{vezzosiCubicalAgdaDependently2019}, allowing automation of
-% complex proofs over finite types.
-% We frame this in terms of the principle of omniscience for finite types.
-% Thanks to the flexibility afforded to us by Cubical Type Theory, we are able to
-% go further than the usual examples of this kind of proof automation: as well as
-% proving properties about functions, we can synthesise functions whole-cloth from
-% their desired properties.
-% Through the unified interface for finite and countable types, we can
-% reuse the automation machinery for \emph{partial} proof search over infinite
-% search spaces.
-% Along the way, we extend the work in \cite{allaisGenericLevelPolymorphic2019} to
-% prove isomorphisms between the curried and uncurried forms of \(n\)-ary
-% dependent functions.
+\section{Introduction}
+In this work, we will explore finite types in Cubical Type Theory
+\cite{cohenCubicalTypeTheory2016}, and expose their relationship to infinite
+types, species, and demonstrate their practical uses for proofs in dependently
+typed programming languages.
+\subsubsection{Strong Finiteness Predicates}
+We will first explore the ``strong'' notions of finiteness (i.e. those at least
+as strong as Kuratowski finiteness \cite{kuratowskiNotionEnsembleFini1920}),
+with a special focus on cardinal finiteness (section~\ref{cardinal}), and
+manifest enumerability (section~\ref{manifest-enumerability}), which is new, to
+our knowledge.
+
+\input{figures/finite-classification}
+Figure~\ref{finite-classification} organises the predicates according to
+their ``strength''; i.e. how much information they provide about a conforming
+type.
+For instance, a proof that some type \(A\) is manifestly Bishop finite (the
+strongest of the notions, explored in section~\ref{manifest-bishop}) also tells
+us that \(A\) is discrete (has decidable equality), and gives us a linear order
+on the type.
+A type that is Kuratowski finite (section~\ref{kuratowski}) has no such extra
+features: indeed, we will see examples of Kuratowski finite types which are not
+even sets, never mind discrete ones.
+
+We will go through each of the predicates, proving how to weaken each (i.e. we
+will provide a proof that every cardinally finite type is Kuratowski finite),
+and how to strengthen them, given the required property.
+In terms of figure~\ref{finite-classification}, this amounts to providing proofs
+for each arrow.
+
+We will---through the use of containers
+\cite{abbottContainersConstructingStrictly2005}---formally prove the equivalence
+these predicates have with the usual function relations i.e. we will show that a
+proof of manifest enumerability is precisely equivalent to a surjection from a
+finite prefix of the natural numbers.
+
+For each predicate, we will also prove its closure over sums and products in
+both dependent and non-dependent forms, if such a closure exists.
+This will culminate in our main result for this section: the formal proof that
+decidable Kuratowski finite sets form a topos.
+More specifically, we show that cardinal finite sets form a topos, that
+decidable Kuratowski finite sets are equivalent in strength to cardinal finite
+sets, and carry the proof over.
+This result relies on proofs on each of the other finiteness predicates.
+\subsubsection{Species}
+\todo[inline]{Redo this next paragraph}
+
+\subsubsection{Infinite Types}
+In section~\ref{infinite-cardinalities}, we will extend our study of finite
+types to infinite but countable types.
+We will see that the finiteness predicates are mirrored with countable
+counterparts, and we will prove closure under the Kleene star and plus.
+
+\subsubsection{Practical Uses of Finiteness}
+Proofs of finiteness have well-known practical applications in
+constructive mathematics \cite{firsovDependentlyTypedProgramming2015}.
+In section~\ref{practical}, we build a library which exploits these uses in
+Cubical Agda \cite{vezzosiCubicalAgdaDependently2019}, allowing automation of
+complex proofs over finite types.
+We frame this in terms of the principle of omniscience for finite types.
+Thanks to the flexibility afforded to us by Cubical Type Theory, we are able to
+go further than the usual examples of this kind of proof automation: as well as
+proving properties about functions, we can synthesise functions whole-cloth from
+their desired properties.
+Through the unified interface for finite and countable types, we can
+reuse the automation machinery for \emph{partial} proof search over infinite
+search spaces.
+Along the way, we extend the work in \cite{allaisGenericLevelPolymorphic2019} to
+prove isomorphisms between the curried and uncurried forms of \(n\)-ary
+dependent functions.
 \section{Split Enumerability} \label{split-enumerable}
@@
 \begin{rm-theorem}
-  Manifest bishop finiteness is closed over dependent functions (\(\prod\)-types).
-
-  Formally, given a type \(A\) and a type family \(U\) on \(A\), when \(A\) is
-  manifestly Bishop finite:
-  \begin{equation}
-    \mathcal{B}\left( A \right)
-  \end{equation}
-  And \(U\) is manifestly Bishop finite over all points of \(A\):
-  \begin{equation}
-    \Pi {(x : A)} , \mathcal{B}\left( U(x) \right)
-  \end{equation}
-  Then we have the following:
-  \begin{equation}
+  Manifest bishop finiteness is closed over dependent functions
+  (\(\prod\)-types).
+  \begin{equation}
+    \frac{
+      \mathcal{B}(A) \; \; \; \Pi {(x : A)} , \mathcal{B}\left( U(x) \right)
+    }{
       \mathcal{B}\left(\Pi {(x : A)} , U(x)\right)
+    }
   \end{equation}
@@
   goal.
-
-  \ExecuteMetaData[agda/Cardinality/Finite/ManifestBishop.tex]{pi-clos}
 \end{proof}
@@
 \end{rm-definition}
-\subsection{Higher Homotopy Levels}
+
 By hiding the position, we have essentially removed the ``decidable'' component
@@
   \(\top\).
-  \ExecuteMetaData[agda/Cardinality/Finite/ManifestEnumerable.tex]{circ-inst}
 \end{proof}
@@
 
-  For the cover proof, we need a way to get the contents out of a truncated
-  value.
-  We can do exactly that with the following lemma, called recompute:
-  given a decision procedure for some type \(A\), and a truncation for the
-  proposition \(A\), we can discount the possibility of \(A\) being false, and
-  therefore extract the true decision.
-
-  For the cover proof, our obligation now becomes constructing a decision for
-  membership of the support list.
-  This is straightforward given decidable equality.
+  For the cover proof, we first need the following function which searches a
+  list for a particular element (given decidable equality on \(A\)).
+  \begin{equation}
+    \decin \; : \Pi(x : A) , \Pi(\mathit{xs} : \mathbf{List}(A)) , \mathbf{Dec}(x \in \mathit{xs})
+  \end{equation}
+  Where \(\mathbf{Dec}(A)\) is a decision on some type \(A\).
+
+  We then need to convert a value of type \(\mathbf{Dec}(x \in \mathit{xs})\) to
+  \(x \in \mathit{xs}\).
+  We use the following to do that:
+  \begin{equation}
+    \text{recompute} : \mathbf{Dec}(A) \rightarrow \lVert A \rVert \rightarrow A
+  \end{equation}
+  A propositionally truncated value can be used to \emph{refute} its negation.
 \end{proof}
@@
   \end{equation}
-  Relying on the fact that \(\textbf{Fin}\) is injective
-  (proposition~\ref{fin-inj}), we can derive that \(n\) and \(m\) are equal.
-\end{proof}
-
-\begin{rm-lemma} \label{fin-inj}
-  \(\textbf{Fin}\) is injective.
-\end{rm-lemma}
-\begin{proof}
-  Though this proof is surprisingly complex, it is a well-known puzzle in
+  All that remains now is to prove that \(\mathbf{Fin}\) is injective.
+  Though the proof is surprisingly complex, it is a well-known puzzle in
   Martin-Löf type theory.
@@
 \begin{proof}
-  Let \(\mathcal{X}\) be the proof of split countability on \(A\), and
-  \(\mathcal{Y}\) be a function of the following type:
-  \begin{equation}
-    \mathcal{Y} : \Pi {(x : A)} , \mathcal{E}!(U(x))
-  \end{equation}
-  Where \(U\) is a type family on \(A\).
-  Our task is to provide the following:
-  \begin{equation}
-    \mathcal{E}!\left(\Sigma {(x : A)} , U(x)\right)
-  \end{equation}
-
   This final proof consists of the support stream, and the proof that the
@@
   We then concatenate these streams into the final support stream.
-
-  Here is the algorithm in Agda:
-  \ExecuteMetaData[agda/Cardinality/Container/Infinite.tex]{conv}
-  \todo[inline]{Finish prose for this proof}
 \end{proof}
@@
 \end{proof}
-\subsubsection{Kleene Star and Plus}
+\subsubsection{Kleene Star}
 While we lose some closures with the inclusion of infinite types, we gain some
 others.
-In particular, we have the Kleene star and plus.
-These allow us to introduce recursion into countable types, by way of lists.
-\begin{rm-definition}[Kleene Lists]
-  It is often useful to have two mutually-defined list types (one for
-  possibly-empty lists and one for nonempty lists).
-  Such a definition closely mirrors the Kleene star and plus.
-  \begin{equation}
-    \begin{aligned}
-      A^+ &= \\
-          &\left\vert \; \text{head} : A^+ \rightarrow A \right; \\
-          &\left\vert \; \text{tail} : A^+ \rightarrow A^\star \right; \\
-      A^\star &= \\
-          &\left\vert \; [] : A^\star \right; \\
-          &\left\vert \; [\cdot] : A^+ \rightarrow A^\star \right;
-    \end{aligned}
-  \end{equation}
-\end{rm-definition}
-Our definition of split countability is closed under these types in particular.
+In particular, we have the Kleene star.
+This means, in effect, that we have closure under lists.
 \begin{rm-theorem}
-  Split countability is closed under Kleene star and plus.
-\end{rm-theorem}
-\begin{proof}
-  \tot[inline]{Proof}
+  Split countability is closed under Kleene star.
+  \begin{equation}
+    \mathcal{E}!(A) \rightarrow \mathcal{E}!(\mathbf{List}(A))
+  \end{equation}
+\end{rm-theorem}
+\begin{proof}
+  As with the proof of closure under \(\sum\), our main task here is to figure
+  out a way to arrange the indices such that 
 \end{proof}
@@
 \section{Practical Uses} \label{practical}
-\todo[inline]{%
-\begin{itemize}
-  \item Synthesize functions according to specifications.
-  \item Nonexhaustive search for the infinite types.
-  \item Lift quantifiers over types other than bishop (i.e. K). Show that
-    quantifiers have to be different (i.e. existence has to be unique).
-  \item Big operators\cite{gustafssonCountingTypeIsomorphisms}
-    \cite{gustafssonFoldableContainersDependent}
-\end{itemize}
-}
 The theory of finite types in constructive mathematics, and in HoTT in
@@
 In particular, \emph{finite} types are omniscient.
-Omniscience is defined in Agda like so:
-\ExecuteMetaData[agda/Relation/Nullary/Omniscience.tex]{omniscient}
 
@@
   \end{equation}
-  \ExecuteMetaData[agda/Relation/Nullary/Omniscience.tex]{exhaustible}
-\end{rm-definition}
+\end{rm-definition}
+
+All of the finiteness predicates we have seen justify exhaustibility.
+We will only prove it once, then, for the weakest:
+\begin{rm-theorem}
+  Kuratowski-finite types are exhaustible.
+\end{rm-theorem}
+\begin{proof}
+  \todo[inline]{Proof}
+\end{proof}
+
 Omniscience is stronger than exhaustibility, as we can derive the latter from
@@
 \begin{proof}
+  For decidable propositions, we know the following:
+  \begin{equation} \label{lem-conversion}
+    \Pi (x : A) , P(x) \leftrightarrow \neg \Sigma (x : A) , \neg P(x) 
+  \end{equation}
+  To derive exhaustibility from omniscience, then, we run the predicate in its
+  negated form, and then subsequently negate the result.
+  The resulting decision over \(\neg \Sigma(x : A) , \neg P(x)\) can be
+  converted into \(\Pi (x : A) , P(x)\).
+\end{proof}
+We cannot derive, however, that any exhaustible type is omniscient, as we do not
+have the inverse of equation~\ref{lem-conversion}:
+\begin{equation} \label{choice-conversion}
+  \Sigma (x : A) , P(x) \leftrightarrow \neg \Pi (x : A) , \neg P(x)
+\end{equation}
+Such an equation would allow us to pick a representative element from any type,
+which is therefore non-constructive.
+In a sense, equation~\ref{lem-conversion} requires a form of LEM on the
+proposition (i.e. requires it to be decidable), whereas
+equation~\ref{choice-conversion} requires a form of choice.
+Those finiteness predicates which are ordered do in fact give us this form of
+choice, so the conversion is valid.
+As such, all of the ordered finiteness predicates imply omniscience.
+Again, we will prove it only for the weakest.
+\begin{rm-theorem}
+  Manifest enumerable types are omniscient.
+\end{rm-theorem}
+\begin{proof}
   \todo[inline]{Proof}
-  \ExecuteMetaData[agda/Relation/Nullary/Omniscience.tex]{weaken}
-  \todo[noline]{stability in appendix}
-\end{proof}
-We cannot derive, however, that any exhaustible type is omniscient: to do so
-would require us to choose a representative from an arbitrary type, which is not
-possible constructively.
-
-\subsubsection{Split Enumerability}
-We now show that split enumerable types are omniscient.
+\end{proof}
+Finally, we do have a form of omniscience for prop-valued predicates, as they do
+not care about the chosen representative.
 \begin{rm-theorem}
-  Any split enumerable type is omniscient.
-\end{rm-theorem}
-\begin{proof}
-  Let \(E_A\) be a proof of split enumerability on a type \(A\), and \(P\) be a
-  decidable predicate on the type \(A\).
-  First, we lift the decidable predicate onto one on lists, using the
-  \(\Diamond\) type described earlier.
-  Then, we run the predicate on the support list of \(E_A\).
-
-  In the case that it is true for any member of the list, we return that element
-  along with its proof.
-
-  If the predicate fails for every member of the list, we say that the whole
-  test has failed.
-  To support this conclusion, we reason that if a predicate is false for every
-  item in a list \(\mathit{xs}\), and some element \(x\) is in that list, then
-  the predicate must be false for that item.
-
-  In Agda:
-  \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{exists-quant}
-\end{proof}
-\begin{rm-theorem}
-  Any manifestly enumerable type is omniscient.
+  Kuratowski finite types are omniscient about prop-valued predicates.
 \end{rm-theorem}
@@
 \end{proof}
-\begin{rm-theorem}
-  Any cardinal finite type is exhaustible.
-\end{rm-theorem}
-\begin{proof}
-  \todo[inline]{Proof}
-\end{proof}
-\begin{rm-theorem}
-  Cardinally finite types are omniscient about prop-valued predicates.
-\end{rm-theorem}
-\begin{rm-theorem}
-  Kuratowski finite types have the same omniscience properties as cardinal
-  finite types.
-\end{rm-theorem}
-
 \subsection{Synthesising Pattern-Matching Proofs}
@@
 
-The alternative is to derive the things we need from \(\Enum\) somehow.
+The alternative is to derive the things we need from \(\Enum!\) somehow.
 First, then, we need an instance for \(\AgdaDatatype{Pauli}\):
 \ExecuteMetaData[agda/Data/Pauli.tex]{inst}
-From here we can already derive decidable equality:
-\ExecuteMetaData[agda/Data/Pauli.tex]{good-eq}
-
-Next, we want to prove some things about the group operation itself:
-\ExecuteMetaData[agda/Data/Pauli.tex]{group-op}
-The first property is the following:
-\ExecuteMetaData[agda/Data/Pauli.tex]{cancel}
-Tough the proof is simple enough to write by hand, it's a good place to start
-for demonstrating the automation technique.
-
-First, we will witness the fact that if a property is true for every element in
-a finite type's support list, it is true for every element in the type (and vice
-versa).
-\ExecuteMetaData[agda/Cardinality/Finite/Search.tex]{all-comb}
-We can also lift a decidable property on elements of the type to a decidable
-property on \emph{lists} of elements of the type.
-\ExecuteMetaData[agda/Data/List/Relation/Unary.tex]{dec-all}
-Combining these, we can turn a decidable predicate on a set into a decidable
-\emph{universal property} on the set.
-\ExecuteMetaData[agda/Cardinality/Finite/Search.tex]{lift-forall}
-
-To use this as a proof, we can run the decision procedure at type checking,
-using the following:
+From here we can already derive decidable equality, a function which requires 16
+cases if implemented manually.
+
+For proof search, the procedure is a well-known one in Agda
+\cite{devrieseBrightSideType2011}: we ask for the result of a decision procedure
+as an \emph{instance argument}, which will demand computation during
+typechecking.
 \ExecuteMetaData[agda/Cardinality/Finite/Search.tex]{auto}
-This function uses the \(\AgdaDatatype{True}\) type, defined like so:
-\ExecuteMetaData[agda/Relation/Nullary.tex]{true-def}
-For a decision \(d\), \(\AgdaDatatype{True}\; d\) will resolve to \(\top\) when
-the decision is \(\AgdaInductiveConstructor{yes}\) (i.e. when the proposition is
-true), or \(\bot\) if the decision is \(\AgdaInductiveConstructor{no}\).
-Then, we require an instance of the resulting type to be in scope: this
-requirement can only be satisfied for \(\top\), thereby providing a proof that
-the decision was indeed \(\AgdaInductiveConstructor{yes}\).
-
 And finally we can prove the property we wanted to on \(\AgdaDatatype{Pauli}\)
@@
 Running decision procedures during typechecking isn't the only use of instances:
-we can also use them similarly to how typeclasses are used in Haskell, to
-automate simple derivation procedures.
-For instance, in Haskell we can define a class for generating textual
-representations of data:
-\ExecuteMetaData[haskell/Introduction.tex]{show-cls}
-With instances that look like this:
-\ExecuteMetaData[haskell/Introduction.tex]{show-inst}
-We can have instances \emph{depend} on other instances, like the following:
-\ExecuteMetaData[haskell/Introduction.tex]{show-deriv}
-In this way, it gets rid of a tremendous about of boilerplate for more complex
-types, automatically picking out the correct instance when it can.
-
-In Agda, as in Haskell, the problem of overlapping instances often presents
-itself when more complex typeclass-based machinery is used.
-While there are ways to deal with overlapping instances, it is usually better to
-avoid them altogether, which is precisely what we do in our approach.
-We provide instances for the semiring-based combinators we defined in
-\todo[noline]{reference here}
-, and require the user to provide an instance for their own type.
-We don't, for example, provide an instance that itself searches for surjections
-from other instances: this instance is not uniquely determined.
-
-After all of that, the change to our \(\AgdaFunction{∀↯}\) is simple: we just
-change the first parameter from explicit to an instance.
+we will also use them to automate the finding of a proof of finiteness for a
+given type.
+The change to our \(\AgdaFunction{∀↯}\) is simple: we just change the first
+parameter from explicit to an instance.
 \ExecuteMetaData[agda/Cardinality/Finite/Search.tex]{auto-inst}
@@
 \ExecuteMetaData[agda/Data/Pauli.tex]{cancel-auto}
-But the real benefit is now we can automate proofs over \emph{tuples}, allowing
-us to prove, for instance, commutativity.
+The real benefit is that we have provided instances for things like tuples,
+meaning we can now prove \(n\)-ary properties.
 \ExecuteMetaData[agda/Data/Pauli.tex]{comm-uncurried}
@@
 \end{itemize}
-A full explanation of our implementation is beyond the scope of this work, but
-we will mention the key parts here.
-First, we define a function arrow generic over the application method:
-\ExecuteMetaData[agda/Data/Product/N-ary.tex]{fun-arr}
-Then, we prove that it is isomorphic to the normal function arrow:
-\ExecuteMetaData[agda/Data/Product/N-ary.tex]{fun-iso}
-This step will allow us to write the curry-uncurry proofs once, and then extend
-them to the three different argument forms without difficulty.
-
-We do the same with dependent function types:
-\ExecuteMetaData[agda/Data/Product/N-ary.tex]{pi-iso}
-
-Next, we need to make our machinery for multiple arguments.
-Here we will define a generic tuple, indexed by some \(\Nat\).
-As observed in \cite{allaisGenericLevelPolymorphic2019}, it's important to
-\emph{not} implement this as an inductively defined vector, e.g.
-\ExecuteMetaData[agda/Data/Vec.tex]{def}
-As this will not give us the correct \(\eta\)-equality we need for unification.
-\todo[noline]{Stick reference to iterated Vec definition}
-Once we have a unification-friendly vector type, we can use it to implement our
-generic (and level-polymorphic) tuples.
-\ExecuteMetaData[agda/Data/Product/N-ary.tex]{tuple}
-We decided against \(\top\)-terminated tuples, as the actual contents of the
-tuple type are exposed in the interface.
-
-Finally, we can prove isomorphisms between the curried and uncurried versions of
-functions:
-\ExecuteMetaData[agda/Data/Product/N-ary.tex]{curry-n-iso}
-And dependent functions:
-\ExecuteMetaData[agda/Data/Product/N-ary.tex]{curry-n-pi-iso}
-With these combinators, we can implement the search functions in an
-arity-generic way:
+A full explanation of our implementation is beyond the scope of this work, so we
+only present the finished interface.
 \ExecuteMetaData[agda/Cardinality/Finite/Search.tex]{all-nary}
-And automate away our proofs:
+Which is used like so:
 \ExecuteMetaData[agda/Data/Pauli.tex]{prf-curried}

### 2020-01-07 paper.tex
more!

@@
 \input{preamble}
-\title{Finiteness, Cardinality, and Combinatorics in Homotopy Type Theory}
+\title{Finiteness and Cardinality in Homotopy Type Theory}
 \author{Donnacha Oisín Kidney}
@@
   We explore five notions of finiteness in Homotopy Type Theory \cite{hottbook}.
-  We prove closure properties about all of these notions, culminating in a proof
-  that decidable Kuratowski-finite sets form a topos.
-
-  We extend the definitions to include infinite types, developing a similar
-  classification of countable types.
-
-  We use the definition of finiteness to formalise \emph{species}, in much
-  the same way as in \cite{yorgeyCombinatorialSpeciesLabelled2014}.
-  A clear duality with containers
-  \cite{abbottContainersConstructingStrictly2005} falls out naturally from our
-  definition.
+  We will classify these notions, and prove their closure properties,
+  culminating in a proof that decidable Kuratowski-finite sets
+  \cite{kuratowskiNotionEnsembleFini1920} form a topos.
+
+  We then develop a parallel classification for the countably infinite types, as
+  well as a proof of the countability of \(A^\star\) for a countable type \(A\).
 
@@
 for each arrow.
+Our main proof in this regard is that we can derive manifest Bishop finiteness
+from Kuratowski finiteness plus a total order.
 
@@
 This result relies on proofs on each of the other finiteness predicates.
-\subsubsection{Species}
-\todo[inline]{Redo this next paragraph}
 
@@
 types to infinite but countable types.
-We will see that the finiteness predicates are mirrored with countable
-counterparts, and we will prove closure under the Kleene star and plus.
-
-\subsubsection{Practical Uses of Finiteness}
-Proofs of finiteness have well-known practical applications in
-constructive mathematics \cite{firsovDependentlyTypedProgramming2015}.
-In section~\ref{practical}, we build a library which exploits these uses in
-Cubical Agda \cite{vezzosiCubicalAgdaDependently2019}, allowing automation of
-complex proofs over finite types.
-We frame this in terms of the principle of omniscience for finite types.
-Thanks to the flexibility afforded to us by Cubical Type Theory, we are able to
-go further than the usual examples of this kind of proof automation: as well as
-proving properties about functions, we can synthesise functions whole-cloth from
-their desired properties.
-Through the unified interface for finite and countable types, we can
-reuse the automation machinery for \emph{partial} proof search over infinite
-search spaces.
-Along the way, we extend the work in \cite{allaisGenericLevelPolymorphic2019} to
-prove isomorphisms between the curried and uncurried forms of \(n\)-ary
-dependent functions.
+We will see that the classification of finiteness predicates mirrors that of
+their countable counterparts, and we will prove closure under the Kleene star.
+
+\subsubsection{Practical Uses}
+Finally, in section~\ref{practical} we will demonstrate some practical uses of
+the finiteness proofs in Cubical Agda.
+We will show how to use well-known techniques
+\cite{firsovDependentlyTypedProgramming2015, devrieseBrightSideType2011} to
+automate proofs of functions with finite (and infinite) domains.
+We further show how to automate the synthesis of functions from desired
+properties.
 \section{Split Enumerability} \label{split-enumerable}
@@
 the set.
-More formally:
-\begin{rm-definition}[Split Enumerable Set]
-  \begin{equation}
-    \mathcal{E}!(A) = \Sigma {(\mathit{xs} : \textbf{List}(A))} , \Pi {(x : A)} , x \in xs
-  \end{equation}
-  We call the first component of this pair the ``support'' list, and the second
-  component the ``cover'' proof.
-\end{rm-definition}
-
-We will at this point take a moment to define some of the types we used to
-define \(\mathcal{E}!\).
-Lists, and membership thereof, are defined using \emph{containers}.
+
+Before giving the definition of the predicate, we will first define lists (and
+membership thereof): we have chosen a \emph{container}-based definition for this
+work.
 \begin{rm-definition}[Container] \label{container-def}
@@
   the elements of \(P(s)\) are called the \emph{positions} of a container.
-  \begin{equation}
-    \begin{aligned}
-      S : \text{Type}, P : S \rightarrow \text{Type} \\
-      \mathbf{Container} = S \triangleright P
-    \end{aligned}
-  \end{equation}
   We ``interpret'' a container into a functor defined like so:
   \begin{equation}
-    \llbracket S \triangleright P \rrbracket = \Pi {(X : \text{Type})} , \Sigma {(s : S)} , \left( P(s) \rightarrow X \right)
+    \llbracket S \triangleright P \rrbracket(A) = \Sigma {(s : S)} , \left( P(s) \rightarrow A \right)
   \end{equation}
@@
   \end{equation}
-\end{rm-definition}
-Containers can be used to define a wide variety of functors (streams, trees,
-etc.): lists are all that interest us now.
+  Where \(\text{fiber}\) is from \cite[definition 4.2.4]{hottbook}.
+\end{rm-definition}
+\begin{rm-definition}[Lists]
+  \begin{equation}
+    \mathbf{List} = \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket
+  \end{equation}
+\end{rm-definition}
 \begin{rm-definition}[\mathbf{Fin}]
@@
     \end{aligned}
-  \end{equation}
-\end{rm-definition}
-\begin{rm-definition}[Lists]
-  The ``shape'' of lists is \(\mathbb{N}\), indicating the length of the list in
-  question.
-  \begin{equation}
-    \mathbf{List} = \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket
   \end{equation}
@@
 our representation.
+
+Finally, we can define formally split enumerability.
+\begin{rm-definition}[Split Enumerable Set]
+  \begin{equation}
+    \mathcal{E}!(A) = \Sigma {(\mathit{xs} : \textbf{List}(A))} , \Pi {(x : A)} , x \in xs
+  \end{equation}
+  We call the first component of this pair the ``support'' list, and the second
+  component the ``cover'' proof.
+\end{rm-definition}
 \subsection{Split Surjections}
-Another, equivalent way to define ``finiteness'' is via a (split) surjection
-from a finite prefix of the natural numbers.
-In this section, we will prove that equivalence, formally.
-
+\begin{rm-definition}[Surjections]
+  \begin{multicols}{2}
+    \begin{equation}
+      A \twoheadrightarrow B = \Sigma (f : A \rightarrow B) , \text{surjective}(f)
+    \end{equation} \break
+    \begin{equation}
+      A \twoheadrightarrow! \; B = \Sigma (f : A \rightarrow B) , \text{split surjective}(f)
+    \end{equation}
+  \end{multicols}
+  Where the definitions of surjective are taken from \cite[definition
+  4.6.1]{hottbook}.
+\end{rm-definition}
+An equivalent way to define ``finiteness'' is via a split surjection from a
+finite prefix of the natural numbers.
 \begin{rm-theorem} \label{split-enum-surj}
@@
 \end{proof}
-\begin{rm-lemma}
-  Any split enumerable type is a set.
-\end{rm-lemma}
-\begin{proof}
-  By Hedberg's theorem \cite{hedbergCoherenceTheoremMartinLof1998}, since split
-  enumerable types have decidable equality
-  (proposition~\ref{split-enum-discrete}), they are sets.
-\end{proof}
+By Hedberg's theorem \cite{hedbergCoherenceTheoremMartinLof1998}, we know that
+this implies split enumerable types must all be sets.
 \subsection{Closure}
@@
 the semiring operations, and dependent sums.
-in order to show closure under exponentials (function arrows), we will need an
+In order to show closure under exponentials (function arrows), we will need an
 equivalence with \(\mathbf{Fin}\), which will be provided in
@@
 \begin{rm-lemma}
-  \(\bot\), \(\top\), and \(\mathbf{Bool}\) are all split enumerable.
+  \(\mathbf{Bool}\), \(\top\), and \(\bot\) are all split enumerable.
 \end{rm-lemma}
@@
   These non-recursive types have similar, simple proofs.
-  For each we first provide the support list: they are \([]\), \([\text{tt}]\),
-  and \([\text{false}, \text{true}]\) respectively.
-  The cover proof should return an index which points at the given element: for
-  \(\bot\) this function is present via the principle of explosion, for \(\top\)
-  we always return a \(0\), and for \(\mathbf{Bool}\) we return a \(0\) for
-  \(\text{false}\), and a \(1\) for \(\text{true}\).
   \begin{multicols}{3}
@@
 \end{proof}
-Next, we will look into how to combine proofs of split enumerability.
 \begin{rm-theorem}
-  Split-enumerability is closed under \(\sum\).
-\end{rm-theorem}
-\begin{proof}
-  Let \(E_A\) be a proof of split enumerability for some type \(A\), and \(E_U\)
-  be a function of the type:
-  \begin{equation}
-    E_U : \Pi {(x : A)} , \mathcal{E}!(U(x))
-  \end{equation}
-  In other words, a function which returns a proof of split enumerability for
-  each member of the family \(U\).
-
+  Split-enumerability is closed under \(\Sigma\).
+  \begin{equation}
+    \frac{
+      \mathcal{E}!(A) \; \; \; \Pi (x : A) , \mathcal{E}!(U(x))
+    }{
+      \mathcal{E}!(\Sigma (x : A) , U(x))
+    }
+  \end{equation}
+\end{rm-theorem}
+\begin{proof}
   To obtain the support list, we concatenate the support lists of all the proofs
@@
 \end{proof}
-\begin{rm-lemma}
-  Split-enumerability is closed under disjoint union (non-dependent sum) and
-  Cartesian product (non-dependent product).
-\end{rm-lemma}
-\begin{proof}
-  Both of these closures can be derived from closure of the \(\sum\) type.
-  The non-dependent product is equivalent to a special case of \(\sum\):
-  \begin{equation}
-    A \times B \simeq \Sigma {(x : A)} , B
-  \end{equation}
-  And disjoint union between two types \(A\) and \(B\) can be represented by the
-  following type:
-  \begin{equation}
-    A + B = \Sigma {(x : \mathbf{Bool})} , \text{if} \; x \; \text{then} \; A \; \text{else} \; B
-  \end{equation}
-  Then, since all of \(\mathbf{Bool}\), \(A\), and \(B\) are split enumerable,
-  the type \(A + B\) is split enumerable.
-\end{proof}
+Closure under disjoint union and Cartesian product both follow from \(\Sigma\),
+as these types can be defined in terms of \(\Sigma\).
+\begin{align}
+  A \times B &= \Sigma {(x : A)} , B \\
+  A + B &= \Sigma {(x : \mathbf{Bool})} , \text{if} \; x \; \text{then} \; A \; \text{else} \; B
+\end{align}
 \section{Manifest Bishop Finiteness} \label{manifest-bishop}
+Where split enumerability was the enumeration form of a split surjection from
+\(\mathbf{Fin}\), manifest Bishop finiteness is the enumeration form of an
+\emph{equivalence} with \(\mathbf{Fin}\).
+
 \begin{rm-definition}[Manifest Bishop Finiteness]
@@
   We say an item \(x\) is ``uniquely in'' some container \(\mathit{xs}\) if its
-  membership in that list is a \emph{contraction}; i.e. its membership proof
-  exists, and all such proofs are equal.
+  membership in that list is a \emph{contraction} \cite[definition
+  3.11.1]{hottbook}; i.e. its membership proof exists, and all such proofs are
+  equal.
   \begin{equation}
@@
 \end{rm-definition}
+By the definition of a contraction, we can always recover the underlying
+membership proof, meaning that we can always derive split enumerability from
+manifest Bishop finiteness.
 
@@
 \subsection{Equivalence}
-Where split enumerability was the enumeration form of a surjection from
-\(\mathbf{Fin}\), we see here that manifest Bishop finiteness is the enumeration
-form of an \emph{equivalence} with \(\mathbf{Fin}\).
 \begin{rm-lemma} \label{bishop-equiv}
@@
   Here we take the version preferred in the Cubical Agda library: contractible
-  maps \cite{hottbook}.
+  maps \cite[definition 4.4.1]{hottbook}.
   Because of the parallels between contractible maps and split surjections,
@@
 enumerability.
-\begin{rm-lemma}
-  Any manifest Bishop finite set is split enumerable.
-\end{rm-lemma}
-\begin{proof}
-  The support set carries over simply, and the cover proof can be taken from the
-  first component of the cover proof from the proof of manifest Bishop
-  finiteness.
-\end{proof}
 \begin{rm-theorem} \label{split-enum-to-manifest-bishop}
@@
 \end{proof}
-Note that while manifest Bishop finiteness as split enumerability are equivalent
+Note that while manifest Bishop finiteness and split enumerability are equivalent
 in ``strength'', the two types are not equivalent.
-In particular, there are infinitely many inhabitants of \(\mathcal{E}!\), while
-for a type \(A\) with \(n\) inhabitants, there are only \(n!\) inhabitants of
+For instance, for a type \(A\) with \(n\) inhabitants, there are infinitely many
+inhabitants of \(\mathcal{E}!(A)\), while there are only \(n!\) inhabitants of
 \(\mathcal{B}(A)\).
@@
 \section{Manifest Enumerability} \label{manifest-enumerability}
-A defining feature of split enumerability and Bishop finiteness is decidability:
-both predicates imply a decidable equality function on the underlying type.
-To find a predicate for finiteness which doesn't imply this, and therefore
-works with types other than just sets, we \emph{truncate} the membership proof.
+Both split enumerability and manifest Bishop finiteness are restricted to sets:
+because they both imply decidable equality, no non-set types can satisfy them as
+predicates.
+To find a more general predicate for finiteness, we \emph{truncate} the
+membership proof.
 \begin{rm-definition}[Manifest Enumerability]
@@
   The type \(\lVert A \rVert\) on some type \(A\) is a propositionally truncated
-  proof of \(A\).
+  proof of \(A\) \cite[3.7]{hottbook}.
   In other words, it is a proof that some \(A\) exists, but it does not tell you
@@
       As a result, \(\lVert \cdot \rVert\) forms a Monad: for our purposes, this
-      simply means tat we can work ``under'' a propositional truncation in an
+      simply means that we can work ``under'' a propositional truncation in an
       ergonomic way.
@@
 
-By hiding the position, we have essentially removed the ``decidable'' component
-from split enumerability.
-Our predicate now becomes general enough to work with non-sets: we will show
-here that the circle is manifestly enumerable.
+We will use the circle as an example non-set type which is nonetheless
+manifestly enumerable.
 \begin{rm-theorem}
@@
 \subsection{Relation to Split Enumerability}
-\begin{rm-lemma}
-  Any split enumerable type is also manifestly enumerable.
-\end{rm-lemma}
-\begin{proof}
-  The proof carries over via truncation of the cover proof.
-\end{proof}
 \begin{rm-theorem} \label{manifest-enum-to-split}
@@
   \end{equation}
-  A propositionally truncated value can be used to \emph{refute} its negation.
+  The function works via case-analysis on \(\mathbf{Dec}\): in the true case, we
+  return the proof; in the false case, we can apply the proof of negation under
+  the truncation, and then extract \(\lVert \bot \rVert\) to \(\bot\), as
+  \(\bot\) is a mere proposition, thus proving the goal via explosion.
 \end{proof}
@@
 \begin{rm-lemma}
-  Manifest enumerability is closed under dependent sum, disjoint union
-  (non-dependent sum), and Cartesian product (non-dependent product).
+  Manifest enumerability is closed under \(\Sigma\).
 \end{rm-lemma}
@@
 \end{proof}
-Notice that we do not have closure under functions: without decidability,
-manifest enumerability is not closed under function arrows.
 \section{Cardinal Finiteness} \label{cardinal}
@@
 \subsection{Closure}
-The closure proofs for cardinal finiteness are especially easy.
-In contrast to manifest enumerability, under the propositional truncation we
-have a full proof of bishop finiteness, meaning that all of the closure proofs
-carry over.
 \begin{rm-lemma}
-  Cardinal finiteness is closed under dependent and non-dependent sums,
-  products, and functions.
+  Cardinal finiteness is closed under \(\Sigma\) and \(\Pi\).
 \end{rm-lemma}
@@
 \begin{rm-theorem}
-  Kuratowski finite sets are closed under \(\sum\).
-\end{rm-theorem}
-\begin{proof}
-  This follows from the \(\sum\) closure proof on manifestly enumerable types.
-\end{proof}
-\section{Species and Containers} \label{species-and-containers}
-\todo[inline]{Figure out this section}
+  Kuratowski finite sets are closed under \(\Sigma\).
+\end{rm-theorem}
+\begin{proof}
+  This follows from the \(\Sigma\) closure proof on manifestly enumerable types.
+\end{proof}
 \section{Infinite Cardinalities} \label{infinite-cardinalities}
@@
 \begin{proof}
-  As with the proof of closure under \(\sum\), our main task here is to figure
+  As with the proof of closure under \(\Sigma\), our main task here is to figure
   out a way to arrange the indices such that 
@@
 \begin{itemize}
-  \item Our generic representation can handle dependent \(\sum\) and \(\prod\)
+  \item Our generic representation can handle dependent \(\Sigma\) and \(\prod\)
     types (rather than their non-dependent counterparts, \(\times\) and

### 2020-01-07 paper.tex
corrected closure proofs on cardinal finiteness

@@
 typed programming languages.
-\subsubsection{Strong Finiteness Predicates}
+\subsection{Strong Finiteness Predicates}
 We will first explore the ``strong'' notions of finiteness (i.e. those at least
@@
 
-\subsubsection{Infinite Types}
+\subsection{Infinite Types}
 In section~\ref{infinite-cardinalities}, we will extend our study of finite
@@
 
-\subsubsection{Practical Uses}
+\subsection{Practical Uses}
 Finally, in section~\ref{practical} we will demonstrate some practical uses of
@@
 \end{proof}
-\begin{rm-theorem}
+\begin{rm-theorem} \label{split-enum-sigma}
   Split-enumerability is closed under \(\Sigma\).
-  \begin{equation}
+  \begin{equation} \label{split-enum-sigma-clos-sig}
     \frac{
@@
 \subsection{Closure}
+Since we don't have an equivalence with \(\mathbf{Fin}\), we don't get closure
+under \(\Pi\).
 \begin{rm-lemma}
@@
 \begin{proof}
-  For these three closures, the proofs on split enumerability consisted of a
-  list manipulation followed by a proof that membership was preserved by the
-  list manipulation.
-  Because we separate these two concerns, the proofs carry over onto manifest
-  enumerability: the support list manipulation stays the same, and the
-  cover proofs are performed ``under'' the truncation.
+  This closure proof is almost the same as theorem~\ref{split-enum-sigma}.
+  The manipulation of the support lists can be carried over as-is; but the type
+  of the cover proof has changed, so it will need to be updated.
+  As it happens, the translation is straightforward: we effectively write the
+  ``monadic'' version of the old function.
 \end{proof}
@@
     \mathcal{C}(A) = \lVert \mathcal{B}(A) \rVert
-  \end{equation}
-\end{rm-definition}
+  \end{equation}%
+\end{rm-definition}%
 \subsection{Closure}
+The closure properties of cardinal finiteness are effectively the non-dependent
+versions of manifest Bishop finiteness.
+To see why, consider equation~\ref{split-enum-sigma-clos-sig}.
+We can ``lift'' the proof (as a binary function) under a propositional
+truncation, giving us equation~\ref{sigma-proof-lifted}, but that doesn't give us
+the desired closure proof (equation~\ref{sigma-proof-desired}).
+\begin{multicols}{2}
+  \noindent
+  \begin{equation} \label{sigma-proof-lifted}
+    \frac{
+      \lVert \mathcal{B}(A) \rVert \; \; \; \lVert \Pi (x : A) , \mathcal{B}(U(x)) \rVert
+    }{
+      \lVert \mathcal{B}(\Sigma(x : A) , U(x)) \rVert
+    }
+  \end{equation}
+  \begin{equation} \label{sigma-proof-desired}
+    \frac{
+      \lVert \mathcal{B}(A) \rVert \; \; \; \Pi (x : A) , \lVert \mathcal{B}(U(x))  \rVert
+    }{
+      \lVert \mathcal{B}(\Sigma(x : A) , U(x)) \rVert
+    }
+  \end{equation}
+\end{multicols}%
+To remedy the mismatch we would need a function of the type:
+\begin{equation}
+  \left( \Pi(x : A) , \lVert \mathcal{B}(U(x)) \rVert \right) \rightarrow
+  \lVert \Pi(x : A) , \mathcal{B}(U(x)) \rVert
+\end{equation}
+Unfortunately, this equation in particular is a form of the axiom of choice
+\cite[equation 3.8.3]{hottbook}.
+This leaves us with closure under only the non-dependent operations.
 \begin{rm-lemma}
-  Cardinal finiteness is closed under \(\Sigma\) and \(\Pi\).
+  Cardinal finiteness is closed under \(\times\), \(+\), and \(\rightarrow\).
 \end{rm-lemma}
 \begin{proof}
-  All closure functions can be lifted under propositional truncation.
-  Therefore, cardinal finiteness has the same closure properties as manifest
-  Bishop finiteness.
+  All of these closure proofs can be lifted directly from their corresponding
+  proofs on manifest Bishop finiteness.
 \end{proof}

### 2020-01-07 paper.tex
down to 16 pages

@@
 \section{Introduction}
-In this work, we will explore finite types in Cubical Type Theory
-\cite{cohenCubicalTypeTheory2016}, and expose their relationship to infinite
-types, species, and demonstrate their practical uses for proofs in dependently
-typed programming languages.
 \subsection{Strong Finiteness Predicates}
@@
 
-\input{figures/finite-classification}
 Figure~\ref{finite-classification} organises the predicates according to
-their ``strength''; i.e. how much information they provide about a conforming
+their strength; i.e. how much information they provide about a conforming
 type.
@@
 even sets, never mind discrete ones.
+
+\input{figures/finite-classification}
 
@@
 decidable Kuratowski finite sets form a topos.
-More specifically, we show that cardinal finite sets form a topos, that
-decidable Kuratowski finite sets are equivalent in strength to cardinal finite
-sets, and carry the proof over.
-This result relies on proofs on each of the other finiteness predicates.
-
 \subsection{Infinite Types}
@@
 \begin{rm-definition}[Lists]
-  \begin{equation}
-    \mathbf{List} = \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket
+  \begin{equation}%
+    \mathbf{List} = \llbracket \mathbb{N} \triangleright \mathbf{Fin} \rrbracket
   \end{equation}
@@
 \begin{rm-definition}[\mathbf{Fin}]
-  \(\mathbf{Fin}\)(n) is the type of natural numbers smaller than \(n\).
-  It is defined inductively, as follows:
-  \begin{equation}
-    \begin{aligned}
-      \mathbf{Fin}(0) &= \bot \\
-      \mathbf{Fin}(n + 1) &= \top + \mathbf{Fin}(n)
-    \end{aligned}
-  \end{equation}
+  \(\mathbf{Fin}(n)\) is the type of natural numbers smaller than \(n\).
+  We define it the standard way, where \(\mathbf{Fin}(0) = \bot\) and
+  \(\mathbf{Fin}(n + 1) &= \top + \mathbf{Fin}(n)\).
 \end{rm-definition}
@@
 \end{rm-definition}
-An equivalent way to define ``finiteness'' is via a split surjection from a
-finite prefix of the natural numbers.
 \begin{rm-theorem} \label{split-enum-surj}
@@
 \end{proof}
-\subsection{Decidable Equality}
+
 \begin{rm-lemma} \label{split-enum-discrete}
@@
 section~\ref{manifest-bishop}.
-\begin{rm-lemma}
+\begin{rm-lemma} \label{split-enum-bool-top-bot}
   \(\mathbf{Bool}\), \(\top\), and \(\bot\) are all split enumerable.
@@
 \begin{proof}
-  Let \(E\) be a proof of split enumerability for some set \(A\).
   From proposition~\ref{split-enum-discrete} we can derive decidable equality on
-  \(A\), and using this we can define a function \(\text{uniques}\) which filters
+  \(A\), and using this we can define a function (called \(\text{uniques}\)) which filters
   out duplicates from lists of \(A\)s.
-  \begin{equation}
-    \text{uniques} : \textbf{List}(A) \rightarrow \textbf{List}(A)
-  \end{equation}
   This gives us our support list.
-
-  It suffices now to prove the following:
+  To generate the cover proof it suffices now to prove the following:
   \begin{equation}
@@
   \end{equation}
-  And from that we can generate our cover proof.
 \end{proof}
-Note that while manifest Bishop finiteness and split enumerability are equivalent
-in ``strength'', the two types are not equivalent.
-For instance, for a type \(A\) with \(n\) inhabitants, there are infinitely many
-inhabitants of \(\mathcal{E}!(A)\), while there are only \(n!\) inhabitants of
-\(\mathcal{B}(A)\).
 \subsection{Closure}
@@
   \end{equation}
-  Seen another way, a permutation is an isomorphism of cover proofs.
 \end{rm-definition}
-We also prove some of the identities you might expect with regards to
-permutations.
 \begin{rm-lemma}
@@
 \end{proof}
-Based on previous proofs, we can derive that if we add decidability to a
-Kuratowski finite type we retrieve cardinal finiteness.
-\begin{rm-lemma}
-  Any Kuratowski finite set with decidable equality is cardinally finite.
-\end{rm-lemma}
-\begin{proof}
-  We first convert to a propositionally truncated proof of manifest
-  enumerability.
-  From here, we can convert to manifest Bishop finiteness under the truncation
-  with our decidable equality proof.
-\end{proof}
 \subsection{Topos}
@@
 \begin{rm-theorem}
-  Kuratowski finite sets are closed under \(\Sigma\).
+  Kuratowski finite sets are closed under \(\times\) and \(+\).
 \end{rm-theorem}
 \begin{proof}
-  This follows from the \(\Sigma\) closure proof on manifestly enumerable types.
+  These follow from their lifted versions on manifestly enumerable types.
 \end{proof}
 \section{Infinite Cardinalities} \label{infinite-cardinalities}
-While the previous section purported to be about finite sets, we can now see
-that it was really only studying surjections and isomorphisms of different
-flavours between types and \(\mathbf{Fin}\).
-The natural next question which arises, then, is if we can extend that work to
-surjections and isomorphisms with \(\mathbb{N}\).
-In more standard language, what we're referring to here is of course
-countable infinity and related concepts.
-However, as with finite types, the ``countably infinite'' types have more
-varieties in constructive mathematics than their classical counterparts.
+In the previous sections we saw different flavours of finiteness which were
+really just different flavours of relations to \(\mathbf{Fin}\).
+In this section we will see that we can construct a similar classification of
+relations to \(\mathbb{N}\), in the form of the countably infinite types.
 \subsection{Split Countable Types}
@@
 We need change only one element: instead of a support \emph{list}, we instead
-have a support \emph{stream}, which is its infinite, coinductive counterpart.
+have a support \emph{stream}, which is its infinite.
 \begin{rm-definition}[Stream]
-  We will work with two isomorphic definitions of streams.
-  The first is the following:
+  As with lists, it is cleaner, theory-wise, to define streams as a container.
   \begin{equation}
-    \mathbf{Stream}(A) = \mathbb{N} \rightarrow A
+    \mathbf{Stream}(A) = \llbracket \top \triangleright \text{const} (\mathbb{N}) \rrbracket(A) \simeq \mathbb{N} \rightarrow A
   \end{equation}
   Conceptually, a stream is like a list without an end.
-  Of course, such a type can not be defined in the same way as a list: it would
-  be impossible to construct a value, as inductive types do not admit
-  infinitely-sized inhabitants.
 \end{rm-definition}
-We can now define the split countable types.
 \begin{rm-definition}[Split Countability]
@@
 \section{Practical Uses} \label{practical}
-The theory of finite types in constructive mathematics, and in HoTT in
-particular, is rich and interesting, as we hope we have demonstrated thus far.
-As well as being theoretically interesting, however, the proofs and combinators
-we have defined here are \emph{practically} useful.
 \subsection{Omniscience}
@@
 \ExecuteMetaData[agda/Data/Pauli.tex]{def}
-For this type, there are several practical tasks we hope to achieve with the
-help of our finiteness combinators:
-\begin{itemize}
-  \item Define decidable equality on the type
-  \item Define the group operation on the type
-  \item Prove properties about the group operation
-\end{itemize}
-Unfortunately, the simple pattern-matching way to do many of these tasks is
-prohibitively verbose.
 As \(\AgdaDatatype{Pauli}\) has 4 constructors, \(n\)-ary functions on
-\(\AgdaDatatype{Pauli}\) may require up to \(4^n\) cases.
-A proof of decidable equality is one such function: it can be seen fully
-worked-through in the appendix.
+\(\AgdaDatatype{Pauli}\) may require up to \(4^n\) cases, making even simple
+proofs prohibitively verbose.
 
-The alternative is to derive the things we need from \(\Enum!\) somehow.
-First, then, we need an instance for \(\AgdaDatatype{Pauli}\):
-\ExecuteMetaData[agda/Data/Pauli.tex]{inst}
+The alternative is to derive the things we need from \(\mathcal{E}!\) somehow.
+As \(\AgdaDatatype{Pauli}\) is a simple finite type, the instance can be defined
+in a similar way to those in lemma~\ref{split-enum-bool-top-bot}.
 From here we can already derive decidable equality, a function which requires 16
@@
 typechecking.
-\ExecuteMetaData[agda/Cardinality/Finite/Search.tex]{auto}
-And finally we can prove the property we wanted to on \(\AgdaDatatype{Pauli}\)
-like so:
+As an example, we can check that the group operation on \(\AgdaDatatype{Pauli}\)
+is idempotent like so:
 \ExecuteMetaData[agda/Data/Pauli.tex]{noninst-prf}
 
-As a quick aside, when the property is \emph{not} true, for instance:
-\ExecuteMetaData[agda/Data/Pauli.tex]{contra}
-Agda will fail by not finding an instance of \(\bot\).
-The error message specifically is:
-\begin{displayquote}
-  \verb+No instance of type ⊥ was found in scope.+
-\end{displayquote}
-We can actually display a counterexample, by defining a custom empty type
-parameterised by the counterexample itself.
-\ExecuteMetaData[agda/Cardinality/Finite/Search.tex]{counterexample-def}
+As a quick aside, when the property is \emph{not} true (for instance \(\forall x
+\rightarrow x \cdot x \equiv x\)), Agda will fail by not finding an instance of
+\(\bot\).
+The error message specifically is
+\verb+No instance of type ⊥ was found in scope+, which isn't very helpful.
+We can do a little better by displaying the particular element on which the
+property failed, by defining a custom empty type parameterised by the
+counterexample itself.
 Combined with some other changes to the combinators, this will give the
 following more helpful error message:
-\begin{displayquote}
-  \verb+No instance of type Counterexample X was found in scope.+
-\end{displayquote}
-\subsection{Instances}
-Running decision procedures during typechecking isn't the only use of instances:
-we will also use them to automate the finding of a proof of finiteness for a
-given type.
-The change to our \(\AgdaFunction{∀↯}\) is simple: we just change the first
-parameter from explicit to an instance.
-\ExecuteMetaData[agda/Cardinality/Finite/Search.tex]{auto-inst}
-
-The change to the \(\AgdaDatatype{Pauli}\) proof of cancellation isn't
-groundbreaking at first:
-\ExecuteMetaData[agda/Data/Pauli.tex]{cancel-auto}
-The real benefit is that we have provided instances for things like tuples,
-meaning we can now prove \(n\)-ary properties.
-\ExecuteMetaData[agda/Data/Pauli.tex]{comm-uncurried}
-\subsection{Generic Currying and Uncurrying}
-While we have arguably removed the bulk of the boilerplate from the automated
-proofs, there is still the case of the ugly noise of currying and uncurrying.
-In this section, we take inspiration from
-\cite{allaisGenericLevelPolymorphic2019} to develop a small interface to generic
-\(n\)-ary functions and properties.
-What we provide here differs from that work in the following ways:
-\begin{itemize}
-  \item Our generic representation can handle dependent \(\Sigma\) and \(\prod\)
+\verb+No instance of type Counterexample X was found in scope+.
+\subsection{Multiple Arguments}
+The automation machinery above only deals with single-argument predicates.
+This is not a problem, as we know that we can work with multiple arguments by
+currying and uncurrying, since all of the finiteness predicates are closed under
+\(\times\).
+To automate away the curry/uncurry noise we will use instance search, building
+on \cite{allaisGenericLevelPolymorphic2019} to develop a small interface to
+generic \(n\)-ary functions and properties.
+Our generic representation can handle dependent \(\Sigma\) and \(\prod\)
 types (rather than their non-dependent counterparts, \(\times\) and
@@
 as the obvious next step. 
-  \item We implement the curry-uncurry combinators as (verified) isomorphisms.
-    Since we are in a cubical setting, this gives us equivalences between the
-    types, a feature not available in standard Agda.
-  \item We deal with implicit and instance arguments generically.
-\end{itemize}
+We also implement the curry-uncurry combinators as (verified) isomorphisms.
+
 A full explanation of our implementation is beyond the scope of this work, so we
-only present the finished interface.
-\ExecuteMetaData[agda/Cardinality/Finite/Search.tex]{all-nary}
-Which is used like so:
+only present the finished interface, which is used like so:
 \ExecuteMetaData[agda/Data/Pauli.tex]{prf-curried}
+
+\todo[inline]{Synthesise functions}
+\todo[inline]{Partial search over infinite types}
 \bibliographystyle{splncs04}

### 2020-01-08 paper.tex
more

@@
     \begin{aligned}
-      \lVert A \rVert &= \\
-        &| \; \lvert \cdot \rvert : A \rightarrow \lVert A \rVert ; \\
-        &| \; \text{squash} : \Pi {(x, y : \lVert A \rVert)} , x \equiv y  ; \\
+      \lVert A \rVert =& \; \lvert \cdot \rvert : A \rightarrow \lVert A \rVert ; \\
+                      |& \; \text{squash} : \Pi {(x, y : \lVert A \rVert)} , x \equiv y  ; 
     \end{aligned}
@@
     \begin{aligned}
-      \mathcal{K}(A) &=  \\
-        &\left\vert \; \cdot \dblcolon \cdot : A \times \mathcal{K}(A) \rightarrow \mathcal{K}(A) \right; \\
-        &\left\vert \; [] : \mathcal{K}(A) \right; \\
-        &\left\vert \; \text{com} : \Pi (x, y: A) , \Pi (\mathit{xs}: \mathcal{K}(A)) , x \dblcolon y \dblcolon \mathit{xs} \equiv y \dblcolon x \dblcolon \mathit{xs} \right; \\
-        &\left\vert \; \text{dup} : \Pi (\mathit{x}: A) , \Pi (\mathit{xs}: \mathcal{K}(A)) , x \dblcolon x \dblcolon \mathit{xs} \equiv x \dblcolon \mathit{xs} \right; \\
-        &\left\vert \; \text{trunc} : \Pi (\mathit{xs}, \mathit{ys}: \mathcal{K}(A)) , \Pi (p, q : \mathit{xs} \equiv \mathit{ys}) , p \equiv q \right;
+      \mathcal{K}(A) =&
+       \; \cdot \dblcolon \cdot : A \times \mathcal{K}(A) \rightarrow \mathcal{K}(A) ; \\
+      \vert& \; [] : \mathcal{K}(A) ; \\
+      \vert& \; \text{com} : \Pi (x, y: A) , \Pi (\mathit{xs}: \mathcal{K}(A)) , x \dblcolon y \dblcolon \mathit{xs} \equiv y \dblcolon x \dblcolon \mathit{xs} ; \\
+      \vert& \; \text{dup} : \Pi (\mathit{x}: A) , \Pi (\mathit{xs}: \mathcal{K}(A)) , x \dblcolon x \dblcolon \mathit{xs} \equiv x \dblcolon \mathit{xs} ; \\
+      \vert& \; \text{trunc} : \Pi (\mathit{xs}, \mathit{ys}: \mathcal{K}(A)) , \Pi (p, q : \mathit{xs} \equiv \mathit{ys}) , p \equiv q ;
     \end{aligned}
@@
   First, we need to provide equations for the two point constructors.
-  \begin{equation}
-    \begin{alignat}{2}
-      x &\in& [] &= \bot \\
-      x &\in& y \dblcolon \mathit{xs} &= \lVert (x \equiv y) + (x \in \mathit{xs}) \rVert
-    \end{alignat}
-  \end{equation}
+  \(x \in [] = \bot\), and \(x \in y \dblcolon \mathit{xs} = \lVert (x \equiv y)
+  + (x \in \mathit{xs}) \rVert \).
   The \(\text{com}\) and \(\text{dup}\) constructors are handled by proving that
@@
 
-\subsection{Strength}
-Since the Kuratowski set is a departure in structure from our previous
-list-based notions of finiteness, it makes sense to first look for the closest
-list-based analogue.
-As it turns out, that analogue is manifestly enumerable finiteness, with the
-order removed.
 \begin{rm-theorem}
@@
 In particular, we see that decidable Kuratowski finite sets form a \emph{topos}.
-\todo[inline]{Fill in rest}
-\subsection{Closure}
 \begin{rm-theorem}
@@
 \end{proof}
+\todo[inline]{Fill in rest}
 \section{Infinite Cardinalities} \label{infinite-cardinalities}

### 2020-01-08 check-all.sh
all working now

@@ -4,7 +4,7 @@ echo '{-# OPTIONS --cubical #-}
 ' > Everything
 echo 'module Everything where
 ' >> Everything
-find -E . -type f -regex '.*\.l?agda' | cut -c 3- | cut -f1 -d'.' | sed 's/\//\./g' | sed 's/^/import /' >> Everything
+find -E . -type f -regex '.*\.l?agda' | cut -c 3- | cut -f1 -d'.' | sed 's/\//\./g' | sed 's/^/open import /' >> Everything
 mv Everything Everything.agda
 agda Everything.agda
 rm Everything.agda

### 2020-01-09 check-all.sh
docs

@@ -1,11 +1,11 @@
 #!/bin/bash
 ( cd agda
-echo '{-# OPTIONS --cubical #-}
+echo '{-# OPTIONS --cubical --safe #-}
 ' > Everything
 echo 'module Everything where
 ' >> Everything
 find -E . -type f -regex '.*\.l?agda' | cut -c 3- | cut -f1 -d'.' | sed 's/\//\./g' | sed 's/^/open import /' >> Everything
 mv Everything Everything.agda
-agda Everything.agda
+agda --html --html-dir=../docs Everything.agda
 rm Everything.agda
 )

### 2020-01-09 paper.tex
initial agda
now totally moved over

@@
   \begin{equation}
-    \mathcal{E}!(A) \simeq \Sigma {(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \twoheadrightarrow ! \; A \right)
+    \mathcal{E}!(A) \iff \Sigma {(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \twoheadrightarrow ! \; A \right)
   \end{equation}
@@
   our goal is simply a reassociation.
-  \ExecuteMetaData[agda/Cardinality/Container/Finite.tex]{split-surj}
+  \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{split-surj}
   To be clear: in Agda, the proof could simply be \(\AgdaFunction{reassoc}\); we
@@
     \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{bool}
-    \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{unit}
-    \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{empty}
+    \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{top}
+    \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{bot}
   \end{multicols}
@@
   In Agda:
-  \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{sigma-enum}
+  \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{sigma-sup}
   ``do-notation'' is available to us as we're working in the list monad.
@@
 In particular, we see that decidable Kuratowski finite sets form a \emph{topos}.
-\begin{rm-theorem}
-  Kuratowski finite sets are closed under \(\times\) and \(+\).
-\end{rm-theorem}
-\begin{proof}
-  These follow from their lifted versions on manifestly enumerable types.
-\end{proof}
-\todo[inline]{Fill in rest}
+We already have most of the components we need: closure under \(\bot\),
+\(\top\), \(\mathbf{Bool}\), \(+\), \(\times\), and \(\rightarrow\).
+What remains is the subobject classifier.
+\begin{definition}
+\end{definition}
 \section{Infinite Cardinalities} \label{infinite-cardinalities}
@@
 typechecking.
-As an example, we can check that the group operation on \(\AgdaDatatype{Pauli}\)
-is idempotent like so:
-\ExecuteMetaData[agda/Data/Pauli.tex]{noninst-prf}
-
-As a quick aside, when the property is \emph{not} true (for instance \(\forall x
-\rightarrow x \cdot x \equiv x\)), Agda will fail by not finding an instance of
-\(\bot\).
-The error message specifically is
-\verb+No instance of type ⊥ was found in scope+, which isn't very helpful.
-We can do a little better by displaying the particular element on which the
-property failed, by defining a custom empty type parameterised by the
-counterexample itself.
-Combined with some other changes to the combinators, this will give the
-following more helpful error message:
-\verb+No instance of type Counterexample X was found in scope+.
 \subsection{Multiple Arguments}
@@
 only present the finished interface, which is used like so:
-\ExecuteMetaData[agda/Data/Pauli.tex]{prf-curried}
+\ExecuteMetaData[agda/Data/Pauli.tex]{assoc-prf}
 

### 2020-01-09 paper.tex
short

@@
 \begin{proof}
-  These non-recursive types have similar, simple proofs.
-  \begin{multicols}{3}
-    \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{bool}
-    \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{top}
-    \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{bot}
-  \end{multicols}
+  Each of these types clearly has a finite number of elements (2, 1, and 0,
+  respectively), and furthermore has a straightforward enumeration.
 \end{proof}
@@
 manifest Bishop finiteness.
-
-A nice consequence of prohibiting duplicates is that now the length of the
-support list is the same as the cardinality of the set.
 \subsection{Equivalence}

### 2020-01-10 figures/pairing-functions.tex
done ish

\begin{figure}
  \centering
  \begin{subfigure}[b]{.5\linewidth}
    \begin{tikzcd}[sep=tiny,font=\footnotesize]
      (1,e) \ar[rdddd, out=-45, in=135] & (2,e) \ar[rdddd, out=-45, in=135] & (3,e) \ar[rdddd, out=-45, in=135] & (4,e) \ar[rdddd, out=-45, in=135] & (5,e)        \\
      (1,d) \ar[u]     & (2,d) \ar[u]    & (3,d) \ar[u]    & (4,d) \ar[u]    & (5,d) \ar[u] \\
      (1,c) \ar[u]     & (2,c) \ar[u]    & (3,c) \ar[u]    & (4,c) \ar[u]    & (5,c) \ar[u] \\
      (1,b) \ar[u]     & (2,b) \ar[u]    & (3,b) \ar[u]    & (4,b) \ar[u]    & (5,b) \ar[u] \\
      (1,a) \ar[u]     & (2,a) \ar[u]    & (3,a) \ar[u]    & (4,a) \ar[u]    & (5,a) \ar[u]
    \end{tikzcd}
    \caption{Cartesian}
    \label{cartesian}
  \end{subfigure}%
  \begin{subfigure}[b]{.5\linewidth}
    \begin{tikzcd}[sep=tiny,font=\footnotesize]
      (1,e) \ar[dr] & (2,e) \ar[dr]  & (3,e) \ar[dr]    & (4,e) \ar[dr] & (5,e) \\
      (1,d) \ar[dr] & (2,d) \ar[dr]  & (3,d) \ar[dr]    & (4,d) \ar[dr] & (5,d) \ar[u] \\
      (1,c) \ar[dr] & (2,c) \ar[dr]  & (3,c) \ar[dr]    & (4,c) \ar[dr] & (5,c) \ar[uul] \\
      (1,b) \ar[dr] & (2,b) \ar[dr]  & (3,b) \ar[dr]    & (4,b) \ar[dr] & (5,b) \ar[uuull, out=130, in=-50] \\
      (1,a) \ar[u]  & (2,a) \ar[uul, out=130, in=-50] & (3,a) \ar[uuull, out=130, in=-50] & (4,a) \ar[uuuulll, out=130, in=-50] & (5,a) \ar[uuuulll, out=130, in=-50]
    \end{tikzcd}
    \caption{Cantor}
    \label{cantor}
  \end{subfigure}
  \caption{Two possible products for the sets \(\left[ 1 \dots 5 \right]\) and
    \(\left[  a \dots e \right]\)}
  \label{pairings}
\end{figure}

### 2020-01-10 paper.tex
done ish

@@
 \subsection{Strong Finiteness Predicates}
+\input{figures/finite-classification}
 We will first explore the ``strong'' notions of finiteness (i.e. those at least
@@
 Figure~\ref{finite-classification} organises the predicates according to
-their strength; i.e. how much information they provide about a conforming
-type.
-For instance, a proof that some type \(A\) is manifestly Bishop finite (the
-strongest of the notions, explored in section~\ref{manifest-bishop}) also tells
-us that \(A\) is discrete (has decidable equality), and gives us a linear order
-on the type.
-A type that is Kuratowski finite (section~\ref{kuratowski}) has no such extra
-features: indeed, we will see examples of Kuratowski finite types which are not
-even sets, never mind discrete ones.
-
-\input{figures/finite-classification}
-
+their strength.
 We will go through each of the predicates, proving how to weaken each (i.e. we
@@
   component the ``cover'' proof.
+  This definition is what is referred to as \verb+Listable+ in
 \end{rm-definition}
+\cite{firsovDependentlyTypedProgramming2015}.
 \subsection{Split Surjections}
@@
 \begin{proof}
-  The proof is surprisingly short: after sufficient inlining, it emerges that
-  our goal is simply a reassociation.
+  After sufficient inlining, it emerges that our goal is simply a reassociation.
   \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{split-surj}
@@
 \end{proof}
-
 \begin{rm-lemma} \label{split-enum-discrete}
@@
 \subsection{Relationship to Split Enumerability}
-We now show that manifest Bishop finiteness has equal strength to split
-enumerability.
 \begin{rm-theorem} \label{split-enum-to-manifest-bishop}
@@
 
-  We will use and consume values of the type \(\lVert A \rVert\) in three main
-  ways.
+  We will use three eliminators from \(\lVert A \rVert\) in this paper.
   \begin{enumerate}
-    \item \label{elim-prop-prop} We can first eliminate from \(\lVert A \rVert\)
-      into any type which is a proposition.
-      In other words, given a function \(f : A \rightarrow B\), and a proof that
-      \(B\) is a proposition, we can construct a function \(\lVert A \rVert
-      \rightarrow B\).
-    \item \label{elim-prop-monad} As a consequence of this first point, we can
-      always eliminate into another propositionally truncated type.
-      As a result, \(\lVert \cdot \rVert\) forms a Monad: for our purposes, this
-      simply means that we can work ``under'' a propositional truncation in an
-      ergonomic way.
+  \item \label{elim-prop-prop} For any function \(A \rightarrow B\), where
+    \(\text{isProp}(B)\), we have a function \(\lVert A \rVert \rightarrow B\).
+  \item \label{elim-prop-monad} A special case of \ref{elim-prop-prop} implies
+    that \(\lVert \cdot \rVert\) forms a monad: this means that we get the usual
+    Monadic operators on \(\lVert \cdot \rVert\) (bind, pure, fmap, etc.).
     \item \label{elim-prop-coh} We can eliminate from \(\lVert A \rVert\) with a
       function \(f : A \rightarrow B\) iff \(f\) ``doesn't care'' about the
-      choice of \(A\).
-      \begin{equation}
-        \Pi {(x , y : A)} , f(x) \equiv f(y)
-      \end{equation}
+      choice of \(A\) (\(\Pi {(x , y : A)} , f(x) \equiv f(y) \)).
       Formally speaking, \(f\) needs to be ``coherently constant''
@@
 \end{rm-lemma}
-\begin{proof}
-  As with the other surjection proof (lemma~\ref{split-enum-surj}), this is
-  simply a reassociation.
-\end{proof}
 \subsection{Relation to Split Enumerability}
@@
 \end{proof}
-\subsection{Decidable Equality}
+\subsection{Strength}
+The following two theorems are proven in
+\cite{yorgeyCombinatorialSpeciesLabelled2014}, in much the same way as we have
+done here.
+Our contribution for this section is simply the formalisation.
 \begin{rm-theorem} \label{cardinal-finite-discrete}
@@
 \begin{proof}
-  We will use eliminator~\ref{elim-prop-prop} from definition~\ref{prop-trunc}.
-  Manifest Bishop finiteness implies decidable equality already, so our task
-  here is to prove that decidable equality itself is a proposition.
-
-  We know that if a type \(A\) is a proposition, then the decision over that
-  type is also a proposition.
-  Then, via Hedberg's theorem, we know that any type with decidable equality is
-  a set, meaning that paths in that type are themselves propositions.
-  Therefore we can derive that a decision of equality on elements with decidable
-  equality is a proposition, and by function extensionality we see that
-  decidable equality is itself a proposition.
+  See \cite[Proposition 2.4.10]{yorgeyCombinatorialSpeciesLabelled2014}.
 \end{proof}
-\subsection{Cardinality}
 \begin{rm-theorem}
@@
 \begin{proof}
-  Our task here is to ``pull out'' the cardinality of the set from under the
-  propositional truncation.
-  In effect, we need the following function:
-  \begin{equation}
-    \left\lVert \Sigma {(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \simeq A \right) \right\rVert \rightarrow
-    \Sigma {(n : \mathbb{N})} , \left\lVert \mathbf{Fin}(n) \simeq A \right\rVert
-  \end{equation}
-
-  We will use eliminator~\ref{elim-prop-coh} from definition~\ref{prop-trunc}.
-  We eliminate with the following function:
-  \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{trunc-alg}
-
-  To show that \(\text{alg}\) is coherently constant, we first notice that the
-  second element of the output pair is propositionally truncated, meaning that
-  it is trivially equal to any other element of the same type.
-  Our task, then, simplifies to demonstrating that the first element of the
-  output pair is coherently constant.
-  \begin{equation}
-    \Pi {(x : \Sigma {(n : \mathbb{N})} , \textbf{Fin}(n) \simeq A)} ,
-    \Pi {(y : \Sigma {(m : \mathbb{N})} , \textbf{Fin}(m) \simeq A)} ,
-    n \equiv m
-  \end{equation}
-
-  Notice that \(\mathbf{Fin}(n)\) and \(\mathbf{Fin}(m)\) are both equivalent to
-  \(A\): we can join these proofs together, giving us the following:
-  \begin{equation}
-    \textbf{Fin}(n) \equiv \textbf{Fin}(m)
-  \end{equation}
-  All that remains now is to prove that \(\mathbf{Fin}\) is injective.
-  Though the proof is surprisingly complex, it is a well-known puzzle in
-  Martin-Löf type theory.
-  Our proof does not differ significantly from standard approaches, so we will
-  not detail it here.
+  \cite[Proposition 2.4.9]{yorgeyCombinatorialSpeciesLabelled2014}.
 \end{proof}
@@
 \begin{proof}
-  Again, this proof is quite complex, so we only give a sketch here.
   First, we prove two properties about insertion sort:
@@
 What remains is the subobject classifier.
-\begin{definition}
-\end{definition}
+\todo[inline]{Topos!}
 \section{Infinite Cardinalities} \label{infinite-cardinalities}
@@
 \begin{rm-definition}[Stream]
-  As with lists, it is cleaner, theory-wise, to define streams as a container.
   \begin{equation}
@@
   \end{equation}
-  Conceptually, a stream is like a list without an end.
 \end{rm-definition}
@@
 \end{rm-definition}
-\subsubsection{\(\Sigma\) Closure}
+\subsubsection{Closure}
 We know that countable infinity is not closed under the exponential (function
@@
 search of the pairings of both lists.
-\begin{figure}
-  \centering
-  \begin{subfigure}[b]{.5\linewidth}
-    \begin{tikzcd}[sep=tiny,font=\footnotesize]
-      (1,e) \ar[rdddd, out=-45, in=135] & (2,e) \ar[rdddd, out=-45, in=135] & (3,e) \ar[rdddd, out=-45, in=135] & (4,e) \ar[rdddd, out=-45, in=135] & (5,e)        \\
-      (1,d) \ar[u]     & (2,d) \ar[u]    & (3,d) \ar[u]    & (4,d) \ar[u]    & (5,d) \ar[u] \\
-      (1,c) \ar[u]     & (2,c) \ar[u]    & (3,c) \ar[u]    & (4,c) \ar[u]    & (5,c) \ar[u] \\
-      (1,b) \ar[u]     & (2,b) \ar[u]    & (3,b) \ar[u]    & (4,b) \ar[u]    & (5,b) \ar[u] \\
-      (1,a) \ar[u]     & (2,a) \ar[u]    & (3,a) \ar[u]    & (4,a) \ar[u]    & (5,a) \ar[u]
-    \end{tikzcd}
-    \caption{Cartesian}
-    \label{cartesian}
-  \end{subfigure}%
-  \begin{subfigure}[b]{.5\linewidth}
-    \begin{tikzcd}[sep=tiny,font=\footnotesize]
-      (1,e) \ar[dr] & (2,e) \ar[dr]  & (3,e) \ar[dr]    & (4,e) \ar[dr] & (5,e) \\
-      (1,d) \ar[dr] & (2,d) \ar[dr]  & (3,d) \ar[dr]    & (4,d) \ar[dr] & (5,d) \ar[u] \\
-      (1,c) \ar[dr] & (2,c) \ar[dr]  & (3,c) \ar[dr]    & (4,c) \ar[dr] & (5,c) \ar[uul] \\
-      (1,b) \ar[dr] & (2,b) \ar[dr]  & (3,b) \ar[dr]    & (4,b) \ar[dr] & (5,b) \ar[uuull, out=130, in=-50] \\
-      (1,a) \ar[u]  & (2,a) \ar[uul, out=130, in=-50] & (3,a) \ar[uuull, out=130, in=-50] & (4,a) \ar[uuuulll, out=130, in=-50] & (5,a) \ar[uuuulll, out=130, in=-50]
-    \end{tikzcd}
-    \caption{Cantor}
-    \label{cantor}
-  \end{subfigure}
-  \caption{Two possible products for the sets \(\left[ 1 \dots 5 \right]\) and
-    \(\left[  a \dots e \right]\)}
-  \label{pairings}
-\end{figure}
+\input{figures/pairing-functions.tex}
 \begin{rm-theorem} \label{split-countability-sigma}
@@
 \begin{proof}
-  This final proof consists of the support stream, and the proof that the
-  support stream covers the input.
-
-  As mentioned, we will have to use a more sophisticated pairing function than
-  the Cartesian product we used before.
-  We instead will mirror the pattern in figure~\ref{cantor}.
-  To greatly simplify the algorithm, we will produce an intermediate stream of
-  lists which consists of the diagonals in the diagram.
-  We then concatenate these streams into the final support stream.
+  The main task here is to find a productive (in the coinductive sense) way to
+  combine the two support streams.
+  A trick which greatly simplifies the algorithm is to first produce a stream of
+  lists, which we then concatenate to a single stream.
+  This intermediate stream, which we will call \(C\), is generated from the two
+  support streams \(\mathit{xs}\) and \(\mathit{ys}\) like so:
+  \begin{equation}
+    C_i = \left\{ \left( \mathit{xs}_j , \mathit{ys}_{i - j} \right) \mid j \in [0,i] \right\}
+  \end{equation}
 \end{proof}
@@
 \todo[inline]{Partial search over infinite types}
+\section{Related Work}
+\cite{spiwackConstructivelyFinite2010} explored the notion of ``constructive
+finite sets'', demonstrating that there is a rich variety of finiteness
+predicates in a constructive setting.
+Some of those predicates where explored more in
+\cite{firsovVariationsNoetherianness2016a},
+\cite{parmannInvestigatingStreamlessSets2015}, and
+\cite{uustaluFinitenessRationalSequences2017a}.
+
+\cite{yorgeyCombinatorialSpeciesLabelled2014} explored some of the definitions
+we have given here (manifest bishop finiteness and cardinal finiteness) in HoTT,
+in order to construct species.
+Some of this paper is a formalisation of the early parts of that work.
+
+\cite{firsovDependentlyTypedProgramming2015} explores finite sets in Agda and
+uses them to automate proof machinery in much the same way as we have here.
+The primary difference between their proof automation and ours is that we can
+talk about finite functions, as we are working in HoTT.
+
+\cite{fruminFiniteSetsHomotopy2018} explores many of the same notions as we do
+here: cardinal finiteness, and Kuratowski finite sets.
+Our work makes extensive use of the computational properties of CuTT, however,
+and we are able to formalise the finiteness of function objects.
 \bibliographystyle{splncs04}

### 2020-01-10 paper.tex
finished

@@
 \end{rm-definition}
+This type is definitionally equal to it surjection equivalent (\(\mathbb{N}
+\twoheadrightarrow ! \; A\)).
 \subsubsection{Closure}
@@
 what's left.
-To do this we have to take a slightly different approach to the functions we
-defined before.
-Figure~\ref{pairings} illustrates the reason why: previously, we used the
-``Cartesian'' product pairing for each support list.
-This diverges if the first list is infinite, never exploring anything other than
-the first element in the second list.
-Instead, we use here the cantor pairing function, which performs a breadth-first
-search of the pairings of both lists.
 \input{figures/pairing-functions.tex}
@@
 \begin{proof}
-  The main task here is to find a productive (in the coinductive sense) way to
-  combine the two support streams.
-  A trick which greatly simplifies the algorithm is to first produce a stream of
-  lists, which we then concatenate to a single stream.
-  This intermediate stream, which we will call \(C\), is generated from the two
-  support streams \(\mathit{xs}\) and \(\mathit{ys}\) like so:
-  \begin{equation}
-    C_i = \left\{ \left( \mathit{xs}_j , \mathit{ys}_{i - j} \right) \mid j \in [0,i] \right\}
-  \end{equation}
-\end{proof}
-\begin{rm-lemma}
-  Split countability is closed under non-dependent product and sum.
-\end{rm-lemma}
-\begin{proof}
-  Follows from theorem~\ref{split-countability-sigma}.
+  The main task here is to figure out a pattern which will pair elements of the
+  two support streams without diverging.
+  Figure~\ref{cartesian} illustrates why the pairing function we used for lists
+  won't work: it needs to exhaust one of the input lists before inspecting
+  anything other than the first element of the other.
+  Of course, we can't exhaust a stream, so this diverges.
+
+  Instead, we will use the pattern in figure~\ref{cantor}.
+  We will produce a stream of lists (where each list is a diagonal), which is
+  then concatenated to the final output stream.
+  The intermediate stream is sometimes called the discrete convolution of the
+  two input streams.
 \end{proof}
@@
 In particular, we have the Kleene star.
-This means, in effect, that we have closure under lists.
+This means that we have closure under lists.
 \begin{rm-theorem}
@@
   As with the proof of closure under \(\Sigma\), our main task here is to figure
-  out a way to arrange the indices such that 
+  out a pattern that reaches any given finite list in finite time.
+  Our first step is to split up the problem as we did for \(\Sigma\)-closure: we
+  produce an intermediate stream of lists, which we then concatenate to give the
+  final result.
+  The intermediate stream is constructed like so:
+  \begin{equation}
+    \mathit{xs}^\star_i = \left\{ \text{map}(\mathit{xs} , x) \mid \mathit{x} \in \mathbf{List}(\mathbb{N}) , \text{sum}(\text{map}(1+, x)) \equiv i \right\}
+  \end{equation}
+  In other words, the \(i\)th element is all lists such that the sum of the
+  successor of the indices in the list is equal to \(i\).
 \end{proof}
-\subsection{Manifest Countability}
-As we can quotient out the position information with finite types, so can we with
-countable types.
-\todo[inline]{fill in rest here}
+\subsection{Other Types of Countability}
+There are of course other forms of countability available to us: manifest
+countability, cardinal, etc.
+We have a similar structure of relationships to the finite types, with the
+exception that we cannot construct an isomorphism with \(\mathbb{N}\) fiven a
+surjection from \(\mathbb{N}\).
 \section{Practical Uses} \label{practical}
@@
 \end{rm-theorem}
-\begin{proof}
-  \todo[inline]{Proof}
-\end{proof}
 
@@
 \end{rm-theorem}
-\begin{proof}
-  \todo[inline]{Proof}
-\end{proof}
 Finally, we do have a form of omniscience for prop-valued predicates, as they do
@@
 \end{rm-theorem}
-\begin{proof}
-  \todo[inline]{Proof}
-\end{proof}
 \subsection{Synthesising Pattern-Matching Proofs}
@@
 \ExecuteMetaData[agda/Data/Pauli.tex]{assoc-prf}
-
-\todo[inline]{Synthesise functions}
-\todo[inline]{Partial search over infinite types}
+\subsection{Synthesising Functions}
+Finally, thanks to extensionality provided to us by HoTT, and the computation
+properties provided by CuTT, we can derive decidable equality on functions over
+finite types.
+We can also use functions in our proof search.
+Here, for instance, is a automated procedure which finds the
+\(\AgdaFunction{not}\) function on \(\mathbf{Bool}\), given a specification.
+\ExecuteMetaData[agda/Data/Pauli.tex]{not-spec}
 \section{Related Work}

### 2020-01-13 paper.tex
dependent function types for cardinal finiteness

@@
 \begin{abstract}
-  We explore five notions of finiteness in Homotopy Type Theory \cite{hottbook}.
-  We will classify these notions, and prove their closure properties,
-  culminating in a proof that decidable Kuratowski-finite sets
-  \cite{kuratowskiNotionEnsembleFini1920} form a topos.
+  We study finiteness in Homotopy Type Theory, classifying five fundamentally
+  different finiteness predicates and proving formally the relationship between
+  them.
+  We also prove the closure properties of each predicate, which includes a proof
+  that decidable Kuratowski finite sets form a topos.
 
@@
 
-  We formalise our work in Cubical Agda
-  \cite{vezzosiCubicalAgdaDependently2019}, and we implement a library for proof
+  We formalise our work in Cubical Agda, and we implement a library for proof
   search (including combinators for level-polymorphic fully generic currying),

### 2020-01-13 paper.tex
new abstract

@@
 \input{preamble}
-\title{Finiteness and Cardinality in Homotopy Type Theory}
+\title{Finiteness in Cubical Type Theory}
 \author{Donnacha Oisín Kidney}
@@
 \begin{abstract}
-  We study finiteness in Homotopy Type Theory, classifying five fundamentally
-  different finiteness predicates and proving formally the relationship between
-  them.
-  We also prove the closure properties of each predicate, which includes a proof
-  that decidable Kuratowski finite sets form a topos.
+  We study five different notions of finiteness in Cubical Type Theory and prove
+  the relationship between them.
+  In particular we show that any totally ordered Kuratowski finite type is
+  manifestly Bishop finite.
+
+  We also prove closure properties for each finite type, and classify them
+  topos-theoretically.
+  For instance we show that the category of decidable Kuratowski finite sets
+  (also called the category of cardinal finite sets) form a \(\Pi\)-pretopos.
 

### 2020-01-14 paper.tex
finite isomorphisms

@@
 \section{Introduction}
+In constructive mathematics, we are often preoccupied with \emph{how} we know
+something.
+Finiteness is one of the prime examples which demonstrates this phenomenon: to
+show that some type is finite, we might provide a bijection with another
+finite set, or a surjection from some finite type.
+We could alternatively say that a type is finite if any collection of its
+elements larger than some bound contains duplicates, or if any stream of its
+elements contains duplicates.
+
+Classically, all of these ways to prove finiteness end up proving the same
+thing: constructively, however, each defines a fundamentally different
+``version'' of finiteness.
+In Martin-Löf type theory \cite{martin-lofIntuitionisticTypeTheory1980}, these
+variations (and more) have been explored in
+\cite{spiwackConstructivelyFinite2010} and
+\cite{firsovVariationsNoetherianness2016}.
+
+Homotopy Type Theory \cite{hottbook} further adds to this story in two ways.
+First, its universe of types includes non-sets: whether or not a finiteness
+predicate includes these gives us another dimension on which to discriminate.
+Secondly,
+
+classifying these types as
+finite or not adds
+
+can step in.
+Through primarily a generalisation of equality, we gain access to a number of
+exotic types which can, among other things, allow us to perform this ``hiding''.
+In the context of finiteness, the act of hiding is worth studying in its own
+right, and further delineates several new finiteness predicates which collapse
+even in standard
+
+The other headline feature of HoTT's generalised equality is univalence: put
+simply, it allows us to treat isomorphic types as equal, giving us in the
+constructive world access to a technique that is commonplace classically.
+It turns out that finite types have a lot to say about isomorphisms, and we will
+rely heavily on this aspect of HoTT to work with them.
+\subsection{Contributions}
+In this work, we will explore finite types in Cubical Type Theory
+\cite{cohenCubicalTypeTheory2016}, and expose their relationship to infinite
+types, species, and demonstrate their practical uses for proofs in dependently
+typed programming languages.
+
+In section~\ref{strong-finiteness}, we will explore the ``strong'' notions of
+finiteness (i.e. those at least as strong as Kuratowski finiteness
+\cite{kuratowskiNotionEnsembleFini1920}), with a special focus on cardinal
+finiteness.
+We will delineate some subtly different finiteness predicates, and prove the
+relationship between each.
+We will prove, then, that every discrete Kuratowski finite type is cardinal
+finite, and subsequently will show that discrete Kuratowski finite sets form a
+topos.
+
+In section~\ref{species-and-containers}, we will redefine our finiteness
+predicates using \emph{containers}, and prove that the new definitions are
+equivalent to the old.
+Using this, we will prove much stronger relationships between our finiteness
+predicates and relation-based definitions of finiteness.
+Finally, We will show how the species-container duality falls out naturally from
+these new definitions.
+
+In section~\ref{infinite-cardinalities}, we will extend our study of finite
+types to infinite but countable types.
+We will see that the finiteness predicates are mirrored with countable
+counterparts, and we will prove closure under the Kleene star and plus.
+
+Proofs of finiteness have well-known practical applications in
+constructive mathematics \cite{firsovDependentlyTypedProgramming2015}.
+In section~\ref{practical}, we build a library which exploits these uses in
+Cubical Agda \cite{vezzosiCubicalAgdaDependently2019}, allowing automation of
+complex proofs over finite types.
+We frame this in terms of the principle of omniscience for finite types.
+Thanks to the flexibility afforded to us by Cubical Type Theory, we are able to
+go further than the usual examples of this kind of proof automation: as well as
+proving properties about functions, we can synthesise functions whole-cloth from
+their desired properties.
+Through the unified interface for finite and countable types, we can
+reuse the automation machinery for \emph{partial} proof search over infinite
+search spaces.
+Along the way, we extend the work in \cite{allaisGenericLevelPolymorphic2019} to
+prove isomorphisms between the curried and uncurried forms of \(n\)-ary
+dependent functions.
+
+% Classically, of course, each of these statements are equivalent.
+% Constructively, though, all four predicates are fundamentally different.
+
+% Part of the difference comes from ``leaking'' information about the finite type
+% in the proof of its finiteness.
+% If I demonstrate a type is finite by showing a surjection to that type, I've
+% revealed a way to construct all of the elements of my type.
+% If I show that it is bounded, I reveal an upper bound on the cardinality of the
+% type.
+
 \subsection{Strong Finiteness Predicates}

### 2020-01-15 paper.tex
more

@@
 predicate includes these gives us another dimension on which to discriminate.
-Secondly,
 
-classifying these types as
-finite or not adds
+The second addition is concerned with the hiding of information.
+Occasionally in a constructive setting you don't want to have to reveal how you
+proved something.
+To model logical disjunction, for instance, we usually use a disjoint union: the
+model isn't entirely correct, though, because a proof of \(A \text{or} B\)
+reveals which of \(A\) or \(B\) we actually proved.
 
-can step in.
-Through primarily a generalisation of equality, we gain access to a number of
-exotic types which can, among other things, allow us to perform this ``hiding''.
-In the context of finiteness, the act of hiding is worth studying in its own
-right, and further delineates several new finiteness predicates which collapse
-even in standard
+logical disjunction, for instance, is usually modelled with a
+disjoint union.
+This always means, however, that to prove ``A or B'' you have to reveal which of
+A or B you did prove.
 

### 2020-01-16 paper.tex
more

@@
 In constructive mathematics, we are often preoccupied with \emph{how} we know
-something.
-Finiteness is one of the prime examples which demonstrates this phenomenon: to
-show that some type is finite, we might provide a bijection with another
-finite set, or a surjection from some finite type.
-We could alternatively say that a type is finite if any collection of its
-elements larger than some bound contains duplicates, or if any stream of its
-elements contains duplicates.
+something, in contrast to the classical focus on the \emph{what}.
+Take finiteness, for example.
+There are a handful of ways to demonstrate some type is finite: we could provide
+a bijection or a surjection from another finite type; we could show that any
+collection of its elements larger than some bound contains duplicates, or that
+any stream of its elements contain duplicates.
 
-Classically, all of these ways to prove finiteness end up proving the same
+Classically, all of these proofs end up proving the same thing: that our type is
+finite.
+Constructively, however, all \emph{four} of the statements above imply a
+different ``version'' of finiteness.
+\emph{How} we show that some type is finite has a significant impact on the
+scope of the things we can do with that type.
+
+That is the first thing this paper sets out to investigate: different finiteness
+predicates.
+The five in particular that we are interested in are organised in
+figure~\ref{finite-classification}.
+We especially want to examine what separates each predicate: what extra
+information do you need, for instance, to go from a proof of Cardinal finiteness
+(section~\ref{cardinal}) to manifest Bishop finiteness
+(section~\ref{manifest-bishop}).
+These predicates aren't just different in proof obligation, either: there are
+concrete types which can be shown to satisfy some, but not all, of these
+finiteness predicates.
+
+\input{figures/finite-classification}
+
+Next, we will examine the category-theoretic and topos-theoretic interpretations
+of each of the predicates.
+
+ways to prove finiteness end up proving the same
 thing: constructively, however, each defines a fundamentally different
@@
 \subsection{Strong Finiteness Predicates}
-\input{figures/finite-classification}
 We will first explore the ``strong'' notions of finiteness (i.e. those at least

### 2020-01-17 paper.tex
categories etc

@@
 (section~\ref{manifest-bishop}).
-These predicates aren't just different in proof obligation, either: there are
-concrete types which can be shown to satisfy some, but not all, of these
-finiteness predicates.
 
@@
 of each of the predicates.
+We will give a ``category of finite sets'' for each finiteness predicate, and
+then prove the various closure properties required to arrive at the relevant
+topos (we show that decidable Kuratowski finite sets form a \(\Pi\)-pretopos,
+for instance).
 

### 2020-01-17 paper.tex
roweorked beginning

--- 2020-01-15 paper.tex
@@
 \section{Introduction}
-In constructive mathematics, we are often preoccupied with \emph{how} we know
-something.
-Finiteness is one of the prime examples which demonstrates this phenomenon: to
-show that some type is finite, we might provide a bijection with another
-finite set, or a surjection from some finite type.
-We could alternatively say that a type is finite if any collection of its
-elements larger than some bound contains duplicates, or if any stream of its
-elements contains duplicates.
+\subsection{Foreword}
+In constructive mathematics we are often preoccupied with \emph{why} something
+is true.
+Take finiteness, for example.
+There are a handful of ways to demonstrate some type is finite: we could provide
+a surjection from another finite type; we could show that any collection of its
+elements larger than some bound contains duplicates; or that any stream of its
+elements contain duplicates.
 
-Classically, all of these ways to prove finiteness end up proving the same
-thing: constructively, however, each defines a fundamentally different
-``version'' of finiteness.
-In Martin-Löf type theory \cite{martin-lofIntuitionisticTypeTheory1980}, these
-variations (and more) have been explored in
-\cite{spiwackConstructivelyFinite2010} and
-\cite{firsovVariationsNoetherianness2016}.
+Classically, all of these proofs end up proving the same thing: that our type is
+finite.
+Constructively (in Martin-Löf Type Theory
+\cite{martin-lofIntuitionisticTypeTheory1980} at least), however, all three
+of the statements above construct a different version of finiteness.
+\emph{How} we show that some type is finite has a significant impact on the
+type of finiteness we end up dealing with.
 
-Homotopy Type Theory \cite{hottbook} further adds to this story in two ways.
-First, its universe of types includes non-sets: whether or not a finiteness
-predicate includes these gives us another dimension on which to discriminate.
+Homotopy Type Theory \cite{hottbook} adds another wrinkle to the story.
+Firstly, in HoTT we cannot assume that every type is a (homotopy) set: this
+means that the finiteness predicates above can be further split into versions
+which apply to sets only, and those that apply to all types.
+Secondly, HoTT gives us a principled and powerful way to construct quotients,
+allowing us to regain some of the flexibility of classical definitions by
+``forgetting'' the parts of a proof we would be forced to remember in MLTT.
+\subsection{Contributions}
+\begin{itemize}
+  \item Relation between different versions of finiteness, especially card to
+    bishop.
+  \item Topos for Kuratowski
+  \item Countable and W-topos
+  \item Search
+\end{itemize}
+\subsection{Related Work}
+\subsection{Notation}
 
-The second addition is concerned with the hiding of information.
-Occasionally in a constructive setting you don't want to have to reveal how you
-proved something.
-To model logical disjunction, for instance, we usually use a disjoint union: the
-model isn't entirely correct, though, because a proof of \(A \text{or} B\)
-reveals which of \(A\) or \(B\) we actually proved.
+% Taking finiteness as an example again, if we prove that some type is finite via
+% bijection, in MLTT such a proof would have to reveal the particular bijection.
+% This becomes problematic when defining, for instance, the category of finite
+% sets: the objects of this category would ideally be types coupled with proofs of
+% their finiteness, but in order to
 
-logical disjunction, for instance, is usually modelled with a
-disjoint union.
-This always means, however, that to prove ``A or B'' you have to reveal which of
-A or B you did prove.
+% The five in particular that we are interested in are organised in
+% figure~\ref{finite-classification}.
+% We especially want to examine what separates each predicate: what extra
+% information do you need, for instance, to go from a proof of Cardinal finiteness
+% (section~\ref{cardinal}) to manifest Bishop finiteness
+% (section~\ref{manifest-bishop}).
 
-The other headline feature of HoTT's generalised equality is univalence: put
-simply, it allows us to treat isomorphic types as equal, giving us in the
-constructive world access to a technique that is commonplace classically.
-It turns out that finite types have a lot to say about isomorphisms, and we will
-rely heavily on this aspect of HoTT to work with them.
-\subsection{Contributions}
-In this work, we will explore finite types in Cubical Type Theory
-\cite{cohenCubicalTypeTheory2016}, and expose their relationship to infinite
-types, species, and demonstrate their practical uses for proofs in dependently
-typed programming languages.
+\input{figures/finite-classification}
 
-In section~\ref{strong-finiteness}, we will explore the ``strong'' notions of
-finiteness (i.e. those at least as strong as Kuratowski finiteness
-\cite{kuratowskiNotionEnsembleFini1920}), with a special focus on cardinal
-finiteness.
-We will delineate some subtly different finiteness predicates, and prove the
-relationship between each.
-We will prove, then, that every discrete Kuratowski finite type is cardinal
-finite, and subsequently will show that discrete Kuratowski finite sets form a
-topos.
+% Next, we will examine the category-theoretic and topos-theoretic interpretations
+% of each of the predicates.
+% We will give a ``category of finite sets'' for each finiteness predicate, and
+% then prove the various closure properties required to arrive at the relevant
+% topos (we show that decidable Kuratowski finite sets form a \(\Pi\)-pretopos,
+% for instance).
 
-In section~\ref{species-and-containers}, we will redefine our finiteness
-predicates using \emph{containers}, and prove that the new definitions are
-equivalent to the old.
-Using this, we will prove much stronger relationships between our finiteness
-predicates and relation-based definitions of finiteness.
-Finally, We will show how the species-container duality falls out naturally from
-these new definitions.
+% ways to prove finiteness end up proving the same
+% thing: constructively, however, each defines a fundamentally different
+% ``version'' of finiteness.
+% In Martin-Löf type theory \cite{martin-lofIntuitionisticTypeTheory1980}, these
+% variations (and more) have been explored in
+% \cite{spiwackConstructivelyFinite2010} and
+% \cite{firsovVariationsNoetherianness2016}.
 
-In section~\ref{infinite-cardinalities}, we will extend our study of finite
-types to infinite but countable types.
-We will see that the finiteness predicates are mirrored with countable
-counterparts, and we will prove closure under the Kleene star and plus.
+% Homotopy Type Theory \cite{hottbook} further adds to this story in two ways.
+% First, its universe of types includes non-sets: whether or not a finiteness
+% predicate includes these gives us another dimension on which to discriminate.
 
-Proofs of finiteness have well-known practical applications in
-constructive mathematics \cite{firsovDependentlyTypedProgramming2015}.
-In section~\ref{practical}, we build a library which exploits these uses in
-Cubical Agda \cite{vezzosiCubicalAgdaDependently2019}, allowing automation of
-complex proofs over finite types.
-We frame this in terms of the principle of omniscience for finite types.
-Thanks to the flexibility afforded to us by Cubical Type Theory, we are able to
-go further than the usual examples of this kind of proof automation: as well as
-proving properties about functions, we can synthesise functions whole-cloth from
-their desired properties.
-Through the unified interface for finite and countable types, we can
-reuse the automation machinery for \emph{partial} proof search over infinite
-search spaces.
-Along the way, we extend the work in \cite{allaisGenericLevelPolymorphic2019} to
-prove isomorphisms between the curried and uncurried forms of \(n\)-ary
-dependent functions.
+% The second addition is concerned with the hiding of information.
+% Occasionally in a constructive setting you don't want to have to reveal how you
+% proved something.
+% To model logical disjunction, for instance, we usually use a disjoint union: the
+% model isn't entirely correct, though, because a proof of \(A \text{or} B\)
+% reveals which of \(A\) or \(B\) we actually proved.
+
+% logical disjunction, for instance, is usually modelled with a
+% disjoint union.
+% This always means, however, that to prove ``A or B'' you have to reveal which of
+% A or B you did prove.
+
+% The other headline feature of HoTT's generalised equality is univalence: put
+% simply, it allows us to treat isomorphic types as equal, giving us in the
+% constructive world access to a technique that is commonplace classically.
+% It turns out that finite types have a lot to say about isomorphisms, and we will
+% rely heavily on this aspect of HoTT to work with them.
+% \subsection{Contributions}
+% In this work, we will explore finite types in Cubical Type Theory
+% \cite{cohenCubicalTypeTheory2016}, and expose their relationship to infinite
+% types, species, and demonstrate their practical uses for proofs in dependently
+% typed programming languages.
+
+% In section~\ref{strong-finiteness}, we will explore the ``strong'' notions of
+% finiteness (i.e. those at least as strong as Kuratowski finiteness
+% \cite{kuratowskiNotionEnsembleFini1920}), with a special focus on cardinal
+% finiteness.
+% We will delineate some subtly different finiteness predicates, and prove the
+% relationship between each.
+% We will prove, then, that every discrete Kuratowski finite type is cardinal
+% finite, and subsequently will show that discrete Kuratowski finite sets form a
+% topos.
+
+% In section~\ref{species-and-containers}, we will redefine our finiteness
+% predicates using \emph{containers}, and prove that the new definitions are
+% equivalent to the old.
+% Using this, we will prove much stronger relationships between our finiteness
+% predicates and relation-based definitions of finiteness.
+% Finally, We will show how the species-container duality falls out naturally from
+% these new definitions.
+
+% In section~\ref{infinite-cardinalities}, we will extend our study of finite
+% types to infinite but countable types.
+% We will see that the finiteness predicates are mirrored with countable
+% counterparts, and we will prove closure under the Kleene star and plus.
+
+% Proofs of finiteness have well-known practical applications in
+% constructive mathematics \cite{firsovDependentlyTypedProgramming2015}.
+% In section~\ref{practical}, we build a library which exploits these uses in
+% Cubical Agda \cite{vezzosiCubicalAgdaDependently2019}, allowing automation of
+% complex proofs over finite types.
+% We frame this in terms of the principle of omniscience for finite types.
+% Thanks to the flexibility afforded to us by Cubical Type Theory, we are able to
+% go further than the usual examples of this kind of proof automation: as well as
+% proving properties about functions, we can synthesise functions whole-cloth from
+% their desired properties.
+% Through the unified interface for finite and countable types, we can
+% reuse the automation machinery for \emph{partial} proof search over infinite
+% search spaces.
+% Along the way, we extend the work in \cite{allaisGenericLevelPolymorphic2019} to
+% prove isomorphisms between the curried and uncurried forms of \(n\)-ary
+% dependent functions.
 

### 2020-01-18 paper.tex
more

@@
   \email{o.kidney@cs.ucc.ie}}
+\author{Gregory Provan}
+\institute{University College Cork \and
+  \email{g.provan@cs.ucc.ie}}
 \begin{document}
@@
 ``forgetting'' the parts of a proof we would be forced to remember in MLTT.
+
+Finally, the other important property of constructive mathematics is that we can
+actually \emph{compute} with proofs.
+Cubical Type Theory \cite{cohenCubicalTypeTheory2016}, and its implementation in
+Cubical Agda \cite{vezzosiCubicalAgdaDependently2019}, realise this property
+even in the presence of univalence, giving computational content to programs
+written in HoTT.
 \subsection{Contributions}
+In this work we will examine five notions of finiteness in Homotopy Type Theory:
+the relationships between them, and their topos-theoretic characterisation.
+We also briefly examine a definition for countable sets, comparing it to the
+manifestly Bishop finite sets.
+Our work is formalised in Cubical Agda, where we also develop a library for
+proof search based on the finiteness predicates.
+
+\input{figures/finite-classification}
+
+The finiteness predicates we are interested in are organised in
+Figure~\ref{finite-classification}.
+We will provide a definition for each, and prove each relation to the other
+finiteness predicates.
+In terms of Figure~\ref{finite-classification}, this means providing a function
+to inhabit each arrows.
+
+We will then characterise each predicate as some form of topos.
+Our proofs follow the structure of \cite[Chapters 9, 10]{hottbook} and
+\cite{rijkeSetsHomotopyType2015}.
+This means we will define first the precategory of sets, and then the category
+of sets, and prove the required closures and limits to get us to a topos.
+We say ``some form of'' topos here because of course the category of sets in
+HoTT do \emph{not}, in fact, form a topos, rather a \(\Pi W\)-pretopos.
+
 \begin{itemize}
@@
 % (section~\ref{manifest-bishop}).
-
-\input{figures/finite-classification}
 

### 2020-01-18 paper.tex
more

@@
 \title{Finiteness in Cubical Type Theory}
-\author{Donnacha Oisín Kidney}
-\authorrunning{D. O. Kidney}
-\institute{University College Cork \and
-  \email{o.kidney@cs.ucc.ie}}
-\author{Gregory Provan}
-\institute{University College Cork \and
-  \email{g.provan@cs.ucc.ie}}
+\author{
+  Donnacha Oisín Kidney \inst{1} \and
+  Gregory Provan \inst{2}
+}
+\institute{
+  University College Cork \email{o.kidney@cs.ucc.ie} \and
+  University College Cork \email{g.provan@cs.ucc.ie}
+}
 \begin{document}
@@
 \subsection{Contributions}
-In this work we will examine five notions of finiteness in Homotopy Type Theory:
+In this work we will examine five notions of finiteness in Homotopy Type Theory,
 the relationships between them, and their topos-theoretic characterisation.
@@
 Figure~\ref{finite-classification}.
-We will provide a definition for each, and prove each relation to the other
-finiteness predicates.
-In terms of Figure~\ref{finite-classification}, this means providing a function
-to inhabit each arrows.
+We will explore two aspects of each predicate: its relation with the other
+predicates, and its topos-theoretic classification.
+
+When we say ``relation'' we are referring to the arrows in
+Figure~\ref{finite-classification}.
+The proofs, then, amount to a function which inhabits each arrow.
+Each unlabelled arrow is a weakening: i.e., every manifest Bishop finite set
+is manifest enumerable.
+The labelled arrows are strengthening proofs: i.e. every manifest enumerable set
+\emph{with decidable equality} is manifest Bishop finite.
 
@@
 
-\begin{itemize}
-  \item Relation between different versions of finiteness, especially card to
-    bishop.
-  \item Topos for Kuratowski
-  \item Countable and W-topos
-  \item Search
-\end{itemize}
+\todo[inline]{redo these last two paragraphs}
+After the finite predicates, we will briefly look at the infinite countable
+types, and classify them in a parallel way to the finite predicates.
+
+All of our work is formalised in Cubical Agda
+\cite{vezzosiCubicalAgdaDependently2019}.
+We will make mention of the few occasions where the formalisation of some proof
+is of interest, but the main place where we will discuss the code is in the
+final section, where we implement a library for proof search, based on
+omniscience and exhaustibility.
+We also use this section to demonstrate some \emph{computation} using
+univalence: our proof search library can deal seamlessly with functions, testing
+them for equality, synthesising them, etc.
 \subsection{Related Work}
@@
 
-% Taking finiteness as an example again, if we prove that some type is finite via
-% bijection, in MLTT such a proof would have to reveal the particular bijection.
-% This becomes problematic when defining, for instance, the category of finite
-% sets: the objects of this category would ideally be types coupled with proofs of
-% their finiteness, but in order to
-
-% The five in particular that we are interested in are organised in
-% figure~\ref{finite-classification}.
-% We especially want to examine what separates each predicate: what extra
-% information do you need, for instance, to go from a proof of Cardinal finiteness
-% (section~\ref{cardinal}) to manifest Bishop finiteness
-% (section~\ref{manifest-bishop}).
-
-% Next, we will examine the category-theoretic and topos-theoretic interpretations
-% of each of the predicates.
-% We will give a ``category of finite sets'' for each finiteness predicate, and
-% then prove the various closure properties required to arrive at the relevant
-% topos (we show that decidable Kuratowski finite sets form a \(\Pi\)-pretopos,
-% for instance).
-
-% ways to prove finiteness end up proving the same
-% thing: constructively, however, each defines a fundamentally different
-% ``version'' of finiteness.
-% In Martin-Löf type theory \cite{martin-lofIntuitionisticTypeTheory1980}, these
-% variations (and more) have been explored in
-% \cite{spiwackConstructivelyFinite2010} and
-% \cite{firsovVariationsNoetherianness2016}.
-
-% Homotopy Type Theory \cite{hottbook} further adds to this story in two ways.
-% First, its universe of types includes non-sets: whether or not a finiteness
-% predicate includes these gives us another dimension on which to discriminate.
-
-% The second addition is concerned with the hiding of information.
-% Occasionally in a constructive setting you don't want to have to reveal how you
-% proved something.
-% To model logical disjunction, for instance, we usually use a disjoint union: the
-% model isn't entirely correct, though, because a proof of \(A \text{or} B\)
-% reveals which of \(A\) or \(B\) we actually proved.
-
-% logical disjunction, for instance, is usually modelled with a
-% disjoint union.
-% This always means, however, that to prove ``A or B'' you have to reveal which of
-% A or B you did prove.
-
-% The other headline feature of HoTT's generalised equality is univalence: put
-% simply, it allows us to treat isomorphic types as equal, giving us in the
-% constructive world access to a technique that is commonplace classically.
-% It turns out that finite types have a lot to say about isomorphisms, and we will
-% rely heavily on this aspect of HoTT to work with them.
-% \subsection{Contributions}
-% In this work, we will explore finite types in Cubical Type Theory
-% \cite{cohenCubicalTypeTheory2016}, and expose their relationship to infinite
-% types, species, and demonstrate their practical uses for proofs in dependently
-% typed programming languages.
-
-% In section~\ref{strong-finiteness}, we will explore the ``strong'' notions of
-% finiteness (i.e. those at least as strong as Kuratowski finiteness
-% \cite{kuratowskiNotionEnsembleFini1920}), with a special focus on cardinal
-% finiteness.
-% We will delineate some subtly different finiteness predicates, and prove the
-% relationship between each.
-% We will prove, then, that every discrete Kuratowski finite type is cardinal
-% finite, and subsequently will show that discrete Kuratowski finite sets form a
-% topos.
-
-% In section~\ref{species-and-containers}, we will redefine our finiteness
-% predicates using \emph{containers}, and prove that the new definitions are
-% equivalent to the old.
-% Using this, we will prove much stronger relationships between our finiteness
-% predicates and relation-based definitions of finiteness.
-% Finally, We will show how the species-container duality falls out naturally from
-% these new definitions.
-
-% In section~\ref{infinite-cardinalities}, we will extend our study of finite
-% types to infinite but countable types.
-% We will see that the finiteness predicates are mirrored with countable
-% counterparts, and we will prove closure under the Kleene star and plus.
-
-% Proofs of finiteness have well-known practical applications in
-% constructive mathematics \cite{firsovDependentlyTypedProgramming2015}.
-% In section~\ref{practical}, we build a library which exploits these uses in
-% Cubical Agda \cite{vezzosiCubicalAgdaDependently2019}, allowing automation of
-% complex proofs over finite types.
-% We frame this in terms of the principle of omniscience for finite types.
-% Thanks to the flexibility afforded to us by Cubical Type Theory, we are able to
-% go further than the usual examples of this kind of proof automation: as well as
-% proving properties about functions, we can synthesise functions whole-cloth from
-% their desired properties.
-% Through the unified interface for finite and countable types, we can
-% reuse the automation machinery for \emph{partial} proof search over infinite
-% search spaces.
-% Along the way, we extend the work in \cite{allaisGenericLevelPolymorphic2019} to
-% prove isomorphisms between the curried and uncurried forms of \(n\)-ary
-% dependent functions.
-
-% Classically, of course, each of these statements are equivalent.
-% Constructively, though, all four predicates are fundamentally different.
-
-% Part of the difference comes from ``leaking'' information about the finite type
-% in the proof of its finiteness.
-% If I demonstrate a type is finite by showing a surjection to that type, I've
-% revealed a way to construct all of the elements of my type.
-% If I show that it is bounded, I reveal an upper bound on the cardinality of the
-% type.
-
-\subsection{Strong Finiteness Predicates}
-We will first explore the ``strong'' notions of finiteness (i.e. those at least
-as strong as Kuratowski finiteness \cite{kuratowskiNotionEnsembleFini1920}),
-with a special focus on cardinal finiteness (section~\ref{cardinal}), and
-manifest enumerability (section~\ref{manifest-enumerability}), which is new, to
-our knowledge.
-
-Figure~\ref{finite-classification} organises the predicates according to
-their strength.
-We will go through each of the predicates, proving how to weaken each (i.e. we
-will provide a proof that every cardinally finite type is Kuratowski finite),
-and how to strengthen them, given the required property.
-In terms of figure~\ref{finite-classification}, this amounts to providing proofs
-for each arrow.
-Our main proof in this regard is that we can derive manifest Bishop finiteness
-from Kuratowski finiteness plus a total order.
-
-We will---through the use of containers
-\cite{abbottContainersConstructingStrictly2005}---formally prove the equivalence
-these predicates have with the usual function relations i.e. we will show that a
-proof of manifest enumerability is precisely equivalent to a surjection from a
-finite prefix of the natural numbers.
-
-For each predicate, we will also prove its closure over sums and products in
-both dependent and non-dependent forms, if such a closure exists.
-This will culminate in our main result for this section: the formal proof that
-decidable Kuratowski finite sets form a topos.
-\subsection{Infinite Types}
-In section~\ref{infinite-cardinalities}, we will extend our study of finite
-types to infinite but countable types.
-We will see that the classification of finiteness predicates mirrors that of
-their countable counterparts, and we will prove closure under the Kleene star.
-
-\subsection{Practical Uses}
-Finally, in section~\ref{practical} we will demonstrate some practical uses of
-the finiteness proofs in Cubical Agda.
-We will show how to use well-known techniques
-\cite{firsovDependentlyTypedProgramming2015, devrieseBrightSideType2011} to
-automate proofs of functions with finite (and infinite) domains.
-We further show how to automate the synthesis of functions from desired
-properties.
-\section{Split Enumerability} \label{split-enumerable}
-We will start with the simplest definition of finiteness: we say a set is
-enumerable if there is a list of its elements which contains every element in
-the set.
-
-Before giving the definition of the predicate, we will first define lists (and
-membership thereof): we have chosen a \emph{container}-based definition for this
-work.
-\begin{rm-definition}[Container] \label{container-def}
-  A container \cite{abbottContainersConstructingStrictly2005} is a pair
-  \(S \triangleright P\) where \(S\) is a type, the elements of which are called
-  the \emph{shapes} of the container, and \(P\) is a type family on \(S\), where
-  the elements of \(P(s)\) are called the \emph{positions} of a container.
-  We ``interpret'' a container into a functor defined like so:
-  \begin{equation}
-    \llbracket S \triangleright P \rrbracket(A) = \Sigma {(s : S)} , \left( P(s) \rightarrow A \right)
-  \end{equation}
-  Membership of a container can be defined like so:
-  \begin{equation}
-    x \in \mathit{xs} = \text{fiber}(\text{snd}(\mathit{xs}), x)
-  \end{equation}
-  Where \(\text{fiber}\) is from \cite[definition 4.2.4]{hottbook}.
-\end{rm-definition}
-\begin{rm-definition}[Lists]
-  \begin{equation}%
-    \mathbf{List} = \llbracket \mathbb{N} \triangleright \mathbf{Fin} \rrbracket
-  \end{equation}
-\end{rm-definition}
-\begin{rm-definition}[\mathbf{Fin}]
-  \(\mathbf{Fin}(n)\) is the type of natural numbers smaller than \(n\).
-  We define it the standard way, where \(\mathbf{Fin}(0) = \bot\) and
-  \(\mathbf{Fin}(n + 1) &= \top + \mathbf{Fin}(n)\).
-\end{rm-definition}
-
-Internally, in our formalisation, we actually use the standard
-inductive definition of lists more often (it tends to work better in more
-complex algorithms, and functions on it seems to satisfy the termination
-checker more readily).
-However, since both types are equivalent, univalence allows us to transport to
-whichever representation is more convenient in a given situation.
-For the higher-level proofs we present here, though, the container-based
-definition greatly simplifies certain steps, which is why we have chosen it as
-our representation.
-
-Finally, we can define formally split enumerability.
+\section{Finiteness Predicates}
+\subsection{Split Enumerable}
 \begin{rm-definition}[Split Enumerable Set]
@@
   This definition is what is referred to as \verb+Listable+ in
+  \cite{firsovDependentlyTypedProgramming2015}.
 \end{rm-definition}
-\cite{firsovDependentlyTypedProgramming2015}.
-\subsection{Split Surjections}
-\begin{rm-definition}[Surjections]
-  \begin{multicols}{2}
-    \begin{equation}
-      A \twoheadrightarrow B = \Sigma (f : A \rightarrow B) , \text{surjective}(f)
-    \end{equation} \break
-    \begin{equation}
-      A \twoheadrightarrow! \; B = \Sigma (f : A \rightarrow B) , \text{split surjective}(f)
-    \end{equation}
-  \end{multicols}
-  Where the definitions of surjective are taken from \cite[definition
-  4.6.1]{hottbook}.
-\end{rm-definition}
-\begin{rm-theorem} \label{split-enum-surj}
-  Split enumerability is equivalent to a split surjection from a finite prefix
-  of the natural numbers.
-  \begin{equation}
-    \mathcal{E}!(A) \iff \Sigma {(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \twoheadrightarrow ! \; A \right)
-  \end{equation}
-\end{rm-theorem}
-\begin{proof}
-  After sufficient inlining, it emerges that our goal is simply a reassociation.
-  \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{split-surj}
-  To be clear: in Agda, the proof could simply be \(\AgdaFunction{reassoc}\); we
-  have written out the extra lines for clarity alone.
-\end{proof}
-\begin{rm-lemma} \label{split-enum-discrete}
-  Any split enumerable type has decidable equality (is discrete).
-\end{rm-lemma}
-\begin{proof}
-  We use a corollary that if there is a split-surjection from \(A\) to \(B\),
-  and \(A\) is discrete, then \(B\) is also discrete.
-\end{proof}
-By Hedberg's theorem \cite{hedbergCoherenceTheoremMartinLof1998}, we know that
-this implies split enumerable types must all be sets.
-\subsection{Closure}
-In this section we will prove closure under various operations for split
-enumerable sets.
-We are working towards a topos proof, which requires us to prove closure under
-a variety of operations: for now, we only have enough machinery to demonstrate
-the semiring operations, and dependent sums.
-In order to show closure under exponentials (function arrows), we will need an
-equivalence with \(\mathbf{Fin}\), which will be provided in
-section~\ref{manifest-bishop}.
-\begin{rm-lemma} \label{split-enum-bool-top-bot}
-  \(\mathbf{Bool}\), \(\top\), and \(\bot\) are all split enumerable.
-\end{rm-lemma}
-\begin{proof}
-  Each of these types clearly has a finite number of elements (2, 1, and 0,
-  respectively), and furthermore has a straightforward enumeration.
-\end{proof}
-\begin{rm-theorem} \label{split-enum-sigma}
-  Split-enumerability is closed under \(\Sigma\).
-  \begin{equation} \label{split-enum-sigma-clos-sig}
-    \frac{
-      \mathcal{E}!(A) \; \; \; \Pi (x : A) , \mathcal{E}!(U(x))
-    }{
-      \mathcal{E}!(\Sigma (x : A) , U(x))
-    }
-  \end{equation}
-\end{rm-theorem}
-\begin{proof}
-  To obtain the support list, we concatenate the support lists of all the proofs
-  of split-finiteness for \(U\) over the support list of \(E_A\).
-  In Agda:
-  \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{sigma-sup}
-  ``do-notation'' is available to us as we're working in the list monad.
-\end{proof}
-Closure under disjoint union and Cartesian product both follow from \(\Sigma\),
-as these types can be defined in terms of \(\Sigma\).
-\begin{align}
-  A \times B &= \Sigma {(x : A)} , B \\
-  A + B &= \Sigma {(x : \mathbf{Bool})} , \text{if} \; x \; \text{then} \; A \; \text{else} \; B
-\end{align}
-\section{Manifest Bishop Finiteness} \label{manifest-bishop}
-Where split enumerability was the enumeration form of a split surjection from
-\(\mathbf{Fin}\), manifest Bishop finiteness is the enumeration form of an
-\emph{equivalence} with \(\mathbf{Fin}\).
-
-\begin{rm-definition}[Manifest Bishop Finiteness]
-  \begin{equation}
-    \mathcal{B}(A) = \Sigma {(\mathit{xs} : \textbf{List}(A))} , \Pi {(x : A)} , x \inunique xs
-  \end{equation}
-\end{rm-definition}
-The only difference between this predicate and split enumerability is the list
-membership term: we use \(\inunique\) here, where \(x \inunique \mathit{xs}\) is
-to be read as ``\(x\) occurs exactly once in \(\mathit{xs}\)''.
-\begin{rm-definition}[Unique Membership]
-  We say an item \(x\) is ``uniquely in'' some container \(\mathit{xs}\) if its
-  membership in that list is a \emph{contraction} \cite[definition
-  3.11.1]{hottbook}; i.e. its membership proof exists, and all such proofs are
-  equal.
-  \begin{equation}
-    x \inunique \mathit{xs} = \text{isContr}(x \in \mathit{xs})
-  \end{equation}
-\end{rm-definition}
-By the definition of a contraction, we can always recover the underlying
-membership proof, meaning that we can always derive split enumerability from
-manifest Bishop finiteness.
-\subsection{Equivalence}
-\begin{rm-lemma} \label{bishop-equiv}
-  A proof of manifest Bishop finiteness is equivalent to an equivalence with a
-  finite prefix of the natural numbers.
-  \begin{equation}
-    \mathcal{B}(A) \simeq \Sigma {(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \simeq A \right)
-  \end{equation}
-\end{rm-lemma}
-\begin{proof}
-  There are many equivalent definitions of equivalence in HoTT.
-  Here we take the version preferred in the Cubical Agda library: contractible
-  maps \cite[definition 4.4.1]{hottbook}.
-  Because of the parallels between contractible maps and split surjections,
-  the proof proceeds much the same as \ref{split-enum-surj}.
-  In other words, the definition of Bishop finiteness is itself a reassociation
-  of a contractible map.
-\end{proof}
-\subsection{Relationship to Split Enumerability}
-\begin{rm-theorem} \label{split-enum-to-manifest-bishop}
-  Any split enumerable set is manifest Bishop finite.
-\end{rm-theorem}
-\begin{proof}
-  From proposition~\ref{split-enum-discrete} we can derive decidable equality on
-  \(A\), and using this we can define a function (called \(\text{uniques}\)) which filters
-  out duplicates from lists of \(A\)s.
-  This gives us our support list.
-  To generate the cover proof it suffices now to prove the following:
-  \begin{equation}
-    \Pi {(x : A)} , \Pi {(\mathit{xs} : \textbf{List}(A))} , x \in \mathit{xs} \rightarrow x \inunique \text{uniques}(\mathit{xs})
-  \end{equation}
-\end{proof}
-\subsection{Closure}
-Proving equal strength of split enumerability and manifest Bishop finiteness
-allows us to carry all of the previous proofs of closure over to manifest Bishop
-finite sets (and vice-versa).
-Missing from our previous proofs was a proof of closure of functions.
-We remedy that here.
-\begin{rm-theorem}
-  Manifest bishop finiteness is closed over dependent functions
-  (\(\prod\)-types).
-  \begin{equation}
-    \frac{
-      \mathcal{B}(A) \; \; \; \Pi {(x : A)} , \mathcal{B}\left( U(x) \right)
-    }{
-      \mathcal{B}\left(\Pi {(x : A)} , U(x)\right)
-    }
-  \end{equation}
-\end{rm-theorem}
-\begin{proof}
-  This proof is essentially the composition of two transport operations, made
-  available to us via univalence.
-
-  First, we will simplify things slightly by working only with split
-  enumerability.
-  As this is equal in strength to manifest Bishop finiteness, any closure proofs
-  carry over.
-
-  Secondly, we will replace \(A\) in all places with \(\mathbf{Fin}(n)\).
-  Since we have already seen an equivalence between these two types, we are
-  permitted to transport along these lines.
-  This is the first transport operation.
-
-  The bulk of the proof now is concerned with proving the following:
-  \begin{equation}
-    \left( \Pi {(x : \mathbf{Fin}(n))} , \mathcal{E}!(A(x)) \right) \rightarrow \mathcal{E}!\left( \Pi {(x : \mathbf{Fin}(n))} , A(x) \right)
-  \end{equation}
-  Our strategy to accomplish this will be to consider functions from
-  \(\mathbf{Fin}(n)\) as \(n\)-tuples over some type family \(T : \mathbb{N}
-  \rightarrow \text{Type}\).
-  \begin{equation}
-    \begin{aligned}
-      \mathbf{Tuple}(T, 0)   &= \top \\
-      \mathbf{Tuple}(T, n+1) &= T(0) \times \mathbf{Tuple}(T \circ \text{suc}, n)
-    \end{aligned}
-  \end{equation}
-  This type is manifestly Bishop finite, as it is constructed only from products
-  and the unit type.
-
-  We then prove an isomorphism between this representation and \(\Pi\)-types.
-  \begin{equation}
-    \mathbf{Tuple}(T, n) \iff \Pi {(x : \mathbf{Fin}(n))} , T(x)
-  \end{equation}
-  This allows us to transport our proof of finiteness on tuples to one on
-  functions from \(\mathbf{Fin}\) (our second transport operation), proving our
-  goal.
-\end{proof}
-\section{Manifest Enumerability} \label{manifest-enumerability}
-Both split enumerability and manifest Bishop finiteness are restricted to sets:
-because they both imply decidable equality, no non-set types can satisfy them as
-predicates.
-To find a more general predicate for finiteness, we \emph{truncate} the
-membership proof.
-\begin{rm-definition}[Manifest Enumerability]
-  \begin{equation}
-    \mathcal{E}(A) = \Sigma {(\mathit{xs} : \mathbf{List}(A))} , \Pi {(x : A)} , \lVert x \in \mathit{xs} \rVert
-  \end{equation}
-\end{rm-definition}
-\begin{rm-definition}[Propositional Truncation]a \label{prop-trunc}
-  The type \(\lVert A \rVert\) on some type \(A\) is a propositionally truncated
-  proof of \(A\) \cite[3.7]{hottbook}.
-  In other words, it is a proof that some \(A\) exists, but it does not tell you
-  \emph{which} \(A\).
-
-  It is defined as a Higher Inductive Type:
-  \begin{equation}
-    \begin{aligned}
-      \lVert A \rVert =& \; \lvert \cdot \rvert : A \rightarrow \lVert A \rVert ; \\
-                      |& \; \text{squash} : \Pi {(x, y : \lVert A \rVert)} , x \equiv y  ; 
-    \end{aligned}
-  \end{equation}
-
-  We will use three eliminators from \(\lVert A \rVert\) in this paper.
-  \begin{enumerate}
-  \item \label{elim-prop-prop} For any function \(A \rightarrow B\), where
-    \(\text{isProp}(B)\), we have a function \(\lVert A \rVert \rightarrow B\).
-  \item \label{elim-prop-monad} A special case of \ref{elim-prop-prop} implies
-    that \(\lVert \cdot \rVert\) forms a monad: this means that we get the usual
-    Monadic operators on \(\lVert \cdot \rVert\) (bind, pure, fmap, etc.).
-    \item \label{elim-prop-coh} We can eliminate from \(\lVert A \rVert\) with a
-      function \(f : A \rightarrow B\) iff \(f\) ``doesn't care'' about the
-      choice of \(A\) (\(\Pi {(x , y : A)} , f(x) \equiv f(y) \)).
-      Formally speaking, \(f\) needs to be ``coherently constant''
-      \cite{krausGeneralUniversalProperty2015}, and \(B\) needs to be an
-      \(n\)-type for some finite \(n\).
-  \end{enumerate}
-\end{rm-definition}
-
-We will use the circle as an example non-set type which is nonetheless
-manifestly enumerable.
-\begin{rm-theorem}
-  The circle \(S^1\) is manifestly enumerable.
-\end{rm-theorem}
-\begin{proof}
-  As the cover proof is a truncated proposition, we need only consider the point
-  constructors, making this poof the same as the proof of split enumerability on
-  \(\top\).
-\end{proof}
-\subsection{Surjections}
-This predicates relation to surjectivity is much the same as split
-enumerability's relation to \emph{split} surjectivity.
-\begin{rm-lemma}
-  A proof of manifest enumerability is equivalent to a surjection from a finite
-  prefix of the natural numbers.
-\end{rm-lemma}
-\subsection{Relation to Split Enumerability}
-\begin{rm-theorem} \label{manifest-enum-to-split}
-  A manifestly enumerable type with decidable equality is split enumerable.
-\end{rm-theorem}
-\begin{proof}
-  The support list stays the same between both enumerability proofs.
-
-  For the cover proof, we first need the following function which searches a
-  list for a particular element (given decidable equality on \(A\)).
-  \begin{equation}
-    \decin \; : \Pi(x : A) , \Pi(\mathit{xs} : \mathbf{List}(A)) , \mathbf{Dec}(x \in \mathit{xs})
-  \end{equation}
-  Where \(\mathbf{Dec}(A)\) is a decision on some type \(A\).
-
-  We then need to convert a value of type \(\mathbf{Dec}(x \in \mathit{xs})\) to
-  \(x \in \mathit{xs}\).
-  We use the following to do that:
-  \begin{equation}
-    \text{recompute} : \mathbf{Dec}(A) \rightarrow \lVert A \rVert \rightarrow A
-  \end{equation}
-  The function works via case-analysis on \(\mathbf{Dec}\): in the true case, we
-  return the proof; in the false case, we can apply the proof of negation under
-  the truncation, and then extract \(\lVert \bot \rVert\) to \(\bot\), as
-  \(\bot\) is a mere proposition, thus proving the goal via explosion.
-\end{proof}
-\subsection{Closure}
-Since we don't have an equivalence with \(\mathbf{Fin}\), we don't get closure
-under \(\Pi\).
-\begin{rm-lemma}
-  Manifest enumerability is closed under \(\Sigma\).
-\end{rm-lemma}
-\begin{proof}
-  This closure proof is almost the same as theorem~\ref{split-enum-sigma}.
-  The manipulation of the support lists can be carried over as-is; but the type
-  of the cover proof has changed, so it will need to be updated.
-  As it happens, the translation is straightforward: we effectively write the
-  ``monadic'' version of the old function.
-\end{proof}
-\section{Cardinal Finiteness} \label{cardinal}
-For manifest enumerability, we removed the need for decidable equality: in these
-next two finiteness predicates, we remove the need for a total order on the
-underlying type.
-\begin{rm-definition}[Cardinal Finiteness]
-  A type \(A\) is cardinally finite, \(\mathcal{C}\), if it has a
-  propositionally-truncated proof of bishop finiteness.
-  \begin{equation}
-    \mathcal{C}(A) = \lVert \mathcal{B}(A) \rVert
-  \end{equation}%
-\end{rm-definition}%
-\subsection{Closure}
-The closure properties of cardinal finiteness are effectively the non-dependent
-versions of manifest Bishop finiteness.
-To see why, consider equation~\ref{split-enum-sigma-clos-sig}.
-We can ``lift'' the proof (as a binary function) under a propositional
-truncation, giving us equation~\ref{sigma-proof-lifted}, but that doesn't give us
-the desired closure proof (equation~\ref{sigma-proof-desired}).
-\begin{multicols}{2}
-  \noindent
-  \begin{equation} \label{sigma-proof-lifted}
-    \frac{
-      \lVert \mathcal{B}(A) \rVert \; \; \; \lVert \Pi (x : A) , \mathcal{B}(U(x)) \rVert
-    }{
-      \lVert \mathcal{B}(\Sigma(x : A) , U(x)) \rVert
-    }
-  \end{equation}
-  \begin{equation} \label{sigma-proof-desired}
-    \frac{
-      \lVert \mathcal{B}(A) \rVert \; \; \; \Pi (x : A) , \lVert \mathcal{B}(U(x))  \rVert
-    }{
-      \lVert \mathcal{B}(\Sigma(x : A) , U(x)) \rVert
-    }
-  \end{equation}
-\end{multicols}%
-To remedy the mismatch we would need a function of the type:
-\begin{equation}
-  \left( \Pi(x : A) , \lVert \mathcal{B}(U(x)) \rVert \right) \rightarrow
-  \lVert \Pi(x : A) , \mathcal{B}(U(x)) \rVert
-\end{equation}
-Unfortunately, this equation in particular is a form of the axiom of choice
-\cite[equation 3.8.3]{hottbook}.
-This leaves us with closure under only the non-dependent operations.
-\begin{rm-lemma}
-  Cardinal finiteness is closed under \(\times\), \(+\), and \(\rightarrow\).
-\end{rm-lemma}
-\begin{proof}
-  All of these closure proofs can be lifted directly from their corresponding
-  proofs on manifest Bishop finiteness.
-\end{proof}
-\subsection{Strength}
-The following two theorems are proven in
-\cite{yorgeyCombinatorialSpeciesLabelled2014}, in much the same way as we have
-done here.
-Our contribution for this section is simply the formalisation.
-\begin{rm-theorem} \label{cardinal-finite-discrete}
-  Any cardinal-finite set has decidable equality.
-\end{rm-theorem}
-\begin{proof}
-  See \cite[Proposition 2.4.10]{yorgeyCombinatorialSpeciesLabelled2014}.
-\end{proof}
-\begin{rm-theorem}
-  Given a cardinally finite type, we can derive the type's cardinality, as well
-  as a propositionally truncated proof of equivalence with \(\textbf{Fin}\)s of
-  the same cardinality.
-  \begin{equation}
-    \mathcal{C}(A) \rightarrow \Sigma {(n : \mathbb{N})} , \lVert \textbf{Fin}(n) \simeq A \rVert
-  \end{equation}
-\end{rm-theorem}
-\begin{proof}
-  \cite[Proposition 2.4.9]{yorgeyCombinatorialSpeciesLabelled2014}.
-\end{proof}
-\subsection{Relation to Manifest Bishop Finiteness}
-Cardinal finiteness tells us that there is an isomorphism between a type and
-\(\mathbf{Fin}\); it just doesn't tell us \emph{which} isomorphism.
-To take a simple example, \(\mathbf{Bool}\) has 2 possible isomorphisms with the
-set \(\mathbf{Fin}(2)\): one where false maps to 0, and true to 1; and another
-where false maps to 1 and true to 0.
-
-To convert from Cardinal finiteness to Bishop finiteness, then, requires that we
-supply enough information to identity a particular isomorphism.
-A total order is sufficient here: it will give us enough to uniquely order the
-support list invariant under permutations.
-This tells us what we already knew in the introduction: manifest Bishop
-finiteness is cardinal finiteness plus an order.
-
-\begin{rm-theorem}
-  Any cardinal finite type with a (decidable) total order is manifestly Bishop
-  finite.
-\end{rm-theorem}
-\begin{proof}
-  This proof is quite involved, and will rely on several subsequent lemmas, so
-  we will give only its outline here.
-  \begin{itemize}
-    \item First, we will convert to manifest enumerability: knowing that the
-      underlying type is discrete (theorem~\ref{cardinal-finite-discrete}) we can
-      go from manifest enumerability to split enumerability
-      (lemma~\ref{manifest-enum-to-split}), and subsequently to manifest Bishop
-      finiteness (lemma~\ref{split-enum-to-manifest-bishop}).
-    \item To convert to manifest enumerability, we need to provide a support
-      list: this cannot simply be the support list hidden under the truncation,
-      since that would violate the hiding promised by the truncation.
-      Instead, we sort the list (using insertion sort).
-      We must, therefore, prove that insertion sort is invariant under all
-      support lists in cardinal finiteness proofs.
-    \item We do this by first showing that all support lists in cardinal
-      finiteness proofs are permutations of each other, and then that insertion
-      sort is invariant under permutations.
-    \item Given our particular definition of permutations, cover proofs
-      transfers naturally between lists which are permutations of each other.
-  \end{itemize} 
-\end{proof}
-Now we will build up the toolkit we need to perform the above steps.
-First, permutations.
-\begin{rm-definition}[List Permutations]
-  We say that two lists are permutations of each other if there is an
-  isomorphism between membership proofs
-  \cite{danielssonBagEquivalenceProofRelevant2012}.
-  \begin{equation}
-    \mathit{xs} \leftrightsquigarrow \mathit{ys} = \Pi {(x : A)} , x \in \mathit{xs} \iff x \in \mathit{ys}
-  \end{equation}
-\end{rm-definition}
-\begin{rm-lemma}
-  Insertion sort is invariant under permutations.
-  \begin{equation}
-    \mathit{xs} \leftrightsquigarrow \mathit{ys} \implies \text{sort}(\mathit{xs}) \equiv \text{sort}(\mathit{ys})
-  \end{equation}
-\end{rm-lemma}
-\begin{proof}
-  First, we prove two properties about insertion sort:
-  \begin{enumerate}
-    \item It returns a sorted list.
-    \item It a returns a list that is a permutation of its input.
-  \end{enumerate}
-  The second of these points allows us to show that \(\text{sort}(\mathit{xs})\)
-  is a permutation of \(\text{sort}(\mathit{ys})\).
-  \begin{equation}
-    \text{sort}(\mathit{xs}) \leftrightsquigarrow \mathit{xs} \leftrightsquigarrow \mathit{ys} \leftrightsquigarrow \text{sort}(\mathit{ys})
-  \end{equation}
-  Then, we show that any lists which are both sorted and permutations of each
-  other are equal.
-  Both of these conditions are true for the output of \(\text{sort}\).
-\end{proof}
-\section{Kuratowski Finiteness} \label{kuratowski}
-Finally we arrive at Kuratowski finiteness
-\cite{kuratowskiNotionEnsembleFini1920}.
-\begin{rm-definition}[Kuratowski-Finite Set]
-  The Kuratowski finite set is a free join semilattice (or, equivalently, a free
-  commutative idempotent monoid).
-  HITs are required to define this type
-  \cite{altenkirchDefinableQuotientsType2011}:
-  \begin{equation}
-    \begin{aligned}
-      \mathcal{K}(A) =&
-       \; \cdot \dblcolon \cdot : A \times \mathcal{K}(A) \rightarrow \mathcal{K}(A) ; \\
-      \vert& \; [] : \mathcal{K}(A) ; \\
-      \vert& \; \text{com} : \Pi (x, y: A) , \Pi (\mathit{xs}: \mathcal{K}(A)) , x \dblcolon y \dblcolon \mathit{xs} \equiv y \dblcolon x \dblcolon \mathit{xs} ; \\
-      \vert& \; \text{dup} : \Pi (\mathit{x}: A) , \Pi (\mathit{xs}: \mathcal{K}(A)) , x \dblcolon x \dblcolon \mathit{xs} \equiv x \dblcolon \mathit{xs} ; \\
-      \vert& \; \text{trunc} : \Pi (\mathit{xs}, \mathit{ys}: \mathcal{K}(A)) , \Pi (p, q : \mathit{xs} \equiv \mathit{ys}) , p \equiv q ;
-    \end{aligned}
-  \end{equation}
-  The \(\text{com}\) and \(\text{dup}\) constructors effectively add
-  commutativity and idempotency to the free monoid (the list), which is made by
-  the first two constructors.
-  The last constructor makes \(\mathcal{K}(A)\) a set.
-
-  To eliminate from \(\mathcal{K}(A)\), we have to provide equations for each of
-  the point constructors which obey the equations of the path constructors.
-  For \(\text{com}\) and \(\text{dup}\), this means ensuring that the fold is
-  commutative and idempotent, whereas \(\text{trunc}\) means we can only
-  eliminate into sets.
-\end{rm-definition}
-
-Other representations of \(\mathcal{K}\) \cite{fruminFiniteSetsHomotopy2018}
-are more explicit constructions of the free join semilattice (i.e. there is a
-point constructors for union instead of cons, and then path constructors for
-the associativity and identity laws), but we have found this representation
-easier to work with.
-Nonetheless, the alternative representation is included in our formalisation,
-and proven equivalent to the representation here.
-\begin{rm-definition}[Membership of \mathcal{K}]
-  First, we need to provide equations for the two point constructors.
-  \(x \in [] = \bot\), and \(x \in y \dblcolon \mathit{xs} = \lVert (x \equiv y)
-  + (x \in \mathit{xs}) \rVert \).
-  The \(\text{com}\) and \(\text{dup}\) constructors are handled by proving that
-  the truncated form of \(+\) is itself commutative and idempotent.
-  The type of propositions is itself a set, satisfying the \(\text{trunc}\)
-  constructor.
-\end{rm-definition}
-\begin{rm-definition}[Kuratowski Finiteness]
-  A type is Kuratowski finite iff there exists a Kuratowski Set which contains
-  all of its elements.
-  \begin{equation}
-    \mathcal{K}^{f}(A) = \Sigma {(\mathit{xs} : \mathcal{K}(A))} , \Pi (x : A) , x \in \mathit{xs}
-  \end{equation}
-\end{rm-definition}
-
-\begin{rm-theorem}
-  A proof of Kuratowski finiteness is equivalent to a propositionally truncated
-  proof of enumerability.
-  \begin{equation}
-    \mathcal{K}^f(A) \simeq \lVert \mathcal{E}(A) \rVert
-  \end{equation}
-\end{rm-theorem}
-\begin{proof}
-  We prove by way of an isomorphism.
-  In the first direction (from \(\mathcal{K}\) to \(\mathcal{E}\)), because we
-  are eliminating into a proposition, we need only deal with the point
-  constructors.
-  For these, we convert the \(\mathcal{K}\) cons to its list counterpart, and
-  similarly for the nil constructor.
-
-  The other direction is proven in \cite{fruminFiniteSetsHomotopy2018}, so we
-  will not describe it here.
-\end{proof}
-\subsection{Topos}
-At this point, we see that a ``decidable Kuratowski finite set'' is precisely
-equivalent to a cardinal finite set.
-From this, we can lift over all of the properties of cardinal finite sets.
-In particular, we see that decidable Kuratowski finite sets form a \emph{topos}.
-We already have most of the components we need: closure under \(\bot\),
-\(\top\), \(\mathbf{Bool}\), \(+\), \(\times\), and \(\rightarrow\).
-What remains is the subobject classifier.
-\todo[inline]{Topos!}
-\section{Infinite Cardinalities} \label{infinite-cardinalities}
-In the previous sections we saw different flavours of finiteness which were
-really just different flavours of relations to \(\mathbf{Fin}\).
-In this section we will see that we can construct a similar classification of
-relations to \(\mathbb{N}\), in the form of the countably infinite types.
-\subsection{Split Countable Types}
-Our first foray into the world of countable types will be a straightforward
-analogue to the split enumerable types.
-We need change only one element: instead of a support \emph{list}, we instead
-have a support \emph{stream}, which is its infinite.
-\begin{rm-definition}[Stream]
-  \begin{equation}
-    \mathbf{Stream}(A) = \llbracket \top \triangleright \text{const} (\mathbb{N}) \rrbracket(A) \simeq \mathbb{N} \rightarrow A
-  \end{equation}
-\end{rm-definition}
-\begin{rm-definition}[Split Countability]
-  \begin{equation}
-    \mathcal{E}!(A) = \Sigma {(\mathit{xs} : \mathbf{Stream}(A))} , \Pi {(x : A)} , x \in \mathit{xs}
-  \end{equation}
-\end{rm-definition}
-This type is definitionally equal to it surjection equivalent (\(\mathbb{N}
-\twoheadrightarrow ! \; A\)).
-\subsubsection{Closure}
-We know that countable infinity is not closed under the exponential (function
-arrow), so the only closure we need to prove is \(\Sigma\) to cover all of
-what's left.
-\input{figures/pairing-functions.tex}
-\begin{rm-theorem} \label{split-countability-sigma}
-  Split countability is closed under \(\Sigma\).
-\end{rm-theorem}
-\begin{proof}
-  The main task here is to figure out a pattern which will pair elements of the
-  two support streams without diverging.
-  Figure~\ref{cartesian} illustrates why the pairing function we used for lists
-  won't work: it needs to exhaust one of the input lists before inspecting
-  anything other than the first element of the other.
-  Of course, we can't exhaust a stream, so this diverges.
-
-  Instead, we will use the pattern in figure~\ref{cantor}.
-  We will produce a stream of lists (where each list is a diagonal), which is
-  then concatenated to the final output stream.
-  The intermediate stream is sometimes called the discrete convolution of the
-  two input streams.
-\end{proof}
-\subsubsection{Kleene Star}
-While we lose some closures with the inclusion of infinite types, we gain some
-others.
-In particular, we have the Kleene star.
-This means that we have closure under lists.
-\begin{rm-theorem}
-  Split countability is closed under Kleene star.
-  \begin{equation}
-    \mathcal{E}!(A) \rightarrow \mathcal{E}!(\mathbf{List}(A))
-  \end{equation}
-\end{rm-theorem}
-\begin{proof}
-  As with the proof of closure under \(\Sigma\), our main task here is to figure
-  out a pattern that reaches any given finite list in finite time.
-  Our first step is to split up the problem as we did for \(\Sigma\)-closure: we
-  produce an intermediate stream of lists, which we then concatenate to give the
-  final result.
-  The intermediate stream is constructed like so:
-  \begin{equation}
-    \mathit{xs}^\star_i = \left\{ \text{map}(\mathit{xs} , x) \mid \mathit{x} \in \mathbf{List}(\mathbb{N}) , \text{sum}(\text{map}(1+, x)) \equiv i \right\}
-  \end{equation}
-  In other words, the \(i\)th element is all lists such that the sum of the
-  successor of the indices in the list is equal to \(i\).
-\end{proof}
-\subsection{Other Types of Countability}
-There are of course other forms of countability available to us: manifest
-countability, cardinal, etc.
-We have a similar structure of relationships to the finite types, with the
-exception that we cannot construct an isomorphism with \(\mathbb{N}\) fiven a
-surjection from \(\mathbb{N}\).
-\section{Practical Uses} \label{practical}
-\subsection{Omniscience}
-In this section we are interested in restricted forms of the limited principle
-of omniscience \cite{myhillErrettBishopFoundations1972}.
-\begin{rm-definition}[Limited Principle of Omniscience]
-  For any type \(A\) and predicate \(P\) on \(A\), the limited principle of
-  omniscience is as follows:
-  \begin{equation}
-    \left( \Pi {(x : A)} , \mathbf{Dec}(P(x)) \right) \rightarrow \mathbf{Dec} \left( \Sigma {(x : A)} , P(x) \right)
-  \end{equation}
-  In other words, for any decidable predicate the existential quantification of
-  that predicate is also decidable.
-\end{rm-definition}
-The limited principle of omniscience is non-constructive, but individual types
-can themselves satisfy omniscience.
-In particular, \emph{finite} types are omniscient.
-
-There is also a universal form of omniscience, which we call exhaustibility.
-\begin{rm-definition}[Exhaustibility]
-  We say a type \(A\) is exhaustible if, for any decidable predicate \(P\) on
-  \(A\), the universal quantification of the predicate is decidable.
-  \begin{equation}
-    \left( \Pi {(x : A)} , \mathbf{Dec}(P(x)) \right) \rightarrow \mathbf{Dec} \left( \Pi {(x : A)} , P(x) \right)
-  \end{equation}
-\end{rm-definition}
-
-All of the finiteness predicates we have seen justify exhaustibility.
-We will only prove it once, then, for the weakest:
-\begin{rm-theorem}
-  Kuratowski-finite types are exhaustible.
-\end{rm-theorem}
-
-Omniscience is stronger than exhaustibility, as we can derive the latter from
-the former:
-\begin{rm-lemma} \label{omniscient-is-exhaustible}
-  Any omniscient type is exhaustible.
-\end{rm-lemma}
-\begin{proof}
-  For decidable propositions, we know the following:
-  \begin{equation} \label{lem-conversion}
-    \Pi (x : A) , P(x) \leftrightarrow \neg \Sigma (x : A) , \neg P(x) 
-  \end{equation}
-  To derive exhaustibility from omniscience, then, we run the predicate in its
-  negated form, and then subsequently negate the result.
-  The resulting decision over \(\neg \Sigma(x : A) , \neg P(x)\) can be
-  converted into \(\Pi (x : A) , P(x)\).
-\end{proof}
-We cannot derive, however, that any exhaustible type is omniscient, as we do not
-have the inverse of equation~\ref{lem-conversion}:
-\begin{equation} \label{choice-conversion}
-  \Sigma (x : A) , P(x) \leftrightarrow \neg \Pi (x : A) , \neg P(x)
-\end{equation}
-Such an equation would allow us to pick a representative element from any type,
-which is therefore non-constructive.
-In a sense, equation~\ref{lem-conversion} requires a form of LEM on the
-proposition (i.e. requires it to be decidable), whereas
-equation~\ref{choice-conversion} requires a form of choice.
-Those finiteness predicates which are ordered do in fact give us this form of
-choice, so the conversion is valid.
-As such, all of the ordered finiteness predicates imply omniscience.
-Again, we will prove it only for the weakest.
-\begin{rm-theorem}
-  Manifest enumerable types are omniscient.
-\end{rm-theorem}
-Finally, we do have a form of omniscience for prop-valued predicates, as they do
-not care about the chosen representative.
-\begin{rm-theorem}
-  Kuratowski finite types are omniscient about prop-valued predicates.
-\end{rm-theorem}
-\subsection{Synthesising Pattern-Matching Proofs}
-In particular, they can automate large proofs by analysing every possible case.
-In \cite{firsovDependentlyTypedProgramming2015}, the \(\AgdaDatatype{Pauli}\)
-group is used as an example.
-\ExecuteMetaData[agda/Data/Pauli.tex]{def}
-As \(\AgdaDatatype{Pauli}\) has 4 constructors, \(n\)-ary functions on
-\(\AgdaDatatype{Pauli}\) may require up to \(4^n\) cases, making even simple
-proofs prohibitively verbose.
-
-The alternative is to derive the things we need from \(\mathcal{E}!\) somehow.
-As \(\AgdaDatatype{Pauli}\) is a simple finite type, the instance can be defined
-in a similar way to those in lemma~\ref{split-enum-bool-top-bot}.
-From here we can already derive decidable equality, a function which requires 16
-cases if implemented manually.
-
-For proof search, the procedure is a well-known one in Agda
-\cite{devrieseBrightSideType2011}: we ask for the result of a decision procedure
-as an \emph{instance argument}, which will demand computation during
-typechecking.
-\subsection{Multiple Arguments}
-The automation machinery above only deals with single-argument predicates.
-This is not a problem, as we know that we can work with multiple arguments by
-currying and uncurrying, since all of the finiteness predicates are closed under
-\(\times\).
-To automate away the curry/uncurry noise we will use instance search, building
-on \cite{allaisGenericLevelPolymorphic2019} to develop a small interface to
-generic \(n\)-ary functions and properties.
-Our generic representation can handle dependent \(\Sigma\) and \(\prod\)
-types (rather than their non-dependent counterparts, \(\times\) and
-\(\rightarrow\)).
-This extension was necessary for our use case: it is mentioned in the paper
-as the obvious next step. 
-We also implement the curry-uncurry combinators as (verified) isomorphisms.
-
-A full explanation of our implementation is beyond the scope of this work, so we
-only present the finished interface, which is used like so:
-\ExecuteMetaData[agda/Data/Pauli.tex]{assoc-prf}
-\subsection{Synthesising Functions}
-Finally, thanks to extensionality provided to us by HoTT, and the computation
-properties provided by CuTT, we can derive decidable equality on functions over
-finite types.
-We can also use functions in our proof search.
-Here, for instance, is a automated procedure which finds the
-\(\AgdaFunction{not}\) function on \(\mathbf{Bool}\), given a specification.
-\ExecuteMetaData[agda/Data/Pauli.tex]{not-spec}
-\section{Related Work}
-\cite{spiwackConstructivelyFinite2010} explored the notion of ``constructive
-finite sets'', demonstrating that there is a rich variety of finiteness
-predicates in a constructive setting.
-Some of those predicates where explored more in
-\cite{firsovVariationsNoetherianness2016a},
-\cite{parmannInvestigatingStreamlessSets2015}, and
-\cite{uustaluFinitenessRationalSequences2017a}.
-
-\cite{yorgeyCombinatorialSpeciesLabelled2014} explored some of the definitions
-we have given here (manifest bishop finiteness and cardinal finiteness) in HoTT,
-in order to construct species.
-Some of this paper is a formalisation of the early parts of that work.
-
-\cite{firsovDependentlyTypedProgramming2015} explores finite sets in Agda and
-uses them to automate proof machinery in much the same way as we have here.
-The primary difference between their proof automation and ours is that we can
-talk about finite functions, as we are working in HoTT.
-
-\cite{fruminFiniteSetsHomotopy2018} explores many of the same notions as we do
-here: cardinal finiteness, and Kuratowski finite sets.
-Our work makes extensive use of the computational properties of CuTT, however,
-and we are able to formalise the finiteness of function objects.
+
+\section{Relations Between Each Finiteness Definition}
+\section{Topos}
+\section{Countably Infinite Types}
+\section{Search}
+
+% % Taking finiteness as an example again, if we prove that some type is finite via
+% % bijection, in MLTT such a proof would have to reveal the particular bijection.
+% % This becomes problematic when defining, for instance, the category of finite
+% % sets: the objects of this category would ideally be types coupled with proofs of
+% % their finiteness, but in order to
+
+% % The five in particular that we are interested in are organised in
+% % figure~\ref{finite-classification}.
+% % We especially want to examine what separates each predicate: what extra
+% % information do you need, for instance, to go from a proof of Cardinal finiteness
+% % (section~\ref{cardinal}) to manifest Bishop finiteness
+% % (section~\ref{manifest-bishop}).
+
+% % Next, we will examine the category-theoretic and topos-theoretic interpretations
+% % of each of the predicates.
+% % We will give a ``category of finite sets'' for each finiteness predicate, and
+% % then prove the various closure properties required to arrive at the relevant
+% % topos (we show that decidable Kuratowski finite sets form a \(\Pi\)-pretopos,
+% % for instance).
+
+% % ways to prove finiteness end up proving the same
+% % thing: constructively, however, each defines a fundamentally different
+% % ``version'' of finiteness.
+% % In Martin-Löf type theory \cite{martin-lofIntuitionisticTypeTheory1980}, these
+% % variations (and more) have been explored in
+% % \cite{spiwackConstructivelyFinite2010} and
+% % \cite{firsovVariationsNoetherianness2016}.
+
+% % Homotopy Type Theory \cite{hottbook} further adds to this story in two ways.
+% % First, its universe of types includes non-sets: whether or not a finiteness
+% % predicate includes these gives us another dimension on which to discriminate.
+
+% % The second addition is concerned with the hiding of information.
+% % Occasionally in a constructive setting you don't want to have to reveal how you
+% % proved something.
+% % To model logical disjunction, for instance, we usually use a disjoint union: the
+% % model isn't entirely correct, though, because a proof of \(A \text{or} B\)
+% % reveals which of \(A\) or \(B\) we actually proved.
+
+% % logical disjunction, for instance, is usually modelled with a
+% % disjoint union.
+% % This always means, however, that to prove ``A or B'' you have to reveal which of
+% % A or B you did prove.
+
+% % The other headline feature of HoTT's generalised equality is univalence: put
+% % simply, it allows us to treat isomorphic types as equal, giving us in the
+% % constructive world access to a technique that is commonplace classically.
+% % It turns out that finite types have a lot to say about isomorphisms, and we will
+% % rely heavily on this aspect of HoTT to work with them.
+% % \subsection{Contributions}
+% % In this work, we will explore finite types in Cubical Type Theory
+% % \cite{cohenCubicalTypeTheory2016}, and expose their relationship to infinite
+% % types, species, and demonstrate their practical uses for proofs in dependently
+% % typed programming languages.
+
+% % In section~\ref{strong-finiteness}, we will explore the ``strong'' notions of
+% % finiteness (i.e. those at least as strong as Kuratowski finiteness
+% % \cite{kuratowskiNotionEnsembleFini1920}), with a special focus on cardinal
+% % finiteness.
+% % We will delineate some subtly different finiteness predicates, and prove the
+% % relationship between each.
+% % We will prove, then, that every discrete Kuratowski finite type is cardinal
+% % finite, and subsequently will show that discrete Kuratowski finite sets form a
+% % topos.
+
+% % In section~\ref{species-and-containers}, we will redefine our finiteness
+% % predicates using \emph{containers}, and prove that the new definitions are
+% % equivalent to the old.
+% % Using this, we will prove much stronger relationships between our finiteness
+% % predicates and relation-based definitions of finiteness.
+% % Finally, We will show how the species-container duality falls out naturally from
+% % these new definitions.
+
+% % In section~\ref{infinite-cardinalities}, we will extend our study of finite
+% % types to infinite but countable types.
+% % We will see that the finiteness predicates are mirrored with countable
+% % counterparts, and we will prove closure under the Kleene star and plus.
+
+% % Proofs of finiteness have well-known practical applications in
+% % constructive mathematics \cite{firsovDependentlyTypedProgramming2015}.
+% % In section~\ref{practical}, we build a library which exploits these uses in
+% % Cubical Agda \cite{vezzosiCubicalAgdaDependently2019}, allowing automation of
+% % complex proofs over finite types.
+% % We frame this in terms of the principle of omniscience for finite types.
+% % Thanks to the flexibility afforded to us by Cubical Type Theory, we are able to
+% % go further than the usual examples of this kind of proof automation: as well as
+% % proving properties about functions, we can synthesise functions whole-cloth from
+% % their desired properties.
+% % Through the unified interface for finite and countable types, we can
+% % reuse the automation machinery for \emph{partial} proof search over infinite
+% % search spaces.
+% % Along the way, we extend the work in \cite{allaisGenericLevelPolymorphic2019} to
+% % prove isomorphisms between the curried and uncurried forms of \(n\)-ary
+% % dependent functions.
+
+% % Classically, of course, each of these statements are equivalent.
+% % Constructively, though, all four predicates are fundamentally different.
+
+% % Part of the difference comes from ``leaking'' information about the finite type
+% % in the proof of its finiteness.
+% % If I demonstrate a type is finite by showing a surjection to that type, I've
+% % revealed a way to construct all of the elements of my type.
+% % If I show that it is bounded, I reveal an upper bound on the cardinality of the
+% % type.
+
+% \subsection{Strong Finiteness Predicates}
+% We will first explore the ``strong'' notions of finiteness (i.e. those at least
+% as strong as Kuratowski finiteness \cite{kuratowskiNotionEnsembleFini1920}),
+% with a special focus on cardinal finiteness (section~\ref{cardinal}), and
+% manifest enumerability (section~\ref{manifest-enumerability}), which is new, to
+% our knowledge.
+
+% Figure~\ref{finite-classification} organises the predicates according to
+% their strength.
+% We will go through each of the predicates, proving how to weaken each (i.e. we
+% will provide a proof that every cardinally finite type is Kuratowski finite),
+% and how to strengthen them, given the required property.
+% In terms of figure~\ref{finite-classification}, this amounts to providing proofs
+% for each arrow.
+% Our main proof in this regard is that we can derive manifest Bishop finiteness
+% from Kuratowski finiteness plus a total order.
+
+% We will---through the use of containers
+% \cite{abbottContainersConstructingStrictly2005}---formally prove the equivalence
+% these predicates have with the usual function relations i.e. we will show that a
+% proof of manifest enumerability is precisely equivalent to a surjection from a
+% finite prefix of the natural numbers.
+
+% For each predicate, we will also prove its closure over sums and products in
+% both dependent and non-dependent forms, if such a closure exists.
+% This will culminate in our main result for this section: the formal proof that
+% decidable Kuratowski finite sets form a topos.
+% \subsection{Infinite Types}
+% In section~\ref{infinite-cardinalities}, we will extend our study of finite
+% types to infinite but countable types.
+% We will see that the classification of finiteness predicates mirrors that of
+% their countable counterparts, and we will prove closure under the Kleene star.
+
+% \subsection{Practical Uses}
+% Finally, in section~\ref{practical} we will demonstrate some practical uses of
+% the finiteness proofs in Cubical Agda.
+% We will show how to use well-known techniques
+% \cite{firsovDependentlyTypedProgramming2015, devrieseBrightSideType2011} to
+% automate proofs of functions with finite (and infinite) domains.
+% We further show how to automate the synthesis of functions from desired
+% properties.
+% \section{Split Enumerability} \label{split-enumerable}
+% We will start with the simplest definition of finiteness: we say a set is
+% enumerable if there is a list of its elements which contains every element in
+% the set.
+
+% Before giving the definition of the predicate, we will first define lists (and
+% membership thereof): we have chosen a \emph{container}-based definition for this
+% work.
+% \begin{rm-definition}[Container] \label{container-def}
+%   A container \cite{abbottContainersConstructingStrictly2005} is a pair
+%   \(S \triangleright P\) where \(S\) is a type, the elements of which are called
+%   the \emph{shapes} of the container, and \(P\) is a type family on \(S\), where
+%   the elements of \(P(s)\) are called the \emph{positions} of a container.
+%   We ``interpret'' a container into a functor defined like so:
+%   \begin{equation}
+%     \llbracket S \triangleright P \rrbracket(A) = \Sigma {(s : S)} , \left( P(s) \rightarrow A \right)
+%   \end{equation}
+%   Membership of a container can be defined like so:
+%   \begin{equation}
+%     x \in \mathit{xs} = \text{fiber}(\text{snd}(\mathit{xs}), x)
+%   \end{equation}
+%   Where \(\text{fiber}\) is from \cite[definition 4.2.4]{hottbook}.
+% \end{rm-definition}
+% \begin{rm-definition}[Lists]
+%   \begin{equation}%
+%     \mathbf{List} = \llbracket \mathbb{N} \triangleright \mathbf{Fin} \rrbracket
+%   \end{equation}
+% \end{rm-definition}
+% \begin{rm-definition}[\mathbf{Fin}]
+%   \(\mathbf{Fin}(n)\) is the type of natural numbers smaller than \(n\).
+%   We define it the standard way, where \(\mathbf{Fin}(0) = \bot\) and
+%   \(\mathbf{Fin}(n + 1) &= \top + \mathbf{Fin}(n)\).
+% \end{rm-definition}
+
+% Internally, in our formalisation, we actually use the standard
+% inductive definition of lists more often (it tends to work better in more
+% complex algorithms, and functions on it seems to satisfy the termination
+% checker more readily).
+% However, since both types are equivalent, univalence allows us to transport to
+% whichever representation is more convenient in a given situation.
+% For the higher-level proofs we present here, though, the container-based
+% definition greatly simplifies certain steps, which is why we have chosen it as
+% our representation.
+
+% Finally, we can define formally split enumerability.
+% \subsection{Split Surjections}
+% \begin{rm-definition}[Surjections]
+%   \begin{multicols}{2}
+%     \begin{equation}
+%       A \twoheadrightarrow B = \Sigma (f : A \rightarrow B) , \text{surjective}(f)
+%     \end{equation} \break
+%     \begin{equation}
+%       A \twoheadrightarrow! \; B = \Sigma (f : A \rightarrow B) , \text{split surjective}(f)
+%     \end{equation}
+%   \end{multicols}
+%   Where the definitions of surjective are taken from \cite[definition
+%   4.6.1]{hottbook}.
+% \end{rm-definition}
+% \begin{rm-theorem} \label{split-enum-surj}
+%   Split enumerability is equivalent to a split surjection from a finite prefix
+%   of the natural numbers.
+%   \begin{equation}
+%     \mathcal{E}!(A) \iff \Sigma {(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \twoheadrightarrow ! \; A \right)
+%   \end{equation}
+% \end{rm-theorem}
+% \begin{proof}
+%   After sufficient inlining, it emerges that our goal is simply a reassociation.
+%   \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{split-surj}
+%   To be clear: in Agda, the proof could simply be \(\AgdaFunction{reassoc}\); we
+%   have written out the extra lines for clarity alone.
+% \end{proof}
+% \begin{rm-lemma} \label{split-enum-discrete}
+%   Any split enumerable type has decidable equality (is discrete).
+% \end{rm-lemma}
+% \begin{proof}
+%   We use a corollary that if there is a split-surjection from \(A\) to \(B\),
+%   and \(A\) is discrete, then \(B\) is also discrete.
+% \end{proof}
+% By Hedberg's theorem \cite{hedbergCoherenceTheoremMartinLof1998}, we know that
+% this implies split enumerable types must all be sets.
+% \subsection{Closure}
+% In this section we will prove closure under various operations for split
+% enumerable sets.
+% We are working towards a topos proof, which requires us to prove closure under
+% a variety of operations: for now, we only have enough machinery to demonstrate
+% the semiring operations, and dependent sums.
+% In order to show closure under exponentials (function arrows), we will need an
+% equivalence with \(\mathbf{Fin}\), which will be provided in
+% section~\ref{manifest-bishop}.
+% \begin{rm-lemma} \label{split-enum-bool-top-bot}
+%   \(\mathbf{Bool}\), \(\top\), and \(\bot\) are all split enumerable.
+% \end{rm-lemma}
+% \begin{proof}
+%   Each of these types clearly has a finite number of elements (2, 1, and 0,
+%   respectively), and furthermore has a straightforward enumeration.
+% \end{proof}
+% \begin{rm-theorem} \label{split-enum-sigma}
+%   Split-enumerability is closed under \(\Sigma\).
+%   \begin{equation} \label{split-enum-sigma-clos-sig}
+%     \frac{
+%       \mathcal{E}!(A) \; \; \; \Pi (x : A) , \mathcal{E}!(U(x))
+%     }{
+%       \mathcal{E}!(\Sigma (x : A) , U(x))
+%     }
+%   \end{equation}
+% \end{rm-theorem}
+% \begin{proof}
+%   To obtain the support list, we concatenate the support lists of all the proofs
+%   of split-finiteness for \(U\) over the support list of \(E_A\).
+%   In Agda:
+%   \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{sigma-sup}
+%   ``do-notation'' is available to us as we're working in the list monad.
+% \end{proof}
+% Closure under disjoint union and Cartesian product both follow from \(\Sigma\),
+% as these types can be defined in terms of \(\Sigma\).
+% \begin{align}
+%   A \times B &= \Sigma {(x : A)} , B \\
+%   A + B &= \Sigma {(x : \mathbf{Bool})} , \text{if} \; x \; \text{then} \; A \; \text{else} \; B
+% \end{align}
+% \section{Manifest Bishop Finiteness} \label{manifest-bishop}
+% Where split enumerability was the enumeration form of a split surjection from
+% \(\mathbf{Fin}\), manifest Bishop finiteness is the enumeration form of an
+% \emph{equivalence} with \(\mathbf{Fin}\).
+
+% \begin{rm-definition}[Manifest Bishop Finiteness]
+%   \begin{equation}
+%     \mathcal{B}(A) = \Sigma {(\mathit{xs} : \textbf{List}(A))} , \Pi {(x : A)} , x \inunique xs
+%   \end{equation}
+% \end{rm-definition}
+% The only difference between this predicate and split enumerability is the list
+% membership term: we use \(\inunique\) here, where \(x \inunique \mathit{xs}\) is
+% to be read as ``\(x\) occurs exactly once in \(\mathit{xs}\)''.
+% \begin{rm-definition}[Unique Membership]
+%   We say an item \(x\) is ``uniquely in'' some container \(\mathit{xs}\) if its
+%   membership in that list is a \emph{contraction} \cite[definition
+%   3.11.1]{hottbook}; i.e. its membership proof exists, and all such proofs are
+%   equal.
+%   \begin{equation}
+%     x \inunique \mathit{xs} = \text{isContr}(x \in \mathit{xs})
+%   \end{equation}
+% \end{rm-definition}
+% By the definition of a contraction, we can always recover the underlying
+% membership proof, meaning that we can always derive split enumerability from
+% manifest Bishop finiteness.
+% \subsection{Equivalence}
+% \begin{rm-lemma} \label{bishop-equiv}
+%   A proof of manifest Bishop finiteness is equivalent to an equivalence with a
+%   finite prefix of the natural numbers.
+%   \begin{equation}
+%     \mathcal{B}(A) \simeq \Sigma {(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \simeq A \right)
+%   \end{equation}
+% \end{rm-lemma}
+% \begin{proof}
+%   There are many equivalent definitions of equivalence in HoTT.
+%   Here we take the version preferred in the Cubical Agda library: contractible
+%   maps \cite[definition 4.4.1]{hottbook}.
+%   Because of the parallels between contractible maps and split surjections,
+%   the proof proceeds much the same as \ref{split-enum-surj}.
+%   In other words, the definition of Bishop finiteness is itself a reassociation
+%   of a contractible map.
+% \end{proof}
+% \subsection{Relationship to Split Enumerability}
+% \begin{rm-theorem} \label{split-enum-to-manifest-bishop}
+%   Any split enumerable set is manifest Bishop finite.
+% \end{rm-theorem}
+% \begin{proof}
+%   From proposition~\ref{split-enum-discrete} we can derive decidable equality on
+%   \(A\), and using this we can define a function (called \(\text{uniques}\)) which filters
+%   out duplicates from lists of \(A\)s.
+%   This gives us our support list.
+%   To generate the cover proof it suffices now to prove the following:
+%   \begin{equation}
+%     \Pi {(x : A)} , \Pi {(\mathit{xs} : \textbf{List}(A))} , x \in \mathit{xs} \rightarrow x \inunique \text{uniques}(\mathit{xs})
+%   \end{equation}
+% \end{proof}
+% \subsection{Closure}
+% Proving equal strength of split enumerability and manifest Bishop finiteness
+% allows us to carry all of the previous proofs of closure over to manifest Bishop
+% finite sets (and vice-versa).
+% Missing from our previous proofs was a proof of closure of functions.
+% We remedy that here.
+% \begin{rm-theorem}
+%   Manifest bishop finiteness is closed over dependent functions
+%   (\(\prod\)-types).
+%   \begin{equation}
+%     \frac{
+%       \mathcal{B}(A) \; \; \; \Pi {(x : A)} , \mathcal{B}\left( U(x) \right)
+%     }{
+%       \mathcal{B}\left(\Pi {(x : A)} , U(x)\right)
+%     }
+%   \end{equation}
+% \end{rm-theorem}
+% \begin{proof}
+%   This proof is essentially the composition of two transport operations, made
+%   available to us via univalence.
+
+%   First, we will simplify things slightly by working only with split
+%   enumerability.
+%   As this is equal in strength to manifest Bishop finiteness, any closure proofs
+%   carry over.
+
+%   Secondly, we will replace \(A\) in all places with \(\mathbf{Fin}(n)\).
+%   Since we have already seen an equivalence between these two types, we are
+%   permitted to transport along these lines.
+%   This is the first transport operation.
+
+%   The bulk of the proof now is concerned with proving the following:
+%   \begin{equation}
+%     \left( \Pi {(x : \mathbf{Fin}(n))} , \mathcal{E}!(A(x)) \right) \rightarrow \mathcal{E}!\left( \Pi {(x : \mathbf{Fin}(n))} , A(x) \right)
+%   \end{equation}
+%   Our strategy to accomplish this will be to consider functions from
+%   \(\mathbf{Fin}(n)\) as \(n\)-tuples over some type family \(T : \mathbb{N}
+%   \rightarrow \text{Type}\).
+%   \begin{equation}
+%     \begin{aligned}
+%       \mathbf{Tuple}(T, 0)   &= \top \\
+%       \mathbf{Tuple}(T, n+1) &= T(0) \times \mathbf{Tuple}(T \circ \text{suc}, n)
+%     \end{aligned}
+%   \end{equation}
+%   This type is manifestly Bishop finite, as it is constructed only from products
+%   and the unit type.
+
+%   We then prove an isomorphism between this representation and \(\Pi\)-types.
+%   \begin{equation}
+%     \mathbf{Tuple}(T, n) \iff \Pi {(x : \mathbf{Fin}(n))} , T(x)
+%   \end{equation}
+%   This allows us to transport our proof of finiteness on tuples to one on
+%   functions from \(\mathbf{Fin}\) (our second transport operation), proving our
+%   goal.
+% \end{proof}
+% \section{Manifest Enumerability} \label{manifest-enumerability}
+% Both split enumerability and manifest Bishop finiteness are restricted to sets:
+% because they both imply decidable equality, no non-set types can satisfy them as
+% predicates.
+% To find a more general predicate for finiteness, we \emph{truncate} the
+% membership proof.
+% \begin{rm-definition}[Manifest Enumerability]
+%   \begin{equation}
+%     \mathcal{E}(A) = \Sigma {(\mathit{xs} : \mathbf{List}(A))} , \Pi {(x : A)} , \lVert x \in \mathit{xs} \rVert
+%   \end{equation}
+% \end{rm-definition}
+% \begin{rm-definition}[Propositional Truncation]a \label{prop-trunc}
+%   The type \(\lVert A \rVert\) on some type \(A\) is a propositionally truncated
+%   proof of \(A\) \cite[3.7]{hottbook}.
+%   In other words, it is a proof that some \(A\) exists, but it does not tell you
+%   \emph{which} \(A\).
+
+%   It is defined as a Higher Inductive Type:
+%   \begin{equation}
+%     \begin{aligned}
+%       \lVert A \rVert =& \; \lvert \cdot \rvert : A \rightarrow \lVert A \rVert ; \\
+%                       |& \; \text{squash} : \Pi {(x, y : \lVert A \rVert)} , x \equiv y  ; 
+%     \end{aligned}
+%   \end{equation}
+
+%   We will use three eliminators from \(\lVert A \rVert\) in this paper.
+%   \begin{enumerate}
+%   \item \label{elim-prop-prop} For any function \(A \rightarrow B\), where
+%     \(\text{isProp}(B)\), we have a function \(\lVert A \rVert \rightarrow B\).
+%   \item \label{elim-prop-monad} A special case of \ref{elim-prop-prop} implies
+%     that \(\lVert \cdot \rVert\) forms a monad: this means that we get the usual
+%     Monadic operators on \(\lVert \cdot \rVert\) (bind, pure, fmap, etc.).
+%     \item \label{elim-prop-coh} We can eliminate from \(\lVert A \rVert\) with a
+%       function \(f : A \rightarrow B\) iff \(f\) ``doesn't care'' about the
+%       choice of \(A\) (\(\Pi {(x , y : A)} , f(x) \equiv f(y) \)).
+%       Formally speaking, \(f\) needs to be ``coherently constant''
+%       \cite{krausGeneralUniversalProperty2015}, and \(B\) needs to be an
+%       \(n\)-type for some finite \(n\).
+%   \end{enumerate}
+% \end{rm-definition}
+
+% We will use the circle as an example non-set type which is nonetheless
+% manifestly enumerable.
+% \begin{rm-theorem}
+%   The circle \(S^1\) is manifestly enumerable.
+% \end{rm-theorem}
+% \begin{proof}
+%   As the cover proof is a truncated proposition, we need only consider the point
+%   constructors, making this poof the same as the proof of split enumerability on
+%   \(\top\).
+% \end{proof}
+% \subsection{Surjections}
+% This predicates relation to surjectivity is much the same as split
+% enumerability's relation to \emph{split} surjectivity.
+% \begin{rm-lemma}
+%   A proof of manifest enumerability is equivalent to a surjection from a finite
+%   prefix of the natural numbers.
+% \end{rm-lemma}
+% \subsection{Relation to Split Enumerability}
+% \begin{rm-theorem} \label{manifest-enum-to-split}
+%   A manifestly enumerable type with decidable equality is split enumerable.
+% \end{rm-theorem}
+% \begin{proof}
+%   The support list stays the same between both enumerability proofs.
+
+%   For the cover proof, we first need the following function which searches a
+%   list for a particular element (given decidable equality on \(A\)).
+%   \begin{equation}
+%     \decin \; : \Pi(x : A) , \Pi(\mathit{xs} : \mathbf{List}(A)) , \mathbf{Dec}(x \in \mathit{xs})
+%   \end{equation}
+%   Where \(\mathbf{Dec}(A)\) is a decision on some type \(A\).
+
+%   We then need to convert a value of type \(\mathbf{Dec}(x \in \mathit{xs})\) to
+%   \(x \in \mathit{xs}\).
+%   We use the following to do that:
+%   \begin{equation}
+%     \text{recompute} : \mathbf{Dec}(A) \rightarrow \lVert A \rVert \rightarrow A
+%   \end{equation}
+%   The function works via case-analysis on \(\mathbf{Dec}\): in the true case, we
+%   return the proof; in the false case, we can apply the proof of negation under
+%   the truncation, and then extract \(\lVert \bot \rVert\) to \(\bot\), as
+%   \(\bot\) is a mere proposition, thus proving the goal via explosion.
+% \end{proof}
+% \subsection{Closure}
+% Since we don't have an equivalence with \(\mathbf{Fin}\), we don't get closure
+% under \(\Pi\).
+% \begin{rm-lemma}
+%   Manifest enumerability is closed under \(\Sigma\).
+% \end{rm-lemma}
+% \begin{proof}
+%   This closure proof is almost the same as theorem~\ref{split-enum-sigma}.
+%   The manipulation of the support lists can be carried over as-is; but the type
+%   of the cover proof has changed, so it will need to be updated.
+%   As it happens, the translation is straightforward: we effectively write the
+%   ``monadic'' version of the old function.
+% \end{proof}
+% \section{Cardinal Finiteness} \label{cardinal}
+% For manifest enumerability, we removed the need for decidable equality: in these
+% next two finiteness predicates, we remove the need for a total order on the
+% underlying type.
+% \begin{rm-definition}[Cardinal Finiteness]
+%   A type \(A\) is cardinally finite, \(\mathcal{C}\), if it has a
+%   propositionally-truncated proof of bishop finiteness.
+%   \begin{equation}
+%     \mathcal{C}(A) = \lVert \mathcal{B}(A) \rVert
+%   \end{equation}%
+% \end{rm-definition}%
+% \subsection{Closure}
+% The closure properties of cardinal finiteness are effectively the non-dependent
+% versions of manifest Bishop finiteness.
+% To see why, consider equation~\ref{split-enum-sigma-clos-sig}.
+% We can ``lift'' the proof (as a binary function) under a propositional
+% truncation, giving us equation~\ref{sigma-proof-lifted}, but that doesn't give us
+% the desired closure proof (equation~\ref{sigma-proof-desired}).
+% \begin{multicols}{2}
+%   \noindent
+%   \begin{equation} \label{sigma-proof-lifted}
+%     \frac{
+%       \lVert \mathcal{B}(A) \rVert \; \; \; \lVert \Pi (x : A) , \mathcal{B}(U(x)) \rVert
+%     }{
+%       \lVert \mathcal{B}(\Sigma(x : A) , U(x)) \rVert
+%     }
+%   \end{equation}
+%   \begin{equation} \label{sigma-proof-desired}
+%     \frac{
+%       \lVert \mathcal{B}(A) \rVert \; \; \; \Pi (x : A) , \lVert \mathcal{B}(U(x))  \rVert
+%     }{
+%       \lVert \mathcal{B}(\Sigma(x : A) , U(x)) \rVert
+%     }
+%   \end{equation}
+% \end{multicols}%
+% To remedy the mismatch we would need a function of the type:
+% \begin{equation}
+%   \left( \Pi(x : A) , \lVert \mathcal{B}(U(x)) \rVert \right) \rightarrow
+%   \lVert \Pi(x : A) , \mathcal{B}(U(x)) \rVert
+% \end{equation}
+% Unfortunately, this equation in particular is a form of the axiom of choice
+% \cite[equation 3.8.3]{hottbook}.
+% This leaves us with closure under only the non-dependent operations.
+% \begin{rm-lemma}
+%   Cardinal finiteness is closed under \(\times\), \(+\), and \(\rightarrow\).
+% \end{rm-lemma}
+% \begin{proof}
+%   All of these closure proofs can be lifted directly from their corresponding
+%   proofs on manifest Bishop finiteness.
+% \end{proof}
+% \subsection{Strength}
+% The following two theorems are proven in
+% \cite{yorgeyCombinatorialSpeciesLabelled2014}, in much the same way as we have
+% done here.
+% Our contribution for this section is simply the formalisation.
+% \begin{rm-theorem} \label{cardinal-finite-discrete}
+%   Any cardinal-finite set has decidable equality.
+% \end{rm-theorem}
+% \begin{proof}
+%   See \cite[Proposition 2.4.10]{yorgeyCombinatorialSpeciesLabelled2014}.
+% \end{proof}
+% \begin{rm-theorem}
+%   Given a cardinally finite type, we can derive the type's cardinality, as well
+%   as a propositionally truncated proof of equivalence with \(\textbf{Fin}\)s of
+%   the same cardinality.
+%   \begin{equation}
+%     \mathcal{C}(A) \rightarrow \Sigma {(n : \mathbb{N})} , \lVert \textbf{Fin}(n) \simeq A \rVert
+%   \end{equation}
+% \end{rm-theorem}
+% \begin{proof}
+%   \cite[Proposition 2.4.9]{yorgeyCombinatorialSpeciesLabelled2014}.
+% \end{proof}
+% \subsection{Relation to Manifest Bishop Finiteness}
+% Cardinal finiteness tells us that there is an isomorphism between a type and
+% \(\mathbf{Fin}\); it just doesn't tell us \emph{which} isomorphism.
+% To take a simple example, \(\mathbf{Bool}\) has 2 possible isomorphisms with the
+% set \(\mathbf{Fin}(2)\): one where false maps to 0, and true to 1; and another
+% where false maps to 1 and true to 0.
+
+% To convert from Cardinal finiteness to Bishop finiteness, then, requires that we
+% supply enough information to identity a particular isomorphism.
+% A total order is sufficient here: it will give us enough to uniquely order the
+% support list invariant under permutations.
+% This tells us what we already knew in the introduction: manifest Bishop
+% finiteness is cardinal finiteness plus an order.
+
+% \begin{rm-theorem}
+%   Any cardinal finite type with a (decidable) total order is manifestly Bishop
+%   finite.
+% \end{rm-theorem}
+% \begin{proof}
+%   This proof is quite involved, and will rely on several subsequent lemmas, so
+%   we will give only its outline here.
+%   \begin{itemize}
+%     \item First, we will convert to manifest enumerability: knowing that the
+%       underlying type is discrete (theorem~\ref{cardinal-finite-discrete}) we can
+%       go from manifest enumerability to split enumerability
+%       (lemma~\ref{manifest-enum-to-split}), and subsequently to manifest Bishop
+%       finiteness (lemma~\ref{split-enum-to-manifest-bishop}).
+%     \item To convert to manifest enumerability, we need to provide a support
+%       list: this cannot simply be the support list hidden under the truncation,
+%       since that would violate the hiding promised by the truncation.
+%       Instead, we sort the list (using insertion sort).
+%       We must, therefore, prove that insertion sort is invariant under all
+%       support lists in cardinal finiteness proofs.
+%     \item We do this by first showing that all support lists in cardinal
+%       finiteness proofs are permutations of each other, and then that insertion
+%       sort is invariant under permutations.
+%     \item Given our particular definition of permutations, cover proofs
+%       transfers naturally between lists which are permutations of each other.
+%   \end{itemize} 
+% \end{proof}
+% Now we will build up the toolkit we need to perform the above steps.
+% First, permutations.
+% \begin{rm-definition}[List Permutations]
+%   We say that two lists are permutations of each other if there is an
+%   isomorphism between membership proofs
+%   \cite{danielssonBagEquivalenceProofRelevant2012}.
+%   \begin{equation}
+%     \mathit{xs} \leftrightsquigarrow \mathit{ys} = \Pi {(x : A)} , x \in \mathit{xs} \iff x \in \mathit{ys}
+%   \end{equation}
+% \end{rm-definition}
+% \begin{rm-lemma}
+%   Insertion sort is invariant under permutations.
+%   \begin{equation}
+%     \mathit{xs} \leftrightsquigarrow \mathit{ys} \implies \text{sort}(\mathit{xs}) \equiv \text{sort}(\mathit{ys})
+%   \end{equation}
+% \end{rm-lemma}
+% \begin{proof}
+%   First, we prove two properties about insertion sort:
+%   \begin{enumerate}
+%     \item It returns a sorted list.
+%     \item It a returns a list that is a permutation of its input.
+%   \end{enumerate}
+%   The second of these points allows us to show that \(\text{sort}(\mathit{xs})\)
+%   is a permutation of \(\text{sort}(\mathit{ys})\).
+%   \begin{equation}
+%     \text{sort}(\mathit{xs}) \leftrightsquigarrow \mathit{xs} \leftrightsquigarrow \mathit{ys} \leftrightsquigarrow \text{sort}(\mathit{ys})
+%   \end{equation}
+%   Then, we show that any lists which are both sorted and permutations of each
+%   other are equal.
+%   Both of these conditions are true for the output of \(\text{sort}\).
+% \end{proof}
+% \section{Kuratowski Finiteness} \label{kuratowski}
+% Finally we arrive at Kuratowski finiteness
+% \cite{kuratowskiNotionEnsembleFini1920}.
+% \begin{rm-definition}[Kuratowski-Finite Set]
+%   The Kuratowski finite set is a free join semilattice (or, equivalently, a free
+%   commutative idempotent monoid).
+%   HITs are required to define this type
+%   \cite{altenkirchDefinableQuotientsType2011}:
+%   \begin{equation}
+%     \begin{aligned}
+%       \mathcal{K}(A) =&
+%        \; \cdot \dblcolon \cdot : A \times \mathcal{K}(A) \rightarrow \mathcal{K}(A) ; \\
+%       \vert& \; [] : \mathcal{K}(A) ; \\
+%       \vert& \; \text{com} : \Pi (x, y: A) , \Pi (\mathit{xs}: \mathcal{K}(A)) , x \dblcolon y \dblcolon \mathit{xs} \equiv y \dblcolon x \dblcolon \mathit{xs} ; \\
+%       \vert& \; \text{dup} : \Pi (\mathit{x}: A) , \Pi (\mathit{xs}: \mathcal{K}(A)) , x \dblcolon x \dblcolon \mathit{xs} \equiv x \dblcolon \mathit{xs} ; \\
+%       \vert& \; \text{trunc} : \Pi (\mathit{xs}, \mathit{ys}: \mathcal{K}(A)) , \Pi (p, q : \mathit{xs} \equiv \mathit{ys}) , p \equiv q ;
+%     \end{aligned}
+%   \end{equation}
+%   The \(\text{com}\) and \(\text{dup}\) constructors effectively add
+%   commutativity and idempotency to the free monoid (the list), which is made by
+%   the first two constructors.
+%   The last constructor makes \(\mathcal{K}(A)\) a set.
+
+%   To eliminate from \(\mathcal{K}(A)\), we have to provide equations for each of
+%   the point constructors which obey the equations of the path constructors.
+%   For \(\text{com}\) and \(\text{dup}\), this means ensuring that the fold is
+%   commutative and idempotent, whereas \(\text{trunc}\) means we can only
+%   eliminate into sets.
+% \end{rm-definition}
+
+% Other representations of \(\mathcal{K}\) \cite{fruminFiniteSetsHomotopy2018}
+% are more explicit constructions of the free join semilattice (i.e. there is a
+% point constructors for union instead of cons, and then path constructors for
+% the associativity and identity laws), but we have found this representation
+% easier to work with.
+% Nonetheless, the alternative representation is included in our formalisation,
+% and proven equivalent to the representation here.
+% \begin{rm-definition}[Membership of \mathcal{K}]
+%   First, we need to provide equations for the two point constructors.
+%   \(x \in [] = \bot\), and \(x \in y \dblcolon \mathit{xs} = \lVert (x \equiv y)
+%   + (x \in \mathit{xs}) \rVert \).
+%   The \(\text{com}\) and \(\text{dup}\) constructors are handled by proving that
+%   the truncated form of \(+\) is itself commutative and idempotent.
+%   The type of propositions is itself a set, satisfying the \(\text{trunc}\)
+%   constructor.
+% \end{rm-definition}
+% \begin{rm-definition}[Kuratowski Finiteness]
+%   A type is Kuratowski finite iff there exists a Kuratowski Set which contains
+%   all of its elements.
+%   \begin{equation}
+%     \mathcal{K}^{f}(A) = \Sigma {(\mathit{xs} : \mathcal{K}(A))} , \Pi (x : A) , x \in \mathit{xs}
+%   \end{equation}
+% \end{rm-definition}
+
+% \begin{rm-theorem}
+%   A proof of Kuratowski finiteness is equivalent to a propositionally truncated
+%   proof of enumerability.
+%   \begin{equation}
+%     \mathcal{K}^f(A) \simeq \lVert \mathcal{E}(A) \rVert
+%   \end{equation}
+% \end{rm-theorem}
+% \begin{proof}
+%   We prove by way of an isomorphism.
+%   In the first direction (from \(\mathcal{K}\) to \(\mathcal{E}\)), because we
+%   are eliminating into a proposition, we need only deal with the point
+%   constructors.
+%   For these, we convert the \(\mathcal{K}\) cons to its list counterpart, and
+%   similarly for the nil constructor.
+
+%   The other direction is proven in \cite{fruminFiniteSetsHomotopy2018}, so we
+%   will not describe it here.
+% \end{proof}
+% \subsection{Topos}
+% At this point, we see that a ``decidable Kuratowski finite set'' is precisely
+% equivalent to a cardinal finite set.
+% From this, we can lift over all of the properties of cardinal finite sets.
+% In particular, we see that decidable Kuratowski finite sets form a \emph{topos}.
+% We already have most of the components we need: closure under \(\bot\),
+% \(\top\), \(\mathbf{Bool}\), \(+\), \(\times\), and \(\rightarrow\).
+% What remains is the subobject classifier.
+% \todo[inline]{Topos!}
+% \section{Infinite Cardinalities} \label{infinite-cardinalities}
+% In the previous sections we saw different flavours of finiteness which were
+% really just different flavours of relations to \(\mathbf{Fin}\).
+% In this section we will see that we can construct a similar classification of
+% relations to \(\mathbb{N}\), in the form of the countably infinite types.
+% \subsection{Split Countable Types}
+% Our first foray into the world of countable types will be a straightforward
+% analogue to the split enumerable types.
+% We need change only one element: instead of a support \emph{list}, we instead
+% have a support \emph{stream}, which is its infinite.
+% \begin{rm-definition}[Stream]
+%   \begin{equation}
+%     \mathbf{Stream}(A) = \llbracket \top \triangleright \text{const} (\mathbb{N}) \rrbracket(A) \simeq \mathbb{N} \rightarrow A
+%   \end{equation}
+% \end{rm-definition}
+% \begin{rm-definition}[Split Countability]
+%   \begin{equation}
+%     \mathcal{E}!(A) = \Sigma {(\mathit{xs} : \mathbf{Stream}(A))} , \Pi {(x : A)} , x \in \mathit{xs}
+%   \end{equation}
+% \end{rm-definition}
+% This type is definitionally equal to it surjection equivalent (\(\mathbb{N}
+% \twoheadrightarrow ! \; A\)).
+% \subsubsection{Closure}
+% We know that countable infinity is not closed under the exponential (function
+% arrow), so the only closure we need to prove is \(\Sigma\) to cover all of
+% what's left.
+% \input{figures/pairing-functions.tex}
+% \begin{rm-theorem} \label{split-countability-sigma}
+%   Split countability is closed under \(\Sigma\).
+% \end{rm-theorem}
+% \begin{proof}
+%   The main task here is to figure out a pattern which will pair elements of the
+%   two support streams without diverging.
+%   Figure~\ref{cartesian} illustrates why the pairing function we used for lists
+%   won't work: it needs to exhaust one of the input lists before inspecting
+%   anything other than the first element of the other.
+%   Of course, we can't exhaust a stream, so this diverges.
+
+%   Instead, we will use the pattern in figure~\ref{cantor}.
+%   We will produce a stream of lists (where each list is a diagonal), which is
+%   then concatenated to the final output stream.
+%   The intermediate stream is sometimes called the discrete convolution of the
+%   two input streams.
+% \end{proof}
+% \subsubsection{Kleene Star}
+% While we lose some closures with the inclusion of infinite types, we gain some
+% others.
+% In particular, we have the Kleene star.
+% This means that we have closure under lists.
+% \begin{rm-theorem}
+%   Split countability is closed under Kleene star.
+%   \begin{equation}
+%     \mathcal{E}!(A) \rightarrow \mathcal{E}!(\mathbf{List}(A))
+%   \end{equation}
+% \end{rm-theorem}
+% \begin{proof}
+%   As with the proof of closure under \(\Sigma\), our main task here is to figure
+%   out a pattern that reaches any given finite list in finite time.
+%   Our first step is to split up the problem as we did for \(\Sigma\)-closure: we
+%   produce an intermediate stream of lists, which we then concatenate to give the
+%   final result.
+%   The intermediate stream is constructed like so:
+%   \begin{equation}
+%     \mathit{xs}^\star_i = \left\{ \text{map}(\mathit{xs} , x) \mid \mathit{x} \in \mathbf{List}(\mathbb{N}) , \text{sum}(\text{map}(1+, x)) \equiv i \right\}
+%   \end{equation}
+%   In other words, the \(i\)th element is all lists such that the sum of the
+%   successor of the indices in the list is equal to \(i\).
+% \end{proof}
+% \subsection{Other Types of Countability}
+% There are of course other forms of countability available to us: manifest
+% countability, cardinal, etc.
+% We have a similar structure of relationships to the finite types, with the
+% exception that we cannot construct an isomorphism with \(\mathbb{N}\) fiven a
+% surjection from \(\mathbb{N}\).
+% \section{Practical Uses} \label{practical}
+% \subsection{Omniscience}
+% In this section we are interested in restricted forms of the limited principle
+% of omniscience \cite{myhillErrettBishopFoundations1972}.
+% \begin{rm-definition}[Limited Principle of Omniscience]
+%   For any type \(A\) and predicate \(P\) on \(A\), the limited principle of
+%   omniscience is as follows:
+%   \begin{equation}
+%     \left( \Pi {(x : A)} , \mathbf{Dec}(P(x)) \right) \rightarrow \mathbf{Dec} \left( \Sigma {(x : A)} , P(x) \right)
+%   \end{equation}
+%   In other words, for any decidable predicate the existential quantification of
+%   that predicate is also decidable.
+% \end{rm-definition}
+% The limited principle of omniscience is non-constructive, but individual types
+% can themselves satisfy omniscience.
+% In particular, \emph{finite} types are omniscient.
+
+% There is also a universal form of omniscience, which we call exhaustibility.
+% \begin{rm-definition}[Exhaustibility]
+%   We say a type \(A\) is exhaustible if, for any decidable predicate \(P\) on
+%   \(A\), the universal quantification of the predicate is decidable.
+%   \begin{equation}
+%     \left( \Pi {(x : A)} , \mathbf{Dec}(P(x)) \right) \rightarrow \mathbf{Dec} \left( \Pi {(x : A)} , P(x) \right)
+%   \end{equation}
+% \end{rm-definition}
+
+% All of the finiteness predicates we have seen justify exhaustibility.
+% We will only prove it once, then, for the weakest:
+% \begin{rm-theorem}
+%   Kuratowski-finite types are exhaustible.
+% \end{rm-theorem}
+
+% Omniscience is stronger than exhaustibility, as we can derive the latter from
+% the former:
+% \begin{rm-lemma} \label{omniscient-is-exhaustible}
+%   Any omniscient type is exhaustible.
+% \end{rm-lemma}
+% \begin{proof}
+%   For decidable propositions, we know the following:
+%   \begin{equation} \label{lem-conversion}
+%     \Pi (x : A) , P(x) \leftrightarrow \neg \Sigma (x : A) , \neg P(x) 
+%   \end{equation}
+%   To derive exhaustibility from omniscience, then, we run the predicate in its
+%   negated form, and then subsequently negate the result.
+%   The resulting decision over \(\neg \Sigma(x : A) , \neg P(x)\) can be
+%   converted into \(\Pi (x : A) , P(x)\).
+% \end{proof}
+% We cannot derive, however, that any exhaustible type is omniscient, as we do not
+% have the inverse of equation~\ref{lem-conversion}:
+% \begin{equation} \label{choice-conversion}
+%   \Sigma (x : A) , P(x) \leftrightarrow \neg \Pi (x : A) , \neg P(x)
+% \end{equation}
+% Such an equation would allow us to pick a representative element from any type,
+% which is therefore non-constructive.
+% In a sense, equation~\ref{lem-conversion} requires a form of LEM on the
+% proposition (i.e. requires it to be decidable), whereas
+% equation~\ref{choice-conversion} requires a form of choice.
+% Those finiteness predicates which are ordered do in fact give us this form of
+% choice, so the conversion is valid.
+% As such, all of the ordered finiteness predicates imply omniscience.
+% Again, we will prove it only for the weakest.
+% \begin{rm-theorem}
+%   Manifest enumerable types are omniscient.
+% \end{rm-theorem}
+% Finally, we do have a form of omniscience for prop-valued predicates, as they do
+% not care about the chosen representative.
+% \begin{rm-theorem}
+%   Kuratowski finite types are omniscient about prop-valued predicates.
+% \end{rm-theorem}
+% \subsection{Synthesising Pattern-Matching Proofs}
+% In particular, they can automate large proofs by analysing every possible case.
+% In \cite{firsovDependentlyTypedProgramming2015}, the \(\AgdaDatatype{Pauli}\)
+% group is used as an example.
+% \ExecuteMetaData[agda/Data/Pauli.tex]{def}
+% As \(\AgdaDatatype{Pauli}\) has 4 constructors, \(n\)-ary functions on
+% \(\AgdaDatatype{Pauli}\) may require up to \(4^n\) cases, making even simple
+% proofs prohibitively verbose.
+
+% The alternative is to derive the things we need from \(\mathcal{E}!\) somehow.
+% As \(\AgdaDatatype{Pauli}\) is a simple finite type, the instance can be defined
+% in a similar way to those in lemma~\ref{split-enum-bool-top-bot}.
+% From here we can already derive decidable equality, a function which requires 16
+% cases if implemented manually.
+
+% For proof search, the procedure is a well-known one in Agda
+% \cite{devrieseBrightSideType2011}: we ask for the result of a decision procedure
+% as an \emph{instance argument}, which will demand computation during
+% typechecking.
+% \subsection{Multiple Arguments}
+% The automation machinery above only deals with single-argument predicates.
+% This is not a problem, as we know that we can work with multiple arguments by
+% currying and uncurrying, since all of the finiteness predicates are closed under
+% \(\times\).
+% To automate away the curry/uncurry noise we will use instance search, building
+% on \cite{allaisGenericLevelPolymorphic2019} to develop a small interface to
+% generic \(n\)-ary functions and properties.
+% Our generic representation can handle dependent \(\Sigma\) and \(\prod\)
+% types (rather than their non-dependent counterparts, \(\times\) and
+% \(\rightarrow\)).
+% This extension was necessary for our use case: it is mentioned in the paper
+% as the obvious next step. 
+% We also implement the curry-uncurry combinators as (verified) isomorphisms.
+
+% A full explanation of our implementation is beyond the scope of this work, so we
+% only present the finished interface, which is used like so:
+% \ExecuteMetaData[agda/Data/Pauli.tex]{assoc-prf}
+% \subsection{Synthesising Functions}
+% Finally, thanks to extensionality provided to us by HoTT, and the computation
+% properties provided by CuTT, we can derive decidable equality on functions over
+% finite types.
+% We can also use functions in our proof search.
+% Here, for instance, is a automated procedure which finds the
+% \(\AgdaFunction{not}\) function on \(\mathbf{Bool}\), given a specification.
+% \ExecuteMetaData[agda/Data/Pauli.tex]{not-spec}
+% \section{Related Work}
+% \cite{spiwackConstructivelyFinite2010} explored the notion of ``constructive
+% finite sets'', demonstrating that there is a rich variety of finiteness
+% predicates in a constructive setting.
+% Some of those predicates where explored more in
+% \cite{firsovVariationsNoetherianness2016a},
+% \cite{parmannInvestigatingStreamlessSets2015}, and
+% \cite{uustaluFinitenessRationalSequences2017a}.
+
+% \cite{yorgeyCombinatorialSpeciesLabelled2014} explored some of the definitions
+% we have given here (manifest bishop finiteness and cardinal finiteness) in HoTT,
+% in order to construct species.
+% Some of this paper is a formalisation of the early parts of that work.
+
+% \cite{firsovDependentlyTypedProgramming2015} explores finite sets in Agda and
+% uses them to automate proof machinery in much the same way as we have here.
+% The primary difference between their proof automation and ours is that we can
+% talk about finite functions, as we are working in HoTT.
+
+% \cite{fruminFiniteSetsHomotopy2018} explores many of the same notions as we do
+% here: cardinal finiteness, and Kuratowski finite sets.
+% Our work makes extensive use of the computational properties of CuTT, however,
+% and we are able to formalise the finiteness of function objects.
 \bibliographystyle{splncs04}

### 2020-01-19 figures/finite-classification.tex
down to 16 pages
more

@@
-\begin{figure}[h]
+\begin{figure}
   \centering
@@
       \; \ref{kuratowski}
-      \ar[ll, start anchor=north west, end anchor=north east, bend right=10,  "\text{linOrd}" description]
       \ar[uu, xshift=1ex, bend right=30,  "\text{Discrete}" description]
@@
   \end{tikzcd}
-  \caption{Classification of the strong definitions of finiteness, according to
-    whether they are discrete (imply decidable equality) and whether they induce
-    a linear order.}
+  \caption{Classification of finiteness predicates according to whether they are
+    discrete (imply decidable equality) and whether they induce a linear order.}
   \label{finite-classification}
-\end{figure}
+\end{figure}%

### 2020-01-19 paper.tex
more

@@
   full functions given desired properties.
+  \keywords{Agda \and Homotopy Type Theory \and Cubical Type Theory \and
+    Dependent Types \and Finiteness \and Topos \and Kuratowski finite}
 \end{abstract}
+
 \section{Introduction}
@@
 Figure~\ref{finite-classification}.
-We will explore two aspects of each predicate: its relation with the other
+We will explore two aspects of each predicate: its relation to the other
 predicates, and its topos-theoretic classification.
@@
 is manifest enumerable.
-The labelled arrows are strengthening proofs: i.e. every manifest enumerable set
+The labelled arrows are strengthening proofs: i.e., every manifest enumerable set
 \emph{with decidable equality} is manifest Bishop finite.
@@
 \subsection{Notation}
+\todo[inline]{Basic MLTT Notation: \(\Pi\), \(\Sigma\), etc}
+\todo[inline]{Paths, Univalence, isomorphisms}
+\todo[inline]{HITs}
 
 \section{Finiteness Predicates}
+\subsection{Kuratowski Finiteness}
+\begin{rm-definition}[Kuratowski Finite Subset]
+  \(\mathcal{K}(A)\) is the type of Kuratowski-finite subsets of \(A\).
+  \begin{equation}
+    \begin{aligned}
+      \mathcal{K}(A) =&
+             \; \eta : A \rightarrow \mathcal{K}(A) ; \\
+      \vert& \; \emptyset : \mathcal{K}(A) ; \\
+      \vert& \; \wc \cup \wc : \mathcal{K}(A) \rightarrow \mathcal{K}(A) \rightarrow \mathcal{K}(A) ; \\
+      \vert& \; \text{associative} : \Pi (\mathit{xs}, \mathit{ys}, \mathit{zs}: \mathcal{K}(A)) , (\mathit{xs} \cup \mathit{ys}) \cup \mathit{zs} \equiv \mathit{xs} \cup (\mathit{ys} \cup \mathit{zs}) ; \\
+      \vert& \; \text{identity} : \Pi (\mathit{xs}: \mathcal{K}(A)) , \emptyset \cup \mathit{xs} \equiv \mathit{xs} ; \\
+      \vert& \; \text{commutative} : \Pi (\mathit{xs}, \mathit{ys}: \mathcal{K}(A)) , \mathit{xs} \cup \mathit{ys} \equiv \mathit{ys} \cup \mathit{xs} ; \\
+      \vert& \; \text{idempotent} : \Pi (\mathit{xs}: \mathcal{K}(A)) , \mathit{xs} \cup \mathit{xs} \equiv \mathit{xs} ; \\
+      \vert& \; \text{trunc} : \Pi (\mathit{xs}, \mathit{ys}: \mathcal{K}(A)) , \Pi (p, q : \mathit{xs} \equiv \mathit{ys}) , p \equiv q ;
+    \end{aligned}
+  \end{equation}
+\end{rm-definition}
+The Kuratowski finite subset is a free join semilattice (or, equivalently, a
+free commutative idempotent monoid).
+Though it would be possible to use this type as-is, we found it easier to work
+with the following type as it has fewer path constructors.
+\begin{equation}
+  \begin{aligned}
+    \mathcal{K}(A) =&
+           \; [] : \mathcal{K}(A) ; \\
+    \vert& \; \wc \dblcolon \wc : A \rightarrow \mathcal{K}(A) \rightarrow \mathcal{K}(A) ; \\
+    \vert& \; \text{com} : \Pi (x, y: A) , \Pi (\mathit{xs} : \mathcal{K}(A)) , x \dblcolon y \dblcolon \mathit{xs} \equiv y \dblcolon x \dblcolon \mathit{xs} ; \\
+    \vert& \; \text{dup} : \Pi (x : A) , \Pi (\mathit{xs} : \mathcal{K}(A)) , x \dblcolon x \dblcolon \mathit{xs} \equiv x \dblcolon \mathit{xs} ; \\
+    \vert& \; \text{trunc} : \Pi (\mathit{xs}, \mathit{ys}: \mathcal{K}(A)) , \Pi (p, q : \mathit{xs} \equiv \mathit{ys}) , p \equiv q ;
+  \end{aligned}
+\end{equation}
+What we've done here is basically pulled out the ``monoid'' part of ``free
+commutative idempotent monoid'': the first two constructors now construct a
+list, which is a representation of the free monoid \emph{without} quotient
+types.
+This obviates the need for the associative and identity constructors.
+We have proven both types equivalent in our formalisation.
+
+Next, we need a way to say that a type is Kuratowski finite, rather than just
+some subset.
+For that, we will need to define membership of \(\mathcal{K}\).
+\begin{rm-definition}[Membership of \mathcal{K}]
+  First, we need to provide equations for the two point constructors.
+  \(x \in [] = \bot\), and \(x \in y \dblcolon \mathit{xs} = \lVert (x \equiv y)
+  + (x \in \mathit{xs}) \rVert \).
+  The \(\text{com}\) and \(\text{dup}\) constructors are handled by proving that
+  the truncated form of \(+\) is itself commutative and idempotent.
+  The type of propositions is itself a set, satisfying the \(\text{trunc}\)
+  constructor.
+\end{rm-definition}
+\begin{rm-definition}[Kuratowski Finiteness]
+  A type is Kuratowski finite iff there exists a Kuratowski finite subset which
+  contains all of its elements.
+  \begin{equation}
+    \mathcal{K}^{f}(A) = \Sigma {(\mathit{xs} : \mathcal{K}(A))} , \Pi (x : A) , x \in \mathit{xs}
+  \end{equation}
+\end{rm-definition}
+
 \subsection{Split Enumerable}
@@
   component the ``cover'' proof.
-  This definition is what is referred to as \verb+Listable+ in
-  \cite{firsovDependentlyTypedProgramming2015}.
 \end{rm-definition}
-
+In this, and the rest of the definitions in the paper, we tend to prefer
+list-based notions of finiteness.
+This is purely a matter of perspective, however: the definition above is
+precisely equivalent to a split surjection from a finite prefix of the natural
+numbers.
+\begin{rm-lemma}
+  \begin{equation}
+    \mathcal{E}!(A) \simeq \Sigma (n : \mathbb{N}) , \left( \mathbf{Fin}(n) \twoheadrightarrow ! \; A \right)
+  \end{equation}
+\end{rm-lemma}
+We have proven this statement in our formalisation, but it is worth pointing out
+that the proof is nearly trivial: after inlining the definitions of
+\(\mathbf{List}\), \(\in\), and \(\twoheadrightarrow!\), it becomes clear that
+the two terms are basically the same anyway, modulo the reassociation of
+\(\Sigma\).
+
+Split enumerability is quite a strong predicate, implying both decidable
+equality and a total order on the underlying type.
+This
+
+\subsection{Manifest Bishop}
+\subsection{Manifest Enumerable}
+\subsection{Cardinal}
 \section{Relations Between Each Finiteness Definition}
@@
 % \cite{kuratowskiNotionEnsembleFini1920}.
-% \begin{rm-definition}[Kuratowski-Finite Set]
-%   The Kuratowski finite set is a free join semilattice (or, equivalently, a free
-%   commutative idempotent monoid).
-%   HITs are required to define this type
-%   \cite{altenkirchDefinableQuotientsType2011}:
-%   \begin{equation}
-%     \begin{aligned}
-%       \mathcal{K}(A) =&
-%        \; \cdot \dblcolon \cdot : A \times \mathcal{K}(A) \rightarrow \mathcal{K}(A) ; \\
-%       \vert& \; [] : \mathcal{K}(A) ; \\
-%       \vert& \; \text{com} : \Pi (x, y: A) , \Pi (\mathit{xs}: \mathcal{K}(A)) , x \dblcolon y \dblcolon \mathit{xs} \equiv y \dblcolon x \dblcolon \mathit{xs} ; \\
-%       \vert& \; \text{dup} : \Pi (\mathit{x}: A) , \Pi (\mathit{xs}: \mathcal{K}(A)) , x \dblcolon x \dblcolon \mathit{xs} \equiv x \dblcolon \mathit{xs} ; \\
-%       \vert& \; \text{trunc} : \Pi (\mathit{xs}, \mathit{ys}: \mathcal{K}(A)) , \Pi (p, q : \mathit{xs} \equiv \mathit{ys}) , p \equiv q ;
-%     \end{aligned}
-%   \end{equation}
-%   The \(\text{com}\) and \(\text{dup}\) constructors effectively add
-%   commutativity and idempotency to the free monoid (the list), which is made by
-%   the first two constructors.
-%   The last constructor makes \(\mathcal{K}(A)\) a set.
-
-%   To eliminate from \(\mathcal{K}(A)\), we have to provide equations for each of
-%   the point constructors which obey the equations of the path constructors.
-%   For \(\text{com}\) and \(\text{dup}\), this means ensuring that the fold is
-%   commutative and idempotent, whereas \(\text{trunc}\) means we can only
-%   eliminate into sets.
-% \end{rm-definition}
 
@@
 % and proven equivalent to the representation here.
-% \begin{rm-definition}[Membership of \mathcal{K}]
-%   First, we need to provide equations for the two point constructors.
-%   \(x \in [] = \bot\), and \(x \in y \dblcolon \mathit{xs} = \lVert (x \equiv y)
-%   + (x \in \mathit{xs}) \rVert \).
-%   The \(\text{com}\) and \(\text{dup}\) constructors are handled by proving that
-%   the truncated form of \(+\) is itself commutative and idempotent.
-%   The type of propositions is itself a set, satisfying the \(\text{trunc}\)
-%   constructor.
-% \end{rm-definition}
-% \begin{rm-definition}[Kuratowski Finiteness]
-%   A type is Kuratowski finite iff there exists a Kuratowski Set which contains
-%   all of its elements.
-%   \begin{equation}
-%     \mathcal{K}^{f}(A) = \Sigma {(\mathit{xs} : \mathcal{K}(A))} , \Pi (x : A) , x \in \mathit{xs}
-%   \end{equation}
-% \end{rm-definition}
 
@@
 % and we are able to formalise the finiteness of function objects.
+
+\todo[inline]{Add funding ack:
+
+  This work has been supported by the Science Foundation Ireland under the
+  following grant: 13/RC/2D94 to Irish Software Research Centre.
+}
 \bibliographystyle{splncs04}

### 2020-01-20 paper.tex
more

@@
 \emph{with decidable equality} is manifest Bishop finite.
+Our most significant result here is the proof that a cardinal finite set with a
+decidable total order is manifestly Bishop finite.
 
@@
 HoTT do \emph{not}, in fact, form a topos, rather a \(\Pi W\)-pretopos.
-
-\todo[inline]{redo these last two paragraphs}
+Our main result here is that the category of decidable Kuratowski finite sets
+forms a \(\Pi\)-pretopos.
+
 After the finite predicates, we will briefly look at the infinite countable
 types, and classify them in a parallel way to the finite predicates.
+We here show that countably finite sets form a \(W\)-pretopos.
 
@@
 \todo[inline]{HITs}
-
 \section{Finiteness Predicates}
+While our primary interest is in decidable Kuratowski finiteness, we found it
+useful to examine the other finiteness predicates as well, if only as a way to
+construct proofs which can then be transferred to proof on Kuratowski
+finiteness.
+This, in turn, is why the relation proofs are so important: in HoTT, if two
+types are proved equivalent, we get access to the extremely powerful univalence
+axiom, which allows us to switch between types freely depending on what is more
+convenient.
+The \(\Pi\)-topos proof, for instance, is defined on cardinal finiteness, not
+decidable Kuratowski finiteness: an equivalence proof between the two types
+means we can transfer it over, formally, and with good justification.
+Similarly, many of the closure proofs are more naturally defined over the
+enumerated finiteness predicates.
 \subsection{Kuratowski Finiteness}
+The first thing we must define is a representation of subsets.
 \begin{rm-definition}[Kuratowski Finite Subset]
   \(\mathcal{K}(A)\) is the type of Kuratowski-finite subsets of \(A\).
-  \begin{equation}
-    \begin{aligned}
-      \mathcal{K}(A) =&
-             \; \eta : A \rightarrow \mathcal{K}(A) ; \\
-      \vert& \; \emptyset : \mathcal{K}(A) ; \\
-      \vert& \; \wc \cup \wc : \mathcal{K}(A) \rightarrow \mathcal{K}(A) \rightarrow \mathcal{K}(A) ; \\
-      \vert& \; \text{associative} : \Pi (\mathit{xs}, \mathit{ys}, \mathit{zs}: \mathcal{K}(A)) , (\mathit{xs} \cup \mathit{ys}) \cup \mathit{zs} \equiv \mathit{xs} \cup (\mathit{ys} \cup \mathit{zs}) ; \\
-      \vert& \; \text{identity} : \Pi (\mathit{xs}: \mathcal{K}(A)) , \emptyset \cup \mathit{xs} \equiv \mathit{xs} ; \\
-      \vert& \; \text{commutative} : \Pi (\mathit{xs}, \mathit{ys}: \mathcal{K}(A)) , \mathit{xs} \cup \mathit{ys} \equiv \mathit{ys} \cup \mathit{xs} ; \\
-      \vert& \; \text{idempotent} : \Pi (\mathit{xs}: \mathcal{K}(A)) , \mathit{xs} \cup \mathit{xs} \equiv \mathit{xs} ; \\
-      \vert& \; \text{trunc} : \Pi (\mathit{xs}, \mathit{ys}: \mathcal{K}(A)) , \Pi (p, q : \mathit{xs} \equiv \mathit{ys}) , p \equiv q ;
-    \end{aligned}
-  \end{equation}
-\end{rm-definition}
-The Kuratowski finite subset is a free join semilattice (or, equivalently, a
-free commutative idempotent monoid).
-Though it would be possible to use this type as-is, we found it easier to work
-with the following type as it has fewer path constructors.
   \begin{equation}
@@
   \end{equation}
+\end{rm-definition}
+The Kuratowski finite subset is a free join semilattice (or, equivalently, a
+free commutative idempotent monoid).
+From a programming perspective, it could well be seen as a finite set (in the
+sense of the data structure).
+In most languages, it might be represented by a hash table or some balanced
+tree: HITs allow us to represent precisely the notion of a finite set, without
+putting any constraints on the contents, or choosing any internal data
+structure.
+
+Though it would be possible to use this type as-is, we found it easier to work
+with the following type as it has fewer path constructors.
 What we've done here is basically pulled out the ``monoid'' part of ``free
@@
 \section{Topos}
+\begin{itemize}
+  \item Complete (pullbacks + terminal object)
+\end{itemize}
 \section{Countably Infinite Types}

### 2020-01-20 paper.tex
more

@@
 omniscience and exhaustibility.
-We also use this section to demonstrate some \emph{computation} using
-univalence: our proof search library can deal seamlessly with functions, testing
-them for equality, synthesising them, etc.
+While proof search based on finiteness is not new, implementing it here does
+give us an opportunity to demonstrate some actual \emph{computation} in a
+univalent setting.
+Many of the finiteness predicates are built using univalence, so for the proof
+search functionality to work at all we require computational content in the
+univalence axiom.
+Furthermore, the extensionality afforded to us by HoTT means that we can have
+\(\Pi\) types in the domain of the search, meaning we can synthesise full
+functions from specifications alone.
 \subsection{Related Work}
@@
 \section{Finiteness Predicates}
-While our primary interest is in decidable Kuratowski finiteness, we found it
-useful to examine the other finiteness predicates as well, if only as a way to
-construct proofs which can then be transferred to proof on Kuratowski
-finiteness.
-This, in turn, is why the relation proofs are so important: in HoTT, if two
-types are proved equivalent, we get access to the extremely powerful univalence
-axiom, which allows us to switch between types freely depending on what is more
-convenient.
-The \(\Pi\)-topos proof, for instance, is defined on cardinal finiteness, not
-decidable Kuratowski finiteness: an equivalence proof between the two types
-means we can transfer it over, formally, and with good justification.
-Similarly, many of the closure proofs are more naturally defined over the
-enumerated finiteness predicates.
+In this section, we will define and briefly describe each of the five predicates
+in Figure~\ref{finite-classification}.
+The reason we explore predicates other than our focus (Kuratowski finiteness) is
+that we can often prove things like closure much more readily on the simpler
+predicates.
+The relations (which we will prove in the next section) then allow us to
+transfer those proofs onto Kuratowski finiteness.
 \subsection{Kuratowski Finiteness}
@@
 free commutative idempotent monoid).
-From a programming perspective, it could well be seen as a finite set (in the
-sense of the data structure).
-In most languages, it might be represented by a hash table or some balanced
-tree: HITs allow us to represent precisely the notion of a finite set, without
-putting any constraints on the contents, or choosing any internal data
-structure.
-
-Though it would be possible to use this type as-is, we found it easier to work
-with the following type as it has fewer path constructors.
-What we've done here is basically pulled out the ``monoid'' part of ``free
-commutative idempotent monoid'': the first two constructors now construct a
-list, which is a representation of the free monoid \emph{without} quotient
-types.
-This obviates the need for the associative and identity constructors.
-We have proven both types equivalent in our formalisation.
-
-Next, we need a way to say that a type is Kuratowski finite, rather than just
-some subset.
+More prosaically, \(\mathcal{K}\) is the abstract data type for finite sets, as
+defined in the Boom hierarchy \cite{boomFurtherThoughtsAbstracto1981,
+  bunkenburgBoomHierarchy1994}.
+However, rather than just being a specification, \(\mathcal{K}\) is fully usable
+as a data type in its own right, thanks to HITs.
+
+Other definitions of \(\mathcal{K}\) exist (such as the one in
+\cite{fruminFiniteSetsHomotopy2018}) which make the fact that \(\mathcal{K}\) is
+the free join semilattice more obvious.
+We have included such a definition in our formalisation, and proven it
+equivalent to the one above.
+
+Next, we need a way to say that an entire type is Kuratowski finite.
 For that, we will need to define membership of \(\mathcal{K}\).
-\begin{rm-definition}[Membership of \mathcal{K}]
+\begin{rm-definition}[Membership of \(\mathcal{K}\)]
   First, we need to provide equations for the two point constructors.
@@
 \end{rm-definition}
+Finally, we have enough background to define Kuratowski finiteness.
 \begin{rm-definition}[Kuratowski Finiteness]
-  A type is Kuratowski finite iff there exists a Kuratowski finite subset which
-  contains all of its elements.
   \begin{equation}
@@
 \end{rm-definition}
+\todo[inline]{Kf is a proposition, circle is Kf}
 

### 2020-01-20 paper.tex
more

@@
   topos-theoretically.
-  For instance we show that the category of decidable Kuratowski finite sets
+  This includes a proof that the category of decidable Kuratowski finite sets
   (also called the category of cardinal finite sets) form a \(\Pi\)-pretopos.
@@
 
-  We formalise our work in Cubical Agda, and we implement a library for proof
-  search (including combinators for level-polymorphic fully generic currying),
-  and demonstrate how it can be used to both prove properties and synthesise
-  full functions given desired properties.
+  We formalise our work in Cubical Agda, where we implement a library for proof
+  search (including combinators for level-polymorphic fully generic currying).
+  Through this library we demonstrate a number of uses for the computational
+  content of the univalence axiom, including searching for and synthesising
+  functions.
   \keywords{Agda \and Homotopy Type Theory \and Cubical Type Theory \and
@@
 \todo[inline]{HITs}
+\todo[inline]{Surjections}
+\todo[inline]{Containers}
+\todo[inline]{Fin}
+\todo[inline]{Biconditional, Isomorphism, Equivalence, Equals}
 \section{Finiteness Predicates}
@@
 in Figure~\ref{finite-classification}.
-The reason we explore predicates other than our focus (Kuratowski finiteness) is
+The reason we explore predicates other than our focus (cardinal finiteness) is
 that we can often prove things like closure much more readily on the simpler
@@
 \subsection{Kuratowski Finiteness}
+Much work has already been done on Kuratowski finiteness in HoTT in
+\cite{fruminFiniteSetsHomotopy2018}.
+As a result, we will not needlessly repeat proofs, rather we will just give a
+brief introduction to the topic and point out where our treatment differs.
+
 The first thing we must define is a representation of subsets.
-\begin{rm-definition}[Kuratowski Finite Subset]
+\begin{romdefinition}[Kuratowski Finite Subset]
   \(\mathcal{K}(A)\) is the type of Kuratowski-finite subsets of \(A\).
@@
   \end{equation}
-\end{rm-definition}
+\end{romdefinition}
 The Kuratowski finite subset is a free join semilattice (or, equivalently, a
@@
 For that, we will need to define membership of \(\mathcal{K}\).
-\begin{rm-definition}[Membership of \(\mathcal{K}\)]
-  First, we need to provide equations for the two point constructors.
-  \(x \in [] = \bot\), and \(x \in y \dblcolon \mathit{xs} = \lVert (x \equiv y)
-  + (x \in \mathit{xs}) \rVert \).
+\begin{romdefinition}[Membership of \(\mathcal{K}\)]
+  Membership is defined by pattern-matching on \(\mathcal{K}\).
+  The two point constructors are handled like so:
+  \begin{equation}
+    \begin{alignedat}{2}
+      x \in&& \; []                      &= \bot \\
+      x \in&& \; y \dblcolon \mathit{ys} &= \lVert x \equiv y \uplus x \in \mathit{ys} \rVert
+    \end{alignedat}
+  \end{equation}
   The \(\text{com}\) and \(\text{dup}\) constructors are handled by proving that
-  the truncated form of \(+\) is itself commutative and idempotent.
+  the truncated form of \(\uplus\) is itself commutative and idempotent.
   The type of propositions is itself a set, satisfying the \(\text{trunc}\)
   constructor.
-\end{rm-definition}
+\end{romdefinition}
 Finally, we have enough background to define Kuratowski finiteness.
-\begin{rm-definition}[Kuratowski Finiteness]
+\begin{romdefinition}[Kuratowski Finiteness]
   \begin{equation}
@@
   \end{equation}
-\end{rm-definition}
-\todo[inline]{Kf is a proposition, circle is Kf}
-
+\end{romdefinition}
+
+We also have the following two lemmas, proven in both
+\cite{fruminFiniteSetsHomotopy2018} and our formalisation.
+\begin{romlemma}
+  \(\mathcal{K}^f\) is a mere proposition.
+\end{romlemma}
+\begin{romlemma}
+  This circle \(S^1\) is Kuratowski finite.
+\end{romlemma}
+The second of these in particular tells us that the ``Kuratowski-finite types''
+are not necessarily sets; it also tells us that we cannot derive decidable
+equality from a proof of Kuratowski finiteness.
 \subsection{Split Enumerable}
-\begin{rm-definition}[Split Enumerable Set]
+\begin{romdefinition}[Split Enumerable Set]
   \begin{equation}
-    \mathcal{E}!(A) = \Sigma {(\mathit{xs} : \textbf{List}(A))} , \Pi {(x : A)} , x \in xs
+    \mathcal{E}!(A) = \Sigma {(\mathit{xs} : \textbf{List}(A))} , \Pi {(x : A)} , x \in \mathit{xs}
   \end{equation}
@@
   component the ``cover'' proof.
-\end{rm-definition}
-In this, and the rest of the definitions in the paper, we tend to prefer
-list-based notions of finiteness.
+  An equivalent version of this predicate was called \verb+Listable+ in
+  \cite{firsovDependentlyTypedProgramming2015}.
+\end{romdefinition}
+
+We tend to prefer list-based definitions of finiteness, rather than ones based
+on bijections or surjections.
 This is purely a matter of perspective, however: the definition above is
@@
 numbers.
-\begin{rm-lemma}
+\begin{romlemma} \label{split-enum-is-split-surj}
   \begin{equation}
-    \mathcal{E}!(A) \simeq \Sigma (n : \mathbb{N}) , \left( \mathbf{Fin}(n) \twoheadrightarrow ! \; A \right)
+    \mathcal{E}!(A) \equiv \Sigma (n : \mathbb{N}) , \left( \mathbf{Fin}(n) \twoheadrightarrow ! \; A \right)
   \end{equation}
-\end{rm-lemma}
-We have proven this statement in our formalisation, but it is worth pointing out
-that the proof is nearly trivial: after inlining the definitions of
-\(\mathbf{List}\), \(\in\), and \(\twoheadrightarrow!\), it becomes clear that
-the two terms are basically the same anyway, modulo the reassociation of
+\end{romlemma}
+\todo[inline]{Put refs for each ``definition of''}
+\begin{proof}
+  \begin{align*}
+    \mathcal{E}!(A)
+    & \equiv \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , x \in \mathit{xs}
+    & \text{Definition of } \mathcal{E}!
+    \\
+    & \equiv \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{fib}_{\text{snd}(\mathit{xs})}(x)
+    & \text{Definition of } \in
+    \\
+    & \equiv \Sigma (\mathit{xs} : \textbf{List}(A)) , \text{sp-surj}(\text{snd}(\mathit{xs}))
+    & \text{Definition of sp-surj}
+    \\
+    & \equiv \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket (A)) , \text{sp-surj}(\text{snd}(\mathit{xs}))
+    & \text{Definition of } \mathbf{List}
+    \\
+    & \equiv \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \mathbf{Fin}(n)) , A) , \text{sp-surj}(\text{snd}(\mathit{xs}))
+    & \text{Definition of } \llbracket \wc \rrbracket
+    \\
+    & \equiv \Sigma (n : \mathbb{N}) , \Sigma (f : \mathbf{Fin}(n) \rightarrow A) , \text{sp-surj}(f)
+    & \text{Reassociation of } \Sigma
+    \\
+    & \equiv \Sigma (n : \mathbb{N}) , ( \mathbf{Fin}(n) \twoheadrightarrow ! \; A )
+    & \text{Definition of } \twoheadrightarrow!
+  \end{align*}
+  \qed
+\end{proof}
+
+In our formalisation, the proof is a single line: most of the steps above are
+simple expansion of definitions.
+The only step which isn't definitional equality is the reassociation of
 \(\Sigma\).
 
-Split enumerability is quite a strong predicate, implying both decidable
-equality and a total order on the underlying type.
-This
-
+Split enumerability implies decidable equality on the underlying type.
+To prove this, we will make use of the following lemma, proven in the
+formalisation:
+\begin{romlemma} \label{discrete-surj}
+  \begin{equation}
+    \frac{
+        A \twoheadrightarrow! \; B \; \; \; \text{Discrete}(A)
+      }{
+       \text{Discrete}(B) 
+      }
+  \end{equation}
+\end{romlemma}
+\begin{romlemma}
+  Every split enumerable type is discrete.
+\end{romlemma}
+\begin{proof}
+  Let \(A\) be a split enumerable type.
+  By lemma~\ref{split-enum-is-split-surj}, there is a surjection from
+  \(\mathbf{Fin}(n)\) for some \(n\).
+  Also, by lemma \todo[noline]{Fin is discrete}, we know that
+  \(\mathbf{Fin}(n)\) is discrete.
+  Therefore, by lemma~\ref{discrete-surj}, \(A\) is discrete.
+  \qed
+\end{proof}
 \subsection{Manifest Bishop}
+\begin{romdefinition}[Manifest Bishop Finiteness]
+  \begin{equation}
+    \mathcal{B}(A) = \Sigma {(\mathit{xs} : \textbf{List}(A))} , \Pi {(x : A)} , x \inunique xs
+  \end{equation}
+\end{romdefinition}
+The only difference between manifest Bishop finiteness and split enumerability
+is the membership term: here we require unique membership (\(\inunique\)),
+rather than simple membership (\(\in\)).
+
+Where split enumerability was the enumeration form of a split surjection from
+\(\mathbf{Fin}\), manifest Bishop finiteness is the enumeration form of an
+\emph{equivalence} with \(\mathbf{Fin}\).
+\begin{romlemma} \label{bishop-equiv}
+  \begin{equation}
+    \mathcal{B}(A) \simeq \Sigma {(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \simeq A \right)
+  \end{equation}
+\end{romlemma}
+\begin{proof}
+  \begin{align*}
+    \mathcal{B}(A)
+    & \equiv \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , x \inunique \mathit{xs}
+    & \text{Definition of } \mathcal{B}
+    \\
+    & \equiv \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{isContr} \left(\text{fib}_{\text{snd}(\mathit{xs})}(x)\right)
+    & \text{Definition of } \inunique
+    \\
+    & \equiv \Sigma (\mathit{xs} : \textbf{List}(A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
+    & \text{\cite[theorem 4.4.5]{hottbook}}
+    \\
+    & \equiv \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket (A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
+    & \text{Definition of } \mathbf{List}
+    \\
+    & \equiv \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \mathbf{Fin}(n)) , A) , \text{isEquiv}(\text{snd}(\mathit{xs}))
+    & \text{Definition of } \llbracket \wc \rrbracket
+    \\
+    & \equiv \Sigma (n : \mathbb{N}) , \Sigma (f : \mathbf{Fin}(n) \rightarrow A) , \text{isEquiv}(f)
+    & \text{Reassociation of } \Sigma
+    \\
+    & \equiv \Sigma (n : \mathbb{N}) , ( \mathbf{Fin}(n) \simeq \; A )
+    & \text{Definition of } \twoheadrightarrow!
+  \end{align*}
+  \qed
+\end{proof}
 \subsection{Manifest Enumerable}
+\begin{romdefinition}[Manifest Enumerability]
+  \begin{equation}
+    \mathcal{E}(A) = \Sigma {(\mathit{xs} : \mathbf{List}(A))} , \Pi {(x : A)} , \lVert x \in \mathit{xs} \rVert
+  \end{equation}
+\end{romdefinition}
+Again, the only difference with this type and split enumerability is the
+membership proof.
+Here we have propositionally truncated it.
+
+% We will use the circle as an example non-set type which is nonetheless
+% manifestly enumerable.
+% \begin{romtheorem}
+%   The circle \(S^1\) is manifestly enumerable.
+% \end{romtheorem}
+% \begin{proof}
+%   As the cover proof is a truncated proposition, we need only consider the point
+%   constructors, making this poof the same as the proof of split enumerability on
+%   \(\top\).
+% \end{proof}
+% \subsection{Surjections}
+% This predicates relation to surjectivity is much the same as split
+% enumerability's relation to \emph{split} surjectivity.
+% \begin{romlemma}
+%   A proof of manifest enumerability is equivalent to a surjection from a finite
+%   prefix of the natural numbers.
+% \end{romlemma}
 \subsection{Cardinal}
@@
 % work.
-% \begin{rm-definition}[Container] \label{container-def}
+% \begin{romdefinition}[Container] \label{container-def}
 %   A container \cite{abbottContainersConstructingStrictly2005} is a pair
@@
 %   Where \(\text{fiber}\) is from \cite[definition 4.2.4]{hottbook}.
-% \end{rm-definition}
-% \begin{rm-definition}[Lists]
+% \end{romdefinition}
+% \begin{romdefinition}[Lists]
 %   \begin{equation}%
@@
 %   \end{equation}
-% \end{rm-definition}
-% \begin{rm-definition}[\mathbf{Fin}]
+% \end{romdefinition}
+% \begin{romdefinition}[\mathbf{Fin}]
 %   \(\mathbf{Fin}(n)\) is the type of natural numbers smaller than \(n\).
@@
 %   \(\mathbf{Fin}(n + 1) &= \top + \mathbf{Fin}(n)\).
-% \end{rm-definition}
+% \end{romdefinition}
 
@@
 % \subsection{Split Surjections}
-% \begin{rm-definition}[Surjections]
+% \begin{romdefinition}[Surjections]
 %   \begin{multicols}{2}
@@
 %   4.6.1]{hottbook}.
-% \end{rm-definition}
-% \begin{rm-theorem} \label{split-enum-surj}
+% \end{romdefinition}
+% \begin{romtheorem} \label{split-enum-surj}
 %   Split enumerability is equivalent to a split surjection from a finite prefix
@@
 %   \end{equation}
-% \end{rm-theorem}
+% \end{romtheorem}
 % \begin{proof}
@@
 % \end{proof}
-% \begin{rm-lemma} \label{split-enum-discrete}
+% \begin{romlemma} \label{split-enum-discrete}
 %   Any split enumerable type has decidable equality (is discrete).
-% \end{rm-lemma}
+% \end{romlemma}
 % \begin{proof}
@@
 % section~\ref{manifest-bishop}.
-% \begin{rm-lemma} \label{split-enum-bool-top-bot}
+% \begin{romlemma} \label{split-enum-bool-top-bot}
 %   \(\mathbf{Bool}\), \(\top\), and \(\bot\) are all split enumerable.
-% \end{rm-lemma}
+% \end{romlemma}
 % \begin{proof}
@@
 % \end{proof}
-% \begin{rm-theorem} \label{split-enum-sigma}
+% \begin{romtheorem} \label{split-enum-sigma}
 %   Split-enumerability is closed under \(\Sigma\).
@@
 
-% \begin{rm-definition}[Manifest Bishop Finiteness]
+% \begin{romdefinition}[Manifest Bishop Finiteness]
 %   \begin{equation}
@@
 %   \end{equation}
-% \end{rm-definition}
+% \end{romdefinition}
 % The only difference between this predicate and split enumerability is the list
@@
 % to be read as ``\(x\) occurs exactly once in \(\mathit{xs}\)''.
-% \begin{rm-definition}[Unique Membership]
+% \begin{romdefinition}[Unique Membership]
 %   We say an item \(x\) is ``uniquely in'' some container \(\mathit{xs}\) if its
@@
 %   \end{equation}
-% \end{rm-definition}
+% \end{romdefinition}
 % By the definition of a contraction, we can always recover the underlying
@@
 % \subsection{Equivalence}
-% \begin{rm-lemma} \label{bishop-equiv}
+% \begin{romlemma} \label{bishop-equiv}
 %   A proof of manifest Bishop finiteness is equivalent to an equivalence with a
@@
 %   \end{equation}
-% \end{rm-lemma}
+% \end{romlemma}
 % \begin{proof}
@@
 % \subsection{Relationship to Split Enumerability}
-% \begin{rm-theorem} \label{split-enum-to-manifest-bishop}
+% \begin{romtheorem} \label{split-enum-to-manifest-bishop}
 %   Any split enumerable set is manifest Bishop finite.
-% \end{rm-theorem}
+% \end{romtheorem}
 % \begin{proof}
@@
 % We remedy that here.
-% \begin{rm-theorem}
+% \begin{romtheorem}
 %   Manifest bishop finiteness is closed over dependent functions
@@
 % membership proof.
-% \begin{rm-definition}[Manifest Enumerability]
-%   \begin{equation}
-%     \mathcal{E}(A) = \Sigma {(\mathit{xs} : \mathbf{List}(A))} , \Pi {(x : A)} , \lVert x \in \mathit{xs} \rVert
-%   \end{equation}
-% \end{rm-definition}
-% \begin{rm-definition}[Propositional Truncation]a \label{prop-trunc}
-%   The type \(\lVert A \rVert\) on some type \(A\) is a propositionally truncated
-%   proof of \(A\) \cite[3.7]{hottbook}.
-%   In other words, it is a proof that some \(A\) exists, but it does not tell you
-%   \emph{which} \(A\).
-
-%   It is defined as a Higher Inductive Type:
-%   \begin{equation}
-%     \begin{aligned}
-%       \lVert A \rVert =& \; \lvert \cdot \rvert : A \rightarrow \lVert A \rVert ; \\
-%                       |& \; \text{squash} : \Pi {(x, y : \lVert A \rVert)} , x \equiv y  ; 
-%     \end{aligned}
-%   \end{equation}
-
-%   We will use three eliminators from \(\lVert A \rVert\) in this paper.
-%   \begin{enumerate}
-%   \item \label{elim-prop-prop} For any function \(A \rightarrow B\), where
-%     \(\text{isProp}(B)\), we have a function \(\lVert A \rVert \rightarrow B\).
-%   \item \label{elim-prop-monad} A special case of \ref{elim-prop-prop} implies
-%     that \(\lVert \cdot \rVert\) forms a monad: this means that we get the usual
-%     Monadic operators on \(\lVert \cdot \rVert\) (bind, pure, fmap, etc.).
-%     \item \label{elim-prop-coh} We can eliminate from \(\lVert A \rVert\) with a
-%       function \(f : A \rightarrow B\) iff \(f\) ``doesn't care'' about the
-%       choice of \(A\) (\(\Pi {(x , y : A)} , f(x) \equiv f(y) \)).
-%       Formally speaking, \(f\) needs to be ``coherently constant''
-%       \cite{krausGeneralUniversalProperty2015}, and \(B\) needs to be an
-%       \(n\)-type for some finite \(n\).
-%   \end{enumerate}
-% \end{rm-definition}
-
-% We will use the circle as an example non-set type which is nonetheless
-% manifestly enumerable.
-% \begin{rm-theorem}
-%   The circle \(S^1\) is manifestly enumerable.
-% \end{rm-theorem}
-% \begin{proof}
-%   As the cover proof is a truncated proposition, we need only consider the point
-%   constructors, making this poof the same as the proof of split enumerability on
-%   \(\top\).
-% \end{proof}
-% \subsection{Surjections}
-% This predicates relation to surjectivity is much the same as split
-% enumerability's relation to \emph{split} surjectivity.
-% \begin{rm-lemma}
-%   A proof of manifest enumerability is equivalent to a surjection from a finite
-%   prefix of the natural numbers.
-% \end{rm-lemma}
 % \subsection{Relation to Split Enumerability}
-% \begin{rm-theorem} \label{manifest-enum-to-split}
+% \begin{romtheorem} \label{manifest-enum-to-split}
 %   A manifestly enumerable type with decidable equality is split enumerable.
-% \end{rm-theorem}
+% \end{romtheorem}
 % \begin{proof}
@@
 % under \(\Pi\).
-% \begin{rm-lemma}
+% \begin{romlemma}
 %   Manifest enumerability is closed under \(\Sigma\).
-% \end{rm-lemma}
+% \end{romlemma}
 % \begin{proof}
@@
 % underlying type.
-% \begin{rm-definition}[Cardinal Finiteness]
+% \begin{romdefinition}[Cardinal Finiteness]
 %   A type \(A\) is cardinally finite, \(\mathcal{C}\), if it has a
@@
 %   \end{equation}%
-% \end{rm-definition}%
+% \end{romdefinition}%
 % \subsection{Closure}
@@
 % This leaves us with closure under only the non-dependent operations.
-% \begin{rm-lemma}
+% \begin{romlemma}
 %   Cardinal finiteness is closed under \(\times\), \(+\), and \(\rightarrow\).
-% \end{rm-lemma}
+% \end{romlemma}
 % \begin{proof}
@@
 % Our contribution for this section is simply the formalisation.
-% \begin{rm-theorem} \label{cardinal-finite-discrete}
+% \begin{romtheorem} \label{cardinal-finite-discrete}
 %   Any cardinal-finite set has decidable equality.
-% \end{rm-theorem}
+% \end{romtheorem}
 % \begin{proof}
@@
 % \end{proof}
-% \begin{rm-theorem}
+% \begin{romtheorem}
 %   Given a cardinally finite type, we can derive the type's cardinality, as well
@@
 
-% \begin{rm-theorem}
+% \begin{romtheorem}
 %   Any cardinal finite type with a (decidable) total order is manifestly Bishop
 %   finite.
-% \end{rm-theorem}
+% \end{romtheorem}
 % \begin{proof}
@@
 % First, permutations.
-% \begin{rm-definition}[List Permutations]
+% \begin{romdefinition}[List Permutations]
 %   We say that two lists are permutations of each other if there is an
@@
 %   \end{equation}
-% \end{rm-definition}
-% \begin{rm-lemma}
+% \end{romdefinition}
+% \begin{romlemma}
 %   Insertion sort is invariant under permutations.
@@
 
-% \begin{rm-theorem}
+% \begin{romtheorem}
 %   A proof of Kuratowski finiteness is equivalent to a propositionally truncated
@@
 % have a support \emph{stream}, which is its infinite.
-% \begin{rm-definition}[Stream]
+% \begin{romdefinition}[Stream]
 %   \begin{equation}
@@
 %   \end{equation}
-% \end{rm-definition}
-% \begin{rm-definition}[Split Countability]
+% \end{romdefinition}
+% \begin{romdefinition}[Split Countability]
 %   \begin{equation}
@@
 %   \end{equation}
-% \end{rm-definition}
+% \end{romdefinition}
 % This type is definitionally equal to it surjection equivalent (\(\mathbb{N}
@@
 % \input{figures/pairing-functions.tex}
-% \begin{rm-theorem} \label{split-countability-sigma}
+% \begin{romtheorem} \label{split-countability-sigma}
 %   Split countability is closed under \(\Sigma\).
-% \end{rm-theorem}
+% \end{romtheorem}
 % \begin{proof}
@@
 % This means that we have closure under lists.
-% \begin{rm-theorem}
+% \begin{romtheorem}
 %   Split countability is closed under Kleene star.
@@
 % of omniscience \cite{myhillErrettBishopFoundations1972}.
-% \begin{rm-definition}[Limited Principle of Omniscience]
+% \begin{romdefinition}[Limited Principle of Omniscience]
 %   For any type \(A\) and predicate \(P\) on \(A\), the limited principle of
@@
 %   that predicate is also decidable.
-% \end{rm-definition}
+% \end{romdefinition}
 % The limited principle of omniscience is non-constructive, but individual types
@@
 % There is also a universal form of omniscience, which we call exhaustibility.
-% \begin{rm-definition}[Exhaustibility]
+% \begin{romdefinition}[Exhaustibility]
 %   We say a type \(A\) is exhaustible if, for any decidable predicate \(P\) on
@@
 %   \end{equation}
-% \end{rm-definition}
+% \end{romdefinition}
 
@@
 % We will only prove it once, then, for the weakest:
-% \begin{rm-theorem}
+% \begin{romtheorem}
 %   Kuratowski-finite types are exhaustible.
-% \end{rm-theorem}
+% \end{romtheorem}
 
@@
 % the former:
-% \begin{rm-lemma} \label{omniscient-is-exhaustible}
+% \begin{romlemma} \label{omniscient-is-exhaustible}
 %   Any omniscient type is exhaustible.
-% \end{rm-lemma}
+% \end{romlemma}
 % \begin{proof}
@@
 % Again, we will prove it only for the weakest.
-% \begin{rm-theorem}
+% \begin{romtheorem}
 %   Manifest enumerable types are omniscient.
-% \end{rm-theorem}
+% \end{romtheorem}
 % Finally, we do have a form of omniscience for prop-valued predicates, as they do
 % not care about the chosen representative.
-% \begin{rm-theorem}
+% \begin{romtheorem}
 %   Kuratowski finite types are omniscient about prop-valued predicates.
-% \end{rm-theorem}
+% \end{romtheorem}
 % \subsection{Synthesising Pattern-Matching Proofs}

### 2020-01-20 paper.tex
more!

@@
 the relationships between them, and their topos-theoretic characterisation.
-We also briefly examine a definition for countable sets, comparing it to the
-manifestly Bishop finite sets.
+We also briefly examine a predicate for countable sets, comparing it to the
+finiteness predicates.
 Our work is formalised in Cubical Agda, where we also develop a library for
@@
 \subsection{Related Work}
-\subsection{Notation}
+\subsection{Notation and Background}
+We work in Cubical Type Theory \cite{cohenCubicalTypeTheory2016}.
+For the various type formers we use the following notation:
+\begin{description}
+  \item[Type] We use \(\mathbf{Type}\) to denote the universe of (small) types.
+    ``Type families'' are functions into type.
+  \item[Universes]
+    Implicitly, types will exist at some point in a universe hierarchy,
+    beginning with \(\mathbf{Type}\) (\(\mathbf{Type} : \mathbf{Type}_1\),
+    \(\mathbf{Type}_1 : \mathbf{Type}_2\), and so on).
+    In our formalisation, every proof is done in the most universe polymorphic
+    way possible: the proof that the cardinally finite sets form a
+    \(\Pi\)-pretopos, for instance, is defined over any universe level.
+  \item[0 , 1 , 2] We call the \(\mathbf{0}\), \(\mathbf{1}\), and
+    \(\mathbf{2}\) types \(\bot\), \(\top\), and \(\mathbf{Bool}\) respectively.
+    The single inhabitant of \(\top\) is tt, and the two inhabitants of
+    \(\mathbf{Bool}\) are false and true.
+  \item[Dependent Sum and Product] We use \(\Sigma\) and \(\Pi\) for the
+    dependent sum and product, respectively.
+    The two projections from \(\Sigma\) are called fst and snd.
+    In the non-dependent case, \(\Sigma\) can be written as \(\times\), and
+    \(\Pi\) as \(\rightarrow\).
+  \item[Disjoint Union] Disjoint union can be defined in terms of \(\Sigma\):
+    \begin{equation}
+      A \uplus B \coloneqq \Sigma(x : \mathbf{Bool}) , \text{if } x \text{ then } A \text{ else } B
+    \end{equation}
+    However, we prefer to use it as an inductively defined type (though the two
+    are equivalent).
+    The two constructors are:
+    \begin{equation}
+      \begin{aligned}
+        \text{inl} &: A \rightarrow A \uplus B \\
+        \text{inr} &: B \rightarrow A \uplus B
+      \end{aligned}
+    \end{equation}
+  \item[Equalities, equivalences, and paths] We use the symbol \(\coloneqq\)
+    for definitions.
+    \(\simeq\) will be used for equivalences, and \(\equiv\) for equalities.
+    Of course, we know that \((A \simeq B) \simeq (A \equiv B)\) by univalence,
+    so the distinction isn't terribly important in usage.
+\end{description}
+
+% \item[Isomorphisms and Biconditionals] A pair of functions \(f : A \rightarrow
+%   B\) and \(g : B \rightarrow A\) will be described by the type \(A
+%   \leftrightarrow B\).
+%   If \(g\) is a quasi-inverse \cite[definition 2.4.6]{hottbook} of \(f\), we
+%   have an isomorphism between \(A\) and \(B\), denoted \(A \iff B\).
+%   Note that while we can derive an equivalence from an isomorphism, we do not
+%   have \((A \iff B) \simeq (A \simeq B)\).
+
+We will use the standard \(\Pi\) and \(\Sigma\) notation for dependent products
+and sums, respectively.
+
 \todo[inline]{Basic MLTT Notation: \(\Pi\), \(\Sigma\), etc}
@@
 transfer those proofs onto Kuratowski finiteness.
-\subsection{Kuratowski Finiteness}
-Much work has already been done on Kuratowski finiteness in HoTT in
-\cite{fruminFiniteSetsHomotopy2018}.
-As a result, we will not needlessly repeat proofs, rather we will just give a
-brief introduction to the topic and point out where our treatment differs.
-
-The first thing we must define is a representation of subsets.
-\begin{romdefinition}[Kuratowski Finite Subset]
-  \(\mathcal{K}(A)\) is the type of Kuratowski-finite subsets of \(A\).
-  \begin{equation}
-    \begin{aligned}
-      \mathcal{K}(A) =&
-      \; [] : \mathcal{K}(A) ; \\
-      \vert& \; \wc \dblcolon \wc : A \rightarrow \mathcal{K}(A) \rightarrow \mathcal{K}(A) ; \\
-      \vert& \; \text{com} : \Pi (x, y: A) , \Pi (\mathit{xs} : \mathcal{K}(A)) , x \dblcolon y \dblcolon \mathit{xs} \equiv y \dblcolon x \dblcolon \mathit{xs} ; \\
-      \vert& \; \text{dup} : \Pi (x : A) , \Pi (\mathit{xs} : \mathcal{K}(A)) , x \dblcolon x \dblcolon \mathit{xs} \equiv x \dblcolon \mathit{xs} ; \\
-      \vert& \; \text{trunc} : \Pi (\mathit{xs}, \mathit{ys}: \mathcal{K}(A)) , \Pi (p, q : \mathit{xs} \equiv \mathit{ys}) , p \equiv q ;
-    \end{aligned}
-  \end{equation}
-\end{romdefinition}
-The Kuratowski finite subset is a free join semilattice (or, equivalently, a
-free commutative idempotent monoid).
-More prosaically, \(\mathcal{K}\) is the abstract data type for finite sets, as
-defined in the Boom hierarchy \cite{boomFurtherThoughtsAbstracto1981,
-  bunkenburgBoomHierarchy1994}.
-However, rather than just being a specification, \(\mathcal{K}\) is fully usable
-as a data type in its own right, thanks to HITs.
-
-Other definitions of \(\mathcal{K}\) exist (such as the one in
-\cite{fruminFiniteSetsHomotopy2018}) which make the fact that \(\mathcal{K}\) is
-the free join semilattice more obvious.
-We have included such a definition in our formalisation, and proven it
-equivalent to the one above.
-
-Next, we need a way to say that an entire type is Kuratowski finite.
-For that, we will need to define membership of \(\mathcal{K}\).
-\begin{romdefinition}[Membership of \(\mathcal{K}\)]
-  Membership is defined by pattern-matching on \(\mathcal{K}\).
-  The two point constructors are handled like so:
-  \begin{equation}
-    \begin{alignedat}{2}
-      x \in&& \; []                      &= \bot \\
-      x \in&& \; y \dblcolon \mathit{ys} &= \lVert x \equiv y \uplus x \in \mathit{ys} \rVert
-    \end{alignedat}
-  \end{equation}
-  The \(\text{com}\) and \(\text{dup}\) constructors are handled by proving that
-  the truncated form of \(\uplus\) is itself commutative and idempotent.
-  The type of propositions is itself a set, satisfying the \(\text{trunc}\)
-  constructor.
-\end{romdefinition}
-Finally, we have enough background to define Kuratowski finiteness.
-\begin{romdefinition}[Kuratowski Finiteness]
-  \begin{equation}
-    \mathcal{K}^{f}(A) = \Sigma {(\mathit{xs} : \mathcal{K}(A))} , \Pi (x : A) , x \in \mathit{xs}
-  \end{equation}
-\end{romdefinition}
-
-We also have the following two lemmas, proven in both
-\cite{fruminFiniteSetsHomotopy2018} and our formalisation.
-\begin{romlemma}
-  \(\mathcal{K}^f\) is a mere proposition.
-\end{romlemma}
-\begin{romlemma}
-  This circle \(S^1\) is Kuratowski finite.
-\end{romlemma}
-The second of these in particular tells us that the ``Kuratowski-finite types''
-are not necessarily sets; it also tells us that we cannot derive decidable
-equality from a proof of Kuratowski finiteness.
 \subsection{Split Enumerable}
@@
   \begin{equation}
-    \mathcal{E}!(A) \equiv \Sigma (n : \mathbb{N}) , \left( \mathbf{Fin}(n) \twoheadrightarrow ! \; A \right)
+    \mathcal{E}!(A) \simeq \Sigma (n : \mathbb{N}) , \left( \mathbf{Fin}(n) \twoheadrightarrow ! \; A \right)
   \end{equation}
@@
     \mathcal{E}!(A)
-    & \equiv \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , x \in \mathit{xs}
+    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , x \in \mathit{xs}
     & \text{Definition of } \mathcal{E}!
     \\
-    & \equiv \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{fib}_{\text{snd}(\mathit{xs})}(x)
+    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{fib}_{\text{snd}(\mathit{xs})}(x)
     & \text{Definition of } \in
     \\
-    & \equiv \Sigma (\mathit{xs} : \textbf{List}(A)) , \text{sp-surj}(\text{snd}(\mathit{xs}))
+    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \text{sp-surj}(\text{snd}(\mathit{xs}))
     & \text{Definition of sp-surj}
     \\
-    & \equiv \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket (A)) , \text{sp-surj}(\text{snd}(\mathit{xs}))
+    & \simeq \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket (A)) , \text{sp-surj}(\text{snd}(\mathit{xs}))
     & \text{Definition of } \mathbf{List}
     \\
-    & \equiv \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \mathbf{Fin}(n)) , A) , \text{sp-surj}(\text{snd}(\mathit{xs}))
+    & \simeq \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \mathbf{Fin}(n)) , A) , \text{sp-surj}(\text{snd}(\mathit{xs}))
     & \text{Definition of } \llbracket \wc \rrbracket
     \\
-    & \equiv \Sigma (n : \mathbb{N}) , \Sigma (f : \mathbf{Fin}(n) \rightarrow A) , \text{sp-surj}(f)
+    & \simeq \Sigma (n : \mathbb{N}) , \Sigma (f : \mathbf{Fin}(n) \rightarrow A) , \text{sp-surj}(f)
     & \text{Reassociation of } \Sigma
     \\
-    & \equiv \Sigma (n : \mathbb{N}) , ( \mathbf{Fin}(n) \twoheadrightarrow ! \; A )
+    & \simeq \Sigma (n : \mathbb{N}) , ( \mathbf{Fin}(n) \twoheadrightarrow ! \; A )
     & \text{Definition of } \twoheadrightarrow!
@@
     \mathcal{B}(A)
-    & \equiv \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , x \inunique \mathit{xs}
+    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , x \inunique \mathit{xs}
     & \text{Definition of } \mathcal{B}
     \\
-    & \equiv \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{isContr} \left(\text{fib}_{\text{snd}(\mathit{xs})}(x)\right)
+    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{isContr} \left(\text{fib}_{\text{snd}(\mathit{xs})}(x)\right)
     & \text{Definition of } \inunique
     \\
-    & \equiv \Sigma (\mathit{xs} : \textbf{List}(A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
+    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
     & \text{\cite[theorem 4.4.5]{hottbook}}
     \\
-    & \equiv \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket (A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
+    & \simeq \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket (A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
     & \text{Definition of } \mathbf{List}
     \\
-    & \equiv \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \mathbf{Fin}(n)) , A) , \text{isEquiv}(\text{snd}(\mathit{xs}))
+    & \simeq \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \mathbf{Fin}(n)) , A) , \text{isEquiv}(\text{snd}(\mathit{xs}))
     & \text{Definition of } \llbracket \wc \rrbracket
     \\
-    & \equiv \Sigma (n : \mathbb{N}) , \Sigma (f : \mathbf{Fin}(n) \rightarrow A) , \text{isEquiv}(f)
+    & \simeq \Sigma (n : \mathbb{N}) , \Sigma (f : \mathbf{Fin}(n) \rightarrow A) , \text{isEquiv}(f)
     & \text{Reassociation of } \Sigma
     \\
-    & \equiv \Sigma (n : \mathbb{N}) , ( \mathbf{Fin}(n) \simeq \; A )
+    & \simeq \Sigma (n : \mathbb{N}) , ( \mathbf{Fin}(n) \simeq \; A )
     & \text{Definition of } \twoheadrightarrow!
@@
 \end{proof}
+\subsection{Cardinal}
+\begin{itemize}
+\item Definition
+\item Cardinality
+\end{itemize}
 \subsection{Manifest Enumerable}
@@
 Again, the only difference with this type and split enumerability is the
-membership proof.
-Here we have propositionally truncated it.
-
-% We will use the circle as an example non-set type which is nonetheless
-% manifestly enumerable.
-% \begin{romtheorem}
-%   The circle \(S^1\) is manifestly enumerable.
-% \end{romtheorem}
-% \begin{proof}
-%   As the cover proof is a truncated proposition, we need only consider the point
-%   constructors, making this poof the same as the proof of split enumerability on
-%   \(\top\).
-% \end{proof}
-% \subsection{Surjections}
-% This predicates relation to surjectivity is much the same as split
-% enumerability's relation to \emph{split} surjectivity.
-% \begin{romlemma}
-%   A proof of manifest enumerability is equivalent to a surjection from a finite
-%   prefix of the natural numbers.
-% \end{romlemma}
-\subsection{Cardinal}
+membership proof: here we have propositionally truncated it.
+This has two effects.
+First, it means that this proof represents a true surjection (rather than a
+split surjection) from \(\mathbf{Fin}\).
+\begin{romlemma}
+  \begin{equation}
+    \mathcal{E}(A) \simeq \Sigma(n : \mathbb{N}) , (\mathbf{Fin}(n) \twoheadrightarrow A)
+  \end{equation}
+\end{romlemma}
+\begin{proof}
+  \todo[inline]{Proof}
+\end{proof}
+
+Secondly, it means the predicate does not imply decidable equality.
+More significantly, it allows the predicate to be defined over non-set types,
+like the circle.
+\begin{romlemma}
+  The circle \(S^1\) is manifestly enumerable.
+\end{romlemma}
+\begin{proof}
+  \todo[inline]{Proof}
+\end{proof}
+
+\subsection{Kuratowski Finiteness}
+Much work has already been done on Kuratowski finiteness in HoTT in
+\cite{fruminFiniteSetsHomotopy2018}.
+As a result, we will not needlessly repeat proofs, rather we will just give a
+brief introduction to the topic and point out where our treatment differs.
+
+The first thing we must define is a representation of subsets.
+\begin{romdefinition}[Kuratowski Finite Subset]
+  \(\mathcal{K}(A)\) is the type of Kuratowski-finite subsets of \(A\).
+  \begin{equation}
+    \begin{aligned}
+      \mathcal{K}(A) =&
+      \; [] : \mathcal{K}(A) ; \\
+      \vert& \; \wc \dblcolon \wc : A \rightarrow \mathcal{K}(A) \rightarrow \mathcal{K}(A) ; \\
+      \vert& \; \text{com} : \Pi (x, y: A) , \Pi (\mathit{xs} : \mathcal{K}(A)) , x \dblcolon y \dblcolon \mathit{xs} \simeq y \dblcolon x \dblcolon \mathit{xs} ; \\
+      \vert& \; \text{dup} : \Pi (x : A) , \Pi (\mathit{xs} : \mathcal{K}(A)) , x \dblcolon x \dblcolon \mathit{xs} \simeq x \dblcolon \mathit{xs} ; \\
+      \vert& \; \text{trunc} : \Pi (\mathit{xs}, \mathit{ys}: \mathcal{K}(A)) , \Pi (p, q : \mathit{xs} \simeq \mathit{ys}) , p \simeq q ;
+    \end{aligned}
+  \end{equation}
+\end{romdefinition}
+The Kuratowski finite subset is a free join semilattice (or, equivalently, a
+free commutative idempotent monoid).
+More prosaically, \(\mathcal{K}\) is the abstract data type for finite sets, as
+defined in the Boom hierarchy \cite{boomFurtherThoughtsAbstracto1981,
+  bunkenburgBoomHierarchy1994}.
+However, rather than just being a specification, \(\mathcal{K}\) is fully usable
+as a data type in its own right, thanks to HITs.
+
+Other definitions of \(\mathcal{K}\) exist (such as the one in
+\cite{fruminFiniteSetsHomotopy2018}) which make the fact that \(\mathcal{K}\) is
+the free join semilattice more obvious.
+We have included such a definition in our formalisation, and proven it
+equivalent to the one above.
+
+Next, we need a way to say that an entire type is Kuratowski finite.
+For that, we will need to define membership of \(\mathcal{K}\).
+\begin{romdefinition}[Membership of \(\mathcal{K}\)]
+  Membership is defined by pattern-matching on \(\mathcal{K}\).
+  The two point constructors are handled like so:
+  \begin{equation}
+    \begin{alignedat}{2}
+      x \in&& \; []                      &= \bot \\
+      x \in&& \; y \dblcolon \mathit{ys} &= \lVert x \simeq y \uplus x \in \mathit{ys} \rVert
+    \end{alignedat}
+  \end{equation}
+  The \(\text{com}\) and \(\text{dup}\) constructors are handled by proving that
+  the truncated form of \(\uplus\) is itself commutative and idempotent.
+  The type of propositions is itself a set, satisfying the \(\text{trunc}\)
+  constructor.
+\end{romdefinition}
+Finally, we have enough background to define Kuratowski finiteness.
+\begin{romdefinition}[Kuratowski Finiteness]
+  \begin{equation}
+    \mathcal{K}^{f}(A) = \Sigma {(\mathit{xs} : \mathcal{K}(A))} , \Pi (x : A) , x \in \mathit{xs}
+  \end{equation}
+\end{romdefinition}
+
+We also have the following two lemmas, proven in both
+\cite{fruminFiniteSetsHomotopy2018} and our formalisation.
+\begin{romlemma}
+  \(\mathcal{K}^f\) is a mere proposition.
+\end{romlemma}
+\begin{romlemma}
+  This circle \(S^1\) is Kuratowski finite.
+\end{romlemma}
+The second of these in particular tells us that the ``Kuratowski-finite types''
+are not necessarily sets; it also tells us that we cannot derive decidable
+equality from a proof of Kuratowski finiteness.
 \section{Relations Between Each Finiteness Definition}
+\begin{itemize}
+  \item SE -> MB (and back)
+  \item MB <-> ME
+  \item MB <-> C
+  \item C <-> K
+  \item ME -> K
+\end{itemize}
 \section{Topos}
 \begin{itemize}
-  \item Complete (pullbacks + terminal object)
+  \item \(\bot\)
+  \item \(\top\)
+  \item Bool
+  \item \(\Sigma\)
+  \item \(\Pi\)
+  \item Pre-categories to categories
+  \item \(\Pi\)-pretopos
 \end{itemize}
 \section{Countably Infinite Types}
+\begin{itemize}
+  \item Definition
+  \item \(\Sigma\)
+  \item Kleene star
+  \item Truncated
+  \item Topos
+\end{itemize}
 \section{Search}
+\begin{itemize}
+  \item Omniscience
+  \item Search
+  \item Curry / uncurry
+  \item Synth function
+\end{itemize}
 
@@
 %   \begin{equation}
-%     \mathit{xs} \leftrightsquigarrow \mathit{ys} \implies \text{sort}(\mathit{xs}) \equiv \text{sort}(\mathit{ys})
+%     \mathit{xs} \leftrightsquigarrow \mathit{ys} \implies \text{sort}(\mathit{xs}) \simeq \text{sort}(\mathit{ys})
 %   \end{equation}
@@
 %   \begin{equation}
-%     \mathit{xs}^\star_i = \left\{ \text{map}(\mathit{xs} , x) \mid \mathit{x} \in \mathbf{List}(\mathbb{N}) , \text{sum}(\text{map}(1+, x)) \equiv i \right\}
+%     \mathit{xs}^\star_i = \left\{ \text{map}(\mathit{xs} , x) \mid \mathit{x} \in \mathbf{List}(\mathbb{N}) , \text{sum}(\text{map}(1+, x)) \simeq i \right\}
 %   \end{equation}

### 2020-01-21 paper.tex
more

@@
 \subsection{Notation and Background}
+\subsubsection{Notation}
 We work in Cubical Type Theory \cite{cohenCubicalTypeTheory2016}.
@@
 \end{description}
+\subsubsection{Cubical Type Theory}
+Cubical Type Theory \cite{cohenCubicalTypeTheory2016} is a constructive type
+theory with an implementation in Cubical Agda
+\cite{vezzosiCubicalAgdaDependently2019}.
+It allows us to do much of the same theory as in HoTT, but crucially the
+univalence ``axiom'' is a \emph{theorem}, rather than an axiom.
+This allows us to actually compute with univalent proofs, a capability missing
+from HoTT.
+\begin{romdefinition}[Path Types]
+  The equality type (which we denote with \(\equiv\)) in CuTT is the type of
+  Paths\footnotemark.
+  The internal structure of paths is largely irrelevant to us here, as we will
+  generally treat \(\equiv\) as a black-box equivalence relation with
+  substitution and congruence.
+\end{romdefinition}
+
+\footnotetext{
+  Actually, CuTT does have an identity type with similar semantics to the
+  identity type in MLTT.
+  We do not use this type anywhere in our work, however, so we will not consider
+  it here.
+}
+\begin{romdefinition}[Homotopy Levels]
+  Types in HoTT and CuTT are not necessarily sets, as they are in MLTT.
+  This means that equalities are not necessarily unique.
+  In fact, we have an entire hierarchy of homotopy levels, of which sets are
+  level 2.
+  \begin{align}
+    \text{isContr}(A)    &\coloneqq \Sigma(x : A) , \Pi(y : A) , (x \equiv y) \\
+    \text{isProp}(A)     &\coloneqq \Pi(x, y : A) , (x \equiv y) \\
+    \text{isSet}(A)      &\coloneqq \Pi(x, y : A) , \text{isProp}(x \equiv y) \\
+    \text{isGroupoid}(A) &\coloneqq \Pi(x, y : A) , \text{isSet}(x \equiv y)
+  \end{align}
+  We can define the above types inductively like so:
+  \begin{align}
+    \text{isOfHLevel}(0 , A) &\coloneqq \text{isContr}(A) \\
+    \text{isOfHLevel}(n+1 , A) &\coloneqq \Pi(x , y : A) , \text{isOfHLevel}(n , x \equiv y)
+  \end{align}
+
+\end{romdefinition}
+
+\begin{romdefinition}[Isomorphism]
+  We say two types \(A\) and \(B\) are isomorphic, denoted with \(A \iff B\), if
+  there is a pair of functions \(f : A \rightarrow B\), \(g : B \rightarrow A\)
+  such that
+  \begin{equation}
+    \begin{aligned}
+      \Pi(x : A) , (g(f(x)) \equiv x) \\
+      \Pi(x : B) , (f(g(x)) \equiv x)
+    \end{aligned}
+  \end{equation}
+  While we can derive an equivalence from an isomorphism:
+  \begin{equation}
+    \text{isoToEquiv} : (A \iff B) \rightarrow (A \simeq B)
+  \end{equation}
+  It is not the case that an isomorphism is the \emph{same} as an equivalence in
+  all cases.
+  In other words, we do not have:
+  \begin{equation}
+    (A \iff B) \simeq (A \simeq B)
+  \end{equation}
+  However, if one of \(A\) or \(B\) is a set, then the above holds.
+\end{romdefinition}
 
@@
 
-\todo[inline]{Basic MLTT Notation: \(\Pi\), \(\Sigma\), etc}
-\todo[inline]{Paths, Univalence, isomorphisms}
 \todo[inline]{HITs}

### 2020-01-21 paper.tex
paper
definitions done

@@
     are equivalent).
-    The two constructors are:
     \begin{equation}
       \begin{aligned}
+        A \uplus B \coloneqq & \;
         \text{inl} &: A \rightarrow A \uplus B \\
-        \text{inr} &: B \rightarrow A \uplus B
+        | & \;  \text{inr} &: B \rightarrow A \uplus B
       \end{aligned}
@@
 from HoTT.
-\begin{romdefinition}[Path Types]
+\begin{romdefinition}[Path Types] \label{path-types}
   The equality type (which we denote with \(\equiv\)) in CuTT is the type of
@@
 }
-\begin{romdefinition}[Homotopy Levels]
+\begin{romdefinition}[Homotopy Levels] \label{homotopy-types}
   Types in HoTT and CuTT are not necessarily sets, as they are in MLTT.
@@
   \end{align}
-
-\end{romdefinition}
-
-\begin{romdefinition}[Isomorphism]
+\end{romdefinition}
+\begin{romdefinition}[Fibers] \label{fibers}
+  A fiber is defined over some function \(f : A \rightarrow B\).
+  \begin{equation}
+    \text{fib}_f(y) = \Sigma(x : A) , (f x \equiv y)
+  \end{equation}
+\end{romdefinition}
+\begin{romdefinition}[Equivalences] \label{equivalences}
+  We will take contractible maps \cite[definition 4.4.1]{hottbook} as our
+  ``default'' definition of equivalences.
+  \begin{equation}
+    A \simeq B \coloneqq \Sigma(f : A \rightarrow B) , \Pi(y : B) , \text{isContr}(\text{fib}_f(y))
+  \end{equation}
+\end{romdefinition}
+\begin{romlemma}
+  Univalence
+  \begin{equation}
+    (A \simeq B) \simeq (A \equiv B)
+  \end{equation}
+\end{romlemma}
+\begin{romdefinition}[Isomorphism] \label{isomorphism}
   We say two types \(A\) and \(B\) are isomorphic, denoted with \(A \iff B\), if
@@
 \end{romdefinition}
-
-% \item[Isomorphisms and Biconditionals] A pair of functions \(f : A \rightarrow
-%   B\) and \(g : B \rightarrow A\) will be described by the type \(A
-%   \leftrightarrow B\).
-%   If \(g\) is a quasi-inverse \cite[definition 2.4.6]{hottbook} of \(f\), we
-%   have an isomorphism between \(A\) and \(B\), denoted \(A \iff B\).
-%   Note that while we can derive an equivalence from an isomorphism, we do not
-%   have \((A \iff B) \simeq (A \simeq B)\).
-
-We will use the standard \(\Pi\) and \(\Sigma\) notation for dependent products
-and sums, respectively.
-
-\todo[inline]{HITs}
-\todo[inline]{Surjections}
-\todo[inline]{Containers}
-\todo[inline]{Fin}
-\todo[inline]{Biconditional, Isomorphism, Equivalence, Equals}
+\begin{romdefinition}[Higher Inductive Types] \label{HITs}
+  \todo[inline]{define}
+\end{romdefinition}
+\begin{romdefinition}[Surjections] \label{surjections}
+  \begin{alignat}{3}
+    &\text{surj}(f)             &&\coloneqq \Pi(y : B) , \lVert \text{fib}_f(y) \rVert \\
+    &A \twoheadrightarrow B     &&\coloneqq \Sigma (f : A \rightarrow B) , \text{surj}(f) \\
+    &\text{sp-surj}(f)          &&\coloneqq \Pi(y : B) , \text{fib}_f(y) \label{sp-surj-eqn} \\
+    &A \twoheadrightarrow! \; B &&\coloneqq \Sigma (f : A \rightarrow B) , \text{sp-surj}(f) \label{sp-surj-arrow-eqn}
+  \end{alignat}
+\end{romdefinition}
+\begin{romdefinition}[Containers] \label{container-def}
+  A container \cite{abbottContainersConstructingStrictly2005} is a pair
+  \(S , P\) where \(S\) is a type, the elements of which are called
+  the \emph{shapes} of the container, and \(P\) is a type family on \(S\), where
+  the elements of \(P(s)\) are called the \emph{positions} of a container.
+  We ``interpret'' a container into a functor defined like so:
+  \begin{equation} \label{container-interp}
+    \llbracket S , P \rrbracket(A) \coloneqq \Sigma {(s : S)} , \left( P(s) \rightarrow A \right)
+  \end{equation}
+  Membership of a container can be defined like so:
+  \begin{equation} \label{container-membership}
+    x \in \mathit{xs} \coloneqq \text{fib}_{\text{snd}(\mathit{xs})}(x)
+  \end{equation}
+\end{romdefinition}
+\begin{romdefinition}[\(\mathbf{List}\)] \label{List}
+  \begin{equation}
+    \mathbf{List} \coloneqq \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket
+  \end{equation}
+\end{romdefinition}
+\begin{romdefinition}[\(\mathbf{Fin}\)] \label{Fin}
+  \(\mathbf{Fin}(n)\) is the type of natural numbers smaller than \(n\).
+  We define it the standard way, where \(\mathbf{Fin}(0) = \bot\) and
+  \(\mathbf{Fin}(n + 1) &= \top + \mathbf{Fin}(n)\).
+\end{romdefinition}
+\begin{romdefinition}[Propositional Truncation] \label{prop-trunc}
+  The type \(\lVert A \rVert\) on some type \(A\) is a propositionally truncated
+  proof of \(A\) \cite[3.7]{hottbook}.
+  In other words, it is a proof that some \(A\) exists, but it does not tell you
+  \emph{which} \(A\).
+
+  It is defined as a Higher Inductive Type:
+  \begin{equation} {
+    \begin{alignat*}{3}
+      \lVert A \rVert \coloneqq & \; \lvert \wc \rvert &&: A \rightarrow \lVert A \rVert ; \\
+                              | & \; \text{squash}     &&: \Pi {(x, y : \lVert A \rVert)} , x \equiv y  ; 
+    \end{alignat*} }
+  \end{equation}
+  We will use three eliminators from \(\lVert A \rVert\) in this paper.
+  \begin{enumerate}
+  \item \label{elim-prop-prop} For any function \(A \rightarrow B\), where
+    \(\text{isProp}(B)\), we have a function \(\lVert A \rVert \rightarrow B\).
+  \item \label{elim-prop-monad} A special case of \ref{elim-prop-prop} implies
+    that \(\lVert \cdot \rVert\) forms a monad: this means that we get the usual
+    Monadic operators on \(\lVert \cdot \rVert\) (bind, pure, fmap, etc.).
+    \item \label{elim-prop-coh} We can eliminate from \(\lVert A \rVert\) with a
+      function \(f : A \rightarrow B\) iff \(f\) ``doesn't care'' about the
+      choice of \(A\) (\(\Pi {(x , y : A)} , f(x) \equiv f(y) \)).
+      Formally speaking, \(f\) needs to be ``coherently constant''
+      \cite{krausGeneralUniversalProperty2015}, and \(B\) needs to be an
+      \(n\)-type for some finite \(n\).
+  \end{enumerate}
+\end{romdefinition}
 \section{Finiteness Predicates}
@@
 \subsection{Split Enumerable}
-\begin{romdefinition}[Split Enumerable Set]
-  \begin{equation}
-    \mathcal{E}!(A) = \Sigma {(\mathit{xs} : \textbf{List}(A))} , \Pi {(x : A)} , x \in \mathit{xs}
+\begin{romdefinition}[Split Enumerable Set] \label{split-enum-def}
+  \begin{equation} \label{split-enum-def-eqn}
+    \mathcal{E}!(A) \coloneqq \Sigma {(\mathit{xs} : \textbf{List}(A))} , \Pi {(x : A)} , x \in \mathit{xs}
   \end{equation}
@@
 \end{romlemma}
-\todo[inline]{Put refs for each ``definition of''}
 \begin{proof}
@@
     & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , x \in \mathit{xs}
-    & \text{Definition of } \mathcal{E}!
+    && \text{Def.~\ref{split-enum-def} }(\mathcal{E}!)
     \\
     & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{fib}_{\text{snd}(\mathit{xs})}(x)
-    & \text{Definition of } \in
+    && \text{Eqn.~\ref{container-membership} } (\in)
     \\
     & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \text{sp-surj}(\text{snd}(\mathit{xs}))
-    & \text{Definition of sp-surj}
+    && \text{Eqn.~\ref{sp-surj-eqn} (sp-surj)}
     \\
     & \simeq \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket (A)) , \text{sp-surj}(\text{snd}(\mathit{xs}))
-    & \text{Definition of } \mathbf{List}
+    && \text{Def.~\ref{List} } (\mathbf{List})
     \\
     & \simeq \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \mathbf{Fin}(n)) , A) , \text{sp-surj}(\text{snd}(\mathit{xs}))
-    & \text{Definition of } \llbracket \wc \rrbracket
+    && \text{Eqn.~\ref{container-interp} } (\llbracket \wc \rrbracket)
     \\
     & \simeq \Sigma (n : \mathbb{N}) , \Sigma (f : \mathbf{Fin}(n) \rightarrow A) , \text{sp-surj}(f)
-    & \text{Reassociation of } \Sigma
+    && \text{Reassociation of } \Sigma
     \\
     & \simeq \Sigma (n : \mathbb{N}) , ( \mathbf{Fin}(n) \twoheadrightarrow ! \; A )
-    & \text{Definition of } \twoheadrightarrow!
+    && \text{Eqn.~\ref{sp-surj-arrow-eqn} } (\twoheadrightarrow!)
   \end{align*}
@@
   \(\mathbf{Fin}(n)\) for some \(n\).
-  Also, by lemma \todo[noline]{Fin is discrete}, we know that
-  \(\mathbf{Fin}(n)\) is discrete.
+  Also, we know that \(\mathbf{Fin}(n)\) is discrete (proven in our
+  formalisation).
   Therefore, by lemma~\ref{discrete-surj}, \(A\) is discrete.
@@
 % work.
-% \begin{romdefinition}[Container] \label{container-def}
-%   A container \cite{abbottContainersConstructingStrictly2005} is a pair
-%   \(S \triangleright P\) where \(S\) is a type, the elements of which are called
-%   the \emph{shapes} of the container, and \(P\) is a type family on \(S\), where
-%   the elements of \(P(s)\) are called the \emph{positions} of a container.
-%   We ``interpret'' a container into a functor defined like so:
-%   \begin{equation}
-%     \llbracket S \triangleright P \rrbracket(A) = \Sigma {(s : S)} , \left( P(s) \rightarrow A \right)
-%   \end{equation}
-%   Membership of a container can be defined like so:
-%   \begin{equation}
-%     x \in \mathit{xs} = \text{fiber}(\text{snd}(\mathit{xs}), x)
-%   \end{equation}
-%   Where \(\text{fiber}\) is from \cite[definition 4.2.4]{hottbook}.
-% \end{romdefinition}
-% \begin{romdefinition}[Lists]
-%   \begin{equation}%
-%     \mathbf{List} = \llbracket \mathbb{N} \triangleright \mathbf{Fin} \rrbracket
-%   \end{equation}
-% \end{romdefinition}
-% \begin{romdefinition}[\mathbf{Fin}]
-%   \(\mathbf{Fin}(n)\) is the type of natural numbers smaller than \(n\).
-%   We define it the standard way, where \(\mathbf{Fin}(0) = \bot\) and
-%   \(\mathbf{Fin}(n + 1) &= \top + \mathbf{Fin}(n)\).
-% \end{romdefinition}
 
@@
 % \subsection{Split Surjections}
-% \begin{romdefinition}[Surjections]
-%   \begin{multicols}{2}
-%     \begin{equation}
-%       A \twoheadrightarrow B = \Sigma (f : A \rightarrow B) , \text{surjective}(f)
-%     \end{equation} \break
-%     \begin{equation}
-%       A \twoheadrightarrow! \; B = \Sigma (f : A \rightarrow B) , \text{split surjective}(f)
-%     \end{equation}
-%   \end{multicols}
-%   Where the definitions of surjective are taken from \cite[definition
-%   4.6.1]{hottbook}.
-% \end{romdefinition}
 % \begin{romtheorem} \label{split-enum-surj}
@@
 %   \begin{equation}
-%     \mathbf{Stream}(A) = \llbracket \top \triangleright \text{const} (\mathbb{N}) \rrbracket(A) \simeq \mathbb{N} \rightarrow A
+%     \mathbf{Stream}(A) = \llbracket \top , \text{const} (\mathbb{N}) \rrbracket(A) \simeq \mathbb{N} \rightarrow A
 %   \end{equation}

### 2020-01-21 paper.tex
equations referenced

@@
   level 2.
-  \begin{align}
-    \text{isContr}(A)    &\coloneqq \Sigma(x : A) , \Pi(y : A) , (x \equiv y) \\
-    \text{isProp}(A)     &\coloneqq \Pi(x, y : A) , (x \equiv y) \\
-    \text{isSet}(A)      &\coloneqq \Pi(x, y : A) , \text{isProp}(x \equiv y) \\
-    \text{isGroupoid}(A) &\coloneqq \Pi(x, y : A) , \text{isSet}(x \equiv y)
-  \end{align}
+  \begin{alignat}{2}
+    &\text{isContr}(A)    &&\coloneqq \Sigma(x : A) , \Pi(y : A) , (x \equiv y) \\
+    &\text{isProp}(A)     &&\coloneqq \Pi(x, y : A) , (x \equiv y) \\
+    &\text{isSet}(A)      &&\coloneqq \Pi(x, y : A) , \text{isProp}(x \equiv y) \\
+    &\text{isGroupoid}(A) &&\coloneqq \Pi(x, y : A) , \text{isSet}(x \equiv y)
+  \end{alignat}
   We can define the above types inductively like so:
-  \begin{align}
-    \text{isOfHLevel}(0 , A) &\coloneqq \text{isContr}(A) \\
-    \text{isOfHLevel}(n+1 , A) &\coloneqq \Pi(x , y : A) , \text{isOfHLevel}(n , x \equiv y)
-  \end{align}
+  \begin{alignat}{2}
+    &\text{isOfHLevel}(0 , A)   &&\coloneqq \text{isContr}(A) \\
+    &\text{isOfHLevel}(n+1 , A) &&\coloneqq \Pi(x , y : A) , \text{isOfHLevel}(n , x \equiv y)
+  \end{alignat}
 \end{romdefinition}
@@
   ``default'' definition of equivalences.
-  \begin{equation}
-    A \simeq B \coloneqq \Sigma(f : A \rightarrow B) , \Pi(y : B) , \text{isContr}(\text{fib}_f(y))
-  \end{equation}
+  \begin{alignat}{2}
+    &\text{isEquiv}(f) &&\coloneqq \Pi(y : B) , \text{isContr}(\text{fib}_f(y)) \label{is-equiv-def} \\
+    &A \simeq B        &&\coloneqq \Sigma(f : A \rightarrow B) , \text{isEquiv}(f)
+  \end{alignat}
 \end{romdefinition}
@@
     &\text{surj}(f)             &&\coloneqq \Pi(y : B) , \lVert \text{fib}_f(y) \rVert \\
-    &A \twoheadrightarrow B     &&\coloneqq \Sigma (f : A \rightarrow B) , \text{surj}(f) \\
+    &A \twoheadrightarrow B     &&\coloneqq \Sigma (f : A \rightarrow B) , \text{surj}(f) \label{surj-arrow-eqn} \\
     &\text{sp-surj}(f)          &&\coloneqq \Pi(y : B) , \text{fib}_f(y) \label{sp-surj-eqn} \\
@@
 transfer those proofs onto Kuratowski finiteness.
-\subsection{Split Enumerable}
+\subsection{Split Enumerability}
 \begin{romdefinition}[Split Enumerable Set] \label{split-enum-def}
@@
 \end{proof}
-\subsection{Manifest Bishop}
+\subsection{Manifest Bishop Finiteness}
 \begin{romdefinition}[Manifest Bishop Finiteness]
-  \begin{equation}
-    \mathcal{B}(A) = \Sigma {(\mathit{xs} : \textbf{List}(A))} , \Pi {(x : A)} , x \inunique xs
+  \begin{equation} \label{bish-def}
+    \mathcal{B}(A) \coloneqq \Sigma {(\mathit{xs} : \textbf{List}(A))} , \Pi {(x : A)} , x \inunique xs
   \end{equation}
@@
 rather than simple membership (\(\in\)).
+\begin{romdefinition}[Unique Membership] \label{uniq-memb-def}
+  \begin{equation}
+    x \inunique \mathit{xs} \coloneqq \text{isContr}(x \in \mathit{xs})
+  \end{equation}
+\end{romdefinition}
 
@@
     & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , x \inunique \mathit{xs}
-    & \text{Definition of } \mathcal{B}
+    && \text{Def.~\ref{bish-def} } (\mathcal{B})
     \\
     & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{isContr} \left(\text{fib}_{\text{snd}(\mathit{xs})}(x)\right)
-    & \text{Definition of } \inunique
+    && \text{Def.~\ref{uniq-memb-def} } (\inunique)
     \\
     & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
-    & \text{\cite[theorem 4.4.5]{hottbook}}
+    && \text{Eqn.~\ref{is-equiv-def} (isEquiv)}
     \\
     & \simeq \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket (A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
-    & \text{Definition of } \mathbf{List}
+    && \text{Def.~\ref{List} } (\mathbf{List})
     \\
     & \simeq \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \mathbf{Fin}(n)) , A) , \text{isEquiv}(\text{snd}(\mathit{xs}))
-    & \text{Definition of } \llbracket \wc \rrbracket
+    && \text{Eqn.~\ref{container-interp} } (\llbracket \wc \rrbracket)
     \\
     & \simeq \Sigma (n : \mathbb{N}) , \Sigma (f : \mathbf{Fin}(n) \rightarrow A) , \text{isEquiv}(f)
-    & \text{Reassociation of } \Sigma
+    && \text{Reassociation of } \Sigma
     \\
     & \simeq \Sigma (n : \mathbb{N}) , ( \mathbf{Fin}(n) \simeq \; A )
-    & \text{Definition of } \twoheadrightarrow!
+    && \text{Eqn.~\ref{surj-arrow-eqn} } (\twoheadrightarrow)
   \end{align*}

### 2020-01-21 paper.tex
cleaned etc

@@
 \end{romdefinition}
+\begin{romdefinition}[\(S^1\)] \label{circle-def}
+  The circle, \(S^1\), can be represented in HoTT as a higher inductive type.
+  \begin{equation}
+    \begin{alignat}{5}
+      S^1 \coloneqq & \; \text{base} &&: S^1 ; \\
+                  | & \; \text{loop} &&: \text{base} \equiv \text{base} ; 
+    \end{alignat}
+  \end{equation}
+  We will use it here as an example of a non-set type, i.e. a type for which not
+  all paths are equal.
+  This also means that it does not have decidable equality.
+\end{romdefinition}
 \section{Finiteness Predicates}
@@
     & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , x \in \mathit{xs}
-    && \text{Def.~\ref{split-enum-def} }(\mathcal{E}!)
+    && \text{def.~\ref{split-enum-def} }(\mathcal{E}!)
     \\
     & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{fib}_{\text{snd}(\mathit{xs})}(x)
-    && \text{Eqn.~\ref{container-membership} } (\in)
+    && \text{eqn.~\ref{container-membership} } (\in)
     \\
     & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \text{sp-surj}(\text{snd}(\mathit{xs}))
-    && \text{Eqn.~\ref{sp-surj-eqn} (sp-surj)}
+    && \text{eqn.~\ref{sp-surj-eqn} (sp-surj)}
     \\
     & \simeq \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket (A)) , \text{sp-surj}(\text{snd}(\mathit{xs}))
-    && \text{Def.~\ref{List} } (\mathbf{List})
+    && \text{def.~\ref{List} } (\mathbf{List})
     \\
     & \simeq \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \mathbf{Fin}(n)) , A) , \text{sp-surj}(\text{snd}(\mathit{xs}))
-    && \text{Eqn.~\ref{container-interp} } (\llbracket \wc \rrbracket)
+    && \text{eqn.~\ref{container-interp} } (\llbracket \wc \rrbracket)
     \\
@@
     & \simeq \Sigma (n : \mathbb{N}) , ( \mathbf{Fin}(n) \twoheadrightarrow ! \; A )
-    && \text{Eqn.~\ref{sp-surj-arrow-eqn} } (\twoheadrightarrow!)
+    && \text{eqn.~\ref{sp-surj-arrow-eqn} } (\twoheadrightarrow!)
   \end{align*}
@@
     & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , x \inunique \mathit{xs}
-    && \text{Def.~\ref{bish-def} } (\mathcal{B})
+    && \text{def.~\ref{bish-def} } (\mathcal{B})
     \\
     & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{isContr} \left(\text{fib}_{\text{snd}(\mathit{xs})}(x)\right)
-    && \text{Def.~\ref{uniq-memb-def} } (\inunique)
+    && \text{def.~\ref{uniq-memb-def} } (\inunique)
     \\
     & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
-    && \text{Eqn.~\ref{is-equiv-def} (isEquiv)}
+    && \text{eqn.~\ref{is-equiv-def} (isEquiv)}
     \\
     & \simeq \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket (A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
-    && \text{Def.~\ref{List} } (\mathbf{List})
+    && \text{def.~\ref{List} } (\mathbf{List})
     \\
     & \simeq \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \mathbf{Fin}(n)) , A) , \text{isEquiv}(\text{snd}(\mathit{xs}))
-    && \text{Eqn.~\ref{container-interp} } (\llbracket \wc \rrbracket)
+    && \text{eqn.~\ref{container-interp} } (\llbracket \wc \rrbracket)
     \\
@@
     & \simeq \Sigma (n : \mathbb{N}) , ( \mathbf{Fin}(n) \simeq \; A )
-    && \text{Eqn.~\ref{surj-arrow-eqn} } (\twoheadrightarrow)
+    && \text{eqn.~\ref{surj-arrow-eqn} } (\twoheadrightarrow)
   \end{align*}
@@
 \subsection{Cardinal}
-\begin{itemize}
-\item Definition
-\item Cardinality
-\end{itemize}
-\subsection{Manifest Enumerable}
+Each finiteness predicate so far has contained an \emph{ordering} of the
+underlying type.
+For our purposes, this is too much information: it means that when constructing
+the ``category of finite sets'' later on, instead of each type having one
+canonical representative, it will have \(n!\), where \(n\) is the cardinality of
+the type.
+
+To remedy the problem, we will use propositional truncation
+(def.~\ref{prop-trunc}).
+\begin{romdefinition}[Cardinal Finiteness]
+  \begin{equation}
+    \mathcal{C}(A) \coloneqq \lVert \mathcal{B}(A) \rVert
+  \end{equation}
+\end{romdefinition}
+At first glance, it might seem that we lose any useful properties we could
+derive from \(\mathcal{B}\).
+Luckily, this is not the case: by eliminator \ref{elim-prop-coh} of
+def.~\ref{prop-trunc}, we need only show that the output is uniquely determined.
+
+The following two lemmas are proven in
+\cite{yorgeyCombinatorialSpeciesLabelled2014} (Proposition 2.4.9 and 2.4.10,
+respectively), in much the same way as we have done here.
+Our contribution for this section is simply the formalisation.
+\begin{romlemma}
+  Given a cardinally finite type, we can derive the type's cardinality, as well
+  as a propositionally truncated proof of equivalence with \(\textbf{Fin}\)s of
+  the same cardinality.
+  \begin{equation}
+    \mathcal{C}(A) \rightarrow \Sigma {(n : \mathbb{N})} , \lVert \textbf{Fin}(n) \simeq A \rVert
+  \end{equation}
+\end{romlemma}
+\begin{romlemma} \label{cardinal-finite-discrete}
+  Any cardinal-finite set has decidable equality.
+\end{romlemma}
+\subsection{Manifest Enumerability}
 \begin{romdefinition}[Manifest Enumerability]
   \begin{equation}
-    \mathcal{E}(A) = \Sigma {(\mathit{xs} : \mathbf{List}(A))} , \Pi {(x : A)} , \lVert x \in \mathit{xs} \rVert
+    \mathcal{E}(A) \coloneqq \Sigma {(\mathit{xs} : \mathbf{List}(A))} , \Pi {(x : A)} , \lVert x \in \mathit{xs} \rVert
   \end{equation}
 \end{romdefinition}
-Again, the only difference with this type and split enumerability is the
-membership proof: here we have propositionally truncated it.
+As with manifest Bishop finiteness, the only difference with this type and split
+enumerability is the membership proof: here we have propositionally truncated
+it.
 This has two effects.
@@
 \end{romlemma}
-\begin{proof}
-  \todo[inline]{Proof}
-\end{proof}
 
@@
 More significantly, it allows the predicate to be defined over non-set types,
-like the circle.
+like the circle (def.~\ref{circle-def}).
 \begin{romlemma}
@@
 \end{romlemma}
-\begin{proof}
-  \todo[inline]{Proof}
-\end{proof}
-
 \subsection{Kuratowski Finiteness}
@@
   \begin{equation}
-    \begin{aligned}
-      \mathcal{K}(A) =&
-      \; [] : \mathcal{K}(A) ; \\
-      \vert& \; \wc \dblcolon \wc : A \rightarrow \mathcal{K}(A) \rightarrow \mathcal{K}(A) ; \\
-      \vert& \; \text{com} : \Pi (x, y: A) , \Pi (\mathit{xs} : \mathcal{K}(A)) , x \dblcolon y \dblcolon \mathit{xs} \simeq y \dblcolon x \dblcolon \mathit{xs} ; \\
-      \vert& \; \text{dup} : \Pi (x : A) , \Pi (\mathit{xs} : \mathcal{K}(A)) , x \dblcolon x \dblcolon \mathit{xs} \simeq x \dblcolon \mathit{xs} ; \\
-      \vert& \; \text{trunc} : \Pi (\mathit{xs}, \mathit{ys}: \mathcal{K}(A)) , \Pi (p, q : \mathit{xs} \simeq \mathit{ys}) , p \simeq q ;
-    \end{aligned}
+    \begin{alignat}{2}
+      \mathcal{K}(A) \coloneqq&
+             \; []                &&: \mathcal{K}(A) ; \\
+      \vert& \; \wc \dblcolon \wc &&: A \rightarrow \mathcal{K}(A) \rightarrow \mathcal{K}(A) ; \\
+      \vert& \; \text{com}        &&: \Pi (x, y: A) , \Pi (\mathit{xs} : \mathcal{K}(A)) , x \dblcolon y \dblcolon \mathit{xs} \equiv y \dblcolon x \dblcolon \mathit{xs} ; \\
+      \vert& \; \text{dup}        &&: \Pi (x : A) , \Pi (\mathit{xs} : \mathcal{K}(A)) , x \dblcolon x \dblcolon \mathit{xs} \equiv x \dblcolon \mathit{xs} ; \\
+      \vert& \; \text{trunc}      &&: \Pi (\mathit{xs}, \mathit{ys}: \mathcal{K}(A)) , \Pi (p, q : \mathit{xs} \equiv \mathit{ys}) , p \equiv q ;
+    \end{alignat}
   \end{equation}
+
+  The first two constructors are point constructors, giving ways to create
+  values of type \(\mathcal{K}(A)\).
+  They are also recognisable as the two constructors for finite lists, a type
+  which represents the free monoid.
+
+  The next two constructors add extra paths to the type: equations that usage of
+  the type must obey.
+  These extra paths turn the free monoid into the free \emph{commutative} (com)
+  \emph{idempotent} (dup) monoid.
+
+  The final constructor enforces that the type \(\mathcal{K}(A)\) must be a set.
 \end{romdefinition}
@@
     \begin{alignedat}{2}
-      x \in&& \; []                      &= \bot \\
-      x \in&& \; y \dblcolon \mathit{ys} &= \lVert x \simeq y \uplus x \in \mathit{ys} \rVert
+      x \in&& \; []                      &\coloneqq \bot \\
+      x \in&& \; y \dblcolon \mathit{ys} &\coloneqq \lVert x \equiv y \uplus x \in \mathit{ys} \rVert
     \end{alignedat}
@@
 \section{Relations Between Each Finiteness Definition}
-\begin{itemize}
-  \item SE -> MB (and back)
-  \item MB <-> ME
-  \item MB <-> C
-  \item C <-> K
-  \item ME -> K
-\end{itemize}
+We now look at under what conditions a given finiteness predicate satisfies
+another.
+\subsection{Split Enumerability and Manifest Bishop Finiteness}
+\subsection{Manifest Bishop Finiteness and Manifest Enumerability}
+\subsection{Manifest Bishop Finiteness and Cardinal Finiteness}
+\subsection{Manifest Enumerability and Kuratowski Finiteness}
+\subsection{Cardinal Finiteness and Kuratowski Finiteness}
 \section{Topos}
@@
 
-% % Taking finiteness as an example again, if we prove that some type is finite via
-% % bijection, in MLTT such a proof would have to reveal the particular bijection.
-% % This becomes problematic when defining, for instance, the category of finite
-% % sets: the objects of this category would ideally be types coupled with proofs of
-% % their finiteness, but in order to
-
-% % The five in particular that we are interested in are organised in
-% % figure~\ref{finite-classification}.
-% % We especially want to examine what separates each predicate: what extra
-% % information do you need, for instance, to go from a proof of Cardinal finiteness
-% % (section~\ref{cardinal}) to manifest Bishop finiteness
-% % (section~\ref{manifest-bishop}).
-
-% % Next, we will examine the category-theoretic and topos-theoretic interpretations
-% % of each of the predicates.
-% % We will give a ``category of finite sets'' for each finiteness predicate, and
-% % then prove the various closure properties required to arrive at the relevant
-% % topos (we show that decidable Kuratowski finite sets form a \(\Pi\)-pretopos,
-% % for instance).
-
-% % ways to prove finiteness end up proving the same
-% % thing: constructively, however, each defines a fundamentally different
-% % ``version'' of finiteness.
-% % In Martin-Löf type theory \cite{martin-lofIntuitionisticTypeTheory1980}, these
-% % variations (and more) have been explored in
-% % \cite{spiwackConstructivelyFinite2010} and
-% % \cite{firsovVariationsNoetherianness2016}.
-
-% % Homotopy Type Theory \cite{hottbook} further adds to this story in two ways.
-% % First, its universe of types includes non-sets: whether or not a finiteness
-% % predicate includes these gives us another dimension on which to discriminate.
-
-% % The second addition is concerned with the hiding of information.
-% % Occasionally in a constructive setting you don't want to have to reveal how you
-% % proved something.
-% % To model logical disjunction, for instance, we usually use a disjoint union: the
-% % model isn't entirely correct, though, because a proof of \(A \text{or} B\)
-% % reveals which of \(A\) or \(B\) we actually proved.
-
-% % logical disjunction, for instance, is usually modelled with a
-% % disjoint union.
-% % This always means, however, that to prove ``A or B'' you have to reveal which of
-% % A or B you did prove.
-
-% % The other headline feature of HoTT's generalised equality is univalence: put
-% % simply, it allows us to treat isomorphic types as equal, giving us in the
-% % constructive world access to a technique that is commonplace classically.
-% % It turns out that finite types have a lot to say about isomorphisms, and we will
-% % rely heavily on this aspect of HoTT to work with them.
-% % \subsection{Contributions}
-% % In this work, we will explore finite types in Cubical Type Theory
-% % \cite{cohenCubicalTypeTheory2016}, and expose their relationship to infinite
-% % types, species, and demonstrate their practical uses for proofs in dependently
-% % typed programming languages.
-
-% % In section~\ref{strong-finiteness}, we will explore the ``strong'' notions of
-% % finiteness (i.e. those at least as strong as Kuratowski finiteness
-% % \cite{kuratowskiNotionEnsembleFini1920}), with a special focus on cardinal
-% % finiteness.
-% % We will delineate some subtly different finiteness predicates, and prove the
-% % relationship between each.
-% % We will prove, then, that every discrete Kuratowski finite type is cardinal
-% % finite, and subsequently will show that discrete Kuratowski finite sets form a
-% % topos.
-
-% % In section~\ref{species-and-containers}, we will redefine our finiteness
-% % predicates using \emph{containers}, and prove that the new definitions are
-% % equivalent to the old.
-% % Using this, we will prove much stronger relationships between our finiteness
-% % predicates and relation-based definitions of finiteness.
-% % Finally, We will show how the species-container duality falls out naturally from
-% % these new definitions.
-
-% % In section~\ref{infinite-cardinalities}, we will extend our study of finite
-% % types to infinite but countable types.
-% % We will see that the finiteness predicates are mirrored with countable
-% % counterparts, and we will prove closure under the Kleene star and plus.
-
-% % Proofs of finiteness have well-known practical applications in
-% % constructive mathematics \cite{firsovDependentlyTypedProgramming2015}.
-% % In section~\ref{practical}, we build a library which exploits these uses in
-% % Cubical Agda \cite{vezzosiCubicalAgdaDependently2019}, allowing automation of
-% % complex proofs over finite types.
-% % We frame this in terms of the principle of omniscience for finite types.
-% % Thanks to the flexibility afforded to us by Cubical Type Theory, we are able to
-% % go further than the usual examples of this kind of proof automation: as well as
-% % proving properties about functions, we can synthesise functions whole-cloth from
-% % their desired properties.
-% % Through the unified interface for finite and countable types, we can
-% % reuse the automation machinery for \emph{partial} proof search over infinite
-% % search spaces.
-% % Along the way, we extend the work in \cite{allaisGenericLevelPolymorphic2019} to
-% % prove isomorphisms between the curried and uncurried forms of \(n\)-ary
-% % dependent functions.
-
-% % Classically, of course, each of these statements are equivalent.
-% % Constructively, though, all four predicates are fundamentally different.
-
-% % Part of the difference comes from ``leaking'' information about the finite type
-% % in the proof of its finiteness.
-% % If I demonstrate a type is finite by showing a surjection to that type, I've
-% % revealed a way to construct all of the elements of my type.
-% % If I show that it is bounded, I reveal an upper bound on the cardinality of the
-% % type.
-
-% \subsection{Strong Finiteness Predicates}
-% We will first explore the ``strong'' notions of finiteness (i.e. those at least
-% as strong as Kuratowski finiteness \cite{kuratowskiNotionEnsembleFini1920}),
-% with a special focus on cardinal finiteness (section~\ref{cardinal}), and
-% manifest enumerability (section~\ref{manifest-enumerability}), which is new, to
-% our knowledge.
-
-% Figure~\ref{finite-classification} organises the predicates according to
-% their strength.
-% We will go through each of the predicates, proving how to weaken each (i.e. we
-% will provide a proof that every cardinally finite type is Kuratowski finite),
-% and how to strengthen them, given the required property.
-% In terms of figure~\ref{finite-classification}, this amounts to providing proofs
-% for each arrow.
-% Our main proof in this regard is that we can derive manifest Bishop finiteness
-% from Kuratowski finiteness plus a total order.
-
-% We will---through the use of containers
-% \cite{abbottContainersConstructingStrictly2005}---formally prove the equivalence
-% these predicates have with the usual function relations i.e. we will show that a
-% proof of manifest enumerability is precisely equivalent to a surjection from a
-% finite prefix of the natural numbers.
-
-% For each predicate, we will also prove its closure over sums and products in
-% both dependent and non-dependent forms, if such a closure exists.
-% This will culminate in our main result for this section: the formal proof that
-% decidable Kuratowski finite sets form a topos.
-% \subsection{Infinite Types}
-% In section~\ref{infinite-cardinalities}, we will extend our study of finite
-% types to infinite but countable types.
-% We will see that the classification of finiteness predicates mirrors that of
-% their countable counterparts, and we will prove closure under the Kleene star.
-
-% \subsection{Practical Uses}
-% Finally, in section~\ref{practical} we will demonstrate some practical uses of
-% the finiteness proofs in Cubical Agda.
-% We will show how to use well-known techniques
-% \cite{firsovDependentlyTypedProgramming2015, devrieseBrightSideType2011} to
-% automate proofs of functions with finite (and infinite) domains.
-% We further show how to automate the synthesis of functions from desired
-% properties.
-% \section{Split Enumerability} \label{split-enumerable}
-% We will start with the simplest definition of finiteness: we say a set is
-% enumerable if there is a list of its elements which contains every element in
-% the set.
-
-% Before giving the definition of the predicate, we will first define lists (and
-% membership thereof): we have chosen a \emph{container}-based definition for this
-% work.
-
-% Internally, in our formalisation, we actually use the standard
-% inductive definition of lists more often (it tends to work better in more
-% complex algorithms, and functions on it seems to satisfy the termination
-% checker more readily).
-% However, since both types are equivalent, univalence allows us to transport to
-% whichever representation is more convenient in a given situation.
-% For the higher-level proofs we present here, though, the container-based
-% definition greatly simplifies certain steps, which is why we have chosen it as
-% our representation.
-
-% Finally, we can define formally split enumerability.
-% \subsection{Split Surjections}
-% \begin{romtheorem} \label{split-enum-surj}
-%   Split enumerability is equivalent to a split surjection from a finite prefix
-%   of the natural numbers.
-%   \begin{equation}
-%     \mathcal{E}!(A) \iff \Sigma {(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \twoheadrightarrow ! \; A \right)
-%   \end{equation}
-% \end{romtheorem}
-% \begin{proof}
-%   After sufficient inlining, it emerges that our goal is simply a reassociation.
-%   \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{split-surj}
-%   To be clear: in Agda, the proof could simply be \(\AgdaFunction{reassoc}\); we
-%   have written out the extra lines for clarity alone.
-% \end{proof}
-% \begin{romlemma} \label{split-enum-discrete}
-%   Any split enumerable type has decidable equality (is discrete).
-% \end{romlemma}
-% \begin{proof}
-%   We use a corollary that if there is a split-surjection from \(A\) to \(B\),
-%   and \(A\) is discrete, then \(B\) is also discrete.
-% \end{proof}
-% By Hedberg's theorem \cite{hedbergCoherenceTheoremMartinLof1998}, we know that
-% this implies split enumerable types must all be sets.
-% \subsection{Closure}
-% In this section we will prove closure under various operations for split
-% enumerable sets.
-% We are working towards a topos proof, which requires us to prove closure under
-% a variety of operations: for now, we only have enough machinery to demonstrate
-% the semiring operations, and dependent sums.
-% In order to show closure under exponentials (function arrows), we will need an
-% equivalence with \(\mathbf{Fin}\), which will be provided in
-% section~\ref{manifest-bishop}.
-% \begin{romlemma} \label{split-enum-bool-top-bot}
-%   \(\mathbf{Bool}\), \(\top\), and \(\bot\) are all split enumerable.
-% \end{romlemma}
-% \begin{proof}
-%   Each of these types clearly has a finite number of elements (2, 1, and 0,
-%   respectively), and furthermore has a straightforward enumeration.
-% \end{proof}
-% \begin{romtheorem} \label{split-enum-sigma}
-%   Split-enumerability is closed under \(\Sigma\).
-%   \begin{equation} \label{split-enum-sigma-clos-sig}
-%     \frac{
-%       \mathcal{E}!(A) \; \; \; \Pi (x : A) , \mathcal{E}!(U(x))
-%     }{
-%       \mathcal{E}!(\Sigma (x : A) , U(x))
-%     }
-%   \end{equation}
-% \end{romtheorem}
-% \begin{proof}
-%   To obtain the support list, we concatenate the support lists of all the proofs
-%   of split-finiteness for \(U\) over the support list of \(E_A\).
-%   In Agda:
-%   \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{sigma-sup}
-%   ``do-notation'' is available to us as we're working in the list monad.
-% \end{proof}
-% Closure under disjoint union and Cartesian product both follow from \(\Sigma\),
-% as these types can be defined in terms of \(\Sigma\).
-% \begin{align}
-%   A \times B &= \Sigma {(x : A)} , B \\
-%   A + B &= \Sigma {(x : \mathbf{Bool})} , \text{if} \; x \; \text{then} \; A \; \text{else} \; B
-% \end{align}
-% \section{Manifest Bishop Finiteness} \label{manifest-bishop}
-% Where split enumerability was the enumeration form of a split surjection from
-% \(\mathbf{Fin}\), manifest Bishop finiteness is the enumeration form of an
-% \emph{equivalence} with \(\mathbf{Fin}\).
-
-% \begin{romdefinition}[Manifest Bishop Finiteness]
-%   \begin{equation}
-%     \mathcal{B}(A) = \Sigma {(\mathit{xs} : \textbf{List}(A))} , \Pi {(x : A)} , x \inunique xs
-%   \end{equation}
-% \end{romdefinition}
-% The only difference between this predicate and split enumerability is the list
-% membership term: we use \(\inunique\) here, where \(x \inunique \mathit{xs}\) is
-% to be read as ``\(x\) occurs exactly once in \(\mathit{xs}\)''.
-% \begin{romdefinition}[Unique Membership]
-%   We say an item \(x\) is ``uniquely in'' some container \(\mathit{xs}\) if its
-%   membership in that list is a \emph{contraction} \cite[definition
-%   3.11.1]{hottbook}; i.e. its membership proof exists, and all such proofs are
-%   equal.
-%   \begin{equation}
-%     x \inunique \mathit{xs} = \text{isContr}(x \in \mathit{xs})
-%   \end{equation}
-% \end{romdefinition}
-% By the definition of a contraction, we can always recover the underlying
-% membership proof, meaning that we can always derive split enumerability from
-% manifest Bishop finiteness.
-% \subsection{Equivalence}
-% \begin{romlemma} \label{bishop-equiv}
-%   A proof of manifest Bishop finiteness is equivalent to an equivalence with a
-%   finite prefix of the natural numbers.
-%   \begin{equation}
-%     \mathcal{B}(A) \simeq \Sigma {(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \simeq A \right)
-%   \end{equation}
-% \end{romlemma}
-% \begin{proof}
-%   There are many equivalent definitions of equivalence in HoTT.
-%   Here we take the version preferred in the Cubical Agda library: contractible
-%   maps \cite[definition 4.4.1]{hottbook}.
-%   Because of the parallels between contractible maps and split surjections,
-%   the proof proceeds much the same as \ref{split-enum-surj}.
-%   In other words, the definition of Bishop finiteness is itself a reassociation
-%   of a contractible map.
-% \end{proof}
-% \subsection{Relationship to Split Enumerability}
-% \begin{romtheorem} \label{split-enum-to-manifest-bishop}
-%   Any split enumerable set is manifest Bishop finite.
-% \end{romtheorem}
-% \begin{proof}
-%   From proposition~\ref{split-enum-discrete} we can derive decidable equality on
-%   \(A\), and using this we can define a function (called \(\text{uniques}\)) which filters
-%   out duplicates from lists of \(A\)s.
-%   This gives us our support list.
-%   To generate the cover proof it suffices now to prove the following:
-%   \begin{equation}
-%     \Pi {(x : A)} , \Pi {(\mathit{xs} : \textbf{List}(A))} , x \in \mathit{xs} \rightarrow x \inunique \text{uniques}(\mathit{xs})
-%   \end{equation}
-% \end{proof}
-% \subsection{Closure}
-% Proving equal strength of split enumerability and manifest Bishop finiteness
-% allows us to carry all of the previous proofs of closure over to manifest Bishop
-% finite sets (and vice-versa).
-% Missing from our previous proofs was a proof of closure of functions.
-% We remedy that here.
-% \begin{romtheorem}
-%   Manifest bishop finiteness is closed over dependent functions
-%   (\(\prod\)-types).
-%   \begin{equation}
-%     \frac{
-%       \mathcal{B}(A) \; \; \; \Pi {(x : A)} , \mathcal{B}\left( U(x) \right)
-%     }{
-%       \mathcal{B}\left(\Pi {(x : A)} , U(x)\right)
-%     }
-%   \end{equation}
-% \end{romtheorem}
-% \begin{proof}
-%   This proof is essentially the composition of two transport operations, made
-%   available to us via univalence.
-
-%   First, we will simplify things slightly by working only with split
-%   enumerability.
-%   As this is equal in strength to manifest Bishop finiteness, any closure proofs
-%   carry over.
-
-%   Secondly, we will replace \(A\) in all places with \(\mathbf{Fin}(n)\).
-%   Since we have already seen an equivalence between these two types, we are
-%   permitted to transport along these lines.
-%   This is the first transport operation.
-
-%   The bulk of the proof now is concerned with proving the following:
-%   \begin{equation}
-%     \left( \Pi {(x : \mathbf{Fin}(n))} , \mathcal{E}!(A(x)) \right) \rightarrow \mathcal{E}!\left( \Pi {(x : \mathbf{Fin}(n))} , A(x) \right)
-%   \end{equation}
-%   Our strategy to accomplish this will be to consider functions from
-%   \(\mathbf{Fin}(n)\) as \(n\)-tuples over some type family \(T : \mathbb{N}
-%   \rightarrow \text{Type}\).
-%   \begin{equation}
-%     \begin{aligned}
-%       \mathbf{Tuple}(T, 0)   &= \top \\
-%       \mathbf{Tuple}(T, n+1) &= T(0) \times \mathbf{Tuple}(T \circ \text{suc}, n)
-%     \end{aligned}
-%   \end{equation}
-%   This type is manifestly Bishop finite, as it is constructed only from products
-%   and the unit type.
-
-%   We then prove an isomorphism between this representation and \(\Pi\)-types.
-%   \begin{equation}
-%     \mathbf{Tuple}(T, n) \iff \Pi {(x : \mathbf{Fin}(n))} , T(x)
-%   \end{equation}
-%   This allows us to transport our proof of finiteness on tuples to one on
-%   functions from \(\mathbf{Fin}\) (our second transport operation), proving our
-%   goal.
-% \end{proof}
-% \section{Manifest Enumerability} \label{manifest-enumerability}
-% Both split enumerability and manifest Bishop finiteness are restricted to sets:
-% because they both imply decidable equality, no non-set types can satisfy them as
-% predicates.
-% To find a more general predicate for finiteness, we \emph{truncate} the
-% membership proof.
-% \subsection{Relation to Split Enumerability}
-% \begin{romtheorem} \label{manifest-enum-to-split}
-%   A manifestly enumerable type with decidable equality is split enumerable.
-% \end{romtheorem}
-% \begin{proof}
-%   The support list stays the same between both enumerability proofs.
-
-%   For the cover proof, we first need the following function which searches a
-%   list for a particular element (given decidable equality on \(A\)).
-%   \begin{equation}
-%     \decin \; : \Pi(x : A) , \Pi(\mathit{xs} : \mathbf{List}(A)) , \mathbf{Dec}(x \in \mathit{xs})
-%   \end{equation}
-%   Where \(\mathbf{Dec}(A)\) is a decision on some type \(A\).
-
-%   We then need to convert a value of type \(\mathbf{Dec}(x \in \mathit{xs})\) to
-%   \(x \in \mathit{xs}\).
-%   We use the following to do that:
-%   \begin{equation}
-%     \text{recompute} : \mathbf{Dec}(A) \rightarrow \lVert A \rVert \rightarrow A
-%   \end{equation}
-%   The function works via case-analysis on \(\mathbf{Dec}\): in the true case, we
-%   return the proof; in the false case, we can apply the proof of negation under
-%   the truncation, and then extract \(\lVert \bot \rVert\) to \(\bot\), as
-%   \(\bot\) is a mere proposition, thus proving the goal via explosion.
-% \end{proof}
-% \subsection{Closure}
-% Since we don't have an equivalence with \(\mathbf{Fin}\), we don't get closure
-% under \(\Pi\).
-% \begin{romlemma}
-%   Manifest enumerability is closed under \(\Sigma\).
-% \end{romlemma}
-% \begin{proof}
-%   This closure proof is almost the same as theorem~\ref{split-enum-sigma}.
-%   The manipulation of the support lists can be carried over as-is; but the type
-%   of the cover proof has changed, so it will need to be updated.
-%   As it happens, the translation is straightforward: we effectively write the
-%   ``monadic'' version of the old function.
-% \end{proof}
-% \section{Cardinal Finiteness} \label{cardinal}
-% For manifest enumerability, we removed the need for decidable equality: in these
-% next two finiteness predicates, we remove the need for a total order on the
-% underlying type.
-% \begin{romdefinition}[Cardinal Finiteness]
-%   A type \(A\) is cardinally finite, \(\mathcal{C}\), if it has a
-%   propositionally-truncated proof of bishop finiteness.
-%   \begin{equation}
-%     \mathcal{C}(A) = \lVert \mathcal{B}(A) \rVert
-%   \end{equation}%
-% \end{romdefinition}%
-% \subsection{Closure}
-% The closure properties of cardinal finiteness are effectively the non-dependent
-% versions of manifest Bishop finiteness.
-% To see why, consider equation~\ref{split-enum-sigma-clos-sig}.
-% We can ``lift'' the proof (as a binary function) under a propositional
-% truncation, giving us equation~\ref{sigma-proof-lifted}, but that doesn't give us
-% the desired closure proof (equation~\ref{sigma-proof-desired}).
-% \begin{multicols}{2}
-%   \noindent
-%   \begin{equation} \label{sigma-proof-lifted}
-%     \frac{
-%       \lVert \mathcal{B}(A) \rVert \; \; \; \lVert \Pi (x : A) , \mathcal{B}(U(x)) \rVert
-%     }{
-%       \lVert \mathcal{B}(\Sigma(x : A) , U(x)) \rVert
-%     }
-%   \end{equation}
-%   \begin{equation} \label{sigma-proof-desired}
-%     \frac{
-%       \lVert \mathcal{B}(A) \rVert \; \; \; \Pi (x : A) , \lVert \mathcal{B}(U(x))  \rVert
-%     }{
-%       \lVert \mathcal{B}(\Sigma(x : A) , U(x)) \rVert
-%     }
-%   \end{equation}
-% \end{multicols}%
-% To remedy the mismatch we would need a function of the type:
-% \begin{equation}
-%   \left( \Pi(x : A) , \lVert \mathcal{B}(U(x)) \rVert \right) \rightarrow
-%   \lVert \Pi(x : A) , \mathcal{B}(U(x)) \rVert
-% \end{equation}
-% Unfortunately, this equation in particular is a form of the axiom of choice
-% \cite[equation 3.8.3]{hottbook}.
-% This leaves us with closure under only the non-dependent operations.
-% \begin{romlemma}
-%   Cardinal finiteness is closed under \(\times\), \(+\), and \(\rightarrow\).
-% \end{romlemma}
-% \begin{proof}
-%   All of these closure proofs can be lifted directly from their corresponding
-%   proofs on manifest Bishop finiteness.
-% \end{proof}
-% \subsection{Strength}
-% The following two theorems are proven in
-% \cite{yorgeyCombinatorialSpeciesLabelled2014}, in much the same way as we have
-% done here.
-% Our contribution for this section is simply the formalisation.
-% \begin{romtheorem} \label{cardinal-finite-discrete}
-%   Any cardinal-finite set has decidable equality.
-% \end{romtheorem}
-% \begin{proof}
-%   See \cite[Proposition 2.4.10]{yorgeyCombinatorialSpeciesLabelled2014}.
-% \end{proof}
-% \begin{romtheorem}
-%   Given a cardinally finite type, we can derive the type's cardinality, as well
-%   as a propositionally truncated proof of equivalence with \(\textbf{Fin}\)s of
-%   the same cardinality.
-%   \begin{equation}
-%     \mathcal{C}(A) \rightarrow \Sigma {(n : \mathbb{N})} , \lVert \textbf{Fin}(n) \simeq A \rVert
-%   \end{equation}
-% \end{romtheorem}
-% \begin{proof}
-%   \cite[Proposition 2.4.9]{yorgeyCombinatorialSpeciesLabelled2014}.
-% \end{proof}
-% \subsection{Relation to Manifest Bishop Finiteness}
-% Cardinal finiteness tells us that there is an isomorphism between a type and
-% \(\mathbf{Fin}\); it just doesn't tell us \emph{which} isomorphism.
-% To take a simple example, \(\mathbf{Bool}\) has 2 possible isomorphisms with the
-% set \(\mathbf{Fin}(2)\): one where false maps to 0, and true to 1; and another
-% where false maps to 1 and true to 0.
-
-% To convert from Cardinal finiteness to Bishop finiteness, then, requires that we
-% supply enough information to identity a particular isomorphism.
-% A total order is sufficient here: it will give us enough to uniquely order the
-% support list invariant under permutations.
-% This tells us what we already knew in the introduction: manifest Bishop
-% finiteness is cardinal finiteness plus an order.
-
-% \begin{romtheorem}
-%   Any cardinal finite type with a (decidable) total order is manifestly Bishop
-%   finite.
-% \end{romtheorem}
-% \begin{proof}
-%   This proof is quite involved, and will rely on several subsequent lemmas, so
-%   we will give only its outline here.
-%   \begin{itemize}
-%     \item First, we will convert to manifest enumerability: knowing that the
-%       underlying type is discrete (theorem~\ref{cardinal-finite-discrete}) we can
-%       go from manifest enumerability to split enumerability
-%       (lemma~\ref{manifest-enum-to-split}), and subsequently to manifest Bishop
-%       finiteness (lemma~\ref{split-enum-to-manifest-bishop}).
-%     \item To convert to manifest enumerability, we need to provide a support
-%       list: this cannot simply be the support list hidden under the truncation,
-%       since that would violate the hiding promised by the truncation.
-%       Instead, we sort the list (using insertion sort).
-%       We must, therefore, prove that insertion sort is invariant under all
-%       support lists in cardinal finiteness proofs.
-%     \item We do this by first showing that all support lists in cardinal
-%       finiteness proofs are permutations of each other, and then that insertion
-%       sort is invariant under permutations.
-%     \item Given our particular definition of permutations, cover proofs
-%       transfers naturally between lists which are permutations of each other.
-%   \end{itemize} 
-% \end{proof}
-% Now we will build up the toolkit we need to perform the above steps.
-% First, permutations.
-% \begin{romdefinition}[List Permutations]
-%   We say that two lists are permutations of each other if there is an
-%   isomorphism between membership proofs
-%   \cite{danielssonBagEquivalenceProofRelevant2012}.
-%   \begin{equation}
-%     \mathit{xs} \leftrightsquigarrow \mathit{ys} = \Pi {(x : A)} , x \in \mathit{xs} \iff x \in \mathit{ys}
-%   \end{equation}
-% \end{romdefinition}
-% \begin{romlemma}
-%   Insertion sort is invariant under permutations.
-%   \begin{equation}
-%     \mathit{xs} \leftrightsquigarrow \mathit{ys} \implies \text{sort}(\mathit{xs}) \simeq \text{sort}(\mathit{ys})
-%   \end{equation}
-% \end{romlemma}
-% \begin{proof}
-%   First, we prove two properties about insertion sort:
-%   \begin{enumerate}
-%     \item It returns a sorted list.
-%     \item It a returns a list that is a permutation of its input.
-%   \end{enumerate}
-%   The second of these points allows us to show that \(\text{sort}(\mathit{xs})\)
-%   is a permutation of \(\text{sort}(\mathit{ys})\).
-%   \begin{equation}
-%     \text{sort}(\mathit{xs}) \leftrightsquigarrow \mathit{xs} \leftrightsquigarrow \mathit{ys} \leftrightsquigarrow \text{sort}(\mathit{ys})
-%   \end{equation}
-%   Then, we show that any lists which are both sorted and permutations of each
-%   other are equal.
-%   Both of these conditions are true for the output of \(\text{sort}\).
-% \end{proof}
-% \section{Kuratowski Finiteness} \label{kuratowski}
-% Finally we arrive at Kuratowski finiteness
-% \cite{kuratowskiNotionEnsembleFini1920}.
-
-% Other representations of \(\mathcal{K}\) \cite{fruminFiniteSetsHomotopy2018}
-% are more explicit constructions of the free join semilattice (i.e. there is a
-% point constructors for union instead of cons, and then path constructors for
-% the associativity and identity laws), but we have found this representation
-% easier to work with.
-% Nonetheless, the alternative representation is included in our formalisation,
-% and proven equivalent to the representation here.
-
-% \begin{romtheorem}
-%   A proof of Kuratowski finiteness is equivalent to a propositionally truncated
-%   proof of enumerability.
-%   \begin{equation}
-%     \mathcal{K}^f(A) \simeq \lVert \mathcal{E}(A) \rVert
-%   \end{equation}
-% \end{romtheorem}
-% \begin{proof}
-%   We prove by way of an isomorphism.
-%   In the first direction (from \(\mathcal{K}\) to \(\mathcal{E}\)), because we
-%   are eliminating into a proposition, we need only deal with the point
-%   constructors.
-%   For these, we convert the \(\mathcal{K}\) cons to its list counterpart, and
-%   similarly for the nil constructor.
-
-%   The other direction is proven in \cite{fruminFiniteSetsHomotopy2018}, so we
-%   will not describe it here.
-% \end{proof}
-% \subsection{Topos}
-% At this point, we see that a ``decidable Kuratowski finite set'' is precisely
-% equivalent to a cardinal finite set.
-% From this, we can lift over all of the properties of cardinal finite sets.
-% In particular, we see that decidable Kuratowski finite sets form a \emph{topos}.
-% We already have most of the components we need: closure under \(\bot\),
-% \(\top\), \(\mathbf{Bool}\), \(+\), \(\times\), and \(\rightarrow\).
-% What remains is the subobject classifier.
-% \todo[inline]{Topos!}
-% \section{Infinite Cardinalities} \label{infinite-cardinalities}
-% In the previous sections we saw different flavours of finiteness which were
-% really just different flavours of relations to \(\mathbf{Fin}\).
-% In this section we will see that we can construct a similar classification of
-% relations to \(\mathbb{N}\), in the form of the countably infinite types.
-% \subsection{Split Countable Types}
-% Our first foray into the world of countable types will be a straightforward
-% analogue to the split enumerable types.
-% We need change only one element: instead of a support \emph{list}, we instead
-% have a support \emph{stream}, which is its infinite.
-% \begin{romdefinition}[Stream]
-%   \begin{equation}
-%     \mathbf{Stream}(A) = \llbracket \top , \text{const} (\mathbb{N}) \rrbracket(A) \simeq \mathbb{N} \rightarrow A
-%   \end{equation}
-% \end{romdefinition}
-% \begin{romdefinition}[Split Countability]
-%   \begin{equation}
-%     \mathcal{E}!(A) = \Sigma {(\mathit{xs} : \mathbf{Stream}(A))} , \Pi {(x : A)} , x \in \mathit{xs}
-%   \end{equation}
-% \end{romdefinition}
-% This type is definitionally equal to it surjection equivalent (\(\mathbb{N}
-% \twoheadrightarrow ! \; A\)).
-% \subsubsection{Closure}
-% We know that countable infinity is not closed under the exponential (function
-% arrow), so the only closure we need to prove is \(\Sigma\) to cover all of
-% what's left.
-% \input{figures/pairing-functions.tex}
-% \begin{romtheorem} \label{split-countability-sigma}
-%   Split countability is closed under \(\Sigma\).
-% \end{romtheorem}
-% \begin{proof}
-%   The main task here is to figure out a pattern which will pair elements of the
-%   two support streams without diverging.
-%   Figure~\ref{cartesian} illustrates why the pairing function we used for lists
-%   won't work: it needs to exhaust one of the input lists before inspecting
-%   anything other than the first element of the other.
-%   Of course, we can't exhaust a stream, so this diverges.
-
-%   Instead, we will use the pattern in figure~\ref{cantor}.
-%   We will produce a stream of lists (where each list is a diagonal), which is
-%   then concatenated to the final output stream.
-%   The intermediate stream is sometimes called the discrete convolution of the
-%   two input streams.
-% \end{proof}
-% \subsubsection{Kleene Star}
-% While we lose some closures with the inclusion of infinite types, we gain some
-% others.
-% In particular, we have the Kleene star.
-% This means that we have closure under lists.
-% \begin{romtheorem}
-%   Split countability is closed under Kleene star.
-%   \begin{equation}
-%     \mathcal{E}!(A) \rightarrow \mathcal{E}!(\mathbf{List}(A))
-%   \end{equation}
-% \end{romtheorem}
-% \begin{proof}
-%   As with the proof of closure under \(\Sigma\), our main task here is to figure
-%   out a pattern that reaches any given finite list in finite time.
-%   Our first step is to split up the problem as we did for \(\Sigma\)-closure: we
-%   produce an intermediate stream of lists, which we then concatenate to give the
-%   final result.
-%   The intermediate stream is constructed like so:
-%   \begin{equation}
-%     \mathit{xs}^\star_i = \left\{ \text{map}(\mathit{xs} , x) \mid \mathit{x} \in \mathbf{List}(\mathbb{N}) , \text{sum}(\text{map}(1+, x)) \simeq i \right\}
-%   \end{equation}
-%   In other words, the \(i\)th element is all lists such that the sum of the
-%   successor of the indices in the list is equal to \(i\).
-% \end{proof}
-% \subsection{Other Types of Countability}
-% There are of course other forms of countability available to us: manifest
-% countability, cardinal, etc.
-% We have a similar structure of relationships to the finite types, with the
-% exception that we cannot construct an isomorphism with \(\mathbb{N}\) fiven a
-% surjection from \(\mathbb{N}\).
-% \section{Practical Uses} \label{practical}
-% \subsection{Omniscience}
-% In this section we are interested in restricted forms of the limited principle
-% of omniscience \cite{myhillErrettBishopFoundations1972}.
-% \begin{romdefinition}[Limited Principle of Omniscience]
-%   For any type \(A\) and predicate \(P\) on \(A\), the limited principle of
-%   omniscience is as follows:
-%   \begin{equation}
-%     \left( \Pi {(x : A)} , \mathbf{Dec}(P(x)) \right) \rightarrow \mathbf{Dec} \left( \Sigma {(x : A)} , P(x) \right)
-%   \end{equation}
-%   In other words, for any decidable predicate the existential quantification of
-%   that predicate is also decidable.
-% \end{romdefinition}
-% The limited principle of omniscience is non-constructive, but individual types
-% can themselves satisfy omniscience.
-% In particular, \emph{finite} types are omniscient.
-
-% There is also a universal form of omniscience, which we call exhaustibility.
-% \begin{romdefinition}[Exhaustibility]
-%   We say a type \(A\) is exhaustible if, for any decidable predicate \(P\) on
-%   \(A\), the universal quantification of the predicate is decidable.
-%   \begin{equation}
-%     \left( \Pi {(x : A)} , \mathbf{Dec}(P(x)) \right) \rightarrow \mathbf{Dec} \left( \Pi {(x : A)} , P(x) \right)
-%   \end{equation}
-% \end{romdefinition}
-
-% All of the finiteness predicates we have seen justify exhaustibility.
-% We will only prove it once, then, for the weakest:
-% \begin{romtheorem}
-%   Kuratowski-finite types are exhaustible.
-% \end{romtheorem}
-
-% Omniscience is stronger than exhaustibility, as we can derive the latter from
-% the former:
-% \begin{romlemma} \label{omniscient-is-exhaustible}
-%   Any omniscient type is exhaustible.
-% \end{romlemma}
-% \begin{proof}
-%   For decidable propositions, we know the following:
-%   \begin{equation} \label{lem-conversion}
-%     \Pi (x : A) , P(x) \leftrightarrow \neg \Sigma (x : A) , \neg P(x) 
-%   \end{equation}
-%   To derive exhaustibility from omniscience, then, we run the predicate in its
-%   negated form, and then subsequently negate the result.
-%   The resulting decision over \(\neg \Sigma(x : A) , \neg P(x)\) can be
-%   converted into \(\Pi (x : A) , P(x)\).
-% \end{proof}
-% We cannot derive, however, that any exhaustible type is omniscient, as we do not
-% have the inverse of equation~\ref{lem-conversion}:
-% \begin{equation} \label{choice-conversion}
-%   \Sigma (x : A) , P(x) \leftrightarrow \neg \Pi (x : A) , \neg P(x)
-% \end{equation}
-% Such an equation would allow us to pick a representative element from any type,
-% which is therefore non-constructive.
-% In a sense, equation~\ref{lem-conversion} requires a form of LEM on the
-% proposition (i.e. requires it to be decidable), whereas
-% equation~\ref{choice-conversion} requires a form of choice.
-% Those finiteness predicates which are ordered do in fact give us this form of
-% choice, so the conversion is valid.
-% As such, all of the ordered finiteness predicates imply omniscience.
-% Again, we will prove it only for the weakest.
-% \begin{romtheorem}
-%   Manifest enumerable types are omniscient.
-% \end{romtheorem}
-% Finally, we do have a form of omniscience for prop-valued predicates, as they do
-% not care about the chosen representative.
-% \begin{romtheorem}
-%   Kuratowski finite types are omniscient about prop-valued predicates.
-% \end{romtheorem}
-% \subsection{Synthesising Pattern-Matching Proofs}
-% In particular, they can automate large proofs by analysing every possible case.
-% In \cite{firsovDependentlyTypedProgramming2015}, the \(\AgdaDatatype{Pauli}\)
-% group is used as an example.
-% \ExecuteMetaData[agda/Data/Pauli.tex]{def}
-% As \(\AgdaDatatype{Pauli}\) has 4 constructors, \(n\)-ary functions on
-% \(\AgdaDatatype{Pauli}\) may require up to \(4^n\) cases, making even simple
-% proofs prohibitively verbose.
-
-% The alternative is to derive the things we need from \(\mathcal{E}!\) somehow.
-% As \(\AgdaDatatype{Pauli}\) is a simple finite type, the instance can be defined
-% in a similar way to those in lemma~\ref{split-enum-bool-top-bot}.
-% From here we can already derive decidable equality, a function which requires 16
-% cases if implemented manually.
-
-% For proof search, the procedure is a well-known one in Agda
-% \cite{devrieseBrightSideType2011}: we ask for the result of a decision procedure
-% as an \emph{instance argument}, which will demand computation during
-% typechecking.
-% \subsection{Multiple Arguments}
-% The automation machinery above only deals with single-argument predicates.
-% This is not a problem, as we know that we can work with multiple arguments by
-% currying and uncurrying, since all of the finiteness predicates are closed under
-% \(\times\).
-% To automate away the curry/uncurry noise we will use instance search, building
-% on \cite{allaisGenericLevelPolymorphic2019} to develop a small interface to
-% generic \(n\)-ary functions and properties.
-% Our generic representation can handle dependent \(\Sigma\) and \(\prod\)
-% types (rather than their non-dependent counterparts, \(\times\) and
-% \(\rightarrow\)).
-% This extension was necessary for our use case: it is mentioned in the paper
-% as the obvious next step. 
-% We also implement the curry-uncurry combinators as (verified) isomorphisms.
-
-% A full explanation of our implementation is beyond the scope of this work, so we
-% only present the finished interface, which is used like so:
-% \ExecuteMetaData[agda/Data/Pauli.tex]{assoc-prf}
-% \subsection{Synthesising Functions}
-% Finally, thanks to extensionality provided to us by HoTT, and the computation
-% properties provided by CuTT, we can derive decidable equality on functions over
-% finite types.
-% We can also use functions in our proof search.
-% Here, for instance, is a automated procedure which finds the
-% \(\AgdaFunction{not}\) function on \(\mathbf{Bool}\), given a specification.
-% \ExecuteMetaData[agda/Data/Pauli.tex]{not-spec}
-% \section{Related Work}
-% \cite{spiwackConstructivelyFinite2010} explored the notion of ``constructive
-% finite sets'', demonstrating that there is a rich variety of finiteness
-% predicates in a constructive setting.
-% Some of those predicates where explored more in
-% \cite{firsovVariationsNoetherianness2016a},
-% \cite{parmannInvestigatingStreamlessSets2015}, and
-% \cite{uustaluFinitenessRationalSequences2017a}.
-
-% \cite{yorgeyCombinatorialSpeciesLabelled2014} explored some of the definitions
-% we have given here (manifest bishop finiteness and cardinal finiteness) in HoTT,
-% in order to construct species.
-% Some of this paper is a formalisation of the early parts of that work.
-
-% \cite{firsovDependentlyTypedProgramming2015} explores finite sets in Agda and
-% uses them to automate proof machinery in much the same way as we have here.
-% The primary difference between their proof automation and ours is that we can
-% talk about finite functions, as we are working in HoTT.
-
-% \cite{fruminFiniteSetsHomotopy2018} explores many of the same notions as we do
-% here: cardinal finiteness, and Kuratowski finite sets.
-% Our work makes extensive use of the computational properties of CuTT, however,
-% and we are able to formalise the finiteness of function objects.
-
 \todo[inline]{Add funding ack:

### 2020-01-22 paper.tex
more

@@
 \begin{romdefinition}[\(\mathbf{Fin}\)] \label{Fin}
-  \(\mathbf{Fin}(n)\) is the type of natural numbers smaller than \(n\).
-  We define it the standard way, where \(\mathbf{Fin}(0) = \bot\) and
-  \(\mathbf{Fin}(n + 1) &= \top + \mathbf{Fin}(n)\).
+  \(\mathbf{Fin}(n)\) is the type of natural numbers smaller than \(n\). We
+  define it the standard way, where \(\mathbf{Fin}(0) \coloneqq \bot\) and
+  \(\mathbf{Fin}(n + 1) &\coloneqq \top \uplus \mathbf{Fin}(n)\).
 \end{romdefinition}
@@
 \end{romlemma}
-\begin{romlemma}
+\begin{romlemma} \label{split-enum-discrete}
   Every split enumerable type is discrete.
@@
 \subsection{Split Enumerability and Manifest Bishop Finiteness}
+While manifest Bishop finiteness might seem stronger than split enumerability,
+it turns out this is not the case.
+Both types imply the other.
+\begin{romlemma}
+  Any manifest Bishop finite type is split enumerable.
+\end{romlemma}
+\begin{proof}
+  The only difference between \(\mathcal{B}\) and \(\mathcal{E}!\) is the
+  membership term.
+  To derive \(\mathcal{E}!\) from \(\mathcal{B}\), then, we need to derive
+  \(\in\) from \(\inunique\).
+  We can do this with fst.
+  \begin{alignat*}{4}
+    &\text{fst} : && \; \Sigma(x : A) , U(x) \rightarrow A \\
+    && \simeq      & \; \text{isContr}(A) \rightarrow A \\
+    && \simeq      & \; (x \in \mathit{xs}) \rightarrow (x \inunique \mathit{xs})
+  \end{alignat*}
+  \qed
+\end{proof}
+
+To derive \(\mathcal{B}\) from \(\mathcal{E}!\) takes significantly more work.
+The ``unique membership'' condition in \(\mathcal{B}\) means that we are not
+permitted duplicates in the support list.
+The first step in the proof, then, is to filter those duplicates out from the
+support list of the \(\mathcal{E}!\) proof: we can do this using the decidable
+equality provided by \(\mathcal{E}!\) (lemma~\ref{split-enum-discrete}).
+From there, all we need to show is that the membership proof carries over
+appropriately.
+\begin{romlemma} \label{split-enum-to-manifest-bishop}
+  Any split enumerable set is manifest Bishop finite.
+\end{romlemma}
 \subsection{Manifest Bishop Finiteness and Manifest Enumerability}
+\begin{romlemma}
+  Any manifestly Bishop finite type is manifestly enumerable.
+\end{romlemma}
+\begin{romlemma}
+  A manifestly enumerable type with decidable equality is manifest Bishop
+  finite.
+\end{romlemma}
 \subsection{Manifest Bishop Finiteness and Cardinal Finiteness}
+\begin{romlemma}
+  Any manifest Bishop finite type is cardinal finite.
+\end{romlemma}
+\begin{romtheorem}
+  Any cardinal finite type with a total order is Bishop finite.
+\end{romtheorem}
 \subsection{Manifest Enumerability and Kuratowski Finiteness}
+\begin{romlemma}
+  Any manifest enumerable type is Kuratowski finite.
+\end{romlemma}
 \subsection{Cardinal Finiteness and Kuratowski Finiteness}
+\begin{romtheorem} \label{kuratowski-cardinal}
+  \begin{equation}
+    \mathcal{K}^f(A) \times \text{Discrete}(A) \simeq \mathcal{C}(A)
+  \end{equation}
+\end{romtheorem}
 \section{Topos}
-\begin{itemize}
-  \item \(\bot\)
-  \item \(\top\)
-  \item Bool
-  \item \(\Sigma\)
-  \item \(\Pi\)
-  \item Pre-categories to categories
-  \item \(\Pi\)-pretopos
-\end{itemize}
+Our main result in this section is the formal proof that decidable Kuratowski
+finite types form a \(Pi\)-pretopos.
+As we saw in theorem~\ref{kuratowski-cardinal}, decidable Kuratowski finite
+types are equivalent to cardinal finite types, so we will work with the latter
+from now on.
+Our first task is to show that cardinal finite types are closed under several
+operations.
+\subsection{Closure}
+For the first two closure proofs, we only consider manifest Bishop finiteness:
+as it is the strongest of the finiteness predicates, we can derive the other
+closure proofs from it.
+\begin{romlemma}
+  \(\bot\) is manifest Bishop finite.
+\end{romlemma}
+\begin{proof}
+  From lemma~\ref{bishop-equiv} we can prove the goal by constructing a term of
+  the type:
+  \begin{equation}
+    \Sigma(n : \mathbb{N}) , (\mathbf{Fin}(n) \simeq \bot)
+  \end{equation}
+  To do so, we take \(n\) to be 0, and by definition~\ref{Fin},
+  \(\mathbf{Fin}(0) = \bot\).
+  The equivalence is simply the identity equivalence.
+\end{proof}
+\begin{romlemma}
+  \(\top\) is manifest Bishop finite.
+\end{romlemma}
+\begin{proof}
+  Again, we need only provide a suitable cardinality (in this case 1), and an
+  equivalence with \(\mathbf{Fin}(1)\) to prove our goal.
+  \begin{align*}
+    \top &\simeq \mathbf{Fin}(1) \\
+          &\simeq \top \uplus \mathbf{Fin}(0) \\
+          &\simeq \top \uplus \bot \\
+          &\simeq \top
+  \end{align*}
+  \qed
+\end{proof}
+\begin{romlemma}
+  \(\mathbf{Bool}\) is manifest Bishop finite.
+\end{romlemma}
+\begin{proof}
+  We prove this by showing an isomorphism with \(\mathbf{Fin}(2)\).
+  \begin{align*}
+    \text{false} & \leftrightarrow 0 \\
+    \text{true} & \leftrightarrow 1
+  \end{align*}
+  \qed
+\end{proof}
+
+\begin{romlemma}
+  Split enumerability is closed under \(\Sigma\),
+\end{romlemma}
+\begin{romlemma}
+  Manifest bishop finiteness is closed over dependent functions
+  (\(\prod\)-types).
+  \begin{equation}
+    \frac{
+      \mathcal{B}(A) \; \; \; \Pi {(x : A)} , \mathcal{B}\left( U(x) \right)
+    }{
+      \mathcal{B}\left(\Pi {(x : A)} , U(x)\right)
+    }
+  \end{equation}
+\end{romlemma}
+\begin{proof}
+  This proof is essentially the composition of two transport operations, made
+  available to us via univalence.
+
+  First, we will simplify things slightly by working only with split
+  enumerability.
+  As this is equal in strength to manifest Bishop finiteness, any closure proofs
+  carry over.
+
+  Secondly, we will replace \(A\) in all places with \(\mathbf{Fin}(n)\).
+  Since we have already seen an equivalence between these two types, we are
+  permitted to transport along these lines.
+  This is the first transport operation.
+
+  The bulk of the proof now is concerned with proving the following:
+  \begin{equation}
+    \left( \Pi {(x : \mathbf{Fin}(n))} , \mathcal{E}!(A(x)) \right) \rightarrow \mathcal{E}!\left( \Pi {(x : \mathbf{Fin}(n))} , A(x) \right)
+  \end{equation}
+  Our strategy to accomplish this will be to consider functions from
+  \(\mathbf{Fin}(n)\) as \(n\)-tuples over some type family \(T : \mathbb{N}
+  \rightarrow \text{Type}\).
+  \begin{equation}
+    \begin{aligned}
+      \mathbf{Tuple}(T, 0)   &= \top \\
+      \mathbf{Tuple}(T, n+1) &= T(0) \times \mathbf{Tuple}(T \circ \text{suc}, n)
+    \end{aligned}
+  \end{equation}
+  This type is manifestly Bishop finite, as it is constructed only from products
+  and the unit type.
+
+  We then prove an isomorphism between this representation and \(\Pi\)-types.
+  \begin{equation}
+    \mathbf{Tuple}(T, n) \iff \Pi {(x : \mathbf{Fin}(n))} , T(x)
+  \end{equation}
+  This allows us to transport our proof of finiteness on tuples to one on
+  functions from \(\mathbf{Fin}\) (our second transport operation), proving our
+  goal.
+\end{proof}
+\subsection{The Category of Finite Sets}
+\subsection{The \(\Pi\)-pretopos of Finite Sets}
 \section{Countably Infinite Types}

### 2020-01-23 paper.tex
more

@@
 \end{romdefinition}
-\begin{romdefinition}[Fibers] \label{fibers}
-  A fiber is defined over some function \(f : A \rightarrow B\).
+\begin{romdefinition}[Fibers] \label{fibres}
+  A fibre is defined over some function \(f : A \rightarrow B\).
   \begin{equation}
@@
 \subsection{Manifest Bishop Finiteness}
+\todo[noline]{Add explanation for word ``manifest''}
 \begin{romdefinition}[Manifest Bishop Finiteness]
@@
 \end{romtheorem}
-\subsection{Manifest Enumerability and Kuratowski Finiteness}
-\begin{romlemma}
-  Any manifest enumerable type is Kuratowski finite.
-\end{romlemma}
-\subsection{Cardinal Finiteness and Kuratowski Finiteness}
-\begin{romtheorem} \label{kuratowski-cardinal}
-  \begin{equation}
-    \mathcal{K}^f(A) \times \text{Discrete}(A) \simeq \mathcal{C}(A)
-  \end{equation}
-\end{romtheorem}
 \section{Topos}
@@
 \end{romlemma}
-\begin{proof}
-  We prove this by showing an isomorphism with \(\mathbf{Fin}(2)\).
-  \begin{align*}
-    \text{false} & \leftrightarrow 0 \\
-    \text{true} & \leftrightarrow 1
-  \end{align*}
-  \qed
-\end{proof}
 
@@
 \subsection{The Category of Finite Sets}
+HoTT and CuTT seem to be especially suitable settings for formalisations of
+category theory.
+The univalence axiom in particular allows us to treat categorical isomorphisms
+as the important, useful objects that they are, upgrading them to full
+equalities.
+
+We follow \cite[chapter 9]{hottbook} in its treatment of
+categories in HoTT, and in its proof that sets do indeed form a category.
+We will first briefly go through the construction of the category
+\(\mathit{Set}\), as it differs slightly from the usual method in type theory.
+
+First, the type of objects and arrows:
+\begin{alignat}{3}
+  &\text{Obj}_\mathit{Set}      &&\coloneqq \Sigma(x : \mathbf{Type}) , \text{isProp}(x) \\
+  &\text{Hom}_\mathit{Set}(x , y) &&\coloneqq  \text{fst}(x) \rightarrow \text{fst}(y)
+\end{alignat}
+As is plain in the type of objects, we have already departed slightly from the
+simpler \(\text{Obj}_\mathit{Set} \coloneqq \mathbf{Type}\) way of doing things:
+of course we have to, as HoTT allows non-set types.
+Furthermore, after proving the usual associativity and identity laws for
+composition (which are definitionally true in this case), we must further show
+\(\text{isSet}(\text{Hom}_\mathit{Set}(x,y))\); even then we only have a
+precategory.
+
+To show that \(\mathit{Set}\) is a category, we must show that categorical
+isomorphisms are equivalent to equivalences.
+In a sense, we must give a univalence rule for the category we are working in.
+
+We have provided formal proofs that \(\mathit{Set}\) does indeed form a
+category, and the following:
+\begin{romtheorem}[The Category of Finite Sets]
+  Finite sets form a category in HoTT when defined like so:
+  \begin{equation}
+    \begin{aligned}
+      &\text{Obj}_\mathit{FinSet}      &&\coloneqq \Sigma(x : \mathbf{Type}) , \mathcal{C}(x) \\
+      &\text{Hom}_\mathit{FinSet}(x , y) &&\coloneqq  \text{fst}(x) \rightarrow \text{fst}(y)
+    \end{aligned}
+  \end{equation}
+\end{romtheorem}
 \subsection{The \(\Pi\)-pretopos of Finite Sets}
+For this proof, we follow again the proof that \(\mathit{Set}\) forms a \(\Pi
+W\)-pretopos from \cite[chapter 10]{hottbook} and
+\cite{rijkeSetsHomotopyType2015}.
 \section{Countably Infinite Types}
+In the previous sections we saw different flavours of finiteness which were
+really just different flavours of relations to \(\mathbf{Fin}\).
+In this section we will see that we can construct a similar classification of
+relations to \(\mathbb{N}\), in the form of the countably infinite types.
+\subsection{Split Countable Types}
+Our first foray into the world of countable types will be a straightforward
+analogue to the split enumerable types.
+We need change only one element: instead of a support \emph{list}, we instead
+have a support \emph{stream}, which is its infinite.
+\begin{romdefinition}[Stream]
+  \begin{equation}
+    \mathbf{Stream}(A) = \llbracket \top \triangleright \text{const} (\mathbb{N}) \rrbracket(A) \simeq \mathbb{N} \rightarrow A
+  \end{equation}
+\end{romdefinition}
+\begin{romdefinition}[Split Countability]
+  \begin{equation}
+    \mathcal{E}!(A) = \Sigma {(\mathit{xs} : \mathbf{Stream}(A))} , \Pi {(x : A)} , x \in \mathit{xs}
+  \end{equation}
+\end{romdefinition}
+This type is definitionally equal to it surjection equivalent (\(\mathbb{N}
+\twoheadrightarrow ! \; A\)).
+\subsubsection{Closure}
+We know that countable infinity is not closed under the exponential (function
+arrow), so the only closure we need to prove is \(\Sigma\) to cover all of
+what's left.
+\input{figures/pairing-functions.tex}
+\begin{romtheorem} \label{split-countability-sigma}
+  Split countability is closed under \(\Sigma\).
+\end{romtheorem}
+\begin{proof}
+  The main task here is to figure out a pattern which will pair elements of the
+  two support streams without diverging.
+  Figure~\ref{cartesian} illustrates why the pairing function we used for lists
+  won't work: it needs to exhaust one of the input lists before inspecting
+  anything other than the first element of the other.
+  Of course, we can't exhaust a stream, so this diverges.
+
+  Instead, we will use the pattern in figure~\ref{cantor}.
+  We will produce a stream of lists (where each list is a diagonal), which is
+  then concatenated to the final output stream.
+  The intermediate stream is sometimes called the discrete convolution of the
+  two input streams.
+\end{proof}
+\subsubsection{Kleene Star}
+While we lose some closures with the inclusion of infinite types, we gain some
+others.
+In particular, we have the Kleene star.
+This means that we have closure under lists.
+\begin{romtheorem}
+  Split countability is closed under Kleene star.
+  \begin{equation}
+    \mathcal{E}!(A) \rightarrow \mathcal{E}!(\mathbf{List}(A))
+  \end{equation}
+\end{romtheorem}
+\begin{proof}
+  As with the proof of closure under \(\Sigma\), our main task here is to figure
+  out a pattern that reaches any given finite list in finite time.
+  Our first step is to split up the problem as we did for \(\Sigma\)-closure: we
+  produce an intermediate stream of lists, which we then concatenate to give the
+  final result.
+  The intermediate stream is constructed like so:
+  \begin{equation}
+    \mathit{xs}^\star_i = \left\{ \text{map}(\mathit{xs} , x) \mid \mathit{x} \in \mathbf{List}(\mathbb{N}) , \text{sum}(\text{map}(1+, x)) \equiv i \right\}
+  \end{equation}
+  In other words, the \(i\)th element is all lists such that the sum of the
+  successor of the indices in the list is equal to \(i\).
+\end{proof}
+\subsection{Other Types of Countability}
+There are of course other forms of countability available to us: manifest
+countability, cardinal, etc.
+We have a similar structure of relationships to the finite types, with the
+exception that we cannot construct an isomorphism with \(\mathbb{N}\) fiven a
+surjection from \(\mathbb{N}\).
 \begin{itemize}
@@
 \section{Search}
-\begin{itemize}
-  \item Omniscience
-  \item Search
-  \item Curry / uncurry
-  \item Synth function
-\end{itemize}
+\subsection{Omniscience}
+In this section we are interested in restricted forms of the limited principle
+of omniscience \cite{myhillErrettBishopFoundations1972}.
+\begin{romdefinition}[Limited Principle of Omniscience]
+  For any type \(A\) and predicate \(P\) on \(A\), the limited principle of
+  omniscience is as follows:
+  \begin{equation}
+    \left( \Pi {(x : A)} , \mathbf{Dec}(P(x)) \right) \rightarrow \mathbf{Dec} \left( \Sigma {(x : A)} , P(x) \right)
+  \end{equation}
+  In other words, for any decidable predicate the existential quantification of
+  that predicate is also decidable.
+\end{romdefinition}
+The limited principle of omniscience is non-constructive, but individual types
+can themselves satisfy omniscience.
+In particular, \emph{finite} types are omniscient.
+
+There is also a universal form of omniscience, which we call exhaustibility.
+\begin{romdefinition}[Exhaustibility]
+  We say a type \(A\) is exhaustible if, for any decidable predicate \(P\) on
+  \(A\), the universal quantification of the predicate is decidable.
+  \begin{equation}
+    \left( \Pi {(x : A)} , \mathbf{Dec}(P(x)) \right) \rightarrow \mathbf{Dec} \left( \Pi {(x : A)} , P(x) \right)
+  \end{equation}
+\end{romdefinition}
+
+All of the finiteness predicates we have seen justify exhaustibility.
+We will only prove it once, then, for the weakest:
+\begin{romtheorem}
+  Kuratowski-finite types are exhaustible.
+\end{romtheorem}
+
+Omniscience is stronger than exhaustibility, as we can derive the latter from
+the former:
+\begin{romlemma} \label{omniscient-is-exhaustible}
+  Any omniscient type is exhaustible.
+\end{romlemma}
+\begin{proof}
+  For decidable propositions, we know the following:
+  \begin{equation} \label{lem-conversion}
+    \Pi (x : A) , P(x) \leftrightarrow \neg \Sigma (x : A) , \neg P(x) 
+  \end{equation}
+  To derive exhaustibility from omniscience, then, we run the predicate in its
+  negated form, and then subsequently negate the result.
+  The resulting decision over \(\neg \Sigma(x : A) , \neg P(x)\) can be
+  converted into \(\Pi (x : A) , P(x)\).
+\end{proof}
+We cannot derive, however, that any exhaustible type is omniscient, as we do not
+have the inverse of equation~\ref{lem-conversion}:
+\begin{equation} \label{choice-conversion}
+  \Sigma (x : A) , P(x) \leftrightarrow \neg \Pi (x : A) , \neg P(x)
+\end{equation}
+Such an equation would allow us to pick a representative element from any type,
+which is therefore non-constructive.
+In a sense, equation~\ref{lem-conversion} requires a form of LEM on the
+proposition (i.e. requires it to be decidable), whereas
+equation~\ref{choice-conversion} requires a form of choice.
+Those finiteness predicates which are ordered do in fact give us this form of
+choice, so the conversion is valid.
+As such, all of the ordered finiteness predicates imply omniscience.
+Again, we will prove it only for the weakest.
+\begin{romtheorem}
+  Manifest enumerable types are omniscient.
+\end{romtheorem}
+Finally, we do have a form of omniscience for prop-valued predicates, as they do
+not care about the chosen representative.
+\begin{romtheorem}
+  Kuratowski finite types are omniscient about prop-valued predicates.
+\end{romtheorem}
+\subsection{Synthesising Pattern-Matching Proofs}
+In particular, they can automate large proofs by analysing every possible case.
+In \cite{firsovDependentlyTypedProgramming2015}, the \(\AgdaDatatype{Pauli}\)
+group is used as an example.
+\ExecuteMetaData[agda/Data/Pauli.tex]{def}
+As \(\AgdaDatatype{Pauli}\) has 4 constructors, \(n\)-ary functions on
+\(\AgdaDatatype{Pauli}\) may require up to \(4^n\) cases, making even simple
+proofs prohibitively verbose.
+
+The alternative is to derive the things we need from \(\mathcal{E}!\) somehow.
+As \(\AgdaDatatype{Pauli}\) is a simple finite type, the instance can be defined
+in a similar way to those in lemma~\ref{split-enum-bool-top-bot}.
+From here we can already derive decidable equality, a function which requires 16
+cases if implemented manually.
+
+For proof search, the procedure is a well-known one in Agda
+\cite{devrieseBrightSideType2011}: we ask for the result of a decision procedure
+as an \emph{instance argument}, which will demand computation during
+typechecking.
+\subsection{Multiple Arguments}
+The automation machinery above only deals with single-argument predicates.
+This is not a problem, as we know that we can work with multiple arguments by
+currying and uncurrying, since all of the finiteness predicates are closed under
+\(\times\).
+To automate away the curry/uncurry noise we will use instance search, building
+on \cite{allaisGenericLevelPolymorphic2019} to develop a small interface to
+generic \(n\)-ary functions and properties.
+Our generic representation can handle dependent \(\Sigma\) and \(\prod\)
+types (rather than their non-dependent counterparts, \(\times\) and
+\(\rightarrow\)).
+This extension was necessary for our use case: it is mentioned in the paper
+as the obvious next step. 
+We also implement the curry-uncurry combinators as (verified) isomorphisms.
+
+A full explanation of our implementation is beyond the scope of this work, so we
+only present the finished interface, which is used like so:
+\ExecuteMetaData[agda/Data/Pauli.tex]{assoc-prf}
+\subsection{Synthesising Functions}
+Finally, thanks to extensionality provided to us by HoTT, and the computation
+properties provided by CuTT, we can derive decidable equality on functions over
+finite types.
+We can also use functions in our proof search.
+Here, for instance, is a automated procedure which finds the
+\(\AgdaFunction{not}\) function on \(\mathbf{Bool}\), given a specification.
+\ExecuteMetaData[agda/Data/Pauli.tex]{not-spec}
 

### 2020-01-23 paper.tex
effective quotients

@@
 \end{alignat}
-As is plain in the type of objects, we have already departed slightly from the
+As the type of objects makes clear, we have already departed slightly from the
 simpler \(\text{Obj}_\mathit{Set} \coloneqq \mathbf{Type}\) way of doing things:
@@
 \cite{rijkeSetsHomotopyType2015}.
+The difference here is that clearly we do not have access to \(W\)-types, as
+they would permit infinitary structures.
+
+We first must show that \(\mathit{Set}\) has an initial object and finite,
+disjoint sums, which are stable under pullback.
+We also must show that \(\mathit{Set}\) is a regular category with effective
+quotients.
+We now have a pretopos: the presence of \(\Pi\) types make it a
+\(\Pi\)-pretopos.
+
 \section{Countably Infinite Types}

### 2020-01-23 paper.tex
more

@@
 
+We have proven the above statements for both \(\mathit{Set}\) and
+\(\mathit{FinSet}\).
+As far as we know, this is the first formalisation of either.
 \section{Countably Infinite Types}
@@
 relations to \(\mathbb{N}\), in the form of the countably infinite types.
-\subsection{Split Countable Types}
-Our first foray into the world of countable types will be a straightforward
-analogue to the split enumerable types.
-We need change only one element: instead of a support \emph{list}, we instead
-have a support \emph{stream}, which is its infinite.
+\subsection{Two Countable Types}
+The two types for countability we will consider are analogous to split
+enumerability and cardinal finiteness.
+The change will be a simple one: in terms of surjections and bijections, we will
+swap out \(\mathbf{Fin}\) for \(\mathbb{N}\).
+In terms of support lists and cover proofs, we will swap out lists for streams.
+A stream can be defined many ways in type theory: we take rather a low-tech
+approach here, saying that a stream is a function from the natural numbers.
 \begin{romdefinition}[Stream]
   \begin{equation}
-    \mathbf{Stream}(A) = \llbracket \top \triangleright \text{const} (\mathbb{N}) \rrbracket(A) \simeq \mathbb{N} \rightarrow A
+    \mathbf{Stream}(A) \coloneqq \mathbb{N} \rightarrow A
   \end{equation}
+  Streams are also definable as containers:
+  \begin{equation}
+    \mathbf{Stream}(A) \simeq \llbracket \top , \text{const}(\mathbb{N}) \rrbracket
+  \end{equation}
+  However the two definitions are clearly equivalent, and furthermore it is
+  simpler to work with the first.
 \end{romdefinition}
+
+From this, we can define ``split countability''.
 \begin{romdefinition}[Split Countability]
   \begin{equation}
-    \mathcal{E}!(A) = \Sigma {(\mathit{xs} : \mathbf{Stream}(A))} , \Pi {(x : A)} , x \in \mathit{xs}
+    \mathcal{E}!(A) \coloneqq \Sigma {(\mathit{xs} : \mathbf{Stream}(A))} , \Pi {(x : A)} , x \in \mathit{xs}
   \end{equation}
@@
 \twoheadrightarrow ! \; A\)).
-\subsubsection{Closure}
+We construct the unordered, propositional version of the predicate in much the
+same way as we constructed cardinal finiteness.
+\begin{romdefinition}[Countability]
+  \begin{equation}
+    \mathcal{E}(A) \coloneqq \lVert \mathcal{E}!(A) \rVert
+  \end{equation}
+\end{romdefinition}
+
+From both of these types we can derive decidable equality.
+\begin{romlemma}
+  Any countable type has decidable equality.
+\end{romlemma}
+\subsection{Closure}
 We know that countable infinity is not closed under the exponential (function
@@
 what's left.
-\input{figures/pairing-functions.tex}
 \begin{romtheorem} \label{split-countability-sigma}
@@
 \end{romtheorem}
-\begin{proof}
-  The main task here is to figure out a pattern which will pair elements of the
-  two support streams without diverging.
-  Figure~\ref{cartesian} illustrates why the pairing function we used for lists
-  won't work: it needs to exhaust one of the input lists before inspecting
-  anything other than the first element of the other.
-  Of course, we can't exhaust a stream, so this diverges.
-
-  Instead, we will use the pattern in figure~\ref{cantor}.
-  We will produce a stream of lists (where each list is a diagonal), which is
-  then concatenated to the final output stream.
-  The intermediate stream is sometimes called the discrete convolution of the
-  two input streams.
-\end{proof}
-\subsubsection{Kleene Star}
-While we lose some closures with the inclusion of infinite types, we gain some
-others.
-In particular, we have the Kleene star.
-This means that we have closure under lists.
 \begin{romtheorem}
@@
 \end{romtheorem}
-\begin{proof}
-  As with the proof of closure under \(\Sigma\), our main task here is to figure
-  out a pattern that reaches any given finite list in finite time.
-  Our first step is to split up the problem as we did for \(\Sigma\)-closure: we
-  produce an intermediate stream of lists, which we then concatenate to give the
-  final result.
-  The intermediate stream is constructed like so:
-  \begin{equation}
-    \mathit{xs}^\star_i = \left\{ \text{map}(\mathit{xs} , x) \mid \mathit{x} \in \mathbf{List}(\mathbb{N}) , \text{sum}(\text{map}(1+, x)) \equiv i \right\}
-  \end{equation}
-  In other words, the \(i\)th element is all lists such that the sum of the
-  successor of the indices in the list is equal to \(i\).
-\end{proof}
-\subsection{Other Types of Countability}
-There are of course other forms of countability available to us: manifest
-countability, cardinal, etc.
-We have a similar structure of relationships to the finite types, with the
-exception that we cannot construct an isomorphism with \(\mathbb{N}\) fiven a
-surjection from \(\mathbb{N}\).
-\begin{itemize}
-  \item Definition
-  \item \(\Sigma\)
-  \item Kleene star
-  \item Truncated
-  \item Topos
-\end{itemize}
+\subsection{Categories}
 \section{Search}

### 2020-01-25 paper.tex
fixed author names
begining edits

@@
 \author{
-  Donnacha Oisín Kidney \inst{1} \and
-  Gregory Provan \inst{2}
+  Donnacha Oisín Kidney\inst{1}\and Gregory Provan \inst{2}
 }
@@
 \end{abstract}
-
 \section{Introduction}
@@
 There are a handful of ways to demonstrate some type is finite: we could provide
-a surjection from another finite type; we could show that any collection of its
-elements larger than some bound contains duplicates; or that any stream of its
-elements contain duplicates.
+a surjection to it from another finite type; we could show that any collection
+of its elements larger than some bound contains duplicates; or we could show
+that any stream of its elements contain duplicates.
 
@@
 
-Finally, the other important property of constructive mathematics is that we can
-actually \emph{compute} with proofs.
+Finally, for a computer scientist constructive mathematics has one invaluable
+feature missing from classical mathematics: computation.
 Cubical Type Theory \cite{cohenCubicalTypeTheory2016}, and its implementation in
@@
 \subsection{Contributions}
-In this work we will examine five notions of finiteness in Homotopy Type Theory,
+In this work we will examine five notions of finiteness in Cubical Type Theory,
 the relationships between them, and their topos-theoretic characterisation.

### 2020-01-25 paper.tex
more

@@
 The finiteness predicates we are interested in are organised in
-Figure~\ref{finite-classification}.
+figure~\ref{finite-classification}, and defined in
+section~\ref{finiteness-predicates}.
 We will explore two aspects of each predicate: its relation to the other
@@
 When we say ``relation'' we are referring to the arrows in
-Figure~\ref{finite-classification}.
-The proofs, then, amount to a function which inhabits each arrow.
+figure~\ref{finite-classification}.
+In section~\ref{relations} we will provide a function which inhabits each arrow.
 Each unlabelled arrow is a weakening: i.e., every manifest Bishop finite set
-is manifest enumerable.
+is manifest enumerable (lemma~\ref{manifest-bishop-to-manifest-enum}).
 The labelled arrows are strengthening proofs: i.e., every manifest enumerable set
-\emph{with decidable equality} is manifest Bishop finite.
+\emph{with decidable equality} is manifest Bishop finite
+(lemma~\ref{manifest-enum-to-manifest-bishop}).
 Our most significant result here is the proof that a cardinal finite set with a
@@
 
-We will then characterise each predicate as some form of topos.
+We will then examine the closure properties of each predicate in
+section~\ref{topos}, culminating in a proof that decidable Kuratowski finite
+sets form a \(\Pi\)-pretopos (theorem~\ref{finite-topos}).
 Our proofs follow the structure of \cite[Chapters 9, 10]{hottbook} and
 \cite{rijkeSetsHomotopyType2015}.
-This means we will define first the precategory of sets, and then the category
-of sets, and prove the required closures and limits to get us to a topos.
-We say ``some form of'' topos here because of course the category of sets in
-HoTT do \emph{not}, in fact, form a topos, rather a \(\Pi W\)-pretopos.
-Our main result here is that the category of decidable Kuratowski finite sets
-forms a \(\Pi\)-pretopos.
 
 After the finite predicates, we will briefly look at the infinite countable
-types, and classify them in a parallel way to the finite predicates.
-We here show that countably finite sets form a \(W\)-pretopos.
+types, and classify them in a parallel way to the finite predicates
+(section~\ref{infinite}).
 
@@
 We will make mention of the few occasions where the formalisation of some proof
-is of interest, but the main place where we will discuss the code is in the
-final section, where we implement a library for proof search, based on
+is of interest, but the main place where we will discuss the code is in
+section~\ref{search}, where we implement a library for proof search, based on
 omniscience and exhaustibility.
@@
 Furthermore, the extensionality afforded to us by HoTT means that we can have
-\(\Pi\) types in the domain of the search, meaning we can synthesise full
-functions from specifications alone.
-\subsection{Related Work}
+\(\Pi\) types in the domain of the search, making our library more powerful than
+comparable libraries.
+We will demonstrate some interesting uses for this feature in particular.
 \subsection{Notation and Background}
@@
 \end{romdefinition}
-\section{Finiteness Predicates}
+\section{Finiteness Predicates} \label{finiteness-predicates}
 In this section, we will define and briefly describe each of the five predicates
@@
 equality from a proof of Kuratowski finiteness.
-\section{Relations Between Each Finiteness Definition}
+\section{Relations Between Each Finiteness Definition} \label{relations}
 We now look at under what conditions a given finiteness predicate satisfies
@@
 Both types imply the other.
-\begin{romlemma}
+\begin{romlemma} \label{manifest-bishop-to-split-enum}
   Any manifest Bishop finite type is split enumerable.
 \end{romlemma}
-\begin{proof}
-  The only difference between \(\mathcal{B}\) and \(\mathcal{E}!\) is the
-  membership term.
-  To derive \(\mathcal{E}!\) from \(\mathcal{B}\), then, we need to derive
-  \(\in\) from \(\inunique\).
-  We can do this with fst.
-  \begin{alignat*}{4}
-    &\text{fst} : && \; \Sigma(x : A) , U(x) \rightarrow A \\
-    && \simeq      & \; \text{isContr}(A) \rightarrow A \\
-    && \simeq      & \; (x \in \mathit{xs}) \rightarrow (x \inunique \mathit{xs})
-  \end{alignat*}
-  \qed
-\end{proof}
 
@@
 \subsection{Manifest Bishop Finiteness and Manifest Enumerability}
-\begin{romlemma}
+\begin{romlemma} \label{manifest-bishop-to-manifest-enum}
   Any manifestly Bishop finite type is manifestly enumerable.
 \end{romlemma}
-\begin{romlemma}
+\begin{romlemma} \label{manifest-enum-to-manifest-bishop}
   A manifestly enumerable type with decidable equality is manifest Bishop
@@
 \subsection{Manifest Bishop Finiteness and Cardinal Finiteness}
-\begin{romlemma}
+\begin{romlemma} \label{manifest-bishop-to-cardinal}
   Any manifest Bishop finite type is cardinal finite.
 \end{romlemma}
-\begin{romtheorem}
+\begin{romtheorem} \label{cardinal-to-manifest-bishop}
   Any cardinal finite type with a total order is Bishop finite.
 \end{romtheorem}
-\section{Topos}
+\subsection{Cardinal Finiteness and Kuratowski Finiteness}
+\begin{romlemma} \label{cardinal-kuratowski}
+  \begin{equation}
+    \mathcal{C}(A) \simeq \mathcal{K}(A) \times \text{Discrete}(A)
+  \end{equation}
+\end{romlemma}
+\section{Topos} \label{topos}
 Our main result in this section is the formal proof that decidable Kuratowski
@@
 As far as we know, this is the first formalisation of either.
-\section{Countably Infinite Types}
+\begin{romtheorem} \label{finite-topos}
+  The category of finite sets, \(\mathit{FinSet}\), forms a \(\Pi\)-topos.
+\end{romtheorem}
+\section{Countably Infinite Types} \label{infinite}
 In the previous sections we saw different flavours of finiteness which were
@@
 \subsection{Categories}
-\section{Search}
+\section{Search} \label{search}
 \subsection{Omniscience}

### 2020-01-25 paper.tex
more

@@
   \item[Type] We use \(\mathbf{Type}\) to denote the universe of (small) types.
-    ``Type families'' are functions into type.
+    ``Type families'' are functions into \(\mathbf{Type}\).
   \item[Universes]
-    Implicitly, types will exist at some point in a universe hierarchy,
-    beginning with \(\mathbf{Type}\) (\(\mathbf{Type} : \mathbf{Type}_1\),
-    \(\mathbf{Type}_1 : \mathbf{Type}_2\), and so on).
+    To avoid Girard's paradox
+    \cite{girardInterpretationFonctionelleElimination1972} we use explicit
+    universe levels.
+    A full explanation is beyond the scope of this work, but it means that there
+    is a hierarchy of types starting with \(\mathbf{Type}\).
+    \begin{equation}
+      \begin{alignat}{3}
+        & 2               &&: \mathbb{N} \\
+        & \mathbb{N}      &&: \mathbf{Type} \\
+        & \mathbf{Type}   &&: \mathbf{Type}_1 \\
+        & \mathbf{Type}_1 &&: \mathbf{Type}_2
+      \end{alignat}
+    \end{equation}
     In our formalisation, every proof is done in the most universe polymorphic
@@
     \(\mathbf{Bool}\) are false and true.
+    The ``negation'' of a type, written \(\neg A\), means \(A \rightarrow
+    \bot\).
   \item[Dependent Sum and Product] We use \(\Sigma\) and \(\Pi\) for the
@@
     Of course, we know that \((A \simeq B) \simeq (A \equiv B)\) by univalence,
-    so the distinction isn't terribly important in usage.
+    so the distinction isn't terribly important in usage: we will only use one
+    or the other as a suggestion of how we constructed it or how it is to be
+    used.
 \end{description}
@@
   \end{alignat}
+  This type is defined similarly in \cite[definition 7.1.1]{hottbook}, although
+  there the numbering starts at \(-2\).
 \end{romdefinition}
-\begin{romdefinition}[Fibers] \label{fibres}
-  A fibre is defined over some function \(f : A \rightarrow B\).
+\begin{romdefinition}[Fibres] \label{fibres}
+  A fibre \cite[definition 4.2.4]{hottbook} is defined over some function \(f :
+  A \rightarrow B\).
   \begin{equation}
@@
   \end{equation}
+  This is the so-called univalence ``axiom'', which is a theorem in CuTT.
 \end{romlemma}
-\begin{romdefinition}[Isomorphism] \label{isomorphism}
-  We say two types \(A\) and \(B\) are isomorphic, denoted with \(A \iff B\), if
-  there is a pair of functions \(f : A \rightarrow B\), \(g : B \rightarrow A\)
-  such that
+\begin{romdefinition}[Decision]
   \begin{equation}
-    \begin{aligned}
-      \Pi(x : A) , (g(f(x)) \equiv x) \\
-      \Pi(x : B) , (f(g(x)) \equiv x)
-    \end{aligned}
+    \mathbf{Dec}(A) \coloneqq A \uplus \neg A
   \end{equation}
-  While we can derive an equivalence from an isomorphism:
+\end{romdefinition}
+\begin{romdefinition}[Discrete]
+  A discrete type is one with decidable equality.
   \begin{equation}
-    \text{isoToEquiv} : (A \iff B) \rightarrow (A \simeq B)
+    \text{Discrete}(A) \coloneqq \Pi(x, y : A) , \mathbf{Dec}(x \equiv y)
   \end{equation}
-  It is not the case that an isomorphism is the \emph{same} as an equivalence in
-  all cases.
-  In other words, we do not have:
-  \begin{equation}
-    (A \iff B) \simeq (A \simeq B)
-  \end{equation}
-  However, if one of \(A\) or \(B\) is a set, then the above holds.
+  By Hedberg's theorem \cite{hedbergCoherenceTheoremMartinLof1998} any discrete
+  type is a set.
 \end{romdefinition}
 \begin{romdefinition}[Higher Inductive Types] \label{HITs}
-  \todo[inline]{define}
-\end{romdefinition}
-\begin{romdefinition}[Surjections] \label{surjections}
-  \begin{alignat}{3}
-    &\text{surj}(f)             &&\coloneqq \Pi(y : B) , \lVert \text{fib}_f(y) \rVert \\
-    &A \twoheadrightarrow B     &&\coloneqq \Sigma (f : A \rightarrow B) , \text{surj}(f) \label{surj-arrow-eqn} \\
-    &\text{sp-surj}(f)          &&\coloneqq \Pi(y : B) , \text{fib}_f(y) \label{sp-surj-eqn} \\
-    &A \twoheadrightarrow! \; B &&\coloneqq \Sigma (f : A \rightarrow B) , \text{sp-surj}(f) \label{sp-surj-arrow-eqn}
-  \end{alignat}
-\end{romdefinition}
-\begin{romdefinition}[Containers] \label{container-def}
-  A container \cite{abbottContainersConstructingStrictly2005} is a pair
-  \(S , P\) where \(S\) is a type, the elements of which are called
-  the \emph{shapes} of the container, and \(P\) is a type family on \(S\), where
-  the elements of \(P(s)\) are called the \emph{positions} of a container.
-  We ``interpret'' a container into a functor defined like so:
-  \begin{equation} \label{container-interp}
-    \llbracket S , P \rrbracket(A) \coloneqq \Sigma {(s : S)} , \left( P(s) \rightarrow A \right)
-  \end{equation}
-  Membership of a container can be defined like so:
-  \begin{equation} \label{container-membership}
-    x \in \mathit{xs} \coloneqq \text{fib}_{\text{snd}(\mathit{xs})}(x)
-  \end{equation}
-\end{romdefinition}
-\begin{romdefinition}[\(\mathbf{List}\)] \label{List}
-  \begin{equation}
-    \mathbf{List} \coloneqq \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket
-  \end{equation}
-\end{romdefinition}
-\begin{romdefinition}[\(\mathbf{Fin}\)] \label{Fin}
-  \(\mathbf{Fin}(n)\) is the type of natural numbers smaller than \(n\). We
-  define it the standard way, where \(\mathbf{Fin}(0) \coloneqq \bot\) and
-  \(\mathbf{Fin}(n + 1) &\coloneqq \top \uplus \mathbf{Fin}(n)\).
+  Normal inductive types have \emph{point} constructors: constructors which
+  construct values of the type.
+  Higher Inductive Types (HITs) also have \emph{path} constructors: ways to
+  construct paths in the type.
 \end{romdefinition}
@@
 \end{romdefinition}
-\begin{romdefinition}[\(S^1\)] \label{circle-def}
-  The circle, \(S^1\), can be represented in HoTT as a higher inductive type.
-  \begin{equation}
-    \begin{alignat}{5}
-      S^1 \coloneqq & \; \text{base} &&: S^1 ; \\
-                  | & \; \text{loop} &&: \text{base} \equiv \text{base} ; 
-    \end{alignat}
-  \end{equation}
-  We will use it here as an example of a non-set type, i.e. a type for which not
-  all paths are equal.
-  This also means that it does not have decidable equality.
-\end{romdefinition}
 \section{Finiteness Predicates} \label{finiteness-predicates}
@@
 
+We used some extra types in the above definition, which we will define here:
+\begin{romdefinition}[Containers] \label{container-def}
+  A container \cite{abbottContainersConstructingStrictly2005} is a pair
+  \(S , P\) where \(S\) is a type, the elements of which are called
+  the \emph{shapes} of the container, and \(P\) is a type family on \(S\), where
+  the elements of \(P(s)\) are called the \emph{positions} of a container.
+  We ``interpret'' a container into a functor defined like so:
+  \begin{equation} \label{container-interp}
+    \llbracket S , P \rrbracket(A) \coloneqq \Sigma {(s : S)} , \left( P(s) \rightarrow A \right)
+  \end{equation}
+  Membership of a container can be defined like so:
+  \begin{equation} \label{container-membership}
+    x \in \mathit{xs} \coloneqq \text{fib}_{\text{snd}(\mathit{xs})}(x)
+  \end{equation}
+\end{romdefinition}
+\begin{romdefinition}[\(\mathbf{List}\)] \label{List}
+  \begin{equation}
+    \mathbf{List} \coloneqq \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket
+  \end{equation}
+\end{romdefinition}
+\begin{romdefinition}[\(\mathbf{Fin}\)] \label{Fin}
+  \(\mathbf{Fin}(n)\) is the type of natural numbers smaller than \(n\). We
+  define it the standard way, where \(\mathbf{Fin}(0) \coloneqq \bot\) and
+  \(\mathbf{Fin}(n + 1) &\coloneqq \top \uplus \mathbf{Fin}(n)\).
+\end{romdefinition}
+
 We tend to prefer list-based definitions of finiteness, rather than ones based
@@
 numbers.
+\begin{romdefinition}[Surjections] \label{surjections}
+  We define both surjections and \emph{split} surjections here \cite[definition
+  4.6.1]{hottbook}.
+  \begin{alignat}{3}
+    &\text{surj}(f)             &&\coloneqq \Pi(y : B) , \lVert \text{fib}_f(y) \rVert \\
+    &A \twoheadrightarrow B     &&\coloneqq \Sigma (f : A \rightarrow B) , \text{surj}(f) \label{surj-arrow-eqn} \\
+    &\text{sp-surj}(f)          &&\coloneqq \Pi(y : B) , \text{fib}_f(y) \label{sp-surj-eqn} \\
+    &A \twoheadrightarrow! \; B &&\coloneqq \Sigma (f : A \rightarrow B) , \text{sp-surj}(f) \label{sp-surj-arrow-eqn}
+  \end{alignat}
+\end{romdefinition}
 \begin{romlemma} \label{split-enum-is-split-surj}
@@
 \subsection{Manifest Bishop Finiteness}
-\todo[noline]{Add explanation for word ``manifest''}
 \begin{romdefinition}[Manifest Bishop Finiteness]
@@
 \end{romdefinition}
+
+We use the word ``manifest'' here to distinguish from another common
+interpretation of Bishop finiteness, which we have called Cardinal finiteness in
+this paper.
+The ``manifest'' refers to the fact that we have a concrete, non-truncated list
+of the elements in the proof.
 
@@
 \end{proof}
-\subsection{Cardinal}
+\subsection{Cardinal Finiteness}
 Each finiteness predicate so far has contained an \emph{ordering} of the
@@
 More significantly, it allows the predicate to be defined over non-set types,
-like the circle (def.~\ref{circle-def}).
+like the circle.
+\begin{romdefinition}[\(S^1\)] \label{circle-def}
+  The circle, \(S^1\), can be represented in HoTT as a higher inductive type.
+  \begin{equation}
+    \begin{alignat}{5}
+      S^1 \coloneqq & \; \text{base} &&: S^1 ; \\
+      | & \; \text{loop} &&: \text{base} \equiv \text{base} ; 
+    \end{alignat}
+  \end{equation}
+  We will use it here as an example of a non-set type, i.e. a type for which not
+  all paths are equal.
+  This also means that it does not have decidable equality.
+\end{romdefinition}
 \begin{romlemma}

### 2020-01-25 figures/finite-classification.tex
more
finished a lot

@@
     &
-    \text{Split Enumerable}
-      \; \ref{split-enumerable}
+    \text{Manifest Bishop}
       \ar[d, xshift=-1ex, bend right=30]
+      \ar[rrd, end anchor=north west, start anchor=east, bend left=10, crossing over]
       &
@@
     &
-    \text{Manifest Bishop}
+    \text{Split Enumerable}
       \ar[u, xshift=1ex, bend right=30]
-      \ar[rr, end anchor=south west, start anchor=south east, bend right=10, crossing over]
-      \; \ref{manifest-bishop}
       &
@@
     \text{Cardinal}
-      \; \ref{cardinal}
-      \ar[ll, start anchor=north west, end anchor=north east, bend right=10, "\text{linOrd}" description]
+      \ar[ll, start anchor=west, end anchor=east, "\text{Ord}" description]
     \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     \text{Manifest Enumerable}
-      \; \ref{manifest-enumerability}
       \ar[from=uu, xshift=-1ex, bend right=30, crossing over]
       \ar[uu, bend right=30,  "\text{Discrete}" description]
-      \ar[rr, start anchor=south east, end anchor=south west, bend right=10, crossing over]
+      \ar[rr, start anchor=east, end anchor=west, crossing over]
       &
@@
     \text{Kuratowski}
-      \; \ref{kuratowski}
       \ar[uu, xshift=1ex, bend right=30,  "\text{Discrete}" description]
@@
   \caption{Classification of finiteness predicates according to whether they are
-    discrete (imply decidable equality) and whether they induce a linear order.}
+    discrete (imply decidable equality) and whether they imply a total order.}
   \label{finite-classification}

### 2020-01-25 paper.tex
finished a lot

@@
 In section~\ref{relations} we will provide a function which inhabits each arrow.
-Each unlabelled arrow is a weakening: i.e., every manifest Bishop finite set
-is manifest enumerable (lemma~\ref{manifest-bishop-to-manifest-enum}).
-The labelled arrows are strengthening proofs: i.e., every manifest enumerable set
-\emph{with decidable equality} is manifest Bishop finite
-(lemma~\ref{manifest-enum-to-manifest-bishop}).
+Each unlabelled arrow is an unconditional implication: every manifest Bishop
+finite set is cardinal finite (lemma~\ref{manifest-bishop-to-cardinal}), for
+instance.
+The labelled arrows are strengthening proofs: every manifest enumerable set
+\emph{with decidable equality} is split enumerable
+(lemma~\ref{manifest-enum-to-split-enum}).
 Our most significant result here is the proof that a cardinal finite set with a
@@
   This means that equalities are not necessarily unique.
-  In fact, we have an entire hierarchy of homotopy levels, of which sets are
-  level 2.
+  In fact, we have an entire hierarchy of homotopy levels, starting with
+  contractions, of which sets are level 2.
   \begin{alignat}{2}
@@
   \begin{equation}
-    \text{fib}_f(y) = \Sigma(x : A) , (f x \equiv y)
+    \text{fib}_f(y) = \Sigma(x : A) , (f (x) \equiv y)
   \end{equation}
@@
 canonical representative, it will have \(n!\), where \(n\) is the cardinality of
-the type.
+the type\footnotemark.
+
+\footnotetext{
+  We actually do get a category (a groupoid, even) from manifest Bishop
+  finiteness \cite{yorgeyCombinatorialSpeciesLabelled2014}: it's the groupoid of
+  finite sets equipped with a linear order, whose morphisms are order-preserving
+  bijections.
+  We do not explore this particular construction in any detail.
+}
 
@@
   \begin{equation}
-    \mathcal{C}(A) \coloneqq \lVert \mathcal{B}(A) \rVert
+    \mathcal{C}(A) \coloneqq \lVert \mathcal{B}(A) \rVert \simeq \lVert \Sigma(n : \mathbb{N}) , (\mathbf{Fin}(n) \simeq A) \rVert
   \end{equation}
@@
   \end{equation}
-
+  We define it as a HIT (definition~\ref{HITs}).
   The first two constructors are point constructors, giving ways to create
@@
 \section{Relations Between Each Finiteness Definition} \label{relations}
-We now look at under what conditions a given finiteness predicate satisfies
-another.
+We will now look at the arrows in figure~\ref{finite-classification}.
 \subsection{Split Enumerability and Manifest Bishop Finiteness}
@@
 \end{romlemma}
+\begin{proof}
+  To construct a proof of split enumerability from one of manifest Bishop
+  finiteness, it suffices to convert a proof of \(x \inunique \mathit{xs}\) to
+  one of \(x \in \mathit{xs}\), for all \(x\) and \(\mathit{xs}\).
+  Since \(\inunique\) is defined as a contraction of \(\in\), such a conversion
+  is simply the \(\text{fst}\) function.
+  \qed
+\end{proof}
 
@@
 \end{romlemma}
-\subsection{Manifest Bishop Finiteness and Manifest Enumerability}
-\begin{romlemma} \label{manifest-bishop-to-manifest-enum}
-  Any manifestly Bishop finite type is manifestly enumerable.
-\end{romlemma}
-\begin{romlemma} \label{manifest-enum-to-manifest-bishop}
-  A manifestly enumerable type with decidable equality is manifest Bishop
-  finite.
-\end{romlemma}
+\subsection{Split Enumerability and Manifest Enumerability}
+\begin{romlemma} \label{split-enum-to-manifest-enum}
+  Any split enumerable type is manifestly enumerable.
+\end{romlemma}
+\begin{romlemma} \label{dec-trunc-refute}
+  For any type \(A\), we can extract a value of type \(A\) from a decision over
+  \(A\) in combination with a propositionally-truncated value of type \(\lVert A \rVert\).
+  \begin{equation}
+    \frac{
+      \mathbf{Dec}(A) \; \; \; \lVert A \rVert
+    }{
+      A
+    }
+  \end{equation}
+\end{romlemma}
+\begin{proof}
+  Let \(d\) be a value of type \(\mathbf{Dec}(A)\), and \(p\) be a value of type
+  \(\lVert A \rVert\).
+  We proceed by case analysis on \(d\).
+  In the case where \(d\) is inl, we have a value of type \(A\), and our goal is
+  satisfied.
+  In the case of inr, we have a value of type \(\neg A\).
+  Using eliminator \ref{elim-prop-prop} from definition~\ref{prop-trunc}, and
+  the fact that \(\bot\) is a proposition, we can run the \(\neg A\) proof on
+  \(p\), getting a contradiction.
+  \qed
+\end{proof}
+\begin{romlemma} \label{manifest-enum-to-split-enum}
+  A manifestly enumerable type with decidable equality is split enumerable.
+\end{romlemma}
+\begin{proof}
+  Similarly to the proof of lemma~\ref{manifest-bishop-to-split-enum}, our
+  obligation for this proof is to provide a conversion function between
+  membership proofs.
+  In this case, we need to show:
+  \begin{equation}
+    \Pi(x : A) , \Pi(\mathit{xs} : \mathbf{List}(A)) , \lVert x \in \mathit{xs} \rVert \rightarrow x \in \mathit{xs}
+  \end{equation}
+  This conversion function can be constructed from the two following functions,
+  provided in the formalisation.
+  First, for any type with decidable equality, decidable membership in a finite
+  list is consequently decidable:
+  \begin{equation}
+    \text{Discrete}(A) \rightarrow \Pi(x : A) , \Pi(\mathit{xs} : \mathbf{List}(A)) , \mathbf{Dec}(x \in \mathit{xs})
+  \end{equation}
+  Then, we use lemma~\ref{dec-trunc-refute}, allowing us to derive \(x \in
+  \mathit{xs}\) from \(\lVert x \in \mathit{xs} \rVert\).
+  \qed
+\end{proof}
 \subsection{Manifest Bishop Finiteness and Cardinal Finiteness}
@@
 \end{romtheorem}
+The proof for this particular theorem is quite involved in the formalisation, so
+we only give its sketch here.
+First, note that we actually convert to manifest enumerability first: this can
+be converted to split enumerability with decidable equality, which is provided
+by cardinal finiteness.
+
+Next, we define permutations.
+\begin{romdefinition}[List Permutations]
+  Two lists are permutations of each other if their membership proofs are all
+  equivalent\footnotemark \cite{danielssonBagEquivalenceProofRelevant2012}.
+  \begin{equation}
+    \mathit{xs} \leftrightsquigarrow \mathit{ys} = \Pi {(x : A)} , x \in \mathit{xs} \simeq x \in \mathit{ys}
+  \end{equation}
+\end{romdefinition}
+
+\footnotetext{
+  The definition in \cite{danielssonBagEquivalenceProofRelevant2012} and our
+  formalisation is slightly different: we say permutations are lists with
+  \emph{isomorphic} membership proofs.
+  The distinction, as it happens, does not affect our work here.
+}
+
+Next, we define a sort function which relies on the provided total order.
+We further prove the following fact about this sort function:
+\begin{equation}
+  \Pi(\mathit{xs}, \mathit{ys} : \mathbf{List}(A)) , \mathit{xs} \leftrightsquigarrow \mathit{ys} \rightarrow \text{sort}(\mathit{xs}) \equiv \text{sort}(\mathit{ys})
+\end{equation}
+
+Next, notice that the support lists of any two proofs of manifest Bishop
+finiteness must be permutations of each other.
+This will allow us to sort the support list of a proof of cardinal finiteness in
+a coherently constant (definition~\ref{prop-trunc},
+eliminator~\ref{elim-prop-coh}) way, pulling the support list out from the
+truncation.
+The cover proof emerges naturally from the definition of the permutation.
 \subsection{Cardinal Finiteness and Kuratowski Finiteness}
@@
 \end{romlemma}
+This proof is constructed by providing a pair of functions: one from
+\(\mathcal{C}(A)\) to \(\mathcal{K}(A) \times \text{Discrete}(A)\), and one the
+other way.
+This pair implies an equivalence, because both source and target are
+propositions.
+The actual functions themselves are proven in our formalisation, as well as in
+\cite{fruminFiniteSetsHomotopy2018}.
 \section{Topos} \label{topos}
-Our main result in this section is the formal proof that decidable Kuratowski
-finite types form a \(Pi\)-pretopos.
-As we saw in theorem~\ref{kuratowski-cardinal}, decidable Kuratowski finite
+In this section we will prove that decidable Kuratowski finite types form a
+\(\Pi\)-pretopos.
+Along the way we will provide closure proofs for a number of the other
+finiteness predicates.
+As we saw in theorem~\ref{cardinal-kuratowski}, decidable Kuratowski finite
 types are equivalent to cardinal finite types, so we will work with the latter
@@
 \begin{romlemma}
-  \(\bot\) is manifest Bishop finite.
-\end{romlemma}
-\begin{proof}
-  From lemma~\ref{bishop-equiv} we can prove the goal by constructing a term of
-  the type:
-  \begin{equation}
-    \Sigma(n : \mathbb{N}) , (\mathbf{Fin}(n) \simeq \bot)
-  \end{equation}
-  To do so, we take \(n\) to be 0, and by definition~\ref{Fin},
-  \(\mathbf{Fin}(0) = \bot\).
-  The equivalence is simply the identity equivalence.
-\end{proof}
+  \(\bot\), \(\top\), and \(\mathbf{Bool}\) are manifest Bishop finite.
+\end{romlemma}
 \begin{romlemma}
-  \(\top\) is manifest Bishop finite.
-\end{romlemma}
-\begin{proof}
-  Again, we need only provide a suitable cardinality (in this case 1), and an
-  equivalence with \(\mathbf{Fin}(1)\) to prove our goal.
-  \begin{align*}
-    \top &\simeq \mathbf{Fin}(1) \\
-          &\simeq \top \uplus \mathbf{Fin}(0) \\
-          &\simeq \top \uplus \bot \\
-          &\simeq \top
-  \end{align*}
-  \qed
-\end{proof}
-\begin{romlemma}
-  \(\mathbf{Bool}\) is manifest Bishop finite.
-\end{romlemma}
-
-\begin{romlemma}
-  Split enumerability is closed under \(\Sigma\),
-\end{romlemma}
+  Split enumerability is closed under \(\Sigma\)
+\end{romlemma}
+To show closure under \(\Sigma\), we take the Cartesian product of the support
+lists.
+
 \begin{romlemma}
@@
 \begin{proof}
-  This proof is essentially the composition of two transport operations, made
-  available to us via univalence.
-
-  First, we will simplify things slightly by working only with split
+  Firstly, we will simplify things slightly by working only with split
   enumerability.
@@
   carry over.
-
-  Secondly, we will replace \(A\) in all places with \(\mathbf{Fin}(n)\).
-  Since we have already seen an equivalence between these two types, we are
-  permitted to transport along these lines.
-  This is the first transport operation.
-
-  The bulk of the proof now is concerned with proving the following:
-  \begin{equation}
-    \left( \Pi {(x : \mathbf{Fin}(n))} , \mathcal{E}!(A(x)) \right) \rightarrow \mathcal{E}!\left( \Pi {(x : \mathbf{Fin}(n))} , A(x) \right)
-  \end{equation}
-  Our strategy to accomplish this will be to consider functions from
-  \(\mathbf{Fin}(n)\) as \(n\)-tuples over some type family \(T : \mathbb{N}
-  \rightarrow \text{Type}\).
-  \begin{equation}
-    \begin{aligned}
-      \mathbf{Tuple}(T, 0)   &= \top \\
-      \mathbf{Tuple}(T, n+1) &= T(0) \times \mathbf{Tuple}(T \circ \text{suc}, n)
-    \end{aligned}
-  \end{equation}
-  This type is manifestly Bishop finite, as it is constructed only from products
-  and the unit type.
-
-  We then prove an isomorphism between this representation and \(\Pi\)-types.
-  \begin{equation}
-    \mathbf{Tuple}(T, n) \iff \Pi {(x : \mathbf{Fin}(n))} , T(x)
-  \end{equation}
-  This allows us to transport our proof of finiteness on tuples to one on
-  functions from \(\mathbf{Fin}\) (our second transport operation), proving our
-  goal.
+  Secondly, we will replace \(A\) in all places with \(\mathbf{Fin}(n)\), using
+  the equivalence provided by lemma~\ref{bishop-equiv}.
+
+  The remainder of the proof now is concerned with proving the following:
+  \begin{equation} \label{given-transp-sublemma}
+    \left( \Pi {(x : \mathbf{Fin}(n))} , \mathcal{E}!(U(x)) \right) \rightarrow \mathcal{E}!\left( \Pi {(x : \mathbf{Fin}(n))} , U(x) \right)
+  \end{equation}
+
+  To prove this goal, consider the following type of \(n\)-tuples.
+  \begin{equation}
+    \begin{alignat}{3}
+      & \mathbf{Tuple}(T, 0)   &&\coloneqq \top \\
+      & \mathbf{Tuple}(T, n+1) &&\coloneqq T(0) \times \mathbf{Tuple}(T \circ \text{suc}, n)
+    \end{alignat}
+  \end{equation}
+  We have the following equivalence, proven in our formalisation.
+  \begin{equation} \label{fin-equiv-sublemma}
+    \mathbf{Tuple}(T, n) \simeq \Pi {(x : \mathbf{Fin}(n))} , T(x)
+  \end{equation}
+
+  By the given proof of finiteness in equation~\ref{given-transp-sublemma}, we
+  know that \(\mathbf{Tuple}(T , n)\) is split enumerable, since it is made of
+  products of the points of \(U\).
+  We can then transport along the equivalence in
+  equation~\ref{fin-equiv-sublemma}, proving our goal.
 \end{proof}
+
+The dependent cases of each of these proofs do not transfer simply to proofs on
+\(\mathcal{C}\).
+In order to transfer them over, we need to prove a variant of the axiom of
+choice on finite sets.
+\begin{romlemma}
+  \begin{equation}
+    \mathcal{C}(A) \rightarrow (\Pi(x : A) , \lVert U(x) \rVert) \rightarrow \lVert \Pi(x : A) , U(x) \rVert
+  \end{equation}
+\end{romlemma}
+This is proven in our formalisation.
 \subsection{The Category of Finite Sets}
@@
 The univalence axiom in particular allows us to treat categorical isomorphisms
-as the important, useful objects that they are, upgrading them to full
-equalities.
+as equalities, saving us from the dreaded ``setoid hell''.
 

### 2020-01-26 paper.tex
nearly finished

@@
 \end{romtheorem}
+This proof does not mirror the proof of \(\Sigma\) closure on finite types,
+surprisingly.
+A different pattern is needed to pair up the members of each support stream: a
+\emph{Cantor} pairing pattern.
+Using the simple finite pairing function would diverge, not properly exploring
+one of the two lists.
+Instead, our function works in two stages.
+First, it builds a stream of non-empty lists.
+From two support streams \(\mathit{xs}\) and \(\mathit{ys}\), this intermediate
+stream \(\mathcal{CV}\) is defined like so:
+\begin{equation}
+  \mathcal{CV}_i \coloneqq \left[ (\mathit{xs}_j , \mathit{ys}_k) \mid j, k \in \mathbb{N}; j + k = i \right]
+\end{equation}
+This is then concatenated to give our result.
+
+Finally, while we have lost certain closure proofs by allowing for infinite
+types, we also \emph{gain} some: in particular the Kleene star.
 \begin{romtheorem}
@@
 \end{romtheorem}
-\subsection{Categories}
+Again, this proof requires a particular pattern to ensure non divergence.
+The pattern here is slightly more complicated, but works on the same basic idea,
+building an intermediate stream \(\mathcal{KV}\) of non-empty lists from the
+input support stream \(\mathit{xs}\).
+\begin{equation}
+  \mathcal{KV}_i \coloneqq \left[ \left[ \mathit{xs}_{j - 1} \mid j \in \mathit{js} \right] \mid \mathit{js} \in \mathbf{List}(\mathbb{N}) ; \text{sum}(\mathit{js}) = i ; 0 \notin \mathit{js}  \right]
+\end{equation}
+Again, this is subsequently flattened.
 \section{Search} \label{search}
@@
 \end{romtheorem}
-
 Omniscience is stronger than exhaustibility, as we can derive the latter from
@@
 \end{romlemma}
-\begin{proof}
-  For decidable propositions, we know the following:
-  \begin{equation} \label{lem-conversion}
-    \Pi (x : A) , P(x) \leftrightarrow \neg \Sigma (x : A) , \neg P(x) 
-  \end{equation}
-  To derive exhaustibility from omniscience, then, we run the predicate in its
-  negated form, and then subsequently negate the result.
-  The resulting decision over \(\neg \Sigma(x : A) , \neg P(x)\) can be
-  converted into \(\Pi (x : A) , P(x)\).
-\end{proof}
-We cannot derive, however, that any exhaustible type is omniscient, as we do not
-have the inverse of equation~\ref{lem-conversion}:
-\begin{equation} \label{choice-conversion}
-  \Sigma (x : A) , P(x) \leftrightarrow \neg \Pi (x : A) , \neg P(x)
-\end{equation}
-Such an equation would allow us to pick a representative element from any type,
-which is therefore non-constructive.
-In a sense, equation~\ref{lem-conversion} requires a form of LEM on the
-proposition (i.e. requires it to be decidable), whereas
-equation~\ref{choice-conversion} requires a form of choice.
-Those finiteness predicates which are ordered do in fact give us this form of
-choice, so the conversion is valid.
-As such, all of the ordered finiteness predicates imply omniscience.
-Again, we will prove it only for the weakest.
+All of the ordered finiteness predicates imply omniscience.
+We will again only prove it for the weakest.
 \begin{romtheorem}
@@
 
-The alternative is to derive the things we need from \(\mathcal{E}!\) somehow.
-As \(\AgdaDatatype{Pauli}\) is a simple finite type, the instance can be defined
-in a similar way to those in lemma~\ref{split-enum-bool-top-bot}.
+The alternative is to derive the things we need from \(\mathcal{E}!\).
 From here we can already derive decidable equality, a function which requires 16
@@
 \ExecuteMetaData[agda/Data/Pauli.tex]{not-spec}
-
-\todo[inline]{Add funding ack:
-
+\section{Related Work}
+\section*{Acknowledgement}
 This work has been supported by the Science Foundation Ireland under the
 following grant: 13/RC/2D94 to Irish Software Research Centre.
-}
 \bibliographystyle{splncs04}

### 2020-01-26 paper.tex
related work done

@@
 \section{Related Work}
+The univalent foundations program is the main basis for this work
+\cite{hottbook}.
+In particular, our formalisation in section~\ref{topos} relied heavily on
+\cite[chapter 10]{hottbook}, and \cite{rijkeSetsHomotopyType2015}, a paper which
+contains much of the same material.
+
+In \cite{fruminFiniteSetsHomotopy2018} the topic of Kuratowski finite sets in
+HoTT is studied extensively: we have focused more on the non-truncated versions
+of finiteness (the ``manifest'' predicates), and we have provided the missing
+\(\Pi\)-pretopos proof of decidable Kuratowski finite sets.
+
+\cite{iversenUnivalentCategoriesFormalization2018} provided a starting point for
+our categorical formalisation: it contains a proof, for instance, that homotopy
+sets form a category.
+
+Finite sets in a constructive setting has been studied extensively before:
+In \cite{coquandConstructivelyFinite2010} four separate predicates for
+finiteness were considered (split-enumerable being the only one explored in this
+work), and \cite{firsovVariationsNoetherianness2016} explores Noetherianness.
+\cite{firsovDependentlyTypedProgramming2015} explored what we have called split
+enumerability and manifest Bishop finiteness (although they are stated slightly
+differently), and they use these to build a library for proof search.
+
+Our formalisation is made possible by Cubical Type Theory
+\cite{cohenCubicalTypeTheory2016} and Cubical Agda
+\cite{vezzosiCubicalAgdaDependently2019}.
 \section*{Acknowledgement}

### 2020-01-26 paper.tex
more

@@
 \subsection{Closure}
-For the first two closure proofs, we only consider manifest Bishop finiteness:
+For the first three closure proofs, we only consider split enumerability:
 as it is the strongest of the finiteness predicates, we can derive the other
 closure proofs from it.
+
 \begin{romlemma}
-  \(\bot\), \(\top\), and \(\mathbf{Bool}\) are manifest Bishop finite.
-\end{romlemma}
+  \(\bot\), \(\top\), and \(\mathbf{Bool}\) are split enumerable.
+\end{romlemma}
+\begin{romlemma} \label{split-enum-sigma}
+  Split enumerability is closed under \(\Sigma\).
+  \begin{equation}
+    \frac{
+      \mathcal{E}!(A) \; \; \; \Pi(x : A) , \mathcal{E}!(U(x))
+    }{
+      \mathcal{E}!(\Sigma(x : A) , U(x))
+    }
+  \end{equation}
+\end{romlemma}
+From this we can derive split enumerability of non-dependent sums and products,
+as both are definable in terms of \(\Sigma\).
+
 \begin{romlemma}
-  Split enumerability is closed under \(\Sigma\)
-\end{romlemma}
-To show closure under \(\Sigma\), we take the Cartesian product of the support
-lists.
-
-\begin{romlemma}
-  Manifest bishop finiteness is closed over dependent functions
-  (\(\prod\)-types).
+  Split enumerability is closed under dependent functions.
+  (\(\Pi\)-types).
   \begin{equation}
     \frac{
-      \mathcal{B}(A) \; \; \; \Pi {(x : A)} , \mathcal{B}\left( U(x) \right)
+      \mathcal{E}!(A) \; \; \; \Pi {(x : A)} , \mathcal{E}!\left( U(x) \right)
     }{
-      \mathcal{B}\left(\Pi {(x : A)} , U(x)\right)
+      \mathcal{E}!\left(\Pi {(x : A)} , U(x)\right)
     }
@@
 \begin{proof}
-  Firstly, we will simplify things slightly by working only with split
-  enumerability.
-  As this is equal in strength to manifest Bishop finiteness, any closure proofs
-  carry over.
-  Secondly, we will replace \(A\) in all places with \(\mathbf{Fin}(n)\), using
-  the equivalence provided by lemma~\ref{bishop-equiv}.
-
-  The remainder of the proof now is concerned with proving the following:
-  \begin{equation} \label{given-transp-sublemma}
-    \left( \Pi {(x : \mathbf{Fin}(n))} , \mathcal{E}!(U(x)) \right) \rightarrow \mathcal{E}!\left( \Pi {(x : \mathbf{Fin}(n))} , U(x) \right)
-  \end{equation}
-
-  To prove this goal, consider the following type of \(n\)-tuples.
+  Let \(A\) be a split enumerable type, and \(U\) be a type family from \(A\),
+  which is split enumerable over all points of \(A\).
+
+  As \(A\) is split enumerable, we know that it is also manifestly Bishop finite
+  (lemma~\ref{split-enum-to-manifest-bishop}), and consequently we know \(A
+  \simeq \mathbf{Fin}(n)\), for some \(n\) (lemma~\ref{bishop-equiv}).
+  We can therefore replace all occurrences of \(A\) with \(\mathbf{Fin}(n)\),
+  changing our goal to:
+  \begin{equation}
+    \frac{
+      \mathcal{E}!(\mathbf{Fin}(n)) \; \; \; \Pi (x : \mathbf{Fin}(n)) , \mathcal{E}!\left( U(x) \right)
+    }{
+      \mathcal{E}!\left(\Pi (x : \mathbf{Fin}(n)) , U(x)\right)
+    }
+  \end{equation}
+
+  We then define the type of \(n\)-tuples over some type family \(T :
+  \mathbf{Fin}(n) \rightarrow \mathbf{Type}\).
   \begin{equation}
     \begin{alignat}{3}
-      & \mathbf{Tuple}(T, 0)   &&\coloneqq \top \\
-      & \mathbf{Tuple}(T, n+1) &&\coloneqq T(0) \times \mathbf{Tuple}(T \circ \text{suc}, n)
+      & \mathbf{Tuple}(0, T)   &&\coloneqq \top \\
+      & \mathbf{Tuple}(n+1, T) &&\coloneqq T(0) \times \mathbf{Tuple}(n, T \circ \text{suc})
     \end{alignat}
   \end{equation}
-  We have the following equivalence, proven in our formalisation.
-  \begin{equation} \label{fin-equiv-sublemma}
-    \mathbf{Tuple}(T, n) \simeq \Pi {(x : \mathbf{Fin}(n))} , T(x)
-  \end{equation}
-
-  By the given proof of finiteness in equation~\ref{given-transp-sublemma}, we
-  know that \(\mathbf{Tuple}(T , n)\) is split enumerable, since it is made of
-  products of the points of \(U\).
-  We can then transport along the equivalence in
-  equation~\ref{fin-equiv-sublemma}, proving our goal.
+  We can show that this type is equivalent to functions (proven in our formalisation):
+  \begin{equation}
+    \Pi(x : \mathbf{Fin}(n)) , U(x) \simeq \mathbf{Tuple}(n, U)
+  \end{equation}
+  And therefore we can simplify again our goal to the following:
+  \begin{equation}
+    \frac{
+      \mathcal{E}!(\mathbf{Fin}(n)) \; \; \; \Pi (x : \mathbf{Fin}(n)) , \mathcal{E}!\left( U(x) \right)
+    }{
+      \mathcal{E}!\left(\mathbf{Tuple}(n, U)\right)
+    }
+  \end{equation}
+
+  We can prove this goal by showing that \(\mathbf{Tuple}(n, U)\) is split
+  enumerable: it is made up of finitely many products of points of \(U\), which
+  are themselves split enumerable, and \(\top\), which is also split enumerable.
+  Lemma~\ref{split-enum-sigma} shows us that the product of finitely many split
+  enumerable types is itself split enumerable, proving our goal.
+  \qed
 \end{proof}

### 2020-01-27 paper.tex
shorter

@@
 finite.
-Constructively (in Martin-Löf Type Theory
+Constructively (in Martin-Löf Type Theory (MLTT)
 \cite{martin-lofIntuitionisticTypeTheory1980} at least), however, all three
@@
 
-Homotopy Type Theory \cite{hottbook} adds another wrinkle to the story.
+Homotopy Type Theory (HoTT) \cite{hottbook} adds another wrinkle to the story.
 Firstly, in HoTT we cannot assume that every type is a (homotopy) set: this
@@
 feature missing from classical mathematics: computation.
-Cubical Type Theory \cite{cohenCubicalTypeTheory2016}, and its implementation in
-Cubical Agda \cite{vezzosiCubicalAgdaDependently2019}, realise this property
-even in the presence of univalence, giving computational content to programs
-written in HoTT.
+Cubical Type Theory (CuTT) \cite{cohenCubicalTypeTheory2016}, and its
+implementation in Cubical Agda \cite{vezzosiCubicalAgdaDependently2019}, realise
+this property even in the presence of univalence, giving computational content
+to programs written in HoTT.
 \subsection{Contributions}
-In this work we will examine five notions of finiteness in Cubical Type Theory,
-the relationships between them, and their topos-theoretic characterisation.
-We also briefly examine a predicate for countable sets, comparing it to the
-finiteness predicates.
-Our work is formalised in Cubical Agda, where we also develop a library for
-proof search based on the finiteness predicates.
-
-\input{figures/finite-classification}
-
 The finiteness predicates we are interested in are organised in
@@
 predicates, and its topos-theoretic classification.
+
+\input{figures/finite-classification}
 
@@
 omniscience and exhaustibility.
-While proof search based on finiteness is not new, implementing it here does
-give us an opportunity to demonstrate some actual \emph{computation} in a
-univalent setting.
-Many of the finiteness predicates are built using univalence, so for the proof
-search functionality to work at all we require computational content in the
-univalence axiom.
-Furthermore, the extensionality afforded to us by HoTT means that we can have
-\(\Pi\) types in the domain of the search, making our library more powerful than
-comparable libraries.
-We will demonstrate some interesting uses for this feature in particular.
+This library relies directly on the computational content of univalence, which
+allows us to (for instance) have \(\Pi\) types in the domain of the search
+space.
 \subsection{Notation and Background}
@@
     ``Type families'' are functions into \(\mathbf{Type}\).
-  \item[Universes]
-    To avoid Girard's paradox
-    \cite{girardInterpretationFonctionelleElimination1972} we use explicit
-    universe levels.
-    A full explanation is beyond the scope of this work, but it means that there
-    is a hierarchy of types starting with \(\mathbf{Type}\).
-    \begin{equation}
-      \begin{alignat}{3}
-        & 2               &&: \mathbb{N} \\
-        & \mathbb{N}      &&: \mathbf{Type} \\
-        & \mathbf{Type}   &&: \mathbf{Type}_1 \\
-        & \mathbf{Type}_1 &&: \mathbf{Type}_2
-      \end{alignat}
-    \end{equation}
-    In our formalisation, every proof is done in the most universe polymorphic
-    way possible: the proof that the cardinally finite sets form a
-    \(\Pi\)-pretopos, for instance, is defined over any universe level.
   \item[0 , 1 , 2] We call the \(\mathbf{0}\), \(\mathbf{1}\), and
@@
     \(\Pi\) as \(\rightarrow\).
-  \item[Disjoint Union] Disjoint union can be defined in terms of \(\Sigma\):
-    \begin{equation}
-      A \uplus B \coloneqq \Sigma(x : \mathbf{Bool}) , \text{if } x \text{ then } A \text{ else } B
-    \end{equation}
-    However, we prefer to use it as an inductively defined type (though the two
-    are equivalent).
+  \item[Disjoint Union] We define disjoint union as an inductive type.
     \begin{equation}
@@
     \end{equation}
+    It is also expressible with only \(\Sigma\):
+    \(A \uplus B \simeq \Sigma(x : \mathbf{Bool}) , \text{if } x \text{ then
+    } A \text{ else } B \).
   \item[Equalities, equivalences, and paths] We use the symbol \(\coloneqq\)
@@
 It allows us to do much of the same theory as in HoTT, but crucially the
-univalence ``axiom'' is a \emph{theorem}, rather than an axiom.
-This allows us to actually compute with univalent proofs, a capability missing
-from HoTT.
+univalence ``axiom'' is a \emph{theorem}, giving it computational content.
 \begin{romdefinition}[Path Types] \label{path-types}
@@
   Types in HoTT and CuTT are not necessarily sets, as they are in MLTT.
-  This means that equalities are not necessarily unique.
-  In fact, we have an entire hierarchy of homotopy levels, starting with
-  contractions, of which sets are level 2.
+  Some have higher homotopies (paths which aren't unique).
+  We actually have a hierarchy of complexity of structure of path spaces in
+  types, starting with the contractions \cite[definition 3.11.1]{hottbook}, then
+  the mere propositions \cite[definition 3.3.1]{hottbook}, and the sets
+  \cite[definition 3.1.1]{hottbook}.
   \begin{alignat}{2}
@@
     &\text{isProp}(A)     &&\coloneqq \Pi(x, y : A) , (x \equiv y) \\
-    &\text{isSet}(A)      &&\coloneqq \Pi(x, y : A) , \text{isProp}(x \equiv y) \\
-    &\text{isGroupoid}(A) &&\coloneqq \Pi(x, y : A) , \text{isSet}(x \equiv y)
+    &\text{isSet}(A)      &&\coloneqq \Pi(x, y : A) , \text{isProp}(x \equiv y)
   \end{alignat}
-  We can define the above types inductively like so:
-  \begin{alignat}{2}
-    &\text{isOfHLevel}(0 , A)   &&\coloneqq \text{isContr}(A) \\
-    &\text{isOfHLevel}(n+1 , A) &&\coloneqq \Pi(x , y : A) , \text{isOfHLevel}(n , x \equiv y)
-  \end{alignat}
-  This type is defined similarly in \cite[definition 7.1.1]{hottbook}, although
-  there the numbering starts at \(-2\).
 \end{romdefinition}
@@
 \end{romdefinition}
-\begin{romlemma}
-  Univalence
-  \begin{equation}
-    (A \simeq B) \simeq (A \equiv B)
-  \end{equation}
-  This is the so-called univalence ``axiom'', which is a theorem in CuTT.
-\end{romlemma}
-\begin{romdefinition}[Decision]
+\begin{romdefinition}[Decidable Types]
   \begin{equation}
@@
 \end{romdefinition}
-\begin{romdefinition}[Discrete]
+\begin{romdefinition}[Discrete Types]
   A discrete type is one with decidable equality.
@@
   \end{equation}
-  We will use three eliminators from \(\lVert A \rVert\) in this paper.
+  We will use two eliminators from \(\lVert A \rVert\) in this paper.
   \begin{enumerate}
@@
     \(\text{isProp}(B)\), we have a function \(\lVert A \rVert \rightarrow B\).
-  \item \label{elim-prop-monad} A special case of \ref{elim-prop-prop} implies
-    that \(\lVert \cdot \rVert\) forms a monad: this means that we get the usual
-    Monadic operators on \(\lVert \cdot \rVert\) (bind, pure, fmap, etc.).
   \item \label{elim-prop-coh} We can eliminate from \(\lVert A \rVert\) with a
@@
 in Figure~\ref{finite-classification}.
-The reason we explore predicates other than our focus (cardinal finiteness) is
-that we can often prove things like closure much more readily on the simpler
-predicates.
+The reason we explore predicates other than our focus (decidable Kuratowski
+finiteness) is that we can often prove things like closure much more readily on
+the simpler predicates.
 The relations (which we will prove in the next section) then allow us to
-transfer those proofs onto Kuratowski finiteness.
+transfer those proofs.
 \subsection{Split Enumerability}
@@
 We use the word ``manifest'' here to distinguish from another common
-interpretation of Bishop finiteness, which we have called Cardinal finiteness in
+interpretation of Bishop finiteness, which we have called cardinal finiteness in
 this paper.
@@
 \end{romlemma}
-\begin{proof}
-  \begin{align*}
-    \mathcal{B}(A)
-    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , x \inunique \mathit{xs}
-    && \text{def.~\ref{bish-def} } (\mathcal{B})
-    \\
-    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{isContr} \left(\text{fib}_{\text{snd}(\mathit{xs})}(x)\right)
-    && \text{def.~\ref{uniq-memb-def} } (\inunique)
-    \\
-    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
-    && \text{eqn.~\ref{is-equiv-def} (isEquiv)}
-    \\
-    & \simeq \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket (A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
-    && \text{def.~\ref{List} } (\mathbf{List})
-    \\
-    & \simeq \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \mathbf{Fin}(n)) , A) , \text{isEquiv}(\text{snd}(\mathit{xs}))
-    && \text{eqn.~\ref{container-interp} } (\llbracket \wc \rrbracket)
-    \\
-    & \simeq \Sigma (n : \mathbb{N}) , \Sigma (f : \mathbf{Fin}(n) \rightarrow A) , \text{isEquiv}(f)
-    && \text{Reassociation of } \Sigma
-    \\
-    & \simeq \Sigma (n : \mathbb{N}) , ( \mathbf{Fin}(n) \simeq \; A )
-    && \text{eqn.~\ref{surj-arrow-eqn} } (\twoheadrightarrow)
-  \end{align*}
-  \qed
-\end{proof}
+This proof is effectively the same as that of
+lemma~\ref{split-enum-is-split-surj}.
 \subsection{Cardinal Finiteness}
@@
 \end{romlemma}
+The proof for this lemma is similar in structure to
+lemma~\ref{split-enum-is-split-surj} and lemma~\ref{bishop-equiv}.
 
@@
 \subsection{Kuratowski Finiteness}
-Much work has already been done on Kuratowski finiteness in HoTT in
-\cite{fruminFiniteSetsHomotopy2018}.
-As a result, we will not needlessly repeat proofs, rather we will just give a
-brief introduction to the topic and point out where our treatment differs.
-
 The first thing we must define is a representation of subsets.
@@
 \end{romlemma}
-The second of these in particular tells us that the ``Kuratowski-finite types''
-are not necessarily sets; it also tells us that we cannot derive decidable
-equality from a proof of Kuratowski finiteness.
 \section{Relations Between Each Finiteness Definition} \label{relations}
@@
 it turns out this is not the case.
-Both types imply the other.
+Both predicates imply the other.
 \begin{romlemma} \label{manifest-bishop-to-split-enum}
@@
 
-To derive \(\mathcal{B}\) from \(\mathcal{E}!\) takes significantly more work.
+\begin{romlemma} \label{split-enum-to-manifest-bishop}
+  Any split enumerable set is manifest Bishop finite.
+\end{romlemma}
+This proof takes significantly more work.
 The ``unique membership'' condition in \(\mathcal{B}\) means that we are not
@@
 appropriately.
-\begin{romlemma} \label{split-enum-to-manifest-bishop}
-  Any split enumerable set is manifest Bishop finite.
-\end{romlemma}
 \subsection{Split Enumerability and Manifest Enumerability}
@@
 \end{romlemma}
-\begin{romlemma} \label{dec-trunc-refute}
-  For any type \(A\), we can extract a value of type \(A\) from a decision over
-  \(A\) in combination with a propositionally-truncated value of type \(\lVert A \rVert\).
-  \begin{equation}
-    \frac{
-      \mathbf{Dec}(A) \; \; \; \lVert A \rVert
-    }{
-      A
-    }
-  \end{equation}
-\end{romlemma}
-\begin{proof}
-  Let \(d\) be a value of type \(\mathbf{Dec}(A)\), and \(p\) be a value of type
-  \(\lVert A \rVert\).
-  We proceed by case analysis on \(d\).
-  In the case where \(d\) is inl, we have a value of type \(A\), and our goal is
-  satisfied.
-  In the case of inr, we have a value of type \(\neg A\).
-  Using eliminator \ref{elim-prop-prop} from definition~\ref{prop-trunc}, and
-  the fact that \(\bot\) is a proposition, we can run the \(\neg A\) proof on
-  \(p\), getting a contradiction.
-  \qed
-\end{proof}
+This lemma is proven by truncating the membership proof in split enumerability.
+
 \begin{romlemma} \label{manifest-enum-to-split-enum}
@@
 \end{romlemma}
-\begin{proof}
-  Similarly to the proof of lemma~\ref{manifest-bishop-to-split-enum}, our
-  obligation for this proof is to provide a conversion function between
-  membership proofs.
-  In this case, we need to show:
-  \begin{equation}
-    \Pi(x : A) , \Pi(\mathit{xs} : \mathbf{List}(A)) , \lVert x \in \mathit{xs} \rVert \rightarrow x \in \mathit{xs}
-  \end{equation}
-  This conversion function can be constructed from the two following functions,
-  provided in the formalisation.
-  First, for any type with decidable equality, decidable membership in a finite
-  list is consequently decidable:
-  \begin{equation}
-    \text{Discrete}(A) \rightarrow \Pi(x : A) , \Pi(\mathit{xs} : \mathbf{List}(A)) , \mathbf{Dec}(x \in \mathit{xs})
-  \end{equation}
-  Then, we use lemma~\ref{dec-trunc-refute}, allowing us to derive \(x \in
-  \mathit{xs}\) from \(\lVert x \in \mathit{xs} \rVert\).
-  \qed
-\end{proof}
+
 \subsection{Manifest Bishop Finiteness and Cardinal Finiteness}
@@
 
-The dependent cases of each of these proofs do not transfer simply to proofs on
-\(\mathcal{C}\).
-In order to transfer them over, we need to prove a variant of the axiom of
-choice on finite sets.
+Lifting the above closure proofs to work on proofs of \(\mathcal{C}\) is not
+necessarily straightforward.
+\(\lVert \wc \rVert\) forms a monad, giving us access to a powerful syntax for
+combining proofs.
+From this, we can derive closure under \(\bot\), \(\top\), \(\mathbf{Bool}\),
+\(\times\), \(\uplus\), and \(\rightarrow\).
+The dependent case is not so straightforward.
+We need the following lemma, a version of the axiom of choice on finite sets,
+prove closure under \(\Sigma\) and \(\Pi\).
 \begin{romlemma}
@@
 \end{romlemma}
-This is proven in our formalisation.
 \subsection{The Category of Finite Sets}
@@
 enumerability and cardinal finiteness.
-The change will be a simple one: in terms of surjections and bijections, we will
-swap out \(\mathbf{Fin}\) for \(\mathbb{N}\).
-In terms of support lists and cover proofs, we will swap out lists for streams.
-A stream can be defined many ways in type theory: we take rather a low-tech
-approach here, saying that a stream is a function from the natural numbers.
-\begin{romdefinition}[Stream]
-  \begin{equation}
-    \mathbf{Stream}(A) \coloneqq \mathbb{N} \rightarrow A
-  \end{equation}
-  Streams are also definable as containers:
-  \begin{equation}
-    \mathbf{Stream}(A) \simeq \llbracket \top , \text{const}(\mathbb{N}) \rrbracket
-  \end{equation}
-  However the two definitions are clearly equivalent, and furthermore it is
-  simpler to work with the first.
-\end{romdefinition}
-
-From this, we can define ``split countability''.
+The change will be a simple one: we will swap out lists for streams.
+\begin{romdefinition}[Streams]
+  \begin{equation}
+    \mathbf{Stream}(A) \coloneqq (\mathbb{N} \rightarrow A)
+    \simeq \llbracket \top , \text{const}(\mathbb{N}) \rrbracket
+  \end{equation}
+\end{romdefinition}
 \begin{romdefinition}[Split Countability]
@@
 surprisingly.
-A different pattern is needed to pair up the members of each support stream: a
-\emph{Cantor} pairing pattern.
-Using the simple finite pairing function would diverge, not properly exploring
-one of the two lists.
-Instead, our function works in two stages.
-First, it builds a stream of non-empty lists.
-From two support streams \(\mathit{xs}\) and \(\mathit{ys}\), this intermediate
-stream \(\mathcal{CV}\) is defined like so:
-\begin{equation}
-  \mathcal{CV}_i \coloneqq \left[ (\mathit{xs}_j , \mathit{ys}_k) \mid j, k \in \mathbb{N}; j + k = i \right]
-\end{equation}
-This is then concatenated to give our result.
+The pattern we used to pair up the relevant support lists for the finite types
+would diverge for infinite types: instead, we need a \emph{pairing} function.
 
@@
 \end{romtheorem}
-Again, this proof requires a particular pattern to ensure non divergence.
-The pattern here is slightly more complicated, but works on the same basic idea,
-building an intermediate stream \(\mathcal{KV}\) of non-empty lists from the
-input support stream \(\mathit{xs}\).
+Again, this proof requires a particular pattern to ensure productivity.
+The pattern here builds an intermediate stream \(\mathcal{KV}\) of non-empty
+lists from the input support stream \(\mathit{xs}\), which is subsequently
+flattened.
 \begin{equation}
@@
 \end{equation}
-Again, this is subsequently flattened.
 \section{Search} \label{search}
 \subsection{Omniscience}
-In this section we are interested in restricted forms of the limited principle
-of omniscience \cite{myhillErrettBishopFoundations1972}.
 \begin{romdefinition}[Limited Principle of Omniscience]
   For any type \(A\) and predicate \(P\) on \(A\), the limited principle of
-  omniscience is as follows:
+  omniscience \cite{myhillErrettBishopFoundations1972} is as follows:
   \begin{equation}
@@
 
-All of the finiteness predicates we have seen justify exhaustibility.
-We will only prove it once, then, for the weakest:
-\begin{romtheorem}
-  Kuratowski-finite types are exhaustible.
-\end{romtheorem}
+All of the finiteness predicates we have seen imply exhaustibility.
 Omniscience is stronger than exhaustibility, as we can derive the latter from
-the former:
-\begin{romlemma} \label{omniscient-is-exhaustible}
-  Any omniscient type is exhaustible.
-\end{romlemma}
+the former.
 All of the ordered finiteness predicates imply omniscience.
-We will again only prove it for the weakest.
-\begin{romtheorem}
-  Manifest enumerable types are omniscient.
-\end{romtheorem}
-Finally, we do have a form of omniscience for prop-valued predicates, as they do
-not care about the chosen representative.
-\begin{romtheorem}
-  Kuratowski finite types are omniscient about prop-valued predicates.
-\end{romtheorem}
-\subsection{Synthesising Pattern-Matching Proofs}
-In particular, they can automate large proofs by analysing every possible case.
-In \cite{firsovDependentlyTypedProgramming2015}, the \(\AgdaDatatype{Pauli}\)
-group is used as an example.
+For the unordered finiteness definitions, we have omniscience for prop-valued
+predicates.
+\subsection{Automating Proofs}
+One use for above constructions is the automation of certain proofs.
+In \cite{firsovDependentlyTypedProgramming2015}, which uses a similar approach
+to ours, the \(\AgdaDatatype{Pauli}\) group is used as an example.
 \ExecuteMetaData[agda/Data/Pauli.tex]{def}
@@
 
-The alternative is to derive the things we need from \(\mathcal{E}!\).
-From here we can already derive decidable equality, a function which requires 16
-cases if implemented manually.
-
+The alternative is to derive the things we need from omniscience, itself derived
+from a finiteness predicate.
 For proof search, the procedure is a well-known one in Agda
@@
 typechecking.
-\subsection{Multiple Arguments}
-The automation machinery above only deals with single-argument predicates.
-This is not a problem, as we know that we can work with multiple arguments by
-currying and uncurrying, since all of the finiteness predicates are closed under
-\(\times\).
-To automate away the curry/uncurry noise we will use instance search, building
-on \cite{allaisGenericLevelPolymorphic2019} to develop a small interface to
-generic \(n\)-ary functions and properties.
-Our generic representation can handle dependent \(\Sigma\) and \(\prod\)
-types (rather than their non-dependent counterparts, \(\times\) and
-\(\rightarrow\)).
-This extension was necessary for our use case: it is mentioned in the paper
-as the obvious next step. 
-We also implement the curry-uncurry combinators as (verified) isomorphisms.
-
-A full explanation of our implementation is beyond the scope of this work, so we
-only present the finished interface, which is used like so:
+Our addition to this technique is a way to handle multiple arguments based on
+fully level-polymorphic dependent currying and uncurrying, building on 
+\cite{allaisGenericLevelPolymorphic2019}.
 \ExecuteMetaData[agda/Data/Pauli.tex]{assoc-prf}
-\subsection{Synthesising Functions}
+
 Finally, thanks to extensionality provided to us by HoTT, and the computation
@@
 
-In \cite{fruminFiniteSetsHomotopy2018} the topic of Kuratowski finite sets in
-HoTT is studied extensively: we have focused more on the non-truncated versions
-of finiteness (the ``manifest'' predicates), and we have provided the missing
-\(\Pi\)-pretopos proof of decidable Kuratowski finite sets.
-
-\cite{iversenUnivalentCategoriesFormalization2018} provided a starting point for
-our categorical formalisation: it contains a proof, for instance, that homotopy
-sets form a category.
-
 Finite sets in a constructive setting has been studied extensively before:
@@
 differently), and they use these to build a library for proof search.
-
-Our formalisation is made possible by Cubical Type Theory
-\cite{cohenCubicalTypeTheory2016} and Cubical Agda
-\cite{vezzosiCubicalAgdaDependently2019}.
+In \cite{fruminFiniteSetsHomotopy2018} the topic of Kuratowski finite sets in
+HoTT is studied extensively: we have focused more on the non-truncated versions
+of finiteness (the ``manifest'' predicates), and we have provided the missing
+\(\Pi\)-pretopos proof of decidable Kuratowski finite sets.
+
+\cite{iversenUnivalentCategoriesFormalization2018} provided a starting point for
+our categorical formalisation: it contains a proof, for instance, that homotopy
+sets form a category.
+
 \section*{Acknowledgement}

### 2020-01-27 paper.tex
fits on single page

@@
 
-Finally, thanks to extensionality provided to us by HoTT, and the computation
-properties provided by CuTT, we can derive decidable equality on functions over
-finite types.
+Finally, we can derive decidable equality on functions over finite types.
 We can also use functions in our proof search.
@@
 
-\section*{Acknowledgement}
+\subsubsection*{Acknowledgement}
 This work has been supported by the Science Foundation Ireland under the

### 2020-03-11 paper.tex
link to formalisation
more rebuttal

@@
 space.
+
+The full code of the formalisation is available at
+\url{https://github.com/oisdk/finiteness-in-cubical-type-theory}.
 \subsection{Notation and Background}
@@
   \begin{equation}
-    \text{fib}_f(y) = \Sigma(x : A) , (f (x) \equiv y)
+    \text{fib}_f(y) \coloneqq \Sigma(x : A) , (f (x) \equiv y)
   \end{equation}
@@
 We need the following lemma, a version of the axiom of choice on finite sets,
-prove closure under \(\Sigma\) and \(\Pi\).
+to prove closure under \(\Sigma\) and \(\Pi\).
 \begin{romlemma}
@@
 \begin{alignat}{3}
-  &\text{Obj}_\mathit{Set}      &&\coloneqq \Sigma(x : \mathbf{Type}) , \text{isProp}(x) \\
+  &\text{Obj}_\mathit{Set}      &&\coloneqq \Sigma(x : \mathbf{Type}) , \text{isSet}(x) \\
   &\text{Hom}_\mathit{Set}(x , y) &&\coloneqq  \text{fst}(x) \rightarrow \text{fst}(y)
@@
 \begin{romtheorem} \label{finite-topos}
-  The category of finite sets, \(\mathit{FinSet}\), forms a \(\Pi\)-topos.
+  The category of finite sets, \(\mathit{FinSet}\), forms a \(\Pi\)-pretopos.
 \end{romtheorem}
@@
 can themselves satisfy omniscience.
-In particular, \emph{finite} types are omniscient.
+In particular, cardinal finite types are omniscient.
 
@@
 We can also use functions in our proof search.
-Here, for instance, is a automated procedure which finds the
+Here, for instance, is an automated procedure which finds the
 \(\AgdaFunction{not}\) function on \(\mathbf{Bool}\), given a specification.

### 2020-03-11 paper.tex
mostly addressed

@@
   \begin{equation}
-    \mathcal{C}(A) \simeq \mathcal{K}(A) \times \text{Discrete}(A)
+    \mathcal{C}(A) \simeq \mathcal{K}^f(A) \times \text{Discrete}(A)
   \end{equation}
@@
 This proof is constructed by providing a pair of functions: one from
-\(\mathcal{C}(A)\) to \(\mathcal{K}(A) \times \text{Discrete}(A)\), and one the
+\(\mathcal{C}(A)\) to \(\mathcal{K}^f(A) \times \text{Discrete}(A)\), and one the
 other way.
@@
 \end{romtheorem}
-This proof does not mirror the proof of \(\Sigma\) closure on finite types,
-surprisingly.
+This proof does not mirror the proof of \(\Sigma\) closure on finite types.
 The pattern we used to pair up the relevant support lists for the finite types

### 2020-05-07 paper.tex
updated to acm format

@@
 \input{preamble}
+\acmConference[TyDe 2020]{Workshop on Type-driven Development}{23 Aug
+  2020}{Jersey City, USA}
+
+\author{Donnacha Oisín Kidney}
+\affiliation{University College Cork}
+\email{o.kidney@cs.ucc.ie}
+\author{Gregory Provan}
+\affiliation{University College Cork}
+\email{g.provan@cs.ucc.ie}
+\author{Nicolas Wu}
+\affiliation{Imperial College London}
+\email{n.wu@imperial.ac.uk}
+
 \title{Finiteness in Cubical Type Theory}
-\author{
-  Donnacha Oisín Kidney\inst{1}\and Gregory Provan \inst{2}
-}
-\institute{
-  University College Cork \email{o.kidney@cs.ucc.ie} \and
-  University College Cork \email{g.provan@cs.ucc.ie}
-}
-\begin{document}
-\maketitle
+
+\keywords{Agda, Homotopy Type Theory, Cubical Type Theory,
+  Dependent Types, Finiteness, Topos, Kuratowski finite}
+
 \begin{abstract}
@@
   functions.
-  \keywords{Agda \and Homotopy Type Theory \and Cubical Type Theory \and
-    Dependent Types \and Finiteness \and Topos \and Kuratowski finite}
 \end{abstract}
+
+\begin{document}
+\maketitle
 \section{Introduction}
@@
 sets form a category.
-
-\subsubsection*{Acknowledgement}
+\begin{acks}
   This work has been supported by the Science Foundation Ireland under the
   following grant: 13/RC/2D94 to Irish Software Research Centre.
-\bibliographystyle{splncs04}
-\bibliography{bibliography.bib}
+\end{acks}
+\printbibliography
 \end{document}

### 2020-05-07 paper.tex
removed non-italic envs (they're already not italic in acm format)

@@
 univalence ``axiom'' is a \emph{theorem}, giving it computational content.
-\begin{romdefinition}[Path Types] \label{path-types}
+\begin{definition}[Path Types] \label{path-types}
   The equality type (which we denote with \(\equiv\)) in CuTT is the type of
@@
   substitution and congruence.
-\end{romdefinition}
+\end{definition}
 
@@
 }
-\begin{romdefinition}[Homotopy Levels] \label{homotopy-types}
+\begin{definition}[Homotopy Levels] \label{homotopy-types}
   Types in HoTT and CuTT are not necessarily sets, as they are in MLTT.
@@
   \end{alignat}
-\end{romdefinition}
-\begin{romdefinition}[Fibres] \label{fibres}
+\end{definition}
+\begin{definition}[Fibres] \label{fibres}
   A fibre \cite[definition 4.2.4]{hottbook} is defined over some function \(f :
@@
   \end{equation}
-\end{romdefinition}
-\begin{romdefinition}[Equivalences] \label{equivalences}
+\end{definition}
+\begin{definition}[Equivalences] \label{equivalences}
   We will take contractible maps \cite[definition 4.4.1]{hottbook} as our
@@
   \end{alignat}
-\end{romdefinition}
-\begin{romdefinition}[Decidable Types]
+\end{definition}
+\begin{definition}[Decidable Types]
   \begin{equation}
@@
   \end{equation}
-\end{romdefinition}
-\begin{romdefinition}[Discrete Types]
+\end{definition}
+\begin{definition}[Discrete Types]
   A discrete type is one with decidable equality.
@@
   type is a set.
-\end{romdefinition}
-\begin{romdefinition}[Higher Inductive Types] \label{HITs}
+\end{definition}
+\begin{definition}[Higher Inductive Types] \label{HITs}
   Normal inductive types have \emph{point} constructors: constructors which
@@
   construct paths in the type.
-\end{romdefinition}
-\begin{romdefinition}[Propositional Truncation] \label{prop-trunc}
+\end{definition}
+\begin{definition}[Propositional Truncation] \label{prop-trunc}
   The type \(\lVert A \rVert\) on some type \(A\) is a propositionally truncated
@@
   \end{enumerate}
-\end{romdefinition}
+\end{definition}
 \section{Finiteness Predicates} \label{finiteness-predicates}
@@
 \subsection{Split Enumerability}
-\begin{romdefinition}[Split Enumerable Set] \label{split-enum-def}
+\begin{definition}[Split Enumerable Set] \label{split-enum-def}
   \begin{equation} \label{split-enum-def-eqn}
@@
   \cite{firsovDependentlyTypedProgramming2015}.
-\end{romdefinition}
+\end{definition}
 
 We used some extra types in the above definition, which we will define here:
-\begin{romdefinition}[Containers] \label{container-def}
+\begin{definition}[Containers] \label{container-def}
   A container \cite{abbottContainersConstructingStrictly2005} is a pair
@@
   \end{equation}
-\end{romdefinition}
-\begin{romdefinition}[\(\mathbf{List}\)] \label{List}
+\end{definition}
+\begin{definition}[\(\mathbf{List}\)] \label{List}
   \begin{equation}
@@
   \end{equation}
-\end{romdefinition}
-\begin{romdefinition}[\(\mathbf{Fin}\)] \label{Fin}
+\end{definition}
+\begin{definition}[\(\mathbf{Fin}\)] \label{Fin}
   \(\mathbf{Fin}(n)\) is the type of natural numbers smaller than \(n\). We
@@
   \(\mathbf{Fin}(n + 1) &\coloneqq \top \uplus \mathbf{Fin}(n)\).
-\end{romdefinition}
+\end{definition}
 
@@
 numbers.
-\begin{romdefinition}[Surjections] \label{surjections}
+\begin{definition}[Surjections] \label{surjections}
   We define both surjections and \emph{split} surjections here \cite[definition
@@
   \end{alignat}
-\end{romdefinition}
-\begin{romlemma} \label{split-enum-is-split-surj}
+\end{definition}
+\begin{lemma} \label{split-enum-is-split-surj}
   \begin{equation}
@@
   \end{equation}
-\end{romlemma}
+\end{lemma}
 \begin{proof}
@@
 formalisation:
-\begin{romlemma} \label{discrete-surj}
+\begin{lemma} \label{discrete-surj}
   \begin{equation}
@@
   \end{equation}
-\end{romlemma}
-\begin{romlemma} \label{split-enum-discrete}
+\end{lemma}
+\begin{lemma} \label{split-enum-discrete}
   Every split enumerable type is discrete.
-\end{romlemma}
+\end{lemma}
 \begin{proof}
@@
 \subsection{Manifest Bishop Finiteness}
-\begin{romdefinition}[Manifest Bishop Finiteness]
+\begin{definition}[Manifest Bishop Finiteness]
   \begin{equation} \label{bish-def}
@@
   \end{equation}
-\end{romdefinition}
+\end{definition}
 The only difference between manifest Bishop finiteness and split enumerability
@@
 rather than simple membership (\(\in\)).
-\begin{romdefinition}[Unique Membership] \label{uniq-memb-def}
+\begin{definition}[Unique Membership] \label{uniq-memb-def}
   \begin{equation}
@@
   \end{equation}
-\end{romdefinition}
+\end{definition}
 
@@
 \emph{equivalence} with \(\mathbf{Fin}\).
-\begin{romlemma} \label{bishop-equiv}
+\begin{lemma} \label{bishop-equiv}
   \begin{equation}
@@
   \end{equation}
-\end{romlemma}
+\end{lemma}
 This proof is effectively the same as that of
@@
 (def.~\ref{prop-trunc}).
-\begin{romdefinition}[Cardinal Finiteness]
+\begin{definition}[Cardinal Finiteness]
   \begin{equation}
@@
   \end{equation}
-\end{romdefinition}
+\end{definition}
 At first glance, it might seem that we lose any useful properties we could
@@
 Our contribution for this section is simply the formalisation.
-\begin{romlemma}
+\begin{lemma}
   Given a cardinally finite type, we can derive the type's cardinality, as well
@@
   \end{equation}
-\end{romlemma}
-\begin{romlemma} \label{cardinal-finite-discrete}
+\end{lemma}
+\begin{lemma} \label{cardinal-finite-discrete}
   Any cardinal-finite set has decidable equality.
-\end{romlemma}
+\end{lemma}
 \subsection{Manifest Enumerability}
-\begin{romdefinition}[Manifest Enumerability]
+\begin{definition}[Manifest Enumerability]
   \begin{equation}
@@
   \end{equation}
-\end{romdefinition}
+\end{definition}
 As with manifest Bishop finiteness, the only difference with this type and split
@@
 split surjection) from \(\mathbf{Fin}\).
-\begin{romlemma}
+\begin{lemma}
   \begin{equation}
@@
   \end{equation}
-\end{romlemma}
+\end{lemma}
 The proof for this lemma is similar in structure to
@@
 like the circle.
-\begin{romdefinition}[\(S^1\)] \label{circle-def}
+\begin{definition}[\(S^1\)] \label{circle-def}
   The circle, \(S^1\), can be represented in HoTT as a higher inductive type.
@@
   This also means that it does not have decidable equality.
-\end{romdefinition}
-\begin{romlemma}
+\end{definition}
+\begin{lemma}
   The circle \(S^1\) is manifestly enumerable.
-\end{romlemma}
+\end{lemma}
 \subsection{Kuratowski Finiteness}
 The first thing we must define is a representation of subsets.
-\begin{romdefinition}[Kuratowski Finite Subset]
+\begin{definition}[Kuratowski Finite Subset]
   \(\mathcal{K}(A)\) is the type of Kuratowski-finite subsets of \(A\).
@@
   The final constructor enforces that the type \(\mathcal{K}(A)\) must be a set.
-\end{romdefinition}
+\end{definition}
 The Kuratowski finite subset is a free join semilattice (or, equivalently, a
@@
 For that, we will need to define membership of \(\mathcal{K}\).
-\begin{romdefinition}[Membership of \(\mathcal{K}\)]
+\begin{definition}[Membership of \(\mathcal{K}\)]
   Membership is defined by pattern-matching on \(\mathcal{K}\).
@@
   constructor.
-\end{romdefinition}
+\end{definition}
 Finally, we have enough background to define Kuratowski finiteness.
-\begin{romdefinition}[Kuratowski Finiteness]
+\begin{definition}[Kuratowski Finiteness]
   \begin{equation}
@@
 \cite{fruminFiniteSetsHomotopy2018} and our formalisation.
-\begin{romlemma}
+\begin{lemma}
   \(\mathcal{K}^f\) is a mere proposition.
-\end{romlemma}
-\begin{romlemma}
+\end{lemma}
+\begin{lemma}
   This circle \(S^1\) is Kuratowski finite.
-\end{romlemma}
+\end{lemma}
 \section{Relations Between Each Finiteness Definition} \label{relations}
@@
 Both predicates imply the other.
-\begin{romlemma} \label{manifest-bishop-to-split-enum}
+\begin{lemma} \label{manifest-bishop-to-split-enum}
   Any manifest Bishop finite type is split enumerable.
-\end{romlemma}
+\end{lemma}
 \begin{proof}
@@
 
-\begin{romlemma} \label{split-enum-to-manifest-bishop}
+\begin{lemma} \label{split-enum-to-manifest-bishop}
   Any split enumerable set is manifest Bishop finite.
-\end{romlemma}
+\end{lemma}
 This proof takes significantly more work.
@@
 \subsection{Split Enumerability and Manifest Enumerability}
-\begin{romlemma} \label{split-enum-to-manifest-enum}
+\begin{lemma} \label{split-enum-to-manifest-enum}
   Any split enumerable type is manifestly enumerable.
-\end{romlemma}
+\end{lemma}
 This lemma is proven by truncating the membership proof in split enumerability.
 
-\begin{romlemma} \label{manifest-enum-to-split-enum}
+\begin{lemma} \label{manifest-enum-to-split-enum}
   A manifestly enumerable type with decidable equality is split enumerable.
-\end{romlemma}
+\end{lemma}
 
 \subsection{Manifest Bishop Finiteness and Cardinal Finiteness}
-\begin{romlemma} \label{manifest-bishop-to-cardinal}
+\begin{lemma} \label{manifest-bishop-to-cardinal}
   Any manifest Bishop finite type is cardinal finite.
-\end{romlemma}
-\begin{romtheorem} \label{cardinal-to-manifest-bishop}
+\end{lemma}
+\begin{theorem} \label{cardinal-to-manifest-bishop}
   Any cardinal finite type with a total order is Bishop finite.
-\end{romtheorem}
+\end{theorem}
 The proof for this particular theorem is quite involved in the formalisation, so
@@
 Next, we define permutations.
-\begin{romdefinition}[List Permutations]
+\begin{definition}[List Permutations]
   Two lists are permutations of each other if their membership proofs are all
@@
 \subsection{Cardinal Finiteness and Kuratowski Finiteness}
-\begin{romlemma} \label{cardinal-kuratowski}
+\begin{lemma} \label{cardinal-kuratowski}
   \begin{equation}
@@
   \end{equation}
-\end{romlemma}
+\end{lemma}
 This proof is constructed by providing a pair of functions: one from
@@
 
-\begin{romlemma}
+\begin{lemma}
   \(\bot\), \(\top\), and \(\mathbf{Bool}\) are split enumerable.
-\end{romlemma}
-\begin{romlemma} \label{split-enum-sigma}
+\end{lemma}
+\begin{lemma} \label{split-enum-sigma}
   Split enumerability is closed under \(\Sigma\).
@@
   \end{equation}
-\end{romlemma}
+\end{lemma}
 From this we can derive split enumerability of non-dependent sums and products,
@@
 
-\begin{romlemma}
+\begin{lemma}
   Split enumerability is closed under dependent functions.
@@
 to prove closure under \(\Sigma\) and \(\Pi\).
-\begin{romlemma}
+\begin{lemma}
   \begin{equation}
@@
   \end{equation}
-\end{romlemma}
+\end{lemma}
 \subsection{The Category of Finite Sets}
@@
 category, and the following:
-\begin{romtheorem}[The Category of Finite Sets]
+\begin{theorem}[The Category of Finite Sets]
   Finite sets form a category in HoTT when defined like so:
@@
   \end{equation}
-\end{romtheorem}
+\end{theorem}
 \subsection{The \(\Pi\)-pretopos of Finite Sets}
@@
 As far as we know, this is the first formalisation of either.
-\begin{romtheorem} \label{finite-topos}
+\begin{theorem} \label{finite-topos}
   The category of finite sets, \(\mathit{FinSet}\), forms a \(\Pi\)-pretopos.
-\end{romtheorem}
+\end{theorem}
 \section{Countably Infinite Types} \label{infinite}
@@
 The change will be a simple one: we will swap out lists for streams.
-\begin{romdefinition}[Streams]
+\begin{definition}[Streams]
   \begin{equation}
@@
   \end{equation}
-\end{romdefinition}
-\begin{romdefinition}[Split Countability]
+\end{definition}
+\begin{definition}[Split Countability]
   \begin{equation}
@@
   \end{equation}
-\end{romdefinition}
+\end{definition}
 This type is definitionally equal to it surjection equivalent (\(\mathbb{N}
@@
 same way as we constructed cardinal finiteness.
-\begin{romdefinition}[Countability]
+\begin{definition}[Countability]
   \begin{equation}
@@
   \end{equation}
-\end{romdefinition}
+\end{definition}
 
 From both of these types we can derive decidable equality.
-\begin{romlemma}
+\begin{lemma}
   Any countable type has decidable equality.
-\end{romlemma}
+\end{lemma}
 \subsection{Closure}
@@
 what's left.
-\begin{romtheorem} \label{split-countability-sigma}
+\begin{theorem} \label{split-countability-sigma}
   Split countability is closed under \(\Sigma\).
-\end{romtheorem}
+\end{theorem}
 This proof does not mirror the proof of \(\Sigma\) closure on finite types.
@@
 types, we also \emph{gain} some: in particular the Kleene star.
-\begin{romtheorem}
+\begin{theorem}
   Split countability is closed under Kleene star.
@@
   \end{equation}
-\end{romtheorem}
+\end{theorem}
 Again, this proof requires a particular pattern to ensure productivity.
@@
 \subsection{Omniscience}
-\begin{romdefinition}[Limited Principle of Omniscience]
+\begin{definition}[Limited Principle of Omniscience]
   For any type \(A\) and predicate \(P\) on \(A\), the limited principle of
@@
   that predicate is also decidable.
-\end{romdefinition}
+\end{definition}
 The limited principle of omniscience is non-constructive, but individual types
@@
 There is also a universal form of omniscience, which we call exhaustibility.
-\begin{romdefinition}[Exhaustibility]
+\begin{definition}[Exhaustibility]
   We say a type \(A\) is exhaustible if, for any decidable predicate \(P\) on

### 2020-05-10 paper.tex
arrempting to compile without errors
compiles without errors
removed redundant qed
narrowed equations, used aleph_0 symbol for countability

@@
     \begin{equation}
-      \begin{aligned}
+      \begin{alignedat}{3}
         A \uplus B \coloneqq & \;
-        \text{inl} &: A \rightarrow A \uplus B \\
-        | & \;  \text{inr} &: B \rightarrow A \uplus B
-      \end{aligned}
+        \text{inl} &: A \rightarrow A \uplus B ; \\
+        | & \;  \text{inr} &: B \rightarrow A \uplus B ;
+      \end{alignedat}
     \end{equation}
@@
   \begin{equation} {
-    \begin{alignat*}{3}
-      \lVert A \rVert \coloneqq & \; \lvert \wc \rvert &&: A \rightarrow \lVert A \rVert ; \\
-                              | & \; \text{squash}     &&: \Pi {(x, y : \lVert A \rVert)} , x \equiv y  ; 
-    \end{alignat*} }
+    \begin{alignedat}{3}
+      \lVert A \rVert \coloneqq & \; \lvert \wc \rvert &:& \; A \rightarrow \lVert A \rVert ; \\
+                              | & \; \text{squash}     &:& \; \Pi {(x, y : \lVert A \rVert)} , x \equiv y  ; 
+    \end{alignedat} }
   \end{equation}
@@
     function \(f : A \rightarrow B\) iff \(f\) ``doesn't care'' about the
-    choice of \(A\) (\(\Pi {(x , y : A)} , f(x) \equiv f(y) \)).
+    choice of \(A\):
+    \[\Pi {(x , y : A)} , f(x) \equiv f(y) \]
     Formally speaking, \(f\) needs to be ``coherently constant''
@@
   \(\mathbf{Fin}(n)\) is the type of natural numbers smaller than \(n\). We
-  define it the standard way, where \(\mathbf{Fin}(0) \coloneqq \bot\) and
-  \(\mathbf{Fin}(n + 1) &\coloneqq \top \uplus \mathbf{Fin}(n)\).
+  define it the standard way:
+  \begin{equation}
+    \begin{alignedat}{3}
+      & \mathbf{Fin}(0)     && \coloneqq \bot ; \\
+      & \mathbf{Fin}(n + 1) && \coloneqq \top \uplus \mathbf{Fin}(n) ;
+    \end{alignedat}
+  \end{equation} 
 \end{definition}
@@
   \end{align*}
-  \qed
 \end{proof}
@@
   Therefore, by lemma~\ref{discrete-surj}, \(A\) is discrete.
-  \qed
 \end{proof}
@@
   \begin{equation}
-    \begin{alignat}{5}
+    \begin{alignedat}{3}
       S^1 \coloneqq & \; \text{base} &&: S^1 ; \\
       | & \; \text{loop} &&: \text{base} \equiv \text{base} ; 
-    \end{alignat}
+    \end{alignedat}
   \end{equation}
@@
   \begin{equation}
-    \begin{alignat}{2}
+    \begin{alignedat}{3}
       \mathcal{K}(A) \coloneqq&
-             \; []                &&: \mathcal{K}(A) ; \\
-      \vert& \; \wc \dblcolon \wc &&: A \rightarrow \mathcal{K}(A) \rightarrow \mathcal{K}(A) ; \\
-      \vert& \; \text{com}        &&: \Pi (x, y: A) , \Pi (\mathit{xs} : \mathcal{K}(A)) , x \dblcolon y \dblcolon \mathit{xs} \equiv y \dblcolon x \dblcolon \mathit{xs} ; \\
-      \vert& \; \text{dup}        &&: \Pi (x : A) , \Pi (\mathit{xs} : \mathcal{K}(A)) , x \dblcolon x \dblcolon \mathit{xs} \equiv x \dblcolon \mathit{xs} ; \\
-      \vert& \; \text{trunc}      &&: \Pi (\mathit{xs}, \mathit{ys}: \mathcal{K}(A)) , \Pi (p, q : \mathit{xs} \equiv \mathit{ys}) , p \equiv q ;
-    \end{alignat}
+             \; []                &:& \; \mathcal{K}(A) ; \\
+      \vert& \; \wc \dblcolon \wc &:& \; A \rightarrow \mathcal{K}(A) \rightarrow \mathcal{K}(A) ; \\
+      \vert& \; \text{com}        &:& \; \Pi (x, y: A) , \Pi (\mathit{xs} : \mathcal{K}(A)) , \\
+                                  &&& \; x \dblcolon y \dblcolon \mathit{xs} \equiv y \dblcolon x \dblcolon \mathit{xs} ; \\
+      \vert& \; \text{dup}        &:& \; \Pi (x : A) , \Pi (\mathit{xs} : \mathcal{K}(A)) , x \dblcolon x \dblcolon \mathit{xs} \equiv x \dblcolon \mathit{xs} ; \\
+      \vert& \; \text{trunc}      &:& \; \Pi (\mathit{xs}, \mathit{ys}: \mathcal{K}(A)) , \Pi (p, q : \mathit{xs} \equiv \mathit{ys}) , p \equiv q ;
+    \end{alignedat}
   \end{equation}
@@
     \begin{alignedat}{2}
-      x \in&& \; []                      &\coloneqq \bot \\
-      x \in&& \; y \dblcolon \mathit{ys} &\coloneqq \lVert x \equiv y \uplus x \in \mathit{ys} \rVert
+      x \in&& \; []                      &\coloneqq \bot ; \\
+      x \in&& \; y \dblcolon \mathit{ys} &\coloneqq \lVert x \equiv y \uplus x \in \mathit{ys} \rVert ;
     \end{alignedat}
@@
   is simply the \(\text{fst}\) function.
-  \qed
 \end{proof}
@@
   \begin{equation}
-    \begin{alignat}{3}
+    \begin{alignedat}{3}
       & \mathbf{Tuple}(0, T)   &&\coloneqq \top \\
       & \mathbf{Tuple}(n+1, T) &&\coloneqq T(0) \times \mathbf{Tuple}(n, T \circ \text{suc})
-    \end{alignat}
+    \end{alignedat}
   \end{equation}
@@
   enumerable types is itself split enumerable, proving our goal.
-  \qed
 \end{proof}
@@
   \begin{equation}
-    \begin{aligned}
+    \begin{alignedat}{3}
       &\text{Obj}_\mathit{FinSet}      &&\coloneqq \Sigma(x : \mathbf{Type}) , \mathcal{C}(x) \\
       &\text{Hom}_\mathit{FinSet}(x , y) &&\coloneqq  \text{fst}(x) \rightarrow \text{fst}(y)
-    \end{aligned}
+    \end{alignedat}
   \end{equation}
@@
   \begin{equation}
-    \mathcal{E}!(A) \coloneqq \Sigma {(\mathit{xs} : \mathbf{Stream}(A))} , \Pi {(x : A)} , x \in \mathit{xs}
+    \aleph_0!(A) \coloneqq \Sigma {(\mathit{xs} : \mathbf{Stream}(A))} , \Pi {(x : A)} , x \in \mathit{xs}
   \end{equation}
@@
   \begin{equation}
-    \mathcal{E}(A) \coloneqq \lVert \mathcal{E}!(A) \rVert
+    \aleph_0(A) \coloneqq \lVert \aleph_0!(A) \rVert
   \end{equation}
@@
   \begin{equation}
-    \mathcal{E}!(A) \rightarrow \mathcal{E}!(\mathbf{List}(A))
+    \aleph_0!(A) \rightarrow \aleph_0!(\mathbf{List}(A))
   \end{equation}
@@
 \end{acks}
-\printbibliography
+
+\bibliographystyle{ACM-Reference-Format}
+\bibliography{bibliography}
 \end{document}

### 2020-05-13 paper.tex
added diagram for proof search

@@
 \section{Countably Infinite Types} \label{infinite}
+\begin{figure*}
+  \centering
+  \begin{subfigure}[b]{.45\linewidth}
+    \centering
+    \begin{tikzcd}
+      (1,e) \ar[rdddd, out=-45, in=135] & (2,e) \ar[rdddd, out=-45, in=135] & (3,e) \ar[rdddd, out=-45, in=135] & (4,e) \ar[rdddd, out=-45, in=135] & (5,e)        \\
+      (1,d) \ar[u]     & (2,d) \ar[u]    & (3,d) \ar[u]    & (4,d) \ar[u]    & (5,d) \ar[u] \\
+      (1,c) \ar[u]     & (2,c) \ar[u]    & (3,c) \ar[u]    & (4,c) \ar[u]    & (5,c) \ar[u] \\
+      (1,b) \ar[u]     & (2,b) \ar[u]    & (3,b) \ar[u]    & (4,b) \ar[u]    & (5,b) \ar[u] \\
+      (1,a) \ar[u]     & (2,a) \ar[u]    & (3,a) \ar[u]    & (4,a) \ar[u]    & (5,a) \ar[u]
+    \end{tikzcd}
+    \caption{Depth-First}
+    \label{depth-first}
+  \end{subfigure} \hfill
+  \begin{subfigure}[b]{.45\linewidth}
+    \centering
+    \begin{tikzcd}
+      (1,e) \ar[dr] & (2,e) \ar[dr]  & (3,e) \ar[dr]    & (4,e) \ar[dr] & (5,e) \\
+      (1,d) \ar[dr] & (2,d) \ar[dr]  & (3,d) \ar[dr]    & (4,d) \ar[dr] & (5,d) \ar[u] \\
+      (1,c) \ar[dr] & (2,c) \ar[dr]  & (3,c) \ar[dr]    & (4,c) \ar[dr] & (5,c) \ar[uul] \\
+      (1,b) \ar[dr] & (2,b) \ar[dr]  & (3,b) \ar[dr]    & (4,b) \ar[dr] & (5,b) \ar[uuull, out=130, in=-50] \\
+      (1,a) \ar[u]  & (2,a) \ar[uul, out=130, in=-50] & (3,a) \ar[uuull, out=130, in=-50] & (4,a) \ar[uuuulll, out=130, in=-50] & (5,a) \ar[uuuulll, out=130, in=-50]
+    \end{tikzcd}
+    \caption{Breadth-First}
+    \label{breadth-first}
+  \end{subfigure}
+  \caption{Two possible products for the sets \(\left[ 1 \dots 5 \right]\) and
+    \(\left[  a \dots e \right]\)}
+  \label{pairings}
+\end{figure*}
 In the previous sections we saw different flavours of finiteness which were
@@
 \end{theorem}
-This proof does not mirror the proof of \(\Sigma\) closure on finite types.
-The pattern we used to pair up the relevant support lists for the finite types
-would diverge for infinite types: instead, we need a \emph{pairing} function.
+We know that countable infinity is not closed under the exponential (function
+arrow), so the only closure we need to prove is \(\Sigma\) to cover all of
+what's left.
+To do this we have to take a slightly different approach to the functions we
+defined before.
+Figure~\ref{pairings} illustrates the reason why: previously, we used the
+depth-first product pairing for each support list.
+This diverges if the first list is infinite, never exploring anything other than
+the first element in the second list.
+Instead, we use here the cantor pairing function, which performs a breadth-first
+search of the pairings of both lists.
 

### 2020-05-27 paper.tex
new format

@@
 \input{preamble}
-\acmConference[TyDe 2020]{Workshop on Type-driven Development}{23 Aug
-  2020}{Jersey City, USA}
 
-\author{Donnacha Oisín Kidney}
-\affiliation{University College Cork}
-\email{o.kidney@cs.ucc.ie}
-\author{Gregory Provan}
-\affiliation{University College Cork}
-\email{g.provan@cs.ucc.ie}
-\author{Nicolas Wu}
-\affiliation{Imperial College London}
-\email{n.wu@imperial.ac.uk}
+\author{
+  Donnacha Oisín Kidney \and
+  Gregory Provan \and
+  Nicolas Wu
+}
+% \affiliation{University College Cork}
+% \email{o.kidney@cs.ucc.ie}
+% \author{Gregory Provan}
+% \affiliation{University College Cork}
+% \email{g.provan@cs.ucc.ie}
+% \author{Nicolas Wu}
+% \affiliation{Imperial College London}
+% \email{n.wu@imperial.ac.uk}
 
@@
 
-\keywords{Agda, Homotopy Type Theory, Cubical Type Theory,
-  Dependent Types, Finiteness, Topos, Kuratowski finite}
+% \keywords{Agda, Homotopy Type Theory, Cubical Type Theory,
+%   Dependent Types, Finiteness, Topos, Kuratowski finite}
 
+\begin{document}
+
+\maketitle
 \begin{abstract}
@@
 
-\begin{document}
-\maketitle
 \section{Introduction}
@@
   \begin{align*}
-    \mathcal{E}!(A)
-    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , x \in \mathit{xs}
+     \mathcal{E}!(A) &
+    \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , x \in \mathit{xs}
     && \text{def.~\ref{split-enum-def} }(\mathcal{E}!)
@@
   \centering
-  \begin{subfigure}[b]{.45\linewidth}
+  \begin{subfigure}[b]{.45\textwidth}
     \centering
@@
   \end{subfigure} \hfill
-  \begin{subfigure}[b]{.45\linewidth}
+  \begin{subfigure}[b]{.45\textwidth}
     \centering
@@
 sets form a category.
-\begin{acks}
-  This work has been supported by the Science Foundation Ireland under the
-  following grant: 13/RC/2D94 to Irish Software Research Centre.
-\end{acks}
+% \begin{acks}
+%   This work has been supported by the Science Foundation Ireland under the
+%   following grant: 13/RC/2D94 to Irish Software Research Centre.
+% \end{acks}
 
-\bibliographystyle{ACM-Reference-Format}
+\bibliographystyle{plain}
 \bibliography{bibliography}

### 2020-05-27 paper.tex
some new intro stuff

@@
 \subsection{Foreword}
+Proofs in constructive mathematics are often more substantial than their
+classical counterparts.
+``Substantial'' here doesn't mean the literal length of the proofs (although the
+double meaning is not lost on anyone who has had to struggle with a modern
+constructive proof assistant), but rather refers to their \emph{content}: a
+constructive proof contains more \emph{stuff} than a classical one.
+
+If we were to prove that a set was finite in classical mathematics, we might
+show that it is in bijection with another finite set, say a finite prefix of the
+natural numbers.
+In semi-formal notation, to say that a set \(A\) is finite we would show the
+following:
+\begin{equation} \label{example-finite}
+  \mathbf{Finite}(A) \coloneqq \exists (n : \mathbb{N}) . \; A \Leftrightarrow \left\{ m \mid m < n\right\}
+\end{equation}
+Where \(A \Leftrightarrow B\) means ``a bijection between \(A\) and \(B\)''.
+
+Classically, a proof of equation~\ref{example-finite} is nothing more than
+\emph{evidence} for the proposition of equation~\ref{example-finite}.
+A constructive proof of \ref{example-finite}, however, could include much more:
+\begin{enumerate}
+  \item A number \(n\) representing the cardinality of \(A\).
+  \item A function from \(A\) to \(\left\{ m \mid m < n \right\}\).
+  \item A function from \(\left\{ m \mid m < n \right\}\) to \(A\).
+\end{enumerate}
+In other words, it makes sense constructively to talk about a function of type
+\(\mathbf{Finite}(A) \rightarrow \mathbb{N}\), which takes a \emph{proof} of the
+finiteness of \(A\) and returns \(A\)'s cardinality.
+
+Such functions are perhaps more concrete than this description suggests:
+aside from being interesting in its own right, much of the attention given to
+constructive mathematics comes from its perceived association with
+computers (themselves a burgeoning topic of study, bound to catch on some day). 
+Indeed many constructive proofs are in fact computer programs in disguise (and
+vice-versa \cite{wadlerPropositionsTypes2015}): the functions referred to above
+correspond to honest-to-goodness functions in a programming language (Agda
+\cite{vezzosiCubicalAgdaDependently2019} in our case) which can be run on a
+computer to produce real output.
+
+All of this added information in our proofs is interesting and useful, yes, but
+it also necessitates some extra caution.
+We will see cases where we have provided \emph{too much} information in a
+constructive proof, and we will need to hide it somehow.
+For instance, in order to construct a category of finite sets we will need to
+describe the objects of that category: constructively this would be ``pairs of
+sets along with proofs that they are finite''.
+This gives us the wrong category, however: for a type \(A\) with cardinality
+\(n\), there are \(n!\) proofs that \(A\) is finite.
+This means that we will have \(n!\) objects in the category for the set \(A\),
+where we only want 1.
+The solution is to hide some aspect of the proof: this is accomplished with
+\emph{truncation}, which we will see in \todo{ref}.
+
+For all of its benefits, constructive mathematics (especially dependently-typed
+constructive mathematics on a computer) still suffers from many painful
+drawbacks.
+As already hinted at, proofs in a constructive setting can tend to be verbose,
+especially when the arbiter of correctness is as strict as a computer, which
+neither understands nor trusts the time-honoured technique of ``the other cases
+are obviously true''.
+While the goal of using a computer to automate the boring parts of a proof still
+seems attainable, too often getting a computer involved in a proof means adding
+thousands of lines of tedium to satisfy its lack of intuition.
+
+In this paper we are interested in two things: systematic generation of data,
+and constructive finiteness.
+
+First, let's explain what we mean by ``systematic generation of data''.
+Consider QuickCheck \cite{claessenQuickCheckLightweightTool2011}: this
+well-known Haskell library allows us to quickly and easily test
+\ExecuteMetaData[haskell/PropertyTests.tex]{sort-stable}
+
 In constructive mathematics we are often preoccupied with \emph{why} something

### 2020-05-27 paper.tex
Update paper.tex

--- 2020-05-27 paper.tex
@@
 \subsection{Foreword}
-In constructive mathematics we are often preoccupied with \emph{why} something
-is true.
-Take finiteness, for example.
-There are a handful of ways to demonstrate some type is finite: we could provide
-a surjection to it from another finite type; we could show that any collection
-of its elements larger than some bound contains duplicates; or we could show
-that any stream of its elements contain duplicates.
-
-Classically, all of these proofs end up proving the same thing: that our type is
-finite.
-Constructively (in Martin-Löf Type Theory (MLTT)
-\cite{martin-lofIntuitionisticTypeTheory1980} at least), however, all three
-of the statements above construct a different version of finiteness.
-\emph{How} we show that some type is finite has a significant impact on the
-type of finiteness we end up dealing with.
-
-Homotopy Type Theory (HoTT) \cite{hottbook} adds another wrinkle to the story.
-Firstly, in HoTT we cannot assume that every type is a (homotopy) set: this
-means that the finiteness predicates above can be further split into versions
-which apply to sets only, and those that apply to all types.
-Secondly, HoTT gives us a principled and powerful way to construct quotients,
-allowing us to regain some of the flexibility of classical definitions by
-``forgetting'' the parts of a proof we would be forced to remember in MLTT.
-
-Finally, for a computer scientist constructive mathematics has one invaluable
-feature missing from classical mathematics: computation.
-Cubical Type Theory (CuTT) \cite{cohenCubicalTypeTheory2016}, and its
-implementation in Cubical Agda \cite{vezzosiCubicalAgdaDependently2019}, realise
-this property even in the presence of univalence, giving computational content
-to programs written in HoTT.
+Proofs in constructive mathematics are often more substantial than their
+classical counterparts.
+``Substantial'' here doesn't mean the literal length of the proofs (although the
+double meaning is not lost on anyone who has had to struggle with a modern
+constructive proof assistant), but rather refers to their \emph{content}: a
+constructive proof contains more \emph{stuff} than a classical one.
+
+If we were to prove that a set was finite in classical mathematics, we might
+show that it is in bijection with another finite set, say a finite prefix of the
+natural numbers.
+In semi-formal notation, to say that a set \(A\) is finite we would show the
+following:
+\begin{equation} \label{example-finite}
+  \mathbf{Finite}(A) \coloneqq \exists (n : \mathbb{N}) . \; A \Leftrightarrow \left\{ m \mid m < n\right\}
+\end{equation}
+Where \(A~\Leftrightarrow~B\) means ``a bijection between \(A\) and \(B\)''.
+
+Classically, a proof of equation~\ref{example-finite} is nothing more than
+\emph{evidence} for the proposition of equation~\ref{example-finite}.
+A constructive proof of \ref{example-finite}, however, could include much more:
+\begin{enumerate}
+  \item A number \(n\) representing the cardinality of \(A\).
+  \item A function from \(A\) to \(\left\{ m \mid m < n \right\}\).
+  \item A function from \(\left\{ m \mid m < n \right\}\) to \(A\).
+\end{enumerate}
+In other words, it makes sense (in constructive mathematics) to talk about a
+function of type \(\mathbf{Finite}(A)~\rightarrow~\mathbb{N}\), which takes a
+\emph{proof} of the finiteness of \(A\) and returns \(A\)'s cardinality.
+
+Such functions are perhaps even more concrete than this description suggests:
+many constructive proofs are in fact computer programs in disguise (and
+vice-versa \cite{wadlerPropositionsTypes2015}).
+All ``proofs'' in this paper correspond to honest-to-goodness proofs in a
+programming language (Cubical Agda \cite{vezzosiCubicalAgdaDependently2019} in
+our case) which can be run on a computer to produce real output.
+
+We have now hopefully established, albeit informally, what it means for
+constructive proofs to be ``substantial''.
+Clearly this extra information carried by proofs can add richness to their
+study: we will see, for instance, no fewer than five finiteness predicates in
+this paper which would collapse to one in a classical setting.
+The extra information also presents a challenge, however: occasionally we will
+see that we have revealed \emph{too much} information while proving some
+property.
+For instance, in order to construct a category of finite sets we will need to
+describe the objects of that category: constructively this would be ``pairs of
+sets along with proofs that they are finite''.
+The category we're trying to construct has one object for every finite set (up
+to unique isomorphism, etc).
+Here's the problem: if we take our finiteness proof to be a naive translation of
+the predicate in equation~\ref{example-finite}, then the proof includes a
+permutation on the set it refers to.
+This means that two objects \((A, \mathbf{Finite}(A))\) and \((A,
+\mathbf{Finite}(A))\) need not be the same, as they could have \emph{different
+  permutations} embedded in the \(\mathbf{Finite}(A)\) term.
+Where we wanted one object per set, we now have \(n!\) objects for every set of
+cardinality \(n\).
+
+Being able to hide the extra information in a constructive proof when we need to
+is clearly important, then.
+HoTT (and CuTT by extension) facilitate this hiding: how exactly it is
+accomplished will be explored in the subsequent sections.
+
+Finally, for all of its benefits, constructive mathematics (especially
+dependently-typed constructive mathematics on a computer) still suffers from
+many painful drawbacks.
+As already hinted at, proofs in a constructive setting can tend to be verbose,
+especially when the arbiter of correctness is as strict as a computer, which
+neither understands nor trusts the time-honoured technique of ``the other cases
+are obviously true''.
+While the goal of using a computer to automate the boring parts of a proof still
+seems attainable, too often getting a computer involved in a proof means adding
+thousands of lines of tedium to satisfy its lack of intuition.
+
+Simple proof search over a finite domain is a simple, if effective, way to
+relieve some of the tedium.
+Many libraries in Agda, Coq, and other languages make use of this fact.
+While the idea for finite proof search is simple, we found that implementing one
+revealed a number of interesting theoretical considerations:
+\begin{description}
+  \item[How should we construct search spaces?]
+    We found that a \(Pi\)-pretopos provided the appropriate ``generator
+    language'' for constructing search spaces for finite types.
+  \item[How do we deal with functions?]
+    We of course would like to be able to search over functions: \todo{?}
+  \item[How do we talk about ``searchability''?]
+    \todo{Omniscience}
+\end{description}
 \subsection{Contributions}

### 2020-05-28 paper.tex
bit more

@@
 \maketitle
-\begin{abstract}
-  We study five different notions of finiteness in Cubical Type Theory and prove
-  the relationship between them.
-  In particular we show that any totally ordered Kuratowski finite type is
-  manifestly Bishop finite.
-
-  We also prove closure properties for each finite type, and classify them
-  topos-theoretically.
-  This includes a proof that the category of decidable Kuratowski finite sets
-  (also called the category of cardinal finite sets) form a \(\Pi\)-pretopos.
-
-  We then develop a parallel classification for the countably infinite types, as
-  well as a proof of the countability of \(A^\star\) for a countable type \(A\).
-
-  We formalise our work in Cubical Agda, where we implement a library for proof
-  search (including combinators for level-polymorphic fully generic currying).
-  Through this library we demonstrate a number of uses for the computational
-  content of the univalence axiom, including searching for and synthesising
-  functions.
-\end{abstract}
+% \begin{abstract}
+%   We study five different notions of finiteness in Cubical Type Theory and prove
+%   the relationship between them.
+%   In particular we show that any totally ordered Kuratowski finite type is
+%   manifestly Bishop finite.
+
+%   We also prove closure properties for each finite type, and classify them
+%   topos-theoretically.
+%   This includes a proof that the category of decidable Kuratowski finite sets
+%   (also called the category of cardinal finite sets) form a \(\Pi\)-pretopos.
+
+%   We then develop a parallel classification for the countably infinite types, as
+%   well as a proof of the countability of \(A^\star\) for a countable type \(A\).
+
+%   We formalise our work in Cubical Agda, where we implement a library for proof
+%   search (including combinators for level-polymorphic fully generic currying).
+%   Through this library we demonstrate a number of uses for the computational
+%   content of the univalence axiom, including searching for and synthesising
+%   functions.
+% \end{abstract}
 
@@
 \emph{evidence} for the proposition of equation~\ref{example-finite}.
-A constructive proof of \ref{example-finite}, however, could include much more:
+A constructive proof of \ref{example-finite}, however, is a mathematical object
+in its own right, containing:
 \begin{enumerate}
@@
 property.
-For instance, in order to construct a category of finite sets we will need to
-describe the objects of that category: constructively this would be ``pairs of
-sets along with proofs that they are finite''.
-The category we're trying to construct has one object for every finite set (up
-to unique isomorphism, etc).
-Here's the problem: if we take our finiteness proof to be a naive translation of
-the predicate in equation~\ref{example-finite}, then the proof includes a
-permutation on the set it refers to.
-This means that two objects \((A, \mathbf{Finite}(A))\) and \((A,
-\mathbf{Finite}(A))\) need not be the same, as they could have \emph{different
-  permutations} embedded in the \(\mathbf{Finite}(A)\) term.
-Where we wanted one object per set, we now have \(n!\) objects for every set of
-cardinality \(n\).
+
+Consider the category of finite sets.
+We will explore this category in much more detail later, but for now we just
+need to know that the objects of this category are---shockingly---finite sets.
+Constructively, the objects of this category can be thought of as \emph{pairs},
+where the first element is a set (like \(\left\{ \text{True}, \text{False}
+\right\}\), the two-element set of the booleans) and the second is a proof that
+that set is finite.
+If we take those proofs to be the kind given in equation~\ref{example-finite},
+we should see that they contain a bijection from the booleans to the set
+\(\left\{ 0, 1 \right\}\): however \emph{two} such bijections exist!
+\begin{figure}[h]
+  \centering
+  \begin{subfigure}[b]{.45\textwidth}
+    \centering
+    \begin{tikzcd}
+      \text{True}  \ar[rr, leftrightarrow] & & 0 \\
+      \text{False} \ar[rr, leftrightarrow] & & 1
+    \end{tikzcd}
+  \end{subfigure}
+  \begin{subfigure}[b]{.45\textwidth}
+    \centering
+    \begin{tikzcd}
+      \text{True}  \ar[rrd, leftrightarrow, start anchor=south east, end anchor=north west] & & 0 \\
+      \text{False} \ar[rru, leftrightarrow, start anchor=north east, end anchor=south west, crossing over] & & 1
+    \end{tikzcd}
+  \end{subfigure}
+  \caption{The Two Bijections Between the Booleans and \(\left\{ 0, 1 \right\}\)}
+\end{figure}
+This means that where we wanted one object to correspond to the finite set
+\(\left\{ \text{True}, \text{False} \right\}\), we instead have two.
+In fact, for any finite set of cardinality \(n\) we will have \(n!\) finiteness
+proofs, and therefore \(n!\) objects.
+We have constructed, in other words, entirely the wrong category.
 
 Being able to hide the extra information in a constructive proof when we need to
-is clearly important, then.
+is clearly important.
 HoTT (and CuTT by extension) facilitate this hiding: how exactly it is
@@
 \end{description}
+Finiteness is a prime example of a phenomenon which happens quite frequently in
+dependently-typed constructive mathematics: through building a practical program
+we unearth theoretically interesting results.
+\todo{rephrase}
 \subsection{Contributions}
-The finiteness predicates we are interested in are organised in
-figure~\ref{finite-classification}, and defined in
-section~\ref{finiteness-predicates}.
-We will explore two aspects of each predicate: its relation to the other
-predicates, and its topos-theoretic classification.
-
-\input{figures/finite-classification}
-
-When we say ``relation'' we are referring to the arrows in
+We are interested in constructive notions of finiteness, their classification as
+a topos, their relation to countability, and their use as a proof search library
+in Cubical Agda.
+\subsubsection{Finiteness Predicates}
+Our focus in this work is decidable Kuratowski finiteness.
+We will characterise it, its strength, and its uses, and in particular we will
+show that it forms a \(\Pi\)-pretopos.
+Our proofs follow the structure of \cite[Chapters 9, 10]{hottbook} and
+\cite{rijkeSetsHomotopyType2015}.
+We will explore this concept in more detail later, but crucially we will need to
+provide proofs of closure under \(\times\), \(\rightarrow\), etc.
+
+In order to prove these closure properties, we will define and use some
+intermediate finiteness predicates, classified in
 figure~\ref{finite-classification}.
+These predicates are also interesting in their own right, though!
+
+The arrows in figure~\ref{finite-classification} are implication functions
+between each predicate.
 In section~\ref{relations} we will provide a function which inhabits each arrow.
@@
 
-We will then examine the closure properties of each predicate in
-section~\ref{topos}, culminating in a proof that decidable Kuratowski finite
-sets form a \(\Pi\)-pretopos (theorem~\ref{finite-topos}).
-Our proofs follow the structure of \cite[Chapters 9, 10]{hottbook} and
-\cite{rijkeSetsHomotopyType2015}.
-
+\input{figures/finite-classification}
+\subsubsection{Countability Predicates}
 After the finite predicates, we will briefly look at the infinite countable
@@
 (section~\ref{infinite}).
-
+\subsubsection{Search}
 All of our work is formalised in Cubical Agda
@@
 space.
+The language of a \(\Pi\)-pretopos turns out to be exactly the right language
+for constructing ``generator''-like expressions.
 
@@
   \centering
+
+  \tikzcdset{
+  cramped,
+  every matrix/.append style={nodes={font=\scriptsize}},
+  row    sep/normal=1em,
+  column sep/normal=1em,
+  }
   \begin{subfigure}[b]{.45\textwidth}

### 2020-05-28 paper.tex
re-lengthening

@@
 \subsection{Contributions}
-We are interested in constructive notions of finiteness, their classification as
-a topos, their relation to countability, and their use as a proof search library
-in Cubical Agda.
+We are interested in constructive notions of finiteness.
+In particular, we are interested in two different aspects of each predicate for
+finiteness: how informative it is (i.e. how many properties of the underlying
+type we can derive from a proof that it is finite), and how restrictive it is
+(i.e. how many types conform to its particular variety of finiteness).
+Once we explore each of five predicates along these two axes, we will then
+construct a category out of one of them (decidable Kuratowski finiteness), and
+show that it forms a \(\Pi\)-pretopos.
+We will then take a brief detour into predicates for countably infinite types,
+and compare its landscape to that of the finite predicates, before jumping into
+a practical application of the work: a library for proof search over finite
+domains.
+We will show how our earlier work gives a good theoretical foundation for this
+library, and allows for a simple and principled API, and we will demonstrate how
+some of the more complex theoretical results allow for a more powerful library
+with capabilities not seen in comparable proof search libraries.
+
 \subsubsection{Finiteness Predicates}
@@
     used.
+  \item[Lambdas] \todo{Figure out how to describe lambdas}
 \end{description}
@@
 \end{proof}
-
-In our formalisation, the proof is a single line: most of the steps above are
-simple expansion of definitions.
+\begin{agdalisting}
+In our formalisation, the proof is a single line:
+\ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{is-split-inj}
 The only step which isn't definitional equality is the reassociation of
 \(\Sigma\).
+\ExecuteMetaData[agda/Data/Sigma/Properties.tex]{reassoc}
+\end{agdalisting}
 
@@
       \vert& \; \wc \dblcolon \wc &:& \; A \rightarrow \mathcal{K}(A) \rightarrow \mathcal{K}(A) ; \\
-      \vert& \; \text{com}        &:& \; \Pi (x, y: A) , \Pi (\mathit{xs} : \mathcal{K}(A)) , \\
-                                  &&& \; x \dblcolon y \dblcolon \mathit{xs} \equiv y \dblcolon x \dblcolon \mathit{xs} ; \\
+      \vert& \; \text{com}        &:& \; \Pi (x, y: A) , \Pi (\mathit{xs} : \mathcal{K}(A)) ,
+                                  x \dblcolon y \dblcolon \mathit{xs} \equiv y \dblcolon x \dblcolon \mathit{xs} ; \\
       \vert& \; \text{dup}        &:& \; \Pi (x : A) , \Pi (\mathit{xs} : \mathcal{K}(A)) , x \dblcolon x \dblcolon \mathit{xs} \equiv x \dblcolon \mathit{xs} ; \\

### 2020-05-29 figures/finite-classification.tex
fix: two column figure
rearranged finite

@@
-\begin{figure}
+\begin{figure*}
   \centering
   \begin{tikzcd}[cramped, row sep=small]
-    &
-    \ar[ddddd, dash, start anchor={[yshift=4ex]}, end anchor={[yshift=-4ex]}] &
+    {} &
+    {} \ar[ddddd, dash, start anchor={[yshift=4ex]}, end anchor={[yshift=-4ex]}] &
+    \text{Unordered} &
+    \ar[ddddd, dash, dashed, start anchor={[yshift=4ex]}, end anchor={[yshift=-4ex]}] &
     \text{Ordered} &
-    \ar[ddddd, dash, dashed, start anchor={[yshift=4ex]}, end anchor={[yshift=-4ex]}] &
-    \text{Unordered}
+    {}
     \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     {}
+    &
     \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    &
+    &
     &
@@
       \ar[d, xshift=-1ex, bend right=30]
-      \ar[rrd, end anchor=north west, start anchor=east, bend left=10, crossing over]
-      &
+      \ar[lld, end anchor=north east, start anchor=west, bend right=10, crossing over]
     &
@@
     &
+    \text{Cardinal}
+    \ar[rr, start anchor=east, end anchor=west, "\text{Ord}" description]
+    &
+    &
     \text{Split Enumerable}
@@
     &
-    &
-    \text{Cardinal}
-      \ar[ll, start anchor=west, end anchor=east, "\text{Ord}" description]
     \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     {}
+    &
     \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     {} &
-    \text{Manifest Enumerable}
+    \text{Kuratowski}
       \ar[from=uu, xshift=-1ex, bend right=30, crossing over]
       \ar[uu, bend right=30,  "\text{Discrete}" description]
-      \ar[rr, start anchor=east, end anchor=west, crossing over]
       &
@@
     &
-    \text{Kuratowski}
+    \text{Manifest Enumerable}
       \ar[uu, xshift=1ex, bend right=30,  "\text{Discrete}" description]
       \ar[from=uu, xshift=-1ex, bend right=30, crossing over]
+      \ar[ll, start anchor=west, end anchor=east, crossing over]
+    &
+    \\
+    & & & & &
+    \\
+    & \ar[rrrr, "\text{More Informative}" description, end anchor={[xshift=-6ex]}]
+    &
+    &
+    &
+    &
+    \ar[uuuuuuu, "\text{More Restrictive}" description]
+    {}
   \end{tikzcd}
@@
   \label{finite-classification}
-\end{figure}%
+\end{figure*}%

### 2020-05-29 paper.tex
rearranged finite

@@
 We are interested in constructive notions of finiteness.
-In particular, we are interested in two different aspects of each predicate for
-finiteness: how informative it is (i.e. how many properties of the underlying
-type we can derive from a proof that it is finite), and how restrictive it is
-(i.e. how many types conform to its particular variety of finiteness).
-Once we explore each of five predicates along these two axes, we will then
-construct a category out of one of them (decidable Kuratowski finiteness), and
-show that it forms a \(\Pi\)-pretopos.
+In contrast to classical finiteness, in a constructive setting there is a wide
+variety of predicates which all have some claim to being the formal
+interpretation of ``finiteness''.
+Broadly, they vary along two axes: informativeness (i.e. how many properties
+of the underlying type we can derive from a proof that it is finite), and
+restrictiveness (i.e. how many types conform to its particular variety of
+finiteness).
+We will explore five of these predicates, proving formally their relations and
+implications.
+
+We will then focus on one of the predicates: decidable Kuratowski finiteness.
+We will construct a category of decidable Kuratowski-finite types, and show that
+it forms a \(\Pi\)-pretopos.
+
 We will then take a brief detour into predicates for countably infinite types,
-and compare its landscape to that of the finite predicates, before jumping into
-a practical application of the work: a library for proof search over finite
-domains.
+and compare its landscape to that of the finite predicates.
+
+Finally, we will present a case-study in a practical application of the
+predicates for finiteness we have constructed: a library for total proof search.
 We will show how our earlier work gives a good theoretical foundation for this

### 2020-05-29 figures/finite-classification.tex
Update finite-classification.tex

@@
     {} \ar[ddddd, dash, start anchor={[yshift=4ex]}, end anchor={[yshift=-4ex]}] &
-    \text{Unordered} &
+    \text{Non Discrete} &
     \ar[ddddd, dash, dashed, start anchor={[yshift=4ex]}, end anchor={[yshift=-4ex]}] &
-    \text{Ordered} &
+    \text{Discrete} &
     {}
@@
     &
+    {}
     \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
     &
+    &
+    \text{Manifest Enumerable}
+    \ar[
+      from=rr,
+      bend left=15,
+      crossing over,
+      start anchor=south west,
+      end anchor=south east
+      ]
+    \ar[
+      rr,
+      bend left=15,
+      "\text{Discrete}" description,
+      start anchor=north east,
+      end anchor=north west
+      ]
+    \ar[ddd, crossing over]
+    &
+    &
+    \text{Split Enumerable}
+      \ar[d, xshift=-1ex, bend right=30]
+    &
+    \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \text{Ordered} &
     &
@@
     \text{Manifest Bishop}
-      \ar[d, xshift=-1ex, bend right=30]
-      \ar[lld, end anchor=north east, start anchor=west, bend right=10, crossing over]
-    &
-    \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \text{Discrete} &
-    &
-    \text{Cardinal}
-    \ar[rr, start anchor=east, end anchor=west, "\text{Ord}" description]
-    &
-    &
-    \text{Split Enumerable}
       \ar[u, xshift=1ex, bend right=30]
@@
     \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \text{Non Discrete} &
+    \text{Unordered} &
     {} &
     \text{Kuratowski}
-      \ar[from=uu, xshift=-1ex, bend right=30, crossing over]
-      \ar[uu, bend right=30,  "\text{Discrete}" description]
+    \ar[rr, crossing over, bend left=15, "\text{Discrete}" description, start anchor=north east, end anchor=north west]
+    \ar[from=rr, crossing over, bend left=15, start anchor=south west, end anchor=south east]
     &
@@
     &
-    \text{Manifest Enumerable}
-      \ar[uu, xshift=1ex, bend right=30,  "\text{Discrete}" description]
-      \ar[from=uu, xshift=-1ex, bend right=30, crossing over]
-      \ar[ll, start anchor=west, end anchor=east, crossing over]
+    \text{Cardinal}
+      \ar[from=uu, crossing over]
+      \ar[uuu, crossing over, "\text{Ord}" description, end anchor=east, start
+      anchor=east, in=-15, out=20]
     &
@@
     \\
-    & \ar[rrrr, "\text{More Informative}" description, end anchor={[xshift=-6ex]}]
+    & \ar[rrrr, "\text{More Restrictive}" description, end anchor={[xshift=-6ex]}]
     &
@@
     &
-    \ar[uuuuuuu, "\text{More Restrictive}" description]
+    \ar[uuuuuu, "\text{More Informative}" {description, near start}]
     {}

### 2020-05-29 paper.tex
split into sections

--- 2020-05-28 paper.tex
@@
 % \end{abstract}
+\include{sections/introduction}
+\include{sections/finiteness-predicates}
+\include{sections/topos}
+\include{sections/countable-predicates}
+\include{sections/search}
+\include{sections/related-work}
 
-\section{Introduction}
-\subsection{Foreword}
-Proofs in constructive mathematics are often more substantial than their
-classical counterparts.
-``Substantial'' here doesn't mean the literal length of the proofs (although the
-double meaning is not lost on anyone who has had to struggle with a modern
-constructive proof assistant), but rather refers to their \emph{content}: a
-constructive proof contains more \emph{stuff} than a classical one.
-
-If we were to prove that a set was finite in classical mathematics, we might
-show that it is in bijection with another finite set, say a finite prefix of the
-natural numbers.
-In semi-formal notation, to say that a set \(A\) is finite we would show the
-following:
-\begin{equation} \label{example-finite}
-  \mathbf{Finite}(A) \coloneqq \exists (n : \mathbb{N}) . \; A \Leftrightarrow \left\{ m \mid m < n\right\}
-\end{equation}
-Where \(A~\Leftrightarrow~B\) means ``a bijection between \(A\) and \(B\)''.
-
-Classically, a proof of equation~\ref{example-finite} is nothing more than
-\emph{evidence} for the proposition of equation~\ref{example-finite}.
-A constructive proof of \ref{example-finite}, however, is a mathematical object
-in its own right, containing:
-\begin{enumerate}
-  \item A number \(n\) representing the cardinality of \(A\).
-  \item A function from \(A\) to \(\left\{ m \mid m < n \right\}\).
-  \item A function from \(\left\{ m \mid m < n \right\}\) to \(A\).
-\end{enumerate}
-In other words, it makes sense (in constructive mathematics) to talk about a
-function of type \(\mathbf{Finite}(A)~\rightarrow~\mathbb{N}\), which takes a
-\emph{proof} of the finiteness of \(A\) and returns \(A\)'s cardinality.
-
-Such functions are perhaps even more concrete than this description suggests:
-many constructive proofs are in fact computer programs in disguise (and
-vice-versa \cite{wadlerPropositionsTypes2015}).
-All ``proofs'' in this paper correspond to honest-to-goodness proofs in a
-programming language (Cubical Agda \cite{vezzosiCubicalAgdaDependently2019} in
-our case) which can be run on a computer to produce real output.
-
-We have now hopefully established, albeit informally, what it means for
-constructive proofs to be ``substantial''.
-Clearly this extra information carried by proofs can add richness to their
-study: we will see, for instance, no fewer than five finiteness predicates in
-this paper which would collapse to one in a classical setting.
-The extra information also presents a challenge, however: occasionally we will
-see that we have revealed \emph{too much} information while proving some
-property.
-
-Consider the category of finite sets.
-We will explore this category in much more detail later, but for now we just
-need to know that the objects of this category are---shockingly---finite sets.
-Constructively, the objects of this category can be thought of as \emph{pairs},
-where the first element is a set (like \(\left\{ \text{True}, \text{False}
-\right\}\), the two-element set of the booleans) and the second is a proof that
-that set is finite.
-If we take those proofs to be the kind given in equation~\ref{example-finite},
-we should see that they contain a bijection from the booleans to the set
-\(\left\{ 0, 1 \right\}\): however \emph{two} such bijections exist!
-\begin{figure}[h]
-  \centering
-  \begin{subfigure}[b]{.45\textwidth}
-    \centering
-    \begin{tikzcd}
-      \text{True}  \ar[rr, leftrightarrow] & & 0 \\
-      \text{False} \ar[rr, leftrightarrow] & & 1
-    \end{tikzcd}
-  \end{subfigure}
-  \begin{subfigure}[b]{.45\textwidth}
-    \centering
-    \begin{tikzcd}
-      \text{True}  \ar[rrd, leftrightarrow, start anchor=south east, end anchor=north west] & & 0 \\
-      \text{False} \ar[rru, leftrightarrow, start anchor=north east, end anchor=south west, crossing over] & & 1
-    \end{tikzcd}
-  \end{subfigure}
-  \caption{The Two Bijections Between the Booleans and \(\left\{ 0, 1 \right\}\)}
-\end{figure}
-This means that where we wanted one object to correspond to the finite set
-\(\left\{ \text{True}, \text{False} \right\}\), we instead have two.
-In fact, for any finite set of cardinality \(n\) we will have \(n!\) finiteness
-proofs, and therefore \(n!\) objects.
-We have constructed, in other words, entirely the wrong category.
-
-Being able to hide the extra information in a constructive proof when we need to
-is clearly important.
-HoTT (and CuTT by extension) facilitate this hiding: how exactly it is
-accomplished will be explored in the subsequent sections.
-
-Finally, for all of its benefits, constructive mathematics (especially
-dependently-typed constructive mathematics on a computer) still suffers from
-many painful drawbacks.
-As already hinted at, proofs in a constructive setting can tend to be verbose,
-especially when the arbiter of correctness is as strict as a computer, which
-neither understands nor trusts the time-honoured technique of ``the other cases
-are obviously true''.
-While the goal of using a computer to automate the boring parts of a proof still
-seems attainable, too often getting a computer involved in a proof means adding
-thousands of lines of tedium to satisfy its lack of intuition.
-
-Simple proof search over a finite domain is a simple, if effective, way to
-relieve some of the tedium.
-Many libraries in Agda, Coq, and other languages make use of this fact.
-While the idea for finite proof search is simple, we found that implementing one
-revealed a number of interesting theoretical considerations:
-\begin{description}
-  \item[How should we construct search spaces?]
-    We found that a \(Pi\)-pretopos provided the appropriate ``generator
-    language'' for constructing search spaces for finite types.
-  \item[How do we deal with functions?]
-    We of course would like to be able to search over functions: \todo{?}
-  \item[How do we talk about ``searchability''?]
-    \todo{Omniscience}
-\end{description}
-Finiteness is a prime example of a phenomenon which happens quite frequently in
-dependently-typed constructive mathematics: through building a practical program
-we unearth theoretically interesting results.
-\todo{rephrase}
-\subsection{Contributions}
-We are interested in constructive notions of finiteness, their classification as
-a topos, their relation to countability, and their use as a proof search library
-in Cubical Agda.
-\subsubsection{Finiteness Predicates}
-Our focus in this work is decidable Kuratowski finiteness.
-We will characterise it, its strength, and its uses, and in particular we will
-show that it forms a \(\Pi\)-pretopos.
-Our proofs follow the structure of \cite[Chapters 9, 10]{hottbook} and
-\cite{rijkeSetsHomotopyType2015}.
-We will explore this concept in more detail later, but crucially we will need to
-provide proofs of closure under \(\times\), \(\rightarrow\), etc.
-
-In order to prove these closure properties, we will define and use some
-intermediate finiteness predicates, classified in
-figure~\ref{finite-classification}.
-These predicates are also interesting in their own right, though!
-
-The arrows in figure~\ref{finite-classification} are implication functions
-between each predicate.
-In section~\ref{relations} we will provide a function which inhabits each arrow.
-Each unlabelled arrow is an unconditional implication: every manifest Bishop
-finite set is cardinal finite (lemma~\ref{manifest-bishop-to-cardinal}), for
-instance.
-The labelled arrows are strengthening proofs: every manifest enumerable set
-\emph{with decidable equality} is split enumerable
-(lemma~\ref{manifest-enum-to-split-enum}).
-Our most significant result here is the proof that a cardinal finite set with a
-decidable total order is manifestly Bishop finite.
-
-\input{figures/finite-classification}
-\subsubsection{Countability Predicates}
-After the finite predicates, we will briefly look at the infinite countable
-types, and classify them in a parallel way to the finite predicates
-(section~\ref{infinite}).
-\subsubsection{Search}
-All of our work is formalised in Cubical Agda
-\cite{vezzosiCubicalAgdaDependently2019}.
-We will make mention of the few occasions where the formalisation of some proof
-is of interest, but the main place where we will discuss the code is in
-section~\ref{search}, where we implement a library for proof search, based on
-omniscience and exhaustibility.
-This library relies directly on the computational content of univalence, which
-allows us to (for instance) have \(\Pi\) types in the domain of the search
-space.
-The language of a \(\Pi\)-pretopos turns out to be exactly the right language
-for constructing ``generator''-like expressions.
-
-The full code of the formalisation is available at
-\url{https://github.com/oisdk/finiteness-in-cubical-type-theory}.
-\subsection{Notation and Background}
-\subsubsection{Notation}
-We work in Cubical Type Theory \cite{cohenCubicalTypeTheory2016}.
-For the various type formers we use the following notation:
-\begin{description}
-  \item[Type] We use \(\mathbf{Type}\) to denote the universe of (small) types.
-    ``Type families'' are functions into \(\mathbf{Type}\).
-  \item[0 , 1 , 2] We call the \(\mathbf{0}\), \(\mathbf{1}\), and
-    \(\mathbf{2}\) types \(\bot\), \(\top\), and \(\mathbf{Bool}\) respectively.
-    The single inhabitant of \(\top\) is tt, and the two inhabitants of
-    \(\mathbf{Bool}\) are false and true.
-    The ``negation'' of a type, written \(\neg A\), means \(A \rightarrow
-    \bot\).
-  \item[Dependent Sum and Product] We use \(\Sigma\) and \(\Pi\) for the
-    dependent sum and product, respectively.
-    The two projections from \(\Sigma\) are called fst and snd.
-    In the non-dependent case, \(\Sigma\) can be written as \(\times\), and
-    \(\Pi\) as \(\rightarrow\).
-  \item[Disjoint Union] We define disjoint union as an inductive type.
-    \begin{equation}
-      \begin{alignedat}{3}
-        A \uplus B \coloneqq & \;
-        \text{inl} &: A \rightarrow A \uplus B ; \\
-        | & \;  \text{inr} &: B \rightarrow A \uplus B ;
-      \end{alignedat}
-    \end{equation}
-    It is also expressible with only \(\Sigma\):
-    \(A \uplus B \simeq \Sigma(x : \mathbf{Bool}) , \text{if } x \text{ then
-    } A \text{ else } B \).
-  \item[Equalities, equivalences, and paths] We use the symbol \(\coloneqq\)
-    for definitions.
-    \(\simeq\) will be used for equivalences, and \(\equiv\) for equalities.
-    Of course, we know that \((A \simeq B) \simeq (A \equiv B)\) by univalence,
-    so the distinction isn't terribly important in usage: we will only use one
-    or the other as a suggestion of how we constructed it or how it is to be
-    used.
-\end{description}
-\subsubsection{Cubical Type Theory}
-Cubical Type Theory \cite{cohenCubicalTypeTheory2016} is a constructive type
-theory with an implementation in Cubical Agda
-\cite{vezzosiCubicalAgdaDependently2019}.
-It allows us to do much of the same theory as in HoTT, but crucially the
-univalence ``axiom'' is a \emph{theorem}, giving it computational content.
-\begin{definition}[Path Types] \label{path-types}
-  The equality type (which we denote with \(\equiv\)) in CuTT is the type of
-  Paths\footnotemark.
-  The internal structure of paths is largely irrelevant to us here, as we will
-  generally treat \(\equiv\) as a black-box equivalence relation with
-  substitution and congruence.
-\end{definition}
-
-\footnotetext{
-  Actually, CuTT does have an identity type with similar semantics to the
-  identity type in MLTT.
-  We do not use this type anywhere in our work, however, so we will not consider
-  it here.
-}
-\begin{definition}[Homotopy Levels] \label{homotopy-types}
-  Types in HoTT and CuTT are not necessarily sets, as they are in MLTT.
-  Some have higher homotopies (paths which aren't unique).
-  We actually have a hierarchy of complexity of structure of path spaces in
-  types, starting with the contractions \cite[definition 3.11.1]{hottbook}, then
-  the mere propositions \cite[definition 3.3.1]{hottbook}, and the sets
-  \cite[definition 3.1.1]{hottbook}.
-  \begin{alignat}{2}
-    &\text{isContr}(A)    &&\coloneqq \Sigma(x : A) , \Pi(y : A) , (x \equiv y) \\
-    &\text{isProp}(A)     &&\coloneqq \Pi(x, y : A) , (x \equiv y) \\
-    &\text{isSet}(A)      &&\coloneqq \Pi(x, y : A) , \text{isProp}(x \equiv y)
-  \end{alignat}
-\end{definition}
-\begin{definition}[Fibres] \label{fibres}
-  A fibre \cite[definition 4.2.4]{hottbook} is defined over some function \(f :
-  A \rightarrow B\).
-  \begin{equation}
-    \text{fib}_f(y) \coloneqq \Sigma(x : A) , (f (x) \equiv y)
-  \end{equation}
-\end{definition}
-\begin{definition}[Equivalences] \label{equivalences}
-  We will take contractible maps \cite[definition 4.4.1]{hottbook} as our
-  ``default'' definition of equivalences.
-  \begin{alignat}{2}
-    &\text{isEquiv}(f) &&\coloneqq \Pi(y : B) , \text{isContr}(\text{fib}_f(y)) \label{is-equiv-def} \\
-    &A \simeq B        &&\coloneqq \Sigma(f : A \rightarrow B) , \text{isEquiv}(f)
-  \end{alignat}
-\end{definition}
-\begin{definition}[Decidable Types]
-  \begin{equation}
-    \mathbf{Dec}(A) \coloneqq A \uplus \neg A
-  \end{equation}
-\end{definition}
-\begin{definition}[Discrete Types]
-  A discrete type is one with decidable equality.
-  \begin{equation}
-    \text{Discrete}(A) \coloneqq \Pi(x, y : A) , \mathbf{Dec}(x \equiv y)
-  \end{equation}
-  By Hedberg's theorem \cite{hedbergCoherenceTheoremMartinLof1998} any discrete
-  type is a set.
-\end{definition}
-\begin{definition}[Higher Inductive Types] \label{HITs}
-  Normal inductive types have \emph{point} constructors: constructors which
-  construct values of the type.
-  Higher Inductive Types (HITs) also have \emph{path} constructors: ways to
-  construct paths in the type.
-\end{definition}
-\begin{definition}[Propositional Truncation] \label{prop-trunc}
-  The type \(\lVert A \rVert\) on some type \(A\) is a propositionally truncated
-  proof of \(A\) \cite[3.7]{hottbook}.
-  In other words, it is a proof that some \(A\) exists, but it does not tell you
-  \emph{which} \(A\).
-
-  It is defined as a Higher Inductive Type:
-  \begin{equation} {
-    \begin{alignedat}{3}
-      \lVert A \rVert \coloneqq & \; \lvert \wc \rvert &:& \; A \rightarrow \lVert A \rVert ; \\
-                              | & \; \text{squash}     &:& \; \Pi {(x, y : \lVert A \rVert)} , x \equiv y  ; 
-    \end{alignedat} }
-  \end{equation}
-  We will use two eliminators from \(\lVert A \rVert\) in this paper.
-  \begin{enumerate}
-  \item \label{elim-prop-prop} For any function \(A \rightarrow B\), where
-    \(\text{isProp}(B)\), we have a function \(\lVert A \rVert \rightarrow B\).
-  \item \label{elim-prop-coh} We can eliminate from \(\lVert A \rVert\) with a
-    function \(f : A \rightarrow B\) iff \(f\) ``doesn't care'' about the
-    choice of \(A\):
-    \[\Pi {(x , y : A)} , f(x) \equiv f(y) \]
-    Formally speaking, \(f\) needs to be ``coherently constant''
-    \cite{krausGeneralUniversalProperty2015}, and \(B\) needs to be an
-    \(n\)-type for some finite \(n\).
-  \end{enumerate}
-\end{definition}
-\section{Finiteness Predicates} \label{finiteness-predicates}
-In this section, we will define and briefly describe each of the five predicates
-in Figure~\ref{finite-classification}.
-The reason we explore predicates other than our focus (decidable Kuratowski
-finiteness) is that we can often prove things like closure much more readily on
-the simpler predicates.
-The relations (which we will prove in the next section) then allow us to
-transfer those proofs.
-\subsection{Split Enumerability}
-\begin{definition}[Split Enumerable Set] \label{split-enum-def}
-  \begin{equation} \label{split-enum-def-eqn}
-    \mathcal{E}!(A) \coloneqq \Sigma {(\mathit{xs} : \textbf{List}(A))} , \Pi {(x : A)} , x \in \mathit{xs}
-  \end{equation}
-  We call the first component of this pair the ``support'' list, and the second
-  component the ``cover'' proof.
-  An equivalent version of this predicate was called \verb+Listable+ in
-  \cite{firsovDependentlyTypedProgramming2015}.
-\end{definition}
-
-We used some extra types in the above definition, which we will define here:
-\begin{definition}[Containers] \label{container-def}
-  A container \cite{abbottContainersConstructingStrictly2005} is a pair
-  \(S , P\) where \(S\) is a type, the elements of which are called
-  the \emph{shapes} of the container, and \(P\) is a type family on \(S\), where
-  the elements of \(P(s)\) are called the \emph{positions} of a container.
-  We ``interpret'' a container into a functor defined like so:
-  \begin{equation} \label{container-interp}
-    \llbracket S , P \rrbracket(A) \coloneqq \Sigma {(s : S)} , \left( P(s) \rightarrow A \right)
-  \end{equation}
-  Membership of a container can be defined like so:
-  \begin{equation} \label{container-membership}
-    x \in \mathit{xs} \coloneqq \text{fib}_{\text{snd}(\mathit{xs})}(x)
-  \end{equation}
-\end{definition}
-\begin{definition}[\(\mathbf{List}\)] \label{List}
-  \begin{equation}
-    \mathbf{List} \coloneqq \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket
-  \end{equation}
-\end{definition}
-\begin{definition}[\(\mathbf{Fin}\)] \label{Fin}
-  \(\mathbf{Fin}(n)\) is the type of natural numbers smaller than \(n\). We
-  define it the standard way:
-  \begin{equation}
-    \begin{alignedat}{3}
-      & \mathbf{Fin}(0)     && \coloneqq \bot ; \\
-      & \mathbf{Fin}(n + 1) && \coloneqq \top \uplus \mathbf{Fin}(n) ;
-    \end{alignedat}
-  \end{equation} 
-\end{definition}
-
-We tend to prefer list-based definitions of finiteness, rather than ones based
-on bijections or surjections.
-This is purely a matter of perspective, however: the definition above is
-precisely equivalent to a split surjection from a finite prefix of the natural
-numbers.
-\begin{definition}[Surjections] \label{surjections}
-  We define both surjections and \emph{split} surjections here \cite[definition
-  4.6.1]{hottbook}.
-  \begin{alignat}{3}
-    &\text{surj}(f)             &&\coloneqq \Pi(y : B) , \lVert \text{fib}_f(y) \rVert \\
-    &A \twoheadrightarrow B     &&\coloneqq \Sigma (f : A \rightarrow B) , \text{surj}(f) \label{surj-arrow-eqn} \\
-    &\text{sp-surj}(f)          &&\coloneqq \Pi(y : B) , \text{fib}_f(y) \label{sp-surj-eqn} \\
-    &A \twoheadrightarrow! \; B &&\coloneqq \Sigma (f : A \rightarrow B) , \text{sp-surj}(f) \label{sp-surj-arrow-eqn}
-  \end{alignat}
-\end{definition}
-\begin{lemma} \label{split-enum-is-split-surj}
-  \begin{equation}
-    \mathcal{E}!(A) \simeq \Sigma (n : \mathbb{N}) , \left( \mathbf{Fin}(n) \twoheadrightarrow ! \; A \right)
-  \end{equation}
-\end{lemma}
-\begin{proof}
-  \begin{align*}
-     \mathcal{E}!(A) &
-    \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , x \in \mathit{xs}
-    && \text{def.~\ref{split-enum-def} }(\mathcal{E}!)
-    \\
-    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{fib}_{\text{snd}(\mathit{xs})}(x)
-    && \text{eqn.~\ref{container-membership} } (\in)
-    \\
-    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \text{sp-surj}(\text{snd}(\mathit{xs}))
-    && \text{eqn.~\ref{sp-surj-eqn} (sp-surj)}
-    \\
-    & \simeq \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket (A)) , \text{sp-surj}(\text{snd}(\mathit{xs}))
-    && \text{def.~\ref{List} } (\mathbf{List})
-    \\
-    & \simeq \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \mathbf{Fin}(n)) , A) , \text{sp-surj}(\text{snd}(\mathit{xs}))
-    && \text{eqn.~\ref{container-interp} } (\llbracket \wc \rrbracket)
-    \\
-    & \simeq \Sigma (n : \mathbb{N}) , \Sigma (f : \mathbf{Fin}(n) \rightarrow A) , \text{sp-surj}(f)
-    && \text{Reassociation of } \Sigma
-    \\
-    & \simeq \Sigma (n : \mathbb{N}) , ( \mathbf{Fin}(n) \twoheadrightarrow ! \; A )
-    && \text{eqn.~\ref{sp-surj-arrow-eqn} } (\twoheadrightarrow!)
-  \end{align*}
-\end{proof}
-
-In our formalisation, the proof is a single line: most of the steps above are
-simple expansion of definitions.
-The only step which isn't definitional equality is the reassociation of
-\(\Sigma\).
-
-Split enumerability implies decidable equality on the underlying type.
-To prove this, we will make use of the following lemma, proven in the
-formalisation:
-\begin{lemma} \label{discrete-surj}
-  \begin{equation}
-    \frac{
-        A \twoheadrightarrow! \; B \; \; \; \text{Discrete}(A)
-      }{
-       \text{Discrete}(B) 
-      }
-  \end{equation}
-\end{lemma}
-\begin{lemma} \label{split-enum-discrete}
-  Every split enumerable type is discrete.
-\end{lemma}
-\begin{proof}
-  Let \(A\) be a split enumerable type.
-  By lemma~\ref{split-enum-is-split-surj}, there is a surjection from
-  \(\mathbf{Fin}(n)\) for some \(n\).
-  Also, we know that \(\mathbf{Fin}(n)\) is discrete (proven in our
-  formalisation).
-  Therefore, by lemma~\ref{discrete-surj}, \(A\) is discrete.
-\end{proof}
-\subsection{Manifest Bishop Finiteness}
-\begin{definition}[Manifest Bishop Finiteness]
-  \begin{equation} \label{bish-def}
-    \mathcal{B}(A) \coloneqq \Sigma {(\mathit{xs} : \textbf{List}(A))} , \Pi {(x : A)} , x \inunique xs
-  \end{equation}
-\end{definition}
-The only difference between manifest Bishop finiteness and split enumerability
-is the membership term: here we require unique membership (\(\inunique\)),
-rather than simple membership (\(\in\)).
-\begin{definition}[Unique Membership] \label{uniq-memb-def}
-  \begin{equation}
-    x \inunique \mathit{xs} \coloneqq \text{isContr}(x \in \mathit{xs})
-  \end{equation}
-\end{definition}
-
-We use the word ``manifest'' here to distinguish from another common
-interpretation of Bishop finiteness, which we have called cardinal finiteness in
-this paper.
-The ``manifest'' refers to the fact that we have a concrete, non-truncated list
-of the elements in the proof.
-
-Where split enumerability was the enumeration form of a split surjection from
-\(\mathbf{Fin}\), manifest Bishop finiteness is the enumeration form of an
-\emph{equivalence} with \(\mathbf{Fin}\).
-\begin{lemma} \label{bishop-equiv}
-  \begin{equation}
-    \mathcal{B}(A) \simeq \Sigma {(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \simeq A \right)
-  \end{equation}
-\end{lemma}
-This proof is effectively the same as that of
-lemma~\ref{split-enum-is-split-surj}.
-\subsection{Cardinal Finiteness}
-Each finiteness predicate so far has contained an \emph{ordering} of the
-underlying type.
-For our purposes, this is too much information: it means that when constructing
-the ``category of finite sets'' later on, instead of each type having one
-canonical representative, it will have \(n!\), where \(n\) is the cardinality of
-the type\footnotemark.
-
-\footnotetext{
-  We actually do get a category (a groupoid, even) from manifest Bishop
-  finiteness \cite{yorgeyCombinatorialSpeciesLabelled2014}: it's the groupoid of
-  finite sets equipped with a linear order, whose morphisms are order-preserving
-  bijections.
-  We do not explore this particular construction in any detail.
-}
-
-To remedy the problem, we will use propositional truncation
-(def.~\ref{prop-trunc}).
-\begin{definition}[Cardinal Finiteness]
-  \begin{equation}
-    \mathcal{C}(A) \coloneqq \lVert \mathcal{B}(A) \rVert \simeq \lVert \Sigma(n : \mathbb{N}) , (\mathbf{Fin}(n) \simeq A) \rVert
-  \end{equation}
-\end{definition}
-At first glance, it might seem that we lose any useful properties we could
-derive from \(\mathcal{B}\).
-Luckily, this is not the case: by eliminator \ref{elim-prop-coh} of
-def.~\ref{prop-trunc}, we need only show that the output is uniquely determined.
-
-The following two lemmas are proven in
-\cite{yorgeyCombinatorialSpeciesLabelled2014} (Proposition 2.4.9 and 2.4.10,
-respectively), in much the same way as we have done here.
-Our contribution for this section is simply the formalisation.
-\begin{lemma}
-  Given a cardinally finite type, we can derive the type's cardinality, as well
-  as a propositionally truncated proof of equivalence with \(\textbf{Fin}\)s of
-  the same cardinality.
-  \begin{equation}
-    \mathcal{C}(A) \rightarrow \Sigma {(n : \mathbb{N})} , \lVert \textbf{Fin}(n) \simeq A \rVert
-  \end{equation}
-\end{lemma}
-\begin{lemma} \label{cardinal-finite-discrete}
-  Any cardinal-finite set has decidable equality.
-\end{lemma}
-\subsection{Manifest Enumerability}
-\begin{definition}[Manifest Enumerability]
-  \begin{equation}
-    \mathcal{E}(A) \coloneqq \Sigma {(\mathit{xs} : \mathbf{List}(A))} , \Pi {(x : A)} , \lVert x \in \mathit{xs} \rVert
-  \end{equation}
-\end{definition}
-As with manifest Bishop finiteness, the only difference with this type and split
-enumerability is the membership proof: here we have propositionally truncated
-it.
-This has two effects.
-First, it means that this proof represents a true surjection (rather than a
-split surjection) from \(\mathbf{Fin}\).
-\begin{lemma}
-  \begin{equation}
-    \mathcal{E}(A) \simeq \Sigma(n : \mathbb{N}) , (\mathbf{Fin}(n) \twoheadrightarrow A)
-  \end{equation}
-\end{lemma}
-The proof for this lemma is similar in structure to
-lemma~\ref{split-enum-is-split-surj} and lemma~\ref{bishop-equiv}.
-
-Secondly, it means the predicate does not imply decidable equality.
-More significantly, it allows the predicate to be defined over non-set types,
-like the circle.
-\begin{definition}[\(S^1\)] \label{circle-def}
-  The circle, \(S^1\), can be represented in HoTT as a higher inductive type.
-  \begin{equation}
-    \begin{alignedat}{3}
-      S^1 \coloneqq & \; \text{base} &&: S^1 ; \\
-      | & \; \text{loop} &&: \text{base} \equiv \text{base} ; 
-    \end{alignedat}
-  \end{equation}
-  We will use it here as an example of a non-set type, i.e. a type for which not
-  all paths are equal.
-  This also means that it does not have decidable equality.
-\end{definition}
-\begin{lemma}
-  The circle \(S^1\) is manifestly enumerable.
-\end{lemma}
-\subsection{Kuratowski Finiteness}
-The first thing we must define is a representation of subsets.
-\begin{definition}[Kuratowski Finite Subset]
-  \(\mathcal{K}(A)\) is the type of Kuratowski-finite subsets of \(A\).
-  \begin{equation}
-    \begin{alignedat}{3}
-      \mathcal{K}(A) \coloneqq&
-             \; []                &:& \; \mathcal{K}(A) ; \\
-      \vert& \; \wc \dblcolon \wc &:& \; A \rightarrow \mathcal{K}(A) \rightarrow \mathcal{K}(A) ; \\
-      \vert& \; \text{com}        &:& \; \Pi (x, y: A) , \Pi (\mathit{xs} : \mathcal{K}(A)) , \\
-                                  &&& \; x \dblcolon y \dblcolon \mathit{xs} \equiv y \dblcolon x \dblcolon \mathit{xs} ; \\
-      \vert& \; \text{dup}        &:& \; \Pi (x : A) , \Pi (\mathit{xs} : \mathcal{K}(A)) , x \dblcolon x \dblcolon \mathit{xs} \equiv x \dblcolon \mathit{xs} ; \\
-      \vert& \; \text{trunc}      &:& \; \Pi (\mathit{xs}, \mathit{ys}: \mathcal{K}(A)) , \Pi (p, q : \mathit{xs} \equiv \mathit{ys}) , p \equiv q ;
-    \end{alignedat}
-  \end{equation}
-  We define it as a HIT (definition~\ref{HITs}).
-  The first two constructors are point constructors, giving ways to create
-  values of type \(\mathcal{K}(A)\).
-  They are also recognisable as the two constructors for finite lists, a type
-  which represents the free monoid.
-
-  The next two constructors add extra paths to the type: equations that usage of
-  the type must obey.
-  These extra paths turn the free monoid into the free \emph{commutative} (com)
-  \emph{idempotent} (dup) monoid.
-
-  The final constructor enforces that the type \(\mathcal{K}(A)\) must be a set.
-\end{definition}
-The Kuratowski finite subset is a free join semilattice (or, equivalently, a
-free commutative idempotent monoid).
-More prosaically, \(\mathcal{K}\) is the abstract data type for finite sets, as
-defined in the Boom hierarchy \cite{boomFurtherThoughtsAbstracto1981,
-  bunkenburgBoomHierarchy1994}.
-However, rather than just being a specification, \(\mathcal{K}\) is fully usable
-as a data type in its own right, thanks to HITs.
-
-Other definitions of \(\mathcal{K}\) exist (such as the one in
-\cite{fruminFiniteSetsHomotopy2018}) which make the fact that \(\mathcal{K}\) is
-the free join semilattice more obvious.
-We have included such a definition in our formalisation, and proven it
-equivalent to the one above.
-
-Next, we need a way to say that an entire type is Kuratowski finite.
-For that, we will need to define membership of \(\mathcal{K}\).
-\begin{definition}[Membership of \(\mathcal{K}\)]
-  Membership is defined by pattern-matching on \(\mathcal{K}\).
-  The two point constructors are handled like so:
-  \begin{equation}
-    \begin{alignedat}{2}
-      x \in&& \; []                      &\coloneqq \bot ; \\
-      x \in&& \; y \dblcolon \mathit{ys} &\coloneqq \lVert x \equiv y \uplus x \in \mathit{ys} \rVert ;
-    \end{alignedat}
-  \end{equation}
-  The \(\text{com}\) and \(\text{dup}\) constructors are handled by proving that
-  the truncated form of \(\uplus\) is itself commutative and idempotent.
-  The type of propositions is itself a set, satisfying the \(\text{trunc}\)
-  constructor.
-\end{definition}
-Finally, we have enough background to define Kuratowski finiteness.
-\begin{definition}[Kuratowski Finiteness]
-  \begin{equation}
-    \mathcal{K}^{f}(A) = \Sigma {(\mathit{xs} : \mathcal{K}(A))} , \Pi (x : A) , x \in \mathit{xs}
-  \end{equation}
-\end{definition}
-
-We also have the following two lemmas, proven in both
-\cite{fruminFiniteSetsHomotopy2018} and our formalisation.
-\begin{lemma}
-  \(\mathcal{K}^f\) is a mere proposition.
-\end{lemma}
-\begin{lemma}
-  This circle \(S^1\) is Kuratowski finite.
-\end{lemma}
-\section{Relations Between Each Finiteness Definition} \label{relations}
-We will now look at the arrows in figure~\ref{finite-classification}.
-\subsection{Split Enumerability and Manifest Bishop Finiteness}
-While manifest Bishop finiteness might seem stronger than split enumerability,
-it turns out this is not the case.
-Both predicates imply the other.
-\begin{lemma} \label{manifest-bishop-to-split-enum}
-  Any manifest Bishop finite type is split enumerable.
-\end{lemma}
-\begin{proof}
-  To construct a proof of split enumerability from one of manifest Bishop
-  finiteness, it suffices to convert a proof of \(x \inunique \mathit{xs}\) to
-  one of \(x \in \mathit{xs}\), for all \(x\) and \(\mathit{xs}\).
-  Since \(\inunique\) is defined as a contraction of \(\in\), such a conversion
-  is simply the \(\text{fst}\) function.
-\end{proof}
-
-\begin{lemma} \label{split-enum-to-manifest-bishop}
-  Any split enumerable set is manifest Bishop finite.
-\end{lemma}
-This proof takes significantly more work.
-The ``unique membership'' condition in \(\mathcal{B}\) means that we are not
-permitted duplicates in the support list.
-The first step in the proof, then, is to filter those duplicates out from the
-support list of the \(\mathcal{E}!\) proof: we can do this using the decidable
-equality provided by \(\mathcal{E}!\) (lemma~\ref{split-enum-discrete}).
-From there, all we need to show is that the membership proof carries over
-appropriately.
-\subsection{Split Enumerability and Manifest Enumerability}
-\begin{lemma} \label{split-enum-to-manifest-enum}
-  Any split enumerable type is manifestly enumerable.
-\end{lemma}
-This lemma is proven by truncating the membership proof in split enumerability.
-
-\begin{lemma} \label{manifest-enum-to-split-enum}
-  A manifestly enumerable type with decidable equality is split enumerable.
-\end{lemma}
-
-\subsection{Manifest Bishop Finiteness and Cardinal Finiteness}
-\begin{lemma} \label{manifest-bishop-to-cardinal}
-  Any manifest Bishop finite type is cardinal finite.
-\end{lemma}
-\begin{theorem} \label{cardinal-to-manifest-bishop}
-  Any cardinal finite type with a total order is Bishop finite.
-\end{theorem}
-The proof for this particular theorem is quite involved in the formalisation, so
-we only give its sketch here.
-First, note that we actually convert to manifest enumerability first: this can
-be converted to split enumerability with decidable equality, which is provided
-by cardinal finiteness.
-
-Next, we define permutations.
-\begin{definition}[List Permutations]
-  Two lists are permutations of each other if their membership proofs are all
-  equivalent\footnotemark \cite{danielssonBagEquivalenceProofRelevant2012}.
-  \begin{equation}
-    \mathit{xs} \leftrightsquigarrow \mathit{ys} = \Pi {(x : A)} , x \in \mathit{xs} \simeq x \in \mathit{ys}
-  \end{equation}
-\end{definition}
-
-\footnotetext{
-  The definition in \cite{danielssonBagEquivalenceProofRelevant2012} and our
-  formalisation is slightly different: we say permutations are lists with
-  \emph{isomorphic} membership proofs.
-  The distinction, as it happens, does not affect our work here.
-}
-
-Next, we define a sort function which relies on the provided total order.
-We further prove the following fact about this sort function:
-\begin{equation}
-  \Pi(\mathit{xs}, \mathit{ys} : \mathbf{List}(A)) , \mathit{xs} \leftrightsquigarrow \mathit{ys} \rightarrow \text{sort}(\mathit{xs}) \equiv \text{sort}(\mathit{ys})
-\end{equation}
-
-Next, notice that the support lists of any two proofs of manifest Bishop
-finiteness must be permutations of each other.
-This will allow us to sort the support list of a proof of cardinal finiteness in
-a coherently constant (definition~\ref{prop-trunc},
-eliminator~\ref{elim-prop-coh}) way, pulling the support list out from the
-truncation.
-The cover proof emerges naturally from the definition of the permutation.
-\subsection{Cardinal Finiteness and Kuratowski Finiteness}
-\begin{lemma} \label{cardinal-kuratowski}
-  \begin{equation}
-    \mathcal{C}(A) \simeq \mathcal{K}^f(A) \times \text{Discrete}(A)
-  \end{equation}
-\end{lemma}
-This proof is constructed by providing a pair of functions: one from
-\(\mathcal{C}(A)\) to \(\mathcal{K}^f(A) \times \text{Discrete}(A)\), and one the
-other way.
-This pair implies an equivalence, because both source and target are
-propositions.
-The actual functions themselves are proven in our formalisation, as well as in
-\cite{fruminFiniteSetsHomotopy2018}.
-\section{Topos} \label{topos}
-In this section we will prove that decidable Kuratowski finite types form a
-\(\Pi\)-pretopos.
-Along the way we will provide closure proofs for a number of the other
-finiteness predicates.
-As we saw in theorem~\ref{cardinal-kuratowski}, decidable Kuratowski finite
-types are equivalent to cardinal finite types, so we will work with the latter
-from now on.
-Our first task is to show that cardinal finite types are closed under several
-operations.
-\subsection{Closure}
-For the first three closure proofs, we only consider split enumerability:
-as it is the strongest of the finiteness predicates, we can derive the other
-closure proofs from it.
-
-\begin{lemma}
-  \(\bot\), \(\top\), and \(\mathbf{Bool}\) are split enumerable.
-\end{lemma}
-\begin{lemma} \label{split-enum-sigma}
-  Split enumerability is closed under \(\Sigma\).
-  \begin{equation}
-    \frac{
-      \mathcal{E}!(A) \; \; \; \Pi(x : A) , \mathcal{E}!(U(x))
-    }{
-      \mathcal{E}!(\Sigma(x : A) , U(x))
-    }
-  \end{equation}
-\end{lemma}
-From this we can derive split enumerability of non-dependent sums and products,
-as both are definable in terms of \(\Sigma\).
-
-\begin{lemma}
-  Split enumerability is closed under dependent functions.
-  (\(\Pi\)-types).
-  \begin{equation}
-    \frac{
-      \mathcal{E}!(A) \; \; \; \Pi {(x : A)} , \mathcal{E}!\left( U(x) \right)
-    }{
-      \mathcal{E}!\left(\Pi {(x : A)} , U(x)\right)
-    }
-  \end{equation}
-\end{lemma}
-\begin{proof}
-  Let \(A\) be a split enumerable type, and \(U\) be a type family from \(A\),
-  which is split enumerable over all points of \(A\).
-
-  As \(A\) is split enumerable, we know that it is also manifestly Bishop finite
-  (lemma~\ref{split-enum-to-manifest-bishop}), and consequently we know \(A
-  \simeq \mathbf{Fin}(n)\), for some \(n\) (lemma~\ref{bishop-equiv}).
-  We can therefore replace all occurrences of \(A\) with \(\mathbf{Fin}(n)\),
-  changing our goal to:
-  \begin{equation}
-    \frac{
-      \mathcal{E}!(\mathbf{Fin}(n)) \; \; \; \Pi (x : \mathbf{Fin}(n)) , \mathcal{E}!\left( U(x) \right)
-    }{
-      \mathcal{E}!\left(\Pi (x : \mathbf{Fin}(n)) , U(x)\right)
-    }
-  \end{equation}
-
-  We then define the type of \(n\)-tuples over some type family \(T :
-  \mathbf{Fin}(n) \rightarrow \mathbf{Type}\).
-  \begin{equation}
-    \begin{alignedat}{3}
-      & \mathbf{Tuple}(0, T)   &&\coloneqq \top \\
-      & \mathbf{Tuple}(n+1, T) &&\coloneqq T(0) \times \mathbf{Tuple}(n, T \circ \text{suc})
-    \end{alignedat}
-  \end{equation}
-  We can show that this type is equivalent to functions (proven in our formalisation):
-  \begin{equation}
-    \Pi(x : \mathbf{Fin}(n)) , U(x) \simeq \mathbf{Tuple}(n, U)
-  \end{equation}
-  And therefore we can simplify again our goal to the following:
-  \begin{equation}
-    \frac{
-      \mathcal{E}!(\mathbf{Fin}(n)) \; \; \; \Pi (x : \mathbf{Fin}(n)) , \mathcal{E}!\left( U(x) \right)
-    }{
-      \mathcal{E}!\left(\mathbf{Tuple}(n, U)\right)
-    }
-  \end{equation}
-
-  We can prove this goal by showing that \(\mathbf{Tuple}(n, U)\) is split
-  enumerable: it is made up of finitely many products of points of \(U\), which
-  are themselves split enumerable, and \(\top\), which is also split enumerable.
-  Lemma~\ref{split-enum-sigma} shows us that the product of finitely many split
-  enumerable types is itself split enumerable, proving our goal.
-\end{proof}
-
-Lifting the above closure proofs to work on proofs of \(\mathcal{C}\) is not
-necessarily straightforward.
-\(\lVert \wc \rVert\) forms a monad, giving us access to a powerful syntax for
-combining proofs.
-From this, we can derive closure under \(\bot\), \(\top\), \(\mathbf{Bool}\),
-\(\times\), \(\uplus\), and \(\rightarrow\).
-The dependent case is not so straightforward.
-We need the following lemma, a version of the axiom of choice on finite sets,
-to prove closure under \(\Sigma\) and \(\Pi\).
-\begin{lemma}
-  \begin{equation}
-    \mathcal{C}(A) \rightarrow (\Pi(x : A) , \lVert U(x) \rVert) \rightarrow \lVert \Pi(x : A) , U(x) \rVert
-  \end{equation}
-\end{lemma}
-\subsection{The Category of Finite Sets}
-HoTT and CuTT seem to be especially suitable settings for formalisations of
-category theory.
-The univalence axiom in particular allows us to treat categorical isomorphisms
-as equalities, saving us from the dreaded ``setoid hell''.
-
-We follow \cite[chapter 9]{hottbook} in its treatment of
-categories in HoTT, and in its proof that sets do indeed form a category.
-We will first briefly go through the construction of the category
-\(\mathit{Set}\), as it differs slightly from the usual method in type theory.
-
-First, the type of objects and arrows:
-\begin{alignat}{3}
-  &\text{Obj}_\mathit{Set}      &&\coloneqq \Sigma(x : \mathbf{Type}) , \text{isSet}(x) \\
-  &\text{Hom}_\mathit{Set}(x , y) &&\coloneqq  \text{fst}(x) \rightarrow \text{fst}(y)
-\end{alignat}
-As the type of objects makes clear, we have already departed slightly from the
-simpler \(\text{Obj}_\mathit{Set} \coloneqq \mathbf{Type}\) way of doing things:
-of course we have to, as HoTT allows non-set types.
-Furthermore, after proving the usual associativity and identity laws for
-composition (which are definitionally true in this case), we must further show
-\(\text{isSet}(\text{Hom}_\mathit{Set}(x,y))\); even then we only have a
-precategory.
-
-To show that \(\mathit{Set}\) is a category, we must show that categorical
-isomorphisms are equivalent to equivalences.
-In a sense, we must give a univalence rule for the category we are working in.
-
-We have provided formal proofs that \(\mathit{Set}\) does indeed form a
-category, and the following:
-\begin{theorem}[The Category of Finite Sets]
-  Finite sets form a category in HoTT when defined like so:
-  \begin{equation}
-    \begin{alignedat}{3}
-      &\text{Obj}_\mathit{FinSet}      &&\coloneqq \Sigma(x : \mathbf{Type}) , \mathcal{C}(x) \\
-      &\text{Hom}_\mathit{FinSet}(x , y) &&\coloneqq  \text{fst}(x) \rightarrow \text{fst}(y)
-    \end{alignedat}
-  \end{equation}
-\end{theorem}
-\subsection{The \(\Pi\)-pretopos of Finite Sets}
-For this proof, we follow again the proof that \(\mathit{Set}\) forms a \(\Pi
-W\)-pretopos from \cite[chapter 10]{hottbook} and
-\cite{rijkeSetsHomotopyType2015}.
-The difference here is that clearly we do not have access to \(W\)-types, as
-they would permit infinitary structures.
-
-We first must show that \(\mathit{Set}\) has an initial object and finite,
-disjoint sums, which are stable under pullback.
-We also must show that \(\mathit{Set}\) is a regular category with effective
-quotients.
-We now have a pretopos: the presence of \(\Pi\) types make it a
-\(\Pi\)-pretopos.
-
-We have proven the above statements for both \(\mathit{Set}\) and
-\(\mathit{FinSet}\).
-As far as we know, this is the first formalisation of either.
-\begin{theorem} \label{finite-topos}
-  The category of finite sets, \(\mathit{FinSet}\), forms a \(\Pi\)-pretopos.
-\end{theorem}
-\section{Countably Infinite Types} \label{infinite}
-\begin{figure*}
-  \centering
-
-  \tikzcdset{
-  cramped,
-  every matrix/.append style={nodes={font=\scriptsize}},
-  row    sep/normal=1em,
-  column sep/normal=1em,
-  }
-  \begin{subfigure}[b]{.45\textwidth}
-    \centering
-    \begin{tikzcd}
-      (1,e) \ar[rdddd, out=-45, in=135] & (2,e) \ar[rdddd, out=-45, in=135] & (3,e) \ar[rdddd, out=-45, in=135] & (4,e) \ar[rdddd, out=-45, in=135] & (5,e)        \\
-      (1,d) \ar[u]     & (2,d) \ar[u]    & (3,d) \ar[u]    & (4,d) \ar[u]    & (5,d) \ar[u] \\
-      (1,c) \ar[u]     & (2,c) \ar[u]    & (3,c) \ar[u]    & (4,c) \ar[u]    & (5,c) \ar[u] \\
-      (1,b) \ar[u]     & (2,b) \ar[u]    & (3,b) \ar[u]    & (4,b) \ar[u]    & (5,b) \ar[u] \\
-      (1,a) \ar[u]     & (2,a) \ar[u]    & (3,a) \ar[u]    & (4,a) \ar[u]    & (5,a) \ar[u]
-    \end{tikzcd}
-    \caption{Depth-First}
-    \label{depth-first}
-  \end{subfigure} \hfill
-  \begin{subfigure}[b]{.45\textwidth}
-    \centering
-    \begin{tikzcd}
-      (1,e) \ar[dr] & (2,e) \ar[dr]  & (3,e) \ar[dr]    & (4,e) \ar[dr] & (5,e) \\
-      (1,d) \ar[dr] & (2,d) \ar[dr]  & (3,d) \ar[dr]    & (4,d) \ar[dr] & (5,d) \ar[u] \\
-      (1,c) \ar[dr] & (2,c) \ar[dr]  & (3,c) \ar[dr]    & (4,c) \ar[dr] & (5,c) \ar[uul] \\
-      (1,b) \ar[dr] & (2,b) \ar[dr]  & (3,b) \ar[dr]    & (4,b) \ar[dr] & (5,b) \ar[uuull, out=130, in=-50] \\
-      (1,a) \ar[u]  & (2,a) \ar[uul, out=130, in=-50] & (3,a) \ar[uuull, out=130, in=-50] & (4,a) \ar[uuuulll, out=130, in=-50] & (5,a) \ar[uuuulll, out=130, in=-50]
-    \end{tikzcd}
-    \caption{Breadth-First}
-    \label{breadth-first}
-  \end{subfigure}
-  \caption{Two possible products for the sets \(\left[ 1 \dots 5 \right]\) and
-    \(\left[  a \dots e \right]\)}
-  \label{pairings}
-\end{figure*}
-In the previous sections we saw different flavours of finiteness which were
-really just different flavours of relations to \(\mathbf{Fin}\).
-In this section we will see that we can construct a similar classification of
-relations to \(\mathbb{N}\), in the form of the countably infinite types.
-\subsection{Two Countable Types}
-The two types for countability we will consider are analogous to split
-enumerability and cardinal finiteness.
-The change will be a simple one: we will swap out lists for streams.
-\begin{definition}[Streams]
-  \begin{equation}
-    \mathbf{Stream}(A) \coloneqq (\mathbb{N} \rightarrow A)
-    \simeq \llbracket \top , \text{const}(\mathbb{N}) \rrbracket
-  \end{equation}
-\end{definition}
-\begin{definition}[Split Countability]
-  \begin{equation}
-    \aleph_0!(A) \coloneqq \Sigma {(\mathit{xs} : \mathbf{Stream}(A))} , \Pi {(x : A)} , x \in \mathit{xs}
-  \end{equation}
-\end{definition}
-This type is definitionally equal to it surjection equivalent (\(\mathbb{N}
-\twoheadrightarrow ! \; A\)).
-We construct the unordered, propositional version of the predicate in much the
-same way as we constructed cardinal finiteness.
-\begin{definition}[Countability]
-  \begin{equation}
-    \aleph_0(A) \coloneqq \lVert \aleph_0!(A) \rVert
-  \end{equation}
-\end{definition}
-
-From both of these types we can derive decidable equality.
-\begin{lemma}
-  Any countable type has decidable equality.
-\end{lemma}
-\subsection{Closure}
-We know that countable infinity is not closed under the exponential (function
-arrow), so the only closure we need to prove is \(\Sigma\) to cover all of
-what's left.
-\begin{theorem} \label{split-countability-sigma}
-  Split countability is closed under \(\Sigma\).
-\end{theorem}
-We know that countable infinity is not closed under the exponential (function
-arrow), so the only closure we need to prove is \(\Sigma\) to cover all of
-what's left.
-To do this we have to take a slightly different approach to the functions we
-defined before.
-Figure~\ref{pairings} illustrates the reason why: previously, we used the
-depth-first product pairing for each support list.
-This diverges if the first list is infinite, never exploring anything other than
-the first element in the second list.
-Instead, we use here the cantor pairing function, which performs a breadth-first
-search of the pairings of both lists.
-
-Finally, while we have lost certain closure proofs by allowing for infinite
-types, we also \emph{gain} some: in particular the Kleene star.
-\begin{theorem}
-  Split countability is closed under Kleene star.
-  \begin{equation}
-    \aleph_0!(A) \rightarrow \aleph_0!(\mathbf{List}(A))
-  \end{equation}
-\end{theorem}
-Again, this proof requires a particular pattern to ensure productivity.
-The pattern here builds an intermediate stream \(\mathcal{KV}\) of non-empty
-lists from the input support stream \(\mathit{xs}\), which is subsequently
-flattened.
-\begin{equation}
-  \mathcal{KV}_i \coloneqq \left[ \left[ \mathit{xs}_{j - 1} \mid j \in \mathit{js} \right] \mid \mathit{js} \in \mathbf{List}(\mathbb{N}) ; \text{sum}(\mathit{js}) = i ; 0 \notin \mathit{js}  \right]
-\end{equation}
-\section{Search} \label{search}
-\subsection{Omniscience}
-\begin{definition}[Limited Principle of Omniscience]
-  For any type \(A\) and predicate \(P\) on \(A\), the limited principle of
-  omniscience \cite{myhillErrettBishopFoundations1972} is as follows:
-  \begin{equation}
-    \left( \Pi {(x : A)} , \mathbf{Dec}(P(x)) \right) \rightarrow \mathbf{Dec} \left( \Sigma {(x : A)} , P(x) \right)
-  \end{equation}
-  In other words, for any decidable predicate the existential quantification of
-  that predicate is also decidable.
-\end{definition}
-The limited principle of omniscience is non-constructive, but individual types
-can themselves satisfy omniscience.
-In particular, cardinal finite types are omniscient.
-
-There is also a universal form of omniscience, which we call exhaustibility.
-\begin{definition}[Exhaustibility]
-  We say a type \(A\) is exhaustible if, for any decidable predicate \(P\) on
-  \(A\), the universal quantification of the predicate is decidable.
-  \begin{equation}
-    \left( \Pi {(x : A)} , \mathbf{Dec}(P(x)) \right) \rightarrow \mathbf{Dec} \left( \Pi {(x : A)} , P(x) \right)
-  \end{equation}
-\end{definition}
-
-All of the finiteness predicates we have seen imply exhaustibility.
-Omniscience is stronger than exhaustibility, as we can derive the latter from
-the former.
-All of the ordered finiteness predicates imply omniscience.
-For the unordered finiteness definitions, we have omniscience for prop-valued
-predicates.
-\subsection{Automating Proofs}
-One use for above constructions is the automation of certain proofs.
-In \cite{firsovDependentlyTypedProgramming2015}, which uses a similar approach
-to ours, the \(\AgdaDatatype{Pauli}\) group is used as an example.
-\ExecuteMetaData[agda/Data/Pauli.tex]{def}
-As \(\AgdaDatatype{Pauli}\) has 4 constructors, \(n\)-ary functions on
-\(\AgdaDatatype{Pauli}\) may require up to \(4^n\) cases, making even simple
-proofs prohibitively verbose.
-
-The alternative is to derive the things we need from omniscience, itself derived
-from a finiteness predicate.
-For proof search, the procedure is a well-known one in Agda
-\cite{devrieseBrightSideType2011}: we ask for the result of a decision procedure
-as an \emph{instance argument}, which will demand computation during
-typechecking.
-Our addition to this technique is a way to handle multiple arguments based on
-fully level-polymorphic dependent currying and uncurrying, building on 
-\cite{allaisGenericLevelPolymorphic2019}.
-\ExecuteMetaData[agda/Data/Pauli.tex]{assoc-prf}
-
-Finally, we can derive decidable equality on functions over finite types.
-We can also use functions in our proof search.
-Here, for instance, is an automated procedure which finds the
-\(\AgdaFunction{not}\) function on \(\mathbf{Bool}\), given a specification.
-\ExecuteMetaData[agda/Data/Pauli.tex]{not-spec}
-\section{Related Work}
-The univalent foundations program is the main basis for this work
-\cite{hottbook}.
-In particular, our formalisation in section~\ref{topos} relied heavily on
-\cite[chapter 10]{hottbook}, and \cite{rijkeSetsHomotopyType2015}, a paper which
-contains much of the same material.
-
-Finite sets in a constructive setting has been studied extensively before:
-In \cite{coquandConstructivelyFinite2010} four separate predicates for
-finiteness were considered (split-enumerable being the only one explored in this
-work), and \cite{firsovVariationsNoetherianness2016} explores Noetherianness.
-\cite{firsovDependentlyTypedProgramming2015} explored what we have called split
-enumerability and manifest Bishop finiteness (although they are stated slightly
-differently), and they use these to build a library for proof search.
-In \cite{fruminFiniteSetsHomotopy2018} the topic of Kuratowski finite sets in
-HoTT is studied extensively: we have focused more on the non-truncated versions
-of finiteness (the ``manifest'' predicates), and we have provided the missing
-\(\Pi\)-pretopos proof of decidable Kuratowski finite sets.
-
-\cite{iversenUnivalentCategoriesFormalization2018} provided a starting point for
-our categorical formalisation: it contains a proof, for instance, that homotopy
-sets form a category.
 % \begin{acks}

### 2020-05-29 sections/countable-predicates.tex
split into sections

\section{Countably Infinite Types} \label{infinite}
\begin{figure*}
  \centering

  \tikzcdset{
  cramped,
  every matrix/.append style={nodes={font=\scriptsize}},
  row    sep/normal=1em,
  column sep/normal=1em,
  }
  \begin{subfigure}[b]{.45\textwidth}
    \centering
    \begin{tikzcd}
      (1,e) \ar[rdddd, out=-45, in=135] & (2,e) \ar[rdddd, out=-45, in=135] & (3,e) \ar[rdddd, out=-45, in=135] & (4,e) \ar[rdddd, out=-45, in=135] & (5,e)        \\
      (1,d) \ar[u]     & (2,d) \ar[u]    & (3,d) \ar[u]    & (4,d) \ar[u]    & (5,d) \ar[u] \\
      (1,c) \ar[u]     & (2,c) \ar[u]    & (3,c) \ar[u]    & (4,c) \ar[u]    & (5,c) \ar[u] \\
      (1,b) \ar[u]     & (2,b) \ar[u]    & (3,b) \ar[u]    & (4,b) \ar[u]    & (5,b) \ar[u] \\
      (1,a) \ar[u]     & (2,a) \ar[u]    & (3,a) \ar[u]    & (4,a) \ar[u]    & (5,a) \ar[u]
    \end{tikzcd}
    \caption{Depth-First}
    \label{depth-first}
  \end{subfigure} \hfill
  \begin{subfigure}[b]{.45\textwidth}
    \centering
    \begin{tikzcd}
      (1,e) \ar[dr] & (2,e) \ar[dr]  & (3,e) \ar[dr]    & (4,e) \ar[dr] & (5,e) \\
      (1,d) \ar[dr] & (2,d) \ar[dr]  & (3,d) \ar[dr]    & (4,d) \ar[dr] & (5,d) \ar[u] \\
      (1,c) \ar[dr] & (2,c) \ar[dr]  & (3,c) \ar[dr]    & (4,c) \ar[dr] & (5,c) \ar[uul] \\
      (1,b) \ar[dr] & (2,b) \ar[dr]  & (3,b) \ar[dr]    & (4,b) \ar[dr] & (5,b) \ar[uuull, out=130, in=-50] \\
      (1,a) \ar[u]  & (2,a) \ar[uul, out=130, in=-50] & (3,a) \ar[uuull, out=130, in=-50] & (4,a) \ar[uuuulll, out=130, in=-50] & (5,a) \ar[uuuulll, out=130, in=-50]
    \end{tikzcd}
    \caption{Breadth-First}
    \label{breadth-first}
  \end{subfigure}
  \caption{Two possible products for the sets \(\left[ 1 \dots 5 \right]\) and
    \(\left[  a \dots e \right]\)}
  \label{pairings}
\end{figure*}
In the previous sections we saw different flavours of finiteness which were
really just different flavours of relations to \(\mathbf{Fin}\).
In this section we will see that we can construct a similar classification of
relations to \(\mathbb{N}\), in the form of the countably infinite types.
\subsection{Two Countable Types}
The two types for countability we will consider are analogous to split
enumerability and cardinal finiteness.
The change will be a simple one: we will swap out lists for streams.
\begin{definition}[Streams]
  \begin{equation}
    \mathbf{Stream}(A) \coloneqq (\mathbb{N} \rightarrow A)
    \simeq \llbracket \top , \text{const}(\mathbb{N}) \rrbracket
  \end{equation}
\end{definition}
\begin{definition}[Split Countability]
  \begin{equation}
    \aleph_0!(A) \coloneqq \Sigma {(\mathit{xs} : \mathbf{Stream}(A))} , \Pi {(x : A)} , x \in \mathit{xs}
  \end{equation}
\end{definition}
This type is definitionally equal to it surjection equivalent (\(\mathbb{N}
\twoheadrightarrow ! \; A\)).
We construct the unordered, propositional version of the predicate in much the
same way as we constructed cardinal finiteness.
\begin{definition}[Countability]
  \begin{equation}
    \aleph_0(A) \coloneqq \lVert \aleph_0!(A) \rVert
  \end{equation}
\end{definition}

From both of these types we can derive decidable equality.
\begin{lemma}
  Any countable type has decidable equality.
\end{lemma}
\subsection{Closure}
We know that countable infinity is not closed under the exponential (function
arrow), so the only closure we need to prove is \(\Sigma\) to cover all of
what's left.
\begin{theorem} \label{split-countability-sigma}
  Split countability is closed under \(\Sigma\).
\end{theorem}
We know that countable infinity is not closed under the exponential (function
arrow), so the only closure we need to prove is \(\Sigma\) to cover all of
what's left.
To do this we have to take a slightly different approach to the functions we
defined before.
Figure~\ref{pairings} illustrates the reason why: previously, we used the
depth-first product pairing for each support list.
This diverges if the first list is infinite, never exploring anything other than
the first element in the second list.
Instead, we use here the cantor pairing function, which performs a breadth-first
search of the pairings of both lists.

Finally, while we have lost certain closure proofs by allowing for infinite
types, we also \emph{gain} some: in particular the Kleene star.
\begin{theorem}
  Split countability is closed under Kleene star.
  \begin{equation}
    \aleph_0!(A) \rightarrow \aleph_0!(\mathbf{List}(A))
  \end{equation}
\end{theorem}
Again, this proof requires a particular pattern to ensure productivity.
The pattern here builds an intermediate stream \(\mathcal{KV}\) of non-empty
lists from the input support stream \(\mathit{xs}\), which is subsequently
flattened.
\begin{equation}
  \mathcal{KV}_i \coloneqq \left[ \left[ \mathit{xs}_{j - 1} \mid j \in \mathit{js} \right] \mid \mathit{js} \in \mathbf{List}(\mathbb{N}) ; \text{sum}(\mathit{js}) = i ; 0 \notin \mathit{js}  \right]
\end{equation}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../paper"
%%% End:

### 2020-05-29 sections/finiteness-predicates.tex
split into sections

\section{Finiteness Predicates} \label{finiteness-predicates}
In this section, we will define and briefly describe each of the five predicates
in Figure~\ref{finite-classification}.
The reason we explore predicates other than our focus (decidable Kuratowski
finiteness) is that we can often prove things like closure much more readily on
the simpler predicates.
The relations (which we will prove in the next section) then allow us to
transfer those proofs.
\subsection{Split Enumerability}
\begin{definition}[Split Enumerable Set] \label{split-enum-def}
  \begin{equation} \label{split-enum-def-eqn}
    \mathcal{E}!(A) \coloneqq \Sigma {(\mathit{xs} : \textbf{List}(A))} , \Pi {(x : A)} , x \in \mathit{xs}
  \end{equation}
  We call the first component of this pair the ``support'' list, and the second
  component the ``cover'' proof.
  An equivalent version of this predicate was called \verb+Listable+ in
  \cite{firsovDependentlyTypedProgramming2015}.
\end{definition}

We used some extra types in the above definition, which we will define here:
\begin{definition}[Containers] \label{container-def}
  A container \cite{abbottContainersConstructingStrictly2005} is a pair
  \(S , P\) where \(S\) is a type, the elements of which are called
  the \emph{shapes} of the container, and \(P\) is a type family on \(S\), where
  the elements of \(P(s)\) are called the \emph{positions} of a container.
  We ``interpret'' a container into a functor defined like so:
  \begin{equation} \label{container-interp}
    \llbracket S , P \rrbracket(A) \coloneqq \Sigma {(s : S)} , \left( P(s) \rightarrow A \right)
  \end{equation}
  Membership of a container can be defined like so:
  \begin{equation} \label{container-membership}
    x \in \mathit{xs} \coloneqq \text{fib}_{\text{snd}(\mathit{xs})}(x)
  \end{equation}
\end{definition}
\begin{definition}[\(\mathbf{List}\)] \label{List}
  \begin{equation}
    \mathbf{List} \coloneqq \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket
  \end{equation}
\end{definition}
\begin{definition}[\(\mathbf{Fin}\)] \label{Fin}
  \(\mathbf{Fin}(n)\) is the type of natural numbers smaller than \(n\). We
  define it the standard way:
  \begin{equation}
    \begin{alignedat}{3}
      & \mathbf{Fin}(0)     && \coloneqq \bot ; \\
      & \mathbf{Fin}(n + 1) && \coloneqq \top \uplus \mathbf{Fin}(n) ;
    \end{alignedat}
  \end{equation} 
\end{definition}

We tend to prefer list-based definitions of finiteness, rather than ones based
on bijections or surjections.
This is purely a matter of perspective, however: the definition above is
precisely equivalent to a split surjection from a finite prefix of the natural
numbers.
\begin{definition}[Surjections] \label{surjections}
  We define both surjections and \emph{split} surjections here \cite[definition
  4.6.1]{hottbook}.
  \begin{alignat}{3}
    &\text{surj}(f)             &&\coloneqq \Pi(y : B) , \lVert \text{fib}_f(y) \rVert \\
    &A \twoheadrightarrow B     &&\coloneqq \Sigma (f : A \rightarrow B) , \text{surj}(f) \label{surj-arrow-eqn} \\
    &\text{sp-surj}(f)          &&\coloneqq \Pi(y : B) , \text{fib}_f(y) \label{sp-surj-eqn} \\
    &A \twoheadrightarrow! \; B &&\coloneqq \Sigma (f : A \rightarrow B) , \text{sp-surj}(f) \label{sp-surj-arrow-eqn}
  \end{alignat}
\end{definition}
\begin{lemma} \label{split-enum-is-split-surj}
  \begin{equation}
    \mathcal{E}!(A) \simeq \Sigma (n : \mathbb{N}) , \left( \mathbf{Fin}(n) \twoheadrightarrow ! \; A \right)
  \end{equation}
\end{lemma}
\begin{proof}
  \begin{align*}
     \mathcal{E}!(A) &
    \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , x \in \mathit{xs}
    && \text{def.~\ref{split-enum-def} }(\mathcal{E}!)
    \\
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{fib}_{\text{snd}(\mathit{xs})}(x)
    && \text{eqn.~\ref{container-membership} } (\in)
    \\
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \text{sp-surj}(\text{snd}(\mathit{xs}))
    && \text{eqn.~\ref{sp-surj-eqn} (sp-surj)}
    \\
    & \simeq \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket (A)) , \text{sp-surj}(\text{snd}(\mathit{xs}))
    && \text{def.~\ref{List} } (\mathbf{List})
    \\
    & \simeq \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \mathbf{Fin}(n)) , A) , \text{sp-surj}(\text{snd}(\mathit{xs}))
    && \text{eqn.~\ref{container-interp} } (\llbracket \wc \rrbracket)
    \\
    & \simeq \Sigma (n : \mathbb{N}) , \Sigma (f : \mathbf{Fin}(n) \rightarrow A) , \text{sp-surj}(f)
    && \text{Reassociation of } \Sigma
    \\
    & \simeq \Sigma (n : \mathbb{N}) , ( \mathbf{Fin}(n) \twoheadrightarrow ! \; A )
    && \text{eqn.~\ref{sp-surj-arrow-eqn} } (\twoheadrightarrow!)
  \end{align*}
\end{proof}
\begin{agdalisting}
In our formalisation, the proof is a single line:
\ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{is-split-inj}
The only step which isn't definitional equality is the reassociation of
\(\Sigma\).
\ExecuteMetaData[agda/Data/Sigma/Properties.tex]{reassoc}
\end{agdalisting}

Split enumerability implies decidable equality on the underlying type.
To prove this, we will make use of the following lemma, proven in the
formalisation:
\begin{lemma} \label{discrete-surj}
  \begin{equation}
    \frac{
        A \twoheadrightarrow! \; B \; \; \; \text{Discrete}(A)
      }{
       \text{Discrete}(B) 
      }
  \end{equation}
\end{lemma}
\begin{lemma} \label{split-enum-discrete}
  Every split enumerable type is discrete.
\end{lemma}
\begin{proof}
  Let \(A\) be a split enumerable type.
  By lemma~\ref{split-enum-is-split-surj}, there is a surjection from
  \(\mathbf{Fin}(n)\) for some \(n\).
  Also, we know that \(\mathbf{Fin}(n)\) is discrete (proven in our
  formalisation).
  Therefore, by lemma~\ref{discrete-surj}, \(A\) is discrete.
\end{proof}
\subsection{Manifest Bishop Finiteness}
\begin{definition}[Manifest Bishop Finiteness]
  \begin{equation} \label{bish-def}
    \mathcal{B}(A) \coloneqq \Sigma {(\mathit{xs} : \textbf{List}(A))} , \Pi {(x : A)} , x \inunique xs
  \end{equation}
\end{definition}
The only difference between manifest Bishop finiteness and split enumerability
is the membership term: here we require unique membership (\(\inunique\)),
rather than simple membership (\(\in\)).
\begin{definition}[Unique Membership] \label{uniq-memb-def}
  \begin{equation}
    x \inunique \mathit{xs} \coloneqq \text{isContr}(x \in \mathit{xs})
  \end{equation}
\end{definition}

We use the word ``manifest'' here to distinguish from another common
interpretation of Bishop finiteness, which we have called cardinal finiteness in
this paper.
The ``manifest'' refers to the fact that we have a concrete, non-truncated list
of the elements in the proof.

Where split enumerability was the enumeration form of a split surjection from
\(\mathbf{Fin}\), manifest Bishop finiteness is the enumeration form of an
\emph{equivalence} with \(\mathbf{Fin}\).
\begin{lemma} \label{bishop-equiv}
  \begin{equation}
    \mathcal{B}(A) \simeq \Sigma {(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \simeq A \right)
  \end{equation}
\end{lemma}
This proof is effectively the same as that of
lemma~\ref{split-enum-is-split-surj}.
\subsection{Cardinal Finiteness}
Each finiteness predicate so far has contained an \emph{ordering} of the
underlying type.
For our purposes, this is too much information: it means that when constructing
the ``category of finite sets'' later on, instead of each type having one
canonical representative, it will have \(n!\), where \(n\) is the cardinality of
the type\footnotemark.

\footnotetext{
  We actually do get a category (a groupoid, even) from manifest Bishop
  finiteness \cite{yorgeyCombinatorialSpeciesLabelled2014}: it's the groupoid of
  finite sets equipped with a linear order, whose morphisms are order-preserving
  bijections.
  We do not explore this particular construction in any detail.
}

To remedy the problem, we will use propositional truncation
(def.~\ref{prop-trunc}).
\begin{definition}[Cardinal Finiteness]
  \begin{equation}
    \mathcal{C}(A) \coloneqq \lVert \mathcal{B}(A) \rVert \simeq \lVert \Sigma(n : \mathbb{N}) , (\mathbf{Fin}(n) \simeq A) \rVert
  \end{equation}
\end{definition}
At first glance, it might seem that we lose any useful properties we could
derive from \(\mathcal{B}\).
Luckily, this is not the case: by eliminator \ref{elim-prop-coh} of
def.~\ref{prop-trunc}, we need only show that the output is uniquely determined.

The following two lemmas are proven in
\cite{yorgeyCombinatorialSpeciesLabelled2014} (Proposition 2.4.9 and 2.4.10,
respectively), in much the same way as we have done here.
Our contribution for this section is simply the formalisation.
\begin{lemma}
  Given a cardinally finite type, we can derive the type's cardinality, as well
  as a propositionally truncated proof of equivalence with \(\textbf{Fin}\)s of
  the same cardinality.
  \begin{equation}
    \mathcal{C}(A) \rightarrow \Sigma {(n : \mathbb{N})} , \lVert \textbf{Fin}(n) \simeq A \rVert
  \end{equation}
\end{lemma}
\begin{lemma} \label{cardinal-finite-discrete}
  Any cardinal-finite set has decidable equality.
\end{lemma}
\subsection{Manifest Enumerability}
\begin{definition}[Manifest Enumerability]
  \begin{equation}
    \mathcal{E}(A) \coloneqq \Sigma {(\mathit{xs} : \mathbf{List}(A))} , \Pi {(x : A)} , \lVert x \in \mathit{xs} \rVert
  \end{equation}
\end{definition}
As with manifest Bishop finiteness, the only difference with this type and split
enumerability is the membership proof: here we have propositionally truncated
it.
This has two effects.
First, it means that this proof represents a true surjection (rather than a
split surjection) from \(\mathbf{Fin}\).
\begin{lemma}
  \begin{equation}
    \mathcal{E}(A) \simeq \Sigma(n : \mathbb{N}) , (\mathbf{Fin}(n) \twoheadrightarrow A)
  \end{equation}
\end{lemma}
The proof for this lemma is similar in structure to
lemma~\ref{split-enum-is-split-surj} and lemma~\ref{bishop-equiv}.

Secondly, it means the predicate does not imply decidable equality.
More significantly, it allows the predicate to be defined over non-set types,
like the circle.
\begin{definition}[\(S^1\)] \label{circle-def}
  The circle, \(S^1\), can be represented in HoTT as a higher inductive type.
  \begin{equation}
    \begin{alignedat}{3}
      S^1 \coloneqq & \; \text{base} &&: S^1 ; \\
      | & \; \text{loop} &&: \text{base} \equiv \text{base} ; 
    \end{alignedat}
  \end{equation}
  We will use it here as an example of a non-set type, i.e. a type for which not
  all paths are equal.
  This also means that it does not have decidable equality.
\end{definition}
\begin{lemma}
  The circle \(S^1\) is manifestly enumerable.
\end{lemma}
\subsection{Kuratowski Finiteness}
The first thing we must define is a representation of subsets.
\begin{definition}[Kuratowski Finite Subset]
  \(\mathcal{K}(A)\) is the type of Kuratowski-finite subsets of \(A\).
  \begin{equation}
    \begin{alignedat}{3}
      \mathcal{K}(A) \coloneqq&
             \; []                &:& \; \mathcal{K}(A) ; \\
      \vert& \; \wc \dblcolon \wc &:& \; A \rightarrow \mathcal{K}(A) \rightarrow \mathcal{K}(A) ; \\
      \vert& \; \text{com}        &:& \; \Pi (x, y: A) , \Pi (\mathit{xs} : \mathcal{K}(A)) ,
                                  x \dblcolon y \dblcolon \mathit{xs} \equiv y \dblcolon x \dblcolon \mathit{xs} ; \\
      \vert& \; \text{dup}        &:& \; \Pi (x : A) , \Pi (\mathit{xs} : \mathcal{K}(A)) , x \dblcolon x \dblcolon \mathit{xs} \equiv x \dblcolon \mathit{xs} ; \\
      \vert& \; \text{trunc}      &:& \; \Pi (\mathit{xs}, \mathit{ys}: \mathcal{K}(A)) , \Pi (p, q : \mathit{xs} \equiv \mathit{ys}) , p \equiv q ;
    \end{alignedat}
  \end{equation}
  We define it as a HIT (definition~\ref{HITs}).
  The first two constructors are point constructors, giving ways to create
  values of type \(\mathcal{K}(A)\).
  They are also recognisable as the two constructors for finite lists, a type
  which represents the free monoid.

  The next two constructors add extra paths to the type: equations that usage of
  the type must obey.
  These extra paths turn the free monoid into the free \emph{commutative} (com)
  \emph{idempotent} (dup) monoid.

  The final constructor enforces that the type \(\mathcal{K}(A)\) must be a set.
\end{definition}
The Kuratowski finite subset is a free join semilattice (or, equivalently, a
free commutative idempotent monoid).
More prosaically, \(\mathcal{K}\) is the abstract data type for finite sets, as
defined in the Boom hierarchy \cite{boomFurtherThoughtsAbstracto1981,
  bunkenburgBoomHierarchy1994}.
However, rather than just being a specification, \(\mathcal{K}\) is fully usable
as a data type in its own right, thanks to HITs.

Other definitions of \(\mathcal{K}\) exist (such as the one in
\cite{fruminFiniteSetsHomotopy2018}) which make the fact that \(\mathcal{K}\) is
the free join semilattice more obvious.
We have included such a definition in our formalisation, and proven it
equivalent to the one above.

Next, we need a way to say that an entire type is Kuratowski finite.
For that, we will need to define membership of \(\mathcal{K}\).
\begin{definition}[Membership of \(\mathcal{K}\)]
  Membership is defined by pattern-matching on \(\mathcal{K}\).
  The two point constructors are handled like so:
  \begin{equation}
    \begin{alignedat}{2}
      x \in&& \; []                      &\coloneqq \bot ; \\
      x \in&& \; y \dblcolon \mathit{ys} &\coloneqq \lVert x \equiv y \uplus x \in \mathit{ys} \rVert ;
    \end{alignedat}
  \end{equation}
  The \(\text{com}\) and \(\text{dup}\) constructors are handled by proving that
  the truncated form of \(\uplus\) is itself commutative and idempotent.
  The type of propositions is itself a set, satisfying the \(\text{trunc}\)
  constructor.
\end{definition}
Finally, we have enough background to define Kuratowski finiteness.
\begin{definition}[Kuratowski Finiteness]
  \begin{equation}
    \mathcal{K}^{f}(A) = \Sigma {(\mathit{xs} : \mathcal{K}(A))} , \Pi (x : A) , x \in \mathit{xs}
  \end{equation}
\end{definition}

We also have the following two lemmas, proven in both
\cite{fruminFiniteSetsHomotopy2018} and our formalisation.
\begin{lemma}
  \(\mathcal{K}^f\) is a mere proposition.
\end{lemma}
\begin{lemma}
  This circle \(S^1\) is Kuratowski finite.
\end{lemma}
\section{Relations Between Each Finiteness Definition} \label{relations}
We will now look at the arrows in figure~\ref{finite-classification}.
\subsection{Split Enumerability and Manifest Bishop Finiteness}
While manifest Bishop finiteness might seem stronger than split enumerability,
it turns out this is not the case.
Both predicates imply the other.
\begin{lemma} \label{manifest-bishop-to-split-enum}
  Any manifest Bishop finite type is split enumerable.
\end{lemma}
\begin{proof}
  To construct a proof of split enumerability from one of manifest Bishop
  finiteness, it suffices to convert a proof of \(x \inunique \mathit{xs}\) to
  one of \(x \in \mathit{xs}\), for all \(x\) and \(\mathit{xs}\).
  Since \(\inunique\) is defined as a contraction of \(\in\), such a conversion
  is simply the \(\text{fst}\) function.
\end{proof}

\begin{lemma} \label{split-enum-to-manifest-bishop}
  Any split enumerable set is manifest Bishop finite.
\end{lemma}
This proof takes significantly more work.
The ``unique membership'' condition in \(\mathcal{B}\) means that we are not
permitted duplicates in the support list.
The first step in the proof, then, is to filter those duplicates out from the
support list of the \(\mathcal{E}!\) proof: we can do this using the decidable
equality provided by \(\mathcal{E}!\) (lemma~\ref{split-enum-discrete}).
From there, all we need to show is that the membership proof carries over
appropriately.
\subsection{Split Enumerability and Manifest Enumerability}
\begin{lemma} \label{split-enum-to-manifest-enum}
  Any split enumerable type is manifestly enumerable.
\end{lemma}
This lemma is proven by truncating the membership proof in split enumerability.

\begin{lemma} \label{manifest-enum-to-split-enum}
  A manifestly enumerable type with decidable equality is split enumerable.
\end{lemma}

\subsection{Manifest Bishop Finiteness and Cardinal Finiteness}
\begin{lemma} \label{manifest-bishop-to-cardinal}
  Any manifest Bishop finite type is cardinal finite.
\end{lemma}
\begin{theorem} \label{cardinal-to-manifest-bishop}
  Any cardinal finite type with a total order is Bishop finite.
\end{theorem}
The proof for this particular theorem is quite involved in the formalisation, so
we only give its sketch here.
First, note that we actually convert to manifest enumerability first: this can
be converted to split enumerability with decidable equality, which is provided
by cardinal finiteness.

Next, we define permutations.
\begin{definition}[List Permutations]
  Two lists are permutations of each other if their membership proofs are all
  equivalent\footnotemark \cite{danielssonBagEquivalenceProofRelevant2012}.
  \begin{equation}
    \mathit{xs} \leftrightsquigarrow \mathit{ys} = \Pi {(x : A)} , x \in \mathit{xs} \simeq x \in \mathit{ys}
  \end{equation}
\end{definition}

\footnotetext{
  The definition in \cite{danielssonBagEquivalenceProofRelevant2012} and our
  formalisation is slightly different: we say permutations are lists with
  \emph{isomorphic} membership proofs.
  The distinction, as it happens, does not affect our work here.
}

Next, we define a sort function which relies on the provided total order.
We further prove the following fact about this sort function:
\begin{equation}
  \Pi(\mathit{xs}, \mathit{ys} : \mathbf{List}(A)) , \mathit{xs} \leftrightsquigarrow \mathit{ys} \rightarrow \text{sort}(\mathit{xs}) \equiv \text{sort}(\mathit{ys})
\end{equation}

Next, notice that the support lists of any two proofs of manifest Bishop
finiteness must be permutations of each other.
This will allow us to sort the support list of a proof of cardinal finiteness in
a coherently constant (definition~\ref{prop-trunc},
eliminator~\ref{elim-prop-coh}) way, pulling the support list out from the
truncation.
The cover proof emerges naturally from the definition of the permutation.
\subsection{Cardinal Finiteness and Kuratowski Finiteness}
\begin{lemma} \label{cardinal-kuratowski}
  \begin{equation}
    \mathcal{C}(A) \simeq \mathcal{K}^f(A) \times \text{Discrete}(A)
  \end{equation}
\end{lemma}
This proof is constructed by providing a pair of functions: one from
\(\mathcal{C}(A)\) to \(\mathcal{K}^f(A) \times \text{Discrete}(A)\), and one the
other way.
This pair implies an equivalence, because both source and target are
propositions.
The actual functions themselves are proven in our formalisation, as well as in
\cite{fruminFiniteSetsHomotopy2018}.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../paper"
%%% End:

### 2020-05-29 sections/introduction.tex
split into sections

\section{Introduction}
\subsection{Foreword}
Proofs in constructive mathematics are often more substantial than their
classical counterparts.
``Substantial'' here doesn't mean the literal length of the proofs (although the
double meaning is not lost on anyone who has had to struggle with a modern
constructive proof assistant), but rather refers to their \emph{content}: a
constructive proof contains more \emph{stuff} than a classical one.

If we were to prove that a set was finite in classical mathematics, we might
show that it is in bijection with another finite set, say a finite prefix of the
natural numbers.
In semi-formal notation, to say that a set \(A\) is finite we would show the
following:
\begin{equation} \label{example-finite}
  \mathbf{Finite}(A) \coloneqq \exists (n : \mathbb{N}) . \; A \Leftrightarrow \left\{ m \mid m < n\right\}
\end{equation}
Where \(A~\Leftrightarrow~B\) means ``a bijection between \(A\) and \(B\)''.

Classically, a proof of equation~\ref{example-finite} is nothing more than
\emph{evidence} for the proposition of equation~\ref{example-finite}.
A constructive proof of \ref{example-finite}, however, is a mathematical object
in its own right, containing:
\begin{enumerate}
  \item A number \(n\) representing the cardinality of \(A\).
  \item A function from \(A\) to \(\left\{ m \mid m < n \right\}\).
  \item A function from \(\left\{ m \mid m < n \right\}\) to \(A\).
\end{enumerate}
In other words, it makes sense (in constructive mathematics) to talk about a
function of type \(\mathbf{Finite}(A)~\rightarrow~\mathbb{N}\), which takes a
\emph{proof} of the finiteness of \(A\) and returns \(A\)'s cardinality.

Such functions are perhaps even more concrete than this description suggests:
many constructive proofs are in fact computer programs in disguise (and
vice-versa \cite{wadlerPropositionsTypes2015}).
All ``proofs'' in this paper correspond to honest-to-goodness proofs in a
programming language (Cubical Agda \cite{vezzosiCubicalAgdaDependently2019} in
our case) which can be run on a computer to produce real output.

We have now hopefully established, albeit informally, what it means for
constructive proofs to be ``substantial''.
Clearly this extra information carried by proofs can add richness to their
study: we will see, for instance, no fewer than five finiteness predicates in
this paper which would collapse to one in a classical setting.
The extra information also presents a challenge, however: occasionally we will
see that we have revealed \emph{too much} information while proving some
property.

Consider the category of finite sets.
We will explore this category in much more detail later, but for now we just
need to know that the objects of this category are---shockingly---finite sets.
Constructively, the objects of this category can be thought of as \emph{pairs},
where the first element is a set (like \(\left\{ \text{True}, \text{False}
\right\}\), the two-element set of the booleans) and the second is a proof that
that set is finite.
If we take those proofs to be the kind given in equation~\ref{example-finite},
we should see that they contain a bijection from the booleans to the set
\(\left\{ 0, 1 \right\}\): however \emph{two} such bijections exist!
\begin{figure}[h]
  \centering
  \begin{subfigure}[b]{.45\textwidth}
    \centering
    \begin{tikzcd}
      \text{True}  \ar[rr, leftrightarrow] & & 0 \\
      \text{False} \ar[rr, leftrightarrow] & & 1
    \end{tikzcd}
  \end{subfigure}
  \begin{subfigure}[b]{.45\textwidth}
    \centering
    \begin{tikzcd}
      \text{True}  \ar[rrd, leftrightarrow, start anchor=south east, end anchor=north west] & & 0 \\
      \text{False} \ar[rru, leftrightarrow, start anchor=north east, end anchor=south west, crossing over] & & 1
    \end{tikzcd}
  \end{subfigure}
  \caption{The Two Bijections Between the Booleans and \(\left\{ 0, 1 \right\}\)}
\end{figure}
This means that where we wanted one object to correspond to the finite set
\(\left\{ \text{True}, \text{False} \right\}\), we instead have two.
In fact, for any finite set of cardinality \(n\) we will have \(n!\) finiteness
proofs, and therefore \(n!\) objects.
We have constructed, in other words, entirely the wrong category.

Being able to hide the extra information in a constructive proof when we need to
is clearly important.
HoTT (and CuTT by extension) facilitate this hiding: how exactly it is
accomplished will be explored in the subsequent sections.

Finally, for all of its benefits, constructive mathematics (especially
dependently-typed constructive mathematics on a computer) still suffers from
many painful drawbacks.
As already hinted at, proofs in a constructive setting can tend to be verbose,
especially when the arbiter of correctness is as strict as a computer, which
neither understands nor trusts the time-honoured technique of ``the other cases
are obviously true''.
While the goal of using a computer to automate the boring parts of a proof still
seems attainable, too often getting a computer involved in a proof means adding
thousands of lines of tedium to satisfy its lack of intuition.

Simple proof search over a finite domain is a simple, if effective, way to
relieve some of the tedium.
Many libraries in Agda, Coq, and other languages make use of this fact.
While the idea for finite proof search is simple, we found that implementing one
revealed a number of interesting theoretical considerations:
\begin{description}
  \item[How should we construct search spaces?]
    We found that a \(Pi\)-pretopos provided the appropriate ``generator
    language'' for constructing search spaces for finite types.
  \item[How do we deal with functions?]
    We of course would like to be able to search over functions: \todo{?}
  \item[How do we talk about ``searchability''?]
    \todo{Omniscience}
\end{description}
Finiteness is a prime example of a phenomenon which happens quite frequently in
dependently-typed constructive mathematics: through building a practical program
we unearth theoretically interesting results.
\todo{rephrase}
\subsection{Contributions}
We are interested in constructive notions of finiteness.
In contrast to classical finiteness, in a constructive setting there is a wide
variety of predicates which all have some claim to being the formal
interpretation of ``finiteness''.
Broadly, they vary along two axes: informativeness (i.e. how many properties
of the underlying type we can derive from a proof that it is finite), and
restrictiveness (i.e. how many types conform to its particular variety of
finiteness).
We will explore five of these predicates, proving formally their relations and
implications.

We will then focus on one of the predicates: decidable Kuratowski finiteness.
We will construct a category of decidable Kuratowski-finite types, and show that
it forms a \(\Pi\)-pretopos.

We will then take a brief detour into predicates for countably infinite types,
and compare its landscape to that of the finite predicates.

Finally, we will present a case-study in a practical application of the
predicates for finiteness we have constructed: a library for total proof search.
We will show how our earlier work gives a good theoretical foundation for this
library, and allows for a simple and principled API, and we will demonstrate how
some of the more complex theoretical results allow for a more powerful library
with capabilities not seen in comparable proof search libraries.

\input{figures/finite-classification}

\subsubsection{Finiteness Predicates}
Our first contribution is a cataloguing and characterisation of five finiteness
predicates, shown in figure~\ref{finite-classification}.

While our focus is decidable Kuratowski finiteness, in order to prove the
closure properties that we are interested in we will need to work with the
intermediate predicates as well (though they are also interesting in their own
right).

Since we have a proof-search library in mind, we will have to consider how to
present a good API for constructing proofs of finiteness.
In this vein we will show that some simple base types are finite
(\(\mathbf{Bool}\), \(\top\), \(\bot\), etc), and we will provide functions to
combine finiteness proofs together into more complex types (i.e. \(\times\),
\(\uplus\), \(\Sigma\), \(\Pi\), etc).
As we will see this choice of toolbox for building finiteness proofs actually
has a deep categorical justification, but at this point it makes sense to look
at it as just a suitably general and powerful API.

Our focus in this work is decidable Kuratowski finiteness.
We will characterise it, its strength, and its uses, and in particular we will
show that it forms a \(\Pi\)-pretopos.
Our proofs follow the structure of \cite[Chapters 9, 10]{hottbook} and
\cite{rijkeSetsHomotopyType2015}.
We will explore this concept in more detail later, but crucially we will need to
provide proofs of closure under \(\times\), \(\rightarrow\), etc.

In order to prove these closure properties, we will define and use some
intermediate finiteness predicates, classified in
figure~\ref{finite-classification}.
These predicates are also interesting in their own right, though!

The arrows in figure~\ref{finite-classification} are implication functions
between each predicate.
In section~\ref{relations} we will provide a function which inhabits each arrow.
Each unlabelled arrow is an unconditional implication: every manifest Bishop
finite set is cardinal finite (lemma~\ref{manifest-bishop-to-cardinal}), for
instance.
The labelled arrows are strengthening proofs: every manifest enumerable set
\emph{with decidable equality} is split enumerable
(lemma~\ref{manifest-enum-to-split-enum}).
Our most significant result here is the proof that a cardinal finite set with a
decidable total order is manifestly Bishop finite.

\subsubsection{Countability Predicates}
After the finite predicates, we will briefly look at the infinite countable
types, and classify them in a parallel way to the finite predicates
(section~\ref{infinite}).
\subsubsection{Search}
All of our work is formalised in Cubical Agda
\cite{vezzosiCubicalAgdaDependently2019}.
We will make mention of the few occasions where the formalisation of some proof
is of interest, but the main place where we will discuss the code is in
section~\ref{search}, where we implement a library for proof search, based on
omniscience and exhaustibility.
This library relies directly on the computational content of univalence, which
allows us to (for instance) have \(\Pi\) types in the domain of the search
space.
The language of a \(\Pi\)-pretopos turns out to be exactly the right language
for constructing ``generator''-like expressions.

The full code of the formalisation is available at
\url{https://github.com/oisdk/finiteness-in-cubical-type-theory}.
\subsection{Notation and Background}
\subsubsection{Notation}
We work in Cubical Type Theory \cite{cohenCubicalTypeTheory2016}.
For the various type formers we use the following notation:
\begin{description}
  \item[Type] We use \(\mathbf{Type}\) to denote the universe of (small) types.
    ``Type families'' are functions into \(\mathbf{Type}\).
  \item[0 , 1 , 2] We call the \(\mathbf{0}\), \(\mathbf{1}\), and
    \(\mathbf{2}\) types \(\bot\), \(\top\), and \(\mathbf{Bool}\) respectively.
    The single inhabitant of \(\top\) is tt, and the two inhabitants of
    \(\mathbf{Bool}\) are false and true.
    The ``negation'' of a type, written \(\neg A\), means \(A \rightarrow
    \bot\).
  \item[Dependent Sum and Product] We use \(\Sigma\) and \(\Pi\) for the
    dependent sum and product, respectively.
    The two projections from \(\Sigma\) are called fst and snd.
    In the non-dependent case, \(\Sigma\) can be written as \(\times\), and
    \(\Pi\) as \(\rightarrow\).
  \item[Disjoint Union] We define disjoint union as an inductive type.
    \begin{equation}
      \begin{alignedat}{3}
        A \uplus B \coloneqq & \;
        \text{inl} &: A \rightarrow A \uplus B ; \\
        | & \;  \text{inr} &: B \rightarrow A \uplus B ;
      \end{alignedat}
    \end{equation}
    It is also expressible with only \(\Sigma\):
    \(A \uplus B \simeq \Sigma(x : \mathbf{Bool}) , \text{if } x \text{ then
    } A \text{ else } B \).
  \item[Equalities, equivalences, and paths] We use the symbol \(\coloneqq\)
    for definitions.
    \(\simeq\) will be used for equivalences, and \(\equiv\) for equalities.
    Of course, we know that \((A \simeq B) \simeq (A \equiv B)\) by univalence,
    so the distinction isn't terribly important in usage: we will only use one
    or the other as a suggestion of how we constructed it or how it is to be
    used.
  \item[Lambdas] \todo{Figure out how to describe lambdas}
\end{description}
\subsubsection{Cubical Type Theory}
Cubical Type Theory \cite{cohenCubicalTypeTheory2016} is a constructive type
theory with an implementation in Cubical Agda
\cite{vezzosiCubicalAgdaDependently2019}.
It allows us to do much of the same theory as in HoTT, but crucially the
univalence ``axiom'' is a \emph{theorem}, giving it computational content.
\begin{definition}[Path Types] \label{path-types}
  The equality type (which we denote with \(\equiv\)) in CuTT is the type of
  Paths\footnotemark.
  The internal structure of paths is largely irrelevant to us here, as we will
  generally treat \(\equiv\) as a black-box equivalence relation with
  substitution and congruence.
\end{definition}

\footnotetext{
  Actually, CuTT does have an identity type with similar semantics to the
  identity type in MLTT.
  We do not use this type anywhere in our work, however, so we will not consider
  it here.
}
\begin{definition}[Homotopy Levels] \label{homotopy-types}
  Types in HoTT and CuTT are not necessarily sets, as they are in MLTT.
  Some have higher homotopies (paths which aren't unique).
  We actually have a hierarchy of complexity of structure of path spaces in
  types, starting with the contractions \cite[definition 3.11.1]{hottbook}, then
  the mere propositions \cite[definition 3.3.1]{hottbook}, and the sets
  \cite[definition 3.1.1]{hottbook}.
  \begin{alignat}{2}
    &\text{isContr}(A)    &&\coloneqq \Sigma(x : A) , \Pi(y : A) , (x \equiv y) \\
    &\text{isProp}(A)     &&\coloneqq \Pi(x, y : A) , (x \equiv y) \\
    &\text{isSet}(A)      &&\coloneqq \Pi(x, y : A) , \text{isProp}(x \equiv y)
  \end{alignat}
\end{definition}
\begin{definition}[Fibres] \label{fibres}
  A fibre \cite[definition 4.2.4]{hottbook} is defined over some function \(f :
  A \rightarrow B\).
  \begin{equation}
    \text{fib}_f(y) \coloneqq \Sigma(x : A) , (f (x) \equiv y)
  \end{equation}
\end{definition}
\begin{definition}[Equivalences] \label{equivalences}
  We will take contractible maps \cite[definition 4.4.1]{hottbook} as our
  ``default'' definition of equivalences.
  \begin{alignat}{2}
    &\text{isEquiv}(f) &&\coloneqq \Pi(y : B) , \text{isContr}(\text{fib}_f(y)) \label{is-equiv-def} \\
    &A \simeq B        &&\coloneqq \Sigma(f : A \rightarrow B) , \text{isEquiv}(f)
  \end{alignat}
\end{definition}
\begin{definition}[Decidable Types]
  \begin{equation}
    \mathbf{Dec}(A) \coloneqq A \uplus \neg A
  \end{equation}
\end{definition}
\begin{definition}[Discrete Types]
  A discrete type is one with decidable equality.
  \begin{equation}
    \text{Discrete}(A) \coloneqq \Pi(x, y : A) , \mathbf{Dec}(x \equiv y)
  \end{equation}
  By Hedberg's theorem \cite{hedbergCoherenceTheoremMartinLof1998} any discrete
  type is a set.
\end{definition}
\begin{definition}[Higher Inductive Types] \label{HITs}
  Normal inductive types have \emph{point} constructors: constructors which
  construct values of the type.
  Higher Inductive Types (HITs) also have \emph{path} constructors: ways to
  construct paths in the type.
\end{definition}
\begin{definition}[Propositional Truncation] \label{prop-trunc}
  The type \(\lVert A \rVert\) on some type \(A\) is a propositionally truncated
  proof of \(A\) \cite[3.7]{hottbook}.
  In other words, it is a proof that some \(A\) exists, but it does not tell you
  \emph{which} \(A\).

  It is defined as a Higher Inductive Type:
  \begin{equation} {
    \begin{alignedat}{3}
      \lVert A \rVert \coloneqq & \; \lvert \wc \rvert &:& \; A \rightarrow \lVert A \rVert ; \\
                              | & \; \text{squash}     &:& \; \Pi {(x, y : \lVert A \rVert)} , x \equiv y  ; 
    \end{alignedat} }
  \end{equation}
  We will use two eliminators from \(\lVert A \rVert\) in this paper.
  \begin{enumerate}
  \item \label{elim-prop-prop} For any function \(A \rightarrow B\), where
    \(\text{isProp}(B)\), we have a function \(\lVert A \rVert \rightarrow B\).
  \item \label{elim-prop-coh} We can eliminate from \(\lVert A \rVert\) with a
    function \(f : A \rightarrow B\) iff \(f\) ``doesn't care'' about the
    choice of \(A\):
    \[\Pi {(x , y : A)} , f(x) \equiv f(y) \]
    Formally speaking, \(f\) needs to be ``coherently constant''
    \cite{krausGeneralUniversalProperty2015}, and \(B\) needs to be an
    \(n\)-type for some finite \(n\).
  \end{enumerate}
\end{definition}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../paper"
%%% End:

### 2020-05-29 sections/search.tex
split into sections

\section{Search} \label{search}
\subsection{Omniscience}
\begin{definition}[Limited Principle of Omniscience]
  For any type \(A\) and predicate \(P\) on \(A\), the limited principle of
  omniscience \cite{myhillErrettBishopFoundations1972} is as follows:
  \begin{equation}
    \left( \Pi {(x : A)} , \mathbf{Dec}(P(x)) \right) \rightarrow \mathbf{Dec} \left( \Sigma {(x : A)} , P(x) \right)
  \end{equation}
  In other words, for any decidable predicate the existential quantification of
  that predicate is also decidable.
\end{definition}
The limited principle of omniscience is non-constructive, but individual types
can themselves satisfy omniscience.
In particular, cardinal finite types are omniscient.

There is also a universal form of omniscience, which we call exhaustibility.
\begin{definition}[Exhaustibility]
  We say a type \(A\) is exhaustible if, for any decidable predicate \(P\) on
  \(A\), the universal quantification of the predicate is decidable.
  \begin{equation}
    \left( \Pi {(x : A)} , \mathbf{Dec}(P(x)) \right) \rightarrow \mathbf{Dec} \left( \Pi {(x : A)} , P(x) \right)
  \end{equation}
\end{definition}

All of the finiteness predicates we have seen imply exhaustibility.
Omniscience is stronger than exhaustibility, as we can derive the latter from
the former.
All of the ordered finiteness predicates imply omniscience.
For the unordered finiteness definitions, we have omniscience for prop-valued
predicates.
\subsection{Automating Proofs}
One use for above constructions is the automation of certain proofs.
In \cite{firsovDependentlyTypedProgramming2015}, which uses a similar approach
to ours, the \(\AgdaDatatype{Pauli}\) group is used as an example.
\ExecuteMetaData[agda/Data/Pauli.tex]{def}
As \(\AgdaDatatype{Pauli}\) has 4 constructors, \(n\)-ary functions on
\(\AgdaDatatype{Pauli}\) may require up to \(4^n\) cases, making even simple
proofs prohibitively verbose.

The alternative is to derive the things we need from omniscience, itself derived
from a finiteness predicate.
For proof search, the procedure is a well-known one in Agda
\cite{devrieseBrightSideType2011}: we ask for the result of a decision procedure
as an \emph{instance argument}, which will demand computation during
typechecking.
Our addition to this technique is a way to handle multiple arguments based on
fully level-polymorphic dependent currying and uncurrying, building on 
\cite{allaisGenericLevelPolymorphic2019}.
\ExecuteMetaData[agda/Data/Pauli.tex]{assoc-prf}

Finally, we can derive decidable equality on functions over finite types.
We can also use functions in our proof search.
Here, for instance, is an automated procedure which finds the
\(\AgdaFunction{not}\) function on \(\mathbf{Bool}\), given a specification.
\ExecuteMetaData[agda/Data/Pauli.tex]{not-spec}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../paper"
%%% End:

### 2020-05-29 sections/topos.tex
split into sections

\section{Topos} \label{topos}
In this section we will prove that decidable Kuratowski finite types form a
\(\Pi\)-pretopos.
Along the way we will provide closure proofs for a number of the other
finiteness predicates.
As we saw in theorem~\ref{cardinal-kuratowski}, decidable Kuratowski finite
types are equivalent to cardinal finite types, so we will work with the latter
from now on.
Our first task is to show that cardinal finite types are closed under several
operations.
\subsection{Closure}
For the first three closure proofs, we only consider split enumerability:
as it is the strongest of the finiteness predicates, we can derive the other
closure proofs from it.

\begin{lemma}
  \(\bot\), \(\top\), and \(\mathbf{Bool}\) are split enumerable.
\end{lemma}
\begin{lemma} \label{split-enum-sigma}
  Split enumerability is closed under \(\Sigma\).
  \begin{equation}
    \frac{
      \mathcal{E}!(A) \; \; \; \Pi(x : A) , \mathcal{E}!(U(x))
    }{
      \mathcal{E}!(\Sigma(x : A) , U(x))
    }
  \end{equation}
\end{lemma}
From this we can derive split enumerability of non-dependent sums and products,
as both are definable in terms of \(\Sigma\).

\begin{lemma}
  Split enumerability is closed under dependent functions.
  (\(\Pi\)-types).
  \begin{equation}
    \frac{
      \mathcal{E}!(A) \; \; \; \Pi {(x : A)} , \mathcal{E}!\left( U(x) \right)
    }{
      \mathcal{E}!\left(\Pi {(x : A)} , U(x)\right)
    }
  \end{equation}
\end{lemma}
\begin{proof}
  Let \(A\) be a split enumerable type, and \(U\) be a type family from \(A\),
  which is split enumerable over all points of \(A\).

  As \(A\) is split enumerable, we know that it is also manifestly Bishop finite
  (lemma~\ref{split-enum-to-manifest-bishop}), and consequently we know \(A
  \simeq \mathbf{Fin}(n)\), for some \(n\) (lemma~\ref{bishop-equiv}).
  We can therefore replace all occurrences of \(A\) with \(\mathbf{Fin}(n)\),
  changing our goal to:
  \begin{equation}
    \frac{
      \mathcal{E}!(\mathbf{Fin}(n)) \; \; \; \Pi (x : \mathbf{Fin}(n)) , \mathcal{E}!\left( U(x) \right)
    }{
      \mathcal{E}!\left(\Pi (x : \mathbf{Fin}(n)) , U(x)\right)
    }
  \end{equation}

  We then define the type of \(n\)-tuples over some type family \(T :
  \mathbf{Fin}(n) \rightarrow \mathbf{Type}\).
  \begin{equation}
    \begin{alignedat}{3}
      & \mathbf{Tuple}(0, T)   &&\coloneqq \top \\
      & \mathbf{Tuple}(n+1, T) &&\coloneqq T(0) \times \mathbf{Tuple}(n, T \circ \text{suc})
    \end{alignedat}
  \end{equation}
  We can show that this type is equivalent to functions (proven in our formalisation):
  \begin{equation}
    \Pi(x : \mathbf{Fin}(n)) , U(x) \simeq \mathbf{Tuple}(n, U)
  \end{equation}
  And therefore we can simplify again our goal to the following:
  \begin{equation}
    \frac{
      \mathcal{E}!(\mathbf{Fin}(n)) \; \; \; \Pi (x : \mathbf{Fin}(n)) , \mathcal{E}!\left( U(x) \right)
    }{
      \mathcal{E}!\left(\mathbf{Tuple}(n, U)\right)
    }
  \end{equation}

  We can prove this goal by showing that \(\mathbf{Tuple}(n, U)\) is split
  enumerable: it is made up of finitely many products of points of \(U\), which
  are themselves split enumerable, and \(\top\), which is also split enumerable.
  Lemma~\ref{split-enum-sigma} shows us that the product of finitely many split
  enumerable types is itself split enumerable, proving our goal.
\end{proof}

Lifting the above closure proofs to work on proofs of \(\mathcal{C}\) is not
necessarily straightforward.
\(\lVert \wc \rVert\) forms a monad, giving us access to a powerful syntax for
combining proofs.
From this, we can derive closure under \(\bot\), \(\top\), \(\mathbf{Bool}\),
\(\times\), \(\uplus\), and \(\rightarrow\).
The dependent case is not so straightforward.
We need the following lemma, a version of the axiom of choice on finite sets,
to prove closure under \(\Sigma\) and \(\Pi\).
\begin{lemma}
  \begin{equation}
    \mathcal{C}(A) \rightarrow (\Pi(x : A) , \lVert U(x) \rVert) \rightarrow \lVert \Pi(x : A) , U(x) \rVert
  \end{equation}
\end{lemma}
\subsection{The Category of Finite Sets}
HoTT and CuTT seem to be especially suitable settings for formalisations of
category theory.
The univalence axiom in particular allows us to treat categorical isomorphisms
as equalities, saving us from the dreaded ``setoid hell''.

We follow \cite[chapter 9]{hottbook} in its treatment of
categories in HoTT, and in its proof that sets do indeed form a category.
We will first briefly go through the construction of the category
\(\mathit{Set}\), as it differs slightly from the usual method in type theory.

First, the type of objects and arrows:
\begin{alignat}{3}
  &\text{Obj}_\mathit{Set}      &&\coloneqq \Sigma(x : \mathbf{Type}) , \text{isSet}(x) \\
  &\text{Hom}_\mathit{Set}(x , y) &&\coloneqq  \text{fst}(x) \rightarrow \text{fst}(y)
\end{alignat}
As the type of objects makes clear, we have already departed slightly from the
simpler \(\text{Obj}_\mathit{Set} \coloneqq \mathbf{Type}\) way of doing things:
of course we have to, as HoTT allows non-set types.
Furthermore, after proving the usual associativity and identity laws for
composition (which are definitionally true in this case), we must further show
\(\text{isSet}(\text{Hom}_\mathit{Set}(x,y))\); even then we only have a
precategory.

To show that \(\mathit{Set}\) is a category, we must show that categorical
isomorphisms are equivalent to equivalences.
In a sense, we must give a univalence rule for the category we are working in.

We have provided formal proofs that \(\mathit{Set}\) does indeed form a
category, and the following:
\begin{theorem}[The Category of Finite Sets]
  Finite sets form a category in HoTT when defined like so:
  \begin{equation}
    \begin{alignedat}{3}
      &\text{Obj}_\mathit{FinSet}      &&\coloneqq \Sigma(x : \mathbf{Type}) , \mathcal{C}(x) \\
      &\text{Hom}_\mathit{FinSet}(x , y) &&\coloneqq  \text{fst}(x) \rightarrow \text{fst}(y)
    \end{alignedat}
  \end{equation}
\end{theorem}
\subsection{The \(\Pi\)-pretopos of Finite Sets}
For this proof, we follow again the proof that \(\mathit{Set}\) forms a \(\Pi
W\)-pretopos from \cite[chapter 10]{hottbook} and
\cite{rijkeSetsHomotopyType2015}.
The difference here is that clearly we do not have access to \(W\)-types, as
they would permit infinitary structures.

We first must show that \(\mathit{Set}\) has an initial object and finite,
disjoint sums, which are stable under pullback.
We also must show that \(\mathit{Set}\) is a regular category with effective
quotients.
We now have a pretopos: the presence of \(\Pi\) types make it a
\(\Pi\)-pretopos.

We have proven the above statements for both \(\mathit{Set}\) and
\(\mathit{FinSet}\).
As far as we know, this is the first formalisation of either.
\begin{theorem} \label{finite-topos}
  The category of finite sets, \(\mathit{FinSet}\), forms a \(\Pi\)-pretopos.
\end{theorem}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../paper"
%%% End:

### 2020-05-30 sections/finiteness-predicates.tex
more rearranging

@@
 in Figure~\ref{finite-classification}.
-The reason we explore predicates other than our focus (decidable Kuratowski
-finiteness) is that we can often prove things like closure much more readily on
-the simpler predicates.
-The relations (which we will prove in the next section) then allow us to
-transfer those proofs.
+As we will see, each of these predicates has subtle differences from the others:
+we will outline how some predicates are too informative, and how others aren't
+powerful enough, for our needs, before settling on decidable Kuratowski
+finiteness as our focus.
+
+As we make our way through each predicate, we will be interested in two aspects:
+how can we build proofs of this predicate (i.e. is the product of two finite
+types finite?) and what do we \emph{get} once we do (i.e. does this predicate
+tell us the number of elements in the finite set?).
 \subsection{Split Enumerability}
+We will start with a simple notion of finiteness, called \emph{split}
+enumerability.
+
 \begin{definition}[Split Enumerable Set] \label{split-enum-def}
+  To say that some type \(A\) is split enumerable is to say that there is a list
+  \(\mathit{support} : \mathbf{List}(A)\) such that any value \(x : A\) is in
+  \(\mathit{support}\).
   \begin{equation} \label{split-enum-def-eqn}
-    \mathcal{E}!(A) \coloneqq \Sigma {(\mathit{xs} : \textbf{List}(A))} , \Pi {(x : A)} , x \in \mathit{xs}
+    \mathcal{E}!(A) \coloneqq \Sigma {(\mathit{support} : \textbf{List}(A))} , \Pi {(x : A)} , x \in \mathit{support}
   \end{equation}
@@
 \end{definition}
+We used some extra types in the above definition, which we will define here:
 
-We used some extra types in the above definition, which we will define here:
+\begin{definition}[\(\mathbf{List}\)] \label{List}
+  In this paper we define lists as \emph{containers}:
+  \begin{equation}
+    \mathbf{List} \coloneqq \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket
+  \end{equation}
+  This may seem quite foreign or complex in comparison to the usual definition
+  of lists as an inductive data type:
+  \begin{inductivetype}{\mathbf{List}(A)}
+    \inductivetypeclause{\left[ \right]}{\mathbf{List}(A)}
+    \inductivetypeclause{\wc \dblcolon \wc}{A \rightarrow \mathbf{List}(A) \rightarrow \mathbf{List}(A)}
+  \end{inductivetype}
+  But it turns out to be quite useful for later proofs.
+
+  Moreover, the inductive type and the container-based type are equivalent, as
+  is proven in our formalisation.
+\end{definition}
 \begin{definition}[Containers] \label{container-def}
@@
 \end{definition}
-\begin{definition}[\(\mathbf{List}\)] \label{List}
-  \begin{equation}
-    \mathbf{List} \coloneqq \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket
-  \end{equation}
-\end{definition}
 \begin{definition}[\(\mathbf{Fin}\)] \label{Fin}
@@
 \end{definition}
+\subsubsection{Instances}
+Now that we have a suitable definition of finiteness, we will next prove that
+some things are finite.
+\begin{lemma}
+  \(\bot\), \(\top\), and \(\mathbf{Bool}\) are split enumerable.
+\end{lemma}
+\begin{proof}
+  These three types are represented by fairly simple finite sets:
+  \begin{align}
+    \bot &= \left\{  \right\} \\
+    \top &= \left\{ \text{tt}  \right\} \\
+    \mathbf{Bool} &= \left\{  \text{false}, \text{true} \right\}
+  \end{align}
+  Which leads to quite simple proofs of finiteness.
+
+  The support list for is \(\left[  \right]\), \(\left[ \text{tt} \right]\), and
+  \(\left[ \text{false}, \text{true} \right]\), respectively, and the cover
+  proof is a function that returns the index of the supplied element.
+\end{proof}
+\begin{agdalisting}
+  The code in our formalisation of these three proofs is quite simple.
+  Just as an example, the proof that \(\mathbf{Bool}\) is split enumerable is as
+  follows:
+  \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{bool-inst}
+\end{agdalisting}
+
+\begin{lemma} \label{split-enum-sigma}
+  Split enumerability is closed under \(\Sigma\).
+  \begin{equation}
+    \frac{
+      \mathcal{E}!(A) \; \; \; \Pi(x : A) , \mathcal{E}!(U(x))
+    }{
+      \mathcal{E}!(\Sigma(x : A) , U(x))
+    }
+  \end{equation}
+\end{lemma}
 
@@
   \(\mathcal{K}(A)\) is the type of Kuratowski-finite subsets of \(A\).
-  \begin{equation}
-    \begin{alignedat}{3}
-      \mathcal{K}(A) \coloneqq&
-             \; []                &:& \; \mathcal{K}(A) ; \\
-      \vert& \; \wc \dblcolon \wc &:& \; A \rightarrow \mathcal{K}(A) \rightarrow \mathcal{K}(A) ; \\
-      \vert& \; \text{com}        &:& \; \Pi (x, y: A) , \Pi (\mathit{xs} : \mathcal{K}(A)) ,
-                                  x \dblcolon y \dblcolon \mathit{xs} \equiv y \dblcolon x \dblcolon \mathit{xs} ; \\
-      \vert& \; \text{dup}        &:& \; \Pi (x : A) , \Pi (\mathit{xs} : \mathcal{K}(A)) , x \dblcolon x \dblcolon \mathit{xs} \equiv x \dblcolon \mathit{xs} ; \\
-      \vert& \; \text{trunc}      &:& \; \Pi (\mathit{xs}, \mathit{ys}: \mathcal{K}(A)) , \Pi (p, q : \mathit{xs} \equiv \mathit{ys}) , p \equiv q ;
-    \end{alignedat}
-  \end{equation}
+  \begin{inductivetype}{\mathcal{K}(A)}
+    \inductivetypeclause{\left[ \right]}{\mathcal{K}(A)}
+    \inductivetypeclause{\wc \dblcolon \wc}{A \rightarrow \mathcal{K}(A) \rightarrow \mathcal{K}(A)}
+    \inductivetypeclause{\text{com}}{\Pi (x, y: A) , \Pi (\mathit{xs} : \mathcal{K}(A)) , x \dblcolon y \dblcolon \mathit{xs} \equiv y \dblcolon x \dblcolon \mathit{xs}}
+    \inductivetypeclause{\text{dup}}{\Pi (x : A) , \Pi (\mathit{xs} : \mathcal{K}(A)) , x \dblcolon x \dblcolon \mathit{xs} \equiv x \dblcolon \mathit{xs}}
+    \inductivetypeclause{\text{trunc}}{\Pi (\mathit{xs}, \mathit{ys}: \mathcal{K}(A)) , \Pi (p, q : \mathit{xs} \equiv \mathit{ys}) , p \equiv q}
+  \end{inductivetype}
   We define it as a HIT (definition~\ref{HITs}).

### 2020-05-30 sections/finiteness-predicates.tex
added prose proof of sigma closure

@@
 
+With the most basic simple types out of the way, the obvious next choice is the
+(non-dependent) sums and products: \(\uplus\) and \(\times\).
+Both of these types can be constructed from the \emph{dependent} sum, however,
+so that is the type we will prove finite.
+From that we can derive a much wider array of finiteness proofs.
 \begin{lemma} \label{split-enum-sigma}
@@
 \end{lemma}
+\begin{proof}
+  Let \(A\) be a type which is split enumerable, and \(U\) be a type family over
+  \(A\) which is split enumerable at every point.
+  Formally, we have the following proofs:
+  \begin{align}
+    \mathcal{E}!_A &: \mathcal{E}!(A) \\
+    \mathcal{E}!_U &: \Pi(x : A) , \mathcal{E}!(U(x))
+  \end{align}
+
+  Our task is to construct a proof of type:
+  \begin{equation}
+    \mathcal{E}!(\Sigma(x : A) , U(x))
+  \end{equation}
+  This proof itself is composed of two components:
+  \begin{align}
+    \mathit{support} &: \mathbf{List}(\Sigma(x : A) , U(x)) \\
+    \mathit{cover} &: \Pi(x : \Sigma(y : A) , U(y)) , x \in \mathit{support}
+  \end{align}
+
+  To construct the support list, we apply the function \(\mathcal{E}!_U\) to
+  every element in the support list of \(\mathcal{E}!_A\), extract the support
+  lists from the resulting finiteness proofs, and concatenate them.
+
+  \begin{agdalisting}
+    This pattern of applying a function to each element in a list and
+    concatenating the result is of course well-known in functional programming,
+    and is in fact the pattern that makes lists a monad.
+    While this insight isn't strictly relevant to our work here, it does mean
+    the implementation of this function can use Agda's do notation, resulting
+    in the following extremely clean implementation:
+    \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{sup-sigma}
+  \end{agdalisting}
+
+  To prove that this support list does in fact cover the entirety of the type
+  \(\Sigma \; A \; U\), we note that any element of type \(\Sigma \; A \; U\)
+  must have a first component in the support list of \(\mathcal{E}!_A\), and its
+  second component must be in the result of applying \(\mathcal{E}!_U\) to that
+  first element (since that support list contains every element of type
+  \(U(x)\)).
+  Therefore, the pair itself must be in our constructed support list.
+\end{proof}
 

### 2020-05-30 sections/finiteness-predicates.tex
Update finiteness-predicates.tex

@@
 \end{proof}
+\subsubsection{Derivations}
+We have a way to construct finiteness proofs, and a semiring-like toolbox to
+combine them.
+What we're now interested in is what we can \emph{derive} from them.
 
-We tend to prefer list-based definitions of finiteness, rather than ones based
-on bijections or surjections.
-This is purely a matter of perspective, however: the definition above is
-precisely equivalent to a split surjection from a finite prefix of the natural
-numbers.
+First, we will look at how this predicate relates to more traditional, classical
+notions of finiteness.
+in a classical setting we likely wouldn't mention ``lists'' or the like, and
+would instead define finiteness based on the existence of some injection or
+surjection.
+As it turns out, our definition of finiteness here is precisely the same as the
+surjection-based one, in quite a deep way!
+
+First, we will need to define our terms: in HoTT, surjections are a little more
+complex than what you'd find in either MLTT or classical mathematics.
 \begin{definition}[Surjections] \label{surjections}
@@
 \end{definition}
+Over sets, the surjections and split surjections are the same thing, but there
+is a difference one we involve non-set types like the circle.
+
+Next, we will show that split enumerability is equivalent to the existence of a
+split surjection from a finite prefix of the natural numbers:
 \begin{lemma} \label{split-enum-is-split-surj}
@@
 \end{agdalisting}
+(The simplicity of this proof, by the way, is why we preferred the
+container-based definition of lists over the traditional one.)
 
@@
 \end{lemma}
+\begin{proof}
+  \todo[inline]{Prose for proof}
+\end{proof}
 \begin{lemma} \label{split-enum-discrete}

### 2020-05-31 sections/finiteness-predicates.tex
Update finiteness-predicates.tex

@@
 formalisation:
+\begin{definition}[Injections]
+  Injective functions are more straightforward to define constructively than
+  surjective ones:
+  \begin{alignat}{3}
+    & \text{injective}(f) &&\coloneqq \Pi (x, y : A) , f \; x \equiv f \; y \rightarrow x \equiv y \\
+    & A \rightarrowtail B &&\coloneqq \Sigma(f : A \rightarrow B) , \text{injective}(f)
+  \end{alignat}
+\end{definition}
+\begin{lemma} \label{split-surj-to-inj}
+  A split-surjection from \(A\) to \(B\) implies an injection from \(B\) to
+  \(A\).
+  \begin{equation}
+    (A \twoheadrightarrow! \; B) \rightarrow (B \rightarrowtail A)
+  \end{equation}
+\end{lemma}
+\begin{lemma} \label{inj-discrete}
+  For any type \(A\) which injects into a discrete type \(B\), \(A\) is
+  discrete.
+  \begin{equation}
+    \frac{
+      A \rightarrowtail B \; \; \; \text{Discrete}(B)
+    }{
+      \text{Discrete}(A)
+    }
+  \end{equation}
+\end{lemma}
+
 \begin{lemma} \label{discrete-surj}
@@
 \begin{proof}
-  \todo[inline]{Prose for proof}
+  This proof is can be straightforwardly derived from lemmas
+  \ref{split-surj-to-inj} and \ref{inj-discrete}.
 \end{proof}
+
 \begin{lemma} \label{split-enum-discrete}
@@
 \subsection{Manifest Bishop Finiteness}
+We mentioned in the introduction that occasionally in constructive mathematics
+proofs will contain ``too much'' information.
+With split enumerability we can see an instance of this.
+Consider our proof that the booleans are finite:
+
 \begin{definition}[Manifest Bishop Finiteness]

### 2020-05-31 sections/finiteness-predicates.tex
up to bishop finiteness in finiteness predicates done

@@
 \end{proof}
+
 \subsection{Manifest Bishop Finiteness}
@@
 With split enumerability we can see an instance of this.
-Consider our proof that the booleans are finite:
 
-\begin{definition}[Manifest Bishop Finiteness]
-  \begin{equation} \label{bish-def}
+\begin{agdalisting} \label{bool-slop}
+  Consider the following two proofs of the finiteness of Bool:
+  \begin{multicols}{2}
+    \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{bool-slop}
+    \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{bool-inst}
+  \end{multicols}
+  While they both represent the ``same'' information (i.e. they both tell us
+  that bool is finite), they are nonetheless not definitionally equal.
+\end{agdalisting}
+
+There is ``slop'' in the type of split enumerability: there are more distinct
+values than there are \emph{usefully} distinct values.
+As we can see in the example above, for instance, split enumerability allows
+duplicate values in the support list.
+To reconcile this, we will disallow duplicates in the support list.
+
+How exactly we should do this is the next question.
+One approach might be to change the definition of \(\mathbf{List}\), or
+introduce a new type \(\mathbf{NoDupeList}\), and use it in the predicate
+instead.
+However, this would mean we lose access to the functions we have defined on
+lists, and we have to change the definition of \(\in\) as well.
+
+There is a much simpler and more elegant solution: we insist that every
+\emph{membership proof} must be unique.
+This would disallow a definition of \(\mathcal{E}!(\mathbf{Bool})\) with
+duplicates, as there are multiple values which inhabit the type \(\text{false}
+\in \left[ \text{false}, \text{true}, \text{false} \right]\).
+It also allows us to keep most of the split enumerability definition unchanged,
+just adding a condition to the returned membership proof in the cover proof.
+
+To specify that a value must exist uniquely in HoTT we can use the concept of a
+\emph{contraction}.
+\begin{definition}[Unique Membership] \label{uniq-memb-def}
+  \begin{equation}
+    x \inunique \mathit{xs} \coloneqq \text{isContr}(x \in \mathit{xs})
+  \end{equation}
+\end{definition}
+With this we can define manifest Bishop finiteness:
+\begin{definition}[Manifest Bishop Finiteness]  \label{bish-def}
+  \begin{equation}
     \mathcal{B}(A) \coloneqq \Sigma {(\mathit{xs} : \textbf{List}(A))} , \Pi {(x : A)} , x \inunique xs
@@
 rather than simple membership (\(\in\)).
-\begin{definition}[Unique Membership] \label{uniq-memb-def}
-  \begin{equation}
-    x \inunique \mathit{xs} \coloneqq \text{isContr}(x \in \mathit{xs})
-  \end{equation}
-\end{definition}
 
@@
 
-Where split enumerability was the enumeration form of a split surjection from
-\(\mathbf{Fin}\), manifest Bishop finiteness is the enumeration form of an
+\subsubsection{The Relationship Between Manifest Bishop Finiteness and Split
+  Enumerability}
+While manifest Bishop finiteness might seem stronger than split enumerability,
+it turns out this is not the case.
+Both predicates imply the other.
+\begin{lemma} \label{manifest-bishop-to-split-enum}
+  Any manifest Bishop finite type is split enumerable.
+\end{lemma}
+\begin{proof}
+  To construct a proof of split enumerability from one of manifest Bishop
+  finiteness, it suffices to convert a proof of \(x \inunique \mathit{xs}\) to
+  one of \(x \in \mathit{xs}\), for all \(x\) and \(\mathit{xs}\).
+  Since \(\inunique\) is defined as a contraction of \(\in\), such a conversion
+  is simply the \(\text{fst}\) function.
+\end{proof}
+
+\begin{lemma} \label{split-enum-to-manifest-bishop}
+  Any split enumerable set is manifest Bishop finite.
+\end{lemma}
+This proof takes significantly more work.
+The ``unique membership'' condition in \(\mathcal{B}\) means that we are not
+permitted duplicates in the support list.
+The first step in the proof, then, is to filter those duplicates out from the
+support list of the \(\mathcal{E}!\) proof: we can do this using the decidable
+equality provided by \(\mathcal{E}!\) (lemma~\ref{split-enum-discrete}).
+From there, we need to show that the membership proof carries over
+appropriately.
+\todo{Provide more info on this proof?}
+
+We have now proved that every manifestly Bishop finite type is split enumerable,
+and vice versa.
+While the types are not \emph{equivalent} (there are more split enumerable
+proofs than there are manifest Bishop finite proofs), they are of equal power,
+so any closure proof we have on one can be transferred to the other.
+In particular, it means that manifest Bishop finiteness is closed under
+\(\Sigma\).
+\subsubsection{From Manifest Bishop Finiteness to Equivalence}
+We have seen that split enumerability was in fact a split-surjection in
+disguise.
+We will now see that manifest Bishop finiteness is in fact an \emph{equivalence}
+in disguise.
 \emph{equivalence} with \(\mathbf{Fin}\).
@@
 \end{lemma}
-This proof is effectively the same as that of
-lemma~\ref{split-enum-is-split-surj}.
+\begin{proof}
+  \begin{align*}
+     \mathcal{B}(A) &
+    \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , x \inunique \mathit{xs}
+    && \text{def.~\ref{bish-def} }(\mathcal{B})
+    \\
+    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{isContr}(x \in \mathit{xs})
+    && \text{eqn.~\ref{uniq-memb-def} } (\inunique)
+    \\
+    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{isContr}(\text{fib}_{\text{snd}(\mathit{xs})}(x))
+    && \text{eqn.~\ref{container-membership} } (\in)
+    \\
+    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
+    && \text{eqn.~\ref{is-equiv-def} (isEquiv)}
+    \\
+    & \simeq \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket (A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
+    && \text{def.~\ref{List} } (\mathbf{List})
+    \\
+    & \simeq \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \mathbf{Fin}(n)) , A) , \text{isEquiv}(\text{snd}(\mathit{xs}))
+    && \text{eqn.~\ref{container-interp} } (\llbracket \wc \rrbracket)
+    \\
+    & \simeq \Sigma (n : \mathbb{N}) , \Sigma (f : \mathbf{Fin}(n) \rightarrow A) , \text{isEquiv}(f)
+    && \text{Reassociation of } \Sigma
+    \\
+    & \simeq \Sigma (n : \mathbb{N}) , ( \mathbf{Fin}(n) \simeq A )
+    && \text{eqn.~\ref{equiv-def} } (\simeq)
+  \end{align*}
+\end{proof}
+This proof is almost identical\footnotemark to the proof for
+lemma~\ref{split-enum-is-split-surj}: it reveals that
+enumeration-based finiteness predicates are simply another perspective on
+relation-based ones.
+\footnotetext{
+  Unfortunately in our formalisation this proof cannot be a single line: for
+  performance reasons \(\simeq\) is defined as a record type with eta-equality
+  disabled, instead of the definition here which uses \(\Sigma\).
+}
 \subsection{Cardinal Finiteness}
@@
 \subsection{Split Enumerability and Manifest Bishop Finiteness}
-While manifest Bishop finiteness might seem stronger than split enumerability,
-it turns out this is not the case.
-Both predicates imply the other.
-\begin{lemma} \label{manifest-bishop-to-split-enum}
-  Any manifest Bishop finite type is split enumerable.
-\end{lemma}
-\begin{proof}
-  To construct a proof of split enumerability from one of manifest Bishop
-  finiteness, it suffices to convert a proof of \(x \inunique \mathit{xs}\) to
-  one of \(x \in \mathit{xs}\), for all \(x\) and \(\mathit{xs}\).
-  Since \(\inunique\) is defined as a contraction of \(\in\), such a conversion
-  is simply the \(\text{fst}\) function.
-\end{proof}
-
-\begin{lemma} \label{split-enum-to-manifest-bishop}
-  Any split enumerable set is manifest Bishop finite.
-\end{lemma}
-This proof takes significantly more work.
-The ``unique membership'' condition in \(\mathcal{B}\) means that we are not
-permitted duplicates in the support list.
-The first step in the proof, then, is to filter those duplicates out from the
-support list of the \(\mathcal{E}!\) proof: we can do this using the decidable
-equality provided by \(\mathcal{E}!\) (lemma~\ref{split-enum-discrete}).
-From there, all we need to show is that the membership proof carries over
-appropriately.
 \subsection{Split Enumerability and Manifest Enumerability}

### 2020-05-31 sections/introduction.tex
up to bishop finiteness in finiteness predicates done

@@
     &\text{isEquiv}(f) &&\coloneqq \Pi(y : B) , \text{isContr}(\text{fib}_f(y)) \label{is-equiv-def} \\
-    &A \simeq B        &&\coloneqq \Sigma(f : A \rightarrow B) , \text{isEquiv}(f)
+    &A \simeq B        &&\coloneqq \Sigma(f : A \rightarrow B) , \text{isEquiv}(f) \label{equiv-def}
   \end{alignat}

### 2020-05-31 sections/finiteness-predicates.tex
finished cardinal

@@
     & \simeq \Sigma (n : \mathbb{N}) , ( \mathbf{Fin}(n) \twoheadrightarrow ! \; A )
-    && \text{eqn.~\ref{sp-surj-arrow-eqn} } (\twoheadrightarrow!)
+    && \text{eqn.~\ref{sp-surj-arrow-eqn} } (\twoheadrightarrow!) \; \qedhere
   \end{align*}
@@
     & \simeq \Sigma (n : \mathbb{N}) , ( \mathbf{Fin}(n) \simeq A )
-    && \text{eqn.~\ref{equiv-def} } (\simeq)
+    && \text{eqn.~\ref{equiv-def} } (\simeq) \; \qedhere
   \end{align*}
@@
 \subsection{Cardinal Finiteness}
+While we have removed some of the unnecessary information from our finiteness
+predicates, one piece still remains.
+\begin{agdalisting}
+  The two following proofs are both valid proofs of the finiteness of
+  \(\mathbf{Bool}\), and both do not include any duplicates.
+  However they still differ:
+  \begin{multicols}{2}
+    \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{bool-inst}
+    \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{bool-rev}
+  \end{multicols}
+\end{agdalisting}
 Each finiteness predicate so far has contained an \emph{ordering} of the
@@
 \end{lemma}
+\begin{proof}
+  Let \(A\) be a cardinally-finite type, with proof \(F : \mathcal{C}(A)\).
+  Our task is to extract a natural number \(n : \mathbb{N}\) representing the
+  cardinality of \(A\), and a propositionally-truncated proof that \(A\) is
+  equivalent to \(\mathbf{Fin}(n)\).
+
+  Extracting the second component of the pair is trivial, as it itself is
+  truncated.
+  We will now focus on extracting the cardinality.
+
+  Without the propositional truncation, \(\text{fst}\) would suffice for this
+  task.
+  Given that the pair is hidden under the truncation, then, we need a way to
+  convert a function \(f : A \rightarrow B\) to \(g : \lVert A \rVert
+  \rightarrow B\).
+  This is precisely what eliminator \ref{elim-prop-coh} gives us.
+  For our case, we need to show the following:
+  \begin{equation}
+    \frac{(n : \mathbb{N}) \; \; \; (p : \mathbf{Fin}(n) \simeq A) \; \; \;
+          (m : \mathbb{N}) \; \; \; (q : \mathbf{Fin}(m) \simeq A)
+        }{
+          n \equiv m
+        }
+  \end{equation}
+  Immediately we can construct the following term:
+  \begin{equation}
+    \begin{alignat}{3}
+      \mathbf{Fin}(n) & \simeq A && (p) \\
+                      & \simeq \mathbf{Fin}(m) && (q)
+    \end{alignat}
+  \end{equation}
+  Given univalence we have \(\mathbf{Fin}(n) \equiv \mathbf{Fin}(m)\),
+  and the rest of our task is to prove:
+  \begin{equation}
+    \frac{\mathbf{Fin}(n) \equiv \mathbf{Fin}(m)}{n \equiv m}
+  \end{equation}
+
+  This is a well-known chestnut in dependently-typed programming, and one that
+  has a surprisingly tricky and complex proof.
+  We do not include it here, since it has already been explored elsewhere, but
+  it is present in our formalisation.
+\end{proof}
 \begin{lemma} \label{cardinal-finite-discrete}
@@
 \end{lemma}
+\begin{proof}
+  Since we can already derive decidable equality from a proof of manifest Bishop
+  finiteness, it suffices to show that decidable equality is itself a
+  proposition.
+  \begin{equation}
+    \text{isProp}(\Pi(x, y : A) , \mathbf{Dec}(x \equiv y))
+  \end{equation}
+  First, it is clear that \(x \equiv y\) is a proposition: since the type \(A\)
+  has decidable equality, by Hedburg's theorem \todo{ref here} it is a set,
+  meaning precisely that \(x \equiv y\) is a proposition.
+
+  Secondly, we know that any decision over a proposition is itself a
+  proposition.
+  For any two terms \(x, y: \mathbf{Dec}(A)\) we cannot have the case that one
+  is a yes decision and the other is no: from that we could derive \(\bot\).
+  If both are no then they are both equal since \(A \rightarrow \bot\) is a
+  proposition through function extensionality.
+  And finally if both are yes then we know they must be equal because the type
+  decided over is itself a proposition.
+
+  Finally, since we know that \(\mathbf{Dec}(x \equiv y)\) is a proposition, we
+  can derive that \(\Pi(x, y : A) , \mathbf{Dec}(x \equiv y)\) is a proposition
+  (through function extensionality), proving our goal.
+\end{proof}
 \subsection{Manifest Enumerability}

### 2020-05-31 sections/countable-predicates.tex
switched to chapters

@@
-\section{Countably Infinite Types} \label{infinite}
+\chapter{Countably Infinite Types} \label{infinite}
 \begin{figure*}
@@
 relations to \(\mathbb{N}\), in the form of the countably infinite types.
-\subsection{Two Countable Types}
+\section{Two Countable Types}
 The two types for countability we will consider are analogous to split
@@
 \end{lemma}
-\subsection{Closure}
+\section{Closure}
 We know that countable infinity is not closed under the exponential (function

### 2020-05-31 sections/finiteness-predicates.tex
switched to chapters

@@
-\section{Finiteness Predicates} \label{finiteness-predicates}
+\chapter{Finiteness Predicates} \label{finiteness-predicates}
 In this section, we will define and briefly describe each of the five predicates
@@
 tell us the number of elements in the finite set?).
-\subsection{Split Enumerability}
+\section{Split Enumerability}
 We will start with a simple notion of finiteness, called \emph{split}
@@
 \end{definition}
-\subsubsection{Instances}
+\subsection{Instances}
 Now that we have a suitable definition of finiteness, we will next prove that
@@
 \end{proof}
-\subsubsection{Derivations}
+\subsection{Derivations}
 We have a way to construct finiteness proofs, and a semiring-like toolbox to
@@
 
-\subsection{Manifest Bishop Finiteness}
+\section{Manifest Bishop Finiteness}
 We mentioned in the introduction that occasionally in constructive mathematics
@@
 
-\subsubsection{The Relationship Between Manifest Bishop Finiteness and Split
+\subsection{The Relationship Between Manifest Bishop Finiteness and Split
   Enumerability}
@@
 \(\Sigma\).
-\subsubsection{From Manifest Bishop Finiteness to Equivalence}
+\subsection{From Manifest Bishop Finiteness to Equivalence}
 We have seen that split enumerability was in fact a split-surjection in
@@
 }
-\subsection{Cardinal Finiteness}
+\section{Cardinal Finiteness}
 While we have removed some of the unnecessary information from our finiteness
@@
 \end{proof}
-\subsection{Manifest Enumerability}
+\subsection{Restrictiveness}
+So far our explorations into finiteness predicates have pushed us in the
+direction of ``less informative'': however, as mentioned in the introduction, we
+can \emph{also} ask how \emph{restrictive} certain predicates are.
+Since split enumerability and manifest Bishop finiteness imply each other we
+know that there can be no type which satisfies one but not the other.
+We also know that manifest Bishop finiteness implies cardinal finiteness, but we
+do \emph{not} have a function in the other direction:
+\begin{equation}
+  \mathcal{C}(A) \rightarrow \mathcal{B}(A)
+\end{equation}
+So the question arises naturally: is there a cardinally finite type which is
+\emph{not} manifest Bishop finite?
+
+It turns out the answer is no!
+\begin{lemma}
+  \begin{equation}
+    \neg (\Sigma(A : \mathbf{Type}) , \mathcal{C}(A) \times \neg \mathcal{B}(A))
+  \end{equation}
+\end{lemma}
+\begin{proof}
+  We will actually prove a slightly more general statement.
+  For any type \(A\), the following holds:
+  \begin{equation}
+    \neg (\lVert A \rVert \times \neg A)
+  \end{equation}
+  The solution becomes more clear if we write out the definition of \(\neg\):
+  \begin{equation}
+    \frac{\lVert A \rVert \;\;\; A \rightarrow \bot }{\bot}
+  \end{equation}
+  We clearly need to apply a function of type \(A \rightarrow \bot\) to a value
+  of type \(\lVert A \rVert\).
+  Luckily, this is permissible, as \(\bot\) is a mere proposition.
+\end{proof}
+\section{Manifest Enumerability}
+Since we have now discovered that all three predicates defined so far have equal
+restrictiveness, the clear next step is to
+
+So far, we have pushed finiteness predicates in the direction of making them
+``less informative'': we have not, however,
+
 \begin{definition}[Manifest Enumerability]
@@
 \end{lemma}
-\subsection{Kuratowski Finiteness}
+\section{Kuratowski Finiteness}
 The first thing we must define is a representation of subsets.
@@
 We will now look at the arrows in figure~\ref{finite-classification}.
-\subsection{Split Enumerability and Manifest Bishop Finiteness}
-\subsection{Split Enumerability and Manifest Enumerability}
+\section{Split Enumerability and Manifest Bishop Finiteness}
+\section{Split Enumerability and Manifest Enumerability}
 \begin{lemma} \label{split-enum-to-manifest-enum}
@@
 
-\subsection{Manifest Bishop Finiteness and Cardinal Finiteness}
+\section{Manifest Bishop Finiteness and Cardinal Finiteness}
 \begin{lemma} \label{manifest-bishop-to-cardinal}
@@
 The cover proof emerges naturally from the definition of the permutation.
-\subsection{Cardinal Finiteness and Kuratowski Finiteness}
+\section{Cardinal Finiteness and Kuratowski Finiteness}
 \begin{lemma} \label{cardinal-kuratowski}

### 2020-05-31 sections/introduction.tex
switched to chapters

@@
-\section{Introduction}
-\subsection{Foreword}
+\chapter{Introduction}
+\section{Foreword}
 Proofs in constructive mathematics are often more substantial than their
@@
 \todo{rephrase}
-\subsection{Contributions}
+\section{Contributions}
 We are interested in constructive notions of finiteness.
@@
 
-\subsubsection{Finiteness Predicates}
+\subsection{Finiteness Predicates}
 Our first contribution is a cataloguing and characterisation of five finiteness
@@
 
-\subsubsection{Countability Predicates}
+\subsection{Countability Predicates}
 After the finite predicates, we will briefly look at the infinite countable
@@
 (section~\ref{infinite}).
-\subsubsection{Search}
+\subsection{Search}
 All of our work is formalised in Cubical Agda
@@
 \url{https://github.com/oisdk/finiteness-in-cubical-type-theory}.
-\subsection{Notation and Background}
-\subsubsection{Notation}
+\section{Notation and Background}
+\subsection{Notation}
 We work in Cubical Type Theory \cite{cohenCubicalTypeTheory2016}.
@@
 \end{description}
-\subsubsection{Cubical Type Theory}
+\subsection{Cubical Type Theory}
 Cubical Type Theory \cite{cohenCubicalTypeTheory2016} is a constructive type

### 2020-05-31 sections/topos.tex
switched to chapters

@@
-\section{Topos} \label{topos}
+\chapter{Topos} \label{topos}
 In this section we will prove that decidable Kuratowski finite types form a
@@
 operations.
-\subsection{Closure}
+\section{Closure}
 For the first three closure proofs, we only consider split enumerability:
@@
 \end{lemma}
-\subsection{The Category of Finite Sets}
+\section{The Category of Finite Sets}
 HoTT and CuTT seem to be especially suitable settings for formalisations of
@@
 \end{theorem}
-\subsection{The \(\Pi\)-pretopos of Finite Sets}
+\section{The \(\Pi\)-pretopos of Finite Sets}
 For this proof, we follow again the proof that \(\mathit{Set}\) forms a \(\Pi

### 2020-06-01 sections/finiteness-predicates.tex
now compiles without errors

@@
 From that we can derive a much wider array of finiteness proofs.
+
 \begin{lemma} \label{split-enum-sigma}
@@
   \begin{equation}
-    \begin{alignat}{3}
+    \begin{alignedat}{3}
       \mathbf{Fin}(n) & \simeq A && (p) \\
                       & \simeq \mathbf{Fin}(m) && (q)
-    \end{alignat}
+    \end{alignedat}
   \end{equation}
@@
   First, it is clear that \(x \equiv y\) is a proposition: since the type \(A\)
-  has decidable equality, by Hedburg's theorem \todo{ref here} it is a set,
+  has decidable equality, by Hedburg's theorem it is a set,
   meaning precisely that \(x \equiv y\) is a proposition.

### 2020-06-01 sections/finiteness-predicates.tex
rearranged more

@@
 }
+
+As we are working in CuTT, a proof of equivalence between two types gives us the
+ability to \emph{transport} proofs from one type to the other.
+This is extremely powerful, as we will see.
+\subsection{Closure Under \(\Pi\)}
+The glaring omission from our closure proofs under type formers so far has been
+the \(\Pi\) type: we have not proved closure under functions, dependent or
+otherwise.
+In MLTT, this is of course not provable: since all of the finiteness predicates
+we have seen so far imply decidable equality, and since we don't have any kind
+of decidable equality on functions in MLTT, we know that we won't be able to
+show that any kind of function is finite; even one like \(\mathbf{Bool}
+\rightarrow \mathbf{Bool}\).
+
+CuTT is not so restricted.
+Since we have things like function extensionality and transport, we can indeed
+prove the finiteness of function types.
+Our proof here makes use directly of the univalence axiom, and makes use
+furthermore of all the previous closure proofs.
+We will prove this closure on split enumerability, rather than on manifest
+Bishop finiteness, as it requires slightly less legwork in the proof itself, but
+of course we can derive the proof on manifest Bishop finiteness in a few lines.
+\begin{lemma} \label{split-enum-pi}
+  Split enumerability is closed under dependent functions.
+  (\(\Pi\)-types).
+  \begin{equation}
+    \frac{
+      \mathcal{E}!(A) \; \; \; \Pi {(x : A)} , \mathcal{E}!\left( U(x) \right)
+    }{
+      \mathcal{E}!\left(\Pi {(x : A)} , U(x)\right)
+    }
+  \end{equation}
+\end{lemma}
+\begin{proof}
+  Let \(A\) be a split enumerable type, and \(U\) be a type family from \(A\),
+  which is split enumerable over all points of \(A\).
+
+  As \(A\) is split enumerable, we know that it is also manifestly Bishop finite
+  (lemma~\ref{split-enum-to-manifest-bishop}), and consequently we know \(A
+  \simeq \mathbf{Fin}(n)\), for some \(n\) (lemma~\ref{bishop-equiv}).
+  We can therefore replace all occurrences of \(A\) with \(\mathbf{Fin}(n)\),
+  changing our goal to:
+  \begin{equation}
+    \frac{
+      \mathcal{E}!(\mathbf{Fin}(n)) \; \; \; \Pi (x : \mathbf{Fin}(n)) , \mathcal{E}!\left( U(x) \right)
+    }{
+      \mathcal{E}!\left(\Pi (x : \mathbf{Fin}(n)) , U(x)\right)
+    }
+  \end{equation}
+
+  We then define the type of \(n\)-tuples over some type family \(T :
+  \mathbf{Fin}(n) \rightarrow \mathbf{Type}\).
+  \begin{equation}
+    \begin{alignedat}{3}
+      & \mathbf{Tuple}(0, T)   &&\coloneqq \top \\
+      & \mathbf{Tuple}(n+1, T) &&\coloneqq T(0) \times \mathbf{Tuple}(n, T \circ \text{suc})
+    \end{alignedat}
+  \end{equation}
+  We can show that this type is equivalent to functions (proven in our formalisation):
+  \begin{equation}
+    \Pi(x : \mathbf{Fin}(n)) , U(x) \simeq \mathbf{Tuple}(n, U)
+  \end{equation}
+  And therefore we can simplify again our goal to the following:
+  \begin{equation}
+    \frac{
+      \mathcal{E}!(\mathbf{Fin}(n)) \; \; \; \Pi (x : \mathbf{Fin}(n)) , \mathcal{E}!\left( U(x) \right)
+    }{
+      \mathcal{E}!\left(\mathbf{Tuple}(n, U)\right)
+    }
+  \end{equation}
+
+  We can prove this goal by showing that \(\mathbf{Tuple}(n, U)\) is split
+  enumerable: it is made up of finitely many products of points of \(U\), which
+  are themselves split enumerable, and \(\top\), which is also split enumerable.
+  Lemma~\ref{split-enum-sigma} shows us that the product of finitely many split
+  enumerable types is itself split enumerable, proving our goal.
+\end{proof}
 \section{Cardinal Finiteness}
@@
 \end{proof}
+
+\subsection{Going from Cardinal Finiteness to Manifest Bishop Finiteness}
+\begin{lemma} \label{manifest-bishop-to-cardinal}
+  Any manifest Bishop finite type is cardinal finite.
+\end{lemma}
+\begin{theorem} \label{cardinal-to-manifest-bishop}
+  Any cardinal finite type with a total order is Bishop finite.
+\end{theorem}
+The proof for this particular theorem is quite involved in the formalisation, so
+we only give its sketch here.
+First, note that we actually convert to manifest enumerability first: this can
+be converted to split enumerability with decidable equality, which is provided
+by cardinal finiteness.
+
+Next, we define permutations.
+\begin{definition}[List Permutations]
+  Two lists are permutations of each other if their membership proofs are all
+  equivalent\footnotemark \cite{danielssonBagEquivalenceProofRelevant2012}.
+  \begin{equation}
+    \mathit{xs} \leftrightsquigarrow \mathit{ys} = \Pi {(x : A)} , x \in \mathit{xs} \simeq x \in \mathit{ys}
+  \end{equation}
+\end{definition}
+
+\footnotetext{
+  The definition in \cite{danielssonBagEquivalenceProofRelevant2012} and our
+  formalisation is slightly different: we say permutations are lists with
+  \emph{isomorphic} membership proofs.
+  The distinction, as it happens, does not affect our work here.
+}
+
+Next, we define a sort function which relies on the provided total order.
+We further prove the following fact about this sort function:
+\begin{equation}
+  \Pi(\mathit{xs}, \mathit{ys} : \mathbf{List}(A)) , \mathit{xs} \leftrightsquigarrow \mathit{ys} \rightarrow \text{sort}(\mathit{xs}) \equiv \text{sort}(\mathit{ys})
+\end{equation}
+
+Next, notice that the support lists of any two proofs of manifest Bishop
+finiteness must be permutations of each other.
+This will allow us to sort the support list of a proof of cardinal finiteness in
+a coherently constant (definition~\ref{prop-trunc},
+eliminator~\ref{elim-prop-coh}) way, pulling the support list out from the
+truncation.
+The cover proof emerges naturally from the definition of the permutation.
+\subsection{Closure}
+Since we don't have a function of type \(\mathcal{C}(A) \rightarrow
+\mathcal{B}(A)\), closure proofs on \(\mathcal{B}\) do not transfer over to
+\(\mathcal{C}\) trivially (unlike with \(\mathcal{E}!\) and \(\mathcal{B}\)).
+The cases for \(\bot\), \(\top\), and \(\mathbf{Bool}\) are simple to adapt: we
+can just propositionally truncate their Bishop finiteness proof.
+
+Non-dependent operators like \(\times\), \(\uplus\), and \(\rightarrow\) are
+also relatively straightforward: since \(\lVert {\wc} \rVert\) forms a monad, we
+can apply \(n\)-ary functions to values inside it, combining them together.
+\begin{agdalisting}
+  The fact that \(\lVert \wc \rVert\) forms a monad means that we can lift
+  \(n\)-ary functions like the following:
+  \ExecuteMetaData[agda/Cardinality/Finite/ManifestBishop.tex]{times-clos-sig}
+  Into a truncated context:
+  \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{times-clos-impl}
+\end{agdalisting}
+
+Unfortunately, for the dependent type formers like \(\Sigma\) and \(\Pi\), the
+same trick does not work.
+We have closure proofs like:
+\begin{equation}
+  \frac{
+    \mathcal{B}(A) \; \; \; \Pi(x : A) , \mathcal{B}(U(x))
+  }{
+    \mathcal{B}(\Pi \; A \; U)
+  }
+\end{equation}
+If we apply the monadic truncation trick we can derive closure proofs like the
+following:
+\begin{equation}
+  \frac{
+    \lVert \mathcal{B}(A) \rVert \; \; \; \lVert \Pi(x : A) , \mathcal{B}(U(x)) \rVert
+  }{
+    \lVert \mathcal{B}(\Pi \; A \; U) \rVert
+  }
+\end{equation}
+However our \emph{desired} closure proof is the following:
+\begin{equation}
+  \frac{
+    \lVert \mathcal{B}(A) \rVert \; \; \; \Pi(x : A) , \lVert \mathcal{B}(U(x)) \rVert
+  }{
+    \lVert \mathcal{B}(\Pi \; A \; U) \rVert
+  }
+\end{equation}
+They don't match!
+
+The solution would be to find a function of the following type:
+\begin{equation}
+  (\Pi(x : A) , \lVert \mathcal{B}(U(x)) \rVert) \rightarrow
+  \lVert \Pi(x : A) , \mathcal{B}(U(x)) \rVert
+\end{equation}
+However we might be disheartened at realising that this is a required goal: the
+above equation is \emph{extremely} similar to the axiom of choice!
+\begin{definition}[Axiom of Choice]
+  In HoTT, the axiom of choice is commonly defined as follows \cite[lemma
+  3.8.2]{hottbook}.
+  For any set \(A\), and a type family \(U\) which is a set at all the points
+  of \(A\), the following function exists:
+  \begin{equation}
+    \left( \Pi(x : A) ,  \lVert U(x) \rVert \right) \rightarrow \lVert \Pi(x : A) , U(x) \rVert
+  \end{equation}
+\end{definition}
+Luckily the axiom of choice \emph{does} hold for cardinally finite types,
+allowing us to prove the following:
+\begin{lemma}
+  \begin{equation}
+    \mathcal{C}(A) \rightarrow (\Pi(x : A) , \lVert U(x) \rVert) \rightarrow \lVert \Pi(x : A) , U(x) \rVert
+  \end{equation}
+\end{lemma}
+\begin{proof}
+  Let \(A\) be a cardinally finite type, \(U\) be a type family on \(A\), and
+  \(f\) be a dependent function of type \(\Pi(x : A) , \lVert U(x) \rVert\).
+
+  First, since our goal is itself propositionally truncated, we have access to
+  values under truncations: put another way, in the context of proving our goal,
+  we can rely on the fact that \(A\) is manifestly Bishop finite.
+  Using the same technique as we did in lemma~\ref{split-enum-pi}, we can switch
+  from working with dependent functions from \(A\) to \(n\)-tuples, where \(n\)
+  is the cardinality of \(A\).
+  This changes our goal to the following:
+  \begin{equation}
+    \mathbf{Tuple}(n, \lVert \wc \rVert \circ U) \rightarrow \lVert \mathbf{Tuple}(n, U) \rVert
+  \end{equation}
+  Since \(\lVert \wc \rVert\) is closed under finite products, this function
+  exists (in fact, using the fact that \(\lVert \wc \rVert\) forms a monad, we
+  can recognise this function as \verb+sequenceA+ from the \verb+Traversable+
+  class in Haskell).
+\end{proof}
+
+This gets us all of the necessary closure proofs on \(\mathcal{C}\).
 \section{Manifest Enumerability}
-Since we have now discovered that all three predicates defined so far have equal
-restrictiveness, the clear next step is to
-
-So far, we have pushed finiteness predicates in the direction of making them
-``less informative'': we have not, however,
+We have now explored quite far in the ``less informative'' direction.
+However, all three predicates we have examined are equally \emph{restrictive}:
+in this section we will see a predicate which is much less restrictive.
+In particular, this predicate ranges over non-set types.
 
@@
   The circle \(S^1\) is manifestly enumerable.
+\end{lemma}
+\subsection{Relation To Split Enumerability }
+\begin{lemma} \label{split-enum-to-manifest-enum}
+  Any split enumerable type is manifestly enumerable.
+\end{lemma}
+This lemma is proven by truncating the membership proof in split enumerability.
+
+\begin{lemma} \label{manifest-enum-to-split-enum}
+  A manifestly enumerable type with decidable equality is split enumerable.
 \end{lemma}
@@
 \end{lemma}
-\section{Relations Between Each Finiteness Definition} \label{relations}
-We will now look at the arrows in figure~\ref{finite-classification}.
-\section{Split Enumerability and Manifest Bishop Finiteness}
-\section{Split Enumerability and Manifest Enumerability}
-\begin{lemma} \label{split-enum-to-manifest-enum}
-  Any split enumerable type is manifestly enumerable.
-\end{lemma}
-This lemma is proven by truncating the membership proof in split enumerability.
-
-\begin{lemma} \label{manifest-enum-to-split-enum}
-  A manifestly enumerable type with decidable equality is split enumerable.
-\end{lemma}
-
-\section{Manifest Bishop Finiteness and Cardinal Finiteness}
-\begin{lemma} \label{manifest-bishop-to-cardinal}
-  Any manifest Bishop finite type is cardinal finite.
-\end{lemma}
-\begin{theorem} \label{cardinal-to-manifest-bishop}
-  Any cardinal finite type with a total order is Bishop finite.
-\end{theorem}
-The proof for this particular theorem is quite involved in the formalisation, so
-we only give its sketch here.
-First, note that we actually convert to manifest enumerability first: this can
-be converted to split enumerability with decidable equality, which is provided
-by cardinal finiteness.
-
-Next, we define permutations.
-\begin{definition}[List Permutations]
-  Two lists are permutations of each other if their membership proofs are all
-  equivalent\footnotemark \cite{danielssonBagEquivalenceProofRelevant2012}.
-  \begin{equation}
-    \mathit{xs} \leftrightsquigarrow \mathit{ys} = \Pi {(x : A)} , x \in \mathit{xs} \simeq x \in \mathit{ys}
-  \end{equation}
-\end{definition}
-
-\footnotetext{
-  The definition in \cite{danielssonBagEquivalenceProofRelevant2012} and our
-  formalisation is slightly different: we say permutations are lists with
-  \emph{isomorphic} membership proofs.
-  The distinction, as it happens, does not affect our work here.
-}
-
-Next, we define a sort function which relies on the provided total order.
-We further prove the following fact about this sort function:
-\begin{equation}
-  \Pi(\mathit{xs}, \mathit{ys} : \mathbf{List}(A)) , \mathit{xs} \leftrightsquigarrow \mathit{ys} \rightarrow \text{sort}(\mathit{xs}) \equiv \text{sort}(\mathit{ys})
-\end{equation}
-
-Next, notice that the support lists of any two proofs of manifest Bishop
-finiteness must be permutations of each other.
-This will allow us to sort the support list of a proof of cardinal finiteness in
-a coherently constant (definition~\ref{prop-trunc},
-eliminator~\ref{elim-prop-coh}) way, pulling the support list out from the
-truncation.
-The cover proof emerges naturally from the definition of the permutation.
-\section{Cardinal Finiteness and Kuratowski Finiteness}
+\subsection{Relation to Cardinal Finiteness}
 \begin{lemma} \label{cardinal-kuratowski}

### 2020-06-01 sections/topos.tex
now compiles without errors
rearranged more

@@
 
-\begin{lemma}
-  \(\bot\), \(\top\), and \(\mathbf{Bool}\) are split enumerable.
-\end{lemma}
-\begin{lemma} \label{split-enum-sigma}
-  Split enumerability is closed under \(\Sigma\).
-  \begin{equation}
-    \frac{
-      \mathcal{E}!(A) \; \; \; \Pi(x : A) , \mathcal{E}!(U(x))
-    }{
-      \mathcal{E}!(\Sigma(x : A) , U(x))
-    }
-  \end{equation}
-\end{lemma}
-From this we can derive split enumerability of non-dependent sums and products,
-as both are definable in terms of \(\Sigma\).
+% \begin{lemma}
+%   \(\bot\), \(\top\), and \(\mathbf{Bool}\) are split enumerable.
+% \end{lemma}
+% \begin{lemma} \label{split-enum-sigma}
+%   Split enumerability is closed under \(\Sigma\).
+%   \begin{equation}
+%     \frac{
+%       \mathcal{E}!(A) \; \; \; \Pi(x : A) , \mathcal{E}!(U(x))
+%     }{
+%       \mathcal{E}!(\Sigma(x : A) , U(x))
+%     }
+%   \end{equation}
+% \end{lemma}
+% From this we can derive split enumerability of non-dependent sums and products,
+% as both are definable in terms of \(\Sigma\).
 
-\begin{lemma}
-  Split enumerability is closed under dependent functions.
-  (\(\Pi\)-types).
-  \begin{equation}
-    \frac{
-      \mathcal{E}!(A) \; \; \; \Pi {(x : A)} , \mathcal{E}!\left( U(x) \right)
-    }{
-      \mathcal{E}!\left(\Pi {(x : A)} , U(x)\right)
-    }
-  \end{equation}
-\end{lemma}
-\begin{proof}
-  Let \(A\) be a split enumerable type, and \(U\) be a type family from \(A\),
-  which is split enumerable over all points of \(A\).
-
-  As \(A\) is split enumerable, we know that it is also manifestly Bishop finite
-  (lemma~\ref{split-enum-to-manifest-bishop}), and consequently we know \(A
-  \simeq \mathbf{Fin}(n)\), for some \(n\) (lemma~\ref{bishop-equiv}).
-  We can therefore replace all occurrences of \(A\) with \(\mathbf{Fin}(n)\),
-  changing our goal to:
-  \begin{equation}
-    \frac{
-      \mathcal{E}!(\mathbf{Fin}(n)) \; \; \; \Pi (x : \mathbf{Fin}(n)) , \mathcal{E}!\left( U(x) \right)
-    }{
-      \mathcal{E}!\left(\Pi (x : \mathbf{Fin}(n)) , U(x)\right)
-    }
-  \end{equation}
-
-  We then define the type of \(n\)-tuples over some type family \(T :
-  \mathbf{Fin}(n) \rightarrow \mathbf{Type}\).
-  \begin{equation}
-    \begin{alignedat}{3}
-      & \mathbf{Tuple}(0, T)   &&\coloneqq \top \\
-      & \mathbf{Tuple}(n+1, T) &&\coloneqq T(0) \times \mathbf{Tuple}(n, T \circ \text{suc})
-    \end{alignedat}
-  \end{equation}
-  We can show that this type is equivalent to functions (proven in our formalisation):
-  \begin{equation}
-    \Pi(x : \mathbf{Fin}(n)) , U(x) \simeq \mathbf{Tuple}(n, U)
-  \end{equation}
-  And therefore we can simplify again our goal to the following:
-  \begin{equation}
-    \frac{
-      \mathcal{E}!(\mathbf{Fin}(n)) \; \; \; \Pi (x : \mathbf{Fin}(n)) , \mathcal{E}!\left( U(x) \right)
-    }{
-      \mathcal{E}!\left(\mathbf{Tuple}(n, U)\right)
-    }
-  \end{equation}
-
-  We can prove this goal by showing that \(\mathbf{Tuple}(n, U)\) is split
-  enumerable: it is made up of finitely many products of points of \(U\), which
-  are themselves split enumerable, and \(\top\), which is also split enumerable.
-  Lemma~\ref{split-enum-sigma} shows us that the product of finitely many split
-  enumerable types is itself split enumerable, proving our goal.
-\end{proof}
-
-Lifting the above closure proofs to work on proofs of \(\mathcal{C}\) is not
-necessarily straightforward.
-\(\lVert \wc \rVert\) forms a monad, giving us access to a powerful syntax for
-combining proofs.
-From this, we can derive closure under \(\bot\), \(\top\), \(\mathbf{Bool}\),
-\(\times\), \(\uplus\), and \(\rightarrow\).
-The dependent case is not so straightforward.
-We need the following lemma, a version of the axiom of choice on finite sets,
-to prove closure under \(\Sigma\) and \(\Pi\).
-\begin{lemma}
-  \begin{equation}
-    \mathcal{C}(A) \rightarrow (\Pi(x : A) , \lVert U(x) \rVert) \rightarrow \lVert \Pi(x : A) , U(x) \rVert
-  \end{equation}
-\end{lemma}
 \section{The Category of Finite Sets}

### 2020-06-01 sections/finiteness-predicates.tex
cleaner format

@@
 \end{proof}
-\begin{agdalisting}
-  The code in our formalisation of these three proofs is quite simple.
-  Just as an example, the proof that \(\mathbf{Bool}\) is split enumerable is as
-  follows:
+\begin{marginfigure}%
+  \begin{dottedbox}%
+    % The code in our formalisation of these three proofs is quite simple.
+    % Just as an example, the proof that \(\mathbf{Bool}\) is split enumerable is as
+    % follows:
     \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{bool-inst}
-\end{agdalisting}
-
+  \end{dottedbox}
+\end{marginfigure}
 With the most basic simple types out of the way, the obvious next choice is the
@@
 
+  To prove that this support list does in fact cover the entirety of the type
+  \(\Sigma \; A \; U\), we note that any element of type \(\Sigma \; A \; U\)
+  must have a first component in the support list of \(\mathcal{E}!_A\), and its
+  second component must be in the result of applying \(\mathcal{E}!_U\) to that
+  first element (since that support list contains every element of type
+  \(U(x)\)).
+  Therefore, the pair itself must be in our constructed support list.
+\end{proof}
+
 \begin{agdalisting}
@@
 \end{agdalisting}
-
-  To prove that this support list does in fact cover the entirety of the type
-  \(\Sigma \; A \; U\), we note that any element of type \(\Sigma \; A \; U\)
-  must have a first component in the support list of \(\mathcal{E}!_A\), and its
-  second component must be in the result of applying \(\mathcal{E}!_U\) to that
-  first element (since that support list contains every element of type
-  \(U(x)\)).
-  Therefore, the pair itself must be in our constructed support list.
-\end{proof}
 \subsection{Derivations}
@@
 
-Next, we will show that split enumerability is equivalent to the existence of a
-split surjection from a finite prefix of the natural numbers:
-\begin{lemma} \label{split-enum-is-split-surj}
-  \begin{equation}
+We will now see that split enumerability is in fact a split surjection in
+another form:
+\begin{lemma} \label{split-enum-is-split-surj}%
+  A proof of split enumerability is equivalent to a split surjection from a
+  finite prefix of the natural numbers.
+  \begin{equation}%
     \mathcal{E}!(A) \simeq \Sigma (n : \mathbb{N}) , \left( \mathbf{Fin}(n) \twoheadrightarrow ! \; A \right)
@@
 \begin{definition}[Unique Membership] \label{uniq-memb-def}
+  Unique list membership is defined in terms of list membership: it is a
+  contraction of it.
   \begin{equation}
@@
 \begin{definition}[Manifest Bishop Finiteness]  \label{bish-def}
+  A type is manifest Bishop finite if there exists a list which contains each
+  value in the type once.
   \begin{equation}
@@
 \begin{lemma} \label{bishop-equiv}
+  Manifest bishop finiteness is equivalent to an equivalence to a finite prefix
+  of the natural numbers.
   \begin{equation}
@@
 \begin{definition}[Cardinal Finiteness]
+  A type \(A\) is cardinally finite if there exists a propositionally truncated
+  proof that \(A\) is manifest Bishop finite or equivalent to a finite prefix of
+  the natural numbers.
   \begin{equation}

### 2020-06-01 sections/finiteness-predicates.tex
agda code works better
fully margined listings

@@
 \end{proof}
-\begin{marginfigure}%
-  \begin{dottedbox}%
-    % The code in our formalisation of these three proofs is quite simple.
-    % Just as an example, the proof that \(\mathbf{Bool}\) is split enumerable is as
-    % follows:
+\begin{agdalisting}
+  The code in our formalisation of these three proofs is quite simple.
+  Just as an example, the proof that \(\mathbf{Bool}\) is split enumerable is as
+  follows:
   \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{bool-inst}
-  \end{dottedbox}
-\end{marginfigure}
+\end{agdalisting}
 With the most basic simple types out of the way, the obvious next choice is the
@@
   \end{align}
-
   To construct the support list, we apply the function \(\mathcal{E}!_U\) to
@@
 \end{agdalisting}
+
 \subsection{Derivations}
@@
   Consider the following two proofs of the finiteness of Bool:
-  \begin{multicols}{2}
   \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{bool-slop}
   \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{bool-inst}
-  \end{multicols}
   While they both represent the ``same'' information (i.e. they both tell us
@@
   However they still differ:
-  \begin{multicols}{2}
   \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{bool-inst}
   \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{bool-rev}
-  \end{multicols}
 \end{agdalisting}

### 2020-06-02 sections/finiteness-predicates.tex
Update finiteness-predicates.tex

@@
   \begin{alignat}{3}
-    &\text{surj}(f)             &&\coloneqq \Pi(y : B) , \lVert \text{fib}_f(y) \rVert \\
+    &\text{surj}(f)             &&\coloneqq \Pi(y : B) , \lVert \text{fib}_f(y) \rVert \label{surj-eqn} \\
     &A \twoheadrightarrow B     &&\coloneqq \Sigma (f : A \rightarrow B) , \text{surj}(f) \label{surj-arrow-eqn} \\
@@
 \begin{definition}[Manifest Enumerability]
+  Manifest enumerability is an enumeration predicate like Bishop finiteness or
+  split enumerability with the only difference being a propositionally truncated
+  membership proof.
   \begin{equation}
@@
 \end{definition}
-As with manifest Bishop finiteness, the only difference with this type and split
-enumerability is the membership proof: here we have propositionally truncated
-it.
-This has two effects.
-First, it means that this proof represents a true surjection (rather than a
-split surjection) from \(\mathbf{Fin}\).
+As a function-based definition, this predicate represents surjections.
 \begin{lemma}
+  Manifest enumerability is equivalent to a surjection from a finite prefix of
+  the natural numbers.
   \begin{equation}
@@
 \end{lemma}
-The proof for this lemma is similar in structure to
-lemma~\ref{split-enum-is-split-surj} and lemma~\ref{bishop-equiv}.
+\begin{proof}
+  \begin{align*}
+     \mathcal{E}(A) &
+    \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \lVert x \in \mathit{xs} \rVert
+    && \text{def.~\ref{split-enum-def} }(\mathcal{E})
+    \\
+    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \lVert \text{fib}_{\text{snd}(\mathit{xs})}(x) \rVert
+    && \text{eqn.~\ref{container-membership} } (\in)
+    \\
+    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \text{surj}(\text{snd}(\mathit{xs}))
+    && \text{eqn.~\ref{surj-eqn} (surj)}
+    \\
+    & \simeq \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket (A)) , \text{surj}(\text{snd}(\mathit{xs}))
+    && \text{def.~\ref{List} } (\mathbf{List})
+    \\
+    & \simeq \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \mathbf{Fin}(n)) , A) , \text{surj}(\text{snd}(\mathit{xs}))
+    && \text{eqn.~\ref{container-interp} } (\llbracket \wc \rrbracket)
+    \\
+    & \simeq \Sigma (n : \mathbb{N}) , \Sigma (f : \mathbf{Fin}(n) \rightarrow A) , \text{surj}(f)
+    && \text{Reassociation of } \Sigma
+    \\
+    & \simeq \Sigma (n : \mathbb{N}) , ( \mathbf{Fin}(n) \twoheadrightarrow A )
+    && \text{eqn.~\ref{surj-arrow-eqn} } (\twoheadrightarrow) \; \qedhere
+  \end{align*}
+\end{proof}
 
+The other implication of this truncation is that the type we have defined does
+not, in fact, imply decidable equality unlike all of those we have seen so far.
 Secondly, it means the predicate does not imply decidable equality.

### 2020-06-02 sections/finiteness-predicates.tex
Update finiteness-predicates.tex

--- 2020-06-01 sections/finiteness-predicates.tex
@@
 \begin{agdalisting} \label{bool-slop}
-  Consider the following two proofs of the finiteness of Bool:
+  Consider the following proof of the finiteness of Bool:
   \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{bool-slop}
-  \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{bool-inst}
-  While they both represent the ``same'' information (i.e. they both tell us
-  that bool is finite), they are nonetheless not definitionally equal.
+  While it represents the ``same'' information as our previous proof, it is
+  clearly not the same \emph{object}.
 \end{agdalisting}
@@
   However they still differ:
-  \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{bool-inst}
+  \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{bool-inst} \smallskip
   \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{bool-rev}
@@
 def.~\ref{prop-trunc}, we need only show that the output is uniquely determined.
-
+\subsection{Deriving Uniquely-Determined Quantities}
 The following two lemmas are proven in
@@
 \end{proof}
+
+In order to prove that cardinal finiteness implies decidable equality, we will
+need to show that decidable equality itself is a proposition.
+In doing that we will use the following lemma:
+\begin{lemma} \label{prop-refute}
+  We can ``refute'' a propositionally-truncated proof of some proposition with a
+  proof that the non-truncated proposition is false.
+  \begin{equation}
+    \frac{\neg A \; \; \; \lVert A \rVert}{\bot}
+  \end{equation}
+\end{lemma}
+\begin{proof}
+  We know we can eliminate from any value of type \(\lVert A \rVert\) into some
+  \(B\) with a function \(A \to B\) if \(B\) is a proposition.
+  That's precisely what we do in this case: \(\neg A\) is a function of type \(A
+  \to \bot\), and we know that \(\bot\) is a proposition.
+\end{proof}
+
 \begin{lemma} \label{cardinal-finite-discrete}
@@
 \end{lemma}
-The proof for this lemma is similar in structure to
-lemma~\ref{split-enum-is-split-surj} and lemma~\ref{bishop-equiv}.
-
-Secondly, it means the predicate does not imply decidable equality.
-More significantly, it allows the predicate to be defined over non-set types,
-like the circle.
+\begin{proof}
+  \begin{align*}
+     \mathcal{E}(A) &
+    \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \lVert x \in \mathit{xs} \rVert
+    && \text{def.~\ref{split-enum-def} }(\mathcal{E})
+    \\
+    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \lVert \text{fib}_{\text{snd}(\mathit{xs})}(x) \rVert
+    && \text{eqn.~\ref{container-membership} } (\in)
+    \\
+    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \text{surj}(\text{snd}(\mathit{xs}))
+    && \text{eqn.~\ref{surj-eqn} (surj)}
+    \\
+    & \simeq \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket (A)) , \text{surj}(\text{snd}(\mathit{xs}))
+    && \text{def.~\ref{List} } (\mathbf{List})
+    \\
+    & \simeq \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \mathbf{Fin}(n)) , A) , \text{surj}(\text{snd}(\mathit{xs}))
+    && \text{eqn.~\ref{container-interp} } (\llbracket \wc \rrbracket)
+    \\
+    & \simeq \Sigma (n : \mathbb{N}) , \Sigma (f : \mathbf{Fin}(n) \rightarrow A) , \text{surj}(f)
+    && \text{Reassociation of } \Sigma
+    \\
+    & \simeq \Sigma (n : \mathbb{N}) , ( \mathbf{Fin}(n) \twoheadrightarrow A )
+    && \text{eqn.~\ref{surj-arrow-eqn} } (\twoheadrightarrow) \; \qedhere
+  \end{align*}
+\end{proof}
+\subsection{Instances for Non-Set Types}
+The truncation has another very important implication: it means that the
+predicate doesn't provide decidable equality on the underlying type.
+Remember, this is how we knew that our previous predicates wouldn't allow for
+non-set types: because they implied decidable equality, they also implied that
+all conforming types had homotopy levels of at most 2. \todo{Are we doing the
+  homotopy levels starting from -2 thing or from 0?}
+This suggests that non-set types like the circle could conform to this
+finiteness predicate.
 \begin{definition}[\(S^1\)] \label{circle-def}
@@
 \end{lemma}
-\subsection{Relation To Split Enumerability }
-\begin{lemma} \label{split-enum-to-manifest-enum}
-  Any split enumerable type is manifestly enumerable.
+\begin{proof}
+  The support list firstly is a list containing the point constructor for the
+  circle.
+  Since the cover proof is truncated, we need only consider the point
+  constructors of the circle: as such, the cover proof is essentially the same
+  as the one for \(\mathcal{E}!(\top)\).
+\end{proof}
+\subsection{Relation To Split Enumerability}
+It is trivially easy to construct a proof that any split enumerable type is
+manifest enumerable: we simply truncate the membership proof.
+Going the other was requires us to extract a non-truncated proof from a
+truncated one.
+This proof relies on the following lemma:
+\begin{lemma}
+  We can ``recompute'' a truncated proof given a decision over a proof of the
+  same type.
+  \begin{equation}
+    \frac{\lVert A \rVert \; \; \; \mathbf{Dec}(A)}{A}
+  \end{equation}
 \end{lemma}
-This lemma is proven by truncating the membership proof in split enumerability.
+\begin{proof}
+  We proceed by case-analysis over the decision over \(A\).
+  In the case where \(A\) is proven, we are done.
+  In the case where \(A\) is disproven, we use lemma~\ref{prop-refute} to
+  derive impossibility.
+\end{proof}
 
@@
 \end{lemma}
+\begin{proof}
+  The only difference between manifest enumerability and split enumerability is
+  the membership proof: therefor our goal for this proof is to construct a
+  function of the following type:
+  \begin{equation}
+    \lVert x \in \mathit{xs} \rVert \rightarrow x \in \mathit{xs}
+  \end{equation}
+  Given decidable equality over the type of \(x\).
+
+  We do this using the previous recompute lemma: that tells us that all we need
+  to construct is a decision for \(x \in \mathit{xs}\), and it will be able to
+  derive the proof itself.
+  Such a decision procedure is not difficult to construct: for any value \(x\)
+  and list \(\mathit{xs}\), we proceed through the list \(\mathit{xs}\), testing
+  if \(x\) is equal to any of its contents.
+  If it is, we return that we have proven the goal, and that \(x\) is indeed
+  present in \(\mathit{xs}\).
+  Otherwise, we know that \(x\) cannot be in \(\mathit{xs}\) (since we've tested
+  every value), so we return that the goal has been disproven.
+\end{proof}
 \section{Kuratowski Finiteness}

### 2020-06-03 sections/finiteness-predicates.tex
kuratowski

@@
 \section{Kuratowski Finiteness}
-The first thing we must define is a representation of subsets.
-\begin{definition}[Kuratowski Finite Subset]
-  \(\mathcal{K}(A)\) is the type of Kuratowski-finite subsets of \(A\).
+The one big missing definition of finiteness to cover is \emph{Kuratowski}
+finiteness.
+While it's quite important, it's also quite different from the definitions we've
+seen so far.
+It starts with an encoding of the free join semilattice.
+\begin{definition}[Free Join Semilattice]
+  \(\mathcal{K}(A)\) is the free join semilattice, or, alternatively, the type
+  of Kuratowski-finite subsets of \(A\).
   \begin{inductivetype}{\mathcal{K}(A)}
@@
 \begin{lemma} \label{cardinal-kuratowski}
+  Cardinal finiteness is equivalent to Kuratowski finiteness over a discrete
+  set.
   \begin{equation}

### 2020-06-06 sections/topos.tex
countdown dection

@@
 \chapter{Topos} \label{topos}
-In this section we will prove that decidable Kuratowski finite types form a
+In this section we will examine the categorical interpretation of finite sets.
+In particular, we will prove that decidable Kuratowski finite types form a
 \(\Pi\)-pretopos.
-Along the way we will provide closure proofs for a number of the other
-finiteness predicates.
-As we saw in theorem~\ref{cardinal-kuratowski}, decidable Kuratowski finite
-types are equivalent to cardinal finite types, so we will work with the latter
-from now on.
-Our first task is to show that cardinal finite types are closed under several
-operations.
+A lot of the work for this proof has been done already: in
+Theorem~\ref{cardinal-kuratowski} we saw that Kuratowski finite types were
+equivalent to Cardinally finite types.
+We will use the latter definition implementation-wise from now on, as it is
+slightly easier to work with: CuTT's transport means we can do this without loss
+of generality.
+
+\section{Categories in HoTT}
+\section{Closure and the Category of Sets}
+\section{The Absence of the Subobject Classifier}
+
 \section{Closure}
@@
 closure proofs from it.
-
-% \begin{lemma}
-%   \(\bot\), \(\top\), and \(\mathbf{Bool}\) are split enumerable.
-% \end{lemma}
-% \begin{lemma} \label{split-enum-sigma}
-%   Split enumerability is closed under \(\Sigma\).
-%   \begin{equation}
-%     \frac{
-%       \mathcal{E}!(A) \; \; \; \Pi(x : A) , \mathcal{E}!(U(x))
-%     }{
-%       \mathcal{E}!(\Sigma(x : A) , U(x))
-%     }
-%   \end{equation}
-% \end{lemma}
-% From this we can derive split enumerability of non-dependent sums and products,
-% as both are definable in terms of \(\Sigma\).
 

### 2020-06-06 sections/countdown.tex

\chapter{Countdown}
Countdown is a well-known functional programming puzzle (with a spin-off TV show
in the UK), first popularised as a puzzle in which to demonstrate functional
algorithms in \cite{huttonCountdownProblem2002}.
The idea is simple: given a list of numbers, contestants must construct an
arithmetic expression (using a small set of functions) using some or all of the
numbers, to reach some target.

\section{Enumerating Binary Trees}
\section{Enumerating Combinations and Permutations}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../paper"
%%% End:

### 2020-06-06 sections/countdown.tex
countdown section

@@
 
+Take the following problem as an example:
+\begin{gather*}
+  \boxed{1} \boxed{3} \boxed{7} \boxed{10} \boxed{25} \boxed{50} \\
+  \boxed{765} \tag{Target}
+\end{gather*}
+It has the following answer:
+\begin{equation}
+  3 \times ((7 \times (50 - 10)) - 25)
+\end{equation}
+Importantly, we do not have to use every number given to get to the target.
+For our problem, we will use the operators \(\times\), \(+\), and \(-\).
+
+In this section we will develop a program/proof which can decide countdown
+problems totally.
+In other words, given a list of numbers and a target, our program will prove
+whether a solution exists, and if so, it will provide just that solution.
+\section{Classifying The Problem}
+So what is a ``solution'' to the countdown problem?
+Given some list of numbers, what type classifies the candidates for testing from
+those numbers?
+What is the type of the following transformation?
+\begin{alignat*}{6}
+  &1 \; \; \;         &&3         &&7    &&10    &&25  &&50 \\
+  &&&3 \times (&&7 \times (&&50 - &&10) - &&25)
+\end{alignat*}
+Firstly, we can notice that not every number was used in our solution.
+So the transformation must information about what to keep and what to discard on
+an item-by-item basis: a subsequence, in other words.
+% \begin{align*}
+%   &1 \; \; \;         &&3         &&7    &&10    &&25  &&50 \\
+%   &         &&3         &&7    &&10    &&25  &&50 \\
+%   &         &&3         &&7    &&50    &&10  &&25 \\
+%   &         &&3   \times      &&7 \times    &&50 -    &&10 -  &&25 \\
+%   &&&3 \times ((&&7 \times (&&50 - &&10)) - &&25)
+% \end{align*}
+
+  \begin{tikzpicture}
+
+    \node at (-1,  3) {1}  ;
+    \node at (0 ,  3) {3}  ;
+    \node at (1 ,  3) {7}  ;
+    \node at (2 ,  3) {10} ;
+    \node at (3 ,  3) {25} ;
+    \node at (4 ,  3) {50} ;
+
+    \node at (-1,  1.5) {\xcancel{1}}  ;
+    \node at (0 ,  1.5) {3}  ;
+    \node at (1 ,  1.5) {7}  ;
+    \node at (2 ,  1.5) {10} ;
+    \node at (3 ,  1.5) {25} ;
+    \node at (4 ,  1.5) {50} ;
+
+    \node(t3)   at (0 ,  0) {3}  ;
+    \node(t7)   at (1 ,  0) {7}  ;
+    \node(t10)  at (2 ,  0) {10} ;
+    \node(t25)  at (3 ,  0) {25} ;
+    \node(t50)  at (4 ,  0) {50} ;
+    \node(b3)   at (0 , -1.5) {3}  ;
+    \node(b7)   at (1 , -1.5) {7}  ;
+    \node(b50)  at (2 , -1.5) {50} ;
+    \node(b10)  at (3 , -1.5) {10} ;
+    \node(b25)  at (4 , -1.5) {25} ;
+
+    \draw [->, rounded corners] (t3.south) --  (b3.north) ;
+    \draw [->, rounded corners] (t7.south) --  (b7.north) ;
+    \draw [->, rounded corners] (t10.south) to[out=-90, in=90] (b10.north) ;
+    \draw [->, rounded corners] (t25.south) to[out=-90, in=90] (b25.north) ;
+    \draw [->, rounded corners] (t50.south) to[out=-90, in=90] (b50.north) ;
+
+    \node[draw] at (0.5 , -2.25) {$\times$} ;
+    \node[draw] at (1.5 , -2.25) {$\times$} ;
+    \node[draw] at (2.5 , -2.25) {$-$} ;
+    \node[draw] at (3.5 , -2.25) {$-$} ;
+
+    \node at (0.5 , -3) {$\times$} ;
+    \node at (1.5 , -3) {$\times$} ;
+    \node at (2.5 , -3) {$-$} ;
+    \node at (3.5 , -3) {$-$} ;
+
+    \node(e3) at (0   , -3) {3}  ;
+    \node(e7) at (1   , -3) {7}  ;
+    \node(e50) at (2   , -3) {50} ;
+    \node(e10) at (3   , -3) {10} ;
+    \node(e25) at (4   , -3) {25} ;
+
+    \draw[-] (e50.south) -- ++(0.5,-0.75 ) ;
+    \draw[-] (e10.south) -- ++(-1 ,-1.5  ) ;
+    \draw[-] (e7.south)  -- ++(1.5,-2.25) ;
+    \draw[-] (e3.south)  -- ++(2  , -3 ) ;
+    \draw[-] (e25.south) -- ++(-2 , -3 ) ;
+  \end{tikzpicture}
+
+\section{Enumerating Combinations and Permutations}
 \section{Enumerating Binary Trees}
-\section{Enumerating Combinations and Permutations}
+\section{Filtering Out Invalid Expressions with the Subobject Classifier}
+\section{Filtering Out Duplicates with Quotients}
 

### 2020-06-07 sections/countdown.tex
Update countdown.tex

@@
 So what is a ``solution'' to the countdown problem?
-Given some list of numbers, what type classifies the candidates for testing from
-those numbers?
+Put simply, it is a way to:
+\begin{enumerate}
+  \item Arrange some or all of the given numbers into a valid expression
+  \item Such that the expression, when evaluated, is equal to the target.
+\end{enumerate}
+
 What is the type of the following transformation?
-\begin{alignat*}{6}
-  &1 \; \; \;         &&3         &&7    &&10    &&25  &&50 \\
-  &&&3 \times (&&7 \times (&&50 - &&10) - &&25)
-\end{alignat*}
+
 Firstly, we can notice that not every number was used in our solution.
@@
 an item-by-item basis: a subsequence, in other words.
-% \begin{align*}
-%   &1 \; \; \;         &&3         &&7    &&10    &&25  &&50 \\
-%   &         &&3         &&7    &&10    &&25  &&50 \\
-%   &         &&3         &&7    &&50    &&10  &&25 \\
-%   &         &&3   \times      &&7 \times    &&50 -    &&10 -  &&25 \\
-%   &&&3 \times ((&&7 \times (&&50 - &&10)) - &&25)
-% \end{align*}
 
+\begin{figure}
+  \tikzset{>=Triangle[open]}
   \begin{tikzpicture}
+    \node (o1)  at (-1, 1.5) { 1} ;
+    \node (o3)  at ( 0, 1.5) { 3} ;
+    \node (o7)  at ( 1, 1.5) { 7} ;
+    \node (o10) at ( 2, 1.5) {10} ;
+    \node (o25) at ( 3, 1.5) {25} ;
+    \node (o50) at ( 4, 1.5) {50} ;
 
-    \node at (-1,  3) {1}  ;
-    \node at (0 ,  3) {3}  ;
-    \node at (1 ,  3) {7}  ;
-    \node at (2 ,  3) {10} ;
-    \node at (3 ,  3) {25} ;
-    \node at (4 ,  3) {50} ;
-
-    \node at (-1,  1.5) {\xcancel{1}}  ;
-    \node at (0 ,  1.5) {3}  ;
-    \node at (1 ,  1.5) {7}  ;
-    \node at (2 ,  1.5) {10} ;
-    \node at (3 ,  1.5) {25} ;
-    \node at (4 ,  1.5) {50} ;
-
+    \node(t1) at (-1,  0) {\xcancel{1}}  ;
     \node(t3)   at (0 ,  0) {3}  ;
@@
     \node(t50)  at (4 ,  0) {50} ;
+
+    \draw [-{Rays[n=4]}] (o1.south)  -- (t1.north) ;
+    \draw [->] (o3.south)  -- (t3.north) ;
+    \draw [->] (o7.south)  -- (t7.north) ;
+    \draw [->] (o10.south) -- (t10.north) ;
+    \draw [->] (o25.south) -- (t25.north) ;
+    \draw [->] (o50.south) -- (t50.north) ;
+
     \node(b3)   at (0 , -1.5) {3}  ;
@@
 
-    \draw[-] (e50.south) -- ++(0.5,-0.75 ) ;
+    \draw[-] (e50.south) -- ++(0.5,-0.75) ;
     \draw[-] (e10.south) -- ++(-1 ,-1.5  ) ;
@@
     \draw[-] (e25.south) -- ++(-2 , -3 ) ;
+
+    % \node(r40)  [circle, inner sep=0pt, fill=white, anchor=center] at (2.5 , -3.75) {40} ;
+    % \node(r280) [circle, inner sep=0pt, fill=white, anchor=center] at (2   , -4.5) {280} ;
+    % \node(r255) [circle, inner sep=0pt, fill=white, anchor=center] at (2.5   , -5.25) {255} ;
+
+    % \draw[->] (e50) -- (r40) ;
+    % \draw[->] (e10) -- (r40) ;
+
+    \node at (2, -6.5) {$765$} ;
+
+    \node at (6, 0.7) {\parbox{3cm}{\subcaption{Filter        \hfill \; \label{countdown-filter}}}} ;
+    \node at (6,-0.8) {\parbox{3cm}{\subcaption{Permutation   \hfill \; \label{countdown-permutation}}}} ;
+    \node at (6,-2.3) {\parbox{3cm}{\subcaption{Operators     \hfill \; \label{countdown-operators}}}} ;
+    \node at (6,-4.8) {\parbox{3cm}{\subcaption{Parenthesising\hfill \; \label{countdown-parens}}}} ;
   \end{tikzpicture}
+  \caption{The components of a transformation which makes up a Countdown
+    candidate solution}
+\end{figure}
 

### 2020-06-07 sections/countdown.tex
countdown figure

--- 2020-06-06 sections/countdown.tex
@@
 \chapter{Countdown}
+\input{figures/countdown-transformation}
 Countdown is a well-known functional programming puzzle (with a spin-off TV show
@@
 
+Take the following problem as an example:
+\begin{gather*}
+  \boxed{1} \boxed{3} \boxed{7} \boxed{10} \boxed{25} \boxed{50} \\
+  \boxed{765} \tag{Target}
+\end{gather*}
+It has the following answer:
+\begin{equation}
+  3 \times ((7 \times (50 - 10)) - 25)
+\end{equation}
+Importantly, we do not have to use every number given to get to the target.
+For our problem, we will use the operators \(\times\), \(+\), and \(-\).
+
+In this section we will develop a program/proof which can decide countdown
+problems totally.
+In other words, given a list of numbers and a target, our program will prove
+whether a solution exists, and if so, it will provide just that solution.
+\section{Classifying The Problem}
+So what is a ``solution'' to the countdown problem?
+Put simply, it is a way to:
+\begin{enumerate}
+  \item Arrange some or all of the given numbers into a valid expression
+  \item Such that the expression, when evaluated, is equal to the target.
+\end{enumerate}
+Our task is to construct a \emph{type} for this transformation.
+We will mainly focus on the first of these two concerns: arranging a list of
+numbers into an expression using the given rules of Countdown.
+
+We can break this transformation into parts, displayed in
+figure~\ref{countdown-transform}.
+\section{Enumerating Combinations and Permutations}
 \section{Enumerating Binary Trees}
-\section{Enumerating Combinations and Permutations}
+\section{Filtering Out Invalid Expressions with the Subobject Classifier}
+\section{Filtering Out Duplicates with Quotients}
 

### 2020-06-07 figures/pairing-functions.tex
figures

@@
-\begin{figure}
+\begin{figure*}
   \centering
-  \begin{subfigure}[b]{.5\linewidth}
-    \begin{tikzcd}[sep=tiny,font=\footnotesize]
+  \tikzcdset{
+  cramped,
+  every matrix/.append style={nodes={font=\scriptsize}},
+  row    sep/normal=1em,
+  column sep/normal=1em,
+  }
+  \begin{subfigure}[b]{.45\textwidth}
+    \centering
+    \begin{tikzcd}
       (1,e) \ar[rdddd, out=-45, in=135] & (2,e) \ar[rdddd, out=-45, in=135] & (3,e) \ar[rdddd, out=-45, in=135] & (4,e) \ar[rdddd, out=-45, in=135] & (5,e)        \\
@@
     \end{tikzcd}
-    \caption{Cartesian}
-    \label{cartesian}
-  \end{subfigure}%
-  \begin{subfigure}[b]{.5\linewidth}
-    \begin{tikzcd}[sep=tiny,font=\footnotesize]
+    \caption{Depth-First}
+    \label{depth-first}
+  \end{subfigure} \hfill
+  \begin{subfigure}[b]{.45\textwidth}
+    \centering
+    \begin{tikzcd}
       (1,e) \ar[dr] & (2,e) \ar[dr]  & (3,e) \ar[dr]    & (4,e) \ar[dr] & (5,e) \\
@@
     \end{tikzcd}
-    \caption{Cantor}
-    \label{cantor}
+    \caption{Breadth-First}
+    \label{breadth-first}
   \end{subfigure}
@@
   \label{pairings}
-\end{figure}
+\end{figure*}

### 2020-06-07 sections/countable-predicates.tex
figures

@@
 \chapter{Countably Infinite Types} \label{infinite}
-\begin{figure*}
-  \centering
-
-  \tikzcdset{
-  cramped,
-  every matrix/.append style={nodes={font=\scriptsize}},
-  row    sep/normal=1em,
-  column sep/normal=1em,
-  }
-  \begin{subfigure}[b]{.45\textwidth}
-    \centering
-    \begin{tikzcd}
-      (1,e) \ar[rdddd, out=-45, in=135] & (2,e) \ar[rdddd, out=-45, in=135] & (3,e) \ar[rdddd, out=-45, in=135] & (4,e) \ar[rdddd, out=-45, in=135] & (5,e)        \\
-      (1,d) \ar[u]     & (2,d) \ar[u]    & (3,d) \ar[u]    & (4,d) \ar[u]    & (5,d) \ar[u] \\
-      (1,c) \ar[u]     & (2,c) \ar[u]    & (3,c) \ar[u]    & (4,c) \ar[u]    & (5,c) \ar[u] \\
-      (1,b) \ar[u]     & (2,b) \ar[u]    & (3,b) \ar[u]    & (4,b) \ar[u]    & (5,b) \ar[u] \\
-      (1,a) \ar[u]     & (2,a) \ar[u]    & (3,a) \ar[u]    & (4,a) \ar[u]    & (5,a) \ar[u]
-    \end{tikzcd}
-    \caption{Depth-First}
-    \label{depth-first}
-  \end{subfigure} \hfill
-  \begin{subfigure}[b]{.45\textwidth}
-    \centering
-    \begin{tikzcd}
-      (1,e) \ar[dr] & (2,e) \ar[dr]  & (3,e) \ar[dr]    & (4,e) \ar[dr] & (5,e) \\
-      (1,d) \ar[dr] & (2,d) \ar[dr]  & (3,d) \ar[dr]    & (4,d) \ar[dr] & (5,d) \ar[u] \\
-      (1,c) \ar[dr] & (2,c) \ar[dr]  & (3,c) \ar[dr]    & (4,c) \ar[dr] & (5,c) \ar[uul] \\
-      (1,b) \ar[dr] & (2,b) \ar[dr]  & (3,b) \ar[dr]    & (4,b) \ar[dr] & (5,b) \ar[uuull, out=130, in=-50] \\
-      (1,a) \ar[u]  & (2,a) \ar[uul, out=130, in=-50] & (3,a) \ar[uuull, out=130, in=-50] & (4,a) \ar[uuuulll, out=130, in=-50] & (5,a) \ar[uuuulll, out=130, in=-50]
-    \end{tikzcd}
-    \caption{Breadth-First}
-    \label{breadth-first}
-  \end{subfigure}
-  \caption{Two possible products for the sets \(\left[ 1 \dots 5 \right]\) and
-    \(\left[  a \dots e \right]\)}
-  \label{pairings}
-\end{figure*}
+\input{figures/pairing-functions}
 In the previous sections we saw different flavours of finiteness which were

### 2020-06-07 sections/countdown.tex
figures

@@
 \chapter{Countdown}
-\input{figures/countdown-transformation}
 Countdown is a well-known functional programming puzzle (with a spin-off TV show
@@
 \end{enumerate}
-Our task is to construct a \emph{type} for this transformation.
-We will mainly focus on the first of these two concerns: arranging a list of
-numbers into an expression using the given rules of Countdown.
+The first part of this solution is a transformation (from a list of numbers into
+an expression), and the second is a filter (into valid expressions which
+evaluate to the target).
 
-We can break this transformation into parts, displayed in
-figure~\ref{countdown-transform}.
-\section{Enumerating Combinations and Permutations}
-\section{Enumerating Binary Trees}
-\section{Filtering Out Invalid Expressions with the Subobject Classifier}
-\section{Filtering Out Duplicates with Quotients}
+In order to find and prove countdown solutions, we will prove that the above
+transformation is \emph{finite}: this will allow us to search exhaustively
+through the available transformations, proving the existence or absence of a
+transformation which is a solution.
+
+\input{figures/countdown-transformation}
+
+Let's first look at the transformation part.
+As shown in figure~\ref{countdown-transform}, we can break the transformation
+into four parts.
+First, we notice that we don't have to use all of the given numbers in the
+solution candidate: keeping or discarding input numbers, then, is the first step
+of the transformation (fig.~\ref{countdown-filter}).
+Next, we are allowed to rearrange the order of the numbers in our output: the
+next step of the transformation, then, applies a permutation to the selected
+numbers (fig.~\ref{countdown-permutation}).
+Thirdly, we can choose the operators we're going to use: note that for \(n\)
+chosen numbers, we can choose \(n - 1\) operators
+(fig.~\ref{countdown-operators}).
+Finally, we have to apply the binary operators with some parenthesising order
+(fig.~\ref{countdown-parens}).
+
+There are finitely many of each of these transformations: finitely many filters
+for \(n\) objects, for instance, or finitely many tuples of \(n\) operators.
+We are going to construct a type for each of these transformations, prove it
+finite, and use that type to search.
+The final, completed transformation will then be a dependent product of each
+constituent step.
+
+% \section{Filtering and Operators}
+% \section{Enumerating Binary Trees}
+% \section{Filtering Out Invalid Expressions with the Subobject Classifier}
+% \section{Filtering Out Duplicates with Quotients}
 

### 2020-06-07 figures/finite-classification.tex
diagrams etc

@@
   \centering
-  \begin{tikzcd}[cramped, row sep=small]
+  \begin{tikzcd}[cramped, row sep=small, column sep=2.3em]
     {} &
@@
     \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \text{Ordered} &
     &
-    &
-    \text{Manifest Enumerable}
+    \text{\parbox{1.8cm}{\centering Manifest Enumerable}}
     \ar[
@@
       start anchor=south west,
-      end anchor=south east
+      end anchor={[yshift=-2ex]east}
       ]
@@
       "\text{Discrete}" description,
-      start anchor=north east,
+      start anchor={[yshift=2ex]east},
       end anchor=north west
@@
     \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \text{Ordered} &
+    &
     &
@@
 \end{figure*}%
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: "../paper"
+%%% End:

### 2020-06-07 sections/countdown.tex
diagrams etc

@@
 
-Take the following problem as an example:
+Here's an example puzzle:
 \begin{gather*}
@@
 \end{gather*}
-It has the following answer:
-\begin{equation}
-  3 \times ((7 \times (50 - 10)) - 25)
-\end{equation}
-Importantly, we do not have to use every number given to get to the target.
-For our problem, we will use the operators \(\times\), \(+\), and \(-\).
+We'll allow the use of \(\times\), \(+\), and \(-\).
+The answer is at the bottom of this page\footnotemark.
+
+\footnotetext{\rotatebox[origin=c]{180}{Answer: \(3 \times ((7 \times (50 - 10)) - 25)\)}}
 
@@
 
-In order to find and prove countdown solutions, we will prove that the above
-transformation is \emph{finite}: this will allow us to search exhaustively
-through the available transformations, proving the existence or absence of a
-transformation which is a solution.
+We are going to define this transformation/filter pair as a type.
+We will then show that this type is itself finite, and from there we will be
+able to enumerate all possible solutions, derive the cardinality of the
+solutions, and so on.
 
@@
 
-Let's first look at the transformation part.
-As shown in figure~\ref{countdown-transform}, we can break the transformation
-into four parts.
-First, we notice that we don't have to use all of the given numbers in the
-solution candidate: keeping or discarding input numbers, then, is the first step
-of the transformation (fig.~\ref{countdown-filter}).
-Next, we are allowed to rearrange the order of the numbers in our output: the
-next step of the transformation, then, applies a permutation to the selected
-numbers (fig.~\ref{countdown-permutation}).
-Thirdly, we can choose the operators we're going to use: note that for \(n\)
-chosen numbers, we can choose \(n - 1\) operators
+Let's first look at the transformation aspect of the solution.
+We have broken this transformation into four steps in
+figure~\ref{countdown-transform}, each one feeding into the next.
+
+First, we select which numbers in the input we are going to use in our solution
+candidate: this is the filtering step of the transformation
+(fig.~\ref{countdown-filter}).
+Next we decide the order of the selected numbers in the output: this is the
+permutation step of the transformation (fig.~\ref{countdown-permutation}).
+We then select \(n-1\) operators to use with our \(n\) chosen numbers
 (fig.~\ref{countdown-operators}).
@@
 
-There are finitely many of each of these transformations: finitely many filters
-for \(n\) objects, for instance, or finitely many tuples of \(n\) operators.
-We are going to construct a type for each of these transformations, prove it
-finite, and use that type to search.
-The final, completed transformation will then be a dependent product of each
-constituent step.
+Each of these transformations can be represented by a finite object: a filter
+for \(n\) items, for instance, can be represented as an \(n\)-tuple of bools.
+The full transformation itself will be a product of each of these intermediate
+steps, which itself is finite by closure under \(\Sigma\).
 
-% \section{Filtering and Operators}
-% \section{Enumerating Binary Trees}
+In this section we will use split-enumerability as our finiteness predicate.
+Most of what we examine would work equally well with the other predicates, but
+split enumerability is the most powerful and the simplest to define.
+Since we will work only with set types, there won't be any need to look at
+non-set predicates like manifest enumerability.
+\section{Filtering and Operators}
+The first component of the transformation is relatively simple: a filter.
+We can represent this transformation as an \(n\)-tuple of bools.
+The filter in fig.~\ref{countdown-filter}, for instance, could be represented
+with the following:
+\begin{equation}
+  \langle 0 , 1 , 1 , 1 , 1 , 1  \rangle
+\end{equation}
+The proof that this type is finite is straightforward: by induction on \(n\), we
+see that any \(n\)-tuple of finite things is itself finite.
+
+At this point we can look forward slightly to an upcoming transformation and
+prove it finite: the selection of \(n\) operators is \emph{also} constructed
+using an \(n\)-tuple.
+Clearly there are finitely many operators available to us, so an \(n\)-tuple of
+operators is finite for the same reason as the filter.
+\section{Permutations}
+\section{Enumerating Binary Trees}
 % \section{Filtering Out Invalid Expressions with the Subobject Classifier}

### 2020-06-10 sections/countdown.tex
switched to acmart

@@
 
-\footnotetext{\rotatebox[origin=c]{180}{Answer: \(3 \times ((7 \times (50 - 10)) - 25)\)}}
+\footnotetext{\rotatebox[origin=c]{180}{Answer: \(3 \times (7 \times (50 - 10) - 25)\)}}
 
@@
 \begin{enumerate}
-  \item Arrange some or all of the given numbers into a valid expression
-  \item Such that the expression, when evaluated, is equal to the target.
+  \item Arrange some or all of the given numbers into an expression
+  \item Such that the expression evaluates to our target.
 \end{enumerate}
@@
 We have broken this transformation into four steps in
-figure~\ref{countdown-transform}, each one feeding into the next.
+figure~\ref{countdown-transform}.
 
 First, we select which numbers in the input we are going to use in our solution
-candidate: this is the filtering step of the transformation
-(fig.~\ref{countdown-filter}).
+candidate: this is the selection step of the transformation
+(fig.~\ref{countdown-selection}).
 Next we decide the order of the selected numbers in the output: this is the
@@
 
-Each of these transformations can be represented by a finite object: a filter
-for \(n\) items, for instance, can be represented as an \(n\)-tuple of bools.
+Each of these transformations can be represented by a finite object: a selection
+from \(n\), for instance, can be represented as an \(n\)-tuple of bools.
 The full transformation itself will be a product of each of these intermediate
@@
 non-set predicates like manifest enumerability.
-\section{Filtering and Operators}
-The first component of the transformation is relatively simple: a filter.
+\section{Selection and Operators}
+The first component of the transformation is relatively simple: we select some
+of the given numbers to use in our solution.
 We can represent this transformation as an \(n\)-tuple of bools.
-The filter in fig.~\ref{countdown-filter}, for instance, could be represented
+The filter in fig.~\ref{countdown-selection}, for instance, could be represented
 with the following:
@@
 Clearly there are finitely many operators available to us, so an \(n\)-tuple of
-operators is finite for the same reason as the filter.
+operators is finite for the same reason as the selection.
 \section{Permutations}
+Permutations are slightly trickier, and have a number of possible
+representations.
+One option, for instance, would be to represent a permutation of size \(n\) as a
+bijection between \(\mathbf{Fin}(n)\) and \(\mathbf{Fin}(n)\).
+
 \section{Enumerating Binary Trees}
+
 % \section{Filtering Out Invalid Expressions with the Subobject Classifier}

### 2020-06-11 sections/finiteness-predicates.tex
linked to sections in diagram

@@
 tell us the number of elements in the finite set?).
-\section{Split Enumerability}
+\section{Split Enumerability} \label{split-enumerability}
 We will start with a simple notion of finiteness, called \emph{split}
@@
 
-\section{Manifest Bishop Finiteness}
+\section{Manifest Bishop Finiteness} \label{manifest-bishop-finiteness}
 We mentioned in the introduction that occasionally in constructive mathematics
@@
 \end{proof}
-\section{Cardinal Finiteness}
+\section{Cardinal Finiteness} \label{cardinal-finiteness}
 While we have removed some of the unnecessary information from our finiteness
@@
 This gets us all of the necessary closure proofs on \(\mathcal{C}\).
-\section{Manifest Enumerability}
+\section{Manifest Enumerability} \label{manifest-enumerability}
 We have now explored quite far in the ``less informative'' direction.
@@
 \end{proof}
-\section{Kuratowski Finiteness}
+\section{Kuratowski Finiteness} \label{kuratowski}
 The one big missing definition of finiteness to cover is \emph{Kuratowski}

### 2020-06-11 sections/introduction.tex
now compiles without errors
cleaner format
linked to sections in diagram

@@
 \section{Foreword}
-Proofs in constructive mathematics are often more substantial than their
-classical counterparts.
-``Substantial'' here doesn't mean the literal length of the proofs (although the
-double meaning is not lost on anyone who has had to struggle with a modern
-constructive proof assistant), but rather refers to their \emph{content}: a
-constructive proof contains more \emph{stuff} than a classical one.
-
-If we were to prove that a set was finite in classical mathematics, we might
-show that it is in bijection with another finite set, say a finite prefix of the
-natural numbers.
-In semi-formal notation, to say that a set \(A\) is finite we would show the
-following:
-\begin{equation} \label{example-finite}
-  \mathbf{Finite}(A) \coloneqq \exists (n : \mathbb{N}) . \; A \Leftrightarrow \left\{ m \mid m < n\right\}
-\end{equation}
-Where \(A~\Leftrightarrow~B\) means ``a bijection between \(A\) and \(B\)''.
-
-Classically, a proof of equation~\ref{example-finite} is nothing more than
-\emph{evidence} for the proposition of equation~\ref{example-finite}.
-A constructive proof of \ref{example-finite}, however, is a mathematical object
-in its own right, containing:
-\begin{enumerate}
-  \item A number \(n\) representing the cardinality of \(A\).
-  \item A function from \(A\) to \(\left\{ m \mid m < n \right\}\).
-  \item A function from \(\left\{ m \mid m < n \right\}\) to \(A\).
-\end{enumerate}
-In other words, it makes sense (in constructive mathematics) to talk about a
-function of type \(\mathbf{Finite}(A)~\rightarrow~\mathbb{N}\), which takes a
-\emph{proof} of the finiteness of \(A\) and returns \(A\)'s cardinality.
-
-Such functions are perhaps even more concrete than this description suggests:
-many constructive proofs are in fact computer programs in disguise (and
-vice-versa \cite{wadlerPropositionsTypes2015}).
-All ``proofs'' in this paper correspond to honest-to-goodness proofs in a
-programming language (Cubical Agda \cite{vezzosiCubicalAgdaDependently2019} in
-our case) which can be run on a computer to produce real output.
-
-We have now hopefully established, albeit informally, what it means for
-constructive proofs to be ``substantial''.
-Clearly this extra information carried by proofs can add richness to their
-study: we will see, for instance, no fewer than five finiteness predicates in
-this paper which would collapse to one in a classical setting.
-The extra information also presents a challenge, however: occasionally we will
-see that we have revealed \emph{too much} information while proving some
-property.
-
-Consider the category of finite sets.
-We will explore this category in much more detail later, but for now we just
-need to know that the objects of this category are---shockingly---finite sets.
-Constructively, the objects of this category can be thought of as \emph{pairs},
-where the first element is a set (like \(\left\{ \text{True}, \text{False}
-\right\}\), the two-element set of the booleans) and the second is a proof that
-that set is finite.
-If we take those proofs to be the kind given in equation~\ref{example-finite},
-we should see that they contain a bijection from the booleans to the set
-\(\left\{ 0, 1 \right\}\): however \emph{two} such bijections exist!
-\begin{figure}[h]
-  \centering
-  \begin{subfigure}[b]{.45\textwidth}
-    \centering
-    \begin{tikzcd}
-      \text{True}  \ar[rr, leftrightarrow] & & 0 \\
-      \text{False} \ar[rr, leftrightarrow] & & 1
-    \end{tikzcd}
-  \end{subfigure}
-  \begin{subfigure}[b]{.45\textwidth}
-    \centering
-    \begin{tikzcd}
-      \text{True}  \ar[rrd, leftrightarrow, start anchor=south east, end anchor=north west] & & 0 \\
-      \text{False} \ar[rru, leftrightarrow, start anchor=north east, end anchor=south west, crossing over] & & 1
-    \end{tikzcd}
-  \end{subfigure}
-  \caption{The Two Bijections Between the Booleans and \(\left\{ 0, 1 \right\}\)}
-\end{figure}
-This means that where we wanted one object to correspond to the finite set
-\(\left\{ \text{True}, \text{False} \right\}\), we instead have two.
-In fact, for any finite set of cardinality \(n\) we will have \(n!\) finiteness
-proofs, and therefore \(n!\) objects.
-We have constructed, in other words, entirely the wrong category.
-
-Being able to hide the extra information in a constructive proof when we need to
-is clearly important.
-HoTT (and CuTT by extension) facilitate this hiding: how exactly it is
-accomplished will be explored in the subsequent sections.
-
-Finally, for all of its benefits, constructive mathematics (especially
-dependently-typed constructive mathematics on a computer) still suffers from
-many painful drawbacks.
-As already hinted at, proofs in a constructive setting can tend to be verbose,
-especially when the arbiter of correctness is as strict as a computer, which
-neither understands nor trusts the time-honoured technique of ``the other cases
-are obviously true''.
-While the goal of using a computer to automate the boring parts of a proof still
-seems attainable, too often getting a computer involved in a proof means adding
-thousands of lines of tedium to satisfy its lack of intuition.
-
-Simple proof search over a finite domain is a simple, if effective, way to
-relieve some of the tedium.
-Many libraries in Agda, Coq, and other languages make use of this fact.
-While the idea for finite proof search is simple, we found that implementing one
-revealed a number of interesting theoretical considerations:
-\begin{description}
-  \item[How should we construct search spaces?]
-    We found that a \(Pi\)-pretopos provided the appropriate ``generator
-    language'' for constructing search spaces for finite types.
-  \item[How do we deal with functions?]
-    We of course would like to be able to search over functions: \todo{?}
-  \item[How do we talk about ``searchability''?]
-    \todo{Omniscience}
-\end{description}
-Finiteness is a prime example of a phenomenon which happens quite frequently in
-dependently-typed constructive mathematics: through building a practical program
-we unearth theoretically interesting results.
-\todo{rephrase}
+\todo[inline]{Foreword}
 \section{Contributions}
-We are interested in constructive notions of finiteness.
+We are interested in constructive notions of finiteness, formalised in Cubical
+Type Theory.
 In contrast to classical finiteness, in a constructive setting there is a wide
@@
 implications.
+
+\input{figures/finite-classification}
 
@@
 with capabilities not seen in comparable proof search libraries.
-
-\input{figures/finite-classification}
 
@@
 between each predicate.
-In section~\ref{relations} we will provide a function which inhabits each arrow.
+We will provide a function which inhabits each arrow.
 Each unlabelled arrow is an unconditional implication: every manifest Bishop

### 2020-06-12 sections/introduction.tex
more introduction

@@
 We are interested in constructive notions of finiteness, formalised in Cubical
-Type Theory.
+Type Theory \cite{cohenCubicalTypeTheory2016}.
+In this paper we will explore five such notions of finiteness, including their
+categorical interpretation, and use them to build a simple proof-search library
+facilitated in a fundamental way by univalence.
+Along the way we will use the Countdown problem
+\cite{huttonCountdownProblem2002} as an example, and provide a program which
+produces verified solutions to the puzzle.
+We will also briefly examine countability, and demonstrate its parallels and
+differences with finiteness.
+\subsection{The Varieties of Finiteness}
+In Section~\ref{finiteness-predicates} we will explore a number of different
+predicates for finiteness.
 In contrast to classical finiteness, in a constructive setting there is a wide
 variety of predicates which all have some claim to being the formal
-interpretation of ``finiteness''.
-Broadly, they vary along two axes: informativeness (i.e. how many properties
-of the underlying type we can derive from a proof that it is finite), and
-restrictiveness (i.e. how many types conform to its particular variety of
-finiteness).
-We will explore five of these predicates, proving formally their relations and
-implications.
+interpretation of ``finiteness'' \cite{coquandConstructivelyFinite2010}.
+The particular predicates we are interested in are organised in
+Figure~\ref{finite-classification}: each arrow in the diagram represents a proof
+that one predicate can be derived from another.
 
@@
 
-We will then focus on one of the predicates: decidable Kuratowski finiteness.
-We will construct a category of decidable Kuratowski-finite types, and show that
-it forms a \(\Pi\)-pretopos.
+These finiteness predicates differ along two main axes: informativeness, and
+restrictiveness.
+More ``informative'' predicates have proofs which contain extraneous information
+other than the finiteness of the underlying type: a proof of split enumerability
+(Section~\ref{split-enumerability}), for instance, comes with a strict total
+order on the underlying type.
+We will prove that a more informative finiteness predicate can be
+derived from a less informative one by providing the missing information
+(Theorem~\ref{cardinal-to-manifest-bishop}).
 
-We will then take a brief detour into predicates for countably infinite types,
-and compare its landscape to that of the finite predicates.
+The ``restrictiveness'' of a predicate refers to how many types it admits into
+its notion of ``finite''.
+There are strictly more Kuratowski finite (Section~\ref{kuratowski}) types than
+there are Cardinally finite (Section~\ref{cardinal-finiteness}).
+We will prove that we can always derive the less restrictive predicate from the
+more restrictive one, and that we can go in the other direction by satisfying
+the missing requirement (decidable equality in all of these cases).
 
-Finally, we will present a case-study in a practical application of the
-predicates for finiteness we have constructed: a library for total proof search.
-We will show how our earlier work gives a good theoretical foundation for this
-library, and allows for a simple and principled API, and we will demonstrate how
-some of the more complex theoretical results allow for a more powerful library
-with capabilities not seen in comparable proof search libraries.
+Proofs coming with extra information is a common theme in constructive
+mathematics: often this extra information is in the form of an algorithm which
+can do something useful related to the proof itself.
+Indeed, our proofs of finiteness here will provide an algorithm to solve the
+countdown puzzle.
+Occasionally, however, the extra information is undesirable: we may want to
+assert the existence of some value \(x : A\) which satisfies a predicate \(P\)
+without revealing \emph{which} \(A\) we're referring to.
+More concretely, we will need in this paper to prove that two types are in
+bijection without specifying a particular bijection.
+This facility is provided by Homotopy Type Theory \cite{hottbook} in the form of
+propositional truncation, and it is what allows us to prove the bulk of
+propositions in this paper.
 
-\subsection{Finiteness Predicates}
-Our first contribution is a cataloguing and characterisation of five finiteness
-predicates, shown in figure~\ref{finite-classification}.
+For each predicate we will also prove its closure properties (i.e. that the
+product of two finite sets is finite).
+The most significant of these closure proofs is that of closure under \(\Pi\)
+(dependent functions) (Theorem~\ref{split-enum-pi}).
+\subsection{Toposes and Finite Sets}
+In Section~\ref{topos}, we will explore the categorical interpretation of
+decidable Kuratowski finite sets.
+The motivation here is partially a practical one: by the end of this work we
+will have provided a library for proof search over finite types, and the
+``language'' of a topos is a reasonable choice for a principled language for
+constructing proofs of finiteness in the style of QuickCheck
+\cite{claessenQuickCheckLightweightTool2011} generators.
 
-While our focus is decidable Kuratowski finiteness, in order to prove the
-closure properties that we are interested in we will need to work with the
-intermediate predicates as well (though they are also interesting in their own
-right).
-
-Since we have a proof-search library in mind, we will have to consider how to
-present a good API for constructing proofs of finiteness.
-In this vein we will show that some simple base types are finite
-(\(\mathbf{Bool}\), \(\top\), \(\bot\), etc), and we will provide functions to
-combine finiteness proofs together into more complex types (i.e. \(\times\),
-\(\uplus\), \(\Sigma\), \(\Pi\), etc).
-As we will see this choice of toolbox for building finiteness proofs actually
-has a deep categorical justification, but at this point it makes sense to look
-at it as just a suitably general and powerful API.
-
-Our focus in this work is decidable Kuratowski finiteness.
-We will characterise it, its strength, and its uses, and in particular we will
-show that it forms a \(\Pi\)-pretopos.
-Our proofs follow the structure of \cite[Chapters 9, 10]{hottbook} and
+Theoretically speaking, showing that sets in Homotopy Type Theory form a topos
+(with some caveats) is an important step in characterising the categorical
+implications of Homotopy Type Theory, first proven in
 \cite{rijkeSetsHomotopyType2015}.
-We will explore this concept in more detail later, but crucially we will need to
-provide proofs of closure under \(\times\), \(\rightarrow\), etc.
-
-In order to prove these closure properties, we will define and use some
-intermediate finiteness predicates, classified in
-figure~\ref{finite-classification}.
-These predicates are also interesting in their own right, though!
-
-The arrows in figure~\ref{finite-classification} are implication functions
-between each predicate.
-We will provide a function which inhabits each arrow.
-Each unlabelled arrow is an unconditional implication: every manifest Bishop
-finite set is cardinal finite (lemma~\ref{manifest-bishop-to-cardinal}), for
-instance.
-The labelled arrows are strengthening proofs: every manifest enumerable set
-\emph{with decidable equality} is split enumerable
-(lemma~\ref{manifest-enum-to-split-enum}).
-Our most significant result here is the proof that a cardinal finite set with a
-decidable total order is manifestly Bishop finite.
-
+Our work is a formalisation of this result (and the first such formalisation
+that we are aware of).
+The proof that decidable Kuratowski finite sets form a \(\Pi\)-pretopos is
+additional to that.
 \subsection{Countability Predicates}
@@
 types, and classify them in a parallel way to the finite predicates
-(section~\ref{infinite}).
+(Section~\ref{infinite}).
+\todo{Expand}
 \subsection{Search}
 All of our work is formalised in Cubical Agda
-\cite{vezzosiCubicalAgdaDependently2019}.
-We will make mention of the few occasions where the formalisation of some proof
-is of interest, but the main place where we will discuss the code is in
-section~\ref{search}, where we implement a library for proof search, based on
-omniscience and exhaustibility.
-This library relies directly on the computational content of univalence, which
-allows us to (for instance) have \(\Pi\) types in the domain of the search
-space.
-The language of a \(\Pi\)-pretopos turns out to be exactly the right language
-for constructing ``generator''-like expressions.
-
-The full code of the formalisation is available at
-\url{https://github.com/oisdk/finiteness-in-cubical-type-theory}.
+\cite{vezzosiCubicalAgdaDependently2019}: as a result, the constructive
+interpretation of each proof is actually a program which can be run on a
+computer.
+In finiteness in particular, these programs are particularly useful for
+exhaustive search.
+\todo{Expand}
 \section{Notation and Background}

### 2020-06-12 paper.tex
rearranged more
cleaner format
countdown
switched to acmart
more introduction
Update paper.tex

@@
 
-\author{
-  Donnacha Oisín Kidney \and
-  Gregory Provan \and
-  Nicolas Wu
-}
-% \affiliation{University College Cork}
-% \email{o.kidney@cs.ucc.ie}
-% \author{Gregory Provan}
-% \affiliation{University College Cork}
-% \email{g.provan@cs.ucc.ie}
-% \author{Nicolas Wu}
-% \affiliation{Imperial College London}
-% \email{n.wu@imperial.ac.uk}
+\begin{document}
 
@@
 
-% \keywords{Agda, Homotopy Type Theory, Cubical Type Theory,
-%   Dependent Types, Finiteness, Topos, Kuratowski finite}
+\author{Donnacha Oisín Kidney}
+\author{Gregory Provan}
+\author{Nicolas Wu}
 
-\begin{document}
+\begin{abstract}
+  We study five different notions of finiteness in Cubical Type Theory and prove
+  the relationship between them.
+  In particular we show that any totally ordered Kuratowski finite type is
+  manifestly Bishop finite.
+
+  We also prove closure properties for each finite type, and classify them
+  topos-theoretically.
+  This includes a proof that the category of decidable Kuratowski finite sets
+  (also called the category of cardinal finite sets) form a \(\Pi\)-pretopos.
+
+  We then develop a parallel classification for the countably infinite types, as
+  well as a proof of the countability of \(A^\star\) for a countable type \(A\).
+
+  We formalise our work in Cubical Agda, where we implement a library for proof
+  search (including combinators for level-polymorphic fully generic currying).
+  Through this library we demonstrate a number of uses for the computational
+  content of the univalence axiom, including searching for and synthesising
+  functions.
+\end{abstract}
+
+\keywords{Agda, Homotopy Type Theory, Cubical Type Theory,
+  Dependent Types, Finiteness, Topos, Kuratowski finite}
 
 \maketitle
-% \begin{abstract}
-%   We study five different notions of finiteness in Cubical Type Theory and prove
-%   the relationship between them.
-%   In particular we show that any totally ordered Kuratowski finite type is
-%   manifestly Bishop finite.
 
-%   We also prove closure properties for each finite type, and classify them
-%   topos-theoretically.
-%   This includes a proof that the category of decidable Kuratowski finite sets
-%   (also called the category of cardinal finite sets) form a \(\Pi\)-pretopos.
-
-%   We then develop a parallel classification for the countably infinite types, as
-%   well as a proof of the countability of \(A^\star\) for a countable type \(A\).
-
-%   We formalise our work in Cubical Agda, where we implement a library for proof
-%   search (including combinators for level-polymorphic fully generic currying).
-%   Through this library we demonstrate a number of uses for the computational
-%   content of the univalence axiom, including searching for and synthesising
-%   functions.
-% \end{abstract}
 \include{sections/introduction}
@@
 \include{sections/search}
+\include{sections/countdown}
 \include{sections/related-work}
@@
 % \end{acks}
-
-\bibliographystyle{plain}
 \bibliography{bibliography}

### 2020-06-12 sections/introduction.tex
skeleton of related work

@@
 (Section~\ref{infinite}).
-\todo{Expand}
+We will see that we lose closure under function arrows, but we gain it under the
+Kleene star (Theorem~\ref{split-countability-sigma}).
 \subsection{Search}
@@
 exhaustive search.
-\todo{Expand}
+
+We will use the countdown problem as a running example throughout the paper: we
+will show how to prove that any given puzzle has a finite number of solutions,
+and from that we will show how to enumerate those solutions, thereby solving the
+puzzle in a verified way.
+
+In Section~\ref{search} we will package up the ``search'' aspect of finiteness
+into a library for proof search: similar libraries have been built in
+\cite{fruminFiniteSetsHomotopy2018} and
+\cite{firsovDependentlyTypedProgramming2015}.
+Our library differs from those in two important ways: firstly, it is strictly
+more powerful, 
 \section{Notation and Background}

### 2020-06-12 sections/introduction.tex
finished contributions

@@
 \cite{firsovDependentlyTypedProgramming2015}.
-Our library differs from those in two important ways: firstly, it is strictly
-more powerful, 
+Our library differs from those in three important ways: firstly, it is strictly
+more powerful, as it allows for search over function types.
+Secondly, finiteness proofs also provide equivalence proofs to any other finite
+type: this allows transport of proofs between types of the same cardinality.
+Finally, through generic programming we provide a simple syntax for stating
+properties which mimics that of QuickCheck.
+We also ground the library in the theoretical notions of omniscience.
 \section{Notation and Background}

### 2020-06-12 sections/introduction.tex
countdown moved

@@
 We also ground the library in the theoretical notions of omniscience.
+\section{Countdown}
+Countdown is a well-known functional programming puzzle (later turned into a TV
+show in the UK), first popularised as a puzzle in which to demonstrate
+functional algorithms in \cite{huttonCountdownProblem2002}.
+As a running example in this paper, we will produce a verified program which
+lists all solutions to a given countdown puzzle: here we will briefly explain
+the game and our strategy for solving it.
+
+The idea behind countdown is simple: given a list of numbers, contestants must
+construct an arithmetic expression (using a small set of functions) using some
+or all of the numbers, to reach some target.
+Here's an example puzzle:
+\begin{gather*}
+  \boxed{1} \boxed{3} \boxed{7} \boxed{10} \boxed{25} \boxed{50} \\
+  \boxed{765} \tag{Target}
+\end{gather*}
+We'll allow the use of \(\times\), \(+\), and \(-\).
+The answer is at the bottom of this page\footnotemark.
+
+\footnotetext{\rotatebox[origin=c]{180}{Answer: \(3 \times (7 \times (50 - 10) - 25)\)}}
+
+Our strategy for finding solutions to a given puzzle is to describe precisely
+the type of solutions to a puzzle, and then show that that type is finite.
+So what is a ``solution'' to
+
 \section{Notation and Background}

### 2020-06-14 sections/introduction.tex
finished countdown in introduction

@@
 the type of solutions to a puzzle, and then show that that type is finite.
-So what is a ``solution'' to
+So what is a ``solution'' to a countdown puzzle?
+Broadly, it has two parts:
+\begin{description}
+  \item[A Transformation] from a list of numbers to an expression.
+  \item[A Predicate] showing that the expression is valid and evaluates to the
+    target.
+\end{description}
+The first part is described in Figure~\ref{countdown-transform}.
 
+\input{figures/countdown-transformation}
+
+This transformation has four steps.
+First (Fig.~\ref{countdown-selection}) we have to pick which numbers we include
+in our solution.
+The representation of this transformation should be fairly simple: to pick from
+\(n\) objects we can have an \(n\)-tuple of booleans.
+
+Secondly (Fig.~\ref{countdown-permutation}) we have to permute the chosen
+numbers.
+The representation for a permutation is a little trickier to envision: proving
+that it's finite is trickier still.
+We will need to rely on some of the more involved lemmas later on for this
+problem.
+
+The third step (Fig.~\ref{countdown-operators}) is quite like the first: it's
+just another \(n\)-tuple of finite objects (in this case operators chosen from
+\(+\), \(\times\), \(-\), and \(\div\)).
+Although it is complicated slightly by the fact that the \(n\) in this
+\(n\)-tuple is dependent on the number of trues in the previously chosen tuple
+of booleans (in terms of types, that means we'll need a \(\Sigma\) rather than a
+\(\times\), explanations of which are forthcoming).
+
+Finally (Fig.~\ref{countdown-parens}), we have to parenthesise the expression in
+a certain way.
+This can be encapsulated by a binary tree with a certain number of leaves:
+proving that that is finite is tricky again.
+
+Once we have proven that there are finitely many transformations for a list of
+numbers, we will then have to filter them down to those transformations which
+are valid, and evaluate to the target.
+This amounts to proving that the decidable subset of a finite set is also
+finite.
+
+Finally, we will also want to optimise our solutions and solver: for this we
+will remove equivalent expressions, which can be accomplished with quotients.
 \section{Notation and Background}

### 2020-06-14 sections/introduction.tex
beginning change to agda code instead of maths

@@
 For the various type formers we use the following notation:
+\subsubsection{Type}
+We use \(\AgdaDatatype{Type}\) to denote the universe of (small) types.
+The universe level is denoted with a subscript number, starting at 0.
+``Type families'' are functions into \(\AgdaDatatype{Type}\).
+\subsubsection{Simple Types}
+The three basic types---often called 0, 1, and 2---in MLTT here will be denoted
+with their more common names: \(\bot\), \(\top\), and \(\mathbf{Bool}\),
+respectively.
+\begin{figure}[h!]
+  \begin{multicols}{3}
+    \begin{agdalisting}
+      \ExecuteMetaData[agda/Snippets/Introduction.tex]{bot}
+    \end{agdalisting}
+    \begin{agdalisting}
+      \ExecuteMetaData[agda/Snippets/Introduction.tex]{top}
+    \end{agdalisting}
+    \begin{agdalisting}
+      \ExecuteMetaData[agda/Snippets/Introduction.tex]{bool}
+    \end{agdalisting}
+  \end{multicols}
+  \caption{Agda Definitions of the Three Basic Types}
+\end{figure}
+\subsubsection{Dependent Sums}
+Dependent sums are denoted with the usual \(\Sigma\) symbol, and are defined in
+Fig.~\ref{sigma-def}.
+\begin{figure}[h!]
+  \begin{multicols}{2}
+    \begin{agdalisting}
+      \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma}
+    \end{agdalisting}
+    \begin{agdalisting}
+      \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma-syntax-1}
+      \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma-syntax-2}
+      \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma-syntax-3}
+    \end{agdalisting}
+  \end{multicols}
+  \caption{Agda Definitions of Dependent Sums}
+  \label{sigma-def}
+\end{figure}
+
 \begin{description}
-  \item[Type] We use \(\mathbf{Type}\) to denote the universe of (small) types.
-    ``Type families'' are functions into \(\mathbf{Type}\).
   \item[0 , 1 , 2] We call the \(\mathbf{0}\), \(\mathbf{1}\), and

### 2020-06-15 sections/introduction.tex
more converted into agda notation

@@
 \section{Notation and Background}
-\subsection{Notation}
 We work in Cubical Type Theory \cite{cohenCubicalTypeTheory2016}.
-For the various type formers we use the following notation:
-\subsubsection{Type}
+The notation we use comes from the conventions used in the Agda standard
+library.
+\begin{definition}[Type]
   We use \(\AgdaDatatype{Type}\) to denote the universe of (small) types.
@@
   ``Type families'' are functions into \(\AgdaDatatype{Type}\).
-\subsubsection{Simple Types}
-The three basic types---often called 0, 1, and 2---in MLTT here will be denoted
-with their more common names: \(\bot\), \(\top\), and \(\mathbf{Bool}\),
-respectively.
-\begin{figure}[h!]
-  \begin{multicols}{3}
+\end{definition}
+\begin{definition}[Basic Types]
+  The three basic types---often called 0, 1, and 2 in MLTT---here will be
+  denoted with their more common names: \(\bot\), \(\top\), and
+  \(\mathbf{Bool}\), respectively.
+  \begin{multicols}{3} \centering
     \begin{agdalisting}
@@
     \end{agdalisting}
+
     \begin{agdalisting}
@@
   \end{multicols}
-  \caption{Agda Definitions of the Three Basic Types}
-\end{figure}
-\subsubsection{Dependent Sums}
-Dependent sums are denoted with the usual \(\Sigma\) symbol, and are defined in
-Fig.~\ref{sigma-def}.
-\begin{figure}[h!]
-  \begin{multicols}{2}
+\end{definition}
+\begin{definition}[The Dependent Sum]
+  Dependent sums are denoted with the usual \(\Sigma\) symbol, and has the
+  following definition in Agda:
+
+  \begin{center}
     \begin{agdalisting}
@@
     \end{agdalisting}
+  \end{center}
+  We will use different notations to refer to this type depending on the
+  setting.
+  The following four expressions all denote the same type:
+
+  \begin{multicols}{4} \centering
     \begin{agdalisting}
       \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma-syntax-1}
+    \end{agdalisting}
+
+    \begin{agdalisting}
       \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma-syntax-2}
+    \end{agdalisting}
+
+    \begin{agdalisting}
       \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma-syntax-3}
     \end{agdalisting}
+
+    \begin{agdalisting}
+      \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma-syntax-4}
+    \end{agdalisting}
   \end{multicols}
-  \caption{Agda Definitions of Dependent Sums}
-  \label{sigma-def}
-\end{figure}
+
+  The non-dependent product is a special instance of the dependent.
+  We indicate a simple pair of types \(A\) and \(B\) as \(A \times B\).
+\end{definition}
+\begin{definition}[Dependent Product]
+  Dependent products (dependent functions) use the \(\Pi\) symbol.
+
+  The three following expressions all denote the same type:
+
+  \begin{multicols}{3} \centering
+    \begin{agdalisting}
+      \ExecuteMetaData[agda/Snippets/Introduction.tex]{pi-syntax-1}
+    \end{agdalisting}
+
+    \begin{agdalisting}
+      \ExecuteMetaData[agda/Snippets/Introduction.tex]{pi-syntax-2}
+    \end{agdalisting}
+
+    \begin{agdalisting}
+      \ExecuteMetaData[agda/Snippets/Introduction.tex]{pi-syntax-3}
+    \end{agdalisting}
+  \end{multicols}
+
+  Again, the non-dependent case is a special case of the dependent case.
+  Non-dependent functions are denoted with the arrow (\(\rightarrow\)).
+\end{definition}
 
 \begin{description}
-  \item[0 , 1 , 2] We call the \(\mathbf{0}\), \(\mathbf{1}\), and
-    \(\mathbf{2}\) types \(\bot\), \(\top\), and \(\mathbf{Bool}\) respectively.
-    The single inhabitant of \(\top\) is tt, and the two inhabitants of
-    \(\mathbf{Bool}\) are false and true.
-    The ``negation'' of a type, written \(\neg A\), means \(A \rightarrow
-    \bot\).
-  \item[Dependent Sum and Product] We use \(\Sigma\) and \(\Pi\) for the
-    dependent sum and product, respectively.
-    The two projections from \(\Sigma\) are called fst and snd.
-    In the non-dependent case, \(\Sigma\) can be written as \(\times\), and
-    \(\Pi\) as \(\rightarrow\).
   \item[Disjoint Union] We define disjoint union as an inductive type.

### 2020-06-16 sections/introduction.tex
more agda conversion

@@
 \section{Notation and Background}
-We work in Cubical Type Theory \cite{cohenCubicalTypeTheory2016}.
-The notation we use comes from the conventions used in the Agda standard
-library.
+We work in Cubical Type Theory \cite{cohenCubicalTypeTheory2016}, specifically
+Cubical Agda \cite{vezzosiCubicalAgdaDependently2019}.
+Cubical Agda is a dependently-typed functional programming language, based on
+Martin-Löf Intuitionistic Type Theory, with a Haskell-like syntax.
+
+Being a dependently-typed language, we'll have to be clear about what we mean
+when we say ``type'' in Agda.
 \begin{definition}[Type]
@@
 \end{definition}
+
+The are two broad ways to define types in Agda: as an inductive
+\(\AgdaKeyword{data}\) type, similar to data type definitions in Haskell, or as
+a \(\AgdaKeyword{record}\).
+Here we'll define the basic type formers used in MLTT.
 \begin{definition}[Basic Types]
@@
 
-  \begin{multicols}{3} \centering
+  \begin{multicols}{3}
     \begin{agdalisting}
@@
 
-\begin{description}
-  \item[Disjoint Union] We define disjoint union as an inductive type.
-    \begin{equation}
-      \begin{alignedat}{3}
-        A \uplus B \coloneqq & \;
-        \text{inl} &: A \rightarrow A \uplus B ; \\
-        | & \;  \text{inr} &: B \rightarrow A \uplus B ;
-      \end{alignedat}
-    \end{equation}
+\begin{definition}[Disjoint Union]
+  We define disjoint union as an inductive type.
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Snippets/Introduction.tex]{disj-union}
+  \end{agdalisting}
   It is also expressible with only \(\Sigma\):
-    \(A \uplus B \simeq \Sigma(x : \mathbf{Bool}) , \text{if } x \text{ then
-    } A \text{ else } B \).
-  \item[Equalities, equivalences, and paths] We use the symbol \(\coloneqq\)
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma-disj-union}
+  \end{agdalisting}
+\end{definition}
+\begin{definition}[Equalities, equivalences, and paths] We use the symbol \(\coloneqq\)
   for definitions.
@@
   used.
-  \item[Lambdas] \todo{Figure out how to describe lambdas}
-\end{description}
-\subsection{Cubical Type Theory}
-Cubical Type Theory \cite{cohenCubicalTypeTheory2016} is a constructive type
-theory with an implementation in Cubical Agda
-\cite{vezzosiCubicalAgdaDependently2019}.
-It allows us to do much of the same theory as in HoTT, but crucially the
-univalence ``axiom'' is a \emph{theorem}, giving it computational content.
+\end{definition}
+
 \begin{definition}[Path Types] \label{path-types}

### 2020-06-16 sections/introduction.tex
more conversion to agda

@@
   The following four expressions all denote the same type:
-
-  \begin{multicols}{4} \centering
+  \begin{multicols}{4}
     \begin{agdalisting}
@@
   \end{multicols}
-
   The non-dependent product is a special instance of the dependent.
@@
 \end{definition}
-\begin{definition}[Equalities, equivalences, and paths] We use the symbol \(\coloneqq\)
-  for definitions.
+\begin{definition}[Equalities, equivalences, and paths]
   \(\simeq\) will be used for equivalences, and \(\equiv\) for equalities.
@@
   \cite[definition 3.1.1]{hottbook}.
-  \begin{alignat}{2}
-    &\text{isContr}(A)    &&\coloneqq \Sigma(x : A) , \Pi(y : A) , (x \equiv y) \\
-    &\text{isProp}(A)     &&\coloneqq \Pi(x, y : A) , (x \equiv y) \\
-    &\text{isSet}(A)      &&\coloneqq \Pi(x, y : A) , \text{isProp}(x \equiv y)
-  \end{alignat}
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Snippets/Introduction.tex]{hlevels}
+  \end{agdalisting}
 \end{definition}
-\begin{definition}[Fibres] \label{fibres}
-  A fibre \cite[definition 4.2.4]{hottbook} is defined over some function \(f :
+
+\begin{definition}[Fibers] \label{fibers}
+  A fiber \cite[definition 4.2.4]{hottbook} is defined over some function \(f :
   A \rightarrow B\).
-  \begin{equation}
-    \text{fib}_f(y) \coloneqq \Sigma(x : A) , (f (x) \equiv y)
-  \end{equation}
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Snippets/Introduction.tex]{fiber}
+  \end{agdalisting}
 \end{definition}

### 2020-06-16 scripts/haskell-from-toplevel.sh
removed haskell code

@@ -1,7 +1,0 @@
-#!/bin/bash
-withcode=$1
-nocode=${withcode#"haskell/"}
-cd haskell || exit
-dir=$(dirname "$nocode")
-file=$(basename "$nocode" .lhs).tex
-lhs2TeX -o "$dir/$file" "$nocode"

### 2020-06-16 scripts/init-lhs.sh

@@ -1,18 +1,0 @@
-#!/bin/bash
-
-if [ ! -f "locallhs2TeX.lhs" ]; then
-    echo "%include polycode.fmt" > "locallhs2TeX.lhs"
-fi
-
-if [ ! -f "locallhs2TeX.sty" ]; then
-    lhs2TeX -o "locallhs2TeX.sty" "locallhs2TeX.lhs"
-fi
-
-find haskell -type f -name '*.lhs' | while read -r code ; do
-    dir=$(dirname "$code")
-    file="$dir"/$(basename "$code" .lhs).tex
-    if [ ! -e "$file" ]
-    then
-        ./haskell-from-toplevel.sh "$code"
-    fi
-done

### 2020-06-16 scripts/check-all.sh
initial

@@ -1,11 +1,0 @@
-#!/bin/bash
-( cd agda
-echo '{-# OPTIONS --cubical --safe #-}
-' > Everything
-echo 'module Everything where
-' >> Everything
-find -E . -type f -regex '.*\.l?agda' | cut -c 3- | cut -f1 -d'.' | sed 's/\//\./g' | sed 's/^/open import /' >> Everything
-mv Everything Everything.agda
-agda --html --html-dir=../docs Everything.agda
-rm Everything.agda
-)

### 2020-06-16 scripts/init-missing-lagda.sh

@@ -1,10 +1,0 @@
-#!/bin/bash
-cd agda || exit
-find . -type f -name '*.lagda' | while read -r code ; do
-    dir=$(dirname "$code")
-    file=$(basename "$code" .lagda).tex
-    if [ ! -e "$dir/$file" ]
-    then
-        agda --latex --latex-dir=. "$code"
-    fi
-done

### 2020-06-17 scripts/agda-from-toplevel.sh
proper paths

@@ -1,4 +1,4 @@
-#!/bin/bash
+#!/usr/bin/env bash
 withcode=$1
 nocode=${withcode#"agda/"}
 cd agda || exit

### 2020-06-17 scripts/check-all.sh

@@ -1,4 +1,4 @@
-#!/bin/bash
+#!/usr/bin/env bash
 ( cd agda
 echo '{-# OPTIONS --cubical --safe #-}
 ' > Everything

### 2020-06-17 scripts/init-missing-lagda.sh

@@ -1,4 +1,4 @@
-#!/bin/bash
+#!/use/bin/env bash
 cd agda || exit
 find . -type f -name '*.lagda' | while read -r code ; do
     dir=$(dirname "$code")

### 2020-06-17 scripts/init-missing-lagda.sh
Update init-missing-lagda.sh

@@ -1,4 +1,4 @@
-#!/use/bin/env bash
+#!/usr/bin/env bash
 cd agda || exit
 find . -type f -name '*.lagda' | while read -r code ; do
     dir=$(dirname "$code")

### 2020-06-17 sections/finiteness-predicates.tex
more introduction
moved definitions to their place of first usage

@@
   \end{equation}
+\end{definition}
+
+\begin{definition}[Fibers] \label{fibers}
+  A fiber \cite[definition 4.2.4]{hottbook} is defined over some function \(f :
+  A \rightarrow B\).
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Snippets/Introduction.tex]{fiber}
+  \end{agdalisting}
+\end{definition}
+
+\begin{definition}[Disjoint Union]
+  We define disjoint union as an inductive type.
+
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Snippets/Introduction.tex]{disj-union}
+  \end{agdalisting}
+
+  It is also expressible with only \(\Sigma\):
+
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma-disj-union}
+  \end{agdalisting}
+
+  Although the inductive type definition is slightly more ergonomic.
 \end{definition}
@@
 complex than what you'd find in either MLTT or classical mathematics.
-\begin{definition}[Surjections] \label{surjections}
-  We define both surjections and \emph{split} surjections here \cite[definition
+\begin{definition}[Split Surjections] \label{split-surjections}
+  We define \emph{split} surjections here \cite[definition
   4.6.1]{hottbook}.
   \begin{alignat}{3}
-    &\text{surj}(f)             &&\coloneqq \Pi(y : B) , \lVert \text{fib}_f(y) \rVert \label{surj-eqn} \\
-    &A \twoheadrightarrow B     &&\coloneqq \Sigma (f : A \rightarrow B) , \text{surj}(f) \label{surj-arrow-eqn} \\
     &\text{sp-surj}(f)          &&\coloneqq \Pi(y : B) , \text{fib}_f(y) \label{sp-surj-eqn} \\
@@
 
+\begin{definition}[Decidable Types]
+  \begin{equation}
+    \mathbf{Dec}(A) \coloneqq A \uplus \neg A
+  \end{equation}
+\end{definition}
+\begin{definition}[Discrete Types]
+  A discrete type is one with decidable equality.
+  \begin{equation}
+    \text{Discrete}(A) \coloneqq \Pi(x, y : A) , \mathbf{Dec}(x \equiv y)
+  \end{equation}
+  By Hedberg's theorem \cite{hedbergCoherenceTheoremMartinLof1998} any discrete
+  type is a set.
+\end{definition}
+
 \begin{lemma} \label{discrete-surj}
@@
 \emph{contraction}.
+
+\begin{definition}[Homotopy Levels] \label{homotopy-types}
+  Types in HoTT and CuTT are not necessarily sets, as they are in MLTT.
+  Some have higher homotopies (paths which aren't unique).
+  We actually have a hierarchy of complexity of structure of path spaces in
+  types, starting with the contractions \cite[definition 3.11.1]{hottbook}, then
+  the mere propositions \cite[definition 3.3.1]{hottbook}, and the sets
+  \cite[definition 3.1.1]{hottbook}.
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Snippets/Introduction.tex]{hlevels}
+  \end{agdalisting}
+\end{definition}
+
 \begin{definition}[Unique Membership] \label{uniq-memb-def}
@@
 \emph{equivalence} with \(\mathbf{Fin}\).
+
+\begin{definition}[Equivalences] \label{equivalences}
+  We will take contractible maps \cite[definition 4.4.1]{hottbook} as our
+  ``default'' definition of equivalences.
+  \begin{alignat}{2}
+    &\text{isEquiv}(f) &&\coloneqq \Pi(y : B) , \text{isContr}(\text{fib}_f(y)) \label{is-equiv-def} \\
+    &A \simeq B        &&\coloneqq \Sigma(f : A \rightarrow B) , \text{isEquiv}(f) \label{equiv-def}
+  \end{alignat}
+\end{definition}
 \begin{lemma} \label{bishop-equiv}
@@
 of course we can derive the proof on manifest Bishop finiteness in a few lines.
-\begin{lemma} \label{split-enum-pi}
+\begin{theorem} \label{split-enum-pi}
   Split enumerability is closed under dependent functions.
@@
   \end{equation}
-\end{lemma}
+\end{theorem}
 \begin{proof}
@@
 (def.~\ref{prop-trunc}).
+
+\begin{definition}[Higher Inductive Types] \label{HITs}
+  Normal inductive types have \emph{point} constructors: constructors which
+  construct values of the type.
+  Higher Inductive Types (HITs) also have \emph{path} constructors: ways to
+  construct paths in the type.
+\end{definition}
+\begin{definition}[Propositional Truncation] \label{prop-trunc}
+  The type \(\lVert A \rVert\) on some type \(A\) is a propositionally truncated
+  proof of \(A\) \cite[3.7]{hottbook}.
+  In other words, it is a proof that some \(A\) exists, but it does not tell you
+  \emph{which} \(A\).
+
+  It is defined as a Higher Inductive Type:
+  \begin{equation} {
+    \begin{alignedat}{3}
+      \lVert A \rVert \coloneqq & \; \lvert \wc \rvert &:& \; A \rightarrow \lVert A \rVert ; \\
+                              | & \; \text{squash}     &:& \; \Pi {(x, y : \lVert A \rVert)} , x \equiv y  ; 
+    \end{alignedat} }
+  \end{equation}
+  We will use two eliminators from \(\lVert A \rVert\) in this paper.
+  \begin{enumerate}
+  \item \label{elim-prop-prop} For any function \(A \rightarrow B\), where
+    \(\text{isProp}(B)\), we have a function \(\lVert A \rVert \rightarrow B\).
+  \item \label{elim-prop-coh} We can eliminate from \(\lVert A \rVert\) with a
+    function \(f : A \rightarrow B\) iff \(f\) ``doesn't care'' about the
+    choice of \(A\):
+    \[\Pi {(x , y : A)} , f(x) \equiv f(y) \]
+    Formally speaking, \(f\) needs to be ``coherently constant''
+    \cite{krausGeneralUniversalProperty2015}, and \(B\) needs to be an
+    \(n\)-type for some finite \(n\).
+  \end{enumerate}
+\end{definition}
 \begin{definition}[Cardinal Finiteness]
@@
 As a function-based definition, this predicate represents surjections.
+
+\begin{definition}[Surjections] \label{surjections}
+  We define proper surjections (not split surjections) here \cite[definition
+  4.6.1]{hottbook}.
+  \begin{alignat}{3}
+    &\text{surj}(f)             &&\coloneqq \Pi(y : B) , \lVert \text{fib}_f(y) \rVert \label{surj-eqn} \\
+    &A \twoheadrightarrow B     &&\coloneqq \Sigma (f : A \rightarrow B) , \text{surj}(f) \label{surj-arrow-eqn}
+  \end{alignat}
+\end{definition}
 \begin{lemma}

### 2020-06-17 sections/introduction.tex
moved out-directory instructions to makefile
moved definitions to their place of first usage

@@
 
-\begin{definition}[Disjoint Union]
-  We define disjoint union as an inductive type.
-  \begin{agdalisting}
-    \ExecuteMetaData[agda/Snippets/Introduction.tex]{disj-union}
-  \end{agdalisting}
-  It is also expressible with only \(\Sigma\):
-  \begin{agdalisting}
-    \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma-disj-union}
-  \end{agdalisting}
-\end{definition}
-\begin{definition}[Equalities, equivalences, and paths]
-  \(\simeq\) will be used for equivalences, and \(\equiv\) for equalities.
-  Of course, we know that \((A \simeq B) \simeq (A \equiv B)\) by univalence,
-  so the distinction isn't terribly important in usage: we will only use one
-  or the other as a suggestion of how we constructed it or how it is to be
-  used.
-\end{definition}
-
 \begin{definition}[Path Types] \label{path-types}
@@
 }
-\begin{definition}[Homotopy Levels] \label{homotopy-types}
-  Types in HoTT and CuTT are not necessarily sets, as they are in MLTT.
-  Some have higher homotopies (paths which aren't unique).
-  We actually have a hierarchy of complexity of structure of path spaces in
-  types, starting with the contractions \cite[definition 3.11.1]{hottbook}, then
-  the mere propositions \cite[definition 3.3.1]{hottbook}, and the sets
-  \cite[definition 3.1.1]{hottbook}.
-  \begin{agdalisting}
-    \ExecuteMetaData[agda/Snippets/Introduction.tex]{hlevels}
-  \end{agdalisting}
-\end{definition}
-
-\begin{definition}[Fibers] \label{fibers}
-  A fiber \cite[definition 4.2.4]{hottbook} is defined over some function \(f :
-  A \rightarrow B\).
-  \begin{agdalisting}
-    \ExecuteMetaData[agda/Snippets/Introduction.tex]{fiber}
-  \end{agdalisting}
-\end{definition}
-\begin{definition}[Equivalences] \label{equivalences}
-  We will take contractible maps \cite[definition 4.4.1]{hottbook} as our
-  ``default'' definition of equivalences.
-  \begin{alignat}{2}
-    &\text{isEquiv}(f) &&\coloneqq \Pi(y : B) , \text{isContr}(\text{fib}_f(y)) \label{is-equiv-def} \\
-    &A \simeq B        &&\coloneqq \Sigma(f : A \rightarrow B) , \text{isEquiv}(f) \label{equiv-def}
-  \end{alignat}
-\end{definition}
-\begin{definition}[Decidable Types]
-  \begin{equation}
-    \mathbf{Dec}(A) \coloneqq A \uplus \neg A
-  \end{equation}
-\end{definition}
-\begin{definition}[Discrete Types]
-  A discrete type is one with decidable equality.
-  \begin{equation}
-    \text{Discrete}(A) \coloneqq \Pi(x, y : A) , \mathbf{Dec}(x \equiv y)
-  \end{equation}
-  By Hedberg's theorem \cite{hedbergCoherenceTheoremMartinLof1998} any discrete
-  type is a set.
-\end{definition}
-\begin{definition}[Higher Inductive Types] \label{HITs}
-  Normal inductive types have \emph{point} constructors: constructors which
-  construct values of the type.
-  Higher Inductive Types (HITs) also have \emph{path} constructors: ways to
-  construct paths in the type.
-\end{definition}
-\begin{definition}[Propositional Truncation] \label{prop-trunc}
-  The type \(\lVert A \rVert\) on some type \(A\) is a propositionally truncated
-  proof of \(A\) \cite[3.7]{hottbook}.
-  In other words, it is a proof that some \(A\) exists, but it does not tell you
-  \emph{which} \(A\).
-
-  It is defined as a Higher Inductive Type:
-  \begin{equation} {
-    \begin{alignedat}{3}
-      \lVert A \rVert \coloneqq & \; \lvert \wc \rvert &:& \; A \rightarrow \lVert A \rVert ; \\
-                              | & \; \text{squash}     &:& \; \Pi {(x, y : \lVert A \rVert)} , x \equiv y  ; 
-    \end{alignedat} }
-  \end{equation}
-  We will use two eliminators from \(\lVert A \rVert\) in this paper.
-  \begin{enumerate}
-  \item \label{elim-prop-prop} For any function \(A \rightarrow B\), where
-    \(\text{isProp}(B)\), we have a function \(\lVert A \rVert \rightarrow B\).
-  \item \label{elim-prop-coh} We can eliminate from \(\lVert A \rVert\) with a
-    function \(f : A \rightarrow B\) iff \(f\) ``doesn't care'' about the
-    choice of \(A\):
-    \[\Pi {(x , y : A)} , f(x) \equiv f(y) \]
-    Formally speaking, \(f\) needs to be ``coherently constant''
-    \cite{krausGeneralUniversalProperty2015}, and \(B\) needs to be an
-    \(n\)-type for some finite \(n\).
-  \end{enumerate}
-\end{definition}
 

### 2020-06-17 sections/introduction.tex
introduction finished (?)

@@
 \chapter{Introduction}
-\section{Foreword}
-\todo[inline]{Foreword}
-\section{Contributions}
 We are interested in constructive notions of finiteness, formalised in Cubical
@@
 differences with finiteness.
-\subsection{The Varieties of Finiteness}
+\section{The Varieties of Finiteness}
 In Section~\ref{finiteness-predicates} we will explore a number of different
 predicates for finiteness.
-In contrast to classical finiteness, in a constructive setting there is a wide
-variety of predicates which all have some claim to being the formal
-interpretation of ``finiteness'' \cite{coquandConstructivelyFinite2010}.
+In contrast to classical finiteness, in a constructive setting there are many
+predicates which all have some claim to being the formal interpretation of
+``finiteness'' \cite{coquandConstructivelyFinite2010}.
 The particular predicates we are interested in are organised in
@@
 that one predicate can be derived from another.
+Each arrow in Figure~\ref{finite-classification} corresponds to a proof of
+implication: cardinal finiteness, for instance, with a strict total order,
+implies split enumerability (Theorem~\ref{cardinal-to-manifest-bishop}).
 
@@
 order on the underlying type.
-We will prove that a more informative finiteness predicate can be
-derived from a less informative one by providing the missing information
-(Theorem~\ref{cardinal-to-manifest-bishop}).
 
@@
 there are Cardinally finite (Section~\ref{cardinal-finiteness}).
-We will prove that we can always derive the less restrictive predicate from the
-more restrictive one, and that we can go in the other direction by satisfying
-the missing requirement (decidable equality in all of these cases).
 
@@
 (dependent functions) (Theorem~\ref{split-enum-pi}).
-\subsection{Toposes and Finite Sets}
+\section{Toposes and Finite Sets}
 In Section~\ref{topos}, we will explore the categorical interpretation of
@@
 additional to that.
-\subsection{Countability Predicates}
+\section{Countability Predicates}
 After the finite predicates, we will briefly look at the infinite countable
@@
 Kleene star (Theorem~\ref{split-countability-sigma}).
-\subsection{Search}
+\section{Search}
 All of our work is formalised in Cubical Agda
@@
 \section{Countdown}
-Countdown is a well-known functional programming puzzle (later turned into a TV
-show in the UK), first popularised as a puzzle in which to demonstrate
-functional algorithms in \cite{huttonCountdownProblem2002}.
+The Countdown problem \cite{huttonCountdownProblem2002} is a well-known puzzle
+in functional programming (which was apparently turned into a TV show).
 As a running example in this paper, we will produce a verified program which
@@
 \end{gather*}
-We'll allow the use of \(\times\), \(+\), and \(-\).
+We'll allow the use of \(+\), \(-\), \(\times\), and \(\div\).
 The answer is at the bottom of this page\footnotemark.
@@
 The representation of this transformation should be fairly simple: to pick from
-\(n\) objects we can have an \(n\)-tuple of booleans.
+\(n\) objects we can have a vector of \(n\) booleans.
 
@@
 The third step (Fig.~\ref{countdown-operators}) is quite like the first: it's
-just another \(n\)-tuple of finite objects (in this case operators chosen from
-\(+\), \(\times\), \(-\), and \(\div\)).
+just another vector of length \(n\) of finite objects (in this case operators
+chosen from \(+\), \(\times\), \(-\), and \(\div\)).
 Although it is complicated slightly by the fact that the \(n\) in this
@@
     \begin{agdalisting}
-      \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma-syntax-2}
-    \end{agdalisting}
-
-    \begin{agdalisting}
       \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma-syntax-3}
@@
     \end{agdalisting}
+
+    \begin{agdalisting}
+      \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma-syntax-2}
+    \end{agdalisting}
   \end{multicols}
   The non-dependent product is a special instance of the dependent.
-  We indicate a simple pair of types \(A\) and \(B\) as \(A \times B\).
+  We denote a simple pair of types \(A\) and \(B\) as \(A \times B\).
 \end{definition}
@@
   Dependent products (dependent functions) use the \(\Pi\) symbol.
-
   The three following expressions all denote the same type:
-
   \begin{multicols}{3}
@@
   \end{multicols}
-
-  Again, the non-dependent case is a special case of the dependent case.
   Non-dependent functions are denoted with the arrow (\(\rightarrow\)).
@@
 
+At this point, as a quick example, we can define the first of our objects for
+the countdown transformation: the vector of Booleans for selection.
+A vector is relatively simple to define: a vector of zero elements is simply a
+unit, a vector of \(n+1\) elements is the product of an element and a vector of
+\(n\) elements.
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Data/Vec/Iterated.tex]{vec-def}
+\end{agdalisting}
+From this we can see that a vector of \(n\) Booleans has the type
+\(\AgdaDatatype{Vec} \; \AgdaDatatype{Bool} \; n\).
+
+Finally, there is one last thing we must define before moving on to the
+finiteness predicates: paths.
 \begin{definition}[Path Types] \label{path-types}
@@
   Paths\footnotemark.
-  The internal structure of paths is largely irrelevant to us here, as we will
-  generally treat \(\equiv\) as a black-box equivalence relation with
-  substitution and congruence.
+  The nature and internal structure of Paths is complex and central to how
+  Cubical Type Theory ``implements'' Homotopy Type Theory, but those details are
+  not relevant to us here.
+  Instead, we only need to know that univalence holds for paths, and path types
+  do indeed compute in Cubical Agda.
 \end{definition}

### 2020-06-17 sections/introduction.tex
applied linter suggestions

@@
 We are interested in constructive notions of finiteness, formalised in Cubical
-Type Theory \cite{cohenCubicalTypeTheory2016}.
+Type Theory~\cite{cohenCubicalTypeTheory2016}.
 In this paper we will explore five such notions of finiteness, including their
@@
 facilitated in a fundamental way by univalence.
-Along the way we will use the Countdown problem
-\cite{huttonCountdownProblem2002} as an example, and provide a program which
-produces verified solutions to the puzzle.
+Along the way we will use the Countdown
+problem~\cite{huttonCountdownProblem2002} as an example, and provide a program
+which produces verified solutions to the puzzle.
 We will also briefly examine countability, and demonstrate its parallels and
@@
 predicates which all have some claim to being the formal interpretation of
-``finiteness'' \cite{coquandConstructivelyFinite2010}.
+``finiteness''~\cite{coquandConstructivelyFinite2010}.
 The particular predicates we are interested in are organised in
@@
 bijection without specifying a particular bijection.
-This facility is provided by Homotopy Type Theory \cite{hottbook} in the form of
+This facility is provided by Homotopy Type Theory~\cite{hottbook} in the form of
 propositional truncation, and it is what allows us to prove the bulk of
@@
 
-For each predicate we will also prove its closure properties (i.e. that the
+For each predicate we will also prove its closure properties (i.e.\ that the
 product of two finite sets is finite).
@@
 ``language'' of a topos is a reasonable choice for a principled language for
-constructing proofs of finiteness in the style of QuickCheck
-\cite{claessenQuickCheckLightweightTool2011} generators.
+constructing proofs of finiteness in the style of
+QuickCheck~\cite{claessenQuickCheckLightweightTool2011} generators.
 
@@
 (with some caveats) is an important step in characterising the categorical
-implications of Homotopy Type Theory, first proven in
-\cite{rijkeSetsHomotopyType2015}.
+implications of Homotopy Type Theory, first proven
+in~\cite{rijkeSetsHomotopyType2015}.
 Our work is a formalisation of this result (and the first such formalisation
@@
 \section{Search}
-All of our work is formalised in Cubical Agda
-\cite{vezzosiCubicalAgdaDependently2019}: as a result, the constructive
+All of our work is formalised in Cubical
+Agda~\cite{vezzosiCubicalAgdaDependently2019}: as a result, the constructive
 interpretation of each proof is actually a program which can be run on a
@@
 In Section~\ref{search} we will package up the ``search'' aspect of finiteness
-into a library for proof search: similar libraries have been built in
-\cite{fruminFiniteSetsHomotopy2018} and
-\cite{firsovDependentlyTypedProgramming2015}.
+into a library for proof search: similar libraries have been built
+in~\cite{fruminFiniteSetsHomotopy2018}
+and~\cite{firsovDependentlyTypedProgramming2015}.
 Our library differs from those in three important ways: firstly, it is strictly
@@
 \section{Countdown}
-The Countdown problem \cite{huttonCountdownProblem2002} is a well-known puzzle
+The Countdown problem~\cite{huttonCountdownProblem2002} is a well-known puzzle
 in functional programming (which was apparently turned into a TV show).
@@
 \section{Notation and Background}
-We work in Cubical Type Theory \cite{cohenCubicalTypeTheory2016}, specifically
-Cubical Agda \cite{vezzosiCubicalAgdaDependently2019}.
+We work in Cubical Type Theory~\cite{cohenCubicalTypeTheory2016}, specifically
+Cubical Agda~\cite{vezzosiCubicalAgdaDependently2019}.
 Cubical Agda is a dependently-typed functional programming language, based on
@@
 a \(\AgdaKeyword{record}\).
-Here we'll define the basic type formers used in MLTT.
+Here we'll define the basic type formers used in MLTT.\@
 \begin{definition}[Basic Types]
@@
 finiteness predicates: paths.
-\begin{definition}[Path Types] \label{path-types}
+\begin{definition}[Path Types]\label{path-types}
   The equality type (which we denote with \(\equiv\)) in CuTT is the type of
@@
 
-\footnotetext{
+\footnotetext{%
   Actually, CuTT does have an identity type with similar semantics to the
-  identity type in MLTT.
+  identity type in MLTT.\@
   We do not use this type anywhere in our work, however, so we will not consider

### 2020-06-17 sections/finiteness-predicates.tex
switched container to sigma

@@
   \(\mathit{support}\).
-  \begin{equation} \label{split-enum-def-eqn}
-    \mathcal{E}!(A) \coloneqq \Sigma {(\mathit{support} : \textbf{List}(A))} , \Pi {(x : A)} , x \in \mathit{support}
-  \end{equation}
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable/Container.tex]{split-enum-def}
+  \end{agdalisting}
   We call the first component of this pair the ``support'' list, and the second
@@
 \end{definition}
+
 We used some extra types in the above definition, which we will define here:

### 2020-06-17 sections/finiteness-predicates.tex
listings now numbered as equations

@@
   \(\mathit{support}\).
-  \begin{agdalisting}
+  \begin{agdalisting} \label{split-enum-def-eqn}
     \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable/Container.tex]{split-enum-def}
@@
   In this paper we define lists as \emph{containers}:
-  \begin{equation}
-    \mathbf{List} \coloneqq \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket
-  \end{equation}
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Container/List.tex]{list-def}
+  \end{agdalisting}
   This may seem quite foreign or complex in comparison to the usual definition
   of lists as an inductive data type:
-  \begin{inductivetype}{\mathbf{List}(A)}
-    \inductivetypeclause{\left[ \right]}{\mathbf{List}(A)}
-    \inductivetypeclause{\wc \dblcolon \wc}{A \rightarrow \mathbf{List}(A) \rightarrow \mathbf{List}(A)}
-  \end{inductivetype}
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Data/List/Base.tex]{list-def}
+  \end{agdalisting}
   But it turns out to be quite useful for later proofs.

### 2020-06-17 sections/introduction.tex
better spacing

@@
   \(\mathbf{Bool}\), respectively.
-  \begin{multicols}{3} \centering
+  \begin{multicols}{3}
     \begin{agdalisting}
@@
       \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma-syntax-1}
-    \end{agdalisting}
-
+    \end{agdalisting} \columnbreak
     \begin{agdalisting}
       \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma-syntax-3}
-    \end{agdalisting}
-
+    \end{agdalisting} \columnbreak
     \begin{agdalisting}
       \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma-syntax-4}
-    \end{agdalisting}
-
+    \end{agdalisting} \columnbreak
     \begin{agdalisting}
@@
     \end{agdalisting}
-  \end{multicols}
+  \end{multicols}\vspace{-1.5\baselineskip}%
   The non-dependent product is a special instance of the dependent.
@@
       \ExecuteMetaData[agda/Snippets/Introduction.tex]{pi-syntax-1}
-    \end{agdalisting}
-
+    \end{agdalisting} \columnbreak
     \begin{agdalisting}
       \ExecuteMetaData[agda/Snippets/Introduction.tex]{pi-syntax-2}
-    \end{agdalisting}
-
+    \end{agdalisting} \columnbreak
     \begin{agdalisting}
@@
     \end{agdalisting}
-  \end{multicols}
+  \end{multicols}\vspace{-1.5\baselineskip}
   Non-dependent functions are denoted with the arrow (\(\rightarrow\)).
@@
 From this we can see that a vector of \(n\) Booleans has the type
-\(\AgdaDatatype{Vec} \; \AgdaDatatype{Bool} \; n\).
+\(\AgdaDatatype{Vec} \; \AgdaDatatype{Bool} \; n\)
 

### 2020-06-18 sections/finiteness-predicates.tex
more agda instead of mathematical notation

@@
 
-We used some extra types in the above definition, which we will define here:
+We have used two types there which we have not yet defined:
+\(\AgdaDatatype{List}\) and \(\AgdaDatatype{\ensuremath{\in}}\).
+We will define them here.
 
 \begin{definition}[\(\mathbf{List}\)] \label{List}
-  In this paper we define lists as \emph{containers}:
+  In this paper we will work with two equivalent definitions of lists.
+  The first is the standard definition as an inductive type:
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Data/List/Base.tex]{list-def}
+  \end{agdalisting}
+
+  The second way to define lists is to define them as a \emph{container}:
   \begin{agdalisting}
@@
   \end{agdalisting}
-  This may seem quite foreign or complex in comparison to the usual definition
-  of lists as an inductive data type:
+  The reason we use this second strange definition is that it turns out to be
+  quite useful in some later proofs.
+  We have proven the two types equivalent in our formalisation, however, so we
+  can switch between them freely without loss of generality.
+\end{definition}
+
+In defining lists we have introduced another concept which needs defining:
+\(\AgdaDatatype{Fin}\).
+\begin{definition}[\(\mathbf{Fin}\)] \label{Fin}
+  \(\mathbf{Fin}(n)\) is the type of natural numbers smaller than \(n\). We
+  define it the standard way:
   \begin{agdalisting}
-    \ExecuteMetaData[agda/Data/List/Base.tex]{list-def}
-  \end{agdalisting}
-  But it turns out to be quite useful for later proofs.
-
-  Moreover, the inductive type and the container-based type are equivalent, as
-  is proven in our formalisation.
-\end{definition}
+    \ExecuteMetaData[agda/Data/Fin/Base.tex]{fin-def}
+  \end{agdalisting}
+\end{definition}
+Here \(\uplus\) refers to the disjoint union of two types.
+\begin{definition}[Disjoint Union]
+  We define disjoint union as an inductive type.
+
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Snippets/Introduction.tex]{disj-union}
+  \end{agdalisting}
+
+  It is also expressible with only \(\Sigma\):
+
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma-disj-union}
+  \end{agdalisting}
+
+  Although the inductive type definition is slightly more ergonomic.
+\end{definition}
+
+After that interlude, we can get back to defining containers.
 \begin{definition}[Containers] \label{container-def}
@@
   We ``interpret'' a container into a functor defined like so:
-  \begin{equation} \label{container-interp}
-    \llbracket S , P \rrbracket(A) \coloneqq \Sigma {(s : S)} , \left( P(s) \rightarrow A \right)
-  \end{equation}
+  \begin{agdalisting} \label{container-interp}
+    \ExecuteMetaData[agda/Container.tex]{container-interp}
+  \end{agdalisting}
   Membership of a container can be defined like so:
-  \begin{equation} \label{container-membership}
-    x \in \mathit{xs} \coloneqq \text{fib}_{\text{snd}(\mathit{xs})}(x)
-  \end{equation}
+  \begin{agdalisting} \label{container-membership}
+    \ExecuteMetaData[agda/Container/Membership.tex]{membership-def}
+  \end{agdalisting}
 \end{definition}
@@
 
-\begin{definition}[Disjoint Union]
-  We define disjoint union as an inductive type.
-
-  \begin{agdalisting}
-    \ExecuteMetaData[agda/Snippets/Introduction.tex]{disj-union}
-  \end{agdalisting}
-
-  It is also expressible with only \(\Sigma\):
-
-  \begin{agdalisting}
-    \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma-disj-union}
-  \end{agdalisting}
-
-  Although the inductive type definition is slightly more ergonomic.
-\end{definition}
-\begin{definition}[\(\mathbf{Fin}\)] \label{Fin}
-  \(\mathbf{Fin}(n)\) is the type of natural numbers smaller than \(n\). We
-  define it the standard way:
-  \begin{equation}
-    \begin{alignedat}{3}
-      & \mathbf{Fin}(0)     && \coloneqq \bot ; \\
-      & \mathbf{Fin}(n + 1) && \coloneqq \top \uplus \mathbf{Fin}(n) ;
-    \end{alignedat}
-  \end{equation} 
-\end{definition}
 \subsection{Instances}
@@
 \begin{lemma}
-  \(\bot\), \(\top\), and \(\mathbf{Bool}\) are split enumerable.
-\end{lemma}
-\begin{proof}
-  These three types are represented by fairly simple finite sets:
-  \begin{align}
-    \bot &= \left\{  \right\} \\
-    \top &= \left\{ \text{tt}  \right\} \\
-    \mathbf{Bool} &= \left\{  \text{false}, \text{true} \right\}
-  \end{align}
-  Which leads to quite simple proofs of finiteness.
-
-  The support list for is \(\left[  \right]\), \(\left[ \text{tt} \right]\), and
-  \(\left[ \text{false}, \text{true} \right]\), respectively, and the cover
-  proof is a function that returns the index of the supplied element.
-\end{proof}
+  \(\bot\), \(\top\), and \(\AgdaDatatype{Bool}\) are split enumerable.
+\end{lemma}
+\begin{proof}
+  These three types are quite obviously finite: we will show only the proof of
+  finiteness for \(\AgdaDatatype{Bool}\) here for brevity's sake.
   \begin{agdalisting}
-  The code in our formalisation of these three proofs is quite simple.
-  Just as an example, the proof that \(\mathbf{Bool}\) is split enumerable is as
-  follows:
     \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{bool-inst}
   \end{agdalisting}
+\end{proof}
 With the most basic simple types out of the way, the obvious next choice is the
@@
 \emph{membership proof} must be unique.
-This would disallow a definition of \(\mathcal{E}!(\mathbf{Bool})\) with
+This would disallow a definition of \(\mathcal{E}!(\AgdaDatatype{Bool})\) with
 duplicates, as there are multiple values which inhabit the type \(\text{false}
@@
 of decidable equality on functions in MLTT, we know that we won't be able to
-show that any kind of function is finite; even one like \(\mathbf{Bool}
-\rightarrow \mathbf{Bool}\).
+show that any kind of function is finite; even one like \(\AgdaDatatype{Bool}
+\rightarrow \AgdaDatatype{Bool}\).
 
@@
   The two following proofs are both valid proofs of the finiteness of
-  \(\mathbf{Bool}\), and both do not include any duplicates.
+  \(\AgdaDatatype{Bool}\), and both do not include any duplicates.
   However they still differ:
@@
 \(\mathcal{C}\) trivially (unlike with \(\mathcal{E}!\) and \(\mathcal{B}\)).
-The cases for \(\bot\), \(\top\), and \(\mathbf{Bool}\) are simple to adapt: we
+The cases for \(\bot\), \(\top\), and \(\AgdaDatatype{Bool}\) are simple to adapt: we
 can just propositionally truncate their Bishop finiteness proof.

### 2020-06-18 sections/finiteness-predicates.tex
Update finiteness-predicates.tex

@@
     \frac{
-      \mathcal{E}!(A) \; \; \; \Pi(x : A) , \mathcal{E}!(U(x))
+      \AgdaDatatype{\ensuremath{\mathcal{E}!}}(A) \; \; \; \Pi(x : A) , \AgdaDatatype{\ensuremath{\mathcal{E}!}}(U(x))
     }{
-      \mathcal{E}!(\Sigma(x : A) , U(x))
+      \AgdaDatatype{\ensuremath{\mathcal{E}!}}(\Sigma(x : A) , U(x))
     }
@@
   \begin{align}
-    \mathcal{E}!_A &: \mathcal{E}!(A) \\
-    \mathcal{E}!_U &: \Pi(x : A) , \mathcal{E}!(U(x))
+    \AgdaDatatype{\ensuremath{\mathcal{E}!}}_A &: \AgdaDatatype{\ensuremath{\mathcal{E}!}}(A) \\
+    \AgdaDatatype{\ensuremath{\mathcal{E}!}}_U &: \Pi(x : A) , \AgdaDatatype{\ensuremath{\mathcal{E}!}}(U(x))
   \end{align}
@@
   \begin{equation}
-    \mathcal{E}!(\Sigma(x : A) , U(x))
+    \AgdaDatatype{\ensuremath{\mathcal{E}!}}(\Sigma(x : A) , U(x))
   \end{equation}
@@
   \end{align}
-  To construct the support list, we apply the function \(\mathcal{E}!_U\) to
-  every element in the support list of \(\mathcal{E}!_A\), extract the support
+  To construct the support list, we apply the function \(\AgdaDatatype{\ensuremath{\mathcal{E}!}}_U\) to
+  every element in the support list of \(\AgdaDatatype{\ensuremath{\mathcal{E}!}}_A\), extract the support
   lists from the resulting finiteness proofs, and concatenate them.
@@
   \(\Sigma \; A \; U\), we note that any element of type \(\Sigma \; A \; U\)
-  must have a first component in the support list of \(\mathcal{E}!_A\), and its
-  second component must be in the result of applying \(\mathcal{E}!_U\) to that
+  must have a first component in the support list of \(\AgdaDatatype{\ensuremath{\mathcal{E}!}}_A\), and its
+  second component must be in the result of applying \(\AgdaDatatype{\ensuremath{\mathcal{E}!}}_U\) to that
   first element (since that support list contains every element of type
@@
   \begin{equation}%
-    \mathcal{E}!(A) \simeq \Sigma (n : \mathbb{N}) , \left( \mathbf{Fin}(n) \twoheadrightarrow ! \; A \right)
+    \AgdaDatatype{\ensuremath{\mathcal{E}!}}(A) \simeq \Sigma (n : \mathbb{N}) , \left( \mathbf{Fin}(n) \twoheadrightarrow ! \; A \right)
   \end{equation}
@@
   \begin{align*}
-     \mathcal{E}!(A) &
+     \AgdaDatatype{\ensuremath{\mathcal{E}!}}(A) &
     \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , x \in \mathit{xs}
-    && \text{def.~\ref{split-enum-def} }(\mathcal{E}!)
+    && \text{def.~\ref{split-enum-def} }(\AgdaDatatype{\ensuremath{\mathcal{E}!}})
     \\
@@
 With split enumerability we can see an instance of this.
-
+Consider the following proof of the finiteness of Bool:
 \begin{agdalisting} \label{bool-slop}
-  Consider the following proof of the finiteness of Bool:
   \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{bool-slop}
+\end{agdalisting}
 While it represents the ``same'' information as our previous proof, it is
 clearly not the same \emph{object}.
-\end{agdalisting}
 
@@
 \emph{membership proof} must be unique.
-This would disallow a definition of \(\mathcal{E}!(\AgdaDatatype{Bool})\) with
+This would disallow a definition of \(\AgdaDatatype{\ensuremath{\mathcal{E}!}}\;
+\AgdaDatatype{Bool}\) with
 duplicates, as there are multiple values which inhabit the type \(\text{false}
@@
 The first step in the proof, then, is to filter those duplicates out from the
-support list of the \(\mathcal{E}!\) proof: we can do this using the decidable
-equality provided by \(\mathcal{E}!\) (lemma~\ref{split-enum-discrete}).
+support list of the \(\AgdaDatatype{\ensuremath{\mathcal{E}!}}\) proof: we can do this using the decidable
+equality provided by \(\AgdaDatatype{\ensuremath{\mathcal{E}!}}\) (lemma~\ref{split-enum-discrete}).
 From there, we need to show that the membership proof carries over
@@
     \frac{
-      \mathcal{E}!(A) \; \; \; \Pi {(x : A)} , \mathcal{E}!\left( U(x) \right)
+      \AgdaDatatype{\ensuremath{\mathcal{E}!}}(A) \; \; \; \Pi {(x : A)} , \AgdaDatatype{\ensuremath{\mathcal{E}!}}\left( U(x) \right)
     }{
-      \mathcal{E}!\left(\Pi {(x : A)} , U(x)\right)
+      \AgdaDatatype{\ensuremath{\mathcal{E}!}}\left(\Pi {(x : A)} , U(x)\right)
     }
@@
     \frac{
-      \mathcal{E}!(\mathbf{Fin}(n)) \; \; \; \Pi (x : \mathbf{Fin}(n)) , \mathcal{E}!\left( U(x) \right)
+      \AgdaDatatype{\ensuremath{\mathcal{E}!}}(\mathbf{Fin}(n)) \; \; \; \Pi (x : \mathbf{Fin}(n)) , \AgdaDatatype{\ensuremath{\mathcal{E}!}}\left( U(x) \right)
     }{
-      \mathcal{E}!\left(\Pi (x : \mathbf{Fin}(n)) , U(x)\right)
+      \AgdaDatatype{\ensuremath{\mathcal{E}!}}\left(\Pi (x : \mathbf{Fin}(n)) , U(x)\right)
     }
@@
     \frac{
-      \mathcal{E}!(\mathbf{Fin}(n)) \; \; \; \Pi (x : \mathbf{Fin}(n)) , \mathcal{E}!\left( U(x) \right)
+      \AgdaDatatype{\ensuremath{\mathcal{E}!}}(\mathbf{Fin}(n)) \; \; \; \Pi (x : \mathbf{Fin}(n)) , \AgdaDatatype{\ensuremath{\mathcal{E}!}}\left( U(x) \right)
     }{
-      \mathcal{E}!\left(\mathbf{Tuple}(n, U)\right)
+      \AgdaDatatype{\ensuremath{\mathcal{E}!}}\left(\mathbf{Tuple}(n, U)\right)
     }
@@
 \mathcal{B}(A)\), closure proofs on \(\mathcal{B}\) do not transfer over to
-\(\mathcal{C}\) trivially (unlike with \(\mathcal{E}!\) and \(\mathcal{B}\)).
+\(\mathcal{C}\) trivially (unlike with \(\AgdaDatatype{\ensuremath{\mathcal{E}!}}\) and \(\mathcal{B}\)).
 The cases for \(\bot\), \(\top\), and \(\AgdaDatatype{Bool}\) are simple to adapt: we
@@
   constructors of the circle: as such, the cover proof is essentially the same
-  as the one for \(\mathcal{E}!(\top)\).
+  as the one for \(\AgdaDatatype{\ensuremath{\mathcal{E}!}}(\top)\).
 \end{proof}

### 2020-06-18 sections/finiteness-predicates.tex
agda multi-line proof

@@
 \end{lemma}
-\begin{proof}
-  \begin{align*}
-     \AgdaDatatype{\ensuremath{\mathcal{E}!}}(A) &
-    \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , x \in \mathit{xs}
-    && \text{def.~\ref{split-enum-def} }(\AgdaDatatype{\ensuremath{\mathcal{E}!}})
-    \\
-    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{fib}_{\text{snd}(\mathit{xs})}(x)
-    && \text{eqn.~\ref{container-membership} } (\in)
-    \\
-    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \text{sp-surj}(\text{snd}(\mathit{xs}))
-    && \text{eqn.~\ref{sp-surj-eqn} (sp-surj)}
-    \\
-    & \simeq \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket (A)) , \text{sp-surj}(\text{snd}(\mathit{xs}))
-    && \text{def.~\ref{List} } (\mathbf{List})
-    \\
-    & \simeq \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \mathbf{Fin}(n)) , A) , \text{sp-surj}(\text{snd}(\mathit{xs}))
-    && \text{eqn.~\ref{container-interp} } (\llbracket \wc \rrbracket)
-    \\
-    & \simeq \Sigma (n : \mathbb{N}) , \Sigma (f : \mathbf{Fin}(n) \rightarrow A) , \text{sp-surj}(f)
-    && \text{Reassociation of } \Sigma
-    \\
-    & \simeq \Sigma (n : \mathbb{N}) , ( \mathbf{Fin}(n) \twoheadrightarrow ! \; A )
-    && \text{eqn.~\ref{sp-surj-arrow-eqn} } (\twoheadrightarrow!) \; \qedhere
-  \end{align*}
-\end{proof}
-\begin{agdalisting}
-In our formalisation, the proof is a single line:
+\begin{proof} \phantom{x}
+
+  \begin{minipage}[t]{.8\textwidth}\vspace{-\baselineskip}
+    \begin{agdalisting*}
       \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{is-split-inj}
-The only step which isn't definitional equality is the reassociation of
-\(\Sigma\).
+    \end{agdalisting*}
+  \end{minipage}
+  \begin{minipage}[t]{.19\textwidth} \setstretch{1.08}
+
+      def.~\ref{split-enum-def} (\AgdaDatatype{\ensuremath{\mathcal{E}!}}) \\
+      eqn.~\ref{container-membership} (\AgdaDatatype{\ensuremath{\in}}) \\
+      eqn.~\ref{sp-surj-eqn}  \\
+      def.~\ref{List} (\(\AgdaDatatype{List}\)) \\
+      eqn.~\ref{container-interp}  \\
+      Reassociation \\
+      eqn.~\ref{sp-surj-arrow-eqn} \qedhere
+  \end{minipage}
+\end{proof}
+In the above proof syntax the
+\(\AgdaDatatype{\ensuremath{\equiv \langle{} \rangle{} }}\) connects lines which
+are definitionally equal, i.e. they are ``obviously'' equal from the type
+checker's perspective.
+Clearly, only one line isn't a definitional equality: 
+\begin{agdalisting}
   \ExecuteMetaData[agda/Data/Sigma/Properties.tex]{reassoc}
@@
 \end{definition}
-
 \begin{definition}[Unique Membership] \label{uniq-memb-def}
@@
   contraction of it.
-  \begin{equation}
-    x \inunique \mathit{xs} \coloneqq \text{isContr}(x \in \mathit{xs})
-  \end{equation}
-\end{definition}
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Container/Membership.tex]{uniq-memb-def}
+  \end{agdalisting}
+\end{definition}
+
 With this we can define manifest Bishop finiteness:
@@
   value in the type once.
-  \begin{equation}
-    \mathcal{B}(A) \coloneqq \Sigma {(\mathit{xs} : \textbf{List}(A))} , \Pi {(x : A)} , x \inunique xs
-  \end{equation}
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Cardinality/Finite/ManifestBishop/Container.tex]{bish-def}
+  \end{agdalisting}
 \end{definition}
@@
 This proof takes significantly more work.
-The ``unique membership'' condition in \(\mathcal{B}\) means that we are not
-permitted duplicates in the support list.
+The ``unique membership'' condition in
+\(\AgdaDatatype{\ensuremath{\mathcal{B}}}\) means that we are not permitted
+duplicates in the support list.
 The first step in the proof, then, is to filter those duplicates out from the
@@
   ``default'' definition of equivalences.
-  \begin{alignat}{2}
-    &\text{isEquiv}(f) &&\coloneqq \Pi(y : B) , \text{isContr}(\text{fib}_f(y)) \label{is-equiv-def} \\
-    &A \simeq B        &&\coloneqq \Sigma(f : A \rightarrow B) , \text{isEquiv}(f) \label{equiv-def}
-  \end{alignat}
+  \begin{agdalisting} \label{is-equiv-def}
+    \ExecuteMetaData[agda/Snippets/Equivalence.tex]{is-equiv-def}
+  \end{agdalisting} \vspace{-.5\baselineskip}
+  \begin{agdalisting} \label{equiv-def}
+    \ExecuteMetaData[agda/Snippets/Equivalence.tex]{equiv-def}
+  \end{agdalisting}
 \end{definition}
@@
   \begin{equation}
-    \mathcal{B}(A) \simeq \Sigma {(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \simeq A \right)
   \end{equation}
@@
   \begin{align*}
-     \mathcal{B}(A) &
+     \AgdaDatatype{\ensuremath{\mathcal{B}}}(A) &
     \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , x \inunique \mathit{xs}
-    && \text{def.~\ref{bish-def} }(\mathcal{B})
+    && \text{def.~\ref{bish-def} }(\AgdaDatatype{\ensuremath{\mathcal{B}}})
     \\
@@
   \begin{equation}
-    \mathcal{C}(A) \coloneqq \lVert \mathcal{B}(A) \rVert \simeq \lVert \Sigma(n : \mathbb{N}) , (\mathbf{Fin}(n) \simeq A) \rVert
+    \mathcal{C}(A) \coloneqq \lVert \AgdaDatatype{\ensuremath{\mathcal{B}}}(A) \rVert \simeq \lVert \Sigma(n : \mathbb{N}) , (\mathbf{Fin}(n) \simeq A) \rVert
   \end{equation}
@@
 At first glance, it might seem that we lose any useful properties we could
-derive from \(\mathcal{B}\).
+derive from \(\AgdaDatatype{\ensuremath{\mathcal{B}}}\).
 Luckily, this is not the case: by eliminator \ref{elim-prop-coh} of
@@
 \begin{equation}
-  \mathcal{C}(A) \rightarrow \mathcal{B}(A)
+  \mathcal{C}(A) \rightarrow \AgdaDatatype{\ensuremath{\mathcal{B}}}(A)
 \end{equation}
@@
   \begin{equation}
-    \neg (\Sigma(A : \mathbf{Type}) , \mathcal{C}(A) \times \neg \mathcal{B}(A))
+    \neg (\Sigma(A : \mathbf{Type}) , \mathcal{C}(A) \times \neg \AgdaDatatype{\ensuremath{\mathcal{B}}}(A))
   \end{equation}
@@
 Since we don't have a function of type \(\mathcal{C}(A) \rightarrow
-\mathcal{B}(A)\), closure proofs on \(\mathcal{B}\) do not transfer over to
-\(\mathcal{C}\) trivially (unlike with \(\AgdaDatatype{\ensuremath{\mathcal{E}!}}\) and \(\mathcal{B}\)).
+\AgdaDatatype{\ensuremath{\mathcal{B}}}(A)\), closure proofs on \(\AgdaDatatype{\ensuremath{\mathcal{B}}}\) do not transfer over to
+\(\mathcal{C}\) trivially (unlike with \(\AgdaDatatype{\ensuremath{\mathcal{E}!}}\) and \(\AgdaDatatype{\ensuremath{\mathcal{B}}}\)).
 The cases for \(\bot\), \(\top\), and \(\AgdaDatatype{Bool}\) are simple to adapt: we
@@
   \frac{
-    \mathcal{B}(A) \; \; \; \Pi(x : A) , \mathcal{B}(U(x))
+    \AgdaDatatype{\ensuremath{\mathcal{B}}}(A) \; \; \; \Pi(x : A) , \AgdaDatatype{\ensuremath{\mathcal{B}}}(U(x))
   }{
-    \mathcal{B}(\Pi \; A \; U)
+    \AgdaDatatype{\ensuremath{\mathcal{B}}}(\Pi \; A \; U)
   }
@@
   \frac{
-    \lVert \mathcal{B}(A) \rVert \; \; \; \lVert \Pi(x : A) , \mathcal{B}(U(x)) \rVert
+    \lVert \AgdaDatatype{\ensuremath{\mathcal{B}}}(A) \rVert \; \; \; \lVert \Pi(x : A) , \AgdaDatatype{\ensuremath{\mathcal{B}}}(U(x)) \rVert
   }{
-    \lVert \mathcal{B}(\Pi \; A \; U) \rVert
+    \lVert \AgdaDatatype{\ensuremath{\mathcal{B}}}(\Pi \; A \; U) \rVert
   }
@@
   \frac{
-    \lVert \mathcal{B}(A) \rVert \; \; \; \Pi(x : A) , \lVert \mathcal{B}(U(x)) \rVert
+    \lVert \AgdaDatatype{\ensuremath{\mathcal{B}}}(A) \rVert \; \; \; \Pi(x : A) , \lVert \AgdaDatatype{\ensuremath{\mathcal{B}}}(U(x)) \rVert
   }{
-    \lVert \mathcal{B}(\Pi \; A \; U) \rVert
+    \lVert \AgdaDatatype{\ensuremath{\mathcal{B}}}(\Pi \; A \; U) \rVert
   }
@@
 \begin{equation}
-  (\Pi(x : A) , \lVert \mathcal{B}(U(x)) \rVert) \rightarrow
-  \lVert \Pi(x : A) , \mathcal{B}(U(x)) \rVert
+  (\Pi(x : A) , \lVert \AgdaDatatype{\ensuremath{\mathcal{B}}}(U(x)) \rVert) \rightarrow
+  \lVert \Pi(x : A) , \AgdaDatatype{\ensuremath{\mathcal{B}}}(U(x)) \rVert
 \end{equation}

### 2020-06-22 sections/finiteness-predicates.tex
Update finiteness-predicates.tex

@@
 tell us the number of elements in the finite set?).
+We will use the Countdown problem throughout as a running example of a thing to
+prove finite.
 \section{Split Enumerability} \label{split-enumerability}
@@
   To say that some type \(A\) is split enumerable is to say that there is a list
-  \(\mathit{support} : \mathbf{List}(A)\) such that any value \(x : A\) is in
+  \(\mathit{support} : \AgdaDatatype{List}(A)\) such that any value \(x : A\) is in
   \(\mathit{support}\).
@@
 
-\begin{definition}[\(\mathbf{List}\)] \label{List}
+\begin{definition}[\(\AgdaDatatype{List}\)] \label{List}
   In this paper we will work with two equivalent definitions of lists.
@@
 \(\AgdaDatatype{Fin}\).
-\begin{definition}[\(\mathbf{Fin}\)] \label{Fin}
-  \(\mathbf{Fin}(n)\) is the type of natural numbers smaller than \(n\). We
+\begin{definition}[\(\AgdaDatatype{Fin}\)] \label{Fin}
+  \(\AgdaDatatype{Fin}\;n\) is the type of natural numbers smaller than \(n\). We
   define it the standard way:
@@
   We define disjoint union as an inductive type.
-
   \begin{agdalisting}
@@
   \end{agdalisting}
-
   It is also expressible with only \(\Sigma\):
-
   \begin{agdalisting}
@@
   \end{agdalisting}
-
   Although the inductive type definition is slightly more ergonomic.
@@
   \end{agdalisting}
+\end{definition}
+
+The definition of container is a little abstract: it is instructive to think of
+it more concretely for the case of lists.
+The container representing finite lists is a pair of a natural number \(n\)
+representing the length of the list, and a function \(\AgdaDatatype{Fin}\;n\to
+A\), representing the indexing function into the list.
+
+\begin{definition}[Container Membership] \label{container-membership}
   Membership of a container can be defined like so:
-  \begin{agdalisting} \label{container-membership}
+  \begin{agdalisting} 
     \ExecuteMetaData[agda/Container/Membership.tex]{membership-def}
   \end{agdalisting}
-\end{definition}
-
+  Where \(x\;\AgdaDatatype{\ensuremath{\in}}\;\mathit{xs}\) is to be read as ``\(x\)
+  is in \(\mathit{xs}\)''.
+\end{definition}
 \begin{definition}[Fibers] \label{fibers}
@@
 \end{definition}
+
+Membership also makes more sense when described concretely in terms of lists.
+Understood this way, \(x\;\AgdaDatatype{\ensuremath{\in}}\;\mathit{xs}\) means
+``there is an index into \(\mathit{xs}\) such that the index points at an item
+equal to \(x\)''.
 
@@
   \begin{align}
-    \mathit{support} &: \mathbf{List}(\Sigma(x : A) , U(x)) \\
+    \mathit{support} &: \AgdaDatatype{List}(\Sigma(x : A) , U(x)) \\
     \mathit{cover} &: \Pi(x : \Sigma(y : A) , U(y)) , x \in \mathit{support}
@@
   \begin{equation}%
-    \AgdaDatatype{\ensuremath{\mathcal{E}!}}(A) \simeq \Sigma (n : \mathbb{N}) , \left( \mathbf{Fin}(n) \twoheadrightarrow ! \; A \right)
+    \AgdaDatatype{\ensuremath{\mathcal{E}!}}(A) \simeq \Sigma (n : \mathbb{N}) , \left( \AgdaDatatype{Fin}\;n \twoheadrightarrow ! \; A \right)
   \end{equation}
@@
   By lemma~\ref{split-enum-is-split-surj}, there is a surjection from
-  \(\mathbf{Fin}(n)\) for some \(n\).
-  Also, we know that \(\mathbf{Fin}(n)\) is discrete (proven in our
+  \(\AgdaDatatype{Fin}\;n\) for some \(n\).
+  Also, we know that \(\AgdaDatatype{Fin}\;n\) is discrete (proven in our
   formalisation).
@@
 How exactly we should do this is the next question.
-One approach might be to change the definition of \(\mathbf{List}\), or
+One approach might be to change the definition of \(\AgdaDatatype{List}\), or
 introduce a new type \(\mathbf{NoDupeList}\), and use it in the predicate
@@
 in disguise.
-\emph{equivalence} with \(\mathbf{Fin}\).
+\emph{equivalence} with \(\AgdaDatatype{Fin}\)
 
@@
     \\
-    & \simeq \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket (A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
-    && \text{def.~\ref{List} } (\mathbf{List})
-    \\
-    & \simeq \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \mathbf{Fin}(n)) , A) , \text{isEquiv}(\text{snd}(\mathit{xs}))
+    & \simeq \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \AgdaDatatype{Fin} \rrbracket (A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
+    && \text{def.~\ref{List} } (\AgdaDatatype{List})
+    \\
+    & \simeq \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \AgdaDatatype{Fin}\;n) , A) , \text{isEquiv}(\text{snd}(\mathit{xs}))
     && \text{eqn.~\ref{container-interp} } (\llbracket \wc \rrbracket)
     \\
-    & \simeq \Sigma (n : \mathbb{N}) , \Sigma (f : \mathbf{Fin}(n) \rightarrow A) , \text{isEquiv}(f)
+    & \simeq \Sigma (n : \mathbb{N}) , \Sigma (f : \AgdaDatatype{Fin}\;n \rightarrow A) , \text{isEquiv}(f)
     && \text{Reassociation of } \Sigma
     \\
-    & \simeq \Sigma (n : \mathbb{N}) , ( \mathbf{Fin}(n) \simeq A )
+    & \simeq \Sigma (n : \mathbb{N}) , ( \AgdaDatatype{Fin}\;n \simeq A )
     && \text{eqn.~\ref{equiv-def} } (\simeq) \; \qedhere
@@
   (lemma~\ref{split-enum-to-manifest-bishop}), and consequently we know \(A
-  \simeq \mathbf{Fin}(n)\), for some \(n\) (lemma~\ref{bishop-equiv}).
-  We can therefore replace all occurrences of \(A\) with \(\mathbf{Fin}(n)\),
+  \simeq \AgdaDatatype{Fin}\;n\), for some \(n\) (lemma~\ref{bishop-equiv}).
+  We can therefore replace all occurrences of \(A\) with \(\AgdaDatatype{Fin}\;n\),
   changing our goal to:
@@
     \frac{
-      \AgdaDatatype{\ensuremath{\mathcal{E}!}}(\mathbf{Fin}(n)) \; \; \; \Pi (x : \mathbf{Fin}(n)) , \AgdaDatatype{\ensuremath{\mathcal{E}!}}\left( U(x) \right)
+      \AgdaDatatype{\ensuremath{\mathcal{E}!}}(\AgdaDatatype{Fin}\;n) \; \; \; \Pi (x : \AgdaDatatype{Fin}\;n) , \AgdaDatatype{\ensuremath{\mathcal{E}!}}\left( U(x) \right)
     }{
-      \AgdaDatatype{\ensuremath{\mathcal{E}!}}\left(\Pi (x : \mathbf{Fin}(n)) , U(x)\right)
+      \AgdaDatatype{\ensuremath{\mathcal{E}!}}\left(\Pi (x : \AgdaDatatype{Fin}\;n) , U(x)\right)
     }
@@
   We then define the type of \(n\)-tuples over some type family \(T :
-  \mathbf{Fin}(n) \rightarrow \mathbf{Type}\).
+  \AgdaDatatype{Fin}\;n \rightarrow \mathbf{Type}\).
   \begin{equation}
@@
   \begin{equation}
-    \Pi(x : \mathbf{Fin}(n)) , U(x) \simeq \mathbf{Tuple}(n, U)
+    \Pi(x : \AgdaDatatype{Fin}\;n) , U(x) \simeq \mathbf{Tuple}(n, U)
   \end{equation}
@@
     \frac{
-      \AgdaDatatype{\ensuremath{\mathcal{E}!}}(\mathbf{Fin}(n)) \; \; \; \Pi (x : \mathbf{Fin}(n)) , \AgdaDatatype{\ensuremath{\mathcal{E}!}}\left( U(x) \right)
+      \AgdaDatatype{\ensuremath{\mathcal{E}!}}(\AgdaDatatype{Fin}\;n) \; \; \; \Pi (x : \AgdaDatatype{Fin}\;n) , \AgdaDatatype{\ensuremath{\mathcal{E}!}}\left( U(x) \right)
     }{
@@
   \begin{equation}
-    \mathcal{C}(A) \coloneqq \lVert \AgdaDatatype{\ensuremath{\mathcal{B}}}(A) \rVert \simeq \lVert \Sigma(n : \mathbb{N}) , (\mathbf{Fin}(n) \simeq A) \rVert
+    \mathcal{C}(A) \coloneqq \lVert \AgdaDatatype{\ensuremath{\mathcal{B}}}(A) \rVert \simeq \lVert \Sigma(n : \mathbb{N}) , (\AgdaDatatype{Fin}\;n \simeq A) \rVert
   \end{equation}
@@
   cardinality of \(A\), and a propositionally-truncated proof that \(A\) is
-  equivalent to \(\mathbf{Fin}(n)\).
+  equivalent to \(\AgdaDatatype{Fin}\;n\).
 
@@
   \begin{equation}
-    \frac{(n : \mathbb{N}) \; \; \; (p : \mathbf{Fin}(n) \simeq A) \; \; \;
-          (m : \mathbb{N}) \; \; \; (q : \mathbf{Fin}(m) \simeq A)
+    \frac{(n : \mathbb{N}) \; \; \; (p : \AgdaDatatype{Fin}\;n \simeq A) \; \; \;
+          (m : \mathbb{N}) \; \; \; (q : \AgdaDatatype{Fin}(m) \simeq A)
         }{
@@
     \begin{alignedat}{3}
-      \mathbf{Fin}(n) & \simeq A && (p) \\
-                      & \simeq \mathbf{Fin}(m) && (q)
+      \AgdaDatatype{Fin}\;n & \simeq A && (p) \\
+                      & \simeq \AgdaDatatype{Fin}(m) && (q)
     \end{alignedat}
   \end{equation}
-  Given univalence we have \(\mathbf{Fin}(n) \equiv \mathbf{Fin}(m)\),
+  Given univalence we have \(\AgdaDatatype{Fin}\;n \equiv \AgdaDatatype{Fin}(m)\),
   and the rest of our task is to prove:
   \begin{equation}
-    \frac{\mathbf{Fin}(n) \equiv \mathbf{Fin}(m)}{n \equiv m}
+    \frac{\AgdaDatatype{Fin}\;n \equiv \AgdaDatatype{Fin}(m)}{n \equiv m}
   \end{equation}
@@
 \begin{equation}
-  \Pi(\mathit{xs}, \mathit{ys} : \mathbf{List}(A)) , \mathit{xs} \leftrightsquigarrow \mathit{ys} \rightarrow \text{sort}(\mathit{xs}) \equiv \text{sort}(\mathit{ys})
+  \Pi(\mathit{xs}, \mathit{ys} : \AgdaDatatype{List}(A)) , \mathit{xs} \leftrightsquigarrow \mathit{ys} \rightarrow \text{sort}(\mathit{xs}) \equiv \text{sort}(\mathit{ys})
 \end{equation}
@@
   \begin{equation}
-    \mathcal{E}(A) \coloneqq \Sigma {(\mathit{xs} : \mathbf{List}(A))} , \Pi {(x : A)} , \lVert x \in \mathit{xs} \rVert
+    \mathcal{E}(A) \coloneqq \Sigma {(\mathit{xs} : \AgdaDatatype{List}(A))} , \Pi {(x : A)} , \lVert x \in \mathit{xs} \rVert
   \end{equation}
@@
   \begin{equation}
-    \mathcal{E}(A) \simeq \Sigma(n : \mathbb{N}) , (\mathbf{Fin}(n) \twoheadrightarrow A)
+    \mathcal{E}(A) \simeq \Sigma(n : \mathbb{N}) , (\AgdaDatatype{Fin}\;n \twoheadrightarrow A)
   \end{equation}
@@
     \\
-    & \simeq \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket (A)) , \text{surj}(\text{snd}(\mathit{xs}))
-    && \text{def.~\ref{List} } (\mathbf{List})
-    \\
-    & \simeq \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \mathbf{Fin}(n)) , A) , \text{surj}(\text{snd}(\mathit{xs}))
+    & \simeq \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \AgdaDatatype{Fin} \rrbracket (A)) , \text{surj}(\text{snd}(\mathit{xs}))
+    && \text{def.~\ref{List} } (\AgdaDatatype{List})
+    \\
+    & \simeq \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \AgdaDatatype{Fin}\;n) , A) , \text{surj}(\text{snd}(\mathit{xs}))
     && \text{eqn.~\ref{container-interp} } (\llbracket \wc \rrbracket)
     \\
-    & \simeq \Sigma (n : \mathbb{N}) , \Sigma (f : \mathbf{Fin}(n) \rightarrow A) , \text{surj}(f)
+    & \simeq \Sigma (n : \mathbb{N}) , \Sigma (f : \AgdaDatatype{Fin}\;n \rightarrow A) , \text{surj}(f)
     && \text{Reassociation of } \Sigma
     \\
-    & \simeq \Sigma (n : \mathbb{N}) , ( \mathbf{Fin}(n) \twoheadrightarrow A )
+    & \simeq \Sigma (n : \mathbb{N}) , ( \AgdaDatatype{Fin}\;n \twoheadrightarrow A )
     && \text{eqn.~\ref{surj-arrow-eqn} } (\twoheadrightarrow) \; \qedhere

### 2020-06-23 sections/finiteness-predicates.tex
added one countdown example

@@
     \frac{
-      \AgdaDatatype{\ensuremath{\mathcal{E}!}}(A) \; \; \; \Pi(x : A) , \AgdaDatatype{\ensuremath{\mathcal{E}!}}(U(x))
+      \AgdaDatatype{\ensuremath{\mathcal{E}!}}\;A \; \; \; (x : A) \rightarrow \AgdaDatatype{\ensuremath{\mathcal{E}!}}(U\;x)
     }{
-      \AgdaDatatype{\ensuremath{\mathcal{E}!}}(\Sigma(x : A) , U(x))
+      \AgdaDatatype{\ensuremath{\mathcal{E}!}}(\AgdaDatatype{\ensuremath{\Sigma [}}\;x \AgdaDatatype{:} A \AgdaDatatype{]} U\;x)
     }
@@
 
-\begin{agdalisting}
 This pattern of applying a function to each element in a list and
@@
 in the following extremely clean implementation:
+\begin{agdalisting}
   \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{sup-sigma}
+\end{agdalisting}
+
+We now have two components we'll need for the proof that the countdown
+transformation is finite.
+The component we'll look at is step~\ref{countdown-operators}: selection of the
+operators.
+We'll first need a type representing the operators available to us.
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Countdown.tex]{ops-def}
+\end{agdalisting}
+Proving that this type is finite takes much the same form as the proof of
+finiteness for bool.
+
+Next, we will need to build a proof of finiteness for vectors of length \(n\).
+This uses the proof of finiteness for \(\Sigma\).
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Countdown.tex]{vec-fin}
 \end{agdalisting}

### 2020-06-23 sections/finiteness-predicates.tex
countdown example into the next section

@@
 
-\begin{lemma} \label{split-enum-sigma}
+\begin{lemma} \label{split-enum-sigma} \todo{Convert to Agda}
   Split enumerability is closed under \(\Sigma\).
@@
 finiteness for bool.
+\begin{agdalisting} \label{op-fin}
+  \ExecuteMetaData[agda/Countdown.tex]{op-fin}
+\end{agdalisting}
 
@@
 \end{agdalisting}
-
 \subsection{Derivations}
@@
   We define \emph{split} surjections here \cite[definition
-  4.6.1]{hottbook}.
+  4.6.1]{hottbook}. \todo{To Agda}
   \begin{alignat}{3}
@@
   A proof of split enumerability is equivalent to a split surjection from a
-  finite prefix of the natural numbers.
-  \begin{equation}%
+  finite prefix of the natural numbers. \todo{To Agda}
+  \begin{equation}
     \AgdaDatatype{\ensuremath{\mathcal{E}!}}(A) \simeq \Sigma (n : \mathbb{N}) , \left( \AgdaDatatype{Fin}\;n \twoheadrightarrow ! \; A \right)
@@
 \end{lemma}
-\begin{proof} \phantom{x}
-
-  \begin{minipage}[t]{.8\textwidth}\vspace{-\baselineskip}
+\begin{proof} 
+  \begin{minipage}[t]{.7\textwidth}\vspace{-\baselineskip}
     \begin{agdalisting*}
@@
   \end{minipage}
-  \begin{minipage}[t]{.19\textwidth} \setstretch{1.08}
+  \begin{minipage}[t]{.19\textwidth} \setstretch{1.1}
 
@@
       Reassociation \\
-      eqn.~\ref{sp-surj-arrow-eqn} \qedhere
+      eqn.~\ref{sp-surj-arrow-eqn}
   \end{minipage}
@@
 With split enumerability we can see an instance of this.
-Consider the following proof of the finiteness of Bool:
-\begin{agdalisting} \label{bool-slop}
-  \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{bool-slop}
-\end{agdalisting}
-While it represents the ``same'' information as our previous proof, it is
-clearly not the same \emph{object}.
+Consider the following proof of the finiteness of the operators for countdown:
+\begin{agdalisting} \label{op-slop}
+  \ExecuteMetaData[agda/Countdown.tex]{op-slop}
+\end{agdalisting}
+While it represents the ``same'' information as the proof in
+equation~\ref{op-fin}, it clearly is not the same \emph{object}.
 
@@
 values than there are \emph{usefully} distinct values.
-As we can see in the example above, for instance, split enumerability allows
-duplicate values in the support list.
+For the purposes of solving countdown this has the undesirable effect of
+duplicating search effort, but more generally this reveals that the predicate
+for finiteness we have doesn't represent in a concise way what we intend it to
+represent.
 To reconcile this, we will disallow duplicates in the support list.
@@
 One approach might be to change the definition of \(\AgdaDatatype{List}\), or
-introduce a new type \(\mathbf{NoDupeList}\), and use it in the predicate
+introduce a new type \(\AgdaDatatype{NoDupeList}\), and use it in the predicate
 instead.
@@
 in disguise.
-\emph{equivalence} with \(\AgdaDatatype{Fin}\)
 

### 2020-06-23 sections/finiteness-predicates.tex
brief mentions of countdown inserted

@@
 \end{proof}
+
+This proof can again give us insight into how to prove finiteness of our
+countdown transformation.
+In the first step (Fig.~\ref{countdown-selection}), we need to select some
+numbers from an input list: this can be described with a function of type
+\(\AgdaDatatype{Fin}\;n \rightarrow \AgdaDatatype{Bool}\), from indices in the
+original list into whether we keep the values or not.
+We now know that we can prove functions finite without difficulty: in this case,
+we can do it even more simply by proving that an \(n\)-tuple of booleans is
+finite.
 \section{Cardinal Finiteness} \label{cardinal-finiteness}

### 2020-06-23 sections/introduction.tex
more agda instead of mathematical notation
brief mentions of countdown inserted

@@
 in our solution.
-The representation of this transformation should be fairly simple: to pick from
-\(n\) objects we can have a vector of \(n\) booleans.
+We will need to show there are finitely many ways to filter \(n\) numbers.
 
@@
 
-The third step (Fig.~\ref{countdown-operators}) is quite like the first: it's
-just another vector of length \(n\) of finite objects (in this case operators
+The third step (Fig.~\ref{countdown-operators}) is a vector of length \(n\) of finite objects (in this case operators
 chosen from \(+\), \(\times\), \(-\), and \(\div\)).
 Although it is complicated slightly by the fact that the \(n\) in this
-\(n\)-tuple is dependent on the number of trues in the previously chosen tuple
-of booleans (in terms of types, that means we'll need a \(\Sigma\) rather than a
+\(n\)-tuple is dependent on the amount of numbers we let through in the filter
+in step one.
+(in terms of types, that means we'll need a \(\Sigma\) rather than a
 \(\times\), explanations of which are forthcoming).

### 2020-06-23 sections/search.tex
switched to chapters
filter-subobject

@@
-\section{Search} \label{search}
-\subsection{Omniscience}
+\chapter{Search} \label{search}
+Finiteness proofs often play a role in proof automation in dependently-typed
+programming languages \cite{firsovDependentlyTypedProgramming2015}.
+Our work here is no exception.
+Our work goes slightly further than similar libraries which already exist,
+however: using CuTT we are able to search over functions as well as finite data
+types.
+\section{Countdown}
+\section{Omniscience}
 \begin{definition}[Limited Principle of Omniscience]
@@
 predicates.
-\subsection{Automating Proofs}
+\section{Automating Proofs}
 One use for above constructions is the automation of certain proofs.

### 2020-06-24 sections/search.tex
search introduction

--- 2020-05-29 sections/search.tex
@@
-\section{Search} \label{search}
-\subsection{Omniscience}
+\chapter{Search} \label{search}
+A common theme in dependently-typed programming is that proofs of interesting
+theoretical things may actually correspond to useful algorithms in some way
+related to that thing.
+Finiteness is one such case: if we have a proof that a type \(A\) is finite,
+we should be able to search through all the elements of that type in a
+systematic, automated way.
+
+As it happens, this kind of search is a very common method of proof automation
+in dependently-typed languages like Agda.
+Proofs of statements like ``the following function is associative''
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Snippets/Bool.tex]{and-def}
+\end{agdalisting}
+can be tedious: the associativity proof in particular would take \(2^3 = 8\)
+cases.
+This is unacceptable!
+There are only finitely many cases to examine, after all, and we're
+\emph{already} on a computer: why not automate it?
+A proof that \(\AgdaDatatype{Bool}\) is finite can get us much of the way to a
+library to do just that.
+
+Similar automation machinery can be leveraged to provide search algorithms for
+certain ``logic programming''-esque problems.
+Using the machinery we will describe in this section, though, when the program
+says it finds a solution to some problem that solution will be accompanied by a
+formal \emph{proof} of its correctness.
+
+In this section, we will describe the theoretical underpinning and
+implementation of a library for proof search over finite domains, based on the
+finiteness predicates we have introduced already.
+The library will be able to prove statements like the proof of associativity
+above, as well as more complex statements.
+As a running example for a ``more complex statement'' we will use the countdown
+problem, which we have been using throughout: we will demonstrate how to
+construct a prover for the existence of, or absence of, a solution to a given
+countdown puzzle.
+
+The API for writing searches over finite domains comes from the language of the
+\(\Pi\)-pretopos: with it we will show how to compose QuickCheck-like generators
+for proof search, with the addition of some automation machinery that allows us
+to prove things like the associativity in a couple of lines:
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Snippets/Bool.tex]{bool-assoc-auto-proof}
+\end{agdalisting}
+
+We have already, in previous sections, explored the theoretical implications of
+Cubical Type Theory on our formalisation.
+With this library for proof search, however, we will see two distinct
+\emph{practical} applications which would simply not be possible without
+computational univalence.
+First and foremost: our proofs of finiteness, constructed with the API we will
+describe, have all the power of full equalities.
+Put another way any proof over a finite type \(A\) can be lifted to any other
+type with the same cardinality.
+Secondly our proof search can range over functions: we could, for instance, have
+asked the prover to find if \emph{any} function over \(\AgdaDatatype{Bool}\) is
+associative, and if so return it to us.
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Snippets/Bool.tex]{some-assoc}
+\end{agdalisting}
+The usefulness of which is dubious, but we will see a more interesting
+application soon.
+\section{Omniscience}
 \begin{definition}[Limited Principle of Omniscience]
@@
 predicates.
-\subsection{Automating Proofs}
+
+\section{Countdown}
+\section{Automating Proofs}
 One use for above constructions is the automation of certain proofs.

### 2020-06-25 sections/search.tex
More example on computation for search

@@
 to prove things like the associativity in a couple of lines:
-\begin{agdalisting}
+\begin{agdalisting} \label{bool-assoc-auto-proof}
   \ExecuteMetaData[agda/Snippets/Bool.tex]{bool-assoc-auto-proof}
@@
 application soon.
+\section{Proof Automation And Search Techniques}
+For this prover we will not resort to reflection or similar techniques: all of
+the automation machinery will run in standard Agda.
+To accomplish it we will leverage a common technique which we will describe
+briefly here.
+
+To understand the technique we should first notice that some proof automation
+\emph{already} happens in Agda, like the following:
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Snippets/Bool.tex]{obvious}
+\end{agdalisting}
+While it's not a particularly impressive example, it does nonetheless
+demonstrate a principle we will exploit: closed terms will compute to a normal
+form.
+We do not have to manually apply each computation rule or pattern match in order
+for the equation to type-check.
+
+So our task is to rewrite proof obligations like the one in
+Equation~\ref{bool-assoc-auto-proof} into ones which can reduce completely. 
+As it turns out, we have already described the type of proofs which can ``reduce
+completely'': \emph{decidable} proofs.
+If we have a decision procedure over some proposition \(P\) we can run that
+decision during type checking, because the decision procedure itself is a proof
+that the decision will terminate.
+In code, we capture this idea with the following pair of functions:
+\begin{multicols}{2}
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Snippets/Bool.tex]{is-true}
+  \end{agdalisting} \columnbreak
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Snippets/Bool.tex]{from-true}
+  \end{agdalisting}
+\end{multicols}
+The first is a function which derives a type from whether a decision is
+successful or not.
+This function is important because if we use the output of this type at any
+point we will effectively force the unifier to run the decision computation.
+
 \section{Omniscience}
+First we will explore the theoretical underpinning for search, and its relation
+to finiteness.
+The property we are looking for is effectively the ability to exhaustively
+search through some type.
+Consider the proof above in Equation~\ref{bool-assoc-auto-proof}: we know the
+proof could be automatic because we can check every configuration of the values
+of \(x\), \(y\), and \(z\).
+
 \begin{definition}[Limited Principle of Omniscience]

### 2020-06-25 sections/search.tex
finished bool example

@@
 \section{Proof Automation And Search Techniques}
-For this prover we will not resort to reflection or similar techniques: all of
-the automation machinery will run in standard Agda.
-To accomplish it we will leverage a common technique which we will describe
-briefly here.
+For this prover we will not resort to reflection or similar techniques: instead,
+we will trick the type checker to do our automation for us.
+This is a relatively common technique, although not so much outside of Agda, so
+we will briefly explain it here.
 
@@
 \end{agdalisting}
-While it's not a particularly impressive example, it does nonetheless
+The type checker does not require us to manually explain each step of evaluation
+of:
+\(\AgdaInductiveConstructor{true}\;\AgdaFunction{∧}\;\AgdaInductiveConstructor{false}\).
+While it's not a particularly impressive example of automation, it does nonetheless
 demonstrate a principle we will exploit: closed terms will compute to a normal
-form.
-We do not have to manually apply each computation rule or pattern match in order
-for the equation to type-check.
+form if they're needed to type check.
 
@@
 point we will effectively force the unifier to run the decision computation.
+The second takes---as an implicit argument---an inhabitant of the type generated
+from the first, and uses it to prove that the decision can only be true, and the
+extracts the resulting proof from that decision.
+All in all, we can use it like this:
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Snippets/Bool.tex]{extremely-obvious}
+\end{agdalisting}
+This technique will allow us to automatically compute any decidable predicate.
+\section{Omniscience}
+So we now know what is needed of us for proof automation: we need to take our
+proofs and make them decidable.
 
-\section{Omniscience}
 First we will explore the theoretical underpinning for search, and its relation

### 2020-06-25 sections/search.tex
intro to exhaustibility

@@
 library to do just that.
+\todo[inline]{These examples so far are pretty focused on the bool associativity
+  example.
+  I'm not sure I can think of a good way to put countdown in instead: will we
+  try switch?
+  Or will we keep the bool for this short bit?
+  }
 
@@
 proofs and make them decidable.
+In particular, our strategy will be to take a proposition which is decidable in
+the individual case (like
+\((x\;\AgdaFunction{∧}\;y)\;\AgdaFunction{∧}\;z\;\AgdaFunction{≡}\;x\;\AgdaFunction{∧}\;(y\;\AgdaFunction{∧}\;z)\)
+for a given \(x\), \(y\), and \(z\)) and make it decidable in the universal case.
+Whether this is possible or not is a property of the \emph{domain} of a
+decidable predicate, and is described formally by the notion of ``Exhaustibility''.
+\begin{definition}[Exhaustibility]
+  We say a type \(A\) is exhaustible if, for any decidable predicate \(P\) on
+  \(A\), the universal quantification of the predicate is decidable.
+  \begin{equation}
+    \left( \Pi {(x : A)} , \mathbf{Dec}(P(x)) \right) \rightarrow \mathbf{Dec} \left( \Pi {(x : A)} , P(x) \right)
+  \end{equation}
+\end{definition}
+Perhaps surprisingly, it is not only finite types which can are exhaustible.
+Certain infinite types can be
+exhaustible~\cite{escardoInfiniteSetsThat2007}.
 
@@
 There is also a universal form of omniscience, which we call exhaustibility.
-\begin{definition}[Exhaustibility]
-  We say a type \(A\) is exhaustible if, for any decidable predicate \(P\) on
-  \(A\), the universal quantification of the predicate is decidable.
-  \begin{equation}
-    \left( \Pi {(x : A)} , \mathbf{Dec}(P(x)) \right) \rightarrow \mathbf{Dec} \left( \Pi {(x : A)} , P(x) \right)
-  \end{equation}
-\end{definition}
 

### 2020-06-25 sections/introduction.tex

@@
 \chapter{Introduction}
+\todo{regroup into old structure}
 We are interested in constructive notions of finiteness, formalised in Cubical
@@
 \section{The Varieties of Finiteness}
+\todo{Make all references parenthetical}
 In Section~\ref{finiteness-predicates} we will explore a number of different
@@
 implications of Homotopy Type Theory, first proven
-in~\cite{rijkeSetsHomotopyType2015}.
+in~\cite{rijkeSetsHomotopyType2015}. \todo{This reference should be citet not citep}
 Our work is a formalisation of this result (and the first such formalisation

### 2020-06-25 sections/search.tex
more on omniscience

@@
 \end{definition}
-Perhaps surprisingly, it is not only finite types which can are exhaustible.
-Certain infinite types can be
-exhaustible~\cite{escardoInfiniteSetsThat2007}.
-
-First we will explore the theoretical underpinning for search, and its relation
-to finiteness.
-The property we are looking for is effectively the ability to exhaustively
-search through some type.
-Consider the proof above in Equation~\ref{bool-assoc-auto-proof}: we know the
-proof could be automatic because we can check every configuration of the values
-of \(x\), \(y\), and \(z\).
-
+This property of \AgdaDatatype{Bool} would allow us to automate the proof of
+associativity, but it is in fact not strong enough to find individual
+representatives of a type which support some property.
+For that we need the more well-known, but related, property of
+\emph{omniscience}.
 \begin{definition}[Limited Principle of Omniscience]
@@
 \end{definition}
-The limited principle of omniscience is non-constructive, but individual types
-can themselves satisfy omniscience.
-In particular, cardinal finite types are omniscient.
 
-There is also a universal form of omniscience, which we call exhaustibility.
+Both of these properties only make sense to talk about in a constructive
+setting: classically speaking, \emph{every} type is omniscient and exhaustible.
+Finite types are omniscient (the reason should be clear, but we will prove this
+fact presently), but perhaps surprisingly, it is not \emph{only} finite types
+which are exhaustible.
+Certain infinite types can be
+exhaustible~\cite{escardoInfiniteSetsThat2007}, but an exploration of that is
+beyond the scope of this work.
 
-All of the finiteness predicates we have seen imply exhaustibility.
-Omniscience is stronger than exhaustibility, as we can derive the latter from
-the former.
-All of the ordered finiteness predicates imply omniscience.
-For the unordered finiteness definitions, we have omniscience for prop-valued
-predicates.
+Omniscience and exhaustibility are not interchangeable: conceptually, for
+omniscience you need some kind of ordering on the underlying type.
+If we had a property, say like the associativity property above, that is
+satisfied by more than one element of the type, we would have to have some way
+to pick a candidate element from the type to return in the case of the property
+being satisfied.
+This requirement of order mirrors the Ord difference in
+Fig.~\ref{finite-classification}.
+
+Omniscience implies exhaustibility, then, but the converse is not true.
+\begin{lemma}
+  Omniscience implies exhaustibility
+\end{lemma}
+\begin{proof}
+  \todo{proof here}
+\end{proof}
+
+And the relation to the finiteness predicates is straightforward: all of the
+finiteness predicates we have seen imply exhaustibility, and all of the ordered
+finiteness predicates imply omniscience.
+We can prove this by showing exhaustibility and omniscience for the weakest
+candidate of finiteness predicates.
+\begin{lemma}
+  Kuratowski finiteness implies exhaustibility
+\end{lemma}
+\begin{lemma}
+  Manifest enumerability implies omniscience
+\end{lemma}
+
+Finally, we can get around the order requirement for prop-valued predicates for
+omniscience.
+\begin{lemma}
+  Omniscience and exhaustibility coincide for prop-valued predicates.
+\end{lemma}
 

### 2020-06-25 sections/search.tex
Update search.tex
agda omniscient definitions

@@
   \(A\), the universal quantification of the predicate is decidable.
-  \begin{equation}
-    \left( \Pi {(x : A)} , \mathbf{Dec}(P(x)) \right) \rightarrow \mathbf{Dec} \left( \Pi {(x : A)} , P(x) \right)
-  \end{equation}
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Relation/Nullary/Omniscience.tex]{exhaustible}
+  \end{agdalisting}
 \end{definition}
@@
   omniscience \cite{myhillErrettBishopFoundations1972} is as follows:
-  \begin{equation}
-    \left( \Pi {(x : A)} , \mathbf{Dec}(P(x)) \right) \rightarrow \mathbf{Dec} \left( \Sigma {(x : A)} , P(x) \right)
-  \end{equation}
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Relation/Nullary/Omniscience.tex]{omniscient}
+  \end{agdalisting}
   In other words, for any decidable predicate the existential quantification of
@@
 \end{definition}
-
 Both of these properties only make sense to talk about in a constructive
@@
 
-Omniscience and exhaustibility are not interchangeable: conceptually, for
-omniscience you need some kind of ordering on the underlying type.
+Omniscience and exhaustibility are not interchangeable: conceptually,
+omniscience needs some kind of ordering on the underlying type.
 If we had a property, say like the associativity property above, that is

### 2020-06-25 sections/search.tex
more on exhaustible

@@
 proofs and make them decidable.
-In particular, our strategy will be to take a proposition which is decidable in
-the individual case (like
-\((x\;\AgdaFunction{∧}\;y)\;\AgdaFunction{∧}\;z\;\AgdaFunction{≡}\;x\;\AgdaFunction{∧}\;(y\;\AgdaFunction{∧}\;z)\)
-for a given \(x\), \(y\), and \(z\)) and make it decidable in the universal case.
-Whether this is possible or not is a property of the \emph{domain} of a
-decidable predicate, and is described formally by the notion of ``Exhaustibility''.
+In particular, we need to be able to ``lift'' decidability back over a
+function arrow.
+For instance, given \(x\), \(y\), and \(z\) we already have
+\(\AgdaDatatype{Dec}\;((x\;\AgdaFunction{∧}\;y)\;\AgdaFunction{∧}\;z\;\AgdaFunction{≡}\;x\;\AgdaFunction{∧}\;(y\;\AgdaFunction{∧}\;z))\)
+(because equality over booleans is decidable).
+In order to turn this into a proof that \AgdaFunction{∧} is associative we need
+\(\AgdaDatatype{Dec}\;(\forall \; x \; y \; z \rightarrow (x\;\AgdaFunction{∧}\;y)\;\AgdaFunction{∧}\;z\;\AgdaFunction{≡}\;x\;\AgdaFunction{∧}\;(y\;\AgdaFunction{∧}\;z))\).
+The ability to do this is described formally by the notion of
+``Exhaustibility''.
 \begin{definition}[Exhaustibility]

### 2020-06-25 sections/search.tex
Update search.tex

@@
 With this library for proof search, however, we will see two distinct
-\emph{practical} applications which would simply not be possible without
+practical applications which would simply not be possible without
 computational univalence.
@@
 \end{definition}
-Both of these properties only make sense to talk about in a constructive
-setting: classically speaking, \emph{every} type is omniscient and exhaustible.
-Finite types are omniscient (the reason should be clear, but we will prove this
-fact presently), but perhaps surprisingly, it is not \emph{only} finite types
+
+These properties are closely related to the axiom of choice: in particular, the
+statement that ``every type is omniscient'' implies the AOC. \todo{Reference for
+this, and possible proof? Also it might not be entirely correct: I think you
+need a truncation somewhere for omniscience to be equivalent to the AOC}
+Because we're constructive, only a select few types are omniscient: finite
+types, for instance (why finite types are omniscient is not too difficult to
+see, but we will explain it soon).
+Perhaps surprisingly, it is not \emph{only} finite types
 which are exhaustible.
@@
 
-Omniscience and exhaustibility are not interchangeable: conceptually,
-omniscience needs some kind of ordering on the underlying type.
-If we had a property, say like the associativity property above, that is
-satisfied by more than one element of the type, we would have to have some way
-to pick a candidate element from the type to return in the case of the property
-being satisfied.
-This requirement of order mirrors the Ord difference in
-Fig.~\ref{finite-classification}.
-
-Omniscience implies exhaustibility, then, but the converse is not true.
+Omniscience and exhaustibility are not interchangeable: every omniscient type is
+exhaustible, but the converse is not true.
+Conceptually, omniscience needs some kind of ordering on the underlying type.
+This is because omniscience returns a candidate satisfying the given predicate:
+there is no requirement, though, that only one element in the underlying type
+satisfies the decidable predicate.
+As a result, omniscience needs some way to choose among all possible candidates:
+this is the ``order'' we are referring to.
+This is also the same ``order'' that we referred to when talking about the
+finiteness predicates: all ordered predicates (in
+Figure~\ref{finite-classification}) imply omniscience, whereas the unordered
+predicates only imply exhaustibility.
 \begin{lemma}

### 2020-06-25 sections/search.tex
Update search.tex

@@
 
-These properties are closely related to the axiom of choice: in particular, the
-statement that ``every type is omniscient'' implies the AOC. \todo{Reference for
-this, and possible proof? Also it might not be entirely correct: I think you
-need a truncation somewhere for omniscience to be equivalent to the AOC}
 Because we're constructive, only a select few types are omniscient: finite
-types, for instance (why finite types are omniscient is not too difficult to
-see, but we will explain it soon).
-Perhaps surprisingly, it is not \emph{only} finite types
-which are exhaustible.
+types, for instance (the law of the excluded middle implies that all types are
+omniscient, meaning that all types are omniscient classically).
+Perhaps surprisingly, it is not \emph{only} finite types which are exhaustible.
 Certain infinite types can be
@@
 
+\begin{lemma}
+  If exhaustibility implies omniscience, then the axiom of choice holds
+\end{lemma}
+\begin{proof}
+  \todo{proof here}
+\end{proof}
+
 And the relation to the finiteness predicates is straightforward: all of the
@@
 \end{lemma}
-
 \section{Countdown}
+We have already introduced and described countdown: in this section, we will
+fill in the remaining parts of the solver, glue the pieces together, and show
+how the finiteness proofs can assist us to write the solver.
+\subsection{Finite Permutations}
+\subsection{Parenthesising}
+\subsection{Filtering}
+\subsection{Putting It All Together}
 \section{Automating Proofs}

### 2020-06-26 sections/search.tex
first part of permutation in countdown explanation

@@
 \subsection{Finite Permutations}
+The first step of the transformation we will represent as a finite object is a
+\emph{permutation}: once we choose the numbers we're going to use for the
+candidate for the solution, we have to order them in some way.
+
+Our first attempt at representing permutations might look something like this:
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Countdown.tex]{wrong-perm}
+\end{agdalisting}
+the idea is that \(\AgdaDatatype{Perm}\;n\) represents a permutation of \(n\)
+things, as a function from positions to positions.
+Unfortunately such a simple answer won't work: there are no restrictions on the
+operation of the function, so it could (for instance), send more than one input
+position into the same output.
+
+What we actually need is not just a function between positions, but an
+\emph{isomorphism} between them.
+In types:
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Countdown.tex]{iso-perm}
+\end{agdalisting}
+Where an isomorphism is defined as follows:
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Countdown.tex]{isomorphism}
+\end{agdalisting}
+\todo{Should this isomorphism definition be put earlier in the intro with the
+  equivalences etc?}
+While it may look complex, this term is actually composed of individual
+components we've already proven finite.
+First we have \(\AgdaDatatype{Fin}\;n\rightarrow\AgdaDatatype{Fin}\;n\):
+functions between finite types are, as we know, finite
+(Theorem~\ref{split-enum-pi}).
+We take a pair of them: pairs of finite things are \emph{also} finite
+(Lemma~\ref{split-enum-sigma}).
+To get the next two components we can filter to the subobject: this requires
+these predicates to be decidable.
+We will construct a term of the following type:
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Cardinality/Finite/ManifestBishop.tex]{dec-type}
+\end{agdalisting}
+So can we construct such a term? Yes!
+
+We basically need to construct decidable equality for functions between
+\(\AgdaDatatype{Fin}\;n\)s: of course, this decidable equality is provided by
+the fact that such functions are themselves finite.
+
+All in all we can now prove that the isomorphism, and by extension the
+permutation, is finite:
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Cardinality/Finite/ManifestBishop.tex]{iso-finite}
+\end{agdalisting}
+
 \subsection{Parenthesising}
-\subsection{Filtering}
 \subsection{Putting It All Together}

### 2020-06-26 sections/search.tex
finished permutations section

@@
 To get the next two components we can filter to the subobject: this requires
-these predicates to be decidable.
+these predicates to be decidable. \todo{Need to do filter subobject in topos section}
 We will construct a term of the following type:
@@
 
+Unfortunately this implementation is too slow to be useful.
+As nice and declarative as it is, fundamentally it builds a list of all possible
+pairs of functions between \(\AgdaDatatype{Fin}\;n\) and itself (an operation
+which takes in the neighbourhood of \(\mathcal{O}(n^n)\) time), and then tests
+each for equality (which is likely worse than \(\mathcal{O}(n^2)\) time).
+We will instead use a factoriadic encoding: this is a relatively simple encoding
+of permutations which will reduce our time to a blazing fast
+\(\mathcal{O}(n!)\).
+It is expressed in Agda as follows:
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Countdown.tex]{perm-def}
+\end{agdalisting}
+It is a vector of positions, each represented with a \(\AgdaDatatype{Fin}\).
+Each position can only refer to the length of the tail of the list at that
+point: this prevents two input positions mapping to the same output point, which
+was the problem with the naive encoding we had previously.
+And it also has a relatively simple proof of finiteness:
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Countdown.tex]{perm-fin}
+\end{agdalisting}
 \subsection{Parenthesising}
+Our next step is figuring out a way to encode the parenthesisation
+\subsection{Filtering}
 \subsection{Putting It All Together}

### 2020-06-26 sections/search.tex
finished intro to dyck & countdown

@@
 \subsection{Parenthesising}
-Our next step is figuring out a way to encode the parenthesisation
-\subsection{Filtering}
+Our next step is figuring out a way to encode the parenthesisation of the
+expression (Fig.~\ref{countdown-parens}). \todo{There's no way
+  ``parenthesisation'' is a real word}
+At this point of the transformation, we already have our numbers picked out, we
+have ordered them a certain way, and we have also selected the operators we're
+interested in.
+We have, in other words, the following:
+\begin{equation}
+  3 \times 7 \times 50 - 10 - 25
+\end{equation}
+Without parentheses, however, (or a religious adherence to BOMDAS) this
+expression is still ambiguous.
+\begin{align}
+  3 \times ((7 \times (50 - 10)) - 25) &= 765 \\
+  (((3 \times 7) \times 50) - 10) - 25 &= 1015
+\end{align}
+The different ways to parenthesise the expression result in different outputs
+of evaluation.
+
+So what data type encapsulates the ``different ways to parenthesise'' a given
+expression?
+That's what we will figure out in this section, and what we will prove finite.
+
+One way to approach the problem is with a binary tree.
+A binary tree with \(n\) leaves corresponds in a straightforward way to a
+parenthesisation of \(n\) numbers (Fig.~\ref{countdown-parens}).
+\todo{Tree diagram? Or link to previous tree?}
+This doesn't get us much closer to a finiteness proof, however: for that we will
+need to rely on \emph{Dyck} words.
+\begin{definition}[Dyck words]
+  A Dyck word is a string of balanced parentheses.
+  In Agda, we can express it as the following:
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Dyck.tex]{dyck-def}
+  \end{agdalisting}
+  A fully balanced string of \(n\) parentheses has the type
+  \(\AgdaDatatype{Dyck}\;\AgdaInductiveConstructor{zero}\;n\).
+  Here are some example strings:
+  \begin{multicols}{2}
+    \begin{agdalisting}
+      \ExecuteMetaData[agda/Dyck.tex]{dyck-0-2}
+    \end{agdalisting}
+    \begin{agdalisting}
+      \ExecuteMetaData[agda/Dyck.tex]{dyck-0-3}
+    \end{agdalisting}
+  \end{multicols}
+  The first parameter on the type represents the amount of unbalanced closing
+  parens, for instance:
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Dyck.tex]{dyck-1-2}
+  \end{agdalisting}
+\end{definition}
+
+Already Dyck words look easier to prove finite than straight binary trees, but
+for that proof to be useful we'll have to relate Dyck words and binary trees
+somehow.
+As it happens, Dyck words of length \(2n\) are isomorphic to binary trees with
+\(n-1\) leaves, but we only need to show this relation in one direction: from
+Dyck to binary tree.
+To demonstrate the algorithm we'll use a simple tree definition:
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Dyck.tex]{tree-simpl-def}
+\end{agdalisting}
+The algorithm itself is quite similar to stack-based parsing algorithms.
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Dyck.tex]{from-dyck}
+\end{agdalisting}
+\subsection{Filtering to Correct Expressions}
 \subsection{Putting It All Together}

### 2020-06-26 sections/finiteness-predicates.tex
moved closure proofs into topos section

--- 2020-06-23 sections/finiteness-predicates.tex
@@
 equal to \(x\)''.
-
-\subsection{Instances}
-Now that we have a suitable definition of finiteness, we will next prove that
-some things are finite.
-\begin{lemma}
-  \(\bot\), \(\top\), and \(\AgdaDatatype{Bool}\) are split enumerable.
-\end{lemma}
-\begin{proof}
-  These three types are quite obviously finite: we will show only the proof of
-  finiteness for \(\AgdaDatatype{Bool}\) here for brevity's sake.
-  \begin{agdalisting}
-    \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{bool-inst}
-  \end{agdalisting}
-\end{proof}
-With the most basic simple types out of the way, the obvious next choice is the
-(non-dependent) sums and products: \(\uplus\) and \(\times\).
-Both of these types can be constructed from the \emph{dependent} sum, however,
-so that is the type we will prove finite.
-From that we can derive a much wider array of finiteness proofs.
-
-\begin{lemma} \label{split-enum-sigma} \todo{Convert to Agda}
-  Split enumerability is closed under \(\Sigma\).
-  \begin{equation}
-    \frac{
-      \AgdaDatatype{\ensuremath{\mathcal{E}!}}\;A \; \; \; (x : A) \rightarrow \AgdaDatatype{\ensuremath{\mathcal{E}!}}(U\;x)
-    }{
-      \AgdaDatatype{\ensuremath{\mathcal{E}!}}(\AgdaDatatype{\ensuremath{\Sigma [}}\;x \AgdaDatatype{:} A \AgdaDatatype{]} U\;x)
-    }
-  \end{equation}
-\end{lemma}
-\begin{proof}
-  Let \(A\) be a type which is split enumerable, and \(U\) be a type family over
-  \(A\) which is split enumerable at every point.
-  Formally, we have the following proofs:
-  \begin{align}
-    \AgdaDatatype{\ensuremath{\mathcal{E}!}}_A &: \AgdaDatatype{\ensuremath{\mathcal{E}!}}(A) \\
-    \AgdaDatatype{\ensuremath{\mathcal{E}!}}_U &: \Pi(x : A) , \AgdaDatatype{\ensuremath{\mathcal{E}!}}(U(x))
-  \end{align}
-
-  Our task is to construct a proof of type:
-  \begin{equation}
-    \AgdaDatatype{\ensuremath{\mathcal{E}!}}(\Sigma(x : A) , U(x))
-  \end{equation}
-  This proof itself is composed of two components:
-  \begin{align}
-    \mathit{support} &: \AgdaDatatype{List}(\Sigma(x : A) , U(x)) \\
-    \mathit{cover} &: \Pi(x : \Sigma(y : A) , U(y)) , x \in \mathit{support}
-  \end{align}
-  To construct the support list, we apply the function \(\AgdaDatatype{\ensuremath{\mathcal{E}!}}_U\) to
-  every element in the support list of \(\AgdaDatatype{\ensuremath{\mathcal{E}!}}_A\), extract the support
-  lists from the resulting finiteness proofs, and concatenate them.
-
-  To prove that this support list does in fact cover the entirety of the type
-  \(\Sigma \; A \; U\), we note that any element of type \(\Sigma \; A \; U\)
-  must have a first component in the support list of \(\AgdaDatatype{\ensuremath{\mathcal{E}!}}_A\), and its
-  second component must be in the result of applying \(\AgdaDatatype{\ensuremath{\mathcal{E}!}}_U\) to that
-  first element (since that support list contains every element of type
-  \(U(x)\)).
-  Therefore, the pair itself must be in our constructed support list.
-\end{proof}
-
-This pattern of applying a function to each element in a list and
-concatenating the result is of course well-known in functional programming,
-and is in fact the pattern that makes lists a monad.
-While this insight isn't strictly relevant to our work here, it does mean
-the implementation of this function can use Agda's do notation, resulting
-in the following extremely clean implementation:
-\begin{agdalisting}
-  \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{sup-sigma}
-\end{agdalisting}
-
-We now have two components we'll need for the proof that the countdown
-transformation is finite.
-The component we'll look at is step~\ref{countdown-operators}: selection of the
-operators.
-We'll first need a type representing the operators available to us.
-\begin{agdalisting}
-  \ExecuteMetaData[agda/Countdown.tex]{ops-def}
-\end{agdalisting}
-Proving that this type is finite takes much the same form as the proof of
-finiteness for bool.
-\begin{agdalisting} \label{op-fin}
-  \ExecuteMetaData[agda/Countdown.tex]{op-fin}
-\end{agdalisting}
-
-Next, we will need to build a proof of finiteness for vectors of length \(n\).
-This uses the proof of finiteness for \(\Sigma\).
-\begin{agdalisting}
-  \ExecuteMetaData[agda/Countdown.tex]{vec-fin}
-\end{agdalisting}
+\todo{Probably should put back in some instances here, i.e. bool and top.
+  Obviously not the full closure stuff, but one or two instances would be
+  instructive.}
 \subsection{Derivations}
@@
 \end{lemma}
+\todo{Use standard equational reasoning syntax}
 \begin{proof}
@@
 This is extremely powerful, as we will see.
-\subsection{Closure Under \(\Pi\)}
-The glaring omission from our closure proofs under type formers so far has been
-the \(\Pi\) type: we have not proved closure under functions, dependent or
-otherwise.
-In MLTT, this is of course not provable: since all of the finiteness predicates
-we have seen so far imply decidable equality, and since we don't have any kind
-of decidable equality on functions in MLTT, we know that we won't be able to
-show that any kind of function is finite; even one like \(\AgdaDatatype{Bool}
-\rightarrow \AgdaDatatype{Bool}\).
-
-CuTT is not so restricted.
-Since we have things like function extensionality and transport, we can indeed
-prove the finiteness of function types.
-Our proof here makes use directly of the univalence axiom, and makes use
-furthermore of all the previous closure proofs.
-We will prove this closure on split enumerability, rather than on manifest
-Bishop finiteness, as it requires slightly less legwork in the proof itself, but
-of course we can derive the proof on manifest Bishop finiteness in a few lines.
-\begin{theorem} \label{split-enum-pi}
-  Split enumerability is closed under dependent functions.
-  (\(\Pi\)-types).
-  \begin{equation}
-    \frac{
-      \AgdaDatatype{\ensuremath{\mathcal{E}!}}(A) \; \; \; \Pi {(x : A)} , \AgdaDatatype{\ensuremath{\mathcal{E}!}}\left( U(x) \right)
-    }{
-      \AgdaDatatype{\ensuremath{\mathcal{E}!}}\left(\Pi {(x : A)} , U(x)\right)
-    }
-  \end{equation}
-\end{theorem}
-\begin{proof}
-  Let \(A\) be a split enumerable type, and \(U\) be a type family from \(A\),
-  which is split enumerable over all points of \(A\).
-
-  As \(A\) is split enumerable, we know that it is also manifestly Bishop finite
-  (lemma~\ref{split-enum-to-manifest-bishop}), and consequently we know \(A
-  \simeq \AgdaDatatype{Fin}\;n\), for some \(n\) (lemma~\ref{bishop-equiv}).
-  We can therefore replace all occurrences of \(A\) with \(\AgdaDatatype{Fin}\;n\),
-  changing our goal to:
-  \begin{equation}
-    \frac{
-      \AgdaDatatype{\ensuremath{\mathcal{E}!}}(\AgdaDatatype{Fin}\;n) \; \; \; \Pi (x : \AgdaDatatype{Fin}\;n) , \AgdaDatatype{\ensuremath{\mathcal{E}!}}\left( U(x) \right)
-    }{
-      \AgdaDatatype{\ensuremath{\mathcal{E}!}}\left(\Pi (x : \AgdaDatatype{Fin}\;n) , U(x)\right)
-    }
-  \end{equation}
-
-  We then define the type of \(n\)-tuples over some type family \(T :
-  \AgdaDatatype{Fin}\;n \rightarrow \mathbf{Type}\).
-  \begin{equation}
-    \begin{alignedat}{3}
-      & \mathbf{Tuple}(0, T)   &&\coloneqq \top \\
-      & \mathbf{Tuple}(n+1, T) &&\coloneqq T(0) \times \mathbf{Tuple}(n, T \circ \text{suc})
-    \end{alignedat}
-  \end{equation}
-  We can show that this type is equivalent to functions (proven in our formalisation):
-  \begin{equation}
-    \Pi(x : \AgdaDatatype{Fin}\;n) , U(x) \simeq \mathbf{Tuple}(n, U)
-  \end{equation}
-  And therefore we can simplify again our goal to the following:
-  \begin{equation}
-    \frac{
-      \AgdaDatatype{\ensuremath{\mathcal{E}!}}(\AgdaDatatype{Fin}\;n) \; \; \; \Pi (x : \AgdaDatatype{Fin}\;n) , \AgdaDatatype{\ensuremath{\mathcal{E}!}}\left( U(x) \right)
-    }{
-      \AgdaDatatype{\ensuremath{\mathcal{E}!}}\left(\mathbf{Tuple}(n, U)\right)
-    }
-  \end{equation}
-
-  We can prove this goal by showing that \(\mathbf{Tuple}(n, U)\) is split
-  enumerable: it is made up of finitely many products of points of \(U\), which
-  are themselves split enumerable, and \(\top\), which is also split enumerable.
-  Lemma~\ref{split-enum-sigma} shows us that the product of finitely many split
-  enumerable types is itself split enumerable, proving our goal.
-\end{proof}
 \section{Cardinal Finiteness} \label{cardinal-finiteness}
@@
 The cover proof emerges naturally from the definition of the permutation.
-\subsection{Closure}
-Since we don't have a function of type \(\mathcal{C}(A) \rightarrow
-\AgdaDatatype{\ensuremath{\mathcal{B}}}(A)\), closure proofs on \(\AgdaDatatype{\ensuremath{\mathcal{B}}}\) do not transfer over to
-\(\mathcal{C}\) trivially (unlike with \(\AgdaDatatype{\ensuremath{\mathcal{E}!}}\) and \(\AgdaDatatype{\ensuremath{\mathcal{B}}}\)).
-The cases for \(\bot\), \(\top\), and \(\AgdaDatatype{Bool}\) are simple to adapt: we
-can just propositionally truncate their Bishop finiteness proof.
-
-Non-dependent operators like \(\times\), \(\uplus\), and \(\rightarrow\) are
-also relatively straightforward: since \(\lVert {\wc} \rVert\) forms a monad, we
-can apply \(n\)-ary functions to values inside it, combining them together.
-\begin{agdalisting}
-  The fact that \(\lVert \wc \rVert\) forms a monad means that we can lift
-  \(n\)-ary functions like the following:
-  \ExecuteMetaData[agda/Cardinality/Finite/ManifestBishop.tex]{times-clos-sig}
-  Into a truncated context:
-  \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{times-clos-impl}
-\end{agdalisting}
-
-Unfortunately, for the dependent type formers like \(\Sigma\) and \(\Pi\), the
-same trick does not work.
-We have closure proofs like:
-\begin{equation}
-  \frac{
-    \AgdaDatatype{\ensuremath{\mathcal{B}}}(A) \; \; \; \Pi(x : A) , \AgdaDatatype{\ensuremath{\mathcal{B}}}(U(x))
-  }{
-    \AgdaDatatype{\ensuremath{\mathcal{B}}}(\Pi \; A \; U)
-  }
-\end{equation}
-If we apply the monadic truncation trick we can derive closure proofs like the
-following:
-\begin{equation}
-  \frac{
-    \lVert \AgdaDatatype{\ensuremath{\mathcal{B}}}(A) \rVert \; \; \; \lVert \Pi(x : A) , \AgdaDatatype{\ensuremath{\mathcal{B}}}(U(x)) \rVert
-  }{
-    \lVert \AgdaDatatype{\ensuremath{\mathcal{B}}}(\Pi \; A \; U) \rVert
-  }
-\end{equation}
-However our \emph{desired} closure proof is the following:
-\begin{equation}
-  \frac{
-    \lVert \AgdaDatatype{\ensuremath{\mathcal{B}}}(A) \rVert \; \; \; \Pi(x : A) , \lVert \AgdaDatatype{\ensuremath{\mathcal{B}}}(U(x)) \rVert
-  }{
-    \lVert \AgdaDatatype{\ensuremath{\mathcal{B}}}(\Pi \; A \; U) \rVert
-  }
-\end{equation}
-They don't match!
-
-The solution would be to find a function of the following type:
-\begin{equation}
-  (\Pi(x : A) , \lVert \AgdaDatatype{\ensuremath{\mathcal{B}}}(U(x)) \rVert) \rightarrow
-  \lVert \Pi(x : A) , \AgdaDatatype{\ensuremath{\mathcal{B}}}(U(x)) \rVert
-\end{equation}
-However we might be disheartened at realising that this is a required goal: the
-above equation is \emph{extremely} similar to the axiom of choice!
-\begin{definition}[Axiom of Choice]
-  In HoTT, the axiom of choice is commonly defined as follows \cite[lemma
-  3.8.2]{hottbook}.
-  For any set \(A\), and a type family \(U\) which is a set at all the points
-  of \(A\), the following function exists:
-  \begin{equation}
-    \left( \Pi(x : A) ,  \lVert U(x) \rVert \right) \rightarrow \lVert \Pi(x : A) , U(x) \rVert
-  \end{equation}
-\end{definition}
-Luckily the axiom of choice \emph{does} hold for cardinally finite types,
-allowing us to prove the following:
-\begin{lemma}
-  \begin{equation}
-    \mathcal{C}(A) \rightarrow (\Pi(x : A) , \lVert U(x) \rVert) \rightarrow \lVert \Pi(x : A) , U(x) \rVert
-  \end{equation}
-\end{lemma}
-\begin{proof}
-  Let \(A\) be a cardinally finite type, \(U\) be a type family on \(A\), and
-  \(f\) be a dependent function of type \(\Pi(x : A) , \lVert U(x) \rVert\).
-
-  First, since our goal is itself propositionally truncated, we have access to
-  values under truncations: put another way, in the context of proving our goal,
-  we can rely on the fact that \(A\) is manifestly Bishop finite.
-  Using the same technique as we did in lemma~\ref{split-enum-pi}, we can switch
-  from working with dependent functions from \(A\) to \(n\)-tuples, where \(n\)
-  is the cardinality of \(A\).
-  This changes our goal to the following:
-  \begin{equation}
-    \mathbf{Tuple}(n, \lVert \wc \rVert \circ U) \rightarrow \lVert \mathbf{Tuple}(n, U) \rVert
-  \end{equation}
-  Since \(\lVert \wc \rVert\) is closed under finite products, this function
-  exists (in fact, using the fact that \(\lVert \wc \rVert\) forms a monad, we
-  can recognise this function as \verb+sequenceA+ from the \verb+Traversable+
-  class in Haskell).
-\end{proof}
-
-This gets us all of the necessary closure proofs on \(\mathcal{C}\).
 \section{Manifest Enumerability} \label{manifest-enumerability}

### 2020-06-26 sections/topos.tex
filter-subobject
moved closure proofs into topos section

@@
 \section{Categories in HoTT}
-\section{Closure and the Category of Sets}
+\section{Closure}
+\todo{This is just the closure bits pulled over from the finiteness predicates
+  section. They need to be reworked to fit in here.}
+\subsection{Split Enumerability Closure}
+Now that we have a suitable definition of finiteness, we will next prove that
+some things are finite.
+\begin{lemma}
+  \(\bot\), \(\top\), and \(\AgdaDatatype{Bool}\) are split enumerable.
+\end{lemma}
+\begin{proof}
+  These three types are quite obviously finite: we will show only the proof of
+  finiteness for \(\AgdaDatatype{Bool}\) here for brevity's sake.
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{bool-inst}
+  \end{agdalisting}
+\end{proof}
+With the most basic simple types out of the way, the obvious next choice is the
+(non-dependent) sums and products: \(\uplus\) and \(\times\).
+Both of these types can be constructed from the \emph{dependent} sum, however,
+so that is the type we will prove finite.
+From that we can derive a much wider array of finiteness proofs.
+
+\begin{lemma} \label{split-enum-sigma} \todo{Convert to Agda}
+  Split enumerability is closed under \(\Sigma\).
+  \begin{equation}
+    \frac{
+      \AgdaDatatype{\ensuremath{\mathcal{E}!}}\;A \; \; \; (x : A) \rightarrow \AgdaDatatype{\ensuremath{\mathcal{E}!}}(U\;x)
+    }{
+      \AgdaDatatype{\ensuremath{\mathcal{E}!}}(\AgdaDatatype{\ensuremath{\Sigma [}}\;x \AgdaDatatype{:} A \AgdaDatatype{]} U\;x)
+    }
+  \end{equation}
+\end{lemma}
+\begin{proof}
+  Let \(A\) be a type which is split enumerable, and \(U\) be a type family over
+  \(A\) which is split enumerable at every point.
+  Formally, we have the following proofs:
+  \begin{align}
+    \AgdaDatatype{\ensuremath{\mathcal{E}!}}_A &: \AgdaDatatype{\ensuremath{\mathcal{E}!}}(A) \\
+    \AgdaDatatype{\ensuremath{\mathcal{E}!}}_U &: \Pi(x : A) , \AgdaDatatype{\ensuremath{\mathcal{E}!}}(U(x))
+  \end{align}
+
+  Our task is to construct a proof of type:
+  \begin{equation}
+    \AgdaDatatype{\ensuremath{\mathcal{E}!}}(\Sigma(x : A) , U(x))
+  \end{equation}
+  This proof itself is composed of two components:
+  \begin{align}
+    \mathit{support} &: \AgdaDatatype{List}(\Sigma(x : A) , U(x)) \\
+    \mathit{cover} &: \Pi(x : \Sigma(y : A) , U(y)) , x \in \mathit{support}
+  \end{align}
+  To construct the support list, we apply the function \(\AgdaDatatype{\ensuremath{\mathcal{E}!}}_U\) to
+  every element in the support list of \(\AgdaDatatype{\ensuremath{\mathcal{E}!}}_A\), extract the support
+  lists from the resulting finiteness proofs, and concatenate them.
+
+  To prove that this support list does in fact cover the entirety of the type
+  \(\Sigma \; A \; U\), we note that any element of type \(\Sigma \; A \; U\)
+  must have a first component in the support list of \(\AgdaDatatype{\ensuremath{\mathcal{E}!}}_A\), and its
+  second component must be in the result of applying \(\AgdaDatatype{\ensuremath{\mathcal{E}!}}_U\) to that
+  first element (since that support list contains every element of type
+  \(U(x)\)).
+  Therefore, the pair itself must be in our constructed support list.
+\end{proof}
+
+This pattern of applying a function to each element in a list and
+concatenating the result is of course well-known in functional programming,
+and is in fact the pattern that makes lists a monad.
+While this insight isn't strictly relevant to our work here, it does mean
+the implementation of this function can use Agda's do notation, resulting
+in the following extremely clean implementation:
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{sup-sigma}
+\end{agdalisting}
+
+We now have two components we'll need for the proof that the countdown
+transformation is finite.
+The component we'll look at is step~\ref{countdown-operators}: selection of the
+operators.
+We'll first need a type representing the operators available to us.
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Countdown.tex]{ops-def}
+\end{agdalisting}
+Proving that this type is finite takes much the same form as the proof of
+finiteness for bool.
+\begin{agdalisting} \label{op-fin}
+  \ExecuteMetaData[agda/Countdown.tex]{op-fin}
+\end{agdalisting}
+
+Next, we will need to build a proof of finiteness for vectors of length \(n\).
+This uses the proof of finiteness for \(\Sigma\).
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Countdown.tex]{vec-fin}
+\end{agdalisting}
+
+\subsection{Manifest Bishop Closure Under \(\Pi\)}
+The glaring omission from our closure proofs under type formers so far has been
+the \(\Pi\) type: we have not proved closure under functions, dependent or
+otherwise.
+In MLTT, this is of course not provable: since all of the finiteness predicates
+we have seen so far imply decidable equality, and since we don't have any kind
+of decidable equality on functions in MLTT, we know that we won't be able to
+show that any kind of function is finite; even one like \(\AgdaDatatype{Bool}
+\rightarrow \AgdaDatatype{Bool}\).
+
+CuTT is not so restricted.
+Since we have things like function extensionality and transport, we can indeed
+prove the finiteness of function types.
+Our proof here makes use directly of the univalence axiom, and makes use
+furthermore of all the previous closure proofs.
+We will prove this closure on split enumerability, rather than on manifest
+Bishop finiteness, as it requires slightly less legwork in the proof itself, but
+of course we can derive the proof on manifest Bishop finiteness in a few lines.
+\begin{theorem} \label{split-enum-pi}
+  Split enumerability is closed under dependent functions.
+  (\(\Pi\)-types).
+  \begin{equation}
+    \frac{
+      \AgdaDatatype{\ensuremath{\mathcal{E}!}}(A) \; \; \; \Pi {(x : A)} , \AgdaDatatype{\ensuremath{\mathcal{E}!}}\left( U(x) \right)
+    }{
+      \AgdaDatatype{\ensuremath{\mathcal{E}!}}\left(\Pi {(x : A)} , U(x)\right)
+    }
+  \end{equation}
+\end{theorem}
+\begin{proof}
+  Let \(A\) be a split enumerable type, and \(U\) be a type family from \(A\),
+  which is split enumerable over all points of \(A\).
+
+  As \(A\) is split enumerable, we know that it is also manifestly Bishop finite
+  (lemma~\ref{split-enum-to-manifest-bishop}), and consequently we know \(A
+  \simeq \AgdaDatatype{Fin}\;n\), for some \(n\) (lemma~\ref{bishop-equiv}).
+  We can therefore replace all occurrences of \(A\) with \(\AgdaDatatype{Fin}\;n\),
+  changing our goal to:
+  \begin{equation}
+    \frac{
+      \AgdaDatatype{\ensuremath{\mathcal{E}!}}(\AgdaDatatype{Fin}\;n) \; \; \; \Pi (x : \AgdaDatatype{Fin}\;n) , \AgdaDatatype{\ensuremath{\mathcal{E}!}}\left( U(x) \right)
+    }{
+      \AgdaDatatype{\ensuremath{\mathcal{E}!}}\left(\Pi (x : \AgdaDatatype{Fin}\;n) , U(x)\right)
+    }
+  \end{equation}
+
+  We then define the type of \(n\)-tuples over some type family \(T :
+  \AgdaDatatype{Fin}\;n \rightarrow \mathbf{Type}\).
+  \begin{equation}
+    \begin{alignedat}{3}
+      & \mathbf{Tuple}(0, T)   &&\coloneqq \top \\
+      & \mathbf{Tuple}(n+1, T) &&\coloneqq T(0) \times \mathbf{Tuple}(n, T \circ \text{suc})
+    \end{alignedat}
+  \end{equation}
+  We can show that this type is equivalent to functions (proven in our formalisation):
+  \begin{equation}
+    \Pi(x : \AgdaDatatype{Fin}\;n) , U(x) \simeq \mathbf{Tuple}(n, U)
+  \end{equation}
+  And therefore we can simplify again our goal to the following:
+  \begin{equation}
+    \frac{
+      \AgdaDatatype{\ensuremath{\mathcal{E}!}}(\AgdaDatatype{Fin}\;n) \; \; \; \Pi (x : \AgdaDatatype{Fin}\;n) , \AgdaDatatype{\ensuremath{\mathcal{E}!}}\left( U(x) \right)
+    }{
+      \AgdaDatatype{\ensuremath{\mathcal{E}!}}\left(\mathbf{Tuple}(n, U)\right)
+    }
+  \end{equation}
+
+  We can prove this goal by showing that \(\mathbf{Tuple}(n, U)\) is split
+  enumerable: it is made up of finitely many products of points of \(U\), which
+  are themselves split enumerable, and \(\top\), which is also split enumerable.
+  Lemma~\ref{split-enum-sigma} shows us that the product of finitely many split
+  enumerable types is itself split enumerable, proving our goal.
+\end{proof}
+
+This proof can again give us insight into how to prove finiteness of our
+countdown transformation.
+In the first step (Fig.~\ref{countdown-selection}), we need to select some
+numbers from an input list: this can be described with a function of type
+\(\AgdaDatatype{Fin}\;n \rightarrow \AgdaDatatype{Bool}\), from indices in the
+original list into whether we keep the values or not.
+We now know that we can prove functions finite without difficulty: in this case,
+we can do it even more simply by proving that an \(n\)-tuple of booleans is
+finite.
+
+\subsection{Closure on Cardinal Finiteness}
+Since we don't have a function of type \(\mathcal{C}(A) \rightarrow
+\AgdaDatatype{\ensuremath{\mathcal{B}}}(A)\), closure proofs on \(\AgdaDatatype{\ensuremath{\mathcal{B}}}\) do not transfer over to
+\(\mathcal{C}\) trivially (unlike with \(\AgdaDatatype{\ensuremath{\mathcal{E}!}}\) and \(\AgdaDatatype{\ensuremath{\mathcal{B}}}\)).
+The cases for \(\bot\), \(\top\), and \(\AgdaDatatype{Bool}\) are simple to adapt: we
+can just propositionally truncate their Bishop finiteness proof.
+
+Non-dependent operators like \(\times\), \(\uplus\), and \(\rightarrow\) are
+also relatively straightforward: since \(\lVert {\wc} \rVert\) forms a monad, we
+can apply \(n\)-ary functions to values inside it, combining them together.
+\begin{agdalisting}
+  The fact that \(\lVert \wc \rVert\) forms a monad means that we can lift
+  \(n\)-ary functions like the following:
+  \ExecuteMetaData[agda/Cardinality/Finite/ManifestBishop.tex]{times-clos-sig}
+  Into a truncated context:
+  \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{times-clos-impl}
+\end{agdalisting}
+
+Unfortunately, for the dependent type formers like \(\Sigma\) and \(\Pi\), the
+same trick does not work.
+We have closure proofs like:
+\begin{equation}
+  \frac{
+    \AgdaDatatype{\ensuremath{\mathcal{B}}}(A) \; \; \; \Pi(x : A) , \AgdaDatatype{\ensuremath{\mathcal{B}}}(U(x))
+  }{
+    \AgdaDatatype{\ensuremath{\mathcal{B}}}(\Pi \; A \; U)
+  }
+\end{equation}
+If we apply the monadic truncation trick we can derive closure proofs like the
+following:
+\begin{equation}
+  \frac{
+    \lVert \AgdaDatatype{\ensuremath{\mathcal{B}}}(A) \rVert \; \; \; \lVert \Pi(x : A) , \AgdaDatatype{\ensuremath{\mathcal{B}}}(U(x)) \rVert
+  }{
+    \lVert \AgdaDatatype{\ensuremath{\mathcal{B}}}(\Pi \; A \; U) \rVert
+  }
+\end{equation}
+However our \emph{desired} closure proof is the following:
+\begin{equation}
+  \frac{
+    \lVert \AgdaDatatype{\ensuremath{\mathcal{B}}}(A) \rVert \; \; \; \Pi(x : A) , \lVert \AgdaDatatype{\ensuremath{\mathcal{B}}}(U(x)) \rVert
+  }{
+    \lVert \AgdaDatatype{\ensuremath{\mathcal{B}}}(\Pi \; A \; U) \rVert
+  }
+\end{equation}
+They don't match!
+
+The solution would be to find a function of the following type:
+\begin{equation}
+  (\Pi(x : A) , \lVert \AgdaDatatype{\ensuremath{\mathcal{B}}}(U(x)) \rVert) \rightarrow
+  \lVert \Pi(x : A) , \AgdaDatatype{\ensuremath{\mathcal{B}}}(U(x)) \rVert
+\end{equation}
+However we might be disheartened at realising that this is a required goal: the
+above equation is \emph{extremely} similar to the axiom of choice!
+\begin{definition}[Axiom of Choice]
+  In HoTT, the axiom of choice is commonly defined as follows \cite[lemma
+  3.8.2]{hottbook}.
+  For any set \(A\), and a type family \(U\) which is a set at all the points
+  of \(A\), the following function exists:
+  \begin{equation}
+    \left( \Pi(x : A) ,  \lVert U(x) \rVert \right) \rightarrow \lVert \Pi(x : A) , U(x) \rVert
+  \end{equation}
+\end{definition}
+Luckily the axiom of choice \emph{does} hold for cardinally finite types,
+allowing us to prove the following:
+\begin{lemma}
+  \begin{equation}
+    \mathcal{C}(A) \rightarrow (\Pi(x : A) , \lVert U(x) \rVert) \rightarrow \lVert \Pi(x : A) , U(x) \rVert
+  \end{equation}
+\end{lemma}
+\begin{proof}
+  Let \(A\) be a cardinally finite type, \(U\) be a type family on \(A\), and
+  \(f\) be a dependent function of type \(\Pi(x : A) , \lVert U(x) \rVert\).
+
+  First, since our goal is itself propositionally truncated, we have access to
+  values under truncations: put another way, in the context of proving our goal,
+  we can rely on the fact that \(A\) is manifestly Bishop finite.
+  Using the same technique as we did in lemma~\ref{split-enum-pi}, we can switch
+  from working with dependent functions from \(A\) to \(n\)-tuples, where \(n\)
+  is the cardinality of \(A\).
+  This changes our goal to the following:
+  \begin{equation}
+    \mathbf{Tuple}(n, \lVert \wc \rVert \circ U) \rightarrow \lVert \mathbf{Tuple}(n, U) \rVert
+  \end{equation}
+  Since \(\lVert \wc \rVert\) is closed under finite products, this function
+  exists (in fact, using the fact that \(\lVert \wc \rVert\) forms a monad, we
+  can recognise this function as \verb+sequenceA+ from the \verb+Traversable+
+  class in Haskell).
+\end{proof}
+
+This gets us all of the necessary closure proofs on \(\mathcal{C}\).
 \section{The Absence of the Subobject Classifier}
+\begin{agdalisting} \label{filter-subobject}
+  \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{subobject}
+\end{agdalisting}
 

### 2020-06-28 sections/finiteness-predicates.tex
bit more work on finiteness predicates section

@@
 in Figure~\ref{finite-classification}.
-As we will see, each of these predicates has subtle differences from the others:
-we will outline how some predicates are too informative, and how others aren't
-powerful enough, for our needs, before settling on decidable Kuratowski
-finiteness as our focus.
-
-As we make our way through each predicate, we will be interested in two aspects:
-how can we build proofs of this predicate (i.e. is the product of two finite
-types finite?) and what do we \emph{get} once we do (i.e. does this predicate
-tell us the number of elements in the finite set?).
-We will use the Countdown problem throughout as a running example of a thing to
-prove finite.
+We will also explain \emph{why} there are five separate predicates: how can it
+be the case that so many different things describe ``finiteness''?
+As we will see, some predicates are too informative (they
+tell us more about the underlying type other than it just being finite), or too
+restrictive (they don't allow certain finite types to be classified as finite).
+These diversions won't be dead-ends, however: the final predicate we will land
+on as the ``correct'' (or, more accurately, most useful) notion of finiteness
+will be built out of all of the others.
 \section{Split Enumerability} \label{split-enumerability}
@@
 
+\subsection{Terms used in the Definition of Split Enumerability}
 We have used two types there which we have not yet defined:
@@
   A proof of split enumerability is equivalent to a split surjection from a
-  finite prefix of the natural numbers. \todo{To Agda}
-  \begin{equation}
-    \AgdaDatatype{\ensuremath{\mathcal{E}!}}(A) \simeq \Sigma (n : \mathbb{N}) , \left( \AgdaDatatype{Fin}\;n \twoheadrightarrow ! \; A \right)
-  \end{equation}
+  finite prefix of the natural numbers.
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{is-split-inj-type}
+  \end{agdalisting}
 \end{lemma}
 \todo{Use standard equational reasoning syntax}
-\begin{proof}
+\begin{proof} \let\qed\relax
   \begin{minipage}[t]{.7\textwidth}\vspace{-\baselineskip}
@@
 
-      def.~\ref{split-enum-def} (\AgdaDatatype{\ensuremath{\mathcal{E}!}}) \\
-      eqn.~\ref{container-membership} (\AgdaDatatype{\ensuremath{\in}}) \\
-      eqn.~\ref{sp-surj-eqn}  \\
-      def.~\ref{List} (\(\AgdaDatatype{List}\)) \\
-      eqn.~\ref{container-interp}  \\
+      Def.~\ref{split-enum-def} (\AgdaDatatype{\ensuremath{\mathcal{E}!}}) \\
+      Eqn.~\ref{container-membership} (\AgdaDatatype{\ensuremath{\in}}) \\
+      Eqn.~\ref{sp-surj-eqn}  \\
+      Def.~\ref{List} (\(\AgdaDatatype{List}\)) \\
+      Eqn.~\ref{container-interp}  \\
       Reassociation \\
-      eqn.~\ref{sp-surj-arrow-eqn}
+      Eqn.~\ref{sp-surj-arrow-eqn}
   \end{minipage}

### 2020-06-28 sections/finiteness-predicates.tex
fixed equational reasoning syntax
bot more organisation of the split enumerable section

@@
 enumerability.
-
+This predicate is perhaps the first definition of ``finite'' that someone might
+come up with (it's certainly the most common in dependently-typed programming):
+put simply, a split enumerable type is a type for which all of its elements can
+be listed.
 \begin{definition}[Split Enumerable Set] \label{split-enum-def}
@@
 
+Simple as it is, this predicate is arguably the most useful of the finiteness
+predicates.
+We will see presently why it doesn't entirely capture what we mean by the term
+``finite'', but we will nonetheless use it throughout the paper.
+Before we dive in to exploring the predicate itself, though, we will need to
+explain some of the terms we used in its definition.
 \subsection{Terms used in the Definition of Split Enumerability}
@@
 equal to \(x\)''.
-\todo{Probably should put back in some instances here, i.e. bool and top.
-  Obviously not the full closure stuff, but one or two instances would be
-  instructive.}
-\subsection{Derivations}
-We have a way to construct finiteness proofs, and a semiring-like toolbox to
-combine them.
-What we're now interested in is what we can \emph{derive} from them.
-
-First, we will look at how this predicate relates to more traditional, classical
+\subsection{Split Surjections}
+We will now look at how this predicate relates to more traditional, classical
 notions of finiteness.
-in a classical setting we likely wouldn't mention ``lists'' or the like, and
+In a classical setting we likely wouldn't mention ``lists'' or the like, and
 would instead define finiteness based on the existence of some injection or
@@
   We define \emph{split} surjections here \cite[definition
-  4.6.1]{hottbook}. \todo{To Agda}
-  \begin{alignat}{3}
-    &\text{sp-surj}(f)          &&\coloneqq \Pi(y : B) , \text{fib}_f(y) \label{sp-surj-eqn} \\
-    &A \twoheadrightarrow! \; B &&\coloneqq \Sigma (f : A \rightarrow B) , \text{sp-surj}(f) \label{sp-surj-arrow-eqn}
-  \end{alignat}
+  4.6.1]{hottbook}.
+  \begin{agdalisting} \label{sp-surj-eqn}
+    \ExecuteMetaData[agda/Function/Surjective/Base.tex]{split-surjective}
+  \end{agdalisting} \vspace{-1.5\baselineskip}
+  \begin{agdalisting} \label{sp-surj-arrow-eqn}
+    \ExecuteMetaData[agda/Function/Surjective/Base.tex]{split-surjection}
+  \end{agdalisting}
 \end{definition}
@@
 \end{lemma}
-\todo{Use standard equational reasoning syntax}
-\begin{proof} \let\qed\relax
+\begin{proof} \let\qed\relax \vspace{-1\baselineskip}
   \begin{minipage}[t]{.7\textwidth}\vspace{-\baselineskip}
@@
 container-based definition of lists over the traditional one.)
-
+\subsection{Instances}
+We have now characterised the informativeness of split enumerability (it's the
+same as a split surjection), so we should also look at its
+\emph{restrictiveness}.
+For this we'll first look at some simple types which are clearly finite, and
+we'll see how to prove that's the fact using split enumerability.
+\begin{lemma}
+  \(\bot\), \(\top\), and \(\AgdaDatatype{Bool}\) are split enumerable.
+\end{lemma}
+\begin{proof}
+  These three types are quite obviously finite: we will show only the proof of
+  finiteness for \(\AgdaDatatype{Bool}\) here for brevity's sake.
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{bool-inst}
+  \end{agdalisting}
+\end{proof}
+\subsection{Decidable Equality}
 Split enumerability implies decidable equality on the underlying type.

### 2020-06-28 sections/topos.tex
bot more organisation of the split enumerable section

@@
 some things are finite.
-\begin{lemma}
-  \(\bot\), \(\top\), and \(\AgdaDatatype{Bool}\) are split enumerable.
-\end{lemma}
-\begin{proof}
-  These three types are quite obviously finite: we will show only the proof of
-  finiteness for \(\AgdaDatatype{Bool}\) here for brevity's sake.
-  \begin{agdalisting}
-    \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{bool-inst}
-  \end{agdalisting}
-\end{proof}
 With the most basic simple types out of the way, the obvious next choice is the

### 2020-06-28 sections/finiteness-predicates.tex
finished instances bit for split enumerability

@@
     \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{bool-inst}
+  \end{agdalisting}
+\end{proof}
+
+As a slightly more complex example, consider the \(\AgdaDatatype{Fin}\) type
+we've been using.
+Remember that split enumerability is in fact the same as a split surjection from
+\(\AgdaDatatype{Fin}\) (Lemma~\ref{split-enum-is-split-surj}): to show that
+\(\AgdaDatatype{Fin}\) is split enumerable, then, we need only show that it has
+a split surjection from itself.
+We'll prove the following slightly more general statement:
+\begin{lemma}
+  Every type has a split surjection into itself: the identity function.
+\end{lemma}
+\begin{proof}
+  We provide the simple proof here in Agda:
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Function/Surjective/Properties.tex]{split-surj-ident}
   \end{agdalisting}

### 2020-06-28 sections/finiteness-predicates.tex
moved circle

@@
 \section{Split Enumerability} \label{split-enumerability}
-We will start with a simple notion of finiteness, called \emph{split}
-enumerability.
+We will start with a simple notion of finiteness, called split enumerability.
 This predicate is perhaps the first definition of ``finite'' that someone might
@@
 \subsection{Decidable Equality}
+Since we're working in HoTT, it's quite natural to ask how these predicates
+relate to non-set types, like the circle.
+\begin{definition}[\(S^1\)] \label{circle-def}
+  The circle, \(S^1\), can be represented in HoTT as a higher inductive type.
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Snippets/Circle.tex]{circle-def}
+  \end{agdalisting}
+  We will use it here as an example of a non-set type, i.e. a type for which not
+  all paths are equal.
+  This also means that it does not have decidable equality.
+\end{definition}
+
 Split enumerability implies decidable equality on the underlying type.
@@
 finiteness predicate.
-\begin{definition}[\(S^1\)] \label{circle-def}
-  The circle, \(S^1\), can be represented in HoTT as a higher inductive type.
-  \begin{equation}
-    \begin{alignedat}{3}
-      S^1 \coloneqq & \; \text{base} &&: S^1 ; \\
-      | & \; \text{loop} &&: \text{base} \equiv \text{base} ; 
-    \end{alignedat}
-  \end{equation}
-  We will use it here as an example of a non-set type, i.e. a type for which not
-  all paths are equal.
-  This also means that it does not have decidable equality.
-\end{definition}
 \begin{lemma}

### 2020-06-28 sections/finiteness-predicates.tex
fixed todonotes

@@
 \subsection{Decidable Equality}
-Since we're working in HoTT, it's quite natural to ask how these predicates
-relate to non-set types, like the circle.
+We have now seen some examples of types which \emph{are} split enumerable: but
+what types are excluded?
+Clearly infinite types like \(\AgdaDatatype{\ensuremath{\mathbb{N}}}\) aren't
+allowed, but we haven't yet answered the question of non-set types.
+Certainly these can be finite in some sense, but can they be split enumerable?
+
+As our running example of a non-set type we will use the circle:
 \begin{definition}[\(S^1\)] \label{circle-def}
@@
 \end{definition}
+
+Any attempt to prove that this type is split enumerable will run into trouble:
+our task here, then, is to show that such a proof is impossible.
+To do will remember Hedberg's theorem
+\cite{hedbergCoherenceTheoremMartinLof1998}:
+\begin{theorem}[Hedberg's Theorem]
+  Every discrete type is a set.
+\end{theorem}
+With this theorem in mind, then, if we prove that every split enumerable type is
+discrete we will have proven that the circle is not split enumerable, because if
+it were it would be discrete and therefore a set, which we know already it is
+not.
+
+At this point we will define what we mean by \emph{discrete}.
+\begin{definition}[Discrete Types]
+  A discrete type is one with decidable equality.
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Relation/Nullary/Discrete/Base.tex]{discrete-def}
+  \end{agdalisting}
+\end{definition}
+We also use here a data type called \AgdaDatatype{Dec}.
+\begin{definition}[Decidable Types]
+  If some proposition is decidable it means that it is proven true or false.
+  We can encapsulate this with a data type\footnotemark:
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Snippets/Dec.tex]{dec-def}
+  \end{agdalisting}
+  A proof of type \(\AgdaDatatype{Dec}\;A\) is a proof that \(A\) is either true
+  or false.
+\end{definition}
+\footnotetext{This is not, as it happens, the definition of
+  \(\AgdaDatatype{Dec}\) you will find in Agda's standard library.
+The version of the type that exists there is a good deal more complicated (for
+computational reasons),
+although its interface behaves}
 
@@
 \end{lemma}
-
-\begin{definition}[Decidable Types]
-  \begin{equation}
-    \mathbf{Dec}(A) \coloneqq A \uplus \neg A
-  \end{equation}
-\end{definition}
-\begin{definition}[Discrete Types]
-  A discrete type is one with decidable equality.
-  \begin{equation}
-    \text{Discrete}(A) \coloneqq \Pi(x, y : A) , \mathbf{Dec}(x \equiv y)
-  \end{equation}
-  By Hedberg's theorem \cite{hedbergCoherenceTheoremMartinLof1998} any discrete
-  type is a set.
-\end{definition}
 

### 2020-06-28 sections/finiteness-predicates.tex
finished (first draft of?) split enumerable section

@@
 \subsection{Decidable Equality}
-We have now seen some examples of types which \emph{are} split enumerable: but
-what types are excluded?
-Clearly infinite types like \(\AgdaDatatype{\ensuremath{\mathbb{N}}}\) aren't
-allowed, but we haven't yet answered the question of non-set types.
+We have seen now some examples of types which \emph{are} split enumerable, but
+we haven't yet explored the space of types which \emph{aren't}.
+Clearly this includes obviously infinite types like
+\(\AgdaDatatype{\ensuremath{\mathbb{N}}}\), but it's the non-set types we're
+interested in for now.
 Certainly these can be finite in some sense, but can they be split enumerable?
@@
   \end{agdalisting}
-  We will use it here as an example of a non-set type, i.e. a type for which not
-  all paths are equal.
-  This also means that it does not have decidable equality.
 \end{definition}
+The presence of the \AgdaInductiveConstructor{loop} constructor is what
+precludes this type from being a set.
+In sets, all paths are equal: in this type, \AgdaInductiveConstructor{loop} is a
+path which is not equal to the usual identity path.
 
@@
 our task here, then, is to show that such a proof is impossible.
-To do will remember Hedberg's theorem
+To do so we will first remember Hedberg's theorem
 \cite{hedbergCoherenceTheoremMartinLof1998}:
@@
 
-Split enumerability implies decidable equality on the underlying type.
-To prove this, we will make use of the following lemma, proven in the
-formalisation:
 \begin{definition}[Injections]
@@
   surjective ones:
-  \begin{alignat}{3}
-    & \text{injective}(f) &&\coloneqq \Pi (x, y : A) , f \; x \equiv f \; y \rightarrow x \equiv y \\
-    & A \rightarrowtail B &&\coloneqq \Sigma(f : A \rightarrow B) , \text{injective}(f)
-  \end{alignat}
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Function/Injective/Base.tex]{injective}
+  \end{agdalisting} \vspace{-1.5\baselineskip}
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Function/Injective/Base.tex]{injection}
+  \end{agdalisting}
 \end{definition}
@@
   \(A\).
-  \begin{equation}
-    (A \twoheadrightarrow! \; B) \rightarrow (B \rightarrowtail A)
-  \end{equation}
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Function/Surjective/Properties.tex]{surj-to-inj}
+  \end{agdalisting}
 \end{lemma}
+\begin{proof} \let\qed\relax
+
+  \begin{minipage}[t]{.9\textwidth}\vspace{-\baselineskip}
+    \begin{agdalisting*}
+      \ExecuteMetaData[agda/Function/Surjective/Properties.tex]{surj-to-inj-proof}
+    \end{agdalisting*}
+  \end{minipage}
+\end{proof}
 \begin{lemma} \label{inj-discrete}
@@
   discrete.
-  \begin{equation}
-    \frac{
-      A \rightarrowtail B \; \; \; \text{Discrete}(B)
-    }{
-      \text{Discrete}(A)
-    }
-  \end{equation}
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Function/Injective/Properties.tex]{inj-discrete}
+  \end{agdalisting}
 \end{lemma}
@@
 \begin{lemma} \label{discrete-surj}
-  \begin{equation}
-    \frac{
-      A \twoheadrightarrow! \; B \; \; \; \text{Discrete}(A)
-    }{
-      \text{Discrete}(B) 
-    }
-  \end{equation}
+  If a discrete type \(A\) has a split surjection into some type \(B\), \(B\) is
+  discrete.
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Function/Surjective/Properties.tex]{discrete-surj}
+  \end{agdalisting}
 \end{lemma}

### 2020-06-28 sections/finiteness-predicates.tex
Update finiteness-predicates.tex

@@
 \end{theorem}
-With this theorem in mind, then, if we prove that every split enumerable type is
-discrete we will have proven that the circle is not split enumerable, because if
-it were it would be discrete and therefore a set, which we know already it is
-not.
+We will in a moment explain what ``discrete'' means, but first we should notice
+that his theorem gives us a possible avenue to prove that the circle is not
+split enumerable.
+We know, for instance, that the circle is \emph{not} a set\todo{proof/reference
+  for this?}.
+If we can show, then, that every split enumerable type is discrete, we will
+have excluded the circle (and indeed any other non-set type).
 
-At this point we will define what we mean by \emph{discrete}.
+Of course this doesn't make much sense without defining discrete: it turns out
+it's just HoTT-specific language for a concept more usually known as ``decidable
+equality''.
 \begin{definition}[Discrete Types]
@@
 \end{definition}
-We also use here a data type called \AgdaDatatype{Dec}.
+Of course we also will need to explain what we mean by ``decidable''.
 \begin{definition}[Decidable Types]
@@
 
+So now we have our task in focus: we need to prove that split enumerability
+implies decidable equality.
+We will do that by first describing split enumerability in terms of injections.
+\todo{This proof could be presented a lot better, I think}
 \begin{definition}[Injections]
@@
 \end{proof}
-
 \section{Manifest Bishop Finiteness} \label{manifest-bishop-finiteness}

### 2020-06-29 sections/finiteness-predicates.tex
more on propositional truncation

@@
   finiteness for \(\AgdaDatatype{Bool}\) here for brevity's sake.
-  \begin{agdalisting}
+  \begin{agdalisting} \label{bool-is-split-enumerable}
     \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{bool-inst}
@@
 
-As our running example of a non-set type we will use the circle:
+As our running example of a non-set type we will use the circle: \todo{HITs are
+  introduced later on. The definition should be moved to here}
 \begin{definition}[\(S^1\)] \label{circle-def}
@@
 With split enumerability we can see an instance of this.
-Consider the following proof of the finiteness of the operators for countdown:
+Consider the following proof of the finiteness of bool:
 \begin{agdalisting} \label{op-slop}
-  \ExecuteMetaData[agda/Countdown.tex]{op-slop}
+  \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{bool-slop}
 \end{agdalisting}
-While it represents the ``same'' information as the proof in
-equation~\ref{op-fin}, it clearly is not the same \emph{object}.
+There is an extra \AgdaInductiveConstructor{false} at the end of the support
+list.
+There's nothing terribly wrong with that: it is still a valid proof of
+finiteness, after all, but it does mean that this proof has some extra
+information which we didn't necessarily intend to encode.
 
@@
 values than there are \emph{usefully} distinct values.
-For the purposes of solving countdown this has the undesirable effect of
-duplicating search effort, but more generally this reveals that the predicate
-for finiteness we have doesn't represent in a concise way what we intend it to
-represent.
 To reconcile this, we will disallow duplicates in the support list.
 
-How exactly we should do this is the next question.
+This is where manifest Bishop finiteness comes in: this is a definition of
+finiteness quite similar to split enumerability in other regards, except that it
+does not allows for duplicates in the support list.
+
+How exactly to prohibit duplicates is the next question.
 One approach might be to change the definition of \(\AgdaDatatype{List}\), or
@@
 However, this would mean we lose access to the functions we have defined on
-lists, and we have to change the definition of \(\in\) as well.
+lists, and we have to change the definition of
+\(\AgdaFunction{\ensuremath{\in}}\) as well.
 
@@
 \AgdaDatatype{Bool}\) with
-duplicates, as there are multiple values which inhabit the type \(\text{false}
-\in \left[ \text{false}, \text{true}, \text{false} \right]\).
+duplicates, as there are multiple values which inhabit the type
+\(\AgdaInductiveConstructor{false}\;\AgdaFunction{\ensuremath{\in}}\;
+\AgdaInductiveConstructor{\ensuremath{\left[ \text{false}, \text{true},
+      \text{false} \right]}}\).
 It also allows us to keep most of the split enumerability definition unchanged,
@@
 While the types are not \emph{equivalent} (there are more split enumerable
-proofs than there are manifest Bishop finite proofs), they are of equal power,
-so any closure proof we have on one can be transferred to the other.
-In particular, it means that manifest Bishop finiteness is closed under
-\(\Sigma\).
+proofs than there are manifest Bishop finite proofs), they are of equal power.
 \subsection{From Manifest Bishop Finiteness to Equivalence}
@@
 \end{lemma}
-\begin{proof}
+\begin{proof} \todo{Convert to Agda}
   \begin{align*}
@@
 predicates, one piece still remains.
-\begin{agdalisting}
 The two following proofs are both valid proofs of the finiteness of
-  \(\AgdaDatatype{Bool}\), and both do not include any duplicates.
-  However they still differ:
-  \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{bool-inst} \smallskip
+\(\AgdaDatatype{Bool}\), and both do not include any duplicates:
+
+\begin{minipage}{\textwidth}
+\begin{multicols}{2}
+  \begin{agdalisting*}
+    \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{bool-inst}
+  \end{agdalisting*} \columnbreak
+  \begin{agdalisting*}
     \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{bool-rev}
-\end{agdalisting}
+  \end{agdalisting*}
+\end{multicols}
+\end{minipage}
+Clearly they're not the same though: the order of their support lists differs.
 Each finiteness predicate so far has contained an \emph{ordering} of the
@@
   It is defined as a Higher Inductive Type:
-  \begin{equation} {
-    \begin{alignedat}{3}
-      \lVert A \rVert \coloneqq & \; \lvert \wc \rvert &:& \; A \rightarrow \lVert A \rVert ; \\
-                              | & \; \text{squash}     &:& \; \Pi {(x, y : \lVert A \rVert)} , x \equiv y  ; 
-    \end{alignedat} }
-  \end{equation}
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Snippets/PropTrunc.tex]{prop-trunc-def}
+  \end{agdalisting}
   We will use two eliminators from \(\lVert A \rVert\) in this paper.

### 2020-06-29 sections/finiteness-predicates.tex
bit more on finiteness predicates section

@@
 container-based definition of lists over the traditional one.)
-\subsection{Instances}
-We have now characterised the informativeness of split enumerability (it's the
-same as a split surjection), so we should also look at its
-\emph{restrictiveness}.
-For this we'll first look at some simple types which are clearly finite, and
-we'll see how to prove that's the fact using split enumerability.
+\subsection{Instances} \todo{More intro to this}
+We will now provide some example proofs of types which are split enumerable.
 \begin{lemma}
@@
 \subsection{Decidable Equality}
-We have seen now some examples of types which \emph{are} split enumerable, but
-we haven't yet explored the space of types which \emph{aren't}.
-Clearly this includes obviously infinite types like
-\(\AgdaDatatype{\ensuremath{\mathbb{N}}}\), but it's the non-set types we're
-interested in for now.
-Certainly these can be finite in some sense, but can they be split enumerable?
-
-As our running example of a non-set type we will use the circle: \todo{HITs are
-  introduced later on. The definition should be moved to here}
-\begin{definition}[\(S^1\)] \label{circle-def}
-  The circle, \(S^1\), can be represented in HoTT as a higher inductive type.
-  \begin{agdalisting}
-    \ExecuteMetaData[agda/Snippets/Circle.tex]{circle-def}
-  \end{agdalisting}
-\end{definition}
-The presence of the \AgdaInductiveConstructor{loop} constructor is what
-precludes this type from being a set.
-In sets, all paths are equal: in this type, \AgdaInductiveConstructor{loop} is a
-path which is not equal to the usual identity path.
-
-Any attempt to prove that this type is split enumerable will run into trouble:
-our task here, then, is to show that such a proof is impossible.
-To do so we will first remember Hedberg's theorem
-\cite{hedbergCoherenceTheoremMartinLof1998}:
-\begin{theorem}[Hedberg's Theorem]
-  Every discrete type is a set.
-\end{theorem}
-We will in a moment explain what ``discrete'' means, but first we should notice
-that his theorem gives us a possible avenue to prove that the circle is not
-split enumerable.
-We know, for instance, that the circle is \emph{not} a set\todo{proof/reference
-  for this?}.
-If we can show, then, that every split enumerable type is discrete, we will
-have excluded the circle (and indeed any other non-set type).
-
-Of course this doesn't make much sense without defining discrete: it turns out
-it's just HoTT-specific language for a concept more usually known as ``decidable
-equality''.
+One thing that characterises all split enumerable types is that they are all
+\emph{discrete}.
+We will see later that this has implications for the space of types we're
+dealing with, but for now it simply provides a useful function on split
+enumerable types.
 \begin{definition}[Discrete Types]
@@
 
-\begin{definition}[Homotopy Levels] \label{homotopy-types}
+\begin{definition}[Contractions, Propositions and Sets] \label{homotopy-types}
   Types in HoTT and CuTT are not necessarily sets, as they are in MLTT.
@@
   \end{agdalisting}
+  First we have the contractions: those types equivalent to
+  \AgdaDatatype{\ensuremath{\top}}.
+  Then the propositions: those types for which every value is equal to every
+  other.
+  Finally the sets: those types for which every \emph{equality} is equal to
+  every other.
+
+  Often these types are defined inductively: going from contractions upwards,
+  every new level has paths in the level before (i.e. the paths of a proposition
+  are all contractions).
+  We can also proceed further upwards, into non-set types.
 \end{definition}
@@
   the natural numbers.
-  \begin{equation}
-    \mathcal{C}(A) \coloneqq \lVert \AgdaDatatype{\ensuremath{\mathcal{B}}}(A) \rVert \simeq \lVert \Sigma(n : \mathbb{N}) , (\AgdaDatatype{Fin}\;n \simeq A) \rVert
-  \end{equation}
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{cardinal-def}
+  \end{agdalisting}
 \end{definition}
@@
 \section{Manifest Enumerability} \label{manifest-enumerability}
+We have now explored quite far in the direction of restricting information:
+indeed, cardinal finiteness, a mere proposition, has the minimal amount of
+information a predicate can have.
+What we haven't seen, however, is a predicate with a different domain to the
+others.
+Since manifest Bishop finiteness and split enumerability have equal strength
+\todo{ref here}, and since we've shown that there is no type which is cardinal
+finite but not manifest Bishop finite, we haven't explored any of the
+``restrictiveness'' axis from Figure~\ref{finite-classification}.
+
+As an example type which may be considered finite in some circumstances, but
+doesn't conform to any of the definitions we have seen so far, consider the
+circle.
+\begin{definition}[\(S^1\)] \label{circle-def}
+  The circle, \(S^1\), can be represented in HoTT as a higher inductive type.
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Snippets/Circle.tex]{circle-def}
+  \end{agdalisting}
+\end{definition}
+The presence of the \AgdaInductiveConstructor{loop} constructor is what
+precludes this type from being a set.
+In sets, all paths are equal: in this type, \AgdaInductiveConstructor{loop} is a
+path which is not equal to the usual identity path.
+
+Any attempt to prove that this type is split enumerable will run into trouble:
+our task here, then, is to show that such a proof is impossible.
+To do so we will first remember Hedberg's theorem
+\cite{hedbergCoherenceTheoremMartinLof1998}:
+\begin{theorem}[Hedberg's Theorem]
+  Every discrete type is a set.
+\end{theorem}
+We will in a moment explain what ``discrete'' means, but first we should notice
+that his theorem gives us a possible avenue to prove that the circle is not
+split enumerable.
+We know, for instance, that the circle is \emph{not} a set\todo{proof/reference
+  for this?}.
+If we can show, then, that every split enumerable type is discrete, we will
+have excluded the circle (and indeed any other non-set type).
+
 We have now explored quite far in the ``less informative'' direction.

### 2020-06-29 sections/finiteness-predicates.tex
down to manifest enumerable

@@
   \end{agdalisting}
-\end{definition}
   The presence of the \AgdaInductiveConstructor{loop} constructor is what
@@
   path which is not equal to the usual identity path.
+\end{definition}
 
-Any attempt to prove that this type is split enumerable will run into trouble:
-our task here, then, is to show that such a proof is impossible.
-To do so we will first remember Hedberg's theorem
+We suggest that this type cannot be proven to conform to any of the finiteness
+predicates we have seen so far (split enumerability, manifest Bishop finiteness,
+or Cardinal finiteness).
+To provide some better evidence for this fact (other than ``we tried quite hard
+and couldn't manage it'') remember Hedberg's theorem
 \cite{hedbergCoherenceTheoremMartinLof1998}:
@@
 \end{theorem}
-We will in a moment explain what ``discrete'' means, but first we should notice
-that his theorem gives us a possible avenue to prove that the circle is not
-split enumerable.
-We know, for instance, that the circle is \emph{not} a set\todo{proof/reference
-  for this?}.
-If we can show, then, that every split enumerable type is discrete, we will
-have excluded the circle (and indeed any other non-set type).
+Since each finiteness predicate we have seen so far \emph{does} imply decidable
+equality, we know that the cardinal finite types can \emph{only} be sets, ruling
+out the circle.
 
-We have now explored quite far in the ``less informative'' direction.
-However, all three predicates we have examined are equally \emph{restrictive}:
-in this section we will see a predicate which is much less restrictive.
-In particular, this predicate ranges over non-set types.
-
+So our tour through the finiteness predicates takes us next to one which allows
+the circle to be called finite: manifest enumerability.
 \begin{definition}[Manifest Enumerability]
@@
   membership proof.
-  \begin{equation}
-    \mathcal{E}(A) \coloneqq \Sigma {(\mathit{xs} : \AgdaDatatype{List}(A))} , \Pi {(x : A)} , \lVert x \in \mathit{xs} \rVert
-  \end{equation}
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Cardinality/Finite/ManifestEnumerable/Container.tex]{manifest-enum-def}
+  \end{agdalisting}
 \end{definition}

### 2020-06-29 sections/finiteness-predicates.tex
kuratowski

@@
 \end{definition}
-As a function-based definition, this predicate represents surjections.
 
+It might not be immediately clear why this definition of enumerability allows
+the circle to conform while the others do not.
+The crux of the issue was that the cover proofs of the previous definitions
+didn't just tell us that some element was in the support list, they told us
+\emph{where} it was in the support list.
+From the position we were able to derive decidable equality: that position is
+precisely what's hidden in manifest enumerability.
+
+And indeed this means that the circle is manifestly enumerable:
+\begin{lemma}
+  The circle \(S^1\) is manifestly enumerable.
+\end{lemma}
+\begin{proof}
+  The support list firstly is a list containing the point constructor for the
+  circle.
+  Since the cover proof is truncated, we need only consider the point
+  constructors of the circle: as such, the cover proof is essentially the same
+  as the one for \(\AgdaDatatype{\ensuremath{\mathcal{E}!}}(\top)\).
+\end{proof}
+\subsection{Surjections}
+We already say that split enumerability was the listed form of a split
+surjection: manifest enumerability is the listed form of a proper surjection.
 \begin{definition}[Surjections] \label{surjections}
@@
   \end{align*}
-\end{proof}
-\subsection{Instances for Non-Set Types}
-The truncation has another very important implication: it means that the
-predicate doesn't provide decidable equality on the underlying type.
-Remember, this is how we knew that our previous predicates wouldn't allow for
-non-set types: because they implied decidable equality, they also implied that
-all conforming types had homotopy levels of at most 2. \todo{Are we doing the
-  homotopy levels starting from -2 thing or from 0?}
-This suggests that non-set types like the circle could conform to this
-finiteness predicate.
-\begin{lemma}
-  The circle \(S^1\) is manifestly enumerable.
-\end{lemma}
-\begin{proof}
-  The support list firstly is a list containing the point constructor for the
-  circle.
-  Since the cover proof is truncated, we need only consider the point
-  constructors of the circle: as such, the cover proof is essentially the same
-  as the one for \(\AgdaDatatype{\ensuremath{\mathcal{E}!}}(\top)\).
 \end{proof}
@@
 \end{proof}
-\section{Kuratowski Finiteness} \label{kuratowski}
+\section{Kuratowski Finiteness} \label{kuratowski} \todo{better intro}
 The one big missing definition of finiteness to cover is \emph{Kuratowski}
@@
   of Kuratowski-finite subsets of \(A\).
-  \begin{inductivetype}{\mathcal{K}(A)}
-    \inductivetypeclause{\left[ \right]}{\mathcal{K}(A)}
-    \inductivetypeclause{\wc \dblcolon \wc}{A \rightarrow \mathcal{K}(A) \rightarrow \mathcal{K}(A)}
-    \inductivetypeclause{\text{com}}{\Pi (x, y: A) , \Pi (\mathit{xs} : \mathcal{K}(A)) , x \dblcolon y \dblcolon \mathit{xs} \equiv y \dblcolon x \dblcolon \mathit{xs}}
-    \inductivetypeclause{\text{dup}}{\Pi (x : A) , \Pi (\mathit{xs} : \mathcal{K}(A)) , x \dblcolon x \dblcolon \mathit{xs} \equiv x \dblcolon \mathit{xs}}
-    \inductivetypeclause{\text{trunc}}{\Pi (\mathit{xs}, \mathit{ys}: \mathcal{K}(A)) , \Pi (p, q : \mathit{xs} \equiv \mathit{ys}) , p \equiv q}
-  \end{inductivetype}
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Algebra/Construct/Free/Semilattice/Definition.tex]{kuratowski-def}
+  \end{agdalisting}
+
   We define it as a HIT (definition~\ref{HITs}).
@@
   Membership is defined by pattern-matching on \(\mathcal{K}\).
-  The two point constructors are handled like so:
+  The two point constructors are handled like so: \todo{to agda}
   \begin{equation}
@@
 Finally, we have enough background to define Kuratowski finiteness.
-\begin{definition}[Kuratowski Finiteness]
-  \begin{equation}
-    \mathcal{K}^{f}(A) = \Sigma {(\mathit{xs} : \mathcal{K}(A))} , \Pi (x : A) , x \in \mathit{xs}
-  \end{equation}
+\begin{definition}[Kuratowski Finiteness] \phantom{break}
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Cardinality/Finite/Kuratowski.tex]{kuratowski-finite-def}
+  \end{agdalisting}
 \end{definition}

### 2020-06-29 sections/finiteness-predicates.tex
much nicer prose/definitions

@@
 
-Simple as it is, this predicate is arguably the most useful of the finiteness
-predicates.
-We will see presently why it doesn't entirely capture what we mean by the term
-``finite'', but we will nonetheless use it throughout the paper.
+This predicate is simple and useful, but we will see later on how it is perhaps
+a little imprecise.
 Before we dive in to exploring the predicate itself, though, we will need to
 explain some of the terms we used in its definition.
-\subsection{Terms used in the Definition of Split Enumerability}
-We have used two types there which we have not yet defined:
-\(\AgdaDatatype{List}\) and \(\AgdaDatatype{\ensuremath{\in}}\).
-We will define them here.
+\subsection{What is a List?}
+We have referred to \(\AgdaDatatype{List}\) in the definition of split
+enumerability.
+In this paper, the type of lists is defined like so:
+\begin{agdalisting} \label{list-def}
+  \ExecuteMetaData[agda/Container/List.tex]{list-def}
+\end{agdalisting}
+This is the definition for a \emph{container} (Definition~\ref{container-def}):
+effectively, the above definition says that ``Lists are a datatype whose shape
+is given by the natural numbers, and which can be indexed by numbers smaller
+than its shape''.
 
-\begin{definition}[\(\AgdaDatatype{List}\)] \label{List}
-  In this paper we will work with two equivalent definitions of lists.
-  The first is the standard definition as an inductive type:
+If that seems needlessly complex, don't worry: this definition is precisely
+equivalent to the usual inductive one. \todo{This reads a little too informal I think}
 \begin{agdalisting}
@@
 \end{agdalisting}
+And this isn't some kind of hand-waving equivalence, either: since we working in
+HoTT, we can (and do) prove that the two types are equal, allowing us to use one
+or the other depending on whichever is more convenient, and
+\(\AgdaFunction{subst}\) in the other representation without loss of generality.
+That said, defining lists as containers will reveal several interesting
+connections and proofs about split enumerability and the other predicates, 
+so for the remainder of the paper whenever we say \AgdaDatatype{List} we will
+mean Equation~\ref{list-def}.
 
-  The second way to define lists is to define them as a \emph{container}:
-  \begin{agdalisting}
-    \ExecuteMetaData[agda/Container/List.tex]{list-def}
-  \end{agdalisting}
-  The reason we use this second strange definition is that it turns out to be
-  quite useful in some later proofs.
-  We have proven the two types equivalent in our formalisation, however, so we
-  can switch between them freely without loss of generality.
-\end{definition}
-
-In defining lists we have introduced another concept which needs defining:
-\(\AgdaDatatype{Fin}\).
-\begin{definition}[\(\AgdaDatatype{Fin}\)] \label{Fin}
-  \(\AgdaDatatype{Fin}\;n\) is the type of natural numbers smaller than \(n\). We
-  define it the standard way:
+Before we get to those interesting proofs, though, there are some other things
+that need defining.
+\AgdaDatatype{Fin}, for instance: \(\AgdaDatatype{Fin}\;n\) is the type of
+natural numbers smaller than \(n\).
+Its definition relies on a type for disjoint union,
+\(\AgdaDatatype{\ensuremath{\uplus}}\).
+\begin{multicols}{2}
   \begin{agdalisting}
     \ExecuteMetaData[agda/Data/Fin/Base.tex]{fin-def}
-  \end{agdalisting}
-\end{definition}
-Here \(\uplus\) refers to the disjoint union of two types.
-\begin{definition}[Disjoint Union]
-  We define disjoint union as an inductive type.
+  \end{agdalisting} \columnbreak
   \begin{agdalisting}
@@
   \end{agdalisting}
-  It is also expressible with only \(\Sigma\):
-  \begin{agdalisting}
-    \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma-disj-union}
-  \end{agdalisting}
-  Although the inductive type definition is slightly more ergonomic.
-\end{definition}
+\end{multicols}
+
+% It is also expressible with only \(\Sigma\):
+% \begin{agdalisting}
+%   \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma-disj-union}
+% \end{agdalisting}
+% Although the inductive type definition is slightly more ergonomic.
 

### 2020-06-29 sections/finiteness-predicates.tex
Update finiteness-predicates.tex

@@
     \ExecuteMetaData[agda/Snippets/Introduction.tex]{disj-union}
-  \end{agdalisting}
-\end{multicols}
-
+  \end{agdalisting}%
   % It is also expressible with only \(\Sigma\):
@@
   % Although the inductive type definition is slightly more ergonomic.
-
-After that interlude, we can get back to defining containers.
+\end{multicols} \noindent
+And containers themselves, of course.
+Containers are a well-studied topic in dependent type theory, with a rich
+theory: we won't dive in to that here.
 \begin{definition}[Containers] \label{container-def}
@@
 The container representing finite lists is a pair of a natural number \(n\)
-representing the length of the list, and a function \(\AgdaDatatype{Fin}\;n\to
-A\), representing the indexing function into the list.
+representing the length (or ``shape'') of the list, and a function
+\(\AgdaDatatype{Fin}\;n\to A\), representing the indexing function into the
+list.
 
-\begin{definition}[Container Membership] \label{container-membership}
-  Membership of a container can be defined like so:
+One of the nice things about containers is it gives us a generic way to define
+``membership'':
+\begin{multicols}{2}
   \begin{agdalisting} 
     \ExecuteMetaData[agda/Container/Membership.tex]{membership-def}
-  \end{agdalisting}
-  Where \(x\;\AgdaDatatype{\ensuremath{\in}}\;\mathit{xs}\) is to be read as ``\(x\)
-  is in \(\mathit{xs}\)''.
-\end{definition}
-\begin{definition}[Fibers] \label{fibers}
-  A fiber \cite[definition 4.2.4]{hottbook} is defined over some function \(f :
-  A \rightarrow B\).
+  \end{agdalisting} \columnbreak
   \begin{agdalisting}
@@
   \end{agdalisting}
-\end{definition}
-
-Membership also makes more sense when described concretely in terms of lists.
-Understood this way, \(x\;\AgdaDatatype{\ensuremath{\in}}\;\mathit{xs}\) means
-``there is an index into \(\mathit{xs}\) such that the index points at an item
-equal to \(x\)''.
+\end{multicols} \noindent 
+Here we're using the homotopy-theory notion of a \AgdaDatatype{fiber} to define
+membership: a fiber for some function \(f\) and some point \(y\) in its codomain
+is a value \(x\) and a proof that \(f \; x \;\AgdaDatatype{≡}\;y\).
+Membership also makes more sense when described concretely in terms of lists:
+\(x\;\AgdaDatatype{\ensuremath{\in}}\;\mathit{xs}\) means ``there is an index
+into \(\mathit{xs}\) such that the index points at an item equal to \(x\)''.
 \subsection{Split Surjections}

### 2020-06-30 sections/finiteness-predicates.tex
down to surjections in split enumerability

@@
 \subsection{What is a List?}
-We have referred to \(\AgdaDatatype{List}\) in the definition of split
-enumerability.
-In this paper, the type of lists is defined like so:
+In this paper we prefer a slightly unusual definition for the type of lists:
 \begin{agdalisting} \label{list-def}
@@
 
+\todo{The rest of this section certainly reads a lot better because of the
+  inlined definitions, but I worry it's a little too informal and (possibly) too
+verbose}
+
 If that seems needlessly complex, don't worry: this definition is precisely
-equivalent to the usual inductive one. \todo{This reads a little too informal I think}
+equivalent to the usual inductive one.
 \begin{agdalisting}
@@
     \ExecuteMetaData[agda/Container.tex]{container-interp}
-  \end{agdalisting}
+  \end{agdalisting} \vspace{-\baselineskip}
 \end{definition}
@@
 \begin{multicols}{2}
-  \begin{agdalisting} 
+  \begin{agdalisting} \label{container-membership}
     \ExecuteMetaData[agda/Container/Membership.tex]{membership-def}
@@
 \subsection{Split Surjections}
-We will now look at how this predicate relates to more traditional, classical
-notions of finiteness.
+Now that we have our terms defined, let's look a little at how split
+enumerability relates to more traditional, classical notions of finiteness.
 In a classical setting we likely wouldn't mention ``lists'' or the like, and
 would instead define finiteness based on the existence of some injection or
-surjection.
-As it turns out, our definition of finiteness here is precisely the same as the
-surjection-based one, in quite a deep way!
-
-First, we will need to define our terms: in HoTT, surjections are a little more
-complex than what you'd find in either MLTT or classical mathematics.
-\begin{definition}[Split Surjections] \label{split-surjections}
-  We define \emph{split} surjections here \cite[definition
-  4.6.1]{hottbook}.
+surjection, say a surjection from a finite prefix of the natural numbers.
+In HoTT, surjections (or, more precisely, \emph{split} surjections
+\cite[definition 4.6.1]{hottbook}), are defined like so:
+\begin{multicols}{2}
   \begin{agdalisting} \label{sp-surj-eqn}
     \ExecuteMetaData[agda/Function/Surjective/Base.tex]{split-surjective}
-  \end{agdalisting} \vspace{-1.5\baselineskip}
+  \end{agdalisting} \columnbreak
   \begin{agdalisting} \label{sp-surj-arrow-eqn}
@@
   \end{agdalisting}
-\end{definition}
-Over sets, the surjections and split surjections are the same thing, but there
-is a difference one we involve non-set types like the circle.
-
-We will now see that split enumerability is in fact a split surjection in
-another form:
+\end{multicols}
+As it turns out, our definition of finiteness here is precisely the same as a
+surjection-based one, in quite a deep way!
 \begin{lemma} \label{split-enum-is-split-surj}%
@@
       Eqn.~\ref{sp-surj-eqn}  \\
-      Def.~\ref{List} (\(\AgdaDatatype{List}\)) \\
+      Eqn.~\ref{list-def} (\(\AgdaDatatype{List}\)) \\
       Eqn.~\ref{container-interp}  \\
@@
 \end{agdalisting}
-(The simplicity of this proof, by the way, is why we preferred the
-container-based definition of lists over the traditional one.)
+This means that we could have in fact written the whole proof as follows:
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{split-enum-is-split-surj-short}
+\end{agdalisting}
+The simplicity of this proof, by the way, is why we preferred the
+container-based definition of lists over the traditional one.
 \subsection{Instances} \todo{More intro to this}
@@
     & \simeq \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \AgdaDatatype{Fin} \rrbracket (A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
-    && \text{def.~\ref{List} } (\AgdaDatatype{List})
+    && \text{def.~\ref{list-def} } (\AgdaDatatype{List})
     \\
@@
     & \simeq \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \AgdaDatatype{Fin} \rrbracket (A)) , \text{surj}(\text{snd}(\mathit{xs}))
-    && \text{def.~\ref{List} } (\AgdaDatatype{List})
+    && \text{def.~\ref{list-def} } (\AgdaDatatype{List})
     \\

### 2020-06-30 sections/finiteness-predicates.tex
finished prose for section 2.1

@@
   \end{agdalisting}
-\end{multicols} \noindent 
+\end{multicols} \vspace{-1\baselineskip}\noindent 
 Here we're using the homotopy-theory notion of a \AgdaDatatype{fiber} to define
@@
   \end{agdalisting} 
-\end{multicols}
+\end{multicols}\vspace{-1\baselineskip}
 As it turns out, our definition of finiteness here is precisely the same as a
@@
 container-based definition of lists over the traditional one.
-\subsection{Instances} \todo{More intro to this}
-We will now provide some example proofs of types which are split enumerable.
-\begin{lemma}
-  \(\bot\), \(\top\), and \(\AgdaDatatype{Bool}\) are split enumerable.
-\end{lemma}
-\begin{proof}
-  These three types are quite obviously finite: we will show only the proof of
-  finiteness for \(\AgdaDatatype{Bool}\) here for brevity's sake.
+\subsection{Instances}
+To actually show that a type \(A\) is finite amounts to constructing a term of
+type \(\AgdaDatatype{\ensuremath{\mathcal{E}!}}\;A\).
+For simple types like \AgdaDatatype{Bool}, that is simple:
 \begin{agdalisting} \label{bool-is-split-enumerable}
@@
 \end{agdalisting}
-\end{proof}
 
@@
 We'll prove the following slightly more general statement:
-\begin{lemma}
-  Every type has a split surjection into itself: the identity function.
-\end{lemma}
-\begin{proof}
-  We provide the simple proof here in Agda:
 \begin{agdalisting}
@@
 \end{agdalisting}
-\end{proof}
 \subsection{Decidable Equality}
 One thing that characterises all split enumerable types is that they are all
-\emph{discrete}.
+\emph{discrete}, i.e. they have decidable equality.  \todo{All multicol agda code should probably be vertically centred like this}
+\begin{multicols}{2}  \null \vfill
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Relation/Nullary/Discrete/Base.tex]{discrete-def}
+  \end{agdalisting} \vfill \null \columnbreak
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Snippets/Dec.tex]{dec-def}
+  \end{agdalisting}
+\end{multicols} \vspace{-\baselineskip} \noindent
 We will see later that this has implications for the space of types we're
@@
 enumerable types.
-\begin{definition}[Discrete Types]
-  A discrete type is one with decidable equality.
-  \begin{agdalisting}
-    \ExecuteMetaData[agda/Relation/Nullary/Discrete/Base.tex]{discrete-def}
-  \end{agdalisting}
-\end{definition}
-Of course we also will need to explain what we mean by ``decidable''.
-\begin{definition}[Decidable Types]
-  If some proposition is decidable it means that it is proven true or false.
-  We can encapsulate this with a data type\footnotemark:
-  \begin{agdalisting}
-    \ExecuteMetaData[agda/Snippets/Dec.tex]{dec-def}
-  \end{agdalisting}
-  A proof of type \(\AgdaDatatype{Dec}\;A\) is a proof that \(A\) is either true
-  or false.
-\end{definition}
-\footnotetext{This is not, as it happens, the definition of
-  \(\AgdaDatatype{Dec}\) you will find in Agda's standard library.
-The version of the type that exists there is a good deal more complicated (for
-computational reasons),
-although its interface behaves}
 
-So now we have our task in focus: we need to prove that split enumerability
-implies decidable equality.
-We will do that by first describing split enumerability in terms of injections.
-\todo{This proof could be presented a lot better, I think}
-\begin{definition}[Injections]
-  Injective functions are more straightforward to define constructively than
-  surjective ones:
+To prove that split enumerability implies decidable equality we'll take a quick
+detour through injections.
+\begin{multicols}{2}
   \begin{agdalisting}
     \ExecuteMetaData[agda/Function/Injective/Base.tex]{injective}
-  \end{agdalisting} \vspace{-1.5\baselineskip}
+  \end{agdalisting} \columnbreak
   \begin{agdalisting}
@@
   \end{agdalisting}
-\end{definition}
-\begin{lemma} \label{split-surj-to-inj}
-  A split-surjection from \(A\) to \(B\) implies an injection from \(B\) to
-  \(A\).
+\end{multicols} \vspace{-\baselineskip} \noindent
+These are useful because we know that any type which injects into a discrete
+type is itself discrete:
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Function/Injective/Properties.tex]{inj-discrete}
+\end{agdalisting}
+And we can turn a split surjection from \(A\) to \(B\) into an injection from
+\(B\) to \(A\): \todo{make these side-wrapped floats?}
 \begin{agdalisting}
@@
 \end{agdalisting}
-\end{lemma}
-\begin{proof} \let\qed\relax
-
-  \begin{minipage}[t]{.9\textwidth}\vspace{-\baselineskip}
-    \begin{agdalisting*}
-      \ExecuteMetaData[agda/Function/Surjective/Properties.tex]{surj-to-inj-proof}
-    \end{agdalisting*}
-  \end{minipage}
-\end{proof}
-\begin{lemma} \label{inj-discrete}
-  For any type \(A\) which injects into a discrete type \(B\), \(A\) is
-  discrete.
-  \begin{agdalisting}
-    \ExecuteMetaData[agda/Function/Injective/Properties.tex]{inj-discrete}
-  \end{agdalisting}
-\end{lemma}
-
-\begin{lemma} \label{discrete-surj}
-  If a discrete type \(A\) has a split surjection into some type \(B\), \(B\) is
-  discrete.
+Yielding a simple proof that any type with a split surjection from a discrete
+type is itself discrete:
 \begin{agdalisting}
@@
 \end{agdalisting}
-\end{lemma}
-\begin{proof}
-  This proof is can be straightforwardly derived from lemmas
-  \ref{split-surj-to-inj} and \ref{inj-discrete}.
-\end{proof}
-
-\begin{lemma} \label{split-enum-discrete}
-  Every split enumerable type is discrete.
-\end{lemma}
-\begin{proof}
-  Let \(A\) be a split enumerable type.
-  By lemma~\ref{split-enum-is-split-surj}, there is a surjection from
-  \(\AgdaDatatype{Fin}\;n\) for some \(n\).
-  Also, we know that \(\AgdaDatatype{Fin}\;n\) is discrete (proven in our
-  formalisation).
-  Therefore, by lemma~\ref{discrete-surj}, \(A\) is discrete.
-\end{proof}
+Since split enumerability is really just a split surjection from
+\AgdaDatatype{Fin}, and since we know that \AgdaDatatype{Fin} is discrete, the
+overall proof resolves quite simply:
+\begin{agdalisting} \label{split-enum-discrete}
+  \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{split-is-discrete}
+\end{agdalisting}
 \section{Manifest Bishop Finiteness} \label{manifest-bishop-finiteness}

### 2020-06-30 scripts/check-all.sh

@@ -6,6 +6,6 @@ echo 'module Everything where
 ' >> Everything
 find -E . -type f -regex '.*\.l?agda' | cut -c 3- | cut -f1 -d'.' | sed 's/\//\./g' | sed 's/^/open import /' >> Everything
 mv Everything Everything.agda
-agda --html --html-dir=../docs Everything.agda
+agda Everything.agda
 rm Everything.agda
 )

### 2020-06-30 sections/finiteness-predicates.tex
added a wrapfig

@@
 
+\begin{wrapfigure}[5]{r}{.4\linewidth} \vspace{-1.6\baselineskip}
+  \ExecuteMetaData[agda/Data/List/Base.tex]{list-def} \vspace{-\baselineskip}
+  \caption{The Usual Inductive Definition}
+\end{wrapfigure}
+
 If that seems needlessly complex, don't worry: this definition is precisely
-equivalent to the usual inductive one.
-\begin{agdalisting}
-  \ExecuteMetaData[agda/Data/List/Base.tex]{list-def}
-\end{agdalisting}
-And this isn't some kind of hand-waving equivalence, either: since we working in
-HoTT, we can (and do) prove that the two types are equal, allowing us to use one
-or the other depending on whichever is more convenient, and
+equivalent to the usual inductive one. %
+% \begin{agdalisting}
+%   \ExecuteMetaData[agda/Data/List/Base.tex]{list-def}
+% \end{agdalisting}
+And this isn't some kind of hand-waving equivalence, either: since we are
+working in HoTT, we can (and do) prove that the two types are equal, allowing us
+to use one or the other depending on whichever is more convenient, and
 \(\AgdaFunction{subst}\) in the other representation without loss of generality.
@@
 \(\AgdaDatatype{\ensuremath{\uplus}}\).
-\begin{multicols}{2}
+\begin{multicols}{2} \null \vfill
   \begin{agdalisting}
     \ExecuteMetaData[agda/Data/Fin/Base.tex]{fin-def}
-  \end{agdalisting} \columnbreak
+  \end{agdalisting} \vfill \null \columnbreak
   \begin{agdalisting}
@@
 We'll prove the following slightly more general statement:
+
 \begin{agdalisting}

### 2020-06-30 sections/finiteness-predicates.tex
more work on bishop section

--- 2020-06-30 sections/finiteness-predicates.tex
@@
 
+% \begin{wrapfigure}[5]{r}{.4\linewidth} \vspace{-1.6\baselineskip}
+%   \ExecuteMetaData[agda/Data/List/Base.tex]{list-def} \vspace{-\baselineskip}
+%   \caption{The Usual Inductive Definition}
+% \end{wrapfigure}
+
 If that seems needlessly complex, don't worry: this definition is precisely
@@
 \end{agdalisting}
-And this isn't some kind of hand-waving equivalence, either: since we working in
-HoTT, we can (and do) prove that the two types are equal, allowing us to use one
-or the other depending on whichever is more convenient, and
+And this isn't some kind of hand-waving equivalence, either: since we are
+working in HoTT, we can (and do) prove that the two types are equal, allowing us
+to use one or the other depending on whichever is more convenient, and
 \(\AgdaFunction{subst}\) in the other representation without loss of generality.
@@
 To specify that a value must exist uniquely in HoTT we can use the concept of a
-\emph{contraction}.
+\emph{contraction} \cite[definition 3.11.1]{hottbook}.
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Snippets/Introduction.tex]{isContr}
+\end{agdalisting}
+A contraction is a type with the least possible amount of information: it
+represents the tautologies.
+All contractions are isomorphic to \AgdaDatatype{\ensuremath{\top}}. \todo{We'll
+now need to define propositions and sets later on}
 
-\begin{definition}[Contractions, Propositions and Sets] \label{homotopy-types}
-  Types in HoTT and CuTT are not necessarily sets, as they are in MLTT.
-  Some have higher homotopies (paths which aren't unique).
-  We actually have a hierarchy of complexity of structure of path spaces in
-  types, starting with the contractions \cite[definition 3.11.1]{hottbook}, then
-  the mere propositions \cite[definition 3.3.1]{hottbook}, and the sets
-  \cite[definition 3.1.1]{hottbook}.
-  \begin{agdalisting}
-    \ExecuteMetaData[agda/Snippets/Introduction.tex]{hlevels}
-  \end{agdalisting}
-  First we have the contractions: those types equivalent to
-  \AgdaDatatype{\ensuremath{\top}}.
-  Then the propositions: those types for which every value is equal to every
-  other.
-  Finally the sets: those types for which every \emph{equality} is equal to
-  every other.
-
-  Often these types are defined inductively: going from contractions upwards,
-  every new level has paths in the level before (i.e. the paths of a proposition
-  are all contractions).
-  We can also proceed further upwards, into non-set types.
-\end{definition}
-\begin{definition}[Unique Membership] \label{uniq-memb-def}
-  Unique list membership is defined in terms of list membership: it is a
-  contraction of it.
-  \begin{agdalisting}
+By saying that a proof of membership is a contraction, we are saying that it
+must be \emph{unique}.
+\begin{agdalisting} \label{uniq-memb-def}
   \ExecuteMetaData[agda/Container/Membership.tex]{uniq-memb-def}
 \end{agdalisting}
-\end{definition}
+Now a proof of \(x\;\AgdaFunction{\ensuremath{\in}!}\;\mathit{xs}\) means that
+\(x\) is not just in \(\mathit{xs}\), but it appears there \emph{only once}.
+
+% \begin{definition}[Contractions, Propositions and Sets] \label{homotopy-types}
+%   Types in HoTT and CuTT are not necessarily sets, as they are in MLTT.
+%   Some have higher homotopies (paths which aren't unique).
+%   We actually have a hierarchy of complexity of structure of path spaces in
+%   types, starting with the contractions , then
+%   the mere propositions \cite[definition 3.3.1]{hottbook}, and the sets
+%   \cite[definition 3.1.1]{hottbook}.
+%   First we have the contractions: those types equivalent to
+%   \AgdaDatatype{\ensuremath{\top}}.
+%   Then the propositions: those types for which every value is equal to every
+%   other.
+%   Finally the sets: those types for which every \emph{equality} is equal to
+%   every other.
+
+%   Often these types are defined inductively: going from contractions upwards,
+%   every new level has paths in the level before (i.e. the paths of a proposition
+%   are all contractions).
+%   We can also proceed further upwards, into non-set types.
+% \end{definition}
+% \begin{definition}[Unique Membership] \label{uniq-memb-def}
+%   Unique list membership is defined in terms of list membership: it is a
+%   contraction of it.
+% \end{definition}
 
@@
   \end{agdalisting}
+  The only difference between manifest Bishop finiteness and split enumerability
+  is the membership term: here we require unique membership
+  (\AgdaFunction{\ensuremath{\in!}}), rather than simple membership
+  (\AgdaFunction{\(\in\)}).
 \end{definition}
-The only difference between manifest Bishop finiteness and split enumerability
-is the membership term: here we require unique membership (\(\inunique\)),
-rather than simple membership (\(\in\)).
 
@@
 interpretation of Bishop finiteness, which we have called cardinal finiteness in
-this paper.
-The ``manifest'' refers to the fact that we have a concrete, non-truncated list
-of the elements in the proof.
+this paper: this version of the proof is ``manifest'' because we have a
+concrete, non-truncated list of the elements in the proof.
 
@@
 Both predicates imply the other.
-\begin{lemma} \label{manifest-bishop-to-split-enum}
-  Any manifest Bishop finite type is split enumerable.
-\end{lemma}
-\begin{proof}
-  To construct a proof of split enumerability from one of manifest Bishop
+
+Going from manifest Bishop finiteness is relatively straightforward:
+to construct a proof of split enumerability from one of manifest Bishop
 finiteness, it suffices to convert a proof of \(x \inunique \mathit{xs}\) to
@@
 is simply the \(\text{fst}\) function.
-\end{proof}
 
+Going the other direction takes significantly more work.
 \begin{lemma} \label{split-enum-to-manifest-bishop}
@@
 \end{lemma}
-This proof takes significantly more work.
-The ``unique membership'' condition in
+We will only sketch the proof here:
+the ``unique membership'' condition in
 \(\AgdaDatatype{\ensuremath{\mathcal{B}}}\) means that we are not permitted
@@
 \subsection{From Manifest Bishop Finiteness to Equivalence}
+
 We have seen that split enumerability was in fact a split-surjection in
@@
 in disguise.
+\begin{multicols}{2}
+  \begin{agdalisting} \label{is-equiv-def}
+    \ExecuteMetaData[agda/Snippets/Equivalence.tex]{is-equiv-def}
+  \end{agdalisting} \columnbreak
+  \begin{agdalisting} \label{equiv-def}
+    \ExecuteMetaData[agda/Snippets/Equivalence.tex]{equiv-def}
+  \end{agdalisting}
+\end{multicols}
 
@@
   ``default'' definition of equivalences.
-  \begin{agdalisting} \label{is-equiv-def}
-    \ExecuteMetaData[agda/Snippets/Equivalence.tex]{is-equiv-def}
-  \end{agdalisting} \vspace{-.5\baselineskip}
-  \begin{agdalisting} \label{equiv-def}
-    \ExecuteMetaData[agda/Snippets/Equivalence.tex]{equiv-def}
-  \end{agdalisting}
 \end{definition}
@@
   of the natural numbers.
-  \begin{equation}
-  \end{equation}
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Cardinality/Finite/ManifestBishop.tex]{bishop-is-equiv-type}
+  \end{agdalisting}
 \end{lemma}
-\begin{proof} \todo{Convert to Agda}
-  \begin{align*}
-     \AgdaDatatype{\ensuremath{\mathcal{B}}}(A) &
-    \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , x \inunique \mathit{xs}
-    && \text{def.~\ref{bish-def} }(\AgdaDatatype{\ensuremath{\mathcal{B}}})
-    \\
-    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{isContr}(x \in \mathit{xs})
-    && \text{eqn.~\ref{uniq-memb-def} } (\inunique)
-    \\
-    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{isContr}(\text{fib}_{\text{snd}(\mathit{xs})}(x))
-    && \text{eqn.~\ref{container-membership} } (\in)
-    \\
-    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
-    && \text{eqn.~\ref{is-equiv-def} (isEquiv)}
-    \\
-    & \simeq \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \AgdaDatatype{Fin} \rrbracket (A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
-    && \text{def.~\ref{list-def} } (\AgdaDatatype{List})
-    \\
-    & \simeq \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \AgdaDatatype{Fin}\;n) , A) , \text{isEquiv}(\text{snd}(\mathit{xs}))
-    && \text{eqn.~\ref{container-interp} } (\llbracket \wc \rrbracket)
-    \\
-    & \simeq \Sigma (n : \mathbb{N}) , \Sigma (f : \AgdaDatatype{Fin}\;n \rightarrow A) , \text{isEquiv}(f)
-    && \text{Reassociation of } \Sigma
-    \\
-    & \simeq \Sigma (n : \mathbb{N}) , ( \AgdaDatatype{Fin}\;n \simeq A )
-    && \text{eqn.~\ref{equiv-def} } (\simeq) \; \qedhere
-  \end{align*}
+
+\begin{proof} \let\qed\relax \vspace{-1\baselineskip}
+  \begin{minipage}[t]{.7\textwidth}\vspace{-\baselineskip}
+    \begin{agdalisting*}
+      \ExecuteMetaData[agda/Cardinality/Finite/ManifestBishop.tex]{bishop-is-equiv}
+    \end{agdalisting*}
+  \end{minipage}
+  \begin{minipage}[t]{.19\textwidth} \setstretch{1.1}
+
+    Def.~\ref{bish-def} (\AgdaDatatype{\ensuremath{\mathcal{B}}}) \\
+    Eqn.~\ref{uniq-memb-def} (\AgdaDatatype{\ensuremath{\in!}}) \\
+    Eqn.~\ref{container-membership} (\AgdaDatatype{\ensuremath{\in}}) \\
+    Eqn.~\ref{is-equiv-def}  \\
+    Eqn.~\ref{list-def} (\(\AgdaDatatype{List}\)) \\
+    Eqn.~\ref{container-interp}  \\
+    Reassociation \\
+    Eqn.~\ref{equiv-def}
+  \end{minipage}
 \end{proof}
+
+% \begin{proof} \todo{Convert to Agda}
+%   \begin{align*}
+%      \AgdaDatatype{\ensuremath{\mathcal{B}}}(A) &
+%     \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , x \inunique \mathit{xs}
+%     && \text{def.~\ref{bish-def} }(\AgdaDatatype{\ensuremath{\mathcal{B}}})
+%     \\
+%     & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{isContr}(x \in \mathit{xs})
+%     && \text{eqn.~\ref{uniq-memb-def} } (\inunique)
+%     \\
+%     & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{isContr}(\text{fib}_{\text{snd}(\mathit{xs})}(x))
+%     && \text{eqn.~\ref{container-membership} } (\in)
+%     \\
+%     & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
+%     && \text{eqn.~\ref{is-equiv-def} (isEquiv)}
+%     \\
+%     & \simeq \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \AgdaDatatype{Fin} \rrbracket (A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
+%     && \text{def.~\ref{list-def} } (\AgdaDatatype{List})
+%     \\
+%     & \simeq \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \AgdaDatatype{Fin}\;n) , A) , \text{isEquiv}(\text{snd}(\mathit{xs}))
+%     && \text{eqn.~\ref{container-interp} } (\llbracket \wc \rrbracket)
+%     \\
+%     & \simeq \Sigma (n : \mathbb{N}) , \Sigma (f : \AgdaDatatype{Fin}\;n \rightarrow A) , \text{isEquiv}(f)
+%     && \text{Reassociation of } \Sigma
+%     \\
+%     & \simeq \Sigma (n : \mathbb{N}) , ( \AgdaDatatype{Fin}\;n \simeq A )
+%     && \text{eqn.~\ref{equiv-def} } (\simeq) \; \qedhere
+%   \end{align*}
+% \end{proof}
 This proof is almost identical\footnotemark to the proof for

### 2020-06-30 sections/finiteness-predicates.tex
finished manifest bishop section

@@
 to construct a proof of split enumerability from one of manifest Bishop
-finiteness, it suffices to convert a proof of \(x \inunique \mathit{xs}\) to
-one of \(x \in \mathit{xs}\), for all \(x\) and \(\mathit{xs}\).
-Since \(\inunique\) is defined as a contraction of \(\in\), such a conversion
-is simply the \(\text{fst}\) function.
+finiteness, it suffices to convert a proof of \(x\;\AgdaFunction{\ensuremath{\in!}}\;\mathit{xs}\) to
+one of \(x\;\AgdaFunction{\ensuremath{\in}}\;\mathit{xs}\), for all \(x\) and \(\mathit{xs}\).
+Since \AgdaFunction{\(\in!\)} is defined as a contraction of \AgdaFunction{\(\in\)}, such a conversion
+is simply the \AgdaField{fst} function.
 
@@
 \subsection{From Manifest Bishop Finiteness to Equivalence}
-
 We have seen that split enumerability was in fact a split-surjection in
@@
 in disguise.
+We define equivalences as contractible maps \cite[definition 4.4.1]{hottbook}:
 \begin{multicols}{2}
@@
 \end{multicols}
-
-\begin{definition}[Equivalences] \label{equivalences}
-  We will take contractible maps \cite[definition 4.4.1]{hottbook} as our
-  ``default'' definition of equivalences.
-\end{definition}
 \begin{lemma} \label{bishop-equiv}
@@
 \end{lemma}
-
 \begin{proof} \let\qed\relax \vspace{-1\baselineskip}
@@
 
-% \begin{proof} \todo{Convert to Agda}
-%   \begin{align*}
-%      \AgdaDatatype{\ensuremath{\mathcal{B}}}(A) &
-%     \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , x \inunique \mathit{xs}
-%     && \text{def.~\ref{bish-def} }(\AgdaDatatype{\ensuremath{\mathcal{B}}})
-%     \\
-%     & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{isContr}(x \in \mathit{xs})
-%     && \text{eqn.~\ref{uniq-memb-def} } (\inunique)
-%     \\
-%     & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{isContr}(\text{fib}_{\text{snd}(\mathit{xs})}(x))
-%     && \text{eqn.~\ref{container-membership} } (\in)
-%     \\
-%     & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
-%     && \text{eqn.~\ref{is-equiv-def} (isEquiv)}
-%     \\
-%     & \simeq \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \AgdaDatatype{Fin} \rrbracket (A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
-%     && \text{def.~\ref{list-def} } (\AgdaDatatype{List})
-%     \\
-%     & \simeq \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \AgdaDatatype{Fin}\;n) , A) , \text{isEquiv}(\text{snd}(\mathit{xs}))
-%     && \text{eqn.~\ref{container-interp} } (\llbracket \wc \rrbracket)
-%     \\
-%     & \simeq \Sigma (n : \mathbb{N}) , \Sigma (f : \AgdaDatatype{Fin}\;n \rightarrow A) , \text{isEquiv}(f)
-%     && \text{Reassociation of } \Sigma
-%     \\
-%     & \simeq \Sigma (n : \mathbb{N}) , ( \AgdaDatatype{Fin}\;n \simeq A )
-%     && \text{eqn.~\ref{equiv-def} } (\simeq) \; \qedhere
-%   \end{align*}
-% \end{proof}
 This proof is almost identical\footnotemark to the proof for

### 2020-06-30 sections/finiteness-predicates.tex
down to definition of cardinal finiteness

@@
 
-This proof is almost identical\footnotemark to the proof for
+This proof is almost identical to the proof for
 lemma~\ref{split-enum-is-split-surj}: it reveals that
@@
 relation-based ones.
-\footnotetext{
-  Unfortunately in our formalisation this proof cannot be a single line: for
-  performance reasons \(\simeq\) is defined as a record type with eta-equality
-  disabled, instead of the definition here which uses \(\Sigma\).
-}
 
@@
 
-To remedy the problem, we will use propositional truncation
-(def.~\ref{prop-trunc}).
-
-\begin{definition}[Higher Inductive Types] \label{HITs}
+To remedy the problem, we will use the following type:
+\begin{agdalisting} \label{prop-trunc}
+  \ExecuteMetaData[agda/Snippets/PropTrunc.tex]{prop-trunc-def}
+\end{agdalisting}
+This is a \emph{higher inductive type}.
 Normal inductive types have \emph{point} constructors: constructors which
 construct values of the type.
-  Higher Inductive Types (HITs) also have \emph{path} constructors: ways to
-  construct paths in the type.
-\end{definition}
-\begin{definition}[Propositional Truncation] \label{prop-trunc}
-  The type \(\lVert A \rVert\) on some type \(A\) is a propositionally truncated
-  proof of \(A\) \cite[3.7]{hottbook}.
-  In other words, it is a proof that some \(A\) exists, but it does not tell you
-  \emph{which} \(A\).
+The first constructor here (\AgdaInductiveConstructor{\ensuremath{\lvert \_
+    \rvert}}), or the constructor \AgdaInductiveConstructor{true} for
+\AgdaDatatype{Bool}, are both ``point'' constructors.
 
-  It is defined as a Higher Inductive Type:
+What makes this type higher inductive is that it also has \emph{path}
+constructors: constructors which add new equalities to the type. 
+The \AgdaInductiveConstructor{squash} constructor here says that all elements of
+\proptrunc{A}
+are equal, regardless of what \(A\) is.
+In this way it allows us to propositionally truncate types, turning
+information-containing proofs into mere propositions.
+Put another way, a proof of type \proptrunc{A}
+is a proof that some \(A\) exists, without revealing \emph{which} \(A\).
+
+To actually use values of this type we have the following two eliminators:
+\begin{multicols}{2} \null \vfill
   \begin{agdalisting}
-    \ExecuteMetaData[agda/Snippets/PropTrunc.tex]{prop-trunc-def}
-  \end{agdalisting}
-  We will use two eliminators from \(\lVert A \rVert\) in this paper.
-  \begin{enumerate}
-  \item \label{elim-prop-prop} For any function \(A \rightarrow B\), where
-    \(\text{isProp}(B)\), we have a function \(\lVert A \rVert \rightarrow B\).
-  \item \label{elim-prop-coh} We can eliminate from \(\lVert A \rVert\) with a
-    function \(f : A \rightarrow B\) iff \(f\) ``doesn't care'' about the
-    choice of \(A\):
-    \[\Pi {(x , y : A)} , f(x) \equiv f(y) \]
-    Formally speaking, \(f\) needs to be ``coherently constant''
-    \cite{krausGeneralUniversalProperty2015}, and \(B\) needs to be an
-    \(n\)-type for some finite \(n\).
-  \end{enumerate}
-\end{definition}
+    \ExecuteMetaData[agda/HITs/PropositionalTruncation/Display.tex]{rec-prop-trunc}
+  \end{agdalisting} \vfill \null \columnbreak \null \vfill
+  \begin{agdalisting} \label{elim-prop-coh}
+    \ExecuteMetaData[agda/HITs/PropositionalTruncation/Display.tex]{rec-prop-trunc-set}
+  \end{agdalisting} \vfill \null
+\end{multicols} \noindent
+While these are not the only eliminators available for \proptrunc{A}, they are
+the only ones we will use in this paper.
+The first says we can eliminate into any proposition: interestingly, this allows
+us to define a monad instance for \AgdaDatatype{\(\lVert \_ \rVert\)}, meaning
+we can use things like do-notation.
+The second says we can eliminate into a set as long as the function we use
+doesn't care about which value it's given: formally, \(f\) in this example has
+to be ``coherently constant'' \cite{krausGeneralUniversalProperty2015}.
+
+With this, we can define cardinal finiteness:
 \begin{definition}[Cardinal Finiteness]
@@
 
-  We define it as a HIT (definition~\ref{HITs}).
+  We define it as a HIT.
   The first two constructors are point constructors, giving ways to create

### 2020-06-30 sections/finiteness-predicates.tex
changed naming of propositional truncation eliminators

@@
 \end{definition}
+\subsection{Deriving Uniquely-Determined Quantities}
 At first glance, it might seem that we lose any useful properties we could
@@
 def.~\ref{prop-trunc}, we need only show that the output is uniquely determined.
-\subsection{Deriving Uniquely-Determined Quantities}
+
 The following two lemmas are proven in
@@
   Given a cardinally finite type, we can derive the type's cardinality, as well
-  as a propositionally truncated proof of equivalence with \(\textbf{Fin}\)s of
+  as a propositionally truncated proof of equivalence with \AgdaDatatype{Fin}s of
   the same cardinality.
-  \begin{equation}
-    \mathcal{C}(A) \rightarrow \Sigma {(n : \mathbb{N})} , \lVert \textbf{Fin}(n) \simeq A \rVert
-  \end{equation}
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{cardinality-is-unique}
+  \end{agdalisting}
 \end{lemma}

### 2020-06-30 sections/finiteness-predicates.tex
some code for cardinality etc

@@
 \end{agdalisting}
-\subsection{Decidable Equality}
+\subsection{Decidable Equality} \todo{Should this be a lemma rather than a subsection?}
 One thing that characterises all split enumerable types is that they are all
@@
 \end{lemma}
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{cardinality-is-unique-impl}
+\end{agdalisting}
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{card-isSet}
+\end{agdalisting}
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{alg}
+\end{agdalisting}
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{const-alg}
+\end{agdalisting}
+
 \begin{proof}
+
   Let \(A\) be a cardinally-finite type, with proof \(F : \mathcal{C}(A)\).

### 2020-06-30 sections/finiteness-predicates.tex
done first proof in cardinal section

@@
 \end{lemma}
+\begin{proof}
+  The high-level overview of our proof is as follows:
   \begin{agdalisting}
@@
   \end{agdalisting}
+  It is the composition of two operations: first, with
+  \AgdaFunction{\ensuremath{\lVert \text{map} \rVert}}, we change the truncated
+  proof of manifest bishop finiteness to a proof of equivalence with fin.
+
+  Then we use the eliminator from Equation~\ref{elim-prop-coh} with three
+  parameters.
+  The first simply proves that that the output is a set:
   \begin{agdalisting}
@@
   \end{agdalisting}
+  The second is the function we apply to the truncated value:
   \begin{agdalisting}
@@
   \end{agdalisting}
+  And the third is a proof that that function is itself coherently constant:
   \begin{agdalisting}
@@
 
-\begin{proof}
-
-  Let \(A\) be a cardinally-finite type, with proof \(F : \mathcal{C}(A)\).
-  Our task is to extract a natural number \(n : \mathbb{N}\) representing the
-  cardinality of \(A\), and a propositionally-truncated proof that \(A\) is
-  equivalent to \(\AgdaDatatype{Fin}\;n\).
-
-  Extracting the second component of the pair is trivial, as it itself is
-  truncated.
-  We will now focus on extracting the cardinality.
-
-  Without the propositional truncation, \(\text{fst}\) would suffice for this
-  task.
-  Given that the pair is hidden under the truncation, then, we need a way to
-  convert a function \(f : A \rightarrow B\) to \(g : \lVert A \rVert
-  \rightarrow B\).
-  This is precisely what eliminator \ref{elim-prop-coh} gives us.
-  For our case, we need to show the following:
+  The tricky part of the proof is \AgdaFunction{const-alg}: here we need to show
+  that \AgdaFunction{alg} returns the same value no matter its input.
+  That output is a pair, the first component of which is the cardinality, and
+  the second the truncated equivalence proof.
+  The truncated proofs in the output are trivially equal by the truncation, so
+  our obligation now has been reduced to:
   \begin{equation}
-    \frac{(n : \mathbb{N}) \; \; \; (p : \AgdaDatatype{Fin}\;n \simeq A) \; \; \;
-          (m : \mathbb{N}) \; \; \; (q : \AgdaDatatype{Fin}(m) \simeq A)
+    \frac{(n :\;\AgdaDatatype{\ensuremath{\mathbb{N}}}) \; \; \; (p : \AgdaDatatype{Fin}\;n\;\AgdaFunction{\ensuremath{\simeq}}\;A) \; \; \;
+      (m :\;\AgdaDatatype{\ensuremath{\mathbb{N}}}) \; \; \; (q : \AgdaDatatype{Fin}\;m\;\AgdaFunction{\ensuremath{\simeq}}\;A)
     }{
-          n \equiv m
+      n\;\AgdaFunction{\ensuremath{\equiv}}\;m
     }
   \end{equation}
-  Immediately we can construct the following term:
-  \begin{equation}
-    \begin{alignedat}{3}
-      \AgdaDatatype{Fin}\;n & \simeq A && (p) \\
-                      & \simeq \AgdaDatatype{Fin}(m) && (q)
-    \end{alignedat}
-  \end{equation}
-  Given univalence we have \(\AgdaDatatype{Fin}\;n \equiv \AgdaDatatype{Fin}(m)\),
+  Given univalence we have \(\AgdaDatatype{Fin}\;n \;\AgdaFunction{\ensuremath{\equiv}}\; \AgdaDatatype{Fin}\;m\),
   and the rest of our task is to prove:
   \begin{equation}
-    \frac{\AgdaDatatype{Fin}\;n \equiv \AgdaDatatype{Fin}(m)}{n \equiv m}
+    \frac{\AgdaDatatype{Fin}\;n\;\AgdaDatatype{\ensuremath{\equiv}}\;\AgdaDatatype{Fin}\;m}{n \equiv m}
   \end{equation}

### 2020-06-30 sections/finiteness-predicates.tex
done second proof in cardinal section

@@
 \end{agdalisting}
-\subsection{Decidable Equality} \todo{Should this be a lemma rather than a subsection?}
+\subsection{Decidable Equality}
 One thing that characterises all split enumerable types is that they are all
@@
 enumerable types.
-
+\begin{lemma} \todo{Don't know if this should be a lemma or just prose}
+  Split enumerability implies decidable equality.
+\end{lemma}
+\begin{proof}
 To prove that split enumerability implies decidable equality we'll take a quick
@@
 \end{agdalisting}
+\end{proof}
 \section{Manifest Bishop Finiteness} \label{manifest-bishop-finiteness}
@@
 \emph{contraction} \cite[definition 3.11.1]{hottbook}.
-\begin{agdalisting}
+\begin{agdalisting} \label{isContr}
   \ExecuteMetaData[agda/Snippets/Introduction.tex]{isContr}
@@
 
+What we want is a proof of finiteness that is a proposition.
+\begin{agdalisting} \label{isProp}
+  \ExecuteMetaData[agda/Snippets/Introduction.tex]{isProp}
+\end{agdalisting}
+The mere propositions are one homotopy level higher than the contractions
+(Equation~\ref{isContr}), the types for which all values are equal to some value.
+They represent the types for which all values are equal, or, the types
+isomorphic to \agdabot\;or \agdatop.
+You can also define propositions in terms of the contractions: propositions are
+the types whose paths are contractions.
+Soon (Equation~\ref{isSet}) we will see the next homotopy level, which are
+defined in terms of the propositions.
+
+Despite now knowing the precise property we want our finiteness predicate to
+have, we're not much closer to achieving it.
 To remedy the problem, we will use the following type:
@@
 \begin{multicols}{2} \null \vfill
-  \begin{agdalisting}
+  \begin{agdalisting} \label{elim-prop}
     \ExecuteMetaData[agda/HITs/PropositionalTruncation/Display.tex]{rec-prop-trunc}
@@
 derive from \(\AgdaDatatype{\ensuremath{\mathcal{B}}}\).
-Luckily, this is not the case: by eliminator \ref{elim-prop-coh} of
-def.~\ref{prop-trunc}, we need only show that the output is uniquely determined.
+Luckily, this is not the case: we will show here how to use the eliminators from
+propositional truncation (Equations \ref{elim-prop} and \ref{elim-prop-coh}) to
+use to of our already-proven derivations from \AgdaFunction{\ensuremath{\mathcal{B}}}
+on \AgdaFunction{\ensuremath{\mathcal{C}}}. \todo{rephrase?}
 
@@
 
-In order to prove that cardinal finiteness implies decidable equality, we will
-need to show that decidable equality itself is a proposition.
-In doing that we will use the following lemma:
-\begin{lemma} \label{prop-refute}
-  We can ``refute'' a propositionally-truncated proof of some proposition with a
-  proof that the non-truncated proposition is false.
-  \begin{equation}
-    \frac{\neg A \; \; \; \lVert A \rVert}{\bot}
-  \end{equation}
-\end{lemma}
-\begin{proof}
-  We know we can eliminate from any value of type \(\lVert A \rVert\) into some
-  \(B\) with a function \(A \to B\) if \(B\) is a proposition.
-  That's precisely what we do in this case: \(\neg A\) is a function of type \(A
-  \to \bot\), and we know that \(\bot\) is a proposition.
-\end{proof}
+Another proof that carries over from manifest Bishop finiteness is decidable
+equality.
+This demonstrates an interesting aspect of propositional truncation: it can
+contain computational information.
+That is in contrast to other ways to ``truncate'' types: \(\neg \neg P\), for
+instance, is a way to provide a ``proof'' of \(P\) without revealing anything
+about \(P\) in MLTT.
+No matter how much we prove that a function from \(P\) doesn't care about which
+\(P\) it got, though, we can never extract any kind of algorithm or computation
+from \(\neg \neg P\). \todo{Is this true? It certainly seems true...}
 
+Here we'll briefly show how to extract ``computation'' from under a
+propositional truncation.
 \begin{lemma} \label{cardinal-finite-discrete}
@@
 \begin{proof}
-  Since we can already derive decidable equality from a proof of manifest Bishop
-  finiteness, it suffices to show that decidable equality is itself a
-  proposition.
-  \begin{equation}
-    \text{isProp}(\Pi(x, y : A) , \mathbf{Dec}(x \equiv y))
-  \end{equation}
-  First, it is clear that \(x \equiv y\) is a proposition: since the type \(A\)
-  has decidable equality, by Hedburg's theorem it is a set,
-  meaning precisely that \(x \equiv y\) is a proposition.
+We can extract plenty of computation from under propositional truncation as long
+as we prove the thing we're extracting is a proposition (Equation~\ref{elim-prop}). 
+Our task, in other words, is to prove the following:
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Relation/Nullary/Discrete/Properties.tex]{is-prop-discrete}
+\end{agdalisting}
 
-  Secondly, we know that any decision over a proposition is itself a
-  proposition.
-  For any two terms \(x, y: \mathbf{Dec}(A)\) we cannot have the case that one
-  is a yes decision and the other is no: from that we could derive \(\bot\).
-  If both are no then they are both equal since \(A \rightarrow \bot\) is a
-  proposition through function extensionality.
-  And finally if both are yes then we know they must be equal because the type
-  decided over is itself a proposition.
+To show that this type is a proposition we must show that any two given members
+of the type are equal, i.e. we are given two proofs of decidable equality on
+\(A\) and we must show that they are equal.
+Remember that \(\AgdaFunction{Discrete}\;A\) is a function of two arguments
+returning a \AgdaDatatype{Dec} of whether those two arguments are equal or not.
+\todo{ref for wherever this definition was}
+By function extensionality, to prove that that is a proposition we have to prove
+that \(\AgdaDatatype{Dec}\;(\agdaequiv{x}{y})\) is a proposition.
+This proof requires that we show that the payload of each of the constructors
+(\AgdaInductiveConstructor{yes} and \AgdaInductiveConstructor{no}) are
+propositions.
+\AgdaInductiveConstructor{no}'s payload is
+\(\agdaequiv{x}{y}\rightarrow\agdabot\), which is a
+proposition because \agdabot\; is a proposition.
 
-  Finally, since we know that \(\mathbf{Dec}(x \equiv y)\) is a proposition, we
-  can derive that \(\Pi(x, y : A) , \mathbf{Dec}(x \equiv y)\) is a proposition
-  (through function extensionality), proving our goal.
+\AgdaInductiveConstructor{yes} is a little more interesting: its payload is
+\agdaequiv{x}{y}.
+How can we prove that the path between \(x\) and \(y\) is a proposition?
+It turns out that there is a class of types for which all paths are
+propositions: the \emph{sets}.
+\begin{agdalisting} \label{isSet}
+  \ExecuteMetaData[agda/Snippets/Introduction.tex]{isSet}
+\end{agdalisting}
+This is the next homotopy level up from the propositions (Equation~\ref{isProp}).
+More importantly, there is an important theorem relating to sets which
+\emph{also} relates to decidable equality: Hedberg's theorem
+\cite{hedbergCoherenceTheoremMartinLof1998}.
+\begin{theorem}[Hedberg's Theorem]
+  Any type with decidable equality is a set.
+\end{theorem} \noindent
+And of course we know that \(A\) here has decidable equality: we were just given
+two proofs of that fact at the beginning of this proof!
 \end{proof}

### 2020-06-30 sections/finiteness-predicates.tex
finished sort proof

@@
 \begin{equation}
-  \mathcal{C}(A) \rightarrow \AgdaDatatype{\ensuremath{\mathcal{B}}}(A)
+  \agdacal{C}\;A\rightarrow\agdacal{B}\;A
 \end{equation}
@@
 It turns out the answer is no!
-\begin{lemma}
-  \begin{equation}
-    \neg (\Sigma(A : \mathbf{Type}) , \mathcal{C}(A) \times \neg \AgdaDatatype{\ensuremath{\mathcal{B}}}(A))
-  \end{equation}
-\end{lemma}
-\begin{proof}
-  We will actually prove a slightly more general statement.
-  For any type \(A\), the following holds:
-  \begin{equation}
-    \neg (\lVert A \rVert \times \neg A)
-  \end{equation}
-  The solution becomes more clear if we write out the definition of \(\neg\):
-  \begin{equation}
-    \frac{\lVert A \rVert \;\;\; A \rightarrow \bot }{\bot}
-  \end{equation}
-  We clearly need to apply a function of type \(A \rightarrow \bot\) to a value
-  of type \(\lVert A \rVert\).
-  Luckily, this is permissible, as \(\bot\) is a mere proposition.
-\end{proof}
+We can prove the following statement:
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{no-gap-card-bishop}
+\end{agdalisting}
+We use the fact that \agdabot\; is itself a proposition to eliminate from under
+the truncation.
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{no-gap-card-bishop-proof}
+\end{agdalisting}
 
+This tells us that manifest bishop finiteness, cardinal finiteness, and split
+enumerability all refer to the same class of types.
 \subsection{Going from Cardinal Finiteness to Manifest Bishop Finiteness}
-\begin{lemma} \label{manifest-bishop-to-cardinal}
-  Any manifest Bishop finite type is cardinal finite.
-\end{lemma}
+We know of course that we can convert any proof of manifest Bishop finiteness to
+a proof of Cardinal finiteness: it's just the truncation function
+\AgdaInductiveConstructor{\(\lvert \_ \rvert\)}.
+It's the other direction which presents a difficulty:
 \begin{theorem} \label{cardinal-to-manifest-bishop}
@@
 we only give its sketch here.
-First, note that we actually convert to manifest enumerability first: this can
-be converted to split enumerability with decidable equality, which is provided
-by cardinal finiteness.
 
-Next, we define permutations.
-\begin{definition}[List Permutations]
+Our strategy will be to \emph{sort} the support list of the proof for Bishop
+finiteness, and then prove that the sorting function is coherently constant,
+thereby satisfying the eliminator in Equation~\ref{elim-prop-coh}.
+We need to show, in other words, that sorting two support lists from proofs of
+manifest Bishop finiteness on the same type with the same order always returns
+the same result.
+The sort we use is a standard insertion sort:
+\begin{multicols}{2}
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Data/List/Sort.tex]{insert}
+  \end{agdalisting} \columnbreak \null \vfill \null \vfill
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Data/List/Sort.tex]{insertion-sort}
+  \end{agdalisting} \vfill \null
+\end{multicols} \vspace{-\baselineskip} \noindent
+And we prove that \AgdaFunction{sort} produces a list which is sorted, and a
+permutation of its input.
+\begin{multicols}{2}
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Data/List/Sort.tex]{sort-sorts}
+  \end{agdalisting} \columnbreak
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Data/List/Sort.tex]{sort-perm}
+  \end{agdalisting}
+\end{multicols} \vspace{-\baselineskip}
+We've introduced permutations here: we take their definition from
+\cite{danielssonBagEquivalenceProofRelevant2012}.
 Two lists are permutations of each other if their membership proofs are all
-  equivalent\footnotemark \cite{danielssonBagEquivalenceProofRelevant2012}.
-  \begin{equation}
-    \mathit{xs} \leftrightsquigarrow \mathit{ys} = \Pi {(x : A)} , x \in \mathit{xs} \simeq x \in \mathit{ys}
-  \end{equation}
-\end{definition}
+equivalent.
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Data/List/Relation/Binary/Permutation.tex]{perm-def}
+\end{agdalisting}
+This definition fits particularly well for two reasons: first, it is defined on
+containers generically, which fits well with our finiteness predicates.
+Secondly, it is extremely straightforward to show that the support lists of any
+two proofs of manifest Bishop finiteness must be permutations of each other:
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{perm-bish}
+\end{agdalisting}
 
-\footnotetext{
-  The definition in \cite{danielssonBagEquivalenceProofRelevant2012} and our
-  formalisation is slightly different: we say permutations are lists with
-  \emph{isomorphic} membership proofs.
-  The distinction, as it happens, does not affect our work here.
-}
-
-Next, we define a sort function which relies on the provided total order.
-We further prove the following fact about this sort function:
-\begin{equation}
-  \Pi(\mathit{xs}, \mathit{ys} : \AgdaDatatype{List}(A)) , \mathit{xs} \leftrightsquigarrow \mathit{ys} \rightarrow \text{sort}(\mathit{xs}) \equiv \text{sort}(\mathit{ys})
-\end{equation}
-
-Next, notice that the support lists of any two proofs of manifest Bishop
-finiteness must be permutations of each other.
-This will allow us to sort the support list of a proof of cardinal finiteness in
-a coherently constant (definition~\ref{prop-trunc},
-eliminator~\ref{elim-prop-coh}) way, pulling the support list out from the
-truncation.
-The cover proof emerges naturally from the definition of the permutation.
+Almost all of the pieces are in place now: we know that the support lists of all
+proofs of \(\agdacal{B}\;A\) are permutations of each other, and we know that
+\AgdaFunction{sort} returns a sorted permutation of its input.
+The final piece of the puzzle is the following:
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Data/List/Sort.tex]{sorted-perm-eq}
+\end{agdalisting}
+If two sorted lists are both permutations of each other they must be equal.
+Connecting up all the pieces we get the following:
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Data/List/Sort.tex]{perm-invar}
+\end{agdalisting}
+Because we know that all support lists of \(\agdacal{B}\;A\) are permutations of
+each other this is enough to prove that \AgdaFunction{sort} is coherently
+constant, and therefore can eliminate from within a truncation.
+The second component of the output pair (the cover proof) follows quite
+naturally from the definition of permutations.
 \section{Manifest Enumerability} \label{manifest-enumerability}

### 2020-07-01 sections/finiteness-predicates.tex
bit more (up to manifest enum instance for circle)

@@
 
-As an example type which may be considered finite in some circumstances, but
-doesn't conform to any of the definitions we have seen so far, consider the
-circle.
-\begin{definition}[\(S^1\)] \label{circle-def}
-  The circle, \(S^1\), can be represented in HoTT as a higher inductive type.
-  \begin{agdalisting}
+What we're looking for is a type which is in some sense finite, but doesn't
+conform to any of the predicates we've seen so far.
+The \emph{circle} is such a type:
+\begin{agdalisting}\label{circle-def}
   \ExecuteMetaData[agda/Snippets/Circle.tex]{circle-def}
 \end{agdalisting}
-  The presence of the \AgdaInductiveConstructor{loop} constructor is what
-  precludes this type from being a set.
-  In sets, all paths are equal: in this type, \AgdaInductiveConstructor{loop} is a
-  path which is not equal to the usual identity path.
-\end{definition}
+The thing that this type has which precludes it from being, say, split
+enumerable, is its \emph{higher homotopy structure}. \todo{is this the right phrase?}
 
-We suggest that this type cannot be proven to conform to any of the finiteness
-predicates we have seen so far (split enumerability, manifest Bishop finiteness,
-or Cardinal finiteness).
-To provide some better evidence for this fact (other than ``we tried quite hard
-and couldn't manage it'') remember Hedberg's theorem
-\cite{hedbergCoherenceTheoremMartinLof1998}:
-\begin{theorem}[Hedberg's Theorem]
-  Every discrete type is a set.
-\end{theorem}
-Since each finiteness predicate we have seen so far \emph{does} imply decidable
-equality, we know that the cardinal finite types can \emph{only} be sets, ruling
-out the circle.
+So far we have seen three levels of homotopy structure: the contractions
+(Equation~\ref{isContr}), the propositions (Equation~\ref{isProp}), and the sets
+(Equation~\ref{isSet}).
+You may have noticed the pattern that each new level is generated by saying its
+paths are members of the previous level; if we apply that pattern again, we get
+to the next homotopy level: the groupoids.
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Snippets/Introduction.tex]{isGroupoid}
+\end{agdalisting}
+These types do not necessarily have unique identity proofs: there is more than
+one value which can inhabit the type \agdaequiv{x}{y}.
+The circle is one of the simplest examples of non-set groupoids: the constructor
+\AgdaInductiveConstructor{loop} is the extra path in the type which isn't the
+identity path.
 
-So our tour through the finiteness predicates takes us next to one which allows
-the circle to be called finite: manifest enumerability.
+We now need to recall two facts: first, Hedberg's theorem tells us that every
+discrete type is a set.
+Second, every finiteness predicate we've seen thus far implies decidable
+equality.
+From this it's clear that all of the previous predicates are restricted to sets,
+and can't include types like the circle.
+
+But the type certainly \emph{seems} finite!
+In order to explore the ``restrictiveness'' axis in
+Figure~\ref{finite-classification}, then, we'll need to construct a predicate
+which admits the circle.
+Manifest enumerability is one such predicate.
 \begin{definition}[Manifest Enumerability]
@@
 \end{lemma}
-\begin{proof}
-  We proceed by case-analysis over the decision over \(A\).
-  In the case where \(A\) is proven, we are done.
-  In the case where \(A\) is disproven, we use lemma~\ref{prop-refute} to
-  derive impossibility.
-\end{proof}
+\todo{proof}
+% \begin{proof} \
+%   We proceed by case-analysis over the decision over \(A\).
+%   In the case where \(A\) is proven, we are done.
+%   In the case where \(A\) is disproven, we use lemma~\ref{prop-refute} to
+%   derive impossibility.
+% \end{proof}
 

### 2020-07-01 sections/finiteness-predicates.tex
finished manifest enumerability

@@
 
-And indeed this means that the circle is manifestly enumerable:
-\begin{lemma}
-  The circle \(S^1\) is manifestly enumerable.
-\end{lemma}
-\begin{proof}
-  The support list firstly is a list containing the point constructor for the
-  circle.
-  Since the cover proof is truncated, we need only consider the point
-  constructors of the circle: as such, the cover proof is essentially the same
-  as the one for \(\AgdaDatatype{\ensuremath{\mathcal{E}!}}(\top)\).
-\end{proof}
+And indeed this means that the circle is manifestly enumerable. \todo{This uses
+  a bunch of cubical primitives!}
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Cardinality/Finite/ManifestEnumerable.tex]{circle-is-manifest-enum}
+\end{agdalisting}
 \subsection{Surjections}
-We already say that split enumerability was the listed form of a split
-surjection: manifest enumerability is the listed form of a proper surjection.
-\begin{definition}[Surjections] \label{surjections}
-  We define proper surjections (not split surjections) here \cite[definition
-  4.6.1]{hottbook}.
-  \begin{alignat}{3}
-    &\text{surj}(f)             &&\coloneqq \Pi(y : B) , \lVert \text{fib}_f(y) \rVert \label{surj-eqn} \\
-    &A \twoheadrightarrow B     &&\coloneqq \Sigma (f : A \rightarrow B) , \text{surj}(f) \label{surj-arrow-eqn}
-  \end{alignat}
-\end{definition}
-\begin{lemma}
+We already saw that split enumerability was the listed form of a split
+surjection: what we didn't explain was why the word ``split'' was placed before
+surjection.
+In the presence of higher homotopies than sets, split surjections are actually
+\emph{not} a satisfactory definition of surjection.
+And we are most certainly in the presence of higher homotopies: just moments ago
+we were introduced to the circle.
+In these cases, the following definition of surjections is
+preferred~\cite[definition 4.6.1]{hottbook}:
+\begin{multicols}{2}
+  \begin{agdalisting} \label{surj-eqn}
+    \ExecuteMetaData[agda/Function/Surjective/Base.tex]{surjective}
+  \end{agdalisting} \columnbreak
+  \begin{agdalisting} \label{surj-arrow-eqn}
+    \ExecuteMetaData[agda/Function/Surjective/Base.tex]{surjection}
+  \end{agdalisting}
+\end{multicols}
+
+Much in the same way that split enumerability were split surjections, our new
+predicate of manifest enumerability corresponds to the proper surjections.
+\begin{lemma} \todo{to agda}
   Manifest enumerability is equivalent to a surjection from a finite prefix of
@@
 manifest enumerable: we simply truncate the membership proof.
-Going the other was requires us to extract a non-truncated proof from a
-truncated one.
-This proof relies on the following lemma:
-\begin{lemma}
-  We can ``recompute'' a truncated proof given a decision over a proof of the
-  same type.
-  \begin{equation}
-    \frac{\lVert A \rVert \; \; \; \mathbf{Dec}(A)}{A}
-  \end{equation}
-\end{lemma}
-\todo{proof}
-% \begin{proof} \
-%   We proceed by case-analysis over the decision over \(A\).
-%   In the case where \(A\) is proven, we are done.
-%   In the case where \(A\) is disproven, we use lemma~\ref{prop-refute} to
-%   derive impossibility.
-% \end{proof}
-
+Going the other way is more difficult, as we need to extract the membership
+proof from under a truncation.
+We do know what we need, however: the key difference between manifest
+enumerability and split enumerability is that the latter implied decidable
+equality.
+So that's the missing piece we should require in order to go from one to the
+other:
 \begin{lemma} \label{manifest-enum-to-split-enum}
@@
 \end{lemma}
-\begin{proof}
-  The only difference between manifest enumerability and split enumerability is
-  the membership proof: therefor our goal for this proof is to construct a
-  function of the following type:
-  \begin{equation}
-    \lVert x \in \mathit{xs} \rVert \rightarrow x \in \mathit{xs}
-  \end{equation}
-  Given decidable equality over the type of \(x\).
 
-  We do this using the previous recompute lemma: that tells us that all we need
-  to construct is a decision for \(x \in \mathit{xs}\), and it will be able to
-  derive the proof itself.
-  Such a decision procedure is not difficult to construct: for any value \(x\)
-  and list \(\mathit{xs}\), we proceed through the list \(\mathit{xs}\), testing
-  if \(x\) is equal to any of its contents.
-  If it is, we return that we have proven the goal, and that \(x\) is indeed
-  present in \(\mathit{xs}\).
-  Otherwise, we know that \(x\) cannot be in \(\mathit{xs}\) (since we've tested
-  every value), so we return that the goal has been disproven.
-\end{proof}
+Now that we know what extra bit of information we are allowed use in this proof,
+the path forward becomes a little more clear.
+In terms of the actual conversion function, the support list will stay the same,
+and only the return type of the cover proof needs to change: from
+\(\AgdaDatatype{\(\lVert\)}\;x\;\AgdaFunction{\(\in\)}\;\mathit{xs}\;\AgdaDatatype{\(\rVert\)}\)
+to \(x\;\AgdaFunction{\(\in\)}\;\mathit{xs}\).
+
+That can be accomplished with the help of the following function:
+\begin{agdalisting}
+  \ExecuteMetaData[agda/HITs/PropositionalTruncation/Properties.tex]{recompute}
+\end{agdalisting}
+Given a decision procedure for some type, and a propositionally truncated value
+of that type, we can construct an element of the type.
+
+In the case of \(x\;\AgdaFunction{\(\in\)}\;\mathit{xs}\) we can construct a
+decision procedure for membership of a list, since we already have decidable
+equality on the elements of the list, proving our obligation.
 \section{Kuratowski Finiteness} \label{kuratowski} \todo{better intro}

### 2020-07-01 sections/finiteness-predicates.tex
finished applying edits

@@
 
-To actually use values of this type we have the following two eliminators:
-\begin{multicols}{2} \null \vfill
+To actually use values of this type we have the following eliminator:
 \begin{agdalisting} \label{elim-prop}
   \ExecuteMetaData[agda/HITs/PropositionalTruncation/Display.tex]{rec-prop-trunc}
-  \end{agdalisting} \vfill \null \columnbreak \null \vfill
-  \begin{agdalisting} \label{elim-prop-coh}
-    \ExecuteMetaData[agda/HITs/PropositionalTruncation/Display.tex]{rec-prop-trunc-set}
-  \end{agdalisting} \vfill \null
-\end{multicols} \noindent
-While these are not the only eliminators available for \proptrunc{A}, they are
-the only ones we will use in this paper.
-The first says we can eliminate into any proposition: interestingly, this allows
+\end{agdalisting}
+This says that we can eliminate into any proposition: interestingly, this allows
 us to define a monad instance for \AgdaDatatype{\(\lVert \_ \rVert\)}, meaning
-we can use things like do-notation.
-The second says we can eliminate into a set as long as the function we use
-doesn't care about which value it's given: formally, \(f\) in this example has
-to be ``coherently constant'' \cite{krausGeneralUniversalProperty2015}.
+we can use things like do-notation. \todo{Bit more on do-notation etc?}
 
@@
 derive from \(\AgdaDatatype{\ensuremath{\mathcal{B}}}\).
-Luckily, this is not the case: we will show here how to use the eliminators from
-propositional truncation (Equations \ref{elim-prop} and \ref{elim-prop-coh}) to
-use to of our already-proven derivations from \AgdaFunction{\ensuremath{\mathcal{B}}}
-on \AgdaFunction{\ensuremath{\mathcal{C}}}. \todo{rephrase?}
-
-The following two lemmas are proven in
+Luckily, this is not the case: we will show here how to derive decidable
+equality (Lemma~\ref{cardinal-finite-discrete}) and cardinality
+(Lemma~\ref{card-finite-cardinality}) out from under the
+truncation.\todo{Rephrase?}
+Those two lemmas are proven in
 \cite{yorgeyCombinatorialSpeciesLabelled2014} (Proposition 2.4.9 and 2.4.10,
@@
 Our contribution for this section is simply the formalisation.
-\begin{lemma}
-  Given a cardinally finite type, we can derive the type's cardinality, as well
-  as a propositionally truncated proof of equivalence with \AgdaDatatype{Fin}s of
-  the same cardinality.
-  \begin{agdalisting}
-    \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{cardinality-is-unique}
-  \end{agdalisting}
-\end{lemma}
-\begin{proof}
-  The high-level overview of our proof is as follows:
-  \begin{agdalisting}
-    \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{cardinality-is-unique-impl}
-  \end{agdalisting}
-  It is the composition of two operations: first, with
-  \AgdaFunction{\ensuremath{\lVert \text{map} \rVert}}, we change the truncated
-  proof of manifest bishop finiteness to a proof of equivalence with fin.
-
-  Then we use the eliminator from Equation~\ref{elim-prop-coh} with three
-  parameters.
-  The first simply proves that that the output is a set:
-  \begin{agdalisting}
-    \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{card-isSet}
-  \end{agdalisting}
-  The second is the function we apply to the truncated value:
-  \begin{agdalisting}
-    \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{alg}
-  \end{agdalisting}
-  And the third is a proof that that function is itself coherently constant:
-  \begin{agdalisting}
-    \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{const-alg}
-  \end{agdalisting}
-
-  The tricky part of the proof is \AgdaFunction{const-alg}: here we need to show
-  that \AgdaFunction{alg} returns the same value no matter its input.
-  That output is a pair, the first component of which is the cardinality, and
-  the second the truncated equivalence proof.
-  The truncated proofs in the output are trivially equal by the truncation, so
-  our obligation now has been reduced to:
-  \begin{equation}
-    \frac{(n :\;\AgdaDatatype{\ensuremath{\mathbb{N}}}) \; \; \; (p : \AgdaDatatype{Fin}\;n\;\AgdaFunction{\ensuremath{\simeq}}\;A) \; \; \;
-      (m :\;\AgdaDatatype{\ensuremath{\mathbb{N}}}) \; \; \; (q : \AgdaDatatype{Fin}\;m\;\AgdaFunction{\ensuremath{\simeq}}\;A)
-    }{
-      n\;\AgdaFunction{\ensuremath{\equiv}}\;m
-    }
-  \end{equation}
-  Given univalence we have \(\AgdaDatatype{Fin}\;n \;\AgdaFunction{\ensuremath{\equiv}}\; \AgdaDatatype{Fin}\;m\),
-  and the rest of our task is to prove:
-  \begin{equation}
-    \frac{\AgdaDatatype{Fin}\;n\;\AgdaDatatype{\ensuremath{\equiv}}\;\AgdaDatatype{Fin}\;m}{n \equiv m}
-  \end{equation}
-
-  This is a well-known chestnut in dependently-typed programming, and one that
-  has a surprisingly tricky and complex proof.
-  We do not include it here, since it has already been explored elsewhere, but
-  it is present in our formalisation.
-\end{proof}
-
-Another proof that carries over from manifest Bishop finiteness is decidable
-equality.
-This demonstrates an interesting aspect of propositional truncation: it can
-contain computational information.
+
+First we'll show that decidable equality carries over from manifest Bishop
+finiteness.
+Before we do, note that the fact that we can do this says something interesting
+about propositional truncation: it has computational, or algorithmic, content.
 That is in contrast to other ways to ``truncate'' types: \(\neg \neg P\), for
@@
 from \(\neg \neg P\). \todo{Is this true? It certainly seems true...}
-
-Here we'll briefly show how to extract ``computation'' from under a
-propositional truncation.
 \begin{lemma} \label{cardinal-finite-discrete}
   Any cardinal-finite set has decidable equality.
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{card-discrete}
+  \end{agdalisting}
 \end{lemma}
 \begin{proof}
-We can extract plenty of computation from under propositional truncation as long
-as we prove the thing we're extracting is a proposition (Equation~\ref{elim-prop}). 
+We already know that manifest Bishop finiteness implies decidable equality
+\todo{ref here}; to apply that proof to cardinal finiteness we'll use the
+eliminator in Equation~\ref{elim-prop}.
 Our task, in other words, is to prove the following:
@@
 \cite{hedbergCoherenceTheoremMartinLof1998}.
-\begin{theorem}[Hedberg's Theorem]
-  Any type with decidable equality is a set.
-\end{theorem} \noindent
+This tells us that any type with decidable equality is a set.
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Relation/Nullary/Discrete/Properties.tex]{discrete-isset}
+\end{agdalisting}
 And of course we know that \(A\) here has decidable equality: we were just given
 two proofs of that fact at the beginning of this proof!
+
+This suffices to prove that decidable equality is itself a proposition, and
+therefore that we can apply Equation~\ref{elim-prop} and the proof that bishop
+finiteness implies decidable equality to cardinal finiteness, proving our goal.
+\end{proof}
+
+The next thing we can derive from underneath the truncation in cardinal
+finiteness is a natural number representing the actual cardinality of the finite
+type.
+Of course \agdambb{N} isn't a proposition, so the eliminator in
+equation~\ref{elim-prop} won't work for us here.
+Instead we will use the following:
+\begin{agdalisting} \label{elim-prop-coh}
+  \ExecuteMetaData[agda/HITs/PropositionalTruncation/Display.tex]{rec-prop-trunc-set}
+\end{agdalisting}
+This says that we can eliminate into a set as long as the function we use
+doesn't care about which value it's given: formally, \(f\) in this example has
+to be ``coherently constant'' \cite{krausGeneralUniversalProperty2015}.
+
+With that, we can move on to the proof:
+\begin{lemma} \label{card-finite-cardinality}
+  Given a cardinally finite type, we can derive the type's cardinality, as well
+  as a propositionally truncated proof of equivalence with \AgdaDatatype{Fin}s of
+  the same cardinality.
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{cardinality-is-unique}
+  \end{agdalisting}
+\end{lemma}
+\begin{proof}
+  The high-level overview of our proof is as follows:
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{cardinality-is-unique-impl}
+  \end{agdalisting}
+  It is the composition of two operations: first, with
+  \AgdaFunction{\ensuremath{\lVert \text{map} \rVert}}, we change the truncated
+  proof of manifest bishop finiteness to a proof of equivalence with fin.
+
+  Then we use the eliminator from Equation~\ref{elim-prop-coh} with three
+  parameters.
+  The first simply proves that that the output is a set:
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{card-isSet}
+  \end{agdalisting}
+  The second is the function we apply to the truncated value:
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{alg}
+  \end{agdalisting}
+  And the third is a proof that that function is itself coherently constant:
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{const-alg}
+  \end{agdalisting}
+
+  The tricky part of the proof is \AgdaFunction{const-alg}: here we need to show
+  that \AgdaFunction{alg} returns the same value no matter its input.
+  That output is a pair, the first component of which is the cardinality, and
+  the second the truncated equivalence proof.
+  The truncated proofs in the output are trivially equal by the truncation, so
+  our obligation now has been reduced to:
+  \begin{equation}
+    \frac{(n :\;\AgdaDatatype{\ensuremath{\mathbb{N}}}) \; \; \; (p : \AgdaDatatype{Fin}\;n\;\AgdaFunction{\ensuremath{\simeq}}\;A) \; \; \;
+      (m :\;\AgdaDatatype{\ensuremath{\mathbb{N}}}) \; \; \; (q : \AgdaDatatype{Fin}\;m\;\AgdaFunction{\ensuremath{\simeq}}\;A)
+    }{
+      \agdaequiv{n}{m}
+    }
+  \end{equation}
+  Given univalence we have \(\AgdaDatatype{Fin}\;n \;\AgdaFunction{\ensuremath{\equiv}}\; \AgdaDatatype{Fin}\;m\),
+  and the rest of our task is to prove:
+  \begin{equation}
+    \frac{\agdaequiv{\AgdaDatatype{Fin}\;n}{\AgdaDatatype{Fin}\;m}}{\agdaequiv{n}{m}}
+  \end{equation}
+
+  This is a well-known puzzle in dependently-typed programming, and one that
+  has a surprisingly tricky and complex proof.
+  We do not include it here, since it has already been explored elsewhere, but
+  it is present in our formalisation.
+\end{proof}
+
+\subsection{Going from Cardinal Finiteness to Manifest Bishop Finiteness}
+We know of course that we can convert any proof of manifest Bishop finiteness to
+a proof of Cardinal finiteness: it's just the truncation function
+\AgdaInductiveConstructor{\(\lvert \_ \rvert\)}.
+It's the other direction which presents a difficulty:
+\begin{theorem} \label{cardinal-to-manifest-bishop}
+  Any cardinal finite type with a total order is Bishop finite. \todo{Figure out
+    how to put a type here}
+\end{theorem}
+\begin{proof}
+The proof for this particular theorem is quite involved in the formalisation, so
+we only give its sketch here \todo{ref to formalisation}.
+
+Our strategy will be to \emph{sort} the support list of the proof for Bishop
+finiteness, and then prove that the sorting function is coherently constant,
+thereby satisfying the eliminator in Equation~\ref{elim-prop-coh}.
+We need to show, in other words, that sorting two support lists from proofs of
+manifest Bishop finiteness on the same type with the same order always returns
+the same result.
+For simplicity's sake we will use insertion sort:
+
+\begin{minipage}{\linewidth}
+\begin{multicols}{2}
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Data/List/Sort.tex]{insert}
+  \end{agdalisting} \columnbreak \vfill \null \vfill
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Data/List/Sort.tex]{insertion-sort}
+  \end{agdalisting} \vfill \null
+\end{multicols}
+\end{minipage} \noindent
+And we prove that \AgdaFunction{sort} produces a list which is sorted, and a
+permutation of its input.
+\begin{multicols}{2}
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Data/List/Sort.tex]{sort-sorts}
+  \end{agdalisting} \columnbreak
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Data/List/Sort.tex]{sort-perm}
+  \end{agdalisting}
+\end{multicols} \vspace{-\baselineskip}
+We've introduced two new types here: \AgdaFunction{Sorted} is a predicate
+enforcing that the given list is sorted, and
+\AgdaFunction{\(\leftrightsquigarrow\)} is a permutation relation between two
+lists.
+We take the definition of permutations from
+\cite{danielssonBagEquivalenceProofRelevant2012}: two lists are permutations of
+each other if their membership proofs are all equivalent.
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Data/List/Relation/Binary/Permutation.tex]{perm-def}
+\end{agdalisting}
+This definition fits particularly well for two reasons: first, it is defined on
+containers generically, which fits well with our finiteness predicates.
+Secondly, it is extremely straightforward to show that the support lists of any
+two proofs of manifest Bishop finiteness must be permutations of each other:
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{perm-bish}
+\end{agdalisting}
+
+Almost all of the pieces are in place now: we know that the support lists of all
+proofs of \(\agdacal{B}\;A\) are permutations of each other, and we know that
+\AgdaFunction{sort} returns a sorted permutation of its input.
+The final piece of the puzzle is the following:
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Data/List/Sort.tex]{sorted-perm-eq}
+\end{agdalisting}
+If two sorted lists are both permutations of each other they must be equal.
+Connecting up all the pieces we get the following:
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Data/List/Sort.tex]{perm-invar}
+\end{agdalisting}
+Because we know that all support lists of \(\agdacal{B}\;A\) are permutations of
+each other this is enough to prove that \AgdaFunction{sort} is coherently
+constant, and therefore can eliminate from within a truncation.
+The second component of the output pair (the cover proof) follows quite
+naturally from the definition of permutations.
 \end{proof}
@@
 do \emph{not} have a function in the other direction:
-\begin{equation}
+\begin{equation} \label{c-to-b}
   \agdacal{C}\;A\rightarrow\agdacal{B}\;A
-\end{equation}
+\end{equation} \todo{Solve the mystery of the mathcal difference for C and B}
 So the question arises naturally: is there a cardinally finite type which is
@@
 It turns out the answer is no!
-We can prove the following statement:
-\begin{agdalisting}
+The proof of this fact is relatively short:
+\begin{agdalisting} \label{no-gap-card-bishop}
   \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{no-gap-card-bishop}
 \end{agdalisting}
-We use the fact that \agdabot\; is itself a proposition to eliminate from under
-the truncation.
-\begin{agdalisting}
-  \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{no-gap-card-bishop-proof}
-\end{agdalisting}
-
+We can apply the function of type \(\agdacal{B}\;A\rightarrow\agdabot\)
+(i.e. \(\AgdaFunction{\(\neg\)}\;\agdacal{B}\;A\)) to the value of type
+\(\AgdaDatatype{\(\lVert\)}\;\agdacal{B}\;A\;\AgdaDatatype{\(\rVert\)}\) (i.e.
+\(\agdacal{C}\;A\)) using Equation~\ref{elim-prop}, since \agdabot\;is itself a
+proposition.
 This tells us that manifest bishop finiteness, cardinal finiteness, and split
 enumerability all refer to the same class of types.
-\subsection{Going from Cardinal Finiteness to Manifest Bishop Finiteness}
-We know of course that we can convert any proof of manifest Bishop finiteness to
-a proof of Cardinal finiteness: it's just the truncation function
-\AgdaInductiveConstructor{\(\lvert \_ \rvert\)}.
-It's the other direction which presents a difficulty:
-\begin{theorem} \label{cardinal-to-manifest-bishop}
-  Any cardinal finite type with a total order is Bishop finite.
-\end{theorem}
-The proof for this particular theorem is quite involved in the formalisation, so
-we only give its sketch here.
-
-Our strategy will be to \emph{sort} the support list of the proof for Bishop
-finiteness, and then prove that the sorting function is coherently constant,
-thereby satisfying the eliminator in Equation~\ref{elim-prop-coh}.
-We need to show, in other words, that sorting two support lists from proofs of
-manifest Bishop finiteness on the same type with the same order always returns
-the same result.
-The sort we use is a standard insertion sort:
-\begin{multicols}{2}
-  \begin{agdalisting}
-    \ExecuteMetaData[agda/Data/List/Sort.tex]{insert}
-  \end{agdalisting} \columnbreak \null \vfill \null \vfill
-  \begin{agdalisting}
-    \ExecuteMetaData[agda/Data/List/Sort.tex]{insertion-sort}
-  \end{agdalisting} \vfill \null
-\end{multicols} \vspace{-\baselineskip} \noindent
-And we prove that \AgdaFunction{sort} produces a list which is sorted, and a
-permutation of its input.
-\begin{multicols}{2}
-  \begin{agdalisting}
-    \ExecuteMetaData[agda/Data/List/Sort.tex]{sort-sorts}
-  \end{agdalisting} \columnbreak
-  \begin{agdalisting}
-    \ExecuteMetaData[agda/Data/List/Sort.tex]{sort-perm}
-  \end{agdalisting}
-\end{multicols} \vspace{-\baselineskip}
-We've introduced permutations here: we take their definition from
-\cite{danielssonBagEquivalenceProofRelevant2012}.
-Two lists are permutations of each other if their membership proofs are all
-equivalent.
-\begin{agdalisting}
-  \ExecuteMetaData[agda/Data/List/Relation/Binary/Permutation.tex]{perm-def}
-\end{agdalisting}
-This definition fits particularly well for two reasons: first, it is defined on
-containers generically, which fits well with our finiteness predicates.
-Secondly, it is extremely straightforward to show that the support lists of any
-two proofs of manifest Bishop finiteness must be permutations of each other:
-\begin{agdalisting}
-  \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{perm-bish}
-\end{agdalisting}
-
-Almost all of the pieces are in place now: we know that the support lists of all
-proofs of \(\agdacal{B}\;A\) are permutations of each other, and we know that
-\AgdaFunction{sort} returns a sorted permutation of its input.
-The final piece of the puzzle is the following:
-\begin{agdalisting}
-  \ExecuteMetaData[agda/Data/List/Sort.tex]{sorted-perm-eq}
-\end{agdalisting}
-If two sorted lists are both permutations of each other they must be equal.
-Connecting up all the pieces we get the following:
-\begin{agdalisting}
-  \ExecuteMetaData[agda/Data/List/Sort.tex]{perm-invar}
-\end{agdalisting}
-Because we know that all support lists of \(\agdacal{B}\;A\) are permutations of
-each other this is enough to prove that \AgdaFunction{sort} is coherently
-constant, and therefore can eliminate from within a truncation.
-The second component of the output pair (the cover proof) follows quite
-naturally from the definition of permutations.
+
+Interestingly, while we cannot construct a function with the type in
+Equation~\ref{c-to-b}, it does exist \emph{classically}.
+In fact we can derive it from Equation~\ref{no-gap-card-bishop} using
+straightforward applications of De Morgan's laws:
+\begin{align*}
+   &\; \neg (\agdacal{C}\;A \times \neg\; \agdacal{B}\;A) \\
+  =&\; \neg\; \agdacal{C}\;A + \neg \neg\; \agdacal{B}\;A \\
+  =&\; \neg\; \agdacal{C}\;A + \agdacal{B}\;A \\
+  =&\; \agdacal{C}\;A \rightarrow \agdacal{B}\;A
+\end{align*}
 \section{Manifest Enumerability} \label{manifest-enumerability}
-We have now explored quite far in the direction of restricting information:
-indeed, cardinal finiteness, a mere proposition, has the minimal amount of
-information a predicate can have.
-What we haven't seen, however, is a predicate with a different domain to the
-others.
-Since manifest Bishop finiteness and split enumerability have equal strength
-\todo{ref here}, and since we've shown that there is no type which is cardinal
-finite but not manifest Bishop finite, we haven't explored any of the
-``restrictiveness'' axis from Figure~\ref{finite-classification}.
-
-What we're looking for is a type which is in some sense finite, but doesn't
+Given that we have just proven that all of our finiteness predicates apply to
+the same types, the natural next step is to try find a predicate which applies
+to a different class of types.
+Let's first talk about what this new class of types might look like:
+what we're looking for is a type which is in some sense finite, but doesn't
 conform to any of the predicates we've seen so far.
@@
 But the type certainly \emph{seems} finite!
+It has finitely many points, for instance.
 In order to explore the ``restrictiveness'' axis in
@@
 
-And indeed this means that the circle is manifestly enumerable. \todo{This uses
-  a bunch of cubical primitives!}
+And indeed this means that the circle is manifestly enumerable.
 \begin{agdalisting}
   \ExecuteMetaData[agda/Cardinality/Finite/ManifestEnumerable.tex]{circle-is-manifest-enum}
+\end{agdalisting}
+We use a lemma here, proven in the Cubical Agda library, that
+\AgdaDatatype{S\(^1\)} is \emph{connected}:
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Cardinality/Finite/ManifestEnumerable.tex]{s1-connected}
 \end{agdalisting}

### 2020-07-01 sections/finiteness-predicates.tex
finished kuratowski finiteness section

@@
 equality on the elements of the list, proving our obligation.
-\section{Kuratowski Finiteness} \label{kuratowski} \todo{better intro}
-The one big missing definition of finiteness to cover is \emph{Kuratowski}
+\section{Kuratowski Finiteness} \label{kuratowski}
+We now finally arrive at the most important definition of finiteness: Kuratowski
 finiteness.
-While it's quite important, it's also quite different from the definitions we've
-seen so far.
-It starts with an encoding of the free join semilattice.
-\begin{definition}[Free Join Semilattice]
-  \(\mathcal{K}(A)\) is the free join semilattice, or, alternatively, the type
-  of Kuratowski-finite subsets of \(A\).
+As a definition, it is quite different from the predicates we've seen (it
+doesn't involve lists, for instance), but it plays a much larger role in the
+literature on finiteness predicates than, say, manifest enumerability.
+
+We start with the definition of Kuratowski-finite subsets.
 \begin{agdalisting}
@@
 \end{agdalisting}
-
-  We define it as a HIT.
 The first two constructors are point constructors, giving ways to create
-  values of type \(\mathcal{K}(A)\).
+values of type \(\agdacal{K}\;A\).
 They are also recognisable as the two constructors for finite lists, a type
 which represents the free monoid.
-
 The next two constructors add extra paths to the type: equations that usage of
 the type must obey.
-  These extra paths turn the free monoid into the free \emph{commutative} (com)
-  \emph{idempotent} (dup) monoid.
-
-  The final constructor enforces that the type \(\mathcal{K}(A)\) must be a set.
-\end{definition}
+These extra paths turn the free monoid into the free \emph{commutative}
+(\AgdaInductiveConstructor{com}) \emph{idempotent}
+(\AgdaInductiveConstructor{dup}) monoid.
+The final constructor truncates the type \(\agdacal{K}\;A\) to a set.
+
 The Kuratowski finite subset is a free join semilattice (or, equivalently, a
 free commutative idempotent monoid).
-More prosaically, \(\mathcal{K}\) is the abstract data type for finite sets, as
+More prosaically, \(\agdacal{K}\) is the abstract data type for finite sets, as
 defined in the Boom hierarchy \cite{boomFurtherThoughtsAbstracto1981,
   bunkenburgBoomHierarchy1994}.
-However, rather than just being a specification, \(\mathcal{K}\) is fully usable
+However, rather than just being a specification, \(\agdacal{K}\) is fully usable
 as a data type in its own right, thanks to HITs.
 
-Other definitions of \(\mathcal{K}\) exist (such as the one in
-\cite{fruminFiniteSetsHomotopy2018}) which make the fact that \(\mathcal{K}\) is
+Other definitions of \(\agdacal{K}\) exist (such as the one in
+\cite{fruminFiniteSetsHomotopy2018}) which make the fact that \(\agdacal{K}\) is
 the free join semilattice more obvious.
@@
 Next, we need a way to say that an entire type is Kuratowski finite.
-For that, we will need to define membership of \(\mathcal{K}\).
-\begin{definition}[Membership of \(\mathcal{K}\)]
-  Membership is defined by pattern-matching on \(\mathcal{K}\).
-  The two point constructors are handled like so: \todo{to agda}
+For that, we will need to define membership of \(\agdacal{K}\).
 \begin{equation}
   \begin{alignedat}{2}
-      x \in&& \; []                      &\coloneqq \bot ; \\
-      x \in&& \; y \dblcolon \mathit{ys} &\coloneqq \lVert x \equiv y \uplus x \in \mathit{ys} \rVert ;
+    x\;\AgdaFunction{\ensuremath{\in}}& \; \AgdaInductiveConstructor{[]}                      &&= \agdabot \\
+    x\;\AgdaFunction{\ensuremath{\in}}& \; y \;\AgdaInductiveConstructor{\ensuremath{\dblcolon}}\; \mathit{ys} &&= \AgdaDatatype{\ensuremath{\lVert}}\;\agdaequiv{x}{y}\;\AgdaDatatype{\ensuremath{\uplus}}\;x\;\AgdaFunction{\ensuremath{\in}}\;\mathit{ys}\;\AgdaDatatype{\ensuremath{\rVert}}
   \end{alignedat}
 \end{equation}
-  The \(\text{com}\) and \(\text{dup}\) constructors are handled by proving that
-  the truncated form of \(\uplus\) is itself commutative and idempotent.
-  The type of propositions is itself a set, satisfying the \(\text{trunc}\)
+The \AgdaInductiveConstructor{com} and \AgdaInductiveConstructor{dup}
+constructors are handled by proving that the truncated form of \AgdaDatatype{\ensuremath{\uplus}}
+itself commutative and idempotent.
+The type of propositions is itself a set, satisfying the \AgdaInductiveConstructor{trunc}
 constructor.
+This gives us enough to define Kuratowski finiteness.
+\begin{definition}[Kuratowski Finiteness]
+  A type is Kuratowski finite if there exists a Kuratowski-finite subset of that
+  type which contains every element of the type.
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Cardinality/Finite/Kuratowski.tex]{kuratowski-finite-def}
+  \end{agdalisting}
 \end{definition}
-Finally, we have enough background to define Kuratowski finiteness.
-\begin{definition}[Kuratowski Finiteness] \phantom{break}
-  \begin{agdalisting}
-    \ExecuteMetaData[agda/Cardinality/Finite/Kuratowski.tex]{kuratowski-finite-def}
-  \end{agdalisting}
-\end{definition}
-
-We also have the following two lemmas, proven in both
-\cite{fruminFiniteSetsHomotopy2018} and our formalisation.
-\begin{lemma}
-  \(\mathcal{K}^f\) is a mere proposition.
-\end{lemma}
-\begin{lemma}
-  This circle \(S^1\) is Kuratowski finite.
-\end{lemma}
-\subsection{Relation to Cardinal Finiteness}
+
+In our formalisation, we have proven a number of other lemmas regarding
+Kuratowski finiteness, including the fact that
+\AgdaFunction{\ensuremath{\mathcal{K}^f}} is a mere proposition, and that the
+circle is Kuratowski finite.
+These are also provided in \citet{fruminFiniteSetsHomotopy2018}, however, and
+aren't directly relevant to what we're doing here so we won't provide the
+details of the proof.
+The key proof we will rely on for the rest of the paper is the following:
 \begin{lemma} \label{cardinal-kuratowski}
@@
   set.
-  \begin{equation}
-    \mathcal{C}(A) \simeq \mathcal{K}^f(A) \times \text{Discrete}(A)
-  \end{equation}
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Cardinality/Finite/Kuratowski.tex]{card-iso-kuratowski}
+  \end{agdalisting}
 \end{lemma}
 This proof is constructed by providing a pair of functions: one from
-\(\mathcal{C}(A)\) to \(\mathcal{K}^f(A) \times \text{Discrete}(A)\), and one the
+\(\mathcal{C}(A)\) to \(\agdacal{K}^f(A) \times \text{Discrete}(A)\), and one the
 other way.
 This pair implies an equivalence, because both source and target are
-propositions.
-The actual functions themselves are proven in our formalisation, as well as in
-\cite{fruminFiniteSetsHomotopy2018}.
+propositions. \todo{is it clear that this lemma is also in frumin et al?}
 

### 2020-07-01 sections/finiteness-predicates.tex
bit more on kuratowski etc

@@
 equivalent to the one above.
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Algebra/Construct/Free/Semilattice/Direct.tex]{direct-def}
+\end{agdalisting}
 
@@
 
-In our formalisation, we have proven a number of other lemmas regarding
-Kuratowski finiteness, including the fact that
-\AgdaFunction{\ensuremath{\mathcal{K}^f}} is a mere proposition, and that the
-circle is Kuratowski finite.
-These are also provided in \citet{fruminFiniteSetsHomotopy2018}, however, and
-aren't directly relevant to what we're doing here so we won't provide the
-details of the proof.
-The key proof we will rely on for the rest of the paper is the following:
-\begin{lemma} \label{cardinal-kuratowski}
-  Cardinal finiteness is equivalent to Kuratowski finiteness over a discrete
-  set.
-  \begin{agdalisting}
+While Kuratowski finiteness is something of the standard formal definition of
+finiteness, it is quite separated from the enumeration-based definitions we have
+presented so far.
+It's difficult to relate to surjections and equivalences, and requires a
+different style of proof to reason about.
+As such, we want to get \emph{away} from Kuratowski finiteness as quickly as
+possible.
+To do so we use the following lemma:
+\begin{lemma}  \label{manifest-enum-kuratowski}
+  Kuratowski finiteness is equivalent to truncated manifest enumerability.
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Cardinality/Finite/Kuratowski.tex]{manifest-enum-kuratowski}
+  \end{agdalisting}
+\end{lemma}
+\begin{proof} \todo{Does this need to be fleshed out?}
+  This proof is constructed by providing a pair of functions, to and from each
+  side of the equivalence.
+  This pair implies an equivalence, because both source and target are
+  propositions.
+  This proof, as well as its auxiliary lemmas, are also provided in
+  \citet{fruminFiniteSetsHomotopy2018}, although there the setting is HoTT
+  rather than CuTT. \todo{Make the relationship here more clear}
+\end{proof}
+By relating Kuratowski finiteness---with a full equivalence, no less---to an
+enumerated predicate, we have made it possible to talk about Kuratowski
+finiteness without interacting with the type at all.
+
+In the next section, we will explore the category of discrete Kuratowski finite
+sets.
+Under the hood, however, we will really be working with cardinal finite sets.
+We can do this in a fully rigorous way because
+Lemma~\ref{manifest-enum-kuratowski} allows us to prove the following:
+\begin{agdalisting} \label{cardinal-kuratowski}
   \ExecuteMetaData[agda/Cardinality/Finite/Kuratowski.tex]{card-iso-kuratowski}
 \end{agdalisting}
-\end{lemma}
-This proof is constructed by providing a pair of functions: one from
-\(\mathcal{C}(A)\) to \(\agdacal{K}^f(A) \times \text{Discrete}(A)\), and one the
-other way.
-This pair implies an equivalence, because both source and target are
-propositions. \todo{is it clear that this lemma is also in frumin et al?}
 

### 2020-07-01 sections/search.tex
fixed todonotes
down to surjections in split enumerability
bit more on kuratowski etc

@@
 A common theme in dependently-typed programming is that proofs of interesting
-theoretical things may actually correspond to useful algorithms in some way
+theoretical things often correspond to useful algorithms in some way
 related to that thing.
@@
 library to do just that.
-\todo[inline]{These examples so far are pretty focused on the bool associativity
+\todo{These examples so far are pretty focused on the bool associativity
   example.
@@
 The type checker does not require us to manually explain each step of evaluation
-of:
+of
 \(\AgdaInductiveConstructor{true}\;\AgdaFunction{∧}\;\AgdaInductiveConstructor{false}\).
@@
 form if they're needed to type check.
+The type checker will perform \(\beta\)-reduction as much as it can.
 
@@
 ``Exhaustibility''.
-\begin{definition}[Exhaustibility]
-  We say a type \(A\) is exhaustible if, for any decidable predicate \(P\) on
-  \(A\), the universal quantification of the predicate is decidable.
 \begin{agdalisting}
@@
 \end{agdalisting}
-\end{definition}
+We say a type \(A\) is exhaustible if, for any decidable predicate \(P\) on
+\(A\), the universal quantification of the predicate is decidable.
+
 This property of \AgdaDatatype{Bool} would allow us to automate the proof of
@@
 representatives of a type which support some property.
-For that we need the more well-known, but related, property of
+For that we need the more well-known related property of
 \emph{omniscience}.
-\begin{definition}[Limited Principle of Omniscience]
-  For any type \(A\) and predicate \(P\) on \(A\), the limited principle of
-  omniscience \cite{myhillErrettBishopFoundations1972} is as follows:
 \begin{agdalisting}
@@
 \end{agdalisting}
-  In other words, for any decidable predicate the existential quantification of
-  that predicate is also decidable.
-\end{definition}
+The ``limited principle of omniscience''
+\cite{bishopFoundationsConstructiveAnalysis1967} is a classical principle which
+says that omniscience holds for all sets.
+It doesn't hold constructively, of course: it lies a little bit below LEM in
+terms of its non-constructiveness, given that it can be derived from LEM but LEM
+cannot be derived from it.
 
-Because we're constructive, only a select few types are omniscient: finite
-types, for instance (the law of the excluded middle implies that all types are
-omniscient, meaning that all types are omniscient classically).
+Omniscience implies exhaustibility: we can use the usual rule of \todo{what is
+  this called/from again?}
+\begin{equation}
+  \neg \exists x. P(x) \iff \forall x. \neg P(x)
+\end{equation}
+to turn omniscience for some predicate \(P\) into exhaustibility for some
+predicate \(\neg \neg P\).
+Usually we don't have double negation elimination constructively, but since
+\(P\) is decidable it's actually present in this case:
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Relation/Nullary/Decidable/Properties.tex]{dec-double-neg-elim}
+\end{agdalisting}
+All together, this gives us the following proof:
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Relation/Nullary/Omniscience.tex]{omniscient-to-exhaustible}
+\end{agdalisting}
+
+Our focus here is on those types for which omniscience \emph{does} hold,
+which includes the (ordered) finite types.
 Perhaps surprisingly, it is not \emph{only} finite types which are exhaustible.
@@
 
-Omniscience and exhaustibility are not interchangeable: every omniscient type is
-exhaustible, but the converse is not true.
-Conceptually, omniscience needs some kind of ordering on the underlying type.
-This is because omniscience returns a candidate satisfying the given predicate:
-there is no requirement, though, that only one element in the underlying type
-satisfies the decidable predicate.
-As a result, omniscience needs some way to choose among all possible candidates:
-this is the ``order'' we are referring to.
-This is also the same ``order'' that we referred to when talking about the
-finiteness predicates: all ordered predicates (in
-Figure~\ref{finite-classification}) imply omniscience, whereas the unordered
-predicates only imply exhaustibility.
+All of the finiteness predicates imply exhaustibility.
+To prove that fact we'll just show that the Kuratowski finite types are
+exhaustible: since it's the weakest predicate, and can be derived from all the
+others.
 \begin{lemma}
-  Omniscience implies exhaustibility
-\end{lemma}
-\begin{proof}
-  \todo{proof here}
-\end{proof}
+  Kuratowski finiteness implies exhaustibility.
+\end{lemma} \todo{Proof?}
+Manifest enumerability is similarly the weakest of the ordered predicates, so we
+will prove here that it implies omniscience.
+\begin{lemma}
+  Manifest enumerability implies omniscience.
+\end{lemma} \todo{proof?}
 
-\begin{lemma}
-  If exhaustibility implies omniscience, then the axiom of choice holds
-\end{lemma}
-\begin{proof}
-  \todo{proof here}
-\end{proof}
+Finally, there is a form of omniscience which works with Kuratowski finiteness:
+propositional omniscience.
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Relation/Nullary/Omniscience.tex]{prop-omniscient}
+\end{agdalisting}
+By truncating the returned \AgdaDatatype{\ensuremath{\Sigma}} we obviate the
+need for 
+\section{Countdown}
+The Countdown problem~\cite{huttonCountdownProblem2002} is a well-known puzzle
+in functional programming (which was apparently turned into a TV show).
+As a running example in this paper, we will produce a verified program which
+lists all solutions to a given countdown puzzle: here we will briefly explain
+the game and our strategy for solving it.
 
-And the relation to the finiteness predicates is straightforward: all of the
-finiteness predicates we have seen imply exhaustibility, and all of the ordered
-finiteness predicates imply omniscience.
-We can prove this by showing exhaustibility and omniscience for the weakest
-candidate of finiteness predicates.
-\begin{lemma}
-  Kuratowski finiteness implies exhaustibility
-\end{lemma}
-\begin{lemma}
-  Manifest enumerability implies omniscience
-\end{lemma}
+The idea behind countdown is simple: given a list of numbers, contestants must
+construct an arithmetic expression (using a small set of functions) using some
+or all of the numbers, to reach some target.
+Here's an example puzzle:
+\begin{gather*}
+  \boxed{1} \boxed{3} \boxed{7} \boxed{10} \boxed{25} \boxed{50} \\
+  \boxed{765} \tag{Target}
+\end{gather*}
+We'll allow the use of \(+\), \(-\), \(\times\), and \(\div\).
+The answer is at the bottom of this page\footnotemark.
 
-Finally, we can get around the order requirement for prop-valued predicates for
-omniscience.
-\begin{lemma}
-  Omniscience and exhaustibility coincide for prop-valued predicates.
-\end{lemma}
-\section{Countdown}
+\footnotetext{\rotatebox[origin=c]{180}{Answer: \(3 \times (7 \times (50 - 10) - 25)\)}}
+
+Our strategy for finding solutions to a given puzzle is to describe precisely
+the type of solutions to a puzzle, and then show that that type is finite.
+So what is a ``solution'' to a countdown puzzle?
+Broadly, it has two parts:
+\begin{description}
+  \item[A Transformation] from a list of numbers to an expression.
+  \item[A Predicate] showing that the expression is valid and evaluates to the
+    target.
+\end{description}
+The first part is described in Figure~\ref{countdown-transform}.
+
+\input{figures/countdown-transformation}
+
+This transformation has four steps.
+First (Fig.~\ref{countdown-selection}) we have to pick which numbers we include
+in our solution.
+We will need to show there are finitely many ways to filter \(n\) numbers.
+
+Secondly (Fig.~\ref{countdown-permutation}) we have to permute the chosen
+numbers.
+The representation for a permutation is a little trickier to envision: proving
+that it's finite is trickier still.
+We will need to rely on some of the more involved lemmas later on for this
+problem.
+
+The third step (Fig.~\ref{countdown-operators}) is a vector of length \(n\) of finite objects (in this case operators
+chosen from \(+\), \(\times\), \(-\), and \(\div\)).
+Although it is complicated slightly by the fact that the \(n\) in this
+\(n\)-tuple is dependent on the amount of numbers we let through in the filter
+in step one.
+(in terms of types, that means we'll need a \(\Sigma\) rather than a
+\(\times\), explanations of which are forthcoming).
+
+Finally (Fig.~\ref{countdown-parens}), we have to parenthesise the expression in
+a certain way.
+This can be encapsulated by a binary tree with a certain number of leaves:
+proving that that is finite is tricky again.
+
+Once we have proven that there are finitely many transformations for a list of
+numbers, we will then have to filter them down to those transformations which
+are valid, and evaluate to the target.
+This amounts to proving that the decidable subset of a finite set is also
+finite.
+
+Finally, we will also want to optimise our solutions and solver: for this we
+will remove equivalent expressions, which can be accomplished with quotients.
 We have already introduced and described countdown: in this section, we will

### 2020-07-02 sections/search.tex
kuratowski omniscience proof

@@
 \end{agdalisting}
-By truncating the returned \AgdaDatatype{\ensuremath{\Sigma}} we obviate the
-need for 
+By truncating the returned \AgdaDatatype{\ensuremath{\Sigma}} we don't reveal
+which \(A\) we've chosen which satisfies the predicate: this means that it can
+be pulled out of the Kuratowski finite subset without issue.
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Cardinality/Finite/Kuratowski.tex]{kuratowski-prop-omniscient}
+\end{agdalisting}
 \section{Countdown}

### 2020-07-02 sections/search.tex
bit more on search

@@
 
+\input{figures/countdown-transformation}
+
 The idea behind countdown is simple: given a list of numbers, contestants must
@@
 Here's an example puzzle:
-\begin{gather*}
-  \boxed{1} \boxed{3} \boxed{7} \boxed{10} \boxed{25} \boxed{50} \\
-  \boxed{765} \tag{Target}
-\end{gather*}
+\begin{displayquote}
+  Using some or all of the numbers 1, 3, 7, 10, 25, and 50 (using each at most
+  once), construct an expression which equals 765.
+\end{displayquote}
 We'll allow the use of \(+\), \(-\), \(\times\), and \(\div\).
@@
 The first part is described in Figure~\ref{countdown-transform}.
-
-\input{figures/countdown-transformation}
 
@@
 how the finiteness proofs can assist us to write the solver.
+\subsection{Finite Vectors}
+We'll start with a simple example: for both the selection
+(Fig.~\ref{countdown-selection}) and operators (Fig.~\ref{countdown-operators})
+section, all we need to show is that a vector of some finite type is itself
+finite.
+To describe which elements to keep from an \(n\)-element list, so instance, we
+only need a vector of Booleans of length \(n\).
+Similarly, to pick \(n\) operators requires us only to provide a vector of \(n\)
+operators.
+And we can prove in a straightforward way that a vector of finite things is
+itself finite.
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Countdown.tex]{vec-fin}
+\end{agdalisting}
 \subsection{Finite Permutations}
-The first step of the transformation we will represent as a finite object is a
-\emph{permutation}: once we choose the numbers we're going to use for the
-candidate for the solution, we have to order them in some way.
+A more complex, and interesting, step of the transformation is the first step
+(Fig.~\ref{countdown-permutation}), where we need to specify the permutation to
+apply to the chosen numbers.
 
@@
 \end{agdalisting}
+\subsection{Putting It All Together}
+At this point we have each of the four components of the transformation defined.
 \subsection{Filtering to Correct Expressions}
-\subsection{Putting It All Together}
 \section{Automating Proofs}
-One use for above constructions is the automation of certain proofs.
-In \cite{firsovDependentlyTypedProgramming2015}, which uses a similar approach
-to ours, the \(\AgdaDatatype{Pauli}\) group is used as an example.
-\ExecuteMetaData[agda/Data/Pauli.tex]{def}
-As \(\AgdaDatatype{Pauli}\) has 4 constructors, \(n\)-ary functions on
-\(\AgdaDatatype{Pauli}\) may require up to \(4^n\) cases, making even simple
-proofs prohibitively verbose.
-
-The alternative is to derive the things we need from omniscience, itself derived
-from a finiteness predicate.
-For proof search, the procedure is a well-known one in Agda
-\cite{devrieseBrightSideType2011}: we ask for the result of a decision procedure
-as an \emph{instance argument}, which will demand computation during
-typechecking.
-Our addition to this technique is a way to handle multiple arguments based on
-fully level-polymorphic dependent currying and uncurrying, building on 
-\cite{allaisGenericLevelPolymorphic2019}.
-\ExecuteMetaData[agda/Data/Pauli.tex]{assoc-prf}
-
-Finally, we can derive decidable equality on functions over finite types.
-We can also use functions in our proof search.
-Here, for instance, is an automated procedure which finds the
-\(\AgdaFunction{not}\) function on \(\mathbf{Bool}\), given a specification.
-\ExecuteMetaData[agda/Data/Pauli.tex]{not-spec}
 

### 2020-07-02 sections/search.tex
expression finite proof

@@
 \end{agdalisting}
+We've already shown that there are finitely many booleans, the fact that there
+are finitely many operators is similarly simple to prove:
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Countdown.tex]{op-fin}
+\end{agdalisting}
 \subsection{Finite Permutations}
@@
 At this point we have each of the four components of the transformation defined.
+From this we can define what an expression is:
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Countdown.tex]{expr-def}
+\end{agdalisting}
+Notice that we don't allow expressions with no numbers.
+
+The proof that this type is finite mirrors its definition closely:
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Countdown.tex]{expr-finite}
+\end{agdalisting}
 \subsection{Filtering to Correct Expressions}

### 2020-07-02 sections/search.tex
changed agda code a bit

@@
 \subsection{Filtering to Correct Expressions}
+We now have a way to construct, formally, every expression we can generate from
+a given list of numbers.
+This is incomplete in two ways, however.
+Firstly, some expressions are invalid: we should not, for instance, be able to
+divide two numbers which do not divide evenly.
+Secondly, we are only interested in those expressions which actually represent
+solutions: those which evaluate to the target, in other words.
+We can write a function which tells us if both of these things hold for a given
+expression like so:
+\begin{multicols}{2}
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Countdown.tex]{eval}
+  \end{agdalisting} \columnbreak
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Countdown.tex]{app-op}
+  \end{agdalisting}
+\end{multicols}
 \section{Automating Proofs}

### 2020-07-03 sections/search.tex
finished countdown section

@@
 \end{multicols}
+
+With this all together, we can finally write down the type of all solutions to a
+given countdown problem.
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Countdown.tex]{solution-type}
+\end{agdalisting}
+And, because the predicate here is decidable and a mere proposition, we can
+prove that there are finitely many solutions:
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Countdown.tex]{finite-solution}
+\end{agdalisting}
+And we can apply this to a particular problem like so:
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Countdown.tex]{example-solution}
+\end{agdalisting}
+
 \section{Automating Proofs}

### 2020-07-03 sections/search.tex
put back in old proof automation section

@@
 application soon.
-\section{Proof Automation And Search Techniques}
-For this prover we will not resort to reflection or similar techniques: instead,
-we will trick the type checker to do our automation for us.
-This is a relatively common technique, although not so much outside of Agda, so
-we will briefly explain it here.
-
-To understand the technique we should first notice that some proof automation
-\emph{already} happens in Agda, like the following:
-\begin{agdalisting}
-  \ExecuteMetaData[agda/Snippets/Bool.tex]{obvious}
-\end{agdalisting}
-The type checker does not require us to manually explain each step of evaluation
-of
-\(\AgdaInductiveConstructor{true}\;\AgdaFunction{∧}\;\AgdaInductiveConstructor{false}\).
-While it's not a particularly impressive example of automation, it does nonetheless
-demonstrate a principle we will exploit: closed terms will compute to a normal
-form if they're needed to type check.
-The type checker will perform \(\beta\)-reduction as much as it can.
-
-So our task is to rewrite proof obligations like the one in
-Equation~\ref{bool-assoc-auto-proof} into ones which can reduce completely. 
-As it turns out, we have already described the type of proofs which can ``reduce
-completely'': \emph{decidable} proofs.
-If we have a decision procedure over some proposition \(P\) we can run that
-decision during type checking, because the decision procedure itself is a proof
-that the decision will terminate.
-In code, we capture this idea with the following pair of functions:
-\begin{multicols}{2}
-  \begin{agdalisting}
-    \ExecuteMetaData[agda/Snippets/Bool.tex]{is-true}
-  \end{agdalisting} \columnbreak
-  \begin{agdalisting}
-    \ExecuteMetaData[agda/Snippets/Bool.tex]{from-true}
-  \end{agdalisting}
-\end{multicols}
-The first is a function which derives a type from whether a decision is
-successful or not.
-This function is important because if we use the output of this type at any
-point we will effectively force the unifier to run the decision computation.
-The second takes---as an implicit argument---an inhabitant of the type generated
-from the first, and uses it to prove that the decision can only be true, and the
-extracts the resulting proof from that decision.
-All in all, we can use it like this:
-\begin{agdalisting}
-  \ExecuteMetaData[agda/Snippets/Bool.tex]{extremely-obvious}
-\end{agdalisting}
-This technique will allow us to automatically compute any decidable predicate.
 \section{Omniscience}
@@
 expression like so:
+
+\begin{minipage}{\linewidth}
   \begin{multicols}{2}
@@
   \end{multicols}
+\end{minipage}
 
@@
 \end{agdalisting}
+\section{Proof Automation}
+So we have shown now how powerful the library is for proof search: it can
+search for functions, subsets of types, products, and so on, completely
+verifiable.
+In this section we will present the more user-friendly interface to the library,
+designed to be used to automate away tedious proofs in an easy way.
+\subsection{How to make the Typechecker do Automation}
+For this prover we will not resort to reflection or similar techniques: instead,
+we will trick the type checker to do our automation for us.
+This is a relatively common technique, although not so much outside of Agda, so
+we will briefly explain it here.
 
-\section{Automating Proofs}
+To understand the technique we should first notice that some proof automation
+\emph{already} happens in Agda, like the following:
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Snippets/Bool.tex]{obvious}
+\end{agdalisting}
+The type checker does not require us to manually explain each step of evaluation
+of
+\(\AgdaInductiveConstructor{true}\;\AgdaFunction{∧}\;\AgdaInductiveConstructor{false}\).
+While it's not a particularly impressive example of automation, it does nonetheless
+demonstrate a principle we will exploit: closed terms will compute to a normal
+form if they're needed to type check.
+The type checker will perform \(\beta\)-reduction as much as it can.
+
+So our task is to rewrite proof obligations like the one in
+Equation~\ref{bool-assoc-auto-proof} into ones which can reduce completely. 
+As it turns out, we have already described the type of proofs which can ``reduce
+completely'': \emph{decidable} proofs.
+If we have a decision procedure over some proposition \(P\) we can run that
+decision during type checking, because the decision procedure itself is a proof
+that the decision will terminate.
+In code, we capture this idea with the following pair of functions:
+\begin{multicols}{2}
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Snippets/Bool.tex]{is-true}
+  \end{agdalisting} \columnbreak
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Snippets/Bool.tex]{from-true}
+  \end{agdalisting}
+\end{multicols}
+The first is a function which derives a type from whether a decision is
+successful or not.
+This function is important because if we use the output of this type at any
+point we will effectively force the unifier to run the decision computation.
+The second takes---as an implicit argument---an inhabitant of the type generated
+from the first, and uses it to prove that the decision can only be true, and the
+extracts the resulting proof from that decision.
+All in all, we can use it like this:
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Snippets/Bool.tex]{extremely-obvious}
+\end{agdalisting}
+This technique will allow us to automatically compute any decidable predicate.
+\subsection{Building an Interface}
+We will next look at a syntax and interface to this proof technique.
+It starts with the following two functions:
+
+\begin{minipage}{\linewidth}
+  \begin{multicols}{2}
+    \begin{agdalisting}
+      \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable/Search.tex]{forall-ask}
+    \end{agdalisting} \columnbreak
+    \begin{agdalisting}
+      \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable/Search.tex]{exists-ask}
+    \end{agdalisting}
+  \end{multicols}
+\end{minipage}
+Clearly they're just restatements of exhaustibility and omniscience.
+However, we can combine these functions with what we've constructed above to
+create an automation procedure for each:
+
+\begin{minipage}{\linewidth}
+  \begin{multicols}{2}
+    \begin{agdalisting}
+      \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable/Search.tex]{forall-zap}
+    \end{agdalisting} \columnbreak
+    \begin{agdalisting}
+      \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable/Search.tex]{exists-zap}
+    \end{agdalisting}
+  \end{multicols}
+\end{minipage}
+
+This automation procedure allows us to state the property succinctly, and have
+the type checker go and run the decision procedure to solve it for us.
+Here's an example of its use:
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Snippets/Bool.tex]{pre-inst-proof}
+\end{agdalisting} \todo{Include counterexample stuff?}
+
+% As a quick aside, when the property is \emph{not} true, for instance:
+% \ExecuteMetaData[agda/Data/Pauli.tex]{contra}
+% Agda will fail by not finding an instance of \(\bot\).
+% The error message specifically is:
+% \begin{displayquote}
+%   \verb+No instance of type ⊥ was found in scope.+
+% \end{displayquote}
+% We can actually display a counterexample, by defining a custom empty type
+% parameterised by the counterexample itself.
+% % \ExecuteMetaData[agda/Cardinality/Finite/Search.tex]{counterexample-def}
+% Combined with some other changes to the combinators, this will give the
+% following more helpful error message:
+% \begin{displayquote}
+%   \verb+No instance of type Counterexample X was found in scope.+
+% \end{displayquote}
+\subsection{Instances}
+One bit of cruft in the above proof is the need to specify the particular
+finiteness proof for bools.
+While this isn't any great burden in this case, it of course becomes more
+difficult in more complex circumstances.
+
+In Agda, as in Haskell, the problem of overlapping instances often presents
+itself when more complex typeclass-based machinery is used.
+While there are ways to deal with overlapping instances, it is usually better to
+avoid them altogether, which is precisely what we do in our approach.
+We provide instances for the semiring-based combinators we defined in
+\todo[noline]{reference here}
+, and require the user to provide an instance for their own type.
+We don't, for example, provide an instance that itself searches for surjections
+from other instances: this instance is not uniquely determined.
+
+After all of that, the change to our \(\AgdaFunction{∀↯}\) is simple: we just
+change the first parameter from explicit to an instance.
+% \ExecuteMetaData[agda/Cardinality/Finite/Search.tex]{auto-inst}
+
+The change to the \(\AgdaDatatype{Pauli}\) proof of cancellation isn't
+groundbreaking at first:
+\ExecuteMetaData[agda/Data/Pauli.tex]{cancel-auto}
+But the real benefit is now we can automate proofs over \emph{tuples}, allowing
+us to prove, for instance, commutativity.
+\ExecuteMetaData[agda/Data/Pauli.tex]{comm-uncurried}
+\subsection{Generic Currying and Uncurrying}
+While we have arguably removed the bulk of the boilerplate from the automated
+proofs, there is still the case of the ugly noise of currying and uncurrying.
+In this section, we take inspiration from
+\cite{allaisGenericLevelPolymorphic2019} to develop a small interface to generic
+\(n\)-ary functions and properties.
+What we provide here differs from that work in the following ways:
+\begin{itemize}
+  \item Our generic representation can handle dependent \(\sum\) and \(\prod\)
+    types (rather than their non-dependent counterparts, \(\times\) and
+    \(\rightarrow\)).
+    This extension was necessary for our use case: it is mentioned in the paper
+    as the obvious next step. 
+  \item We implement the curry-uncurry combinators as (verified) isomorphisms.
+    Since we are in a cubical setting, this gives us equivalences between the
+    types, a feature not available in standard Agda.
+  \item We deal with implicit and instance arguments generically.
+\end{itemize}
+A full explanation of our implementation is beyond the scope of this work, but
+we will mention the key parts here.
+First, we define a function arrow generic over the application method:
+% \ExecuteMetaData[agda/Data/Product/N-ary.tex]{fun-arr}
+Then, we prove that it is isomorphic to the normal function arrow:
+% \ExecuteMetaData[agda/Data/Product/N-ary.tex]{fun-iso}
+This step will allow us to write the curry-uncurry proofs once, and then extend
+them to the three different argument forms without difficulty.
+
+We do the same with dependent function types:
+% \ExecuteMetaData[agda/Data/Product/N-ary.tex]{pi-iso}
+
+Next, we need to make our machinery for multiple arguments.
+Here we will define a generic tuple, indexed by some \(\Nat\).
+As observed in \cite{allaisGenericLevelPolymorphic2019}, it's important to
+\emph{not} implement this as an inductively defined vector, e.g.
+% \ExecuteMetaData[agda/Data/Vec.tex]{def}
+As this will not give us the correct \(\eta\)-equality we need for unification.
+\todo[noline]{Stick reference to iterated Vec definition}
+Once we have a unification-friendly vector type, we can use it to implement our
+generic (and level-polymorphic) tuples.
+% \ExecuteMetaData[agda/Data/Product/N-ary.tex]{tuple}
+We decided against \(\top\)-terminated tuples, as the actual contents of the
+tuple type are exposed in the interface.
+
+Finally, we can prove isomorphisms between the curried and uncurried versions of
+functions:
+% \ExecuteMetaData[agda/Data/Product/N-ary.tex]{curry-n-iso}
+And dependent functions:
+% \ExecuteMetaData[agda/Data/Product/N-ary.tex]{curry-n-pi-iso}
+With these combinators, we can implement the search functions in an
+arity-generic way:
+% \ExecuteMetaData[agda/Cardinality/Finite/Search.tex]{all-nary}
+And automate away our proofs:
+\ExecuteMetaData[agda/Data/Pauli.tex]{prf-curried}
 

### 2020-07-05 sections/search.tex
much more towards instances for pi

@@
 
-% As a quick aside, when the property is \emph{not} true, for instance:
-% \ExecuteMetaData[agda/Data/Pauli.tex]{contra}
-% Agda will fail by not finding an instance of \(\bot\).
-% The error message specifically is:
-% \begin{displayquote}
-%   \verb+No instance of type ⊥ was found in scope.+
-% \end{displayquote}
-% We can actually display a counterexample, by defining a custom empty type
-% parameterised by the counterexample itself.
-% % \ExecuteMetaData[agda/Cardinality/Finite/Search.tex]{counterexample-def}
-% Combined with some other changes to the combinators, this will give the
-% following more helpful error message:
-% \begin{displayquote}
-%   \verb+No instance of type Counterexample X was found in scope.+
-% \end{displayquote}
 \subsection{Instances}
@@
 difficult in more complex circumstances.
+
+To solve this we can use Agda's instance search.
+This changes the definitions of
+our automation functions to the following:
+
+\begin{minipage}{\linewidth}
+  \begin{multicols}{2}
+    \begin{agdalisting}
+      \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable/Search.tex]{forall-zap-inst}
+    \end{agdalisting} \columnbreak
+    \begin{agdalisting}
+      \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable/Search.tex]{exists-zap-inst}
+    \end{agdalisting}
+  \end{multicols}
+\end{minipage}
+And this also changes the idempotency proof to the following:
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Snippets/Bool.tex]{with-inst-proof}
+\end{agdalisting}
+
+Again, there's not any great revelation in ease of use here, but more complex
+examples really benefit.
+Especially when we build the full set of instances: any expression built out of
 

### 2020-07-05 sections/search.tex
figured out pi and sigma instances

@@
 Especially when we build the full set of instances: any expression built out of
-
-In Agda, as in Haskell, the problem of overlapping instances often presents
-itself when more complex typeclass-based machinery is used.
-While there are ways to deal with overlapping instances, it is usually better to
-avoid them altogether, which is precisely what we do in our approach.
-We provide instances for the semiring-based combinators we defined in
-\todo[noline]{reference here}
-, and require the user to provide an instance for their own type.
-We don't, for example, provide an instance that itself searches for surjections
-from other instances: this instance is not uniquely determined.
-
-After all of that, the change to our \(\AgdaFunction{∀↯}\) is simple: we just
-change the first parameter from explicit to an instance.
-% \ExecuteMetaData[agda/Cardinality/Finite/Search.tex]{auto-inst}
-
-The change to the \(\AgdaDatatype{Pauli}\) proof of cancellation isn't
-groundbreaking at first:
-\ExecuteMetaData[agda/Data/Pauli.tex]{cancel-auto}
-But the real benefit is now we can automate proofs over \emph{tuples}, allowing
-us to prove, for instance, commutativity.
-\ExecuteMetaData[agda/Data/Pauli.tex]{comm-uncurried}
+products and sums will automatically have an instance.
+This will allow us, for instance, to perform proof search over tuples, which
+gives us some degree of automation for proof search in tuples.
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Snippets/Bool.tex]{and-comm}
+\end{agdalisting}
 \subsection{Generic Currying and Uncurrying}

### 2020-07-05 sections/search.tex
bit more

@@
 In this section, we take inspiration from
-\cite{allaisGenericLevelPolymorphic2019} to develop a small interface to generic
+\citet{allaisGenericLevelPolymorphic2019} to develop a small interface to generic
 \(n\)-ary functions and properties.
+We will describe it briefly here.
+
+The basic idea of currying and uncurrying generically is to allow ourselves to
+work with a generic and flexible representation of function arguments which can
+be manipulated more easily than a simple function itself.
+Our first task, then, is to define that representation of function arguments.
+As in \citet{allaisGenericLevelPolymorphic2019}, our representation is a tuple
+which is in some sense a ``second order'' indexed type.
+By second order here we mean that it is an indexed type indexed by another
+indexed type.
+The reason for this complexity is that our solution is to be fully
+level-polymorphic.
+To start, we define a type representing a vector of universe levels:
+
+\begin{minipage}{\linewidth}
+  \begin{multicols}{2}
+    \begin{agdalisting}
+      \ExecuteMetaData[agda/Data/Product/NAry.tex]{levels}
+    \end{agdalisting} \columnbreak
+    \begin{agdalisting}
+      \ExecuteMetaData[agda/Data/Product/NAry.tex]{max-level}
+    \end{agdalisting}
+  \end{multicols}
+\end{minipage}
+
+This will be used to assign our tuple the correct universe level generically.
+Next, we define the list of types (this type is indexed by the list of universe
+levels of each type):
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Data/Product/NAry.tex]{types}
+\end{agdalisting}
+And finally, the tuple, indexed by its list of types:
+
+\begin{minipage}{\linewidth}
+  \begin{multicols}{2}
+    \begin{agdalisting}
+      \ExecuteMetaData[agda/Data/Product/NAry.tex]{tuple-plus}
+    \end{agdalisting} \columnbreak
+    \begin{agdalisting}
+      \ExecuteMetaData[agda/Data/Product/NAry.tex]{tuple}
+    \end{agdalisting}
+  \end{multicols}
+\end{minipage}
+The reason for two separate functions here is to avoid the
+\(\AgdaDatatype{\(\top\)}\)-terminated tuples we would need if we just had one.
+This means that, for instance, to represent a tuple of a \AgdaDatatype{Bool} and
+\agdambb{N} we can write
+\((\AgdaInductiveConstructor{true}\;\AgdaInductiveConstructor{,}\;\AgdaNumber{2})\)
+instead of 
+\((\AgdaInductiveConstructor{true}\;\AgdaInductiveConstructor{,}\;\AgdaNumber{2}\;\AgdaInductiveConstructor{,}\;\AgdaInductiveConstructor{tt})\).
+
+Next we turn to how we will represent functions.
+In Agda there are three ways to pass function arguments: explicitly,
+implicitly, and as an instance.
+We will represent these three different versions with a data type:
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Data/Product/NAry.tex]{arg-form}
+\end{agdalisting}
+
 What we provide here differs from that work in the following ways:

### 2020-07-06 sections/search.tex
finished explanation of n-ary product code

@@
 \end{agdalisting}
+And then we can make a type for functions in the general sense: a type which has
+this sum type as a parameter.
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Data/Product/NAry.tex]{generic-function}
+\end{agdalisting}
+And we can show that this is isomorphic to a normal function:
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Data/Product/Nary.tex]{generic-function-iso}
+\end{agdalisting}
+
+This of course is only a representation of \emph{non}-dependent functions.
+Dependent functions are defined in a similar way:
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Data/Product/NAry.tex]{pi-iso}
+\end{agdalisting}
+
+Using both of these things, we can now define a generic type for multi-argument
+functions:
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Data/Product/NAry.tex]{multi-generic}
+\end{agdalisting}
+We can also define multi-argument dependent functions in a similar way.
+Similarly to how we had to define two tuple types in order to avoid the
+\(\top\)-terminated tuples, we have two definitions for multi-argument dependent
+functions.
+We only include the nonempty version here for brevity.
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Data/Product/NAry.tex]{pi-arrs-plus}
+\end{agdalisting}
+
+Finally, this all allows us to define an isomorphism between generic
+multi-argument dependent functions and their uncurried forms.
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Data/Product/NAry.tex]{full-iso}
+\end{agdalisting}
 

### 2020-07-07 sections/topos.tex
small intro to topos

@@
 A lot of the work for this proof has been done already: in
-Theorem~\ref{cardinal-kuratowski} we saw that Kuratowski finite types were
-equivalent to Cardinally finite types.
+Theorem~\ref{cardinal-kuratowski} we saw that discrete Kuratowski finite types
+were equivalent to Cardinally finite types.
 We will use the latter definition implementation-wise from now on, as it is
@@
 
+There are two reasons we're interested in the categorical and topos-theoretic
+interpretation of finite sets: first, it's an important theoretical grounding
+for finite sets, which allows us to understand them in the context of other
+set-like constructions.
+Secondly, and more practically, the language of a topos is (or in our case the
+\(\Pi\)-pretopos) is a common standard framework for doing mathematics
+generally.
+This makes it a good basis for an API for building QuickCheck-like generators,
+for example.
 \section{Categories in HoTT}

### 2020-07-07 sections/topos.tex
"categories in HoTT" summary

@@
 \section{Categories in HoTT}
+At first glance, HoTT seems like a perfect setting for category theory: the
+univalence axiom identifies isomorphisms with equality, a useful tool for
+category theory missing from MLTT.
+While this initial impression is broadly true, the construction of categories in
+HoTT is unfortunately quite complex and involved (much of the following is a
+summary of \citet[chapter 9]{hottbook}).
+
+First, we need to think about the type of objects and arrows.
+We cannot, unfortunately, leave them unrestricted: because of the potential for
+higher homotopy in HoTT types \todo{This sentence is a tongue twister}, we have
+to restrict the type of arrows to just the sets.
+This notion: that of a category with all the usual laws such that arrows are a
+set, is called a \emph{precategory}.
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Categories.tex]{precategory}
+\end{agdalisting}
+We will use long arrows to refer to morphisms within a category:
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Categories.tex]{morph-arrow}
+\end{agdalisting}
+
+From here, we can define a notion of isomorphisms.
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Categories.tex]{isomorphism}
+\end{agdalisting}
+It's a condition on this type which separates the precategories from the
+categories: if it satisfies a form of univalence, it the precategory is a full
+category.
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Categories.tex]{cat-univalence}
+\end{agdalisting}
+
 \section{Closure}

### 2020-07-08 sections/topos.tex
added some references
bit more on topos

@@
 
+\todo{references here are tricky, need to disentangle the contributions quite
+  precisely}
+Much of this section is simply a summary of parts of \citet[chapter
+9]{hottbook}.
+The formal proofs we provide are part translation of those proofs in that
+chapter, part from \cite{iversenFredefoxCat2018}
+\cite{huProofrelevantCategoryTheory2020}, and part our own.
+
 First, we need to think about the type of objects and arrows.
@@
 \end{agdalisting}
+\section{The Category of Sets}
+Next we'll look at how to construct the category of sets (in the HoTT sense).
+Much of this work comes directly from \citet[chapter 10]{hottbook} and
+\citet{rijkeSetsHomotopyType2015}.
+The formalisation, however, is novel, as far as we know.
 
+The objects are represented by a \(\Sigma\):
+\begin{agdalisting}
+  \ExecuteMetaData[agda/Snippets/Category.tex]{hset}
+\end{agdalisting}
+This will be quite similar to our objects for finite sets.
+
+Since sets in HoTT don't form a topos, there are quite a few smaller lemmas we
+need to prove to get as close as we can (a \(\Pi W\)-pretopos): we won't include
+them here, other than the closure proofs in the following section.
 \section{Closure}

### 2020-07-08 sections/topos.tex
down to closure proofs on ordered types

@@
 \section{Closure}
-\todo{This is just the closure bits pulled over from the finiteness predicates
-  section. They need to be reworked to fit in here.}
-\subsection{Split Enumerability Closure}
-Now that we have a suitable definition of finiteness, we will next prove that
-some things are finite.
-With the most basic simple types out of the way, the obvious next choice is the
-(non-dependent) sums and products: \(\uplus\) and \(\times\).
-Both of these types can be constructed from the \emph{dependent} sum, however,
-so that is the type we will prove finite.
-From that we can derive a much wider array of finiteness proofs.
+The two most involved proofs for showing that discrete Kuratowski sets form a
+\(\Pi\)-pretopos are those proofs that show closure under \(\Pi\) and
+\(\Sigma\).
+We will describe them here.
+\subsection{Closure of the Ordered Predicates}
+First, we will show that split enumerability (and, by extension, manifest
+enumerability) are closed under \(\Pi\) and \(\Sigma\).
+This is the first stepping stone on our way to prove that cardinal finiteness is
+closed under the same.
+
+Practically speaking, these proofs also open up a wide number of other closure
+proofs to us.
+By proving that dependent products and sums are finite, we get the non-dependent
+cases for free.
 
@@
   Split enumerability is closed under \(\Sigma\).
-  \begin{equation}
-    \frac{
-      \AgdaDatatype{\ensuremath{\mathcal{E}!}}\;A \; \; \; (x : A) \rightarrow \AgdaDatatype{\ensuremath{\mathcal{E}!}}(U\;x)
-    }{
-      \AgdaDatatype{\ensuremath{\mathcal{E}!}}(\AgdaDatatype{\ensuremath{\Sigma [}}\;x \AgdaDatatype{:} A \AgdaDatatype{]} U\;x)
-    }
-  \end{equation}
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{split-enum-sigma}
+  \end{agdalisting}
 \end{lemma}
 \begin{proof}
-  Let \(A\) be a type which is split enumerable, and \(U\) be a type family over
-  \(A\) which is split enumerable at every point.
-  Formally, we have the following proofs:
-  \begin{align}
-    \AgdaDatatype{\ensuremath{\mathcal{E}!}}_A &: \AgdaDatatype{\ensuremath{\mathcal{E}!}}(A) \\
-    \AgdaDatatype{\ensuremath{\mathcal{E}!}}_U &: \Pi(x : A) , \AgdaDatatype{\ensuremath{\mathcal{E}!}}(U(x))
-  \end{align}
-
-  Our task is to construct a proof of type:
-  \begin{equation}
-    \AgdaDatatype{\ensuremath{\mathcal{E}!}}(\Sigma(x : A) , U(x))
-  \end{equation}
-  This proof itself is composed of two components:
-  \begin{align}
-    \mathit{support} &: \AgdaDatatype{List}(\Sigma(x : A) , U(x)) \\
-    \mathit{cover} &: \Pi(x : \Sigma(y : A) , U(y)) , x \in \mathit{support}
-  \end{align}
-  To construct the support list, we apply the function \(\AgdaDatatype{\ensuremath{\mathcal{E}!}}_U\) to
-  every element in the support list of \(\AgdaDatatype{\ensuremath{\mathcal{E}!}}_A\), extract the support
-  lists from the resulting finiteness proofs, and concatenate them.
-
-  To prove that this support list does in fact cover the entirety of the type
-  \(\Sigma \; A \; U\), we note that any element of type \(\Sigma \; A \; U\)
-  must have a first component in the support list of \(\AgdaDatatype{\ensuremath{\mathcal{E}!}}_A\), and its
-  second component must be in the result of applying \(\AgdaDatatype{\ensuremath{\mathcal{E}!}}_U\) to that
-  first element (since that support list contains every element of type
-  \(U(x)\)).
-  Therefore, the pair itself must be in our constructed support list.
-\end{proof}
-
-This pattern of applying a function to each element in a list and
-concatenating the result is of course well-known in functional programming,
-and is in fact the pattern that makes lists a monad.
-While this insight isn't strictly relevant to our work here, it does mean
-the implementation of this function can use Agda's do notation, resulting
-in the following extremely clean implementation:
+  Our task is to construct the two components of the output pair: the support
+  list, and the cover proof.
+  We'll start with the support list: this is constructed by taking the Cartesian
+  product of the input support lists.
   \begin{agdalisting}
@@
   \end{agdalisting}
+  We use do notation here because we're working the list monad: this applies the
+  latter function (\(ys\)) to every element of the list \(xs\), and concatenates
+  the results.
 
-We now have two components we'll need for the proof that the countdown
-transformation is finite.
-The component we'll look at is step~\ref{countdown-operators}: selection of the
-operators.
-We'll first need a type representing the operators available to us.
-\begin{agdalisting}
-  \ExecuteMetaData[agda/Countdown.tex]{ops-def}
-\end{agdalisting}
-Proving that this type is finite takes much the same form as the proof of
-finiteness for bool.
-\begin{agdalisting} \label{op-fin}
-  \ExecuteMetaData[agda/Countdown.tex]{op-fin}
-\end{agdalisting}
+  To show that this does indeed cover every element of the target type is a
+  little intricate, but not necessarily difficult. \todo{Should a proof of this
+    be included?}
+\end{proof}
 
-Next, we will need to build a proof of finiteness for vectors of length \(n\).
-This uses the proof of finiteness for \(\Sigma\).
-\begin{agdalisting}
-  \ExecuteMetaData[agda/Countdown.tex]{vec-fin}
-\end{agdalisting}
-
-\subsection{Manifest Bishop Closure Under \(\Pi\)}
-The glaring omission from our closure proofs under type formers so far has been
-the \(\Pi\) type: we have not proved closure under functions, dependent or
-otherwise.
+Next we'll look at closure under \(\Pi\).
 In MLTT, this is of course not provable: since all of the finiteness predicates
@@
 furthermore of all the previous closure proofs.
-We will prove this closure on split enumerability, rather than on manifest
-Bishop finiteness, as it requires slightly less legwork in the proof itself, but
-of course we can derive the proof on manifest Bishop finiteness in a few lines.
 \begin{theorem} \label{split-enum-pi}
-  Split enumerability is closed under dependent functions.
+  Split enumerability is closed under dependent functions
   (\(\Pi\)-types).
-  \begin{equation}
-    \frac{
-      \AgdaDatatype{\ensuremath{\mathcal{E}!}}(A) \; \; \; \Pi {(x : A)} , \AgdaDatatype{\ensuremath{\mathcal{E}!}}\left( U(x) \right)
-    }{
-      \AgdaDatatype{\ensuremath{\mathcal{E}!}}\left(\Pi {(x : A)} , U(x)\right)
-    }
-  \end{equation}
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Cardinality/Finite/ManifestBishop.tex]{pi-clos}
+  \end{agdalisting}
 \end{theorem}
@@
     \frac{
-      \AgdaDatatype{\ensuremath{\mathcal{E}!}}(\AgdaDatatype{Fin}\;n) \; \; \; \Pi (x : \AgdaDatatype{Fin}\;n) , \AgdaDatatype{\ensuremath{\mathcal{E}!}}\left( U(x) \right)
+      \AgdaDatatype{\ensuremath{\mathcal{E}!}}\;(\AgdaDatatype{Fin}\;n) \; \; \; \left((x : \AgdaDatatype{Fin}\;n) \rightarrow \AgdaDatatype{\ensuremath{\mathcal{E}!}}\;\left( U\;x \right)\right)
     }{
-      \AgdaDatatype{\ensuremath{\mathcal{E}!}}\left(\Pi (x : \AgdaDatatype{Fin}\;n) , U(x)\right)
+      \AgdaDatatype{\ensuremath{\mathcal{E}!}}\left((x : \AgdaDatatype{Fin}\;n) \rightarrow U\;x\right)
     }
@@
 
-  We then define the type of \(n\)-tuples over some type family \(T :
-  \AgdaDatatype{Fin}\;n \rightarrow \mathbf{Type}\).
-  \begin{equation}
-    \begin{alignedat}{3}
-      & \mathbf{Tuple}(0, T)   &&\coloneqq \top \\
-      & \mathbf{Tuple}(n+1, T) &&\coloneqq T(0) \times \mathbf{Tuple}(n, T \circ \text{suc})
-    \end{alignedat}
-  \end{equation}
+  We then define the type of \(n\)-tuples over some type family.
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Data/Tuple/UniverseMonomorphic.tex]{tuple-def}
+  \end{agdalisting}
   We can show that this type is equivalent to functions (proven in our formalisation):
-  \begin{equation}
-    \Pi(x : \AgdaDatatype{Fin}\;n) , U(x) \simeq \mathbf{Tuple}(n, U)
-  \end{equation}
+  \begin{agdalisting}
+    \ExecuteMetaData[agda/Data/Tuple/UniverseMonomorphic.tex]{tuple-iso}
+  \end{agdalisting}
   And therefore we can simplify again our goal to the following:
@@
     \frac{
-      \AgdaDatatype{\ensuremath{\mathcal{E}!}}(\AgdaDatatype{Fin}\;n) \; \; \; \Pi (x : \AgdaDatatype{Fin}\;n) , \AgdaDatatype{\ensuremath{\mathcal{E}!}}\left( U(x) \right)
+      \AgdaDatatype{\ensuremath{\mathcal{E}!}}\;(\AgdaDatatype{Fin}\;n) \; \; \; ((x : \AgdaDatatype{Fin}\;n) \rightarrow \AgdaDatatype{\ensuremath{\mathcal{E}!}}\left( U\;x \right))
     }{
-      \AgdaDatatype{\ensuremath{\mathcal{E}!}}\left(\mathbf{Tuple}(n, U)\right)
+      \AgdaFunction{\ensuremath{\mathcal{E}!}}\;\left(\AgdaFunction{Tuple}\;n\;U\right)
     }
@@
 
-  We can prove this goal by showing that \(\mathbf{Tuple}(n, U)\) is split
+  We can prove this goal by showing that \(\AgdaFunction{Tuple}\;n\;U\) is split
   enumerable: it is made up of finitely many products of points of \(U\), which
-  are themselves split enumerable, and \(\top\), which is also split enumerable.
+  are themselves split enumerable, and \agdatop, which is also split enumerable.
   Lemma~\ref{split-enum-sigma} shows us that the product of finitely many split
@@
 \end{proof}
-
-This proof can again give us insight into how to prove finiteness of our
-countdown transformation.
-In the first step (Fig.~\ref{countdown-selection}), we need to select some
-numbers from an input list: this can be described with a function of type
-\(\AgdaDatatype{Fin}\;n \rightarrow \AgdaDatatype{Bool}\), from indices in the
-original list into whether we keep the values or not.
-We now know that we can prove functions finite without difficulty: in this case,
-we can do it even more simply by proving that an \(n\)-tuple of booleans is
-finite.
-
 \subsection{Closure on Cardinal Finiteness}

### 2020-07-08 sections/topos.tex
fixed agda snippet
Update topos.tex

@@
 Since we don't have a function of type \(\mathcal{C}(A) \rightarrow
-\AgdaDatatype{\ensuremath{\mathcal{B}}}(A)\), closure proofs on \(\AgdaDatatype{\ensuremath{\mathcal{B}}}\) do not transfer over to
+\AgdaDatatype{\ensuremath{\mathcal{B}}}\;A\), closure proofs on \(\AgdaDatatype{\ensuremath{\mathcal{B}}}\) do not transfer over to
 \(\mathcal{C}\) trivially (unlike with \(\AgdaDatatype{\ensuremath{\mathcal{E}!}}\) and \(\AgdaDatatype{\ensuremath{\mathcal{B}}}\)).
@@
 Non-dependent operators like \(\times\), \(\uplus\), and \(\rightarrow\) are
-also relatively straightforward: since \(\lVert {\wc} \rVert\) forms a monad, we
+also relatively straightforward: since \(\AgdaDatatype{\ensuremath{\lVert\_\rVert}}\) forms a monad, we
 can apply \(n\)-ary functions to values inside it, combining them together.
 \begin{agdalisting}
-  The fact that \(\lVert \wc \rVert\) forms a monad means that we can lift
-  \(n\)-ary functions like the following:
   \ExecuteMetaData[agda/Cardinality/Finite/ManifestBishop.tex]{times-clos-sig}
+\end{agdalisting}
 Into a truncated context:
+\begin{agdalisting}
   \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{times-clos-impl}
@@
   \frac{
-    \AgdaDatatype{\ensuremath{\mathcal{B}}}(A) \; \; \; \Pi(x : A) , \AgdaDatatype{\ensuremath{\mathcal{B}}}(U(x))
+    \AgdaDatatype{\ensuremath{\mathcal{B}}}\;A \; \; \; ((x : A) \rightarrow \AgdaDatatype{\ensuremath{\mathcal{B}}}\;(U\;x))
   }{
-    \AgdaDatatype{\ensuremath{\mathcal{B}}}(\Pi \; A \; U)
+    \AgdaDatatype{\ensuremath{\mathcal{B}}}\;((x : A) \rightarrow U\;x)
   }
@@
   \frac{
-    \lVert \AgdaDatatype{\ensuremath{\mathcal{B}}}(A) \rVert \; \; \; \lVert \Pi(x : A) , \AgdaDatatype{\ensuremath{\mathcal{B}}}(U(x)) \rVert
+    \AgdaDatatype{\ensuremath{\lVert}}\; \AgdaDatatype{\ensuremath{\mathcal{B}}}\;A \;\AgdaDatatype{\ensuremath{\rVert}} \; \; \; \AgdaDatatype{\ensuremath{\lVert}}\; ((x : A) \rightarrow \AgdaDatatype{\ensuremath{\mathcal{B}}}\;(U\;x)) \;\AgdaDatatype{\ensuremath{\rVert}}
   }{
-    \lVert \AgdaDatatype{\ensuremath{\mathcal{B}}}(\Pi \; A \; U) \rVert
+    \AgdaDatatype{\ensuremath{\lVert}}\; \AgdaDatatype{\ensuremath{\mathcal{B}}}\;((x : A) \rightarrow U\;x) \;\AgdaDatatype{\ensuremath{\rVert}}
   }
@@
   \frac{
-    \lVert \AgdaDatatype{\ensuremath{\mathcal{B}}}(A) \rVert \; \; \; \Pi(x : A) , \lVert \AgdaDatatype{\ensuremath{\mathcal{B}}}(U(x)) \rVert
+    \AgdaDatatype{\ensuremath{\lVert}}\; \AgdaDatatype{\ensuremath{\mathcal{B}}}\;A \;\AgdaDatatype{\ensuremath{\rVert}} \; \; \; ((x : A) \rightarrow \AgdaDatatype{\ensuremath{\lVert}}\; \AgdaDatatype{\ensuremath{\mathcal{B}}}\;(U\;x) \;\AgdaDatatype{\ensuremath{\rVert}})
   }{
-    \lVert \AgdaDatatype{\ensuremath{\mathcal{B}}}(\Pi \; A \; U) \rVert
+    \AgdaDatatype{\ensuremath{\lVert}}\; \AgdaDatatype{\ensuremath{\mathcal{B}}}\;((x : A) \rightarrow U\;x) \;\AgdaDatatype{\ensuremath{\rVert}}
   }
@@
 \begin{equation}
-  (\Pi(x : A) , \lVert \AgdaDatatype{\ensuremath{\mathcal{B}}}(U(x)) \rVert) \rightarrow
-  \lVert \Pi(x : A) , \AgdaDatatype{\ensuremath{\mathcal{B}}}(U(x)) \rVert
+  ((x : A) \rightarrow \AgdaDatatype{\ensuremath{\lVert}}\; \AgdaDatatype{\ensuremath{\mathcal{B}}}\;(U\;x) \;\AgdaDatatype{\ensuremath{\rVert}}) \rightarrow
+  \AgdaDatatype{\ensuremath{\lVert}}\; (x : A) \rightarrow \AgdaDatatype{\ensuremath{\mathcal{B}}}\;(U\;x) \;\AgdaDatatype{\ensuremath{\rVert}}
 \end{equation}
@@
   \begin{equation}
-    \left( \Pi(x : A) ,  \lVert U(x) \rVert \right) \rightarrow \lVert \Pi(x : A) , U(x) \rVert
+    \left( (x : A) \rightarrow  \AgdaDatatype{\ensuremath{\lVert}}\; U(x) \;\AgdaDatatype{\ensuremath{\rVert}} \right) \rightarrow \AgdaDatatype{\ensuremath{\lVert}}\; (x : A) \rightarrow U(x) \;\AgdaDatatype{\ensuremath{\rVert}}
   \end{equation}
@@
   \begin{equation}
-    \mathcal{C}(A) \rightarrow (\Pi(x : A) , \lVert U(x) \rVert) \rightarrow \lVert \Pi(x : A) , U(x) \rVert
+    \agdacal{C}\;A \rightarrow ((x : A) \rightarrow \AgdaDatatype{\ensuremath{\lVert}}\; U(x) \;\AgdaDatatype{\ensuremath{\rVert}}) \rightarrow \AgdaDatatype{\ensuremath{\lVert}}\; (x : A) \rightarrow U(x) \;\AgdaDatatype{\ensuremath{\rVert}}
   \end{equation}
@@
   Let \(A\) be a cardinally finite type, \(U\) be a type family on \(A\), and
-  \(f\) be a dependent function of type \(\Pi(x : A) , \lVert U(x) \rVert\).
+  \(f\) be a dependent function of type \(\Pi(x : A) , \AgdaDatatype{\ensuremath{\lVert}}\; U(x) \;\AgdaDatatype{\ensuremath{\rVert}}\).
 
@@
   \begin{equation}
-    \mathbf{Tuple}(n, \lVert \wc \rVert \circ U) \rightarrow \lVert \mathbf{Tuple}(n, U) \rVert
+    \mathbf{Tuple}(n, \AgdaDatatype{\ensuremath{\lVert}}\; \wc \;\AgdaDatatype{\ensuremath{\rVert}} \circ U) \rightarrow \AgdaDatatype{\ensuremath{\lVert}}\; \mathbf{Tuple}(n, U) \;\AgdaDatatype{\ensuremath{\rVert}}
   \end{equation}
-  Since \(\lVert \wc \rVert\) is closed under finite products, this function
-  exists (in fact, using the fact that \(\lVert \wc \rVert\) forms a monad, we
+  Since \(\AgdaDatatype{\ensuremath{\lVert\_ \rVert}}\) is closed under finite products, this function
+  exists (in fact, using the fact that \(\AgdaDatatype{\ensuremath{\lVert\_ \rVert}}\) forms a monad, we
   can recognise this function as \verb+sequenceA+ from the \verb+Traversable+
